### 2017-11-09 paper/paper.tex
[Paper] Initial setup

\documentclass[conference,10pt]{IEEEtran}

%% IEEE CNS addition:
\makeatletter
\def\ps@headings{%
	\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
	\def\@evenhead{\scriptsize\thepage \hfil \leftmark\mbox{}}%
	\def\@oddfoot{}%
	\def\@evenfoot{}}
\makeatother
\pagestyle{empty}

\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}

\usepackage[binary-units]{siunitx}
\sisetup{range-phrase=--, range-units=single}

\usepackage[basic]{complexity}
\usepackage[super,negative]{nth}

\usepackage{cleveref}
\newcommand{\crefrangeconjunction}{--}

\usepackage{booktabs}
\usepackage{microtype}

%bib
\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,backend=bibtex]{biblatex}
%\addbibresource{papers-off.bib}
%\addbibresource{confs-off.bib}
%\addbibresource{rfc.bib}

%picky abt et al.
\usepackage{xpatch}

\xpatchbibmacro{name:andothers}{%
	\bibstring{andothers}%
}{%
	\bibstring[\emph]{andothers}%
}{}{}

%opening!

\usepackage{hyperref}
\usepackage{url}
\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black,
	pdftitle={Literature + Ideas},
	pdfauthor={Kyle A. Simpson}
}
\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} }

%% Enable /thanks
\IEEEoverridecommandlockouts
\makeatletter
\def\footnoterule{\relax%
	\kern-5pt
	\hbox to \columnwidth{\hfill\vrule width 0.5\columnwidth height 0.4pt\hfill}
	\kern4.6pt}
\makeatother

%-------------------------------------%
%-------------------------------------%

\title{Dynamically Annealed Reinforcement Learning for Intrusion Prevention}
\author{Kyle A. Simpson\thanks{This work was supported by the Engineering and Physical Sciences
		Research Council [grant number EP/M508056/1]},\\\emph{University of Glasgow, Glasgow, Scotland},\\
		\email{k.simpson.1@research.gla.ac.uk}}

\begin{document}

%% If needed, make urls typewritery
\urlstyle{tt}

\maketitle

\begin{abstract}
	Make it look convincing!
\end{abstract}

\section{Introduction}

Okay here is a... paper?

\section{Initial Definitions}

Testing to see if I get the right header behaviour.

\end{document}

### 2017-11-09 paper/paper.tex
[Paper] More initial setup.

@@
 %bib
-\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,backend=bibtex]{biblatex}
+\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,style=ieee,sortcites,backend=bibtex]{biblatex}
 %\addbibresource{papers-off.bib}
@@
 	urlcolor=black,
-	pdftitle={Literature + Ideas},
+	pdftitle={Dynamically Annealed Reinforcement Learning for Intrusion Prevention},
 	pdfauthor={Kyle A. Simpson}
@@
 \newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} }
+
+\usepackage{titling}
+\settowidth{\thanksmarkwidth}{*}
+\setlength{\thanksmargin}{-\thanksmarkwidth}
 
@@
 		\email{k.simpson.1@research.gla.ac.uk}}
+\date{}
 
@@
 %% If needed, make urls typewritery
-\urlstyle{tt}
+%\urlstyle{tt}
 
@@
 
+\printbibliography
+
 \end{document}

### 2017-11-09 paper/paper.tex
[Paper] A to-do list for consideration

@@
 		\email{k.simpson.1@research.gla.ac.uk}}
+
+% Remove date, leave no spacing.
+\predate{}
+\postdate{}
 \date{}
@@
 Testing to see if I get the right header behaviour.
+Another sentence to bring out the edge cases in this style.
+
+Does next paragraph have proper indent? Please!
+
+\section{A Plan, of Sorts}
+
+\begin{enumerate}
+	\item Establish criteria used to inform system temp---what could make my system increase $\epsilon$ again late into training?
+	\item What are my chosen features? Reward Function?
+	\item Test against MARL and others---probably essential.
+	\item Model at least three scenarios: standard (under attack), normality shifts, attack changes.
+	\item \textsc{Ideally:} test adversarial work (what do such examples look like for my classifier, can an attacker actually operate on them?). How do results change with/without pretraining?
+\end{enumerate}
 

### 2018-01-10 paper/paper.tex
[Paper] Forgot something
[Paper] Slight nindent fix
[Paper] Fix on start template
[Paper] Updated thoughts non direction/approach

@@
 
+%% Fix indent in new section...
+\newcommand{\subparagraph}{}
+\usepackage{titlesec}
+\titlespacing*{\section}{0pt}{1.5ex}{0.7ex}
+
 %bib
@@
 
+\newcommand{\mytitle}{%Dynamically Annealed Reinforcement Learning for Intrusion Prevention
+Dynamic Exploration via the Wavelet Transform}
+
 \usepackage{hyperref}
@@
 	urlcolor=black,
-	pdftitle={Dynamically Annealed Reinforcement Learning for Intrusion Prevention},
+	pdftitle={\mytitle{}},
 	pdfauthor={Kyle A. Simpson}
@@
 
-\title{Dynamically Annealed Reinforcement Learning for Intrusion Prevention}
+\title{\mytitle{}}
 \author{Kyle A. Simpson\thanks{This work was supported by the Engineering and Physical Sciences
@@
 	\item Establish criteria used to inform system temp---what could make my system increase $\epsilon$ again late into training?
+	\begin{enumerate}
+		\item Current thoughts: Wavelet analysis of the reward signal.
+		\item Intuition: multiscale representation of variation in reward.
+		\item What do we see? Hopefully, state-to-state variation as well as longer-term characteristics.
+		\item Expect that overall worsening on the coarsest scales should indicate pressing, global need for exploration  (i.e., non-stationarity, changing problem).
+		\item We can perhaps derive a different uncertainty for each scale -- how do we choose? Combine answers, or (in turn) randomly select an $\epsilon$ from the derived set?
+	\end{enumerate}
 	\item What are my chosen features? Reward Function?
+	\item What actions do I give the agents? Just more probabilistic packet dropping?
 	\item Test against MARL and others---probably essential.

### 2018-02-08 marl-report/report.tex
Initial work on a lab report of sorts

\documentclass[10pt, a4paper]{article}

\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}

\usepackage[binary-units]{siunitx}
\sisetup{range-phrase=--, range-units=single}

\usepackage[basic]{complexity}
\usepackage[super,negative]{nth}

\usepackage{booktabs}
\usepackage{microtype}

%% Fix indent in new section...
\usepackage{titlesec}
\titlespacing*{\section}{0pt}{1.5ex}{0.7ex}

%bib
\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,style=ieee,sortcites,backend=bibtex]{biblatex}
\addbibresource{report.bib}

%picky abt et al.
\usepackage{xpatch}

\xpatchbibmacro{name:andothers}{%
	\bibstring{andothers}%
}{%
	\bibstring[\emph]{andothers}%
}{}{}

%opening!

\newcommand{\mytitle}{Lab Report on reimplementation of MARL}

\usepackage{varioref}
\usepackage{hyperref}
\usepackage{url}
\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black,
	pdftitle={\mytitle{}},
	pdfauthor={Kyle A. Simpson}
}
\usepackage{cleveref}
\newcommand{\crefrangeconjunction}{--}

\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} }

\usepackage{titling}
\settowidth{\thanksmarkwidth}{*}
\setlength{\thanksmargin}{-\thanksmarkwidth}

%-------------------------------------%
%-------------------------------------%

\title{\mytitle{}}
\author{Kyle A. Simpson}

\begin{document}

%% If needed, make urls typewritery
%\urlstyle{tt}

\maketitle

\section{Introduction}

Goal: reimplementation of \textcite{DBLP:journals/eaai/MalialisK15}.

\section{Empirical study}

Rough hypotheses:
\begin{itemize}
	\item The findings of \citeauthor{DBLP:journals/eaai/MalialisK15} are replicable in a `real' network environment.

	\item Distributed MARL is tested against an abnormal (ISP-like) network where the network administrator is assumed to have control of nodes which are local to few hosts.
	DDoS mitigation via a `direct-control' reinforcement learning approach over blanket pushback will be expected to cause significant collateral damage, and accordingly have significantly lower performance.

	\item Learners taught with a gradually/linearly annealed $\epsilon$ in this problem domain will be unable to generalise to the case where host distribution, connection patterns etc.\ exhibit explicit non-stationarity.
	\begin{itemize}
		\item For instance, changepoints at later episodes concerning attacker distribution, host distribution, load profiles...
	\end{itemize}
\end{itemize}

\subsection{Methodology}

Yeahh..

\paragraph{Network topology.}
As in the paper. (describe here? TikZ graph?)

?? Simple enough flow rules, just send up the tree indiscriminately. The model doesn't concern the server having to reply to its clients, mysteriously. Oh well.

\paragraph{Traffic generation.}
Traffic is generated using \emph{TCPReplay} at each host, playing back traffic indefinitely from a predefined Pcap file.
??Control over the send rate is managed by either TC, or tcpreplay. I'm having trouble with both at the moment!

The MARL network model requires that hosts be designated `good' or `bad'.
Calculating the reward function for the current network state requires knowledge of which traffic is (il)legitimate, by heuristic or otherwise.
As \citeauthor{DBLP:journals/eaai/MalialisK15} assume perfect knowledge in their training process, I rewrite the traffic from each host to encode this in the last octet of the source IP address.
The scheme is simple: if the octet is even, the packet is `good' (and vice versa).

\paragraph{Pushback.}
Mininet's switches run \emph{Open vSwitch} (OvS), allowing them to be remotely configured via \emph{OpenFlow} messages.
To enable pushback via probabilistic packet dropping as in the original experiment, OvS was modified to include a custom action to do just this for any given flow.

Typically, OvS switches register with a single controller: when a packet which can't be routed arrives, the switch forwards this to the controller, which is responsible for replying with new rules to make delivery possible.
The control application acts outside of the typical Switch--Controller relationship: given that rule updates must be periodically posted \emph{at the control application's discretion}, the flow updates are manually composed and sent directly.

\paragraph{Load monitoring.}
All switches in Mininet appear to occupy the same namespace, and so a daemon running on any Mininet switch is allowed to monitor the interfaces of any other.
This fact is used to easily monitor each interface in a centralised manner.

The monitoring daemon spawns one thread for each interface it's told to subscribe to.
Each of these threads then handles the \emph{libpcap} events attached to that interface, determining whether a packet is `good' or `bad' and appending the packet length to a dedicated location in a shared store.
Every timestep, the calling program locks the shared store and reads out the time difference, traffic counts, and classifications for each interface.
These statistics enable a rough calculation of the current load in \si{\mega\bit\per\second}.

\paragraph{Reinforcement learning.} Opportunistic SARSA! ($\gamma = 0$, no discounting)

For an action $a-t$ chosen by the current value function $Q_t$ while in state $s_t$ ??CLEANUP:

$$ Q_{t+1}(s_t,a_t) = Q_t(s_t,a_t) + \alpha [R_{t+1} + \gamma Q_t(s_{t+1},a_{t+1}) - Q_t(s_t,a_t)] $$
Actions received from each learner cause a different level of throttling to be applied within the network.

?? Tabular approximation, tile coding, maybe list the params here?!

\paragraph{Changes for experiments 2... 3...} Undecided.

\subsection{Results}

TBA. Currently awaiting debugging of anomalous mininet behaviour.

\section{Discussion}

Yeah, those sure were some results huh.

\section{Conclusion}

Yikes!

\printbibliography

\end{document}

### 2018-02-14 marl-report/report.tex
Report progress/fixes

@@
 \usepackage{graphicx}
+\usepackage{amsmath}
 
@@
 \usepackage{booktabs}
-\usepackage{microtype}
+\usepackage[activate={true,nocompatibility},final,tracking=true,kerning=true,spacing=true]{microtype}
 
@@
 
-Yeahh..
+\paragraph{Network topology.}
+As in the paper:
+\begin{itemize}
+	\item The topology contains one server node which has a fixed bandwidth $U_s$ to a dedicated switch.
 
-\paragraph{Network topology.}
-As in the paper. (describe here? TikZ graph?)
+	\item The task of DDoS defence is divided between $n$ teams.
+	Each team's leader switch connects to the server's switch.
 
-?? Simple enough flow rules, just send up the tree indiscriminately. The model doesn't concern the server having to reply to its clients, mysteriously. Oh well.
+	\item Each team then has $m$ intermediate switches, which are connected to it.
+
+	\item Every intermediate router has $k$ connected switches which each host a reinforcement learning agent.
+	An `external' router is connected to each learner---this is an adaptation to move from the previous ISP-like notion of a topology to be more extensible.
+
+	\item Each `external' router then admits $\ell$ hosts.
+	Each host is \emph{legitimate} or an \emph{attacker} according to some fixed probability, which governs the volume of bandwidth it sends to the server.
+\end{itemize}
+This model does not require that the server respond or serve its clients, allowing some simplification: the only flow rules forward packets towards the server indiscriminately.
 
@@
 
-\paragraph{Pushback.}
-Mininet's switches run \emph{Open vSwitch} (OvS), allowing them to be remotely configured via \emph{OpenFlow} messages.
-To enable pushback via probabilistic packet dropping as in the original experiment, OvS was modified to include a custom action to do just this for any given flow.
-
-Typically, OvS switches register with a single controller: when a packet which can't be routed arrives, the switch forwards this to the controller, which is responsible for replying with new rules to make delivery possible.
-The control application acts outside of the typical Switch--Controller relationship: given that rule updates must be periodically posted \emph{at the control application's discretion}, the flow updates are manually composed and sent directly.
-
 \paragraph{Load monitoring.}
@@
 
-The monitoring daemon spawns one thread for each interface it's told to subscribe to.
-Each of these threads then handles the \emph{libpcap} events attached to that interface, determining whether a packet is `good' or `bad' and appending the packet length to a dedicated location in a shared store.
+The monitoring daemon spawns one thread for each interface it told to subscribe to.
+Each of these threads then handles the \emph{libpcap} events attached to that interface, determining whether every received packet is `good' or `bad' and appending the packet length to a dedicated location in a shared store.
 Every timestep, the calling program locks the shared store and reads out the time difference, traffic counts, and classifications for each interface.
-These statistics enable a rough calculation of the current load in \si{\mega\bit\per\second}.
+These statistics enable a rough calculation of the current load across each interface in \si{\mega\bit\per\second}.
 
-\paragraph{Reinforcement learning.} Opportunistic SARSA! ($\gamma = 0$, no discounting)
+\paragraph{Reinforcement learning.} As in the original experiment, I use opportunistic SARSA to learn the state-action value function.
+In this variant, the discount factor is removed ($\gamma = 0$), so the model does not encode or learn any longer term planning (i.e.\ take a worse state to eventually reach a more optimal state).
 
-For an action $a-t$ chosen by the current value function $Q_t$ while in state $s_t$ ??CLEANUP:
+Action selection and learning are simple.
+The agent maintains a \emph{value function} $Q(s, a)$, encoding the \emph{value} of taking any action from a given state.
+Here, we consider an $\epsilon$-greedy process: a random action is chosen with probability $\epsilon$, otherwise the action with the highest value in the current state is chosen.
+This allows for exploration, and is thus linearly annealed.
+If exploration is to stop after $t_{end}$ iterations, we write:
+\begin{equation}
+\epsilon(t) = \epsilon_0(1 - \max[1, t/t_{end}])
+\end{equation}
+For an action $a_t$ chosen by the current value function $Q$ while in state $s_t$, we observe a \emph{reward} $R_{t+1}$ at the next timestep along with state $s_{t+1}$.
+The model may then be updated:
+\begin{align}
+\begin{split}
+Q(s_t,a_t) \leftarrow Q(s_t,a_t) + \alpha [R_{t+1} + \gamma Q(s_{t+1},a_{t+1}) - Q(s_t,a_t)] \\
+= Q(s_t,a_t) + \alpha [R_{t+1} - Q(s_t,a_t)] 
+\end{split}
+\end{align}
+When discounting is zero as above, SARSA and Q-learning become equivalent.
+The learned policy is thus unaware of the action-selection strategy.
 
-$$ Q_{t+1}(s_t,a_t) = Q_t(s_t,a_t) + \alpha [R_{t+1} + \gamma Q_t(s_{t+1},a_{t+1}) - Q_t(s_t,a_t)] $$
-Actions received from each learner cause a different level of throttling to be applied within the network.
+%?? Tabular approximation, tile coding, maybe list the params here?!
+The current state for a learner is initially a vector containing the bandwidths observed along the route from that learner to the server.
+This is discretised using \emph{tile coding}.
+The reward $R_t$ for any learner is computed from the proportions of `good' traffic received at the server switch and the team leader, with explicit punishment for saturating the core link.
 
-?? Tabular approximation, tile coding, maybe list the params here?!
+\paragraph{Pushback.}
+Actions received from each learner cause a different level of throttling to be applied within the network---dropping \SIrange{0}{90}{\percent} of traffic forwarded from each learning agent's switch.
+This functionality is introduced to Mininet by modification of \emph{Open vSwitch} (OVS).
 
-\paragraph{Changes for experiments 2... 3...} Undecided.
+Mininet's switches run OVS, allowing them to be remotely configured via \emph{OpenFlow} messages.
+To enable pushback via probabilistic packet dropping as in the original experiment, the kernel module of OVS was modified to include a custom action to do just this for any given flow.
+This action takes the form $\operatorname{PDrop(\mathit{prob})}$, for some integer $\mathit{prob} \in [0, 2^{32}-1]$: for some $x$ uniformly drawn from the same domain, if $x < \mathit{prob}$ then the rest of the actions execute as normal.
+
+Typically, OVS switches register with a single controller: when a packet which can't be routed arrives, the switch forwards this to the controller, which is responsible for replying with new rules to make delivery possible.
+The control application acts outside of the typical Switch--Controller relationship: given that rule updates must be periodically posted \emph{at the control application's discretion}, the flow updates are manually composed and sent directly.
+
+\paragraph{Changes for experiments 2... 3...}
+Undecided.
+Experiment 2 concerns (in theory) adding more and more hosts and seing if the performance ceiling is lowered.
+(I.e.\ the bound on how much legit traffic you can get to the server decreases)
+
+Issue w/ experiment 3 is that it's hard to get any meaningful variance if we can't have a large ``max-pool'' of hosts.
+I suppose you could keep the total the same, but play about with how the hosts distribute such that there isn't an overload if that always leads to failure anyway.
 
@@
 
-Yikes!
+Inconclusive.
 

### 2018-02-14 code/marl/marl.py
Made host bw controls... controllable

@@ -51,6 +51,15 @@ def marlExperiment(
 
 		dt = 0.001,
 
+		# These should govern tc on the links at various points in the network.
+		#
+		# Final hop is just what you'd think, old_style governs
+		# whether sender rates are mediated by tcpreplay or by tc on the link
+		# over from said host.
+		old_style = False,
+		force_host_tc = False,
+		protect_final_hop = True,
+
 		rf = "ctl"
 	):
 
@@ -261,7 +270,7 @@ def marlExperiment(
 			good = np.random.uniform() < P_good
 			bw = (np.random.uniform(*(good_range if good else evil_range)))
 
-			link = trackedLink(extern, new_host, {})#"bw": bw})
+			link = trackedLink(extern, new_host, {"bw": bw} if (old_style or force_host_tc) else {})
 
 			# Make up a wonderful IP.
 			# Last byte => goodness. Even = good.
@@ -437,8 +446,9 @@ def marlExperiment(
 		
 		# Update master link's bandwidth limit after hosts init.
 		capacity = calc_max_capacity(len(all_hosts))
-		core_link.intf1.config(bw=capacity)
-		core_link.intf2.config(bw=capacity)
+		if protect_final_hop:
+			core_link.intf1.config(bw=capacity)
+			core_link.intf2.config(bw=capacity)
 
 		# Track the rewards, total load observed and legit rates per step (split by episode)
 		rewards.append([])
@@ -462,13 +472,17 @@ def marlExperiment(
 		# gen traffic at each host. This MUST happen after the bootstrap.
 		for (_, _, _, _, hosts, _) in teams:
 			for (host, good, bw, link, ip) in hosts:
-				host.sendCmd(
+				cmd = [
 					"tcpreplay-edit",
 					"-i", host.intfNames()[0],
 					"-l", str(0),
-					"-S", "0.0.0.0/0:{}/32".format(ip),
-					"-M", str(bw),
-					(good_file if good else bad_file)
+					"-S", "0.0.0.0/0:{}/32".format(ip)
+				] + (
+					[] if old_style else ["-M", str(bw)]
+				) + [(good_file if good else bad_file)]
+				 
+				host.sendCmd(
+					*cmd
 				)
 
 		# Let the pcaps come to life.

### 2018-04-02 interim-talk/interim-talk.tex
Initial notes on what I want to talk about at FATA

\documentclass[aspectratio=169,xcolor={dvipsnames}
,hide notes
%,show only notes
%,show notes on second screen=right
]{beamer}
\usetheme[background=light, numbering=fraction]{metropolis}
\usepackage{appendixnumberbeamer}

%\usepackage{fontspec}
%\setsansfont{Fira Sans Mono}

\usepackage[binary-units]{siunitx}
\sisetup{range-phrase=--, range-units=single}

\usepackage[UKenglish]{babel}
\usepackage{csquotes}

\usepackage{amssymb}

\usepackage{lipsum}
\usepackage[basic]{complexity}
\usepackage[super,negative]{nth}

\usepackage{booktabs}

%bib
\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,sortcites
,backend=bibtex
,style=authortitle
]{biblatex}
\addbibresource{papers-off.bib}
\addbibresource{confs-off.bib}
\addbibresource{books-off.bib}
\addbibresource{rfc.bib}
\addbibresource{misc.bib}

%picky abt et al.
\usepackage{xpatch}

\xpatchbibmacro{name:andothers}{%
	\bibstring{andothers}%
}{%
	\bibstring[\emph]{andothers}%
}{}{}

%opening!

\usepackage{cleveref}
\newcommand{\crefrangeconjunction}{--}

\usepackage{fontawesome}

%-------------------------------------%
%-------------------------------------%

\title{Reinforcement Learning in Network Defence/Control (WIP)}
\author{Kyle A. Simpson\\
	\tiny{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}}
\institute{University of Glasgow}
\date{\nth{10} April, 2018}

\begin{document}

\maketitle

\section{Reinforcement learning}

\begin{frame}{An introduction}
	Yeah. \footcite{RL2E}

	Provide a rough explanation.
\end{frame}

\begin{frame}{Why have RL in networks?}
	Link to recent work in data-driven networking/optimisation?

	Link to that one GMM paper w/ RL communication.
\end{frame}

\section{What I'm working on}

\begin{frame}{MARL for DDoS prevention}
	Network Model

	Pushback
\end{frame}

\begin{frame}{The case for finer granularity}
	Reduced service guarantees by nature of \emph{pushback} model.
	This is exacerbated by TCP congestion avoidance---legit hosts will be punished far more severely, bad actors don't care!

	?? Add my plots of throughput?

	Alongside this higher-granularity view, reformulation to include flow characteristics in the decision-making process.
\end{frame}

\section{The adversarial setting}

\begin{frame}{Attacks}
	Poisoning attacks---explain.

	Evasion attacks---yes. These are what are often known as `adversarial examples'!
\end{frame}

\begin{frame}{The capabilities of attackers}
	Evasion for \emph{polymorphic blending attacks} on past models are known \NP-Complete. ?? How do they differ from modern perturbations, which are solved by stochastic optimisation?

	We have empirically measured bounds on the effectiveness of poisoning attacks wrt.\ what percentage of inbound traffic they can control.

	We don't have results for this on e.g., Deep RL in the network.

	We also lack results on how/if evasion attacks can transfer for learners focused on non-stationarity/constant learning.
\end{frame}

\begin{frame}[standout]{Conclusion}
	We've looked at...
	\begin{itemize}
		\item A quick introduction to RL, and its \alert{importance to future networks}.
		\item A `direct control' approach to intrusion prevention, and \alert{intended improvements}.
		\item Adversarial work, security, and open questions.
	\end{itemize}

	\alert{Questions?}
\end{frame}

\appendix

\begin{frame}[allowframebreaks]{References}
\printbibliography[heading=none]
\end{frame}

\end{document}

### 2018-04-03 interim-talk/interim-talk.tex
Minor talk changes

@@
 \author{Kyle A. Simpson\\
-	\tiny{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}}
+	\small{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}}
 \institute{University of Glasgow}
@@
 
-\begin{frame}{An introduction}
-	Yeah. \footcite{RL2E}
+\begin{frame}{Introducing: Reinforcement Learning}
+	\begin{itemize}
+		\item Goal: train an agent to make optimal decisions based on observed state.
+		\begin{itemize}
+			\item Formally, maximise the \alert{expected discounted reward}\footcite{RL2E}.
+		\end{itemize}
 
-	Provide a rough explanation.
+		\item Underlying theory: systems as (discrete-time) \alert{Markov Decision Processes}---states, actions, rewards and transition probabilities.
+		\begin{itemize}
+			\item I.e., choosing action $a_t$ from a policy in state $s_t$, $a_t \sim \pi(s_t)$, induces the next state $s_{t+1}$ and an associated reward $r_{t+1}$ stochastically.
+			\item Generalises to \alert{value} $Q(s,a)$---how much reward can we \emph{eventually} expect from choosing each action currently available?
+		\end{itemize}
+	\end{itemize}
 \end{frame}

### 2018-04-04 interim-talk/interim-talk.tex
Plot added.

@@
 \usepackage{appendixnumberbeamer}
+
+\usepackage[labelfont=bf,textfont={it}]{caption}
+\usepackage{subcaption}
+\captionsetup[figure]{justification=centering}
+\captionsetup[subfigure]{justification=centering}
 
@@
 	Pushback
+
+	Group reward functions etc.
 \end{frame}
@@
 \begin{frame}{The case for finer granularity}
-	Reduced service guarantees by nature of \emph{pushback} model.
-	This is exacerbated by TCP congestion avoidance---legit hosts will be punished far more severely, bad actors don't care!
+	\begin{columns}
+	\begin{column}{0.45\linewidth}
+		\begin{itemize}
+			\item Reduced service guarantees by nature of \emph{pushback} model.
+			\begin{itemize}
+				\item \alert{Worse with good-faith TCP congestion avoidance}.
+			\end{itemize}
+			%This is exacerbated by TCP congestion avoidance---legit hosts will be punished far more severely, bad actors don't care!
 
-	?? Add my plots of throughput?
-
-	Alongside this higher-granularity view, reformulation to include flow characteristics in the decision-making process.
+			\item More granular $\implies$ should focus on flow stats, \alert{not aggregates} like now!
+%Alongside this higher-granularity view, reformulation to include flow characteristics in the decision-making process.
+		\end{itemize}
+	\end{column}
+	\begin{column}{0.5\linewidth}
+		\begin{figure}
+			\includegraphics[width=\linewidth]{../plots/online.pdf}
+			\caption{Service quality decreases as actions become less granular.}
+		\end{figure}
+	\end{column}
+	\end{columns}
 \end{frame}

### 2018-04-06 interim-talk/interim-talk.tex
Almost done!

@@
 
-\begin{frame}{Introducing: Reinforcement Learning}
+\begin{frame}{Introducing: Reinforcement Learning (RL)}
 	\begin{itemize}
@@
 \begin{frame}{Why have RL in networks?}
-	Link to recent work in data-driven networking/optimisation?
+	%Link to recent work in data-driven networking/optimisation?
+	%
+	%Link to that one GMM paper w/ RL communication.
 
-	Link to that one GMM paper w/ RL communication.
+	\begin{itemize}
+		\item \alert{Data-driven networking.} Effectively applied to intra-domain routing \footcite{DBLP:conf/hotnets/ValadarskySST17}, task allocation \footcite{DBLP:conf/hotnets/MaoAMK16} and more, each with general and domain-specific insights.
+
+		\item \alert{In intrusion detectors?} Optimising information sharing in distributed statistical model training \footcite{DBLP:conf/paisi/XuSH07}.
+	\end{itemize}
 \end{frame}
@@
 
-\begin{frame}{MARL for DDoS prevention}
-	Network Model
+\begin{frame}{Multiagent RL for DDoS prevention}
+%	Network Model \footcite{DBLP:journals/eaai/MalialisK15}
+%	
+%	Pushback \footcite{DBLP:journals/ccr/MahajanBFIPS02a}
+%	
+%	Group reward functions etc.
+%
+%	Its weaknesses? Strong assumptions about what knowledge the learners really have...
 
-	Pushback
+	\begin{itemize}
+		\item Reimplementing (and poking holes in) MARL\footcite{DBLP:journals/eaai/MalialisK15}.
+		% How? Violation of ISP-like criterion,
 
-	Group reward functions etc.
+		\item Network model
+		\begin{itemize}
+			\item Hosts have a fixed probability of being benign/malicious.
+			\item $n$ hosts per learner, $i$ learners to a team, $j$ teams, one server.
+			\item Per-team rewards: \alert{coordinated team learning}.
+			\item Action: (per-timestep) choose $p$, s.t. each learner drops $p\%$ of external traffic.
+		\end{itemize}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Flaws/Risks of MARL?}	
+	\begin{itemize}		
+		\item Action mechanism is \emph{pushback}\footcite{DBLP:journals/ccr/MahajanBFIPS02a} (probabilistic packet drop).
+
+		\item Strong assumptions on benign/malicious packet estimation quality. 
+		\item Pushback is known to be flawed in practice.
+		\item Rigid ISP-like network assumptions.
+	\end{itemize}
 \end{frame}
@@
 		\begin{itemize}
+			\item Learner/host ratio (action/host ratio) affects host QoS.
+
 			\item Reduced service guarantees by nature of \emph{pushback} model.
@@
 \begin{frame}{Attacks}
-	Poisoning attacks---explain.
+	\begin{description}
+		\item[Poisoning attacks] Affect the training of a continuously learning model (RL agents, collaborative learning).
 
-	Evasion attacks---yes. These are what are often known as `adversarial examples'!
+		\item[Evasion attacks] Make data/state appear to be different. These are what are often known as \alert{adversarial examples}!
+	\end{description}
 \end{frame}
@@
 \begin{frame}{The capabilities of attackers}
-	Evasion for \emph{polymorphic blending attacks} on past models are known \NP-Complete. ?? How do they differ from modern perturbations, which are solved by stochastic optimisation?
+	\begin{itemize}
+		\item Evasion for \emph{polymorphic blending attacks} (i.e., evasion) on past models are known \NP-Complete\footcite{DBLP:conf/ccs/FoglaL06}.
 
-	We have empirically measured bounds on the effectiveness of poisoning attacks wrt.\ what percentage of inbound traffic they can control.
+		\item We have empirical bounds on the effectiveness of poisoning attacks wrt.\ what percentage of inbound traffic an attacker controls\footcite{DBLP:journals/jmlr/KloftL10}.
 
-	We don't have results for this on e.g., Deep RL in the network.
+		\item Careless model exposure $\implies$ black-box model extraction\footcite{DBLP:conf/uss/TramerZJRR16} $\implies$ evasion\footcite{DBLP:conf/sp/Carlini017}.
 
-	We also lack results on how/if evasion attacks can transfer for learners focused on non-stationarity/constant learning.
+		%\item Promising work on defences\footcite{DBLP:conf/acsac/CaoG17}.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Unknowns}
+	\begin{itemize}
+		\item We don't have results for poisoning on e.g., Deep RL in the network.
+
+		\item We also lack results on how/if evasion attacks can transfer for learners focused on non-stationarity/constant learning.
+
+	\end{itemize}
 \end{frame}

### 2018-04-09 interim-talk/interim-talk.tex
Some minor changes.

@@
 	\begin{itemize}
-		\item Evasion for \emph{polymorphic blending attacks} (i.e., evasion) on past models are known \NP-Complete\footcite{DBLP:conf/ccs/FoglaL06}.
+		\item \emph{Polymorphic blending attacks} (i.e., evasion) on past models are known \alert{\NP-Complete} by reduction to (s)FSA\footcite{DBLP:conf/ccs/FoglaL06}.
 
@@
 
-		\item Careless model exposure $\implies$ black-box model extraction\footcite{DBLP:conf/uss/TramerZJRR16} $\implies$ evasion\footcite{DBLP:conf/sp/Carlini017}.
+		\item Careless model exposure $\implies$ \alert{black-box model extraction}\footcite{DBLP:conf/uss/TramerZJRR16} $\implies$ evasion\footcite{DBLP:conf/sp/Carlini017}.
 
@@
 		\item We don't have results for poisoning on e.g., Deep RL in the network.
+		\begin{itemize}
+			\item Can an adversary precisely affect state at all measurement points?
+			% Networks are unstable ``bursty'', and attacker has limited control over inbound traffic.
+		\end{itemize}
 
-		\item We also lack results on how/if evasion attacks can transfer for learners focused on non-stationarity/constant learning.
+		\item We also lack results on whether evasion attacks can transfer between policies which evolve over time.
+		\begin{itemize}
+			\item E.g., $\pi_t$ to $\pi_{t+\delta}$ for reinforcement learners focused on non-stationarity.
+		\end{itemize}
 

### 2018-05-08 code/marl/marl.py
Online pt 2 for 2, 7

@@ -461,6 +461,7 @@ def marlExperiment(
 		
 		# Update master link's bandwidth limit after hosts init.
 		capacity = calc_max_capacity(len(all_hosts))
+		print capacity, bw_all
 		if protect_final_hop:
 			core_link.intf1.config(bw=capacity)
 			core_link.intf2.config(bw=capacity)

### 2018-05-08 code/marl/online-expr-11.py

@@ -1,0 +1,32 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 11
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+
+	dt = 0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-mod-11.csv", results)
+
+makeResultsAverage("../../results/online-mod-11.csv", "../../results/online-mod-avg-11.csv")
+

### 2018-05-08 code/marl/online-expr-mod-even.py

@@ -27,7 +27,7 @@ results = marlExperiment(
 	rand_seed = 0xcafed00d
 )
 
-writeResults("../../results/online-mod.csv", results)
+writeResults("../../results/online-mod-even.csv", results)
 
-makeResultsAverage("../../results/online-mod.csv", "../../results/online-mod-avg.csv")
+makeResultsAverage("../../results/online-mod-even.csv", "../../results/online-mod-avg-even.csv")
 

### 2018-05-08 code/marl/online-expr-true.py

@@ -25,6 +25,6 @@ results = marlExperiment(
 	rf = "ctl"
 )
 
-writeResults("../../results/online.csv", results)
+#writeResults("../../results/online.csv", results)
 
-makeResultsAverage("../../results/online.csv", "../../results/online-avg.csv")
+#makeResultsAverage("../../results/online.csv", "../../results/online-avg.csv")

### 2018-05-08 code/marl/online-expr-11.py
Actually saved/redone 14

@@ -26,7 +26,7 @@ results = marlExperiment(
 	rand_seed = 0xcafed00d
 )
 
-writeResults("../../results/online-mod-11.csv", results)
+writeResults("../../results/online-11.csv", results)
 
-makeResultsAverage("../../results/online-mod-11.csv", "../../results/online-mod-avg-11.csv")
+makeResultsAverage("../../results/online-11.csv", "../../results/online-11-avg.csv")
 

### 2018-05-08 code/marl/online-expr-mod-more.py

@@ -26,7 +26,7 @@ results = marlExperiment(
 	rand_seed = 0xcafed00d
 )
 
-#writeResults("../../results/online-mod-more.csv", results)
+writeResults("../../results/online-mod-more.csv", results)
 
-#makeResultsAverage("../../results/online-mod-more.csv", "../../results/online-mod-avg-more.csv")
+makeResultsAverage("../../results/online-mod-more.csv", "../../results/online-mod-more-avg.csv")
 

### 2018-05-10 code/marl/sarsa.py
Mplex o'clock

@@ -6,9 +6,12 @@ class SarsaLearner:
 	"""Learning agent powered by Sarsa and Tile Coding, with e-greedy
 			Assumes that actions are discretised, but states are continuous.
 	"""
+	# Tilings taken from
+	# http://etheses.whiterose.ac.uk/8109/1/phd-thesis-malialis.pdf
+	# Section 4.3
 	def __init__(self, max_bw, vec_size, actions,
 				epsilon=0.3, learn_rate=0.05, discount=0,
-				tile_c=16, tilings_c=3, default_q=0.0,
+				tile_c=6, tilings_c=8, default_q=0.0,
 				epsilon_falloff=1000,
 				break_equal=False):
 		state_range = [[0 for i in xrange(vec_size)], [max_bw for i in xrange(vec_size)]]

### 2018-05-10 code/multiplex-test/mplex.py

@@ -10,7 +10,9 @@ import twink.ofp5.build as ofpb
 import twink.ofp5.oxm as ofpm
 import twink.ofp5.parse as ofpp
 
+import csv
 import itertools
+import json
 import numpy as np
 import os
 import random
@@ -34,7 +36,9 @@ def mplexExperiment(
 		playback = False,
 
 		linkopts = {},
-		base_iperf_port = 5201
+		base_iperf_port = 5201,
+
+		csv_dir = None,
 	):
 
 	if inc_function is None:
@@ -276,25 +280,50 @@ def mplexExperiment(
 	executeRouteQueue()
 
 	server_procs = [server.popen(["iperf3", "-s", "-p", str(base_iperf_port+x)], stdin=PIPE, stderr=sys.stderr) for x in xrange(n)]
-	host_procs = []
+	host_procs = [host.popen(
+		["iperf3",
+		"-c", server.IP(),
+		"-p", str(base_iperf_port+i),
+		"-b", "{}M".format(bw),
+		"-t", str(32),
+		"-J"] +
+		["-u"] if udp else [], stdin=PIPE, stdout=PIPE, stderr=sys.stderr) for i, (host, bw) in enumerate(hosts)
+	]
 
 	#net.interact()
 
-	for p in fxrange(pdrop_min, pdrop_max, pdrop_stride):
-		updateUpstreamRoute(server_switch, ac_prob=p)
+	time.sleep(0.5)
 
-		host_procs = [host.popen(["iperf3", "-c", server.IP(), "-p", str(base_iperf_port+i), "-b", "{}M".format(bw)] + ["-u"] if udp else [], stdin=PIPE, stdout=PIPE, stderr=sys.stderr) for i, (host, bw) in enumerate(hosts)]
+	for step in xrange(10):
+		updateUpstreamRoute(server_switch, ac_prob=step/10.0)
 
-		print "=========="
-		print "stats for p_drop={:.2f}".format(p)
-		print "=========="
+		time.sleep(3 * step)
 
-		for proc, (host, bw) in zip(host_procs, hosts):
-			print "bw={}".format(bw)
-			print "----------"
-			print proc.communicate()[0]
+	# for p in fxrange(pdrop_min, pdrop_max, pdrop_stride):
+	# 	updateUpstreamRoute(server_switch, ac_prob=p)
 
-		time.sleep(1)
[…]

### 2018-05-10 code/marl/marl.py
New mplex stuff

@@ -524,6 +524,7 @@ def marlExperiment(
 			]
 
 			observed = load_mbps[0][0] + load_mbps[0][1]
+			print observed, bw_all[2]
 			ratio = observed / bw_all[2]
 
 		last_traffic_ratio = 0.0
@@ -610,8 +611,8 @@ def marlExperiment(
 
 		store_sarsas = team_sarsas
 
-		#for sar in store_sarsas:
-		#	print sar[0].values
+		for sar in store_sarsas:
+			print sar[0].values
 
 	# Okay, done!
 	# Run interesting stats stuff here? Just save the results? SAVE THE LEARNED MODEL?!

### 2018-05-10 code/multiplex-test/mplex.py

@@ -279,56 +279,53 @@ def mplexExperiment(
 	alive = True
 	executeRouteQueue()
 
+	print "starting procs"
 	server_procs = [server.popen(["iperf3", "-s", "-p", str(base_iperf_port+x)], stdin=PIPE, stderr=sys.stderr) for x in xrange(n)]
 	host_procs = [host.popen(
 		["iperf3",
 		"-c", server.IP(),
+		"-i", str(0.5),
 		"-p", str(base_iperf_port+i),
 		"-b", "{}M".format(bw),
 		"-t", str(32),
 		"-J"] +
-		["-u"] if udp else [], stdin=PIPE, stdout=PIPE, stderr=sys.stderr) for i, (host, bw) in enumerate(hosts)
+		(["-u"] if udp else []), 
+		stdin=PIPE, stdout=PIPE, stderr=sys.stderr) for i, (host, bw) in enumerate(hosts)
 	]
 
 	#net.interact()
 
 	time.sleep(0.5)
 
-	for step in xrange(10):
-		updateUpstreamRoute(server_switch, ac_prob=step/10.0)
+	for step in xrange(20):
+		updateUpstreamRoute(server_switch, ac_prob=step/20.0)
+		print "pushing drop rate", step/20.0, "at t =",0.5+1.5*(step)
+		time.sleep(1.5)
 
-		time.sleep(3 * step)
-
-	# for p in fxrange(pdrop_min, pdrop_max, pdrop_stride):
-	# 	updateUpstreamRoute(server_switch, ac_prob=p)
-
-	# 	host_procs = [host.popen(["iperf3", "-c", server.IP(), "-p", str(base_iperf_port+i), "-b", "{}M".format(bw)] + ["-u"] if udp else [], stdin=PIPE, stdout=PIPE, stderr=sys.stderr) for i, (host, bw) in enumerate(hosts)]
-
-	# 	print "=========="
-	# 	print "stats for p_drop={:.2f}".format(p)
-	# 	print "=========="
-
-	# 	for proc, (host, bw) in zip(host_procs, hosts):
-	# 		print "bw={}".format(bw)
-	# 		print "----------"
-	# 		print proc.communicate()[0]
-
-	# 	time.sleep(1)
-
-	datas = [json.loads(proc.communicate()[0]) for proc in host_procs]
+	print "gathering stats..."
+	updateUpstreamRoute(server_switch, ac_prob=0.0)
+	datas_pre = [proc.communicate() for proc in host_procs]
[…]

### 2018-05-11 code/multiplex-test/mplex.py
Better mplex results to plot (averaged)

@@ -35,10 +35,15 @@ def mplexExperiment(
 		playback_file = "../../data/pcaps/bigFlows.pcap",
 		playback = False,
 
-		linkopts = {},
+		linkopts = {"delay": 20},
 		base_iperf_port = 5201,
 
 		csv_dir = None,
+
+		result_interval = 1.0,
+		max_pdrop = 0.4,
+		step_interval = 3.0,
+		n_steps = 10,
 	):
 
 	if inc_function is None:
@@ -244,7 +249,10 @@ def mplexExperiment(
 			bw = inc_function(last_bw, i, n_hosts)
 			last_bw = bw
 
-			trackedLink(server_switch, h)
+			opts = linkopts
+			opts["bw"] = bw
+
+			trackedLink(server_switch, h, extras=opts)
 
 			hosts.append((h, bw))
 			assignIP(h)
@@ -284,7 +292,7 @@ def mplexExperiment(
 	host_procs = [host.popen(
 		["iperf3",
 		"-c", server.IP(),
-		"-i", str(0.5),
+		"-i", str(result_interval),
 		"-p", str(base_iperf_port+i),
 		"-b", "{}M".format(bw),
 		"-t", str(32),
@@ -297,10 +305,10 @@ def mplexExperiment(
 
 	time.sleep(0.5)
 
-	for step in xrange(20):
-		updateUpstreamRoute(server_switch, ac_prob=step/20.0)
-		print "pushing drop rate", step/20.0, "at t =",0.5+1.5*(step)
-		time.sleep(1.5)
+	for step in xrange(n_steps):
+		updateUpstreamRoute(server_switch, ac_prob=max_pdrop * (step/float(n_steps)))
+		print "pushing drop rate", max_pdrop * (step/float(n_steps)), "at t =",0.5+step_interval*(step)
+		time.sleep(step_interval)
 
 	print "gathering stats..."
 	updateUpstreamRoute(server_switch, ac_prob=0.0)
@@ -308,14 +316,13 @@ def mplexExperiment(
 	datas = [json.loads(data[0]) for data in datas_pre]
 	print "gathered"
 
-	if csv_dir is not None:
-		with open(csv_dir, 'w') as csv_file:
-			out = csv.writer(csv_file)
-			for (_, bw), data in zip(hosts, datas):
-				for inter in data["intervals"]:
-					inner = inter["sum"]
-					out.writerow([bw, inner["end"], float(inner["bits_per_second"])/1e6])
[…]

### 2018-06-01 code/marl/online-expr-14-even.py
New results, new names.

@@ -1,0 +1,33 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 14
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+	break_equal = True,
+
+	dt = 0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-14-even.csv", results)
+
+makeResultsAverage("../../results/online-14-even.csv", "../../results/online-14-even-avg.csv")
+

### 2018-06-01 code/marl/online-expr-14.py

@@ -1,0 +1,32 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 14
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+
+	dt = 0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-14.csv", results)
+
+makeResultsAverage("../../results/online-14.csv", "../../results/online-14-avg.csv")
+

### 2018-06-01 code/marl/online-expr-2-even.py

@@ -1,0 +1,30 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [2, 2],
+
+	explore_episodes = 0.3,
+	episodes = 10,#50,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+	break_equal = True,
+
+	dt = 0.05,#0.01,
+
+#	old_style=True,
+
+	rf = "ctl"
+)
+
+writeResults("../../results/online-2-even.csv", results)
+
+makeResultsAverage("../../results/online-2-even.csv", "../../results/online-2-even-avg.csv")

### 2018-06-01 code/marl/online-expr-2.py

@@ -1,0 +1,29 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [2, 2],
+
+	explore_episodes = 0.3,
+	episodes = 10,#50,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+
+	dt = 0.05,#0.01,
+
+#	old_style=True,
+
+	rf = "ctl"
+)
+
+writeResults("../../results/online-2.csv", results)
+
+makeResultsAverage("../../results/online-2.csv", "../../results/online-2-avg.csv")

### 2018-06-01 code/marl/online-expr-7-even.py

@@ -1,0 +1,33 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 7
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+	break_equal = True,
+
+	dt = 0.05,#0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-7-even.csv", results)
+
+makeResultsAverage("../../results/online-7-even.csv", "../../results/online-7-even-avg.csv")
+

### 2018-06-01 code/marl/online-expr-7.py

@@ -1,0 +1,32 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 7
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+
+	dt = 0.05,#0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-7.csv", results)
+
+makeResultsAverage("../../results/online-7.csv", "../../results/online-7-avg.csv")
+

### 2018-06-01 code/marl/online-expr-even.py

@@ -1,30 +1,0 @@
-from marl import *
-from writer import writeResults, makeResultsAverage
-
-results = marlExperiment(
-	n_teams = 2,#5,
-
-	n_inters = 2,
-	n_learners = 3,
-	host_range = [2, 2],
-
-	explore_episodes = 0.3,
-	episodes = 10,#50,#500, Since mininet keeps running out of files even e/ cleanup
-	episode_length = 10000,
-	separate_episodes = True,
-
-	alpha = 0.05,
-	epsilon = 0.2,
-	discount = 0,
-	break_equal = True,
-
-	dt = 0.05,#0.01,
-
-#	old_style=True,
-
-	rf = "ctl"
-)
-
-writeResults("../../results/online-standard-even.csv", results)
-
-makeResultsAverage("../../results/online-standard-even.csv", "../../results/online-standard-avg-even.csv")

### 2018-06-01 code/marl/online-expr-mod-even.py

@@ -1,33 +1,0 @@
-from marl import *
-from writer import writeResults, makeResultsAverage
-
-hosts_p = 7
-
-results = marlExperiment(
-	n_teams = 2,#5,
-
-	n_inters = 2,
-	n_learners = 3,
-	host_range = [hosts_p, hosts_p],
-
-	explore_episodes = 0.3,
-	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
-	episode_length = 10000,
-	separate_episodes = True,
-
-	alpha = 0.05,
-	epsilon = 0.2,
-	discount = 0,
-	break_equal = True,
-
-	dt = 0.05,#0.01,
-
-	rf = "ctl",
-
-	rand_seed = 0xcafed00d
-)
-
-writeResults("../../results/online-mod-even.csv", results)
-
-makeResultsAverage("../../results/online-mod-even.csv", "../../results/online-mod-avg-even.csv")
-

### 2018-06-01 code/marl/online-expr-mod-more-even.py

@@ -1,33 +1,0 @@
-from marl import *
-from writer import writeResults, makeResultsAverage
-
-hosts_p = 14
-
-results = marlExperiment(
-	n_teams = 2,#5,
-
-	n_inters = 2,
-	n_learners = 3,
-	host_range = [hosts_p, hosts_p],
-
-	explore_episodes = 0.3,
-	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
-	episode_length = 10000,
-	separate_episodes = True,
-
-	alpha = 0.05,
-	epsilon = 0.2,
-	discount = 0,
-	break_equal = True,
-
-	dt = 0.01,
-
-	rf = "ctl",
-
-	rand_seed = 0xcafed00d
-)
-
-writeResults("../../results/online-mod-more-even.csv", results)
-
-makeResultsAverage("../../results/online-mod-more-even.csv", "../../results/online-mod-avg-more-even.csv")
-

### 2018-06-01 code/marl/online-expr-mod-more.py

@@ -1,32 +1,0 @@
-from marl import *
-from writer import writeResults, makeResultsAverage
-
-hosts_p = 14
-
-results = marlExperiment(
-	n_teams = 2,#5,
-
-	n_inters = 2,
-	n_learners = 3,
-	host_range = [hosts_p, hosts_p],
-
-	explore_episodes = 0.3,
-	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
-	episode_length = 10000,
-	separate_episodes = True,
-
-	alpha = 0.05,
-	epsilon = 0.2,
-	discount = 0,
-
-	dt = 0.01,
-
-	rf = "ctl",
-
-	rand_seed = 0xcafed00d
-)
-
-writeResults("../../results/online-mod-more.csv", results)
-
-makeResultsAverage("../../results/online-mod-more.csv", "../../results/online-mod-more-avg.csv")
-

### 2018-06-01 code/marl/online-expr-mod.py

@@ -1,32 +1,0 @@
-from marl import *
-from writer import writeResults, makeResultsAverage
-
-hosts_p = 7
-
-results = marlExperiment(
-	n_teams = 2,#5,
-
-	n_inters = 2,
-	n_learners = 3,
-	host_range = [hosts_p, hosts_p],
-
-	explore_episodes = 0.3,
-	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
-	episode_length = 10000,
-	separate_episodes = True,
-
-	alpha = 0.05,
-	epsilon = 0.2,
-	discount = 0,
-
-	dt = 0.05,#0.01,
-
-	rf = "ctl",
-
-	rand_seed = 0xcafed00d
-)
-
-writeResults("../../results/online-mod.csv", results)
-
-makeResultsAverage("../../results/online-mod.csv", "../../results/online-mod-avg.csv")
-

### 2018-06-01 code/marl/online-expr.py

@@ -1,29 +1,0 @@
-from marl import *
-from writer import writeResults, makeResultsAverage
-
-results = marlExperiment(
-	n_teams = 2,#5,
-
-	n_inters = 2,
-	n_learners = 3,
-	host_range = [2, 2],
-
-	explore_episodes = 0.3,
-	episodes = 10,#50,#500, Since mininet keeps running out of files even e/ cleanup
-	episode_length = 10000,
-	separate_episodes = True,
-
-	alpha = 0.05,
-	epsilon = 0.2,
-	discount = 0,
-
-	dt = 0.05,#0.01,
-
-#	old_style=True,
-
-	rf = "ctl"
-)
-
-writeResults("../../results/online-standard.csv", results)
-
-makeResultsAverage("../../results/online-standard.csv", "../../results/online-standard-avg.csv")

### 2018-06-04 code/marl/online-expr-11-even.py
Reruns

@@ -1,0 +1,33 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 11
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+	break_equal = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+
+	dt = 0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-11-even.csv", results)
+
+makeResultsAverage("../../results/online-11-even.csv", "../../results/online-11-even-avg.csv")
+

### 2018-06-04 code/marl/online-expr-4-even.py
Bolstering results.

@@ -1,0 +1,33 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 4
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+	break_equal = True,
+
+	dt = 0.05,#0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-4-even.csv", results)
+
+makeResultsAverage("../../results/online-4-even.csv", "../../results/online-4-even-avg.csv")
+

### 2018-06-04 code/marl/online-expr-4.py

@@ -1,0 +1,32 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 4
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+
+	dt = 0.05,#0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-4.csv", results)
+
+makeResultsAverage("../../results/online-4.csv", "../../results/online-4-avg.csv")
+

### 2018-06-04 code/marl/granularity-expr.sh
More results.

@@ -1,5 +1,5 @@
 #!/bin/bash
-for count in 2 7 11 14;
+for count in 2 4 7 11 14;
 do
 	python "online-expr-$count.py"
 	python "online-expr-$count-even.py"

### 2018-07-05 paper/paper.tex
Ideas for preliminary paper, plots...

--- 2017-11-09 paper/paper.tex
@@
-\documentclass[conference,10pt]{IEEEtran}
+\documentclass[conference, letterpaper, 10pt, times]{IEEEtran}
 
@@
 
+\usepackage[T1]{fontenc}
+
 \usepackage[dvipsnames]{xcolor}
@@
 
-\usepackage{cleveref}
-\newcommand{\crefrangeconjunction}{--}
-
 \usepackage{booktabs}
 \usepackage{microtype}
+
+%% Fix indent in new section...
+\newcommand{\subparagraph}{}
+\usepackage{titlesec}
+\titlespacing*{\section}{0pt}{1.5ex}{0.7ex}
+%\titleformat*{\filcenter\scshape}
 
@@
 
+\newcommand{\mytitle}{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention}
+
+\usepackage{url}
 \usepackage{hyperref}
-\usepackage{url}
+\usepackage{cleveref}
+\newcommand{\crefrangeconjunction}{--}
+
 \hypersetup{
@@
 
-\title{Dynamically Annealed Reinforcement Learning for Intrusion Prevention}
+\title{\mytitle{}}
 \author{Kyle A. Simpson\thanks{This work was supported by the Engineering and Physical Sciences
-		Research Council [grant number EP/M508056/1]},\\\emph{University of Glasgow, Glasgow, Scotland},\\
+		Research Council [grant number EP/M508056/1]}\\\emph{University of Glasgow, Glasgow, Scotland},\\
 		\email{k.simpson.1@research.gla.ac.uk}}
+
+% Remove date, leave no spacing.
+\predate{}
+\postdate{}
 \date{}
@@
 
-Okay here is a... paper?
+What is the story I'm trying to push?
 
-\section{Initial Definitions}
+?? Threat model yada yada everything sucks all the time here are some statistics about things being
 
-Testing to see if I get the right header behaviour.
+?? Reference to the possible mistreatment of RL in our domain? The problem of ``blind application'' so hated by Sommer and Paxson.
+
+?? RL as a mechanism to protect systems by monitoring performance characteristics---overcome the detection problem (different handle on the key problems of burstiness, evolution, non-stationarity).
+
+This paper contributes:
+\begin{itemize}
+	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks.
+	\item Important weaknesses and flaws in the past design and evaluation of similar techniques.
+	\item ?? Advancing the use/representation of state-space in direct-control RL of networked applications.
+\end{itemize}
+
+\section{Background}
+
+?? Introduce RL, related definitions etc.
+
+??
+Is this \emph{actually} just sarsa?
+We're using fn approx (of course), but this is fraught with its own difficulties.
+Is it strictly speaking correct to describe it as Sarsa at this point?
+It's, at the very least, 1-step semi-gradient Sarsa given that it is clearly on-policy...
+?? IDEA: try out average reward, td-$\lambda$ methods...
+
+?? Discuss mininet? Networking terms?
+
+\section{A Plan, of Sorts}
+
+\begin{enumerate}
+	\item The main case for contribution in what I have so far:
+	\begin{itemize}
+		\item Past work reliant on unrealistic network models: tcp-like behaviour (and its effects on collateral damage) not captured, disjoint ranges of traffic distribution (no benign heavy-hitters).
+		\item I offer more realistic network emulation environment, better treatment of protocol/traffic characteristics.
+	\end{itemize}
+	\item Forthcoming: rethinking state/action spaces to operate at a finer level of granularity.
+New network model (live tcp back-and-forth), allows us to test collateral damage assumptions in a more realistic manner (and show clear case for moving beyond work of malialis and kudenko)
+\end{enumerate}
+
+\section{Results and Evaluation}
+
+\begin{figure}[ht]
+	\includegraphics[width=\linewidth]{../plots/online-varyN-uneven}
+	\caption{
+		MARL pushback control system performance plotted over time, for various settings of $n$ hosts per learner.
+		This plot shows that, from the perspective of benign hosts, service guarantees degrade as inference and applied actions become less granular.
+		This generalises for all behavioural discriminators: even a perfect agent \emph{must} punish benign flows if they are grouped with malicious actors.
+		\label{fig:marl-granularity}
+	}
+\end{figure}
+
+See above: a result.
+
+?? Observations: training time lengthier by nature of emulated environment.
+?? Risk of going for a simulation (i.e. numeric)? Always going to be interesting behaviour that is missed out on (in theory), at the cost of training time/limits of simulation speed
+
+\section{Related Work}
+
+This is a new, modern paper, and so I acknowledge others \emph{after} myself.
+How modest!
+
+?? Abuses of RL
+
+?? Direct-control work
+
+?? Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works.
+
+\section{Conclusion}
+
+It all ends here, Mr.\ Bond...
+
+?? Please write me once everything else is in place.
+
+?? Future Work? I.e., \emph{everything}: no one else is really looking at/interested in this specific kind of application of RL yet. \emph{Yet}.
+
+\section*{Acknowledgements}
+My thanks go to Dr.\ Dimitrios Pezaros for his supervision, guidance, and comments relating to the realisation of this work.
+Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
 

### 2018-07-09 code/marl/offline-internal.py
Trying 1k again w/ epsilon fix

@@ -7,7 +7,8 @@ state = ([], [], [], [], None)
 
 filedir = sys.argv[1]
 block_size = int(sys.argv[2])
-should_read = bool(int(sys.argv[3]))
+total_size = int(sys.argv[3])
+should_read = bool(int(sys.argv[4]))
 
 if should_read:
 	with open(filedir, "rb") as infile:
@@ -22,7 +23,7 @@ results = marlExperiment(
 	n_learners = 3,
 	host_range = [2, 2],
 
-	explore_episodes = 80,#000,
+	explore_episodes = 0.8*total_size,
 	episodes = block_size,
 	episode_length = 1000,
 	separate_episodes = False,
@@ -46,6 +47,8 @@ results = marlExperiment(
 	store_sarsas = store_sarsas,
 	rand_state = random_state,
 )
+
+# FIXME:don't pickle results, only the sarsas + rng state!
 
 with open(filedir, "wb") as outfile:
 	cPickle.dump(results, outfile)

### 2018-07-09 code/marl/offline.py

@@ -4,13 +4,13 @@ import subprocess
 import cPickle
 
 pickle_name = "../../results/offline.pkl"
-total_eps = 10000#0
+total_eps = 1000#0#0
 block_size = 50
 
 for i in xrange(0, total_eps, block_size):
 	remaining = total_eps - i
 	local_block = block_size if remaining >= block_size else remaining
-	subprocess.call(["python", "offline-internal.py", pickle_name, str(local_block), str(int(i != 0))])
+	subprocess.call(["python", "offline-internal.py", pickle_name, str(local_block), str(total_eps), str(int(i != 0))])
 
 # process data from loaded pickle
 with open(pickle_name, "rb") as in_file:

### 2018-07-09 paper/paper.tex

@@
 \usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,style=ieee,sortcites,backend=bibtex]{biblatex}
-%\addbibresource{papers-off.bib}
-%\addbibresource{confs-off.bib}
-%\addbibresource{rfc.bib}
+\addbibresource{papers-off.bib}
+\addbibresource{confs-off.bib}
+\addbibresource{books-off.bib}
+\addbibresource{rfc.bib}
+\addbibresource{misc.bib}
 
@@
 \begin{abstract}
-	Make it look convincing!
+Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-heralded, but face problems hampering effective deployment.
+The detection problem in this domain is fraught with difficulty; it is an evolving, non-stationary problem as usage patterns shift, new protocols and applications are introduced, compounded by burstiness and seasonal variation.
+
+\emph{Reinforcement learning} (RL) may overcome the detection problem for certain classes of anomaly by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available).
+
+I present...
+?? Contribs
+
+?? Taking up space to figure out how much room I have for an intro
+
+?? still taking up space...
+
+?? still going...
+
+?? done...
 \end{abstract}
@@
 
-?? Reference to the possible mistreatment of RL in our domain? The problem of ``blind application'' so hated by Sommer and Paxson.
+?? Reference to the possible mistreatment of RL in our domain? The problem of ``blind application'' so hated by \textcite{DBLP:conf/sp/SommerP10}.
 
@@
 	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks.
-	\item Important weaknesses and flaws in the past design and evaluation of similar techniques.
-	\item ?? Advancing the use/representation of state-space in direct-control RL of networked applications.
+	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---deconstruction of the ISP-like formulation, an acknowledgement of the flaws and risk factors of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
+	\item ?? TODO: Advancing the use/representation of state-space in direct-control RL of networked applications.
 \end{itemize}
@@
 
-?? Introduce RL, related definitions etc.
+%?? Introduce RL, related definitions etc.
+\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to perform optimally in the pursuit of a given task \cite{RL2E}.
+We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time it knows which \emph{state} it is in, the set of \emph{actions} which are available to it, and a numeric \emph{reward} obtained from the last action chosen if available.
+RL methods use this information to gradually refine their \emph{valuation} of states or state-action pairs---allowing the inference of policies which choose actions to maximise the total reward.
+
+?? Formalise?
+
+?? Include some details of function approximation in the formalisation?
+
+There exists immense variation in \emph{how} policies and values may be learned, reliant upon the particular convergence guarantees required.
 
@@
 It's, at the very least, 1-step semi-gradient Sarsa given that it is clearly on-policy...
-?? IDEA: try out average reward, td-$\lambda$ methods...
+?? IDEA: try out average reward, td($\lambda$) methods...
 
-?? Discuss mininet? Networking terms?
+?? Discuss mininet? Networking terms? SDN stuff?
 
@@
 	\begin{itemize}
-		\item Past work reliant on unrealistic network models: tcp-like behaviour (and its effects on collateral damage) not captured, disjoint ranges of traffic distribution (no benign heavy-hitters).
+		\item Past work reliant on unrealistic network models: tcp-like behaviour (and its effects on collateral damage) not captured, disjoint ranges of traffic distribution (no benign heavy-hitters), ISP-like topology.
 		\item I offer more realistic network emulation environment, better treatment of protocol/traffic characteristics.
@@
 \end{enumerate}
+
+\section{Methodology}
+
+?? Either say ``I did it like \textcite{DBLP:journals/eaai/MalialisK15}, here are the differences (i.e., the \emph{real} innovations)'' or just spend time going over it again.
+
+?? STRUCTURE: Either (methodology) -> (results) OR (their stuff + its weaknesses) -> (a new direction) [second is way punchier, but not yet ready to go with that.]
 
@@
 
-This is a new, modern paper, and so I acknowledge others \emph{after} myself.
-How modest!
+?? Other recent work in the prevention of DDoS? (i.e., non-RL).
 
-?? Abuses of RL
+%?? Abuses of RL 
+Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
+Past work exhibits abuses of the paradigm as a glorified classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
 
-?? Direct-control work
+?? Good Direct-control work \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}
 
-?? Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works.
+?? Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works. \cite{DBLP:conf/hotnets/ValadarskySST17, DBLP:conf/hotnets/MaoAMK16}.
+The HMM stuff---learning when best to \emph{communicate} and share knowledge between explicit detector models \cite{DBLP:conf/paisi/XuSH07}.
+The latter example's position is slightly weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
 

### 2018-07-09 paper/paper.tex
Init project for traffic, further writeup

@@
 \usepackage[T1]{fontenc}
+\usepackage{amsfonts}
+\usepackage{amsmath}
 
@@
 \emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to perform optimally in the pursuit of a given task \cite{RL2E}.
-We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time it knows which \emph{state} it is in, the set of \emph{actions} which are available to it, and a numeric \emph{reward} obtained from the last action chosen if available.
-RL methods use this information to gradually refine their \emph{valuation} of states or state-action pairs---allowing the inference of policies which choose actions to maximise the total reward.
+We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\mathcal{A} = A(S_t)$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_t \in \mathcal{A}$).
+RL methods combine this information with a \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
+The policy is then further refined by, for instance, updating value estimates for state-action pairs, meaning that RL-based approaches learn adaptively and online.
+Formally, the goal is to maximise the \emph{expected discounted return}
 
@@
 
-\section{Results and Evaluation}
+\section{Performance in an emulated environment}
 
@@
 
+\begin{figure}[ht]
+	\includegraphics[width=\linewidth]{../plots/mplex}
+	\caption{
+		Average TCP upload rate of hosts targeting different bandwidths in \emph{mininet}, computed over 10 runs.
+		As the rate of packet drop is linearly increased from \SIrange{0}{40}{\percent}, all flows begin to converge on an extremely low upload rate, regardless of the target bandwidth.
+		This behaviour is to be expected, and is to the credit of the emulated environment---prior treatments have ignored such phenomena, which have strong implications on agent and network design.
+		This confirms that packet drop does not affect benign TCP flows linearly, causing a higher rate reduction than what malicious flows will experience.
+		\label{fig:mplex-tcp}
+	}
+\end{figure}
+
 See above: a result.
@@
 ?? Risk of going for a simulation (i.e. numeric)? Always going to be interesting behaviour that is missed out on (in theory), at the cost of training time/limits of simulation speed
+
+?? TCP perf characteristics under packet drop known by the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}:
+$$ \operatorname{BW} = \frac{\operatorname{MSS}}{\operatorname{RTT}} \frac{C}{\sqrt{p}},$$
+for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C$ dependent on system assumptions and the probability $p \in (0, 1]$ that a packet is lost.
 
@@
 
+??
+Benefit of the more realistic emulation environ is that it is far closer in behaviour and architecture (i.e.
+viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance.
+It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques.
+Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
+
+??
+Overarching goal: works which respect the complexity of the network environment (rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution.
+Call-to-action?
+
 \section*{Acknowledgements}

### 2018-07-09 code/traffic-host/src/lib.rs

@@ -1,0 +1,16 @@
+extern crate reqwest;
+extern crate tokio;
+
+pub fn run() {
+	println!("Hello, world!");
+
+	match reqwest::get("https://mcfelix.me") {
+		Ok(mut resp) => {
+			println!("url: {}", &resp.url());
+			println!("status: {}", &resp.status());
+			println!("headers: {}", &resp.headers());
+			println!("text: {}", &resp.text().unwrap_or("None".into()));
+		},
+		Err(e) => println!("Error encountered: {:?}", e)
+	}
+}

### 2018-07-09 code/traffic-host/src/main.rs

@@ -1,0 +1,5 @@
+extern crate traffic_host;
+
+fn main() {
+    traffic_host::run();
+}

### 2018-07-10 paper/paper.tex
Stop repeatedly saving & loading stats

@@
 $$ \operatorname{BW} = \frac{\operatorname{MSS}}{\operatorname{RTT}} \frac{C}{\sqrt{p}},$$
-for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C$ dependent on system assumptions and the probability $p \in (0, 1]$ that a packet is lost.
+for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions and the probability $p \in (0, 1]$ that a packet is lost.
+The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size.
 

### 2018-07-10 code/marl/offline-internal.py

@@ -3,7 +3,7 @@ from writer import writeResults, makeResultsAverage
 import cPickle
 import sys
 
-state = ([], [], [], [], None)
+state = ([], None)
 
 filedir = sys.argv[1]
 block_size = int(sys.argv[2])
@@ -14,7 +14,7 @@ if should_read:
 	with open(filedir, "rb") as infile:
 		state = cPickle.load(infile)
 
-(rewards, good_traffic_percents, total_loads, store_sarsas, random_state) = state
+(store_sarsas, random_state) = state
 
 results = marlExperiment(
 	n_teams = 5,
@@ -41,15 +41,14 @@ results = marlExperiment(
 
 	rf = "ctl",
 
-	rewards = rewards,
-	good_traffic_percents = good_traffic_percents,
-	total_loads = total_loads,
 	store_sarsas = store_sarsas,
 	rand_state = random_state,
 )
 
-# FIXME:don't pickle results, only the sarsas + rng state!
+(_, _, _, store_sarsas, random_state) = results
+
+writeResults("../../results/offline.csv", results, append=True)
 
 with open(filedir, "wb") as outfile:
-	cPickle.dump(results, outfile)
+	cPickle.dump((store_sarsas, random_state), outfile)
 

### 2018-07-10 code/marl/offline.py

@@ -4,17 +4,17 @@ import subprocess
 import cPickle
 
 pickle_name = "../../results/offline.pkl"
-total_eps = 1000#0#0
+total_eps = 10000#0
 block_size = 50
 
+# Create the file
+writeResults("../../results/offline.csv", ([], [], [], [], None))
+
+# Run these, natch into sets of 50.
 for i in xrange(0, total_eps, block_size):
 	remaining = total_eps - i
 	local_block = block_size if remaining >= block_size else remaining
 	subprocess.call(["python", "offline-internal.py", pickle_name, str(local_block), str(total_eps), str(int(i != 0))])
 
-# process data from loaded pickle
-with open(pickle_name, "rb") as in_file:
-	results = cPickle.load(in_file)
-
-writeResults("../../results/offline.csv", results)
+# Combine final results.
 lastTimestepsAndEpAverages("../../results/offline.csv", "../../results/offline-avg.csv")

### 2018-07-10 code/marl/writer.py

@@ -14,16 +14,20 @@ def mkdir_p(path):
         else:
             raise
 
-def writeResults(results_file, results, sarsa_dir=None):
+def writeResults(results_file, results, sarsa_dir=None, append=False):
 	(rewards, good_traffic_percents, total_loads, store_sarsas, rng_state) = results
+
+	mode = "a" if append else "w"
 
 	# First, handle the actual results.
 	mkdir_p(os.path.split(results_file)[0])
-	with open(results_file, "w") as csvfile:
+	with open(results_file, mode) as csvfile:
 		out = csv.writer(csvfile)
 		all_t = 0
 		# Key/headers
-		out.writerow(["episode", "t", "global_t", "g_reward", "legit_traffic", "total_load"])
+		if not append:
+			out.writerow(["episode", "t", "global_t", "g_reward", "legit_traffic", "total_load"])
+			
 		for ep, (rewards_ep, g_traf_ep, load_ep) in enumerate(zip(rewards, good_traffic_percents, total_loads)):
 			for t, (reward, g_traffic, load) in enumerate(zip(rewards_ep, g_traf_ep, load_ep)):
 				out.writerow([ep, t, all_t, reward, g_traffic, load])

### 2018-07-10 paper/paper.tex
Way more content

@@
 \documentclass[conference, letterpaper, 10pt, times]{IEEEtran}
-
-%% IEEE CNS addition:
-\makeatletter
-\def\ps@headings{%
-	\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
-	\def\@evenhead{\scriptsize\thepage \hfil \leftmark\mbox{}}%
-	\def\@oddfoot{}%
-	\def\@evenfoot{}}
-\makeatother
-\pagestyle{empty}
 
@@
 \titlespacing*{\section}{0pt}{1.5ex}{0.7ex}
+\titlespacing*{\subsection}{0pt}{1.5ex}{0.7ex}
 %\titleformat*{\filcenter\scshape}
@@
 }{}{}
+\DeclareFieldFormat[inproceedings]{url}{}
+\DeclareFieldFormat[inproceedings]{article}{}
 
@@
 \author{Kyle A. Simpson\thanks{This work was supported by the Engineering and Physical Sciences
-		Research Council [grant number EP/M508056/1]}\\\emph{University of Glasgow, Glasgow, Scotland},\\
+		Research Council [grant number EP/M508056/1]}\\University of Glasgow, Glasgow, Scotland,\\
 		\email{k.simpson.1@research.gla.ac.uk}}
@@
 
-What is the story I'm trying to push?
+Network anomaly detection, intrusion detection and intrusion prevention are continually evolving problems, compounded by the partial, non-IID view of data at each point in the network.
+Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, or availability without being detected.
+These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
 
-?? Threat model yada yada everything sucks all the time here are some statistics about things being
+While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \textcite{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems---to quite a large extent, this remains the case today.
+Their application is made difficult due to significant operational differences from standard ML tasks, alongside certain characteristics of network traffic.
+Briefly, these are the diversity of network traffic across varying timescales \cite{DBLP:conf/sp/SommerP10} and significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
+Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
+This is compounded by the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges of learning from unlabelled (often partial) data.
+All of these factors greatly inflate the difficulty of the detection problem.
+
+For certain classes of problem like non-transit DDoS attacks, \emph{reinforcement learning} (RL) is best-posed to adaptively react to certain classes of threat by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
+In a sense, this allows us to ``overcome'' the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time, by looking for the effect rather than the cause.
+The intent is to augment what existing solutions can provide, by automatically alerting, recording and controlling what are believed to be illegal system states.
+%Whether it takes direct control of the network, or is used indirectly to optimise a key part of another system, more powerful `deep' RL techniques (and well-founded action spaces) aren't yet well explored for network IDS/IPS.
+%These range from more modern training algorithms \cite{DBLP:journals/corr/SchulmanWDRK17, DBLP:conf/icml/SchulmanLAJM15}, to evolutionary strategies \cite{DBLP:journals/corr/SalimansHCS17, DBLP:journals/corr/abs-1802-08842}, hierarchical action composition \cite{DBLP:journals/corr/abs-1710-09767}, and competitive multi-agent learning \cite{DBLP:journals/corr/abs-1710-03748}.
 
@@
 	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks.
-	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---deconstruction of the ISP-like formulation, an acknowledgement of the flaws and risk factors of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
-	\item ?? TODO: Advancing the use/representation of state-space in direct-control RL of networked applications.
+	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction of the ISP-like formulation, an acknowledgement of the flaws and risk factors of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
+	\item ?? TODO: Advance the use/representation/granularity of state-space in direct-control RL of networked applications.
 \end{itemize}
@@
 %?? Introduce RL, related definitions etc.
-\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to perform optimally in the pursuit of a given task \cite{RL2E}.
+\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
 We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\mathcal{A} = A(S_t)$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_t \in \mathcal{A}$).
-RL methods combine this information with a \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
-The policy is then further refined by, for instance, updating value estimates for state-action pairs, meaning that RL-based approaches learn adaptively and online.
-Formally, the goal is to maximise the \emph{expected discounted return}
+RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
+The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in their deployment environment.
+From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{return}, $G_t = R_{t+1} + \ldots + R_{t+T}$ for some (possibly infinite) choice of $T$.
+Formally then, the goal is to maximise the \emph{expected return} $\operatorname{\mathbb{E}_{\pi}}[G_t | S_t]$ from any and all states $S_t$.
 
-?? Formalise?
-
-?? Include some details of function approximation in the formalisation?
+?? Include some details of function approximation in the formalisation? I.e. tile coding, stability and convergence guarantees...
 
@@
 Is it strictly speaking correct to describe it as Sarsa at this point?
-It's, at the very least, 1-step semi-gradient Sarsa given that it is clearly on-policy...
-?? IDEA: try out average reward, td($\lambda$) methods...
+It's, at the very least, 1-step semi-gradient Sarsa given that it is clearly on-policy w/ fn approx...
+?? IDEA: try out average reward, TD($\lambda$) methods as future work...
 
@@
 
-\section{Methodology}
-
-?? Either say ``I did it like \textcite{DBLP:journals/eaai/MalialisK15}, here are the differences (i.e., the \emph{real} innovations)'' or just spend time going over it again.
+\section{Environment and RL Algorithm}
 
@@
 
-\section{Performance in an emulated environment}
+\subsection{Topology}
+The network itself is tree-structured, where one server connects to $k$ team leaders, each connected to $l$ intermediate switches, which in turn each connect to $m$ egress switches.
+Agents are co-located with these egress switches, who control the proportion of upstream packets from $n$ external hosts to discard according to load statistics observed along their path to the server.
+Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
+
+?? Diagram? TikZ?
+
+Reimplementation of the network model was performed using mininet \cite{mininet}, a Python-based network emulator.
+Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
+While unrealistic with respect to packet content, it was assumed that this would provide representative load and packet inter-arrival characteristics.
+A sufficiently well-trained agent needs only to follow the policy it has learned in a greedy manner, allowing pre-training using a priori knowledge by simulation to transfer to reality.
+
+\subsection{Agent Definitions}
+
+?? Specify update rules here? Sarsa \cite[pp.\ \numrange{129}{131}]{RL2E}? Semi-gradient Sarsa? enhanced by tile-coding with identical parameter choices \cite{DBLP:phd/ethos/Malialis14}.
+
+?? Specify reward function here?
+
+?? M\&K Seem not to have a grasp on the importance of up/down traffic splits. Hmm......... (i.e., both could be relevant?)
+
+\section{Performance in an Emulated Environment}
+
+??
+System specs of execution environment: Intel Core i7-6700K clocked at \SI{4.2}{\giga\hertz}, with \SI{32}{\gibi\byte} of RAM.
+All code is available on a public Github repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
 
@@
 for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions and the probability $p \in (0, 1]$ that a packet is lost.
-The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size.
+The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size and is an expected part of normal operation.
 
@@
 
+?? Security? I suspect that the very qualities that make inference difficult in IDS/IPS also increase the level of challenge an advanced threat must overcome.
+System state which is dependent on many signals drawn from across a wide network, in concert with their burstiness and unpredictability, may have substantial effects on an attacker's capabilities.
+
 \section*{Acknowledgements}
-My thanks go to Dr.\ Dimitrios Pezaros for his supervision, guidance, and comments relating to the realisation of this work.
+My thanks go to Dr.\ Dimitrios Pezaros for his supervision, guidance, and comments relating to the realisation of this work, and to Mircea Iordache for his comments and technical assistance.
 Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.

### 2018-07-11 paper/paper.tex
Mathing it up.

@@
 \usepackage{amsmath}
+\usepackage{bm}
 
@@
 
-\usepackage[binary-units]{siunitx}
+\usepackage[binary-units, per-mode=symbol]{siunitx}
 \sisetup{range-phrase=--, range-units=single}
@@
 \DeclareFieldFormat[inproceedings]{url}{}
-\DeclareFieldFormat[inproceedings]{article}{}
+\DeclareFieldFormat[article]{url}{}
 
@@
 \makeatother
+
+% make math easy
+\newcommand{\acval}[3]{\ensuremath{\operatorname{\hat{q}}(#1, #2, #3)}}
+\newcommand{\wvec}[1]{\ensuremath{\bm{w}_{#1}}}
 
@@
 	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks.
-	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction of the ISP-like formulation, an acknowledgement of the flaws and risk factors of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
+	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction of their ISP-like formulation, and an acknowledgement of the flaws and risk factors of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
 	\item ?? TODO: Advance the use/representation/granularity of state-space in direct-control RL of networked applications.
@@
 
-There exists immense variation in \emph{how} policies and values may be learned, reliant upon the particular convergence guarantees required.
+There exists immense variation in \emph{how} policies and/or values may be learned, reliant upon the particular convergence guarantees required.
 
@@
 \subsection{Topology}
-The network itself is tree-structured, where one server connects to $k$ team leaders, each connected to $l$ intermediate switches, which in turn each connect to $m$ egress switches.
+The network itself is tree-structured, where one server connects through a dedicated switch to $k$ team leader switches, each connected to $l$ intermediate switches, which in turn each connect to $m$ egress switches.
 Agents are co-located with these egress switches, who control the proportion of upstream packets from $n$ external hosts to discard according to load statistics observed along their path to the server.
 Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
+Each link has a delay of \SI{10}{\milli\second}.
+All links have unbounded capacity, save for the server-switch connection which is capped at a fixed $U_s$ \si{\mega\bit\per\second} in both directions.
 
@@
 Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
-While unrealistic with respect to packet content, it was assumed that this would provide representative load and packet inter-arrival characteristics.
-A sufficiently well-trained agent needs only to follow the policy it has learned in a greedy manner, allowing pre-training using a priori knowledge by simulation to transfer to reality.
+While unrealistic with respect to packet content, it was assumed that this would provide representative load and packet inter-arrival characteristics, since the agents as initially defined are blind to such information.
 
@@
 
-?? Specify update rules here? Sarsa \cite[pp.\ \numrange{129}{131}]{RL2E}? Semi-gradient Sarsa? enhanced by tile-coding with identical parameter choices \cite{DBLP:phd/ethos/Malialis14}.
+?? Specify update rules here? Semi-gradient Sarsa \cite[pp.\ \numrange{243}{244}]{RL2E}? enhanced by tile-coding with identical parameter choices \cite{DBLP:phd/ethos/Malialis14}.
 
-?? Specify reward function here?
+\begin{subequations}
+\begin{align}
+\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
+\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
+\end{align}
+\end{subequations}
+observing that the gradient is the binary vector of tile activations.
+
+The reward functions as I specify them here were introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, I make use of the global reward as a performance metric and the \emph{coordinated team learning} (CTL) reward for training.
+Using the Iverson bracket notation:
+\begin{subequations}
+\begin{align}
+c_t^{\mathit{f}} = [ > U_s],\\
+c_t^{\mathit{tf}} = [ > U_s/k],\\
+R_{t}^{\mathit{Global}} = (1 - c_t^{\mathit{f}})\frac{a}{b} - c_t^{\mathit{f}},\\
+R_{t}^{\mathit{CTL}} = (1 - c_t^{\mathit{f}}c_t^{\mathit{tf}})\frac{a}{b} - c_t^{\mathit{f}}c_t^{\mathit{tf}},
+\end{align}
+\end{subequations}
+
+?? Note, these are dependent on the current agent! Need to find a succinct way to express this somehow...
 
 ?? M\&K Seem not to have a grasp on the importance of up/down traffic splits. Hmm......... (i.e., both could be relevant?)
+
+?? wrt reward function relying on heuristic elements or perfect knowledge:
+A sufficiently well-trained agent needs only to follow the policy it has learned in a greedy manner, allowing pre-training using a priori knowledge by simulation to transfer to reality.
 
@@
 ??
-System specs of execution environment: Intel Core i7-6700K clocked at \SI{4.2}{\giga\hertz}, with \SI{32}{\gibi\byte} of RAM.
+System specs of execution environment: 4-core Intel Core i7-6700K clocked at \SI{4.2}{\giga\hertz}, with \SI{32}{\gibi\byte} of RAM.
 All code is available on a public Github repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
@@
 ?? TCP perf characteristics under packet drop known by the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}:
-$$ \operatorname{BW} = \frac{\operatorname{MSS}}{\operatorname{RTT}} \frac{C}{\sqrt{p}},$$
+\begin{equation}
+\operatorname{BW} = \frac{\operatorname{MSS}}{\operatorname{RTT}} \frac{C}{\sqrt{p}},
+\end{equation}
 for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions and the probability $p \in (0, 1]$ that a packet is lost.
@@
 \section*{Acknowledgements}
-My thanks go to Dr.\ Dimitrios Pezaros for his supervision, guidance, and comments relating to the realisation of this work, and to Mircea Iordache for his comments and technical assistance.
+My thanks go to Dr.\ Dimitrios Pezaros for his supervision, guidance, and comments relating to the realisation of this work, and to Mircea Iordache and Richard Cziva for their comments and technical assistance.
 Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.

### 2018-07-11 paper/paper.tex
Better math

@@
 
-?? Include some details of function approximation in the formalisation? I.e. tile coding, stability and convergence guarantees...
+%?? Include some details of function approximation in the formalisation? I.e. tile coding, stability and convergence guarantees...
 
-There exists immense variation in \emph{how} policies and/or values may be learned, reliant upon the particular convergence guarantees required.
+There exists immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
+In particular, we shall focus on methods which choose actions according to their value estimates from the current state.
+To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times A$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a) \forall a \in \mathcal{A}$ by subdividing it into a number of overlapping $d$-dimensional grids with an optional bias component.
+This transform, as an example of linear function approximation, then necessitates the use of \emph{semi-gradient Sarsa} \cite[pp.\ \numrange{243}{244}]{RL2E}.
+Given a weight vector $\wvec{0}=\bm{0}$ which defines :
+\begin{subequations}
+	\begin{gather}
+	\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a),\\
+	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
+	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}}.
+	\end{gather}
+	\label{eqn:sg-sarsa}
+\end{subequations}
+
+Observing that $\nabla{\acval{\cdot}{\cdot}{\cdot}} = \operatorname{\mathbf{x}}(s, a)$, and $\operatorname{\mathbf{x}}(s, a)$ .
+Furthermore, if the action set remains
 
@@
 \subsection{Topology}
-The network itself is tree-structured, where one server connects through a dedicated switch to $k$ team leader switches, each connected to $l$ intermediate switches, which in turn each connect to $m$ egress switches.
+The network itself is tree-structured, where one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $l$ intermediate switches, which in turn each connect to $m$ egress switches.
 Agents are co-located with these egress switches, who control the proportion of upstream packets from $n$ external hosts to discard according to load statistics observed along their path to the server.
@@
 
-?? Specify update rules here? Semi-gradient Sarsa \cite[pp.\ \numrange{243}{244}]{RL2E}? enhanced by tile-coding with identical parameter choices \cite{DBLP:phd/ethos/Malialis14}.
+Agents receive state input from monitors in the environment, a vector of load values (in \si{\mega\bit\per\second}) observed along their path to $s$: this is tile-coded with identical parameter choices to those of \textcite{DBLP:phd/ethos/Malialis14}.
 
-\begin{subequations}
-\begin{align}
-\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
-\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
-\end{align}
-\end{subequations}
-observing that the gradient is the binary vector of tile activations.
+At every time step, each agent installs an action onto the switch which hosts it, instructing it to drop $p\si{\percent}$ of traffic.
 
 The reward functions as I specify them here were introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, I make use of the global reward as a performance metric and the \emph{coordinated team learning} (CTL) reward for training.
-Using the Iverson bracket notation:
+These assume the presence of a classifier (exact or heuristic) which estimates the volume of `legitimate traffic' received at any point, $g(\cdot)$.
+At time $t$ an agent $a$ (with team leader $l_a$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
 \begin{subequations}
-\begin{align}
-c_t^{\mathit{f}} = [ > U_s],\\
-c_t^{\mathit{tf}} = [ > U_s/k],\\
-R_{t}^{\mathit{Global}} = (1 - c_t^{\mathit{f}})\frac{a}{b} - c_t^{\mathit{f}},\\
-R_{t}^{\mathit{CTL}} = (1 - c_t^{\mathit{f}}c_t^{\mathit{tf}})\frac{a}{b} - c_t^{\mathit{f}}c_t^{\mathit{tf}},
-\end{align}
+\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
+\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
+\begin{gather}
+\cond{f} = [\load{s} > U_s],\\
+\cond{c}{(a)} = [\load{l_a} > U_s/k],\\
+R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\\
+R_{t}^{\mathit{CTL}} = (1 - \cond{f}\cond{c}{(a)}) \frac{g(\load{l_a})}{g(\load{\mathit{team}})} \allowbreak - \cond{f} \cond{c}{(a)},
+\end{gather}
+where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $a$'s team, respectively.
 \end{subequations}
 
-?? Note, these are dependent on the current agent! Need to find a succinct way to express this somehow...
+The tile-coded state and chosen reward function are then used as input for the \emph{semi-gradient Sarsa} algorithm described in \cref{eqn:sg-sarsa}.
 

### 2018-07-12 paper/paper.tex
Even more paper content

@@
 \usepackage{graphicx}
+\usepackage{tikz}
+\usetikzlibrary{arrows.meta, calc, fit, positioning}
 
@@
 
+\makeatletter
+\DeclareRobustCommand{\rvdots}{%
+	\vbox{
+		\baselineskip4\p@\lineskiplimit\z@
+		\kern-\p@
+		\hbox{.}\hbox{.}\hbox{.}
+}}
+\makeatother
+
+% Official colours!
+
+\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}
+
+\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
+\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
+\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
+\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
+\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
+
+\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
+\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
+\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
+\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
+\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
+\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
+\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
+\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}
+
+\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
+\definecolor{uofgforest}{rgb}{0, 0.317647, 0.2}
+\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
+\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
+
 %-------------------------------------%
@@
 \emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
-We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\mathcal{A} = A(S_t)$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_t \in \mathcal{A}$).
+We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
+This formulation of interaction with a system is best described as a \emph{Markov decision problem} (MDP).
 RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
 The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in their deployment environment.
-From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{return}, $G_t = R_{t+1} + \ldots + R_{t+T}$ for some (possibly infinite) choice of $T$.
-Formally then, the goal is to maximise the \emph{expected return} $\operatorname{\mathbb{E}_{\pi}}[G_t | S_t]$ from any and all states $S_t$.
+From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+2} + \ldots$, choosing the discount factor $\gamma \in [0,1]$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
+Formally then, the goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.
 
@@
 
-There exists immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
+There is immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
 In particular, we shall focus on methods which choose actions according to their value estimates from the current state.
-To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times A$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a) \forall a \in \mathcal{A}$ by subdividing it into a number of overlapping $d$-dimensional grids with an optional bias component.
+To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+|\mathcal{A}|)$-dimensional grids with an optional bias component.
+Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state action pair exists within the corresponding tile.
 This transform, as an example of linear function approximation, then necessitates the use of \emph{semi-gradient Sarsa} \cite[pp.\ \numrange{243}{244}]{RL2E}.
-Given a weight vector $\wvec{0}=\bm{0}$ which defines :
+Given a weight vector $\wvec{0}=\bm{0}$ and learning rate $\alpha$, we may approximate an action $a$'s value in any state $s$ by $\operatorname{q}(s, a) \approx \acval{s}{a}{\wvec{}}$:
 \begin{subequations}
@@
 \end{subequations}
+By computing the approximate value of every action available in the current state, actions with maximal value can be chosen each time (the \emph{greedy} policy), we might take a random action with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might some use some other mechanism.
 
-Observing that $\nabla{\acval{\cdot}{\cdot}{\cdot}} = \operatorname{\mathbf{x}}(s, a)$, and $\operatorname{\mathbf{x}}(s, a)$ .
-Furthermore, if the action set remains
+If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), then the implementation of the policy and update rules may be simplified by storing a vector of action values for each state encountered.
+Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, we may gain further optimisations by considering that a tile-coded feature vector is a boolean vector of constant Hamming weight (and so is amenable to representation as an array of indices).
+Moreover, we need not store vectors of action values for tiles which have not yet been visited.
 
-??
+%??
 Is this \emph{actually} just sarsa?
@@
 All links have unbounded capacity, save for the server-switch connection which is capped at a fixed $U_s$ \si{\mega\bit\per\second} in both directions.
+\Cref{fig:marl-topol} demonstrates this visually.
 
-?? Diagram? TikZ?
+\begin{figure}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tikzpicture}[
+texts/.style = {text=black},
+labeltexts/.style = {text=uofgsandstone},
+treeline/.style = {draw=uofgburgundy},
+treenode/.style = {texts, circle, centered, fill=white, treeline},
+external/.style = {fill=uofgrust},
+hideline/.style = {draw=uofgsandstone!40!white},
+hidenode/.style = {treenode, hideline},
+grow'=right
+]
+\node[treenode, label={[texts]above:Server}] (root) {}
+	child [treeline] { node [treenode, label={[texts]above:Core}] (sswitch) {}
+		child [treeline] { node [treenode, label={[texts]above:Leader}] (teaml) {} 
+			child [treeline] { node [treenode, label={[texts]above:Intermediate}] (inter) {}
+				child [treeline] { node [treenode, label={[texts]above:Agent/Egress}] (agent) {}
+					child [treeline] { node [treenode, external] (extern) {}
+						child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
+						child [hideline] { node [hidenode] (endhost) {} }
+					}
+				}
+				child [hideline] { node [hidenode] (endagent) {} }
+			}
+			child [hideline] { node [hidenode] (endinter) {} }
+		}
+		child [hideline] { node [hidenode] (endteaml) {} }
+		edge from parent
+		node[below, labeltexts] {$U_s$}
+	};
+
+%\draw[-] (teaml) -- (endteaml);
+\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
+\node [labeltexts, right = -0.1cm of kdots] {$k$};
+\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
+\node [labeltexts, right = -0.1cm of ldots] {$l$};
+\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
+\node [labeltexts, right = -0.1cm of mdots] {$m$};
+\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
+\node [labeltexts, right = -0.1cm of ndots] {$n$};
+\end{tikzpicture}
+}
+\caption{
+	Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $l$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
+	Unfilled nodes are considered to be internal.
+	\label{fig:marl-topol}
+}
+\end{figure}
 
@@
 
-Agents receive state input from monitors in the environment, a vector of load values (in \si{\mega\bit\per\second}) observed along their path to $s$: this is tile-coded with identical parameter choices to those of \textcite{DBLP:phd/ethos/Malialis14}.
+Agents receive state input from monitors in the environment, a vector of the four load values (in \si{\mega\bit\per\second}) observed along their path to $s$: this is tile-coded with identical parameter choices to those of \textcite{DBLP:phd/ethos/Malialis14}.
 
-At every time step, each agent installs an action onto the switch which hosts it, instructing it to drop $p\si{\percent}$ of traffic.
+At every time step, each agent installs an action via OpenFlow, instructing its host switch to drop each packet with probability $p$.
+Specifically, the agent chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
 
 The reward functions as I specify them here were introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, I make use of the global reward as a performance metric and the \emph{coordinated team learning} (CTL) reward for training.
-These assume the presence of a classifier (exact or heuristic) which estimates the volume of `legitimate traffic' received at any point, $g(\cdot)$.
-At time $t$ an agent $a$ (with team leader $l_a$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
+These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
+At time $t$ an agent $v$ (with team leader $l_v$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
 \begin{subequations}
@@
 \cond{f} = [\load{s} > U_s],\\
-\cond{c}{(a)} = [\load{l_a} > U_s/k],\\
+\cond{c}{(v)} = [\load{l_v} > U_s/k],\\
 R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\\
-R_{t}^{\mathit{CTL}} = (1 - \cond{f}\cond{c}{(a)}) \frac{g(\load{l_a})}{g(\load{\mathit{team}})} \allowbreak - \cond{f} \cond{c}{(a)},
+R_{t}^{\mathit{CTL}}(v) = (1 - \cond{f}\cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{f} \cond{c}{(v)},
 \end{gather}
-where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $a$'s team, respectively.
+where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.
 \end{subequations}
 
-The tile-coded state and chosen reward function are then used as input for the \emph{semi-gradient Sarsa} algorithm described in \cref{eqn:sg-sarsa}.
+The tile-coded state and chosen reward function are then used as input for the \emph{semi-gradient Sarsa} algorithm described in \cref{eqn:sg-sarsa}, making use of $\epsilon$-greedy action selection.
 
@@
 ?? wrt reward function relying on heuristic elements or perfect knowledge:
-A sufficiently well-trained agent needs only to follow the policy it has learned in a greedy manner, allowing pre-training using a priori knowledge by simulation to transfer to reality.
+A sufficiently well-trained agent needs only to follow the policy it has learned in a greedy manner, allowing pre-training by a simulated environment to transfer to reality.
 
@@
 
-??
-System specs of execution environment: 4-core Intel Core i7-6700K clocked at \SI{4.2}{\giga\hertz}, with \SI{32}{\gibi\byte} of RAM.
-All code is available on a public Github repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
+The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the accurate simulation of TCP-like behaviour (and its effects on collateral damage), disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), and assumes a rigid ISP-like topology.
 
-\begin{figure}[ht]
+All experiments described were executed using an Ubuntu 16.04.4 LTS machine with a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
+All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
+
+\subsection{Service Guarantees and Host Density}
+The .
+In practice, well-established network topologies such as the \emph{fat-tree topology} fan out internally (to provide efficient replication and scaling of resources) while having comparatively few egress points.
+It is this latter condition which I feel is most worth examining: the case where the host/agent ratio is heavily skewed in favour of hosts, and actions as they have been defined are forced to become less granular.
+
+The base `online' experiment I replicate configures the network topology using $k=5$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 7, 11, 14\}$ hosts per learner.
+I set $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$ and linearly annealing $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
+Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts are redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
+$U_s$ was fixed at $klmn+2\si{\mega\bit\per\second}$.
+The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting $\wvec{}=\bm{0}$ between episodes).
+Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
+
+\begin{figure}
 	\includegraphics[width=\linewidth]{../plots/online-varyN-uneven}
@@
 		This generalises for all behavioural discriminators: even a perfect agent \emph{must} punish benign flows if they are grouped with malicious actors.
+		%		?? Effect on max perf AND learning rate??
+		Crucially, both the maximum achievable performance and learning rate are negatively impacted when granularity is coarse.
 		\label{fig:marl-granularity}
@@
 
-\begin{figure}[ht]
+\Cref{fig:marl-granularity} shows that allocating more hosts to each learner in the recreated experiment leads to a decrease in the amount of benign traffic observed at the server.
+Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
+
+Notably, modelling the environment in Mininet makes evaluation rather slow as any interaction, coordination and communication must take place in real-time.
+While I'm sure this enhances the quality of model learned due to the explicit inclusion of more accurate network behaviour, a more abstract simulator based on observed physical properties could provide data at less computational cost.
+
+\subsection{TCP Multiplexing and Fallback Behaviour}
+While Tcpreplay was useful in recreating the original work, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
+To validate the operation of the pdrop OpenFlow action and to investigate the effects of collateral damage upon benign TCP flows, I conducted an iperf3 test using mininet.
+Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
+During the test, the rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured.
+
+\begin{figure}
 	\includegraphics[width=\linewidth]{../plots/mplex}
@@
 		As the rate of packet drop is linearly increased from \SIrange{0}{40}{\percent}, all flows begin to converge on an extremely low upload rate, regardless of the target bandwidth.
-		This behaviour is to be expected, and is to the credit of the emulated environment---prior treatments have ignored such phenomena, which have strong implications on agent and network design.
-		This confirms that packet drop does not affect benign TCP flows linearly, causing a higher rate reduction than what malicious flows will experience.
+		This behaviour is expected and well-understood, and is present in this emulated environment---prior treatments have ignored such phenomena, which have strong implications on agent and network design.
+		Malicious flows are not expected to obey the Mathis equation in this way, worsening collateral damage experienced by benign TCP flows.
 		\label{fig:mplex-tcp}
@@
 
-See above: a result.
-
-?? Observations: training time lengthier by nature of emulated environment.
-?? Risk of going for a simulation (i.e. numeric)? Always going to be interesting behaviour that is missed out on (in theory), at the cost of training time/limits of simulation speed
+\Cref{fig:mplex-tcp} shows that benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is significant.
+Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
+%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
+UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.
 
@@
 The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size and is an expected part of normal operation.
+
+\subsection{Synthesis}
+
+%It'd be really good if I could figure out a rough proof for the service property...
+%Okay... now say what this means for future viability! What changes can I make? Flow-specific metrics, of course! Increase granularity as much as I can!
+Both of these experiments suggest that the most viable way to enhance and deploy this MARL approach is to consider flows individually to maximise performance for legitimate hosts and to detect malicious hosts in a more principled manner.
+While this will add significant computational cost, this may be worked around with intelligent sampling and monitoring while taking multiple actions per-timestep \cite{DBLP:conf/hotnets/MaoAMK16}.
+
+?? Observations: training time lengthier by nature of emulated environment.
+?? Risk of going for a simulation (i.e. numeric)? Always going to be interesting behaviour that is missed out on (in theory), at the cost of training time/limits of simulation speed
+
+\section{A New Normal}
+
+?? TODO: everything
+
+\section{Rethinking the State Space}
+
+?? TODO: everything
+
+?? What concessions will we have to make in order to make per-flow processing more viable? Intelligent sampling/reanalysis of flows when needed (i.e. an external heuristic guiding method)?
+
+\section{The Results of Doing So}
+
+?? They'd probably look real nice...
 

### 2018-07-13 paper/paper.tex
Far more fleshed out. Gaps very rel.

@@
 \usepackage[binary-units, per-mode=symbol]{siunitx}
-\sisetup{range-phrase=--, range-units=single}
+\sisetup{detect-all, range-phrase=--, range-units=single}
 
@@
 \usepackage{booktabs}
-\usepackage{microtype}
+\usepackage[
+activate={true,nocompatibility},
+final,
+tracking=true,
+%kerning=true,spacing=true
+]{microtype}
+\microtypecontext{spacing=nonfrench}
 
@@
 These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
+This issue is inherent to \emph{misuse-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making use of statistical measures.
 
@@
 Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
-This is compounded by the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges of learning from unlabelled (often partial) data.
+Further issues arise from the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges of learning from unlabelled (often partial) data.
 All of these factors greatly inflate the difficulty of the detection problem.
 
-For certain classes of problem like non-transit DDoS attacks, \emph{reinforcement learning} (RL) is best-posed to adaptively react to certain classes of threat by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
+For certain classes of problem like flooding-based DDoS attacks, \emph{reinforcement learning} (RL) offers another perspective for adaptively reacting to certain classes of threat by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
 In a sense, this allows us to ``overcome'' the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time, by looking for the effect rather than the cause.
-The intent is to augment what existing solutions can provide, by automatically alerting, recording and controlling what are believed to be illegal system states.
+The intent is to augment what existing misuse-based solutions can provide, by automatically alerting, recording and controlling what are believed to be illegal system states.
 %Whether it takes direct control of the network, or is used indirectly to optimise a key part of another system, more powerful `deep' RL techniques (and well-founded action spaces) aren't yet well explored for network IDS/IPS.
@@
 \section{Background}
-
 %?? Introduce RL, related definitions etc.
@@
 We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
-This formulation of interaction with a system is best described as a \emph{Markov decision problem} (MDP).
+This formulation of interaction with a system is best described as a \emph{Markov decision process} (MDP).
 RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
@@
 In particular, we shall focus on methods which choose actions according to their value estimates from the current state.
-To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+|\mathcal{A}|)$-dimensional grids with an optional bias component.
-Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state action pair exists within the corresponding tile.
-This transform, as an example of linear function approximation, then necessitates the use of \emph{semi-gradient Sarsa} \cite[pp.\ \numrange{243}{244}]{RL2E}.
-Given a weight vector $\wvec{0}=\bm{0}$ and learning rate $\alpha$, we may approximate an action $a$'s value in any state $s$ by $\operatorname{q}(s, a) \approx \acval{s}{a}{\wvec{}}$:
+To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+ \dim{\mathcal{A}})$-dimensional grids with an optional bias component.
+Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state-action pair exists within the corresponding tile.
+This transform, as an example of linear function approximation, then necessitates the use of \emph{1-step semi-gradient Sarsa} \cite[pp.\ \numrange{243}{244}]{RL2E}.
+Given a weight vector $\wvec{0}=\bm{0}$ and learning rate $\alpha$, we may approximate an action $a$'s value in any state $s$ taking $\operatorname{q}(s, a) \approx \acval{s}{a}{\wvec{}}$ by continually updating $\wvec{t}$ as follows:
 \begin{subequations}
@@
 	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
-	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}}.
+	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
 	\end{gather}
 	\label{eqn:sg-sarsa}
+	taking $\nabla$ with respect to $\wvec{}$.
 \end{subequations}
-By computing the approximate value of every action available in the current state, actions with maximal value can be chosen each time (the \emph{greedy} policy), we might take a random action with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might some use some other mechanism.
+
+Computing the approximate value of every action available in the current state forms the basis of a policy.
+Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might some use some other mechanism.
 
 If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), then the implementation of the policy and update rules may be simplified by storing a vector of action values for each state encountered.
-Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, we may gain further optimisations by considering that a tile-coded feature vector is a boolean vector of constant Hamming weight (and so is amenable to representation as an array of indices).
-Moreover, we need not store vectors of action values for tiles which have not yet been visited.
+Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, we may gain further optimisations by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices).
+Action values are then simply the sum of all activated tile weight vectors.
+Moreover, we need not store action values for tiles which have not yet been visited.
 
@@
 
-\section{A Plan, of Sorts}
-
-\begin{enumerate}
-	\item The main case for contribution in what I have so far:
-	\begin{itemize}
-		\item Past work reliant on unrealistic network models: tcp-like behaviour (and its effects on collateral damage) not captured, disjoint ranges of traffic distribution (no benign heavy-hitters), ISP-like topology.
-		\item I offer more realistic network emulation environment, better treatment of protocol/traffic characteristics.
-	\end{itemize}
-	\item Forthcoming: rethinking state/action spaces to operate at a finer level of granularity.
+%\section{A Plan, of Sorts}
+%
+%\begin{enumerate}
+%	\item The main case for contribution in what I have so far:
+%	\begin{itemize}
+%		\item Past work reliant on unrealistic network models: tcp-like behaviour (and its effects on collateral damage) not captured, disjoint ranges of traffic distribution (no benign heavy-hitters), ISP-like topology.
+%		\item I offer more realistic network emulation environment, better treatment of protocol/traffic characteristics.
+%	\end{itemize}
+%	\item Forthcoming: rethinking state/action spaces to operate at a finer level of granularity.
 New network model (live tcp back-and-forth), allows us to test collateral damage assumptions in a more realistic manner (and show clear case for moving beyond work of malialis and kudenko)
-\end{enumerate}
+%\end{enumerate}
 
@@
 	Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $l$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
-	Unfilled nodes are considered to be internal.
+	Empty nodes are considered to be internal.
 	\label{fig:marl-topol}
@@
 Specifically, the agent chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
+In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
 
@@
 The tile-coded state and chosen reward function are then used as input for the \emph{semi-gradient Sarsa} algorithm described in \cref{eqn:sg-sarsa}, making use of $\epsilon$-greedy action selection.
+Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their experience or weight vector updates with one another.
 
@@
 \section{Performance in an Emulated Environment}
-
-The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the accurate simulation of TCP-like behaviour (and its effects on collateral damage), disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), and assumes a rigid ISP-like topology.
+The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and assumes a rigid ISP-like topology.
+The latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
 
@@
 
+?? TODO: Make repo public...
+
 \subsection{Service Guarantees and Host Density}
-The .
-In practice, well-established network topologies such as the \emph{fat-tree topology} fan out internally (to provide efficient replication and scaling of resources) while having comparatively few egress points.
-It is this latter condition which I feel is most worth examining: the case where the host/agent ratio is heavily skewed in favour of hosts, and actions as they have been defined are forced to become less granular.
+In the default topology, $n=2$ hosts per egress point is not a parameter that can be described as realistic for many web service deployments.
+In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
+I feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions as defined become less granular.
 
-The base `online' experiment I replicate configures the network topology using $k=5$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 7, 11, 14\}$ hosts per learner.
-I set $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$ and linearly annealing $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
-Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts are redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
-$U_s$ was fixed at $klmn+2\si{\mega\bit\per\second}$.
-The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting $\wvec{}=\bm{0}$ between episodes).
+To test this, I configured the network topology using $k=2$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 7, 11, 14\}$ hosts per learner.
+I set $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$ and linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
+Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
+$U_s$ was fixed at $klmn+2$ \si{\mega\bit\per\second}.
+The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
 Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
@@
 	\caption{
-		MARL pushback control system performance plotted over time, for various settings of $n$ hosts per learner.
+		MARL pushback control system performance plotted over time, for various settings of $n$ hosts per agent.
 		This plot shows that, from the perspective of benign hosts, service guarantees degrade as inference and applied actions become less granular.
@@
 
-\Cref{fig:marl-granularity} shows that allocating more hosts to each learner in the recreated experiment leads to a decrease in the amount of benign traffic observed at the server.
+\Cref{fig:marl-granularity} shows that allocating more hosts to each learner leads to a guaranteed decrease in the amount of benign traffic observed at the server.
+This manifests in two ways: the best-achievable performance decreases, as does the learning rate.
 Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
 
-Notably, modelling the environment in Mininet makes evaluation rather slow as any interaction, coordination and communication must take place in real-time.
-While I'm sure this enhances the quality of model learned due to the explicit inclusion of more accurate network behaviour, a more abstract simulator based on observed physical properties could provide data at less computational cost.
-
 \subsection{TCP Multiplexing and Fallback Behaviour}
-While Tcpreplay was useful in recreating the original work, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
-To validate the operation of the pdrop OpenFlow action and to investigate the effects of collateral damage upon benign TCP flows, I conducted an iperf3 test using mininet.
+While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
+To investigate the effects of collateral damage upon benign TCP flows, an iperf3 test was conducted using mininet.
 Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
@@
 
-\Cref{fig:mplex-tcp} shows that benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is significant.
+\Cref{fig:mplex-tcp} shows that benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
 Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
@@
 
-?? TCP perf characteristics under packet drop known by the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}:
+The performance characteristics of TCP under packet drop  are well-known and well-understood through the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}:
 \begin{equation}
@@
 \end{equation}
-for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions and the probability $p \in (0, 1]$ that a packet is lost.
-The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size and is an expected part of normal operation.
+for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions, and the probability $p \in (0, 1]$ that a packet is lost.
+%The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size and is an expected part of normal operation.
+TCP flows in Mininet are found to obey this law, as expected.
+This reveals that the original work \emph{does not capture this behaviour}, or else the initial experiment (UDP-like traffic when $n=2$) wouldn't have achieved similar results to those which were reported.
 
-\subsection{Synthesis}
-
-%It'd be really good if I could figure out a rough proof for the service property...
-%Okay... now say what this means for future viability! What changes can I make? Flow-specific metrics, of course! Increase granularity as much as I can!
-Both of these experiments suggest that the most viable way to enhance and deploy this MARL approach is to consider flows individually to maximise performance for legitimate hosts and to detect malicious hosts in a more principled manner.
-While this will add significant computational cost, this may be worked around with intelligent sampling and monitoring while taking multiple actions per-timestep \cite{DBLP:conf/hotnets/MaoAMK16}.
-
-?? Observations: training time lengthier by nature of emulated environment.
-?? Risk of going for a simulation (i.e. numeric)? Always going to be interesting behaviour that is missed out on (in theory), at the cost of training time/limits of simulation speed
+%\subsection{Synthesis}
+%%It'd be really good if I could figure out a rough proof for the service property...
+%%Okay... now say what this means for future viability! What changes can I make? Flow-specific metrics, of course! Increase granularity as much as I can!
+%Both of these experiments suggest that the most viable way to enhance and deploy this MARL approach is to consider flows individually to maximise performance for legitimate hosts and to detect malicious hosts in a more principled manner.
+%While this will add significant computational cost, this may be worked around with intelligent sampling and monitoring while taking multiple actions per-timestep \cite{DBLP:conf/hotnets/MaoAMK16}.
+%
+%?? Observations: training time lengthier by nature of emulated environment.
+%?? Risk of going for a simulation (i.e. numeric)? Always going to be interesting behaviour that is missed out on (in theory), at the cost of training time/limits of simulation speed
 
@@
 ?? Other recent work in the prevention of DDoS? (i.e., non-RL).
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} had it as oart of their evaluation, BGP-like rerouting techniques (specifically for transit-link attacks, but also for `normal' attacks) \cite{RoutingAroundCongestion}
 
@@
 
-?? Good Direct-control work \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}
+For categorising how RL fits into solving problems, I label works as direct- or indirect-control RL.
+A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
+?? Good Direct-control work? What this work is based on (which directly captures the spirit of ) \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}
 
-?? Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works. \cite{DBLP:conf/hotnets/ValadarskySST17, DBLP:conf/hotnets/MaoAMK16}.
-The HMM stuff---learning when best to \emph{communicate} and share knowledge between explicit detector models \cite{DBLP:conf/paisi/XuSH07}.
-The latter example's position is slightly weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
+An \emph{indirect-control} RL problem, where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
+%Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works. \cite{DBLP:conf/hotnets/ValadarskySST17, DBLP:conf/hotnets/MaoAMK16}.
+A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
+The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
+Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
 
@@
 
-It all ends here, Mr.\ Bond...
+It all ends here...
 
@@
 It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques.
-Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
+THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
 
 ??
-Overarching goal: works which respect the complexity of the network environment (rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution.
-Call-to-action?
+Overarching goal: works which \emph{respect the complexity of the network environment} (rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution.
+\emph{Call-to-action}?
 
@@
 System state which is dependent on many signals drawn from across a wide network, in concert with their burstiness and unpredictability, may have substantial effects on an attacker's capabilities.
+??
+Might want to mention it in related work above, but the recent attention on adversarial examples/tricking models needs to be looked into for RL.
+Poisoning attacks relevant for online techniques: old bounds exist \textcite{DBLP:journals/jmlr/KloftL10}, new stuff concerns collaborative learners \cite{DBLP:conf/acsac/ShenTS16}, nothing for rl.
+Hot topic in deep networks \cite{DBLP:conf/eurosp/PapernotMJFCS16, SpInMl}, but naturally still relevant with even linear approximations or exact tabular case due to limits of the PAC assumption.
+??
+evasion attacks by \textcite{DBLP:conf/sp/Carlini017}---all of these are computed by way of a general stochastic optimiser, such as \emph{Adam} \cite{DBLP:journals/corr/KingmaB14}.
+possible to apply something similar to our learned model to assess its security?
+would the suggested states even be valid?
+(i.e.
+since they're monotonically increasing for the most part).
 

### 2018-08-07 code/marl/marl.py
Fixes + rebalanced experiments.
Binary progression makes more sense than arbitrary choices for n, some say.

@@ -306,7 +306,8 @@ def marlExperiment(
 #			ofpb.ofp_match(None, None, None),
 			ofpb.ofp_match(ofp.OFPMT_OXM, None, [
 				ofpm.build(None, ofp.OFPXMT_OFB_ETH_TYPE, False, 0, 0x0800, None),
-				ofpm.build(None, ofp.OFPXMT_OFB_IPV4_DST, False, 0, socket.inet_aton(ip), None)
+				#ofpm.build(None, ofp.OFPXMT_OFB_IPV4_DST, False, 0, socket.inet_aton(ip), None)
+				ofpm.build(None, ofp.OFPXMT_OFB_IPV4_DST, True, 0, netip(ip, subnet), socket.inet_aton(subnet))
 			]),
 			ofpb.ofp_instruction_actions(ofp.OFPIT_WRITE_ACTIONS, None, [
 				# Looks like 29 is the number I picked for Pdrop.

### 2018-08-07 code/marl/online-expr-16.py

@@ -1,0 +1,32 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 16
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+
+	dt = 0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-16.csv", results)
+
+makeResultsAverage("../../results/online-16.csv", "../../results/online-16-avg.csv")
+

### 2018-08-07 code/marl/online-expr-8.py

@@ -1,0 +1,32 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 8
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+
+	dt = 0.05,#0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-8.csv", results)
+
+makeResultsAverage("../../results/online-8.csv", "../../results/online-8-avg.csv")
+

### 2018-08-07 code/marl/online-expr-32.py
32-hpl results

@@ -1,0 +1,32 @@
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 32
+
+results = marlExperiment(
+	n_teams = 2,#5,
+
+	n_inters = 2,
+	n_learners = 3,
+	host_range = [hosts_p, hosts_p],
+
+	explore_episodes = 0.3,
+	episodes = 10,#500, Since mininet keeps running out of files even e/ cleanup
+	episode_length = 10000,
+	separate_episodes = True,
+
+	alpha = 0.05,
+	epsilon = 0.2,
+	discount = 0,
+
+	dt = 0.01,
+
+	rf = "ctl",
+
+	rand_seed = 0xcafed00d
+)
+
+writeResults("../../results/online-32.csv", results)
+
+makeResultsAverage("../../results/online-32.csv", "../../results/online-32-avg.csv")
+

### 2018-08-07 paper/paper.tex
Report enhancements.

@@
 
+\usepackage[british]{babel}
+\usepackage{csquotes}
+
 \usepackage{booktabs}
@@
 %bib
-\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,style=ieee,sortcites,backend=bibtex]{biblatex}
+\usepackage[style=ieee,maxnames=3,mincitenames=1,maxcitenames=2,maxbibnames=99,mincrossrefs=5,sortcites,backend=bibtex,uniquelist=false]{biblatex}
 \addbibresource{papers-off.bib}
@@
 %picky abt et al.
-\usepackage{xpatch}
+%\usepackage{xpatch}
 
-\xpatchbibmacro{name:andothers}{%
-	\bibstring{andothers}%
-}{%
-	\bibstring[\emph]{andothers}%
-}{}{}
+%\xpatchbibmacro{name:andothers}{%
+%	\bibstring{andothers}%
+%}{%
+%	\bibstring[\emph]{andothers}%
+%}{}{}
+
+% emph'd et al. for ieee style
+\DefineBibliographyStrings{english}{%
+	andothers = {\emph{et al}\adddot}
+}
 \DeclareFieldFormat[inproceedings]{url}{}
@@
 	pdftitle={\mytitle{}},
-	pdfauthor={Kyle A. Simpson}
+	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros}
 }
@@
 \title{\mytitle{}}
-\author{Kyle A. Simpson\thanks{This work was supported by the Engineering and Physical Sciences
-		Research Council [grant number EP/M508056/1]}\\University of Glasgow, Glasgow, Scotland,\\
-		\email{k.simpson.1@research.gla.ac.uk}}
+\author{
+	Kyle A.\ Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
+	Dimitrios P.\ Pezaros\\
+	University of Glasgow, Glasgow, Scotland,\\
+	\email{k.simpson.1@research.gla.ac.uk}
+}
 
@@
 These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
-This issue is inherent to \emph{misuse-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making use of statistical measures.
+This issue is inherent to \emph{misuse-} or \emph{signature-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making use of statistical measures.
 
@@
 \begin{itemize}
-	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks.
-	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction of their ISP-like formulation, and an acknowledgement of the flaws and risk factors of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
-	\item ?? TODO: Advance the use/representation/granularity of state-space in direct-control RL of networked applications.
+	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:rethinking-the-state-space}).
+	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction of their formulation, and an acknowledgement of the flaws and risk factors of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a} (\cref{sec:performance-in-an-emulated-environment}).
+	\item ?? TODO: Advance the use/representation/granularity of state-space in direct-control RL of networked applications (\cref{sec:rethinking-the-state-space}).
 \end{itemize}
@@
 Computing the approximate value of every action available in the current state forms the basis of a policy.
-Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might some use some other mechanism.
+Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might use some other mechanism.
 
-If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), then the implementation of the policy and update rules may be simplified by storing a vector of action values for each state encountered.
-Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, we may gain further optimisations by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices).
-Action values are then simply the sum of all activated tile weight vectors.
-Moreover, we need not store action values for tiles which have not yet been visited.
+If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties of the problem allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
+Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices).
+Action values for any state are then obtained by summation of the weight vectors from all activated tiles.
+Moreover, we need not store action values for tiles which have not yet been visited, conserving memory.
 
@@
 It's, at the very least, 1-step semi-gradient Sarsa given that it is clearly on-policy w/ fn approx...
-?? IDEA: try out average reward, TD($\lambda$) methods as future work...
 
 ?? Discuss mininet? Networking terms? SDN stuff?
+
+?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
 
@@
 
-\section{Environment and RL Algorithm}
+\section{Environment and RL Algorithm}\label{sec:environment-and-rl-algorithm}
 
-?? STRUCTURE: Either (methodology) -> (results) OR (their stuff + its weaknesses) -> (a new direction) [second is way punchier, but not yet ready to go with that.]
+?? Probably rethink the title of this section.
+
+%?? State the problem and what we want to achieve.
+To best discover how the use of RL techniques in the field of network intrusion prevention may bet be improved, it's prudent that we start by looking for weaknesses in work which exists today.
+The closest approach which is available within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant:  their work helped to show the viability of live, adaptive, feedback loop-like control of the network to detect and prevent DDoS attacks.
+We begin with a reimplementation and more principled re-examination of their work to further develop the field.
+
+%?? Explicitly say the ``REIMPLEMENTING M\&K THING'' over here, to make later paras make a reasonable amount of sense.
+
+%??
+Why are we comparing against/building on/reimplementing M\&K?
+What was their main contribution?
+Answer: they helped to show the viability of live, adaptive control of the network to detect/prevent DDoS attacks...
 
 \subsection{Topology}
+
 The network itself is tree-structured, where one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $l$ intermediate switches, which in turn each connect to $m$ egress switches.
@@
 
+Why use this topology, a fan-out tree, in particular?
+While this is taken verbatim from the work we use as a basis, it is presented within the confines of a framework given by \textcite{DBLP:journals/ton/YauLLY05}.
+This framework is, in turn, an extension of the class of topologies examined in the original pushback paper \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
+In actuality, this topology is far closer to the latter; crucially, it seems to not actually represent or correspond to any realistic deployment at all.
+In the words of \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}:
+\begin{displaycquote}{DBLP:journals/ccr/MahajanBFIPS02a}
+	These simulations do not pretend to use realistic topologies or traffic mixes... the simple simulations in this scenario are instead intended... as a first step towards more rigorous evaluation.
+\end{displaycquote}
+%The topology chosen to evaluate our predecessor, indeed, matches theirs.
+%?? So, why does this proliferate, exist? They effectively disavowed it in their own paper. Reading comprehension is hard, I guess.
+Remarkably, it seems to have propagated in a way which was not originally intended.
+It is clear that such a network topology allows for functional testing, but it is hard to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
+We aim to address this, to some extent.
+
 Reimplementation of the network model was performed using mininet \cite{mininet}, a Python-based network emulator.
@@
 In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
+It's worth noting that there are various ways that this could be done, and that the application of \emph{programmable data planes} to this end are suggested as future work.
 
-The reward functions as I specify them here were introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, I make use of the global reward as a performance metric and the \emph{coordinated team learning} (CTL) reward for training.
+The reward functions as specified here were introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of the global reward as a performance metric and the \emph{coordinated team learning} (CTL) reward for training.
 These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
@@
 
-?? M\&K Seem not to have a grasp on the importance of up/down traffic splits. Hmm......... (i.e., both could be relevant?)
+%?? M\&K Seem not to have a grasp on the importance of up/down traffic splits. Hmm......... (i.e., both could be relevant?) Man, who knows.
 
@@
 
-\section{Performance in an Emulated Environment}
+\section{Performance in an Emulated Environment}\label{sec:performance-in-an-emulated-environment}
 The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and assumes a rigid ISP-like topology.
@@
 In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
-I feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions as defined become less granular.
+We feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions as defined become less granular.
 
-To test this, I configured the network topology using $k=2$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 7, 11, 14\}$ hosts per learner.
-I set $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$ and linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
+To test this, we configured the network topology using $k=2$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
+The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
 Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
@@
 \begin{figure}
-	\includegraphics[width=\linewidth]{../plots/online-varyN-uneven}
+	\includegraphics[width=\linewidth]{../plots/online-varyN-binary}
 	\caption{
@@
 \Cref{fig:marl-granularity} shows that allocating more hosts to each learner leads to a guaranteed decrease in the amount of benign traffic observed at the server.
+?? TODO: rephrase.
 This manifests in two ways: the best-achievable performance decreases, as does the learning rate.
 Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
+?? Expand according to the summary you posted in the notes document.
 
@@
 
+?? Full Openflow controller setup (in mininet)
 ?? TODO: everything
 
-\section{Rethinking the State Space}
+??
+ANGLE: set up an environment to test \emph{specific} behaviours to examine \emph{specific} problems in past work.
+I make no claims that it is perfect or representative for all traffic, just for this (likely common) behaviour which I expect to plague almost all legit TCP flows.
+
+\section{Rethinking the State Space}\label{sec:rethinking-the-state-space}
 
@@
 ?? What concessions will we have to make in order to make per-flow processing more viable? Intelligent sampling/reanalysis of flows when needed (i.e. an external heuristic guiding method)?
+
+??
+NOTE: think about how I might need to change/alter the tile-coding when we add extra discriminative features...
+(i.e., are there better ways to handle covariance between properties of system/flow state, properly capture generalisation?)
 
@@
 ?? Other recent work in the prevention of DDoS? (i.e., non-RL).
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} had it as oart of their evaluation, BGP-like rerouting techniques (specifically for transit-link attacks, but also for `normal' attacks) \cite{RoutingAroundCongestion}
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} had it as part of their evaluation, BGP-like rerouting techniques (specifically for transit-link attacks, but also for `normal' attacks) \cite{RoutingAroundCongestion}
 
@@
 Past work exhibits abuses of the paradigm as a glorified classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
+?? Explain/be less incendiary...
 
-For categorising how RL fits into solving problems, I label works as direct- or indirect-control RL.
+For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
 A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
@@
 
+?? IDEA: try out average reward, TD($\lambda$) methods as future work...
+
 ??
-Benefit of the more realistic emulation environ is that it is far closer in behaviour and architecture (i.e.
+IDEA: Apply these techniques to programmable data planes etc.
+While it's pretty neat that what we have works assuming that ach router is a software (x86) switch running OVS, what might we need to consider when applying this to `real' switches?
+``PDP can allow this to be added to real routers to make it efficient to keep \& process state in the manner we require, as well as enabling more adaptive deployment''.
+Cite P4, BPFabric, other work on PDPs?
+
+??
+Benefit of the more realistic emulation environment is that it is far closer in behaviour and architecture (i.e.
 viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance.
@@
 Poisoning attacks relevant for online techniques: old bounds exist \textcite{DBLP:journals/jmlr/KloftL10}, new stuff concerns collaborative learners \cite{DBLP:conf/acsac/ShenTS16}, nothing for rl.
-Hot topic in deep networks \cite{DBLP:conf/eurosp/PapernotMJFCS16, SpInMl}, but naturally still relevant with even linear approximations or exact tabular case due to limits of the PAC assumption.
+Hot topic in deep networks \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18}, but naturally still relevant with even linear approximations or exact tabular case due to limits of the PAC assumption.
+There is now examination of evasion attacks wrt.\ RL \cite{DBLP:journals/corr/HuangPGDA17}!
 ??
@@
 \section*{Acknowledgements}
-My thanks go to Dr.\ Dimitrios Pezaros for his supervision, guidance, and comments relating to the realisation of this work, and to Mircea Iordache and Richard Cziva for their comments and technical assistance.
+My thanks go to Mircea Iordache and Richard Cziva for their comments and technical assistance.
 Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.

### 2018-08-07 code/traffic-host/src/config.rs
More initial work on http client thing.

@@ -1,0 +1,5 @@
+#[derive(Debug)]
+pub struct Config {
+    
+}
+

### 2018-08-07 code/traffic-host/src/lib.rs

@@ -1,8 +1,17 @@
 extern crate reqwest;
 extern crate tokio;
 
-pub fn run() {
+mod config;
+
+pub use config::Config;
+
+pub fn run(options: config::Config) {
 	println!("Hello, world!");
+
+    println!("{:?}", options);
+
+    // TODO: swap this over to use libcurl since that's nicer
+    // and lets me ratelimit etc in a way that hyper doesn't. :)
 
 	match reqwest::get("https://mcfelix.me") {
 		Ok(mut resp) => {

### 2018-08-07 code/traffic-host/src/main.rs

@@ -1,5 +1,44 @@
+extern crate clap;
 extern crate traffic_host;
 
+use clap::{Arg, App};
+
 fn main() {
-    traffic_host::run();
+    let matches =
+        App::new("MARL Traffic Generator")
+            .version("0.10")
+            .author("Kyle Simpson <k.simpson.1@research.gla.ac.uk>")
+            .about("Traffic generator (HTTP, others) for testing MARL.")
+            .arg(Arg::with_name("server")
+                 .short("s")
+                 .long("server")
+                 .value_name("SERVER")
+                 .help("Server to send requests to.")
+                 .takes_value(true)
+                 .default_value("10.0.0.1"))
+            .arg(Arg::with_name("http-dir")
+                 .short("h")
+                 .long("http-dir")
+                 .value_name("PATH")
+                 .help("Local directory tree to use as a basis for queries.")
+                 .takes_value(true)
+                 .default_value("htdocs"))
+            .arg(Arg::with_name("MAX_DOWN")
+                 .help("Maximum download rate from the target server.")
+                 .required(true)
+                 .index(1))
+            .arg(Arg::with_name("MAX_UP")
+                 .help("Maximum upload rate to the target server.")
+                 .index(2))
+            .get_matches();
+
+    // e.g., (before conversion to a real URL/address for library use)
+    let _url = matches.value_of("server")
+        .expect("Server URL always guaranteed to exist...");
+
+    let config = traffic_host::Config {
+
+    };
+
+    traffic_host::run(config);
 }

### 2018-08-08 paper/paper.tex
More ideas + refs

@@
 
+??
+Probably want to talk about NIDS/IPS, DDoS attack variants (leading into characteristics, supporting features).
+Amplification (UDP \cite{DBLP:conf/ndss/Rossow14}, TCP \cite{DBLP:conf/uss/KuhrerHRH14}), Transit-link (Crossfire \cite{DBLP:conf/sp/KangLG13}, Coremelt \cite{DBLP:conf/esorics/StuderP09}).
+Mirai botnet's involvement \cite{DBLP:conf/uss/AntonakakisABBB17}.
+
 ?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
+??
+Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows.
+This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e.
+an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
+
+??
+Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
+?? Need to find a good survey which suggests indicative features.
+?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
 
@@
 To test this, we configured the network topology using $k=2$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
+This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single-server with a fan-out network.
 The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
@@
 
-\Cref{fig:marl-granularity} shows that allocating more hosts to each learner leads to a guaranteed decrease in the amount of benign traffic observed at the server.
-?? TODO: rephrase.
+\Cref{fig:marl-granularity} shows that as more hosts are allocated to each learner, the fraction of traffic believed to be good (as observed at the server) decreases.
 This manifests in two ways: the best-achievable performance decreases, as does the learning rate.
 Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
-?? Expand according to the summary you posted in the notes document.
+From the findings of \cref{sec:tcp-multiplexing-and-fallback-behaviour} and the Mathis equation, their results should be considerably worse in practice.
+%?? Expand according to the summary you posted in the notes document.
+%That the use of UDP-like traffic replicates their experiments, compared against how real tcp behaves in mininet and according to the Mathis equation, shows that they really did not consider the behaviour of benign TCP in this environ (i.e., their results should be considerably worse in practice).
 
-\subsection{TCP Multiplexing and Fallback Behaviour}
+\subsection{TCP Multiplexing and Fallback Behaviour}\label{sec:tcp-multiplexing-and-fallback-behaviour}
 While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
@@
 Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
-During the test, the rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured.
+The rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured, to show a trend and determine for the mininet environment whether throughput would decrease as expected, at a super-linear rate.
 
@@
 	\caption{
-		Average TCP upload rate of hosts targeting different bandwidths in \emph{mininet}, computed over 10 runs.
-		As the rate of packet drop is linearly increased from \SIrange{0}{40}{\percent}, all flows begin to converge on an extremely low upload rate, regardless of the target bandwidth.
-		This behaviour is expected and well-understood, and is present in this emulated environment---prior treatments have ignored such phenomena, which have strong implications on agent and network design.
+		Average TCP upload rate of hosts targeting different bandwidths in \emph{mininet}, computed over 10 runs, for choices of packet drop rate up until throughput is reduced to almost \SI{0}{\mega\bit\per\second}.
+		As the rate of packet drop increases from \SIrange{0}{40}{\percent}, all flows converge on an extremely low throughput, regardless of the target bandwidth---a super-linear decrease for legitimate TCP traffic.
+		Prior treatments have ignored such phenomena, which have strong implications on agent and network design.
 		Malicious flows are not expected to obey the Mathis equation in this way, worsening collateral damage experienced by benign TCP flows.
@@
 
-\Cref{fig:mplex-tcp} shows that benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
+\Cref{fig:mplex-tcp} shows that this trend is present.
+Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
 Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
@@
 
+??
+Existing sims used for testing such applications reliant on traces, or not sophisticated enough to capture interactive, back-and-forth (correlated) behaviours---possibly discarded as second-hand effects by past work when these are so crucial given user traffic patterns (and the nature of the control signal we choose to enact).
+
 \section{Rethinking the State Space}\label{sec:rethinking-the-state-space}
@@
 Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
-Past work exhibits abuses of the paradigm as a glorified classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-?? Explain/be less incendiary...
+Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
+Given that arguably the main strength of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
 
@@
 A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
-?? Good Direct-control work? What this work is based on (which directly captures the spirit of ) \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}
+?? Good Direct-control work? What this work is based on (which directly captures the spirit of RL) \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}
 
@@
 Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+?? Also adaptively choosing video bitrates based on network conditions \cite{DBLP:conf/sigcomm/MaoNA17}.
 These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.

### 2018-08-28 paper/paper.tex
More background, use nice new plots.

@@
 \subsection{TCP Multiplexing and Fallback Behaviour}\label{sec:tcp-multiplexing-and-fallback-behaviour}
+
+?? TODO: Gut this section now...
 While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
@@
 
+%\begin{figure}
+%	\includegraphics[width=\linewidth]{../plots/mplex}
+%	\caption{
+%		Average TCP upload rate of hosts targeting different bandwidths in \emph{mininet}, computed over 10 runs, for choices of packet drop rate up until throughput is reduced to almost \SI{0}{\mega\bit\per\second}.
+%		As the rate of packet drop increases from \SIrange{0}{40}{\percent}, all flows converge on an extremely low throughput, regardless of the target bandwidth---a super-linear decrease for legitimate TCP traffic.
+%		Prior treatments have ignored such phenomena, which have strong implications on agent and network design.
+%		Malicious flows are not expected to obey the Mathis equation in this way, worsening collateral damage experienced by benign TCP flows.
+%		\label{fig:mplex-tcp}
+%	}
+%\end{figure}
+
 \begin{figure}
-	\includegraphics[width=\linewidth]{../plots/mplex}
+	\includegraphics[width=\linewidth]{../plots/online-varyN-nginx}
 	\caption{
-		Average TCP upload rate of hosts targeting different bandwidths in \emph{mininet}, computed over 10 runs, for choices of packet drop rate up until throughput is reduced to almost \SI{0}{\mega\bit\per\second}.
-		As the rate of packet drop increases from \SIrange{0}{40}{\percent}, all flows converge on an extremely low throughput, regardless of the target bandwidth---a super-linear decrease for legitimate TCP traffic.
-		Prior treatments have ignored such phenomena, which have strong implications on agent and network design.
-		Malicious flows are not expected to obey the Mathis equation in this way, worsening collateral damage experienced by benign TCP flows.
-		\label{fig:mplex-tcp}
+		A more convincing and/or to-the-point figure for getting people to believe I'm doing god's work.
+		\label{fig:nginx-coffin-nail}
 	}
@@
 
-\Cref{fig:mplex-tcp} shows that this trend is present.
+%\Cref{fig:mplex-tcp} shows that this trend is present.
+\Cref{fig:nginx-coffin-nail} shows that this trend is present.
 Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
@@
 ?? Also adaptively choosing video bitrates based on network conditions \cite{DBLP:conf/sigcomm/MaoNA17}.
+
+?? Completely reword, yikes.
+\citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s \emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} uses deep RL (the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}).
+Crucially, they find that decisions need to be made for the vast majority of (short-lived) flows, sub-millisecond.
+Their work falls between the above two paradigms: to overcome the high latency of action computation via neural network, two agents are trained.
+The first learns to optimise flow size thresholds governing priority (and accordingly demarcate long and short flows).
+All short flows are routed by ECMP.
+The second agent makes decisions about routing, prioritisation etc.\ for these long flows, who are likely to live long enough that any decisions made will have a significant impact.
+
 These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.

### 2018-08-30 code/marl/marl.py
Fix pseudorandom allocation of states.

@@ -84,7 +84,7 @@ def marlExperiment(
 
 		rf = "ctl",
 
-		rand_seed = None,
+		rand_seed = 0xcafed00d,
 		rand_state = None,
 		force_cmd_routes = False,
 
@@ -509,6 +509,7 @@ def marlExperiment(
 		for i in xrange(host_count):
 			good = random.random() < P_good
 			bw = (random.uniform(*(good_range if good else evil_range)))
+			print "drew: good={}, bw={}".format(good, bw)
 
 			# Make up a wonderful IP.
 			# Last byte => goodness. Even = good.

### 2018-08-30 paper/paper.tex
More paper work.

@@
 \titlespacing*{\subsection}{0pt}{1.5ex}{0.7ex}
+\titlespacing*{\paragraph}{0pt}{1.5ex}{0.7ex}
 %\titleformat*{\filcenter\scshape}
+
+\usepackage{enumitem}
+\setlist[description]{leftmargin=8em,style=nextline}
 
@@
 \section{Performance in an Emulated Environment}\label{sec:performance-in-an-emulated-environment}
-The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and assumes a rigid ISP-like topology.
+The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and assumes a rigid and unrepresentative topology.
 The latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
@@
 ?? TODO: Gut this section now...
-While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
-To investigate the effects of collateral damage upon benign TCP flows, an iperf3 test was conducted using mininet.
-Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
-The rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured, to show a trend and determine for the mininet environment whether throughput would decrease as expected, at a super-linear rate.
+%While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
+%To investigate the effects of collateral damage upon benign TCP flows, an iperf3 test was conducted using mininet.
+%Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
+%The rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured, to show a trend and determine for the mininet environment whether throughput would decrease as expected, at a super-linear rate.
 
@@
 		A more convincing and/or to-the-point figure for getting people to believe I'm doing god's work.
+		?? TODO: describe 4 real.
 		\label{fig:nginx-coffin-nail}
@@
 
+?? Quic will probably behave much like TCP.
+
 The performance characteristics of TCP under packet drop are well-known and well-understood through the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}:
@@
 %The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size and is an expected part of normal operation.
-TCP flows in Mininet are found to obey this law, as expected.
+%TCP flows in Mininet are found to obey this law, as expected.
 This reveals that the original work \emph{does not capture this behaviour}, or else the initial experiment (UDP-like traffic when $n=2$) wouldn't have achieved similar results to those which were reported.
+
+\Cref{sec:a-new-normal} is where the new normal model arrives. Wow!
+
+\subsection{Computational Cost}
+??
+Consider talking about the execution times of the old MARL approach here?
+They're real nice (as expected), so we have lots of room to play around with while (hopefully) remaining under the 1ms target time given by \textcite{DBLP:conf/sigcomm/ChenL0L18}.
+
+We're in \SIrange{80}{100}{\micro\second}
 
@@
 
-\section{A New Normal}
+\section{A New Normal}\label{sec:a-new-normal}
 
-?? Full Openflow controller setup (in mininet)
-?? TODO: everything
+In establishing...
+
+?? How will I structure this?
+?? Motivation -> Model -> Results?
+?? OR Use the results of the last section to springboard into here?
+
+??
+Remember, the motivation is clear.
+We don't care so much that it is "representative" wrt a specific deployment location or network type.
+The whole purpose of this is that we aim to test specific behaviour which traces cannot replicate (i.e., correlated back-and-forth, dynamics introduced to congestion-aware protocols, ...)
+?? If we need to, we can codition the distribution of requests according to statistics mined form an existing trace if reviewer number 2 needs that extra push to be convinced.
+
+??
+Full Openflow controller setup (in mininet).
+All internal routers are primed with knowledge of the shortest path to any internal host, while new inbound flows register the ``way back'' for each hop used to maintain consistency over traffic conditions.
+If such information is lost, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
+The controller is also responsible for responding to ARP traffic, as a gateway might
+The necessity of this arises due to the realities of relying upon Linux's IP stack, and would not need to be considered for purely trace-based evaluation.
+?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
+?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
+
+\Cref{fig:nginx-coffin-nail} shows the effects of using
+
+?? Actually describe what I have here, now. I can do this.
 
@@
 
+?? What do we want?
+
+\begin{IEEEdescription}[\IEEEsetlabelwidth{Global state}]
+	\item[Global state] space space space
+	\item[Local state]
+\end{IEEEdescription}
+
+\textbf{Global state}---what came before (load statistics from various network locations)
+
+\textbf{Local state}---per-flow data.
+
+??
+Last action taken: encodes the notion of current belief of flow maliciousness (default at 0).
+Potentially allows forgiveness: a mistakenly malicious flow 's falloff after punishment could be used for such purpose (if we're lucky...)
+
+?? Correspondence ratio: ratio (in bytes) between upstream/downstream traffic for a src/dst pair. (Cite the work that suggested this)
+
+?? Source IP, port, protocol (*zzz*)
+
+?? Flow duration, size
+
+?? Change in send/recv rates since last action/check?
+
+?? More things... (look into other works, actually research this...)
+
+\paragraph{Done with that...}
+
 ?? TODO: everything
@@
 (i.e., are there better ways to handle covariance between properties of system/flow state, properly capture generalisation?)
+
+??
+What might we do for a reward function in the absence of heuristic estimates and/or explicit a priori knowledge?
+I think a good candidate is the sum of up and down throughputs (normalised by capacity sum), so long as \emph{neither exceeds the link capacity}.
+We can extend the team-based formulations similarly.
+This, in theory, promotes traffic diversity since it's not like flooding-based DDoS attacks are going to submit meaningful work to a server.
+The intuition, I suppose, is that certain classes of flow will have a small footprint in one direction which causes a sizeable increase in the other!
 
@@
 
+?? More future work --- share knowledge between agents. ``Knowledge bases'' for this purpose? (see: Qianru).
+
 \section*{Acknowledgements}
-My thanks go to Mircea Iordache and Richard Cziva for their comments and technical assistance.
+My thanks go to Mircea Iordache, Qianru Zhou, and Richard Cziva for their comments and technical assistance.
 Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.

### 2018-08-31 code/marl/compare-expr.sh
Results ahoy (w/ fair instances)

@@ -1,0 +1,7 @@
+#!/bin/bash
+#for count in 2 4 7 11 14;
+for name in "online-expr-2" "all-udp" "baseline-2" "online-expr-2-nginx" "online-expr-2-e-8";
+do
+	python "$name.py"
+done
+

### 2018-08-31 code/marl/granularity-expr.sh

@@ -1,7 +1,8 @@
 #!/bin/bash
-for count in 2 4 7 11 14;
+#for count in 2 4 7 11 14;
+for count in 2 4 8 16;
 do
 	python "online-expr-$count.py"
-	python "online-expr-$count-even.py"
+	#python "online-expr-$count-even.py"
 done
 

### 2018-08-31 code/marl/online-expr-2-e-8.py

@@ -9,7 +9,7 @@ results = marlExperiment(
 	host_range = [2, 2],
 
 	explore_episodes = 0.8,
-	episodes = 10,#50,#500, Since mininet keeps running out of files even e/ cleanup
+	episodes = 15,#50,#500, Since mininet keeps running out of files even e/ cleanup
 	episode_length = 10000,
 	separate_episodes = True,
 

### 2018-08-31 code/marl/online-expr-2.py

@@ -9,7 +9,7 @@ results = marlExperiment(
 	host_range = [2, 2],
 
 	explore_episodes = 0.3,
-	episodes = 10,#50,#500, Since mininet keeps running out of files even e/ cleanup
+	episodes = 15,#50,#500, Since mininet keeps running out of files even e/ cleanup
 	episode_length = 10000,
 	separate_episodes = True,
 

### 2018-08-31 scone-talk/scone-talk.tex

--- 2018-04-06 interim-talk/interim-talk.tex
@@
+% !TeX program = lualatex
+
 \documentclass[aspectratio=169,xcolor={dvipsnames}
@@
 \usepackage{appendixnumberbeamer}
+
+%\usepackage[T1]{fontenc}
 
@@
 ]{biblatex}
-\addbibresource{papers-off.bib}
-\addbibresource{confs-off.bib}
-\addbibresource{books-off.bib}
-\addbibresource{rfc.bib}
-\addbibresource{misc.bib}
+\addbibresource{../paper/papers-off.bib}
+\addbibresource{../paper/confs-off.bib}
+\addbibresource{../paper/books-off.bib}
+\addbibresource{../paper/rfc.bib}
+\addbibresource{../paper/misc.bib}
 
@@
 
-\title{Reinforcement Learning in Network Defence/Control (WIP)}
+\title{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention (WIP)}
 \author{Kyle A. Simpson\\
@@
 \institute{University of Glasgow}
-\date{\nth{10} April, 2018}
+\date{\nth{3} September, 2018}
 
@@
 
-\section{Reinforcement learning}
+\begin{frame}{Introduction}
+	\begin{itemize}
+		\item Network IDS/IPS backed by machine learning haven't taken off as hoped---particularly anomaly-based work.
+		\item Detection problem tricky in this domain:
+		\begin{itemize}
+			\item Evolving: usage shifts, new protocols, new applications.
+			\item Burstiness, seasonal variation.
+			\item Need for correctness, almost no false-positive tolerance.
+			\item Labelling issues.
+		\end{itemize}
+	\end{itemize}
+\end{frame}
 
-\begin{frame}{Introducing: Reinforcement Learning (RL)}
+\begin{frame}{Introduction: Part II}
+\begin{itemize}
+	\item Classes of problem like flooding-based DDoS attacks manifest as a service degradation.
+	\begin{itemize}
+		\item Can these be controlled via feedback loop?
+		\item \alert{``Overcome'' the difficulties of the detection problem} by monitoring and adapting to \emph{performance characteristics and consequences} in real-time
+	\end{itemize}
+	\item Goal: augment signature-based approaches as a last-ditch effort.
+\end{itemize}
+\end{frame}
+
+%\section{Reinforcement learning}
+
+\begin{frame}{RL: The Main Idea\texttrademark}
 	\begin{itemize}
@@
 		\begin{itemize}
-			\item Formally, maximise the \alert{expected discounted reward}\footcite{RL2E}.
+			\item Formally, learn a \alert{policy} to maximise the \alert{expected discounted reward}\footcite{RL2E}.
 		\end{itemize}
@@
 		\begin{itemize}
-			\item I.e., choosing action $a_t$ from a policy in state $s_t$, $a_t \sim \pi(s_t)$, induces the next state $s_{t+1}$ and an associated reward $r_{t+1}$ stochastically.
+			\item I.e., choosing action $a_t$ from a policy in state $s_t$, $a_t \sim \pi(s_t)$, induces the next state $s_{t+1}$ and an associated reward $r_{t+1}$.
 			\item Generalises to \alert{value} $Q(s,a)$---how much reward can we \emph{eventually} expect from choosing each action currently available?
+		\end{itemize}
+
+%		\item The world/environment is believed to be modelled by stochastic state-transition probabilities, reward function distributions...
+%		\begin{itemize}
+%			\item But we don't need to model that!
+%		\end{itemize}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{RL: The Main Benefits\texttrademark}
+\begin{itemize}
+	\item We can learn the optimal policy \alert{without modelling the world ourselves}.
+	\item Formulation allows learning adaptively and online, so long as a reward signal is available.
+	\item Variation in available algorithms, update mechanisms, function approximations, dependence on value functions, action selection, exploration...
+	\begin{itemize}
+		\item Orthogonal concerns, allowing tunable algorithm design.
 	\end{itemize}
@@
 
-\begin{frame}{Why have RL in networks?}
+\begin{frame}{Where has RL succeeded in networks?}
 	%Link to recent work in data-driven networking/optimisation?
@@
 	\begin{itemize}
-		\item \alert{Data-driven networking.} Effectively applied to intra-domain routing \footcite{DBLP:conf/hotnets/ValadarskySST17}, task allocation \footcite{DBLP:conf/hotnets/MaoAMK16} and more, each with general and domain-specific insights.
+		\item \alert{Data-driven networking.} Effectively applied to intra-domain routing \footcite{DBLP:conf/hotnets/ValadarskySST17}, task allocation \footcite{DBLP:conf/hotnets/MaoAMK16}, traffic optimisation \footcite{DBLP:conf/sigcomm/ChenL0L18} and more, each with general and domain-specific insights.
 
@@
 
-\section{What I'm working on}
+\begin{frame}{What might we gain from RL in DDoS prevention?}
+content...
+\end{frame}
+
+%\section{Existing work}
 
@@
 		\item Rigid ISP-like network assumptions.
+
+		\item FIXME WITH THE NEW GRIPES
 	\end{itemize}
@@
 		\begin{figure}
-			\includegraphics[width=\linewidth]{../plots/online.pdf}
-			\caption{Service quality decreases as actions become less granular.}
+%			\includegraphics[width=\linewidth]{../plots/online-pres.pdf}
+			\caption{Service quality decreases as actions become less granular. ??FIXME: MISSING PLOT}
 		\end{figure}
@@
 
-\section{The adversarial setting}
-
-\begin{frame}{Attacks}
-	\begin{description}
-		\item[Poisoning attacks] Affect the training of a continuously learning model (RL agents, collaborative learning).
-
-		\item[Evasion attacks] Make data/state appear to be different. These are what are often known as \alert{adversarial examples}!
-	\end{description}
+\begin{frame}{The Immediate Future}
+content...
 \end{frame}
 
-\begin{frame}{The capabilities of attackers}
+\begin{frame}{The Far Future}
 \begin{itemize}
-		\item Evasion for \emph{polymorphic blending attacks} (i.e., evasion) on past models are known \NP-Complete\footcite{DBLP:conf/ccs/FoglaL06}.
-
-		\item We have empirical bounds on the effectiveness of poisoning attacks wrt.\ what percentage of inbound traffic an attacker controls\footcite{DBLP:journals/jmlr/KloftL10}.
-
-		\item Careless model exposure $\implies$ black-box model extraction\footcite{DBLP:conf/uss/TramerZJRR16} $\implies$ evasion\footcite{DBLP:conf/sp/Carlini017}.
-
-		%\item Promising work on defences\footcite{DBLP:conf/acsac/CaoG17}.
+	\item Other problems.
+	\begin{itemize}
+		\item New action spaces, careful consideration.
 	\end{itemize}
-\end{frame}
-
-\begin{frame}{Unknowns}
-	\begin{itemize}
-		\item We don't have results for poisoning on e.g., Deep RL in the network.
-
-		\item We also lack results on how/if evasion attacks can transfer for learners focused on non-stationarity/constant learning.
-
 \end{itemize}
@@
 		\item A `direct control' approach to intrusion prevention, and \alert{intended improvements}.
-		\item Adversarial work, security, and open questions.
+		\item ?? FIXME
 	\end{itemize}

### 2018-09-02 scone-talk/scone-talk.tex
Talk beta versio.

@@
 
+\usepackage{tikz}
+\usetikzlibrary{arrows.meta, calc, fit, positioning}
+
 %\usepackage{fontspec}
@@
 \usepackage[binary-units]{siunitx}
-\sisetup{range-phrase=--, range-units=single}
+\sisetup{detect-all, range-phrase=--, range-units=single}
 
@@
 \addbibresource{../paper/misc.bib}
+
+\makeatletter
+\DeclareRobustCommand{\rvdots}{%
+	\vbox{
+		\baselineskip4\p@\lineskiplimit\z@
+		\kern-\p@
+		\hbox{.}\hbox{.}\hbox{.}
+}}
+\makeatother
+
+% official colours
+\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}
+
+\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
+\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
+\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
+\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
+\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
+
+\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
+\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
+\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
+\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
+\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
+\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
+\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
+\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}
+
+\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
+\definecolor{uofgforest}{rgb}{0, 0.317647, 0.2}
+\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
+\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
 
@@
 	\end{itemize}
-	\item Goal: augment signature-based approaches as a last-ditch effort.
+	\item Goal: augment signature-based approaches to provide a last line of defence.
 \end{itemize}
@@
 	\begin{itemize}
-		\item Goal: train an agent to make optimal decisions based on observed state.
-		\begin{itemize}
-			\item Formally, learn a \alert{policy} to maximise the \alert{expected discounted reward}\footcite{RL2E}.
-		\end{itemize}
-
 		\item Underlying theory: systems as (discrete-time) \alert{Markov Decision Processes}---states, actions, rewards and transition probabilities.
@@
 			\item Generalises to \alert{value} $Q(s,a)$---how much reward can we \emph{eventually} expect from choosing each action currently available?
+		\end{itemize}
+
+	\item Goal: train an agent to make optimal decisions based on observed state.
+	\begin{itemize}
+		\item Formally, learn a \alert{policy} to maximise the \alert{expected discounted reward}\footcite{RL2E}.
 	\end{itemize}
@@
 
-		\item \alert{In intrusion detectors?} Optimising information sharing in distributed statistical model training \footcite{DBLP:conf/paisi/XuSH07}.
+		\item \alert{In anomaly detection?} Optimising information sharing in distributed statistical model training \footcite{DBLP:conf/paisi/XuSH07}.
 	\end{itemize}
@@
 
-\begin{frame}{What might we gain from RL in DDoS prevention?}
-content...
-\end{frame}
+%\begin{frame}{What might we gain from RL in DDoS prevention?}
+%content...
+%\end{frame}
 
@@
 		\end{itemize}
+
+		\item Implemented in mininet with Ryu controller, traffic generated by replaying traces.
+		\begin{itemize}
+			\item Packet content unimportant---only need accurate load stats/queuing behaviour.
+			\item Alternate model featuring live HTTP traffic.
+		\end{itemize}
 	\end{itemize}
@@
 
-\begin{frame}{Flaws/Risks of MARL?}	
+\begin{frame}{Multiagent RL for DDoS prevention}
+\begin{columns}
+	\begin{column}{0.45\linewidth}
 		\begin{itemize}
-		\item Action mechanism is \emph{pushback}\footcite{DBLP:journals/ccr/MahajanBFIPS02a} (probabilistic packet drop).
+			\item Algorithm: Semi-gradient Sarsa, linear fn approx.
+			\item Actions: Drop [0, 10, ... 90]\si{\percent} upstream traffic.
+			\item State: load vectors of agent and parents ($\mathbb{R}^4$) $\rightarrow$ tile-coded (fixed-weight binary vector).
+			\item Rewards: \num{-1} if $\mathit{load} > U_s$, else fraction of surviving legit traffic.
+		\end{itemize}
+	\end{column}
+	\begin{column}{0.5\linewidth}
+		\begin{figure}
+			\centering
+			\resizebox{\linewidth}{!}{
+				\begin{tikzpicture}[
+				texts/.style = {text=black},
+				labeltexts/.style = {text=uofgsandstone},
+				treeline/.style = {draw=uofgburgundy},
+				treenode/.style = {texts, circle, centered, fill=white, treeline},
+				load/.style = {fill=uofgcobalt},
+				external/.style = {fill=uofgrust},
+				hideline/.style = {draw=uofgsandstone!40!white},
+				hidenode/.style = {treenode, hideline},
+				grow'=right
+				]
+				\node[treenode, label={[texts]above:Server}] (root) {}
+				child [treeline] { node [treenode, load, label={[texts]above:Core}] (sswitch) {}
+					child [treeline] { node [treenode, load, label={[texts]above:Leader}] (teaml) {} 
+						child [treeline] { node [treenode, load, label={[texts]above:Intermediate}] (inter) {}
+							child [treeline] { node [treenode, load, label={[texts]above:Agent/Egress}] (agent) {}
+								child [treeline] { node [treenode, external] (extern) {}
+									child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
+									child [hideline] { node [hidenode] (endhost) {} }
+								}
+							}
+							child [hideline] { node [hidenode] (endagent) {} }
+						}
+						child [hideline] { node [hidenode] (endinter) {} }
+					}
+					child [hideline] { node [hidenode] (endteaml) {} }
+					edge from parent
+					node[below, labeltexts] {$U_s$}
+				};
 
-		\item Strong assumptions on benign/malicious packet estimation quality. 
-		\item Pushback is known to be flawed in practice.
-		\item Rigid ISP-like network assumptions.
-
-		\item FIXME WITH THE NEW GRIPES
-	\end{itemize}
+				%\draw[-] (teaml) -- (endteaml);
+				\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
+				\node [labeltexts, right = -0.1cm of kdots] {$k$};
+				\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
+				\node [labeltexts, right = -0.1cm of ldots] {$l$};
+				\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
+				\node [labeltexts, right = -0.1cm of mdots] {$m$};
+				\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
+				\node [labeltexts, right = -0.1cm of ndots] {$n$};
+				\end{tikzpicture}
+			}
+			\caption{
+				Network topology diagram.
+				Red nodes are external, blue nodes feature in the state vector.
+				Any packet drop occurs when forwarding packets from an egress switch to its parent (intermediate) switch.
+			}
+		\end{figure}
+	\end{column}
+\end{columns}
 \end{frame}
@@
 
-			\item More granular $\implies$ should focus on flow stats, \alert{not aggregates} like now!
+%			\item More granular $\implies$ should focus on flow stats, \alert{not aggregates} like now!
 %Alongside this higher-granularity view, reformulation to include flow characteristics in the decision-making process.
@@
 		\begin{figure}
-%			\includegraphics[width=\linewidth]{../plots/online-pres.pdf}
-			\caption{Service quality decreases as actions become less granular. ??FIXME: MISSING PLOT}
+			\includegraphics[width=\linewidth]{../plots/online-varyN-binary.pdf}
+			\caption{Service quality decreases as actions become less granular (affecting $n$ hosts at once).}
 		\end{figure}
@@
 
-\begin{frame}{The Immediate Future}
-content...
+\begin{frame}{The case for finer granularity}
+\centering
+\includegraphics[width=0.8\linewidth]{../plots/online-varyN-binary.pdf}
+\end{frame}
+
+\begin{frame}{On collateral damage}
+\begin{columns}
+	\begin{column}{0.5\linewidth}
+		\begin{figure}
+			\includegraphics[width=\linewidth]{../plots/online-varyN-nginx.pdf}
+			\caption{Explicit UDP traffic matches replayed traces (tcpreplay vs hping3). TCP traffic (nginx) is severely punished.}
+		\end{figure}
+	\end{column}
+	\begin{column}{0.45\linewidth}
+		\begin{itemize}
+			\item Is the simulation environment of past work complete?
+			\item \alert{No.} It's reliant on a numerical simulator, derived from observations of \emph{traces}.
+			\item UDP benign traffic similar trend to replayed TCP traces, which matches the original results.
+			\item Live TCP responds very badly.
+		\end{itemize}
+	\end{column}
+\end{columns}
+\end{frame}
+
+\begin{frame}{On collateral damage}
+\centering
+\includegraphics[width=0.8\linewidth]{../plots/online-varyN-nginx.pdf}
+\end{frame}
+
+\begin{frame}{And the replication reveals...}	
+\begin{itemize}
+	\item Network topology has no basis in reality---admitted by its \emph{own} source work \footcite{DBLP:journals/ccr/MahajanBFIPS02a}.
+
+	\item Action granularity causes more collateral damage than we'd like...
+
+	\item ...and the picture is worse still for legitimate TCP flows.
+
+	\item Reward function needs a priori knowledge/reliable estimates to learn online.
+
+	\item \alert{But on tbe plus side, action computation is fast: \SIrange{80}{100}{\micro\second}.}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{How can we use these observations? (The Immediate Future)}
+\begin{itemize}
+	\item \textbf{\alert{Why not take actions on a per-flow basis}}?
+	\begin{itemize}
+		\item Solves the granularity issues by construction.
+		\item Allows different treatment by flow features (i.e., protocol).
+	\end{itemize}
+	\item Need to rethink state space: more costly computation, but we have room to work in.
+	\begin{itemize}
+		\item We need any additions to be justified beyond just ``more data'', since \alert{changes affect training time and execution time}.
+	\end{itemize}
+	\item How do we select flows to act upon?
+\end{itemize}
+\end{frame}
+
+\begin{frame}{A candidate state space}
+\begin{description}
+	\item[Global State] The existing state space.
+	\item[Local State] (At least) the following:
+	\begin{itemize}
+		\item \textbf{Src/dst IP, Port, Protocol}---identification.
+		\item \textbf{Flow size, duration, rate}---standard features.
+		\item \textbf{\alert{Last action taken}}---encode belief/forgiveness.
+		\item \textbf{\alert{Correspondence ratio}}---explicitly capture asymmetry.
+		\item \textbf{\alert{$\Delta$rate}}---model how a flow's behaviour changes post-action.
+		\item Other features?
+	\end{itemize}
+\end{description}
+
+And then finding a suitable discretisation...
+\end{frame}
+
+\begin{frame}{The Near Future}
+\begin{itemize}
+	\item Flow selection strategies (guided action calculation).
+	\item Reward functions without dependence on ahead-of-time knowledge.
+	\begin{itemize}
+		\item I.e., for certain distributions of communication we might want to maximise link utilisation in both directions.
+	\end{itemize}
+
+\item Deriving normal model behaviour from traces.
+\begin{itemize}
+	\item We only need to simulate specific behaviour to test these enhancements, but that can become more representative.
+\end{itemize}
+\end{itemize}
 \end{frame}
@@
 	\end{itemize}
+\item Adversarial capabilities---evasion and poisoning attacks.
+\item Knowledge-sharing between agents: cost-modelling and optimisation.
+\item Test deployments in real networks.
 \end{itemize}
@@
 	\begin{itemize}
-		\item A quick introduction to RL, and its \alert{importance to future networks}.
-		\item A `direct control' approach to intrusion prevention, and \alert{intended improvements}.
-		\item ?? FIXME
+		\item A quick introduction to RL, and its \alert{importance to future networks} for optimisation and control of certain classes of problem.
+		\item A recent `direct control' approach to intrusion prevention, and \alert{its significant weaknesses}.
+		\item \alert{Intended improvements} specifically targeting these weaknesses.
 	\end{itemize}

### 2018-09-02 scone-talk/scone-talk.tex
Finish talk with maths.

@@
 %\usepackage[T1]{fontenc}
+
+\usepackage{bm}
 
@@
 		\begin{itemize}
-			\item Algorithm: Semi-gradient Sarsa, linear fn approx.
+			\item Algorithm: Semi-gradient Sarsa, linear fn approx, $\epsilon$-greedy selection.
 			\item Actions: Drop [0, 10, ... 90]\si{\percent} upstream traffic.
@@
 \end{frame}
+\newcommand{\acval}[3]{\ensuremath{\operatorname{\hat{q}}(#1, #2, #3)}}
+\newcommand{\wvec}[1]{\ensuremath{\bm{w}_{#1}}}
+\begin{frame}{The algorithm?}
+% make math easy
+\begin{subequations}
+A little terse, but the main update rule is:
+\begin{gather}
+\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
+\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla_{\wvec{}}{\acval{S_t}{A_t}{\wvec{t}}}
+\end{gather}
+with linear function approximation:
+\begin{gather}
+\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a), \\
+\nabla_{\wvec{}}{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)
+\end{gather}
+\end{subequations}
+\end{frame}
 
@@
 		\begin{figure}
-			\includegraphics[width=\linewidth]{../plots/online-varyN-binary.pdf}
+			\includegraphics[width=\linewidth]{../plots/online-varyN-binary-pres.pdf}
 			\caption{Service quality decreases as actions become less granular (affecting $n$ hosts at once).}
@@
 \centering
-\includegraphics[width=0.8\linewidth]{../plots/online-varyN-binary.pdf}
+\includegraphics[width=0.8\linewidth]{../plots/online-varyN-binary-pres.pdf}
 \end{frame}
@@
 		\begin{figure}
-			\includegraphics[width=\linewidth]{../plots/online-varyN-nginx.pdf}
-			\caption{Explicit UDP traffic matches replayed traces (tcpreplay vs hping3). TCP traffic (nginx) is severely punished.}
+			\includegraphics[width=\linewidth]{../plots/online-varyN-nginx-pres.pdf}
+			\caption{Explicit UDP traffic matches replayed traces (hping3 vs tcpreplay). TCP traffic (nginx) is severely punished.}
 		\end{figure}
@@
 \centering
-\includegraphics[width=0.8\linewidth]{../plots/online-varyN-nginx.pdf}
+\includegraphics[width=0.8\linewidth]{../plots/online-varyN-nginx-pres.pdf}
 \end{frame}
@@
 \begin{itemize}
-	\item \textbf{\alert{Why not take actions on a per-flow basis}}?
+	\item \textbf{\alert{Why not take actions on a per-flow basis?}}
 	\begin{itemize}
@@
 \end{itemize}
+
+\item Investigate other RL algorithms.
+\begin{itemize}
+	\item ``Deep learning'' probably not feasible.
+	\item \alert{$\operatorname{TD}(\lambda)$, actor-critic methods}.
+\end{itemize}
 \end{itemize}

### 2018-09-03 paper/paper.tex
Service theorem proof.

@@
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{bm}
@@
 \newcommand{\wvec}[1]{\ensuremath{\bm{w}_{#1}}}
+
+%thm environments
+\newtheorem{thm}{Theorem}
+\newtheorem{corr}{Corollary}[thm]
 
@@
 
+\section{Threat Model}
+
+?? The fine folks at most security conferences will want to see one. I suppose it serves its purpose by delineating what an attacker can/cannot do.
+
+?? TODO: threat model.
+
 \section{Environment and RL Algorithm}\label{sec:environment-and-rl-algorithm}
@@
 This manifests in two ways: the best-achievable performance decreases, as does the learning rate.
+The first consequence arises analytically.
+Given the probability that a host is legitimate, $P_G \in [0,1]$, it follows that a host will be malicious with probability $P_B = 1 - P_G$.
+Defining \emph{imperfect service} to mean any case where all $n$ hosts connecting over a switch do not share the same classification (i.e., a mixture), then the probability that a switch is delivering imperfect service is $P_{M,n} = 1 - (P_G^2 - P_B^2)$.
+\begin{thm}
+	As $n$ increases, it is more likely that a switch will exhibit imperfect service, i.e., $\forall n \in \mathbb{Z}^{+}, P_{M,n} \le P_{M,n+1}$.
+\end{thm}
+\begin{proof}
+	\emph{Base case:} $P_{M,1}=0, P_{M,2} = 1 - P_G^2 - P_B^2 > 0$.
+
+	\emph{Inductive step:} Assume that the theorem holds for $n$. Observe that $P_G^n \ge P_G^{n+1}$ (resp.\ $P_B$). It then follows that:
+	\begin{align*}
+	P_G^n + P_B^n &\ge P_G^{n+1} + P_B^{n+1}\\
+	1 - (P_G^n + P_B^n) &\le 1 - (P_G^{n+1} + P_B^{n+1})\\
+	P_{M,n} &\le P_{M,n+1}
+	\end{align*}
+\end{proof}
+\begin{corr}
+	Restricting $P_G \in (0,1)$ so that both $P_G$ and $P_B$ are non-zero ensures strict inequality: $P_{M,n} < P_{M,n+1}$.
+\end{corr}
+
 Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.

### 2018-10-04 paper/paper.tex
Extension, fixes, typos to paper

@@
 Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-heralded, but face problems hampering effective deployment.
-The detection problem in this domain is fraught with difficulty; it is an evolving, non-stationary problem as usage patterns shift, new protocols and applications are introduced, compounded by burstiness and seasonal variation.
+The detection problem in this domain is fraught with difficulty; it is an evolving, non-stationary problem as usage patterns shift, new protocols and applications are introduced, and is compounded by burstiness and seasonal variation.
 
@@
 
-I present...
+We present...
 ?? Contribs
@@
 These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
-This issue is inherent to \emph{misuse-} or \emph{signature-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making use of statistical measures.
+This issue is inherent to \emph{misuse-} or \emph{signature-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making effective use of statistical measures.
 
@@
 Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
-Further issues arise from the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges of learning from unlabelled (often partial) data.
+Further issues arise from the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges encountered when learning from unlabelled (often partial) data.
 All of these factors greatly inflate the difficulty of the detection problem.
 
-For certain classes of problem like flooding-based DDoS attacks, \emph{reinforcement learning} (RL) offers another perspective for adaptively reacting to certain classes of threat by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
-In a sense, this allows us to ``overcome'' the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time, by looking for the effect rather than the cause.
+For certain classes of problem like flooding-based DDoS attacks, \emph{reinforcement learning} (RL) offers another perspective.
+The role of an RL agent differs from that of a standard classifier, adaptively reacting to threats by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
+In a sense, this allows us to ``overcome'' the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time; by looking for the effect rather than the cause.
 The intent is to augment what existing misuse-based solutions can provide, by automatically alerting, recording and controlling what are believed to be illegal system states.
@@
 
-?? RL as a mechanism to protect systems by monitoring performance characteristics---overcome the detection problem (different handle on the key problems of burstiness, evolution, non-stationarity).
-
 This paper contributes:
@@
 	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:rethinking-the-state-space}).
-	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction of their formulation, and an acknowledgement of the flaws and risk factors of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a} (\cref{sec:performance-in-an-emulated-environment}).
-	\item ?? TODO: Advance the use/representation/granularity of state-space in direct-control RL of networked applications (\cref{sec:rethinking-the-state-space}).
+	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws and risk factors concerning many traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
+	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
+	\item A flow-level granularity approach to RL-driven DDoS prevention, improving upon past aggregate-based approaches (\cref{sec:rethinking-the-state-space}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
 \end{itemize}
@@
 %?? Introduce RL, related definitions etc.
+\subsection{Reinforcement Learning}
 \emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
 We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
-This formulation of interaction with a system is best described as a \emph{Markov decision process} (MDP).
+This model of system interaction is best described as a \emph{Markov decision process} (MDP).
 RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
-The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in their deployment environment.
+The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
 From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+2} + \ldots$, choosing the discount factor $\gamma \in [0,1]$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
-Formally then, the goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.
+Formally, an agent's goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.
 
@@
 
-If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties of the problem allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
+If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
 Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices).
@@
 
-?? Discuss mininet? Networking terms? SDN stuff?
+%\subsection{Intrusion Detection}
+%Probably want to talk about NIDS/IPS,
+%?? Discuss mininet? Networking terms? SDN stuff?
 
-??
-Probably want to talk about NIDS/IPS, DDoS attack variants (leading into characteristics, supporting features).
+\subsection{Distributed Denial of Service (DDoS)}
+%??
+DDoS attack variants (leading into characteristics, supporting features).
 Amplification (UDP \cite{DBLP:conf/ndss/Rossow14}, TCP \cite{DBLP:conf/uss/KuhrerHRH14}), Transit-link (Crossfire \cite{DBLP:conf/sp/KangLG13}, Coremelt \cite{DBLP:conf/esorics/StuderP09}).
@@
 
-?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
+%?? Explain amplification attack, maybe transit-link?
+
+DDoS attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
+Attackers achieve this either by exploiting peculiarities of operating system or application behaviour (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{flooding-based attacks}).
+Hosts often participate unwillingly, often having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
+
+Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
+\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most vulnerable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
+Malicious hosts send many small requests, spoofed to appear as though the originated from the victim, significantly increasing a botnet's throughput while masking the identity of each participant.
+\emph{Transit-link attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
+
+\subsection{Motivation}
+%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
+%??
+Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows.
+This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e.
+an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
+Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete benefits associated with this problem.
 ??
@@
 
-?? The fine folks at most security conferences will want to see one. I suppose it serves its purpose by delineating what an attacker can/cannot do.
+%?? The fine folks at most security conferences will want to see one. I suppose it serves its purpose by delineating what an attacker can/cannot do.
 
-?? TODO: threat model.
+We assume that attacks are flooding-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a \emph{transit-link attack}).
+The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.
+Finally, we do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
+While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
+The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
 
@@
 A sufficiently well-trained agent needs only to follow the policy it has learned in a greedy manner, allowing pre-training by a simulated environment to transfer to reality.
+?? BUT, if we don't have reward functions which are available at runtime, then we've lost out on what is arguably one of the main benefits that RL can give us!
+?? Need to tie in the concept of ``meaningful work''?
 
@@
 Given the probability that a host is legitimate, $P_G \in [0,1]$, it follows that a host will be malicious with probability $P_B = 1 - P_G$.
-Defining \emph{imperfect service} to mean any case where all $n$ hosts connecting over a switch do not share the same classification (i.e., a mixture), then the probability that a switch is delivering imperfect service is $P_{M,n} = 1 - (P_G^2 - P_B^2)$.
+Defining \emph{imperfect service} to mean any case where all $n$ hosts connecting over a switch do not share the same classification (i.e., a mixture), then the probability that a switch is delivering imperfect service is $P_{M,n} = 1 - (P_G^n + P_B^n)$.
 \begin{thm}
-	As $n$ increases, it is more likely that a switch will exhibit imperfect service, i.e., $\forall n \in \mathbb{Z}^{+}, P_{M,n} \le P_{M,n+1}$.
+	As $n$ increases, it is more likely that a throttling switch will exhibit imperfect service: $\forall n \in \mathbb{Z}^{+}, P_{M,n} \le P_{M,n+1}$.
 \end{thm}
@@
 	\emph{Base case:} $P_{M,1}=0, P_{M,2} = 1 - P_G^2 - P_B^2 > 0$.
-
 	\emph{Inductive step:} Assume that the theorem holds for $n$. Observe that $P_G^n \ge P_G^{n+1}$ (resp.\ $P_B$). It then follows that:
@@
 
+??
+Mention Colin/Qianru's concerns about Quic, IPv6.
+At this point, src/dst IP is the only ``traditional'' characteristics.
+Even then, IPv6 can have a different IP for each application ON A HOST.
+Port number meaningless...
+
 ?? What do we want?
@@
 
-\section{The Results of Doing So}
+\section{The Results of Doing So}\label{sec:the-results-of-doing-so}
 

### 2018-10-11 code/marl-bwmon/marl-bwmon.cc
Fixes to crucial logic, socket talk

@@ -14,12 +14,15 @@
 
 #include <arpa/inet.h>
 #include <fcntl.h>
+#define __STDC_FORMAT_MACROS
+#include <inttypes.h>
 #include <netinet/in.h>
 #include <pcap/pcap.h>
 #include <poll.h>
 #include <sys/select.h>
 #include <sys/socket.h>
 #include <sys/time.h>
+#include <sys/un.h>
 #include <unistd.h>
 
 namespace ch = std::chrono;
@@ -32,6 +35,9 @@ ch::high_resolution_clock::time_point stdifyTimeval(const struct timeval tv) {
 }
 
 const int LISTEN_PORT = 9932;
+const bool TIME_PRINT_ERR = false;
+const bool USE_UNIX_SOCK = true;
+const char *SOCKET_PATH = "bwmon-sock";
 
 // Store and update some stats about float64s.
 struct Stat {
@@ -122,6 +128,9 @@ struct FlowStats
 	ch::high_resolution_clock::time_point flow_start;
 	ch::high_resolution_clock::time_point last_entry;
 
+	// unseen?
+	bool unseen = true;
+
 	FlowStats() {
 		//interarrivals.report_update = true;
 	};
@@ -182,10 +191,11 @@ struct FlowStats
 		flow_size_in_prev = flow_size_in;
 		flow_size_out_prev = flow_size_out;
 
+		unseen = false;
 		return true;
 	}
 
-	std::optional<FlowMeasurement> clearAndRetrieveStats(uint32_t ip, char *ip_str, ch::high_resolution_clock::time_point startTime, ch::high_resolution_clock::time_point endTime) {
+	std::optional<std::pair<bool, FlowMeasurement>> clearAndRetrieveStats(uint32_t ip, char *ip_str, ch::high_resolution_clock::time_point startTime, ch::high_resolution_clock::time_point endTime) {
 		auto prune_age = ch::seconds(10);
 
 		// prune here if last packet rx'd was of a certain age,
@@ -221,7 +231,10 @@ struct FlowStats
 		flow_size_in_prev = flow_size_in;
 		flow_size_out_prev = flow_size_out;
 
-		return std::optional<FlowMeasurement> {fm};
+		auto new_data = unseen;
+		unseen = false;
+
+		return std::optional<std::pair<bool, FlowMeasurement>> {std::make_pair(new_data, fm)};
 	}
 
 	void clear() {
[…]

### 2018-10-11 code/marl/marl.py

@@ -12,6 +12,7 @@ import twink.ofp5.parse as ofpp
 
 from contextlib import closing
 import cPickle as pickle
+import errno
 import itertools
 import math
 import networkx as nx
@@ -19,6 +20,7 @@ import numpy as np
 import os
 import random
 from sarsa import SarsaLearner
+import select
 import signal
 import socket
 import struct
@@ -100,7 +102,9 @@ def marlExperiment(
 		state_direction = "in",
 		actions_target_flows = False,
 
-		bw_mon_socketed = True,
+		bw_mon_socketed = False,
+		unix_sock = True,
+		print_times = False,
 	):
 
 	linkopts_core = linkopts
@@ -327,7 +331,7 @@ def marlExperiment(
 		return int(prob * 0xffffffff)
 
 	def netip(ip, subnet):
-		(lhs,) = struct.unpack("I", socket.inet_aton(ip))
+		(lhs,) = struct.unpack("I", socket.inet_aton(ip)) if isinstance(ip, str) else (ip,)
 		(rhs,) = struct.unpack("I", socket.inet_aton(subnet))
 		return struct.pack("I", lhs & rhs)
 
@@ -527,11 +531,15 @@ def marlExperiment(
 
 	def enactActions(learners, sarsas, flow_action_sets):
 		if actions_target_flows:
+			intime = time.time()
 			for (node, sarsa, maps) in zip(learners, sarsas, flow_action_sets):
 				for ip, state in maps.iteritems():
 					(_, action, _) = state
 					a = action if override_action is None else override_action
 					updateUpstreamRoute(node, ac_prob=a, target_ip=ip)
+			outtime = time.time()
+			if print_times:
+				print "do_acs:", outtime-intime
 		else:
 			for (node, sarsa) in zip(learners, sarsas):
 				(_, action, _) = sarsa.last_act
@@ -912,89 +920,122 @@ def marlExperiment(
 		#  pkt_out_sz_mean, pkt_out_sz_var, pkt_out_count, [8-10]
 		#  wnd_iat_mean, wnd_iat_var [11-12]
 		# )
+		bw_sock = None
 		if bw_mon_socketed:
 			# need to mix this with connection management
 			# i.e. open a socket out here, have this function make use of it...
 			# remember to close this at the end...!
-			bw_sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
[…]

### 2018-10-11 paper/paper.tex

@@
 Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete benefits associated with this problem.
-??
-Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows.
-This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e.
-an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
+We have seen that for other domains in particular, misclassification is a serious problem, which can introduce \emph{collateral damage} in the context of DDoS prevention.
+In theory, the feedback-loop-like model allows us to monitor flows \emph{after} an action is taken to allow forgiveness of mistakenly punished flows.
+This does rely upon the ability to take a flow-by-flow view of the state space, but if we can combine knowledge of current state with the last applied action, then perhaps a flow which falls off identically to a legitimate flow can be rescued.
 
+Which features might be best suited to this problem?
+%??
+Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
 ??
 Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
+%?
+%? Need to find a good survey which suggests indicative features.
 ?? Need to find a good survey which suggests indicative features.
+%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
 ?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
@@
 %?? State the problem and what we want to achieve.
-To best discover how the use of RL techniques in the field of network intrusion prevention may bet be improved, it's prudent that we start by looking for weaknesses in work which exists today.
+To best discover how the use of RL techniques in the field of network intrusion prevention may yet be improved, it's prudent that we start by looking for weaknesses in work which exists today.
 The closest approach which is available within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant: their work helped to show the viability of live, adaptive, feedback loop-like control of the network to detect and prevent DDoS attacks.
@@
 While this is taken verbatim from the work we use as a basis, it is presented within the confines of a framework given by \textcite{DBLP:journals/ton/YauLLY05}.
-This framework is, in turn, an extension of the class of topologies examined in the original pushback paper \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
-In actuality, this topology is far closer to the latter; crucially, it seems to not actually represent or correspond to any realistic deployment at all.
+This framework is, in turn, an extension of the class of topologies examined in the original pushback paper \cite{DBLP:journals/ccr/MahajanBFIPS02a}; this topology is far closer to the latter.
+Crucially, it seems to not actually represent or correspond to any realistic deployment at all.
 In the words of \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}:
@@
 Remarkably, it seems to have propagated in a way which was not originally intended.
-It is clear that such a network topology allows for functional testing, but it is hard to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
+It is clear that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network, but it is hard to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
 We aim to address this, to some extent.
 
-Reimplementation of the network model was performed using mininet \cite{mininet}, a Python-based network emulator.
+Reimplementation of this network model was performed using mininet \cite{mininet}, a Python-based network emulator.
 Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
@@
 
-The reward functions as specified here were introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of the global reward as a performance metric and the \emph{coordinated team learning} (CTL) reward for training.
+The reward functions as specified here were introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of their global reward as a performance metric and their \emph{coordinated team learning} (CTL) reward for training.
 These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
@@
 \section{Performance in an Emulated Environment}\label{sec:performance-in-an-emulated-environment}
-The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and assumes a rigid and unrepresentative topology.
+The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and assumes a rigid and unrepresentative topology.
 The latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
@@
 \subsection{Service Guarantees and Host Density}
-In the default topology, $n=2$ hosts per egress point is not a parameter that can be described as realistic for many web service deployments.
+The default topology chooses $n=2$ hosts per egress point, which isn't a parameter setting that can be described as realistic for many web service deployments.
 In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
@@
 	\caption{
-		MARL pushback control system performance plotted over time, for various settings of $n$ hosts per agent.
-		This plot shows that, from the perspective of benign hosts, service guarantees degrade as inference and applied actions become less granular.
+		MARL pushback control system performance plotted over time for various settings of $n$ hosts per agent.
+		This plot shows that, from the perspective of benign hosts, service guarantees degrade as inference and actions become less granular.
 		This generalises for all behavioural discriminators: even a perfect agent \emph{must} punish benign flows if they are grouped with malicious actors.
@@
 \Cref{fig:marl-granularity} shows that as more hosts are allocated to each learner, the fraction of traffic believed to be good (as observed at the server) decreases.
-This manifests in two ways: the best-achievable performance decreases, as does the learning rate.
+This manifests in two ways: the best-achievable performance drops, and so too does the learning rate.
 The first consequence arises analytically.
@@
 
-Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
-From the findings of \cref{sec:tcp-multiplexing-and-fallback-behaviour} and the Mathis equation, their results should be considerably worse in practice.
 %?? Expand according to the summary you posted in the notes document.
@@
 
-?? TODO: Gut this section now...
 %While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
@@
 	\caption{
-		A more convincing and/or to-the-point figure for getting people to believe I'm doing god's work.
-		?? TODO: describe 4 real.
+		Comparison of preserved traffic across different traffic models, choosing $n=2$ hosts per egress node.
+		No actions are taken in the ``baseline'' entries.
+		The agents examined here perform above the baseline when traffic sources pay no mind to congestion control, but perform significantly worse for live TCP traffic (nginx).
 		\label{fig:nginx-coffin-nail}
@@
 %\Cref{fig:mplex-tcp} shows that this trend is present.
-\Cref{fig:nginx-coffin-nail} shows that this trend is present.
-Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
-Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
-%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
-UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.
-
-?? Quic will probably behave much like TCP.
-
-The performance characteristics of TCP under packet drop are well-known and well-understood through the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}:
+Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
+From the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}, through which the performance characteristics of TCP under packet drop are well-known and well-understood, their results should be considerably worse in practice:
 \begin{equation}
@@
 for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions, and the probability $p \in (0, 1]$ that a packet is lost.
-%The lower bound on $p$ is deliberately open, as packet drop acts as a control signal for the TCP send window size and is an expected part of normal operation.
-%TCP flows in Mininet are found to obey this law, as expected.
-This reveals that the original work \emph{does not capture this behaviour}, or else the initial experiment (UDP-like traffic when $n=2$) wouldn't have achieved similar results to those which were reported.
-
-\Cref{sec:a-new-normal} is where the new normal model arrives. Wow!
+Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
+Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
+\Cref{fig:nginx-coffin-nail} shows that this trend is present.
+%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
+The TCP traffic as it appears here is generated according to \Cref{sec:a-new-normal}.
+UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.
+As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP.
 
 \subsection{Computational Cost}
-??
+%??
 Consider talking about the execution times of the old MARL approach here?
@@
 
-We're in \SIrange{80}{100}{\micro\second}
+Measurements during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
+This is reassuring when measured alongside the insights from other work.
+\textcite{DBLP:conf/sigcomm/ChenL0L18} observed that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
+%Most flows are short, and flow-size follows a heavy-tailed distribution.
+That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
 
@@
 
-In establishing...
-
-?? How will I structure this?
-?? Motivation -> Model -> Results?
-?? OR Use the results of the last section to springboard into here?
+%In establishing...
+%
+%?? How will I structure this?
+%?? Motivation -> Model -> Results?
+%?? OR Use the results of the last section to springboard into here?
 
@@
 If such information is lost, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
-The controller is also responsible for responding to ARP traffic, as a gateway might
+The controller is also responsible for responding to ARP traffic, as a gateway might do.
 The necessity of this arises due to the realities of relying upon Linux's IP stack, and would not need to be considered for purely trace-based evaluation.
@@
 
-\Cref{fig:nginx-coffin-nail} shows the effects of using
-
-?? Actually describe what I have here, now. I can do this.
+?? Actually describe what I have here, now.
 
@@
 
+?? Two modes: long-flow (i.e., groundhog day loop of GCC8 download) and nice request-chain mode for ``representativeness''(??)
+
 \section{Rethinking the State Space}\label{sec:rethinking-the-state-space}
@@
 Mention Colin/Qianru's concerns about Quic, IPv6.
-At this point, src/dst IP is the only ``traditional'' characteristics.
+At this point, src/dst IP are the only ``traditional'' characteristics that can be relied upon.
 Even then, IPv6 can have a different IP for each application ON A HOST.
@@
 
-\begin{IEEEdescription}[\IEEEsetlabelwidth{Global state}]
-	\item[Global state] space space space
-	\item[Local state]
-\end{IEEEdescription}
-
 \textbf{Global state}---what came before (load statistics from various network locations)
 
-\textbf{Local state}---per-flow data.
+\textbf{Local state}---per-flow data. Specifically, right now I'm using:
+
+?? \emph{IP}
 
 ??
-Last action taken: encodes the notion of current belief of flow maliciousness (default at 0).
+\emph{Last action taken}: encodes the notion of current belief of flow maliciousness (default at 0).
 Potentially allows forgiveness: a mistakenly malicious flow 's falloff after punishment could be used for such purpose (if we're lucky...)
 
-?? Correspondence ratio: ratio (in bytes) between upstream/downstream traffic for a src/dst pair. (Cite the work that suggested this)
+?? \emph{Flow duration, size}
 
-?? Source IP, port, protocol (*zzz*)
+?? \emph{Correspondence ratio}: ratio (in bytes) between upstream/downstream traffic for a src/dst pair. (Cite the work that suggested this)
 
-?? Flow duration, size
+?? \emph{Change in send/recv rates} since last action/check.
 
-?? Change in send/recv rates since last action/check?
+?? \emph{Mean inter-arrival time}.
 
-?? More things... (look into other works, actually research this...)
+\subsection{What's done with these?}
 
-\paragraph{Done with that...}
+?? Tile coded together, action taken per-flow every timestep.
 
@@
 
+\renewcommand*{\bibfont}{\small}
 \printbibliography

### 2018-10-12 paper/paper.tex
Normal model description.

@@
 
-\subsection{Topology}
+\subsection{Topology}\label{sec:topology}
 
@@
 
-??
+From what we have seen, it is difficult (or impossible) for trace-based or numerical simulations to correctly capture certain dynamics without an extraordinary amount of care or consideration.
+%As it turns out, 
+Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
+In particular, we are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
+We make no claims that it is perfect or representative for all traffic, only that it captures some of the behaviour which we expect will plague most legitimate TCP flows; if need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.
+
+%??
+ANGLE: set up an environment to test \emph{specific} behaviours to examine \emph{specific} problems in past work.
+I make no claims that it is perfect or representative for all traffic, just for this (likely common) behaviour which I expect to plague almost all legit TCP flows.
+
+%??
+Existing sims used for testing such applications reliant on traces, or not sophisticated enough to capture interactive, back-and-forth (correlated) behaviours---possibly discarded as second-hand effects by past work when these are so crucial given user traffic patterns (and the nature of the control signal we choose to enact).
+
+%??
 Remember, the motivation is clear.
@@
 The whole purpose of this is that we aim to test specific behaviour which traces cannot replicate (i.e., correlated back-and-forth, dynamics introduced to congestion-aware protocols, ...)
-?? If we need to, we can codition the distribution of requests according to statistics mined form an existing trace if reviewer number 2 needs that extra push to be convinced.
+%?? If we need to, we can condition the distribution of requests according to statistics mined from an existing trace if reviewer number 2 needs that extra push to be convinced.
 
-??
-Full Openflow controller setup (in mininet).
-All internal routers are primed with knowledge of the shortest path to any internal host, while new inbound flows register the ``way back'' for each hop used to maintain consistency over traffic conditions.
-If such information is lost, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
-The controller is also responsible for responding to ARP traffic, as a gateway might do.
-The necessity of this arises due to the realities of relying upon Linux's IP stack, and would not need to be considered for purely trace-based evaluation.
-?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
-?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
+\subsection{Network Design}
+We make use of a fully software-defined network, built using Openflow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
+All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used to maintain consistency over traffic conditions.
+If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
+The controller is also responsible for computing how switches respond to ARP requests: the necessity of this arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
+We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.
 
-?? Actually describe what I have here, now.
+\subsection{Traffic Model}
+%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
+To model legitimate traffic the server node runs an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
+Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
+This provides two modes of operation:
+\begin{itemize}
+	\item \emph{long flow} testing has each host constantly request a large binary, such as a tarball of GCC8;
+	\item \emph{request-chain} testing has each host request a random file known to exist within the website, followed by any dependent resources (stylesheets, images, etc.).
+This presents a more balanced distribution of flow duration and size.
+\end{itemize}
+Each host's download rate is limited to match whatever maximum bandwidth was assigned to it.
+Accordingly, reward functions aim to maximise the downstream throughput.
 
-??
-ANGLE: set up an environment to test \emph{specific} behaviours to examine \emph{specific} problems in past work.
-I make no claims that it is perfect or representative for all traffic, just for this (likely common) behaviour which I expect to plague almost all legit TCP flows.
-
-??
-Existing sims used for testing such applications reliant on traces, or not sophisticated enough to capture interactive, back-and-forth (correlated) behaviours---possibly discarded as second-hand effects by past work when these are so crucial given user traffic patterns (and the nature of the control signal we choose to enact).
-
-?? Two modes: long-flow (i.e., groundhog day loop of GCC8 download) and nice request-chain mode for ``representativeness''(??)
+%?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
+Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
+Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that simulation remains tractable.
 
@@
 ??
+Explicitly mention \emph{why} we don't make use of neural networks, and instead favour more classical function approximation.
+Latency concerns, energy usage, graphics/TPU/whatever hardware availability or lack thereof across switches/routers/etc.
+
+??
 NOTE: think about how I might need to change/alter the tile-coding when we add extra discriminative features...
@@
 Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-Given that arguably the main strength of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
+Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
 
@@
 Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-?? Also adaptively choosing video bitrates based on network conditions \cite{DBLP:conf/sigcomm/MaoNA17}.
+Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
 
@@
 \section*{Acknowledgements}
-My thanks go to Mircea Iordache, Qianru Zhou, and Richard Cziva for their comments and technical assistance.
-Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
+Our thanks go to Mircea Iordache, Qianru Zhou, and Richard Cziva for their comments and technical assistance.
+%Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
 

### 2018-10-19 paper/paper.tex
Paperial Evolution (?)

@@
+% !TeX document-id = {e7fef4b9-afbc-4e81-94aa-c4847183d047}
+% !BIB program = biber
 \documentclass[conference, letterpaper, 10pt, times]{IEEEtran}
 
-\usepackage[T1]{fontenc}
+%\usepackage[T1]{fontenc}
 \usepackage{amsfonts}
@@
 %bib
-\usepackage[style=ieee,maxnames=3,mincitenames=1,maxcitenames=2,maxbibnames=99,mincrossrefs=5,sortcites,backend=bibtex,uniquelist=false]{biblatex}
+\usepackage[style=ieee,maxnames=3,mincitenames=1,maxcitenames=2,maxbibnames=99,
+mincrossrefs=99,minxrefs=99,
+sortcites,
+%backend=bibtex,
+uniquelist=false]{biblatex}
 \addbibresource{papers-off.bib}
@@
 \newtheorem{corr}{Corollary}[thm]
+
+\newcommand{\fakepara}[1]{\noindent\textbf{#1:}}
 
@@
 
-?? Reference to the possible mistreatment of RL in our domain? The problem of ``blind application'' so hated by \textcite{DBLP:conf/sp/SommerP10}.
-
-This paper contributes:
+To date, there have been few applications of this class of algorithms towards intrusion prevention which make use of their full potential for online control, rather than using them as the basis of a classifier.
+We aim to take steps to redress this and establish their proper capabilities, beyond simple ``blind application''.
+In service of this goal, this paper contributes:
 \begin{itemize}
 	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:rethinking-the-state-space}).
-	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws and risk factors concerning many traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
+	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws, and risk factors concerning many traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
 	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
@@
 
-\section{Background}
+\section{Background and Threat Model}
 %?? Introduce RL, related definitions etc.
@@
 To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+ \dim{\mathcal{A}})$-dimensional grids with an optional bias component.
+The number of tilings has a direct effect on the resolution of the discretisation.
 Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state-action pair exists within the corresponding tile.
@@
 Attackers achieve this either by exploiting peculiarities of operating system or application behaviour (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{flooding-based attacks}).
-Hosts often participate unwillingly, often having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
+Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
 
@@
 Malicious hosts send many small requests, spoofed to appear as though the originated from the victim, significantly increasing a botnet's throughput while masking the identity of each participant.
-\emph{Transit-link attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
+\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
 
@@
 an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
-Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete benefits associated with this problem.
+Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete reasons for their use here.
 We have seen that for other domains in particular, misclassification is a serious problem, which can introduce \emph{collateral damage} in the context of DDoS prevention.
@@
 
-\section{Threat Model}
+\subsection{Threat Model}
 
@@
 
-We assume that attacks are flooding-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a \emph{transit-link attack}).
+?? Oultine attacker's goals, particularly wrt SPIFFY.
+??
+They want to impact fair-share bandwidth allocation to flows (minimise), they have no worry about being unmasked (there are many available reflectors, which don't need to be bought etc...) [cite the rossow papers about amplification hell].
+We assume that attacks are bandwidth-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a \emph{transit-link attack}).
 The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.
+
 Finally, we do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
@@
 
-\section{Environment and RL Algorithm}\label{sec:environment-and-rl-algorithm}
-
-?? Probably rethink the title of this section.
-
+\section{Analysing Existing Work}\label{sec:environment-and-rl-algorithm}
 %?? State the problem and what we want to achieve.
 To best discover how the use of RL techniques in the field of network intrusion prevention may yet be improved, it's prudent that we start by looking for weaknesses in work which exists today.
-The closest approach which is available within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant: their work helped to show the viability of live, adaptive, feedback loop-like control of the network to detect and prevent DDoS attacks.
-We begin with a reimplementation and more principled re-examination of their work to further develop the field.
+The closest approach which is available within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant: their work helped to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
+We begin with a reimplementation and more principled re-examination of their work to further develop the state-of-the-art.
 
@@
 
-The reward functions as specified here were introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of their global reward as a performance metric and their \emph{coordinated team learning} (CTL) reward for training.
+We employ the reward functions introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of their global reward as a performance metric and their \emph{coordinated team learning} (CTL) reward for training.
 These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
@@
 
-?? TODO: Make repo public...
+%?? TODO: Make repo public...
 
@@
 In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
-We feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions as defined become less granular.
+We feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions (as defined above) become less granular.
 
@@
 UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.
+
+These results are of particular importance due to TCP's prevalence within the Internet: it carries \SIrange{79}{94}{\percent} of packets, corresponding to \SIrange{89}{98}{\percent} of data transferred by volume \cite{DBLP:conf/saint/ZhangDJC09}.
 As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP.
@@
 
-Measurements during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
+Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
+Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
 This is reassuring when measured alongside the insights from other work.
-\textcite{DBLP:conf/sigcomm/ChenL0L18} observed that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
+\Textcite{DBLP:conf/sigcomm/ChenL0L18} observe that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
 %Most flows are short, and flow-size follows a heavy-tailed distribution.
@@
 We make use of a fully software-defined network, built using Openflow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
-All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used to maintain consistency over traffic conditions.
+All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
 If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
@@
 \begin{itemize}
-	\item \emph{long flow} testing has each host constantly request a large binary, such as a tarball of GCC8;
+	\item \emph{long flow} testing has each host constantly request a large binary file, such as a tarball of GCC8;
 	\item \emph{request-chain} testing has each host request a random file known to exist within the website, followed by any dependent resources (stylesheets, images, etc.).
@@
 ?? What do we want?
+
+?? note---probably put in the experiments that Simon suggested at this point as justification for why we chose what we did...
 
@@
 
-?? What concessions will we have to make in order to make per-flow processing more viable? Intelligent sampling/reanalysis of flows when needed (i.e. an external heuristic guiding method)?
+??
+What concessions will we have to make in order to make per-flow processing more viable?
+Intelligent sampling/reanalysis of flows when needed (i.e.
+an external heuristic guiding method)?
+In SPIFFY's \cite{DBLP:conf/ndss/KangGS16} evaluation, we see clearly that it can take around \SI{2}{\second} for a flow to react fully to a rate increase---I think for the TCP step it may be wise to factor this in, too!
 
@@
 ??
+What other consequences are there?
+We're training each agent on all the flows it's acting on which means slight theoretical changes (we have multiple experience traces, and each uses its last state + new state + new target value (reward) as an update to the \emph{current model}).
+
+??
 What might we do for a reward function in the absence of heuristic estimates and/or explicit a priori knowledge?
@@
 The intuition, I suppose, is that certain classes of flow will have a small footprint in one direction which causes a sizeable increase in the other!
+Alternatively, monitor the health of canary flows which cross the team boundary (i.e.
+only one in-out link).
 
@@
 
-?? Other recent work in the prevention of DDoS? (i.e., non-RL).
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} had it as part of their evaluation, BGP-like rerouting techniques (specifically for transit-link attacks, but also for `normal' attacks) \cite{RoutingAroundCongestion}
+%?? Try and compare my work here when possible?
+
+\fakepara{DDoS Prevention}
+%?? Other recent work in the prevention of DDoS? (i.e., non-RL).
+\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
+Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
+%?? What does it do? detects state, not which flows are contributing...
+
+\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
+\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
+%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
+Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
+
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k nearest-neighbours} classifier to detect individual attack flows.
+Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
+
+\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
+This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
+The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
+Because of this, traditional flooding-based attacks aren't fully alleviated.
 
 %?? Abuses of RL 
+\fakepara{RL in Networks}
 Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
@@
 A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
-?? Good Direct-control work? What this work is based on (which directly captures the spirit of RL) \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}
+We feel there is a shortage of work in this category at present, at least in the field of networks; whether this is due to disinterest or difficulty is as-yet unknown.
+To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.
 
@@
 
-?? Completely reword, yikes.
-\citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s \emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} uses deep RL (the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}).
-Crucially, they find that decisions need to be made for the vast majority of (short-lived) flows, sub-millisecond.
-Their work falls between the above two paradigms: to overcome the high latency of action computation via neural network, two agents are trained.
-The first learns to optimise flow size thresholds governing priority (and accordingly demarcate long and short flows).
-All short flows are routed by ECMP.
-The second agent makes decisions about routing, prioritisation etc.\ for these long flows, who are likely to live long enough that any decisions made will have a significant impact.
+%?? Completely reword, yikes.
+\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
+Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
+To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
+The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
+The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.
 
-These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
+%These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
 
@@
 
+?? Can probably fold a lot of this out into discussion or something.
+
 \section*{Acknowledgements}
-Our thanks go to Mircea Iordache, Qianru Zhou, and Richard Cziva for their comments and technical assistance.
+Our thanks go to Simon Rogers, Mircea Iordache, Qianru Zhou, and Richard Cziva for their advice, comments and technical assistance.
 %Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.

### 2018-10-22 paper/paper.tex
Expansion + Anonymisation

@@
 	pdftitle={\mytitle{}},
-	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros}
+%	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros}
+	pdfauthor={Anonymous Giraffe and Badger}
 }
@@
 \author{
-	Kyle A.\ Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
-	Dimitrios P.\ Pezaros\\
-	University of Glasgow, Glasgow, Scotland,\\
-	\email{k.simpson.1@research.gla.ac.uk}
+%	Kyle A.\ Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
+%	Dimitrios P.\ Pezaros\\
+%	University of Glasgow, Glasgow, Scotland,\\
+%	\email{k.simpson.1@research.gla.ac.uk}
+Anonymous Giraffe\thanks{This work was supported by a nameless, shapeless funder [grant number 7].},
+Anonymous Badger\\
+Unnamed Department, Nowhere\\
+\email{giraffe.a@unnamed.zoo}
 }
@@
 
-Which features might be best suited to this problem?
+%Which features might be best suited to this problem?
 %??
 Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
-??
-Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
-%?
-%? Need to find a good survey which suggests indicative features.
-?? Need to find a good survey which suggests indicative features.
+Other studies suggest that there are particularly useful features which make the task of online DDoS flow identification feasible.
+Aggregate network load observed at various locations suggests the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest asymmetry and in many cases illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
+Generic volume-based statistics (counts, counts per duration, average packet sizes) have seen effectiveness in such as $k$-nearest neighbours classifiers \cite{DBLP:conf/dsn/LeeKSPY17}.
+Most importantly, there is evidence showing behavioural changes in response to bandwidth expansion \cite{DBLP:conf/ndss/KangGS16}, suggesting similar artefacts might arise after throttling, packet drop, or other interference.
 %?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
-?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
+%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
 
@@
 
-?? Oultine attacker's goals, particularly wrt SPIFFY.
-??
+%?? Oultine attacker's goals, particularly wrt SPIFFY.
+%??
 They want to impact fair-share bandwidth allocation to flows (minimise), they have no worry about being unmasked (there are many available reflectors, which don't need to be bought etc...) [cite the rossow papers about amplification hell].
+An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, we they are expected to act rationally in its pursuit.
 We assume that attacks are bandwidth-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a \emph{transit-link attack}).
 The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.
+The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
 
@@
 All experiments described were executed using an Ubuntu 16.04.4 LTS machine with a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
-All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
+All code underpinning these findings is available on a public repository\footnote{\url{https://aweb.site/a-repo.}}.
+%All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
 

### 2018-11-01 paper/paper.tex

@@
 \sisetup{detect-all, range-phrase=--, range-units=single}
+
+\makeatletter
+\let\MYcaption\@makecaption
+\makeatother
+
+\usepackage[font=footnotesize]{subcaption}
+
+\makeatletter
+\let\@makecaption\MYcaption
+\makeatother
 
@@
 Unnamed Department, Nowhere\\
-\email{giraffe.a@unnamed.zoo}
+\email{giraffe.a@unnamed.com}
 }
@@
 	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
-	\item A flow-level granularity approach to RL-driven DDoS prevention, improving upon past aggregate-based approaches (\cref{sec:rethinking-the-state-space}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
+	\item A source-level granularity approach to RL-driven DDoS prevention, improving upon past aggregate-based approaches (\cref{sec:rethinking-the-state-space}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
 \end{itemize}
@@
 They want to impact fair-share bandwidth allocation to flows (minimise), they have no worry about being unmasked (there are many available reflectors, which don't need to be bought etc...) [cite the rossow papers about amplification hell].
-An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, we they are expected to act rationally in its pursuit.
-We assume that attacks are bandwidth-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a \emph{transit-link attack}).
+An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
+Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
+We assume that attacks will be bandwidth-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a \emph{transit-link attack}).
 The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.
-The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
+The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
 
-Finally, we do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
+We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
 While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
 The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
+These are APT-level capabilities, whose exploration presents a rich source of future work.
 
@@
 treenode/.style = {texts, circle, centered, fill=white, treeline},
+load/.style = {fill=uofgcobalt},
+loadhide/.style = {fill=uofgcobalt!40!white},
 external/.style = {fill=uofgrust},
+externalhide/.style = {fill=uofgrust!40!white},
 hideline/.style = {draw=uofgsandstone!40!white},
@@
 \node[treenode, label={[texts]above:Server}] (root) {}
-	child [treeline] { node [treenode, label={[texts]above:Core}] (sswitch) {}
-		child [treeline] { node [treenode, label={[texts]above:Leader}] (teaml) {} 
-			child [treeline] { node [treenode, label={[texts]above:Intermediate}] (inter) {}
-				child [treeline] { node [treenode, label={[texts]above:Agent/Egress}] (agent) {}
+child [treeline] { node [treenode, load, label={[texts]above:Core}] (sswitch) {}
+	child [treeline] { node [treenode, load, label={[texts]above:Leader}] (teaml) {} 
+		child [treeline] { node [treenode, load, label={[texts]above:Intermediate}] (inter) {}
+			child [treeline] { node [treenode, load, label={[texts]above:Agent/Egress}] (agent) {}
 				child [treeline] { node [treenode, external] (extern) {}
 					child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
-						child [hideline] { node [hidenode] (endhost) {} }
+					child [hideline] { node [hidenode, externalhide] (endhost) {} }
 				}
 			}
-				child [hideline] { node [hidenode] (endagent) {} }
+			child [hideline] { node [hidenode, loadhide] (endagent) {} }
 		}
-			child [hideline] { node [hidenode] (endinter) {} }
+		child [hideline] { node [hidenode, loadhide] (endinter) {} }
 	}
-		child [hideline] { node [hidenode] (endteaml) {} }
+	child [hideline] { node [hidenode, loadhide] (endteaml) {} }
 	edge from parent
@@
 	Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $l$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
-	Empty nodes are considered to be internal.
+%	Empty nodes are considered to be internal.
+	Red nodes are considered to be external, and the load of each blue node features in the state vector.
 	\label{fig:marl-topol}
@@
 
-?? wrt reward function relying on heuristic elements or perfect knowledge:
-A sufficiently well-trained agent needs only to follow the policy it has learned in a greedy manner, allowing pre-training by a simulated environment to transfer to reality.
-?? BUT, if we don't have reward functions which are available at runtime, then we've lost out on what is arguably one of the main benefits that RL can give us!
-?? Need to tie in the concept of ``meaningful work''?
+%?? wrt reward function relying on heuristic elements or perfect knowledge:
+While it can be observed that reliance upon heuristic elements or perfect knowledge as in the above reward functions hampers their real-world applicability, a sufficiently well-trained agent needs only to follow the policy it has learned in a greedy manner, allowing pre-training by a simulated environment to transfer to reality.
+%?? BUT, if we don't have reward functions which are available at runtime, then we've lost out on what is arguably one of the main benefits that RL can give us!
+%?? Need to tie in the concept of ``meaningful work''?
+The lack of a truly online reward function is sorely felt, however.
+This causes the loss of some of RL's main benefits, such as online learning with regards to an evolving problem context.
 
@@
 
-All experiments described were executed using an Ubuntu 16.04.4 LTS machine with a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
-All code underpinning these findings is available on a public repository\footnote{\url{https://aweb.site/a-repo.}}.
+All experiments described were executed on Ubuntu 16.04.5 LTS (GNU/Linux 4.4.3-040403-generic x86\_64), using a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
+All code underpinning these findings is available on a public repository\footnote{Link removed for double-blind reviewing.}.
 %All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
@@
 	\item \emph{long flow} testing has each host constantly request a large binary file, such as a tarball of GCC8;
-	\item \emph{request-chain} testing has each host request a random file known to exist within the website, followed by any dependent resources (stylesheets, images, etc.).
+	\item \emph{request-chain} testing has each host request several random files known to exist within the website, followed by any dependent resources for each (stylesheets, images, etc.).
+On completion, a host changes its IP to minimise downtime of inbound data while separating statistics per-flow.
 This presents a more balanced distribution of flow duration and size.
@@
 Each host's download rate is limited to match whatever maximum bandwidth was assigned to it.
-Accordingly, reward functions aim to maximise the downstream throughput.
 
@@
 Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that simulation remains tractable.
+When using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows.
 
-\section{Rethinking the State Space}\label{sec:rethinking-the-state-space}
+\section{Acting On Individual Sources}
+
+?? High-level overview...
+
+\subsection{Models}
+
+?? Talk about the modifications to standard: multiple actions per-timestep, updating many traces from one reward...
+
+?? Timed Random Sequential flow visitation.
+
+?? Accordingly, reward functions aim to maximise the downstream throughput. Reward function direction for overload + protected loads.
+
+??
+Talk about SPF -- Influenced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
+Why?
+To demonstrate the design of a model which includes some aspects of real-world knowledge, and to see whether making an agent less ``free-form'' might allow us to trade power for learning speed (smaller action space etc.).
+
+\subsection{Rethinking the State Space}\label{sec:rethinking-the-state-space}
+
+\begin{figure*}
+	\centering
+	\begin{subfigure}{0.32\linewidth}
+		\includegraphics[width=\linewidth]{../plots/ftprep-good}
+	\end{subfigure}
+	\begin{subfigure}{0.32\linewidth}
+		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
+	\end{subfigure}
+	\begin{subfigure}{0.32\linewidth}
+		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
+	\end{subfigure}
+	\caption{
+		Features: how good are they really? These are for UDP. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
+	\label{fig:udp-feature-plots}
+}
+\end{figure*}
+
+%\begin{figure*}
+%	\centering
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-cap-good}
+%	\end{subfigure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-cap-good-2}
+%	\end{subfigure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-cap-good-2}
+%	\end{subfigure}
+%	\caption{
+%		Features: how good are they really? These are for UDP. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
+%		\label{fig:udp-cap-feature-plots}
+%	}
+%\end{figure*}
+
+\begin{figure*}
+	\begin{subfigure}{0.32\linewidth}
+		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good}
+	\end{subfigure}
+	\begin{subfigure}{0.32\linewidth}
+		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good-2}
+	\end{subfigure}
+	\begin{subfigure}{0.32\linewidth}
+		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good-3}
+	\end{subfigure}
+
+	\caption{
+		Features: how good are they really (part 2)? These are for TCP. These are less conclusive...
+		\label{fig:tcp-feature-plots}
+	}
+\end{figure*}
+
+\begin{figure*}
+	\begin{subfigure}{0.32\linewidth}
+		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good}
+	\end{subfigure}
+	\begin{subfigure}{0.32\linewidth}
+		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-2}
+	\end{subfigure}
+	\begin{subfigure}{0.32\linewidth}
+		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-2}
+	\end{subfigure}
+
+	\caption{
+		Features: how good are they really (part 3)? These are for TCP, but capped. These are less conclusive...
+		\label{fig:tcp-cap-feature-plots}
+	}
+\end{figure*}
 
@@
 
+??
+NEED TO SAY SOMEWHERE: Baselines use TC for establishing interleaving/buffers filling etc., when MARL+ or SPF are in play then disabling TC (thus allowing illegal states, with \num{-1} reward marking them as such---``negative reinforcement'') may accelerate learning.
+We have a choice---but the results show that for TCP we WANT TC enabled...
+
 ?? What do we want?
@@
 
-\textbf{Local state}---per-flow data. Specifically, right now I'm using:
+\textbf{Local state}---per-source data. Specifically, right now I'm using:
 
@@
 \emph{Last action taken}: encodes the notion of current belief of flow maliciousness (default at 0).
-Potentially allows forgiveness: a mistakenly malicious flow 's falloff after punishment could be used for such purpose (if we're lucky...)
+Potentially allows forgiveness: a mistakenly malicious flow's falloff after punishment could be used for such purpose (if we're lucky...)
 
@@
 
-?? \emph{Correspondence ratio}: ratio (in bytes) between upstream/downstream traffic for a src/dst pair. (Cite the work that suggested this)
+??
+\emph{Correspondence ratio}: ratio (in bytes) between upstream/downstream traffic for a src/dst pair.
+(Cite the work that suggested this) $C_X = \min(\mathit{in}, \mathit{out})/\max(\mathit{in}, \mathit{out})$
 
@@
 ?? \emph{Mean inter-arrival time}.
+
+?? \Cref{fig:udp-feature-plots} shows the effectiveness of each for UDP (resp.\ \cref{fig:udp-feature-plots} for TCP).
+
+\begin{table}
+	\centering
+	\caption{Tile coding windows for each feature.}
+
+	\begin{tabular}{@{}ll@{}}
+		\toprule
+		New Feature (unit) & Range \\
+		\midrule
+		Load (\si{\mega\bit\per\second}) & $[0, U_s]$ \\
+		IP & $[0, 2^{32}-1]$ \\
+		Last Action (\si{\percent}) & $[0, 1]$ \\
+		Duration (\si{\milli\second}) & $[0, \num{2000}]$ \\
+		Size (\si{\mebi\byte}) & $[0,10]$ \\
+		Correspondence Ratio & $[0,1]$ \\
+		Mean IAT (\si{\milli\second}) & $[0, \num{10000}]$ \\
+		$\Delta$In/Out Rate (\si{\mega\bit\per\second}) & $[-50, 50]$ \\
+		Packets In/Out & $[0, 7000]$ \\
+		Packets In/Out Window & $[0, 2000]$ \\
+		Mean In/Out Packet Size (\si{\byte}) & $[0, 1560]$ \\
+		\bottomrule
+	\end{tabular}
+\end{table}
 
@@
 
+??
+Have some more things worth testing now.
+FSA with transitions based on the theory underlying SPIFFY.
+5 states (start in low pdrop) from \SIrange{0}{100}{\percent} drop, actions move up, down or hold on the level of drop.
+Why like this?
+Chosen deployment environments might not be able to reroute to get the requisite bandwidth epansion factor, so punish everyone to begin with.
+
 \section{The Results of Doing So}\label{sec:the-results-of-doing-so}
@@
 ?? They'd probably look real nice...
+
+??
+"We chose to take a separate tile-coding over each feature, ignoring combinatorial effects for now.
+Combining them all as one tile-coding led to poor performance due to an inability to generalise value across similar states."
 
@@
 
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k nearest-neighbours} classifier to detect individual attack flows.
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
 Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
@@
 %These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
+
+\section{Discussion}
+
+Talk about flaws here, what could go wrong...
+
+\subsection{Future Work}
+
+Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.
 
@@
 \section*{Acknowledgements}
-Our thanks go to Simon Rogers, Mircea Iordache, Qianru Zhou, and Richard Cziva for their advice, comments and technical assistance.
+%Our thanks go to Simon Rogers, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Richard Cziva for their advice, comments and technical assistance.
+Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
 %Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.

### 2018-11-01 code/marl/bonus-feature-test-cap-prep.py
FT TCP Bonus results

@@ -1,0 +1,96 @@
+import cPickle
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 2
+
+def run(restrict, state, rewards, good_traffic_percents, total_loads):
+	results = marlExperiment(
+		n_teams = 2,#5,
+
+		n_inters = 2,
+		n_learners = 3,
+		host_range = [hosts_p, hosts_p],
+
+		explore_episodes = 0.8,
+		episodes = 1,
+		episode_length = 10000,
+		separate_episodes = True,
+
+		alpha = 0.05,
+		epsilon = 0.2,
+		discount = 0,
+
+		dt = 0.05,
+
+	#	old_style=True,
+
+		rf = "ctl",
+		use_controller = True,
+		actions_target_flows = restrict is not None,
+
+		restrict = restrict,
+
+		rand_state = state,
+		rewards = rewards,
+		good_traffic_percents = good_traffic_percents,
+		total_loads = total_loads,
+
+		feature_max = 18,
+		estimate_const_limit = True,
+	)
+
+	return results
+
+# Run one set to old mode, run 8 on restrict [4+i for i in xrange(8)]
+# Build up results each time, should have 9 sets. collect RNG states for each set
+
+results_dir = "../../results/"
+file_dir = results_dir + "bonus-ftprep-cap.pickle"
+n_features = 8
+n_new_features = 6
+n_episodes = 10
+restrict_sets = [[4 + n_features + i] for i in xrange(n_new_features)]
+out_names = ["f{}".format(n_features + i) for i in xrange(n_new_features)]
+
+print restrict_sets, out_names
+
+result_sets = [([], [], []) for i in xrange(n_new_features)]
+
+if __name__ == "__main__":
+	things_to_pickle = []
+
+	randstate = None
+	for i in xrange(n_episodes):
+		state = (i, randstate)
+		agents = []
+		for j, restriction in enumerate(restrict_sets):
+			(rewards, good_traffic_percents, total_loads) = result_sets[j]
+			(rs, gs, ls, store_sarsas, rng_state, _) = run(
+				restriction,
+				state[1],
+				rewards, good_traffic_percents, total_loads
+			)
+
+			result_sets[j] = (rs, gs, ls)
+			randstate = rng_state
+
+			# list of: each set of agents, and the filtered
+			agents.append((store_sarsas, restriction))
+
[…]

### 2018-11-01 code/marl/bonus-feature-test-tcp-cap-prep.py

@@ -1,0 +1,103 @@
+import cPickle
+from marl import *
+from writer import writeResults, makeResultsAverage
+
+hosts_p = 2
+
+def run(restrict, state, rewards, good_traffic_percents, total_loads):
+	results = marlExperiment(
+		n_teams = 2,#5,
+
+		n_inters = 2,
+		n_learners = 3,
+		host_range = [hosts_p, hosts_p],
+
+
+		explore_episodes = 0.8,
+		episodes = 1,
+		episode_length = 10000,
+		separate_episodes = True,
+
+		alpha = 0.05,
+		epsilon = 0.2,
+		discount = 0,
+
+
+		dt = 0.05,
+
+	#	old_style=True,
+
+		rf = "ctl",
+		use_controller = True,
+		actions_target_flows = restrict is not None,
+
+		restrict = restrict,
+
+		rand_state = state,
+		rewards = rewards,
+		good_traffic_percents = good_traffic_percents,
+		total_loads = total_loads,
+
+		evil_range = [4,7],
+		model = "nginx",
+		reward_direction = "out",
+		randomise = True,
+		randomise_count = 3,
+		randomise_new_ip = True,
+
+		feature_max = 18,
+		estimate_const_limit = True,
+	)
+
+	return results
+
+# Run one set to old mode, run 8 on restrict [4+i for i in xrange(8)]
+# Build up results each time, should have 9 sets. collect RNG states for each set
+
+results_dir = "../../results/"
+file_dir = results_dir + "bonus-ftprep-tcp-cap.pickle"
+n_features = 8
+n_new_features = 6
+n_episodes = 10
+restrict_sets = [[4 + n_features + i] for i in xrange(n_new_features)]
+out_names = ["f{}".format(n_features + i) for i in xrange(n_new_features)]
+
+result_sets = [([], [], []) for i in xrange(n_new_features)]
+
+if __name__ == "__main__":
+	things_to_pickle = []
+
+	randstate = None
+	for i in xrange(n_episodes):
+		state = (i, randstate)
+		agents = []
+		for j, restriction in enumerate(restrict_sets):
+			(rewards, good_traffic_percents, total_loads) = result_sets[j]
+			(rs, gs, ls, store_sarsas, rng_state, _) = run(
+				restriction,
+				state[1],
+				rewards, good_traffic_percents, total_loads
+			)
+
+			result_sets[j] = (rs, gs, ls)
+			randstate = rng_state
+
[…]

### 2018-11-04 paper/paper.tex
Improvements + SPF explained.

@@
 %	\email{k.simpson.1@research.gla.ac.uk}
-Anonymous Giraffe\thanks{This work was supported by a nameless, shapeless funder [grant number 7].},
+Anonymous Giraffe\thanks{This work was supported by a nameless funder [grant number 7].},
 Anonymous Badger\\
@@
 \cond{c}{(v)} = [\load{l_v} > U_s/k],\\
-R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\\
-R_{t}^{\mathit{CTL}}(v) = (1 - \cond{f}\cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{f} \cond{c}{(v)},
+R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\label{eqn:reward-rt}\\
+R_{t}^{\mathit{CTL}}(v) = (1 - \cond{f}\cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{f} \cond{c}{(v)},\label{eqn:reward-rctl}
 \end{gather}
-where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.
+where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.\label{eqn:reward}
 \end{subequations}
@@
 		No actions are taken in the ``baseline'' entries.
-		The agents examined here perform above the baseline when traffic sources pay no mind to congestion control, but perform significantly worse for live TCP traffic (nginx).
+		The agents examined here perform above the baseline in all cases, but perform significantly worse for live TCP traffic (nginx).
 		\label{fig:nginx-coffin-nail}
@@
 Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
+Each host's download rate is limited to match whatever maximum bandwidth was assigned to it.
 This provides two modes of operation:
@@
 \end{itemize}
-Each host's download rate is limited to match whatever maximum bandwidth was assigned to it.
 
@@
 \section{Acting On Individual Sources}
+Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per source.
+However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
+We describe and justify our enhancements to the RL model of \textcite{DBLP:journals/eaai/MalialisK15}, present a new model which draws on domain knowledge leveraged by SPIFFY \cite{DBLP:conf/ndss/KangGS16}, and offer an empirical investigation into the effectiveness of flow-level statistics on live decision-making.
 
-?? High-level overview...
+\subsection{Marl++}\label{sec:marl-plus}
+%\fakepara{Multiple actions}
+%?? Talk about the modifications to standard: multiple actions per-timestep, updating many traces from one reward...
+Modifying the frequency with which actions and value functions are updated has important consequences for not only overall system performance, but also in \emph{how} an agent learns.
+We choose to adapt these algorithms to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to facilitate effective learning.
+To this end, we allow agents to make many decisions per timestep.
+We maintain the last state-action pair associated with each source IP, and calculate any actions for the flows which still exist.
+Finally, we update $\wvec{}$ according to \cref{eqn:sg-sarsa} with each available trace and the sole reward measurement an agent has at that time.
+As exploration still occurs for each action, this approach reduces $\epsilon$ multiple steps for each $t$.
+In turn, we increase the annealing window for $\epsilon$ to \num{8000} action choices to preserve exploration over time, while accounting for the greater volume of decisions being made.
 
-\subsection{Models}
+%\fakepara{Timed Random Sequential actions}
+Taking actions in this manner means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
+This introduces numerous potential issues: inability to response to unexpected changes in flow state, delayed service of new flows and risks that flow states become outdated; at their worst, these present additional attack surface to an adversary.
+To adapt to these problems, we make use of \emph{timed random sequential} updates.
+Each agent begins with an empty work list.
+For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
+Uncompleted work is passed on to the next timestep, until it is emptied, at which point is is repopulated using the set of available measurements.
+To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
+State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
+Following \textcite{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 8 flows in this time.
+We expect this should be tuned based on the frequency at which statistics arrive.
+%?? Prune old? We can get through 8 per iteration, might use this as a basis for knowing when to say ``enough is enough''.
 
-?? Talk about the modifications to standard: multiple actions per-timestep, updating many traces from one reward...
+{
+\newcommand{\load}[2]{\operatorname{load}^{#2}_{t}(#1)}
+\newcommand{\uload}[1]{\load{#1}{\uparrow}}
+\newcommand{\dload}[1]{\load{#1}{\downarrow}}
+\newcommand{\bload}[1]{\load{#1}{\updownarrow}}
+\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
+%\fakepara{Reward function directionality}
+The reward functions, as defined, do not take traffic direction into account.
+We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
+Denoting the upstream, downstream and combined loads $\uload{s}, \dload{s}, \bload{s}$, we modify \cref{eqn:reward}:
+\begin{subequations}
+\begin{gather}
+\cond{f} = [\max(\uload{s}, \dload{s}) > U_s],\\
+\cond{c}{(v)} = [\max(\uload{l_v}, \dload{l_v}) > U_s/k],
+\end{gather}
+\end{subequations}
+replacing $\load{t}{}$ in \crefrange{eqn:reward-rt}{eqn:reward-rctl} with whichever directional load is prioritised.
+We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, though we expect that $\bload{\cdot}$ would be the most suitable for general deployment or heterogeneous traffic patterns.
+}
 
-?? Timed Random Sequential flow visitation.
+\subsection{SPF: SPIFFY Lite}
 
-?? Accordingly, reward functions aim to maximise the downstream throughput. Reward function direction for overload + protected loads.
+An important question to ask while exploring the design space of RL agents is whether (and how) domain knowledge can improve the performance or learning rate of a defence system.
+This often comes as a trade-off: expressive power and unfettered control, weighed against the smaller action spaces afforded by existing insight or structure.
+We present a second model, relying on the measurements of \textcite{DBLP:conf/ndss/KangGS16} which suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
+While it is not stated whether this behaviour applies to amplification attacks, we expect this to remain the case; the upload rate of any amplifier will be fixed at its upper bound given that upstream links of bots are saturated.
+Moreover, bots receive no feedback on the effective throughput of their attack flows.
 
-??
+How might we apply their observations within the RL paradigm, while tailoring them towards non-ISP deployment?
+Our approach is to model how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, reducing the action set to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ in single steps.
+To enable temporary bandwidth expansion in resource constrained deployments, every flow is initially placed under light packet drop ($p=0.05$, or state 1), corresponding to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
+This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and non-congestion-aware traffic as required.
+We combine this with the enhancements described in \cref{sec:marl-plus}.
+%??
 Talk about SPF -- Influenced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
@@
 To demonstrate the design of a model which includes some aspects of real-world knowledge, and to see whether making an agent less ``free-form'' might allow us to trade power for learning speed (smaller action space etc.).
+Also, because we might be able to use its insights in an environ where we don't have spare bandwidth?
+Such a model means we need an agents need to ``plan ahead'', i.e.
+have $\gamma \ne 0$.
+We found $\gamma = 0.8$ to be the most effective choice.
+We also use most of the techniques from above.
+
+This model is significantly different from the perspective of RL parameter tuning: each agent now requires the capability to plan ahead, since they may no longer choose $p$ directly from any state.
+This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
+We found $\gamma = 0.8$ to be the most effective choice during exploratory testing.
 
@@
 %	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-cap-good-2}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-cap-good-3}
 %	\end{subfigure}
 %	\caption{
-%		Features: how good are they really? These are for UDP. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
+%		Features: how good are they really? These are for UDP, but capped. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
 %		\label{fig:udp-cap-feature-plots}
@@
 	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-2}
+		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-3}
 	\end{subfigure}
@@
 We have a choice---but the results show that for TCP we WANT TC enabled...
+For UDP, we actually get a slight degradation in performance.
+This suggests a kind of ``banding'', perhaps?
 
@@
 
+?? What if there are too many flows and a decision set is never completed? Uh... trap door and skip to next set after 50 flow decisions or something?
+
 \subsection{Future Work}

### 2018-11-05 paper/paper.tex
Cleanup + Less Incendiary

@@
 The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
-From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+2} + \ldots$, choosing the discount factor $\gamma \in [0,1]$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
+From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+2} + \ldots$, choosing the discount factor $\gamma \in [0,1)$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
 Formally, an agent's goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.
@@
 Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
-\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most vulnerable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
-Malicious hosts send many small requests, spoofed to appear as though the originated from the victim, significantly increasing a botnet's throughput while masking the identity of each participant.
+\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most exploitable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
+Malicious hosts send many small requests, spoofed to appear as though they originated from the victim, causing many large replies to be sent to the intended target---significantly increasing a botnet's throughput while masking the identity of each participant.
 \emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
@@
 Aggregate network load observed at various locations suggests the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest asymmetry and in many cases illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
-Generic volume-based statistics (counts, counts per duration, average packet sizes) have seen effectiveness in such as $k$-nearest neighbours classifiers \cite{DBLP:conf/dsn/LeeKSPY17}.
+Generic volume-based statistics (counts, counts per duration, average packet sizes) have seen effectiveness in such as $k$-nearest neighbours classifiers trained to detect DDoS attacks in progress \cite{DBLP:conf/dsn/LeeKSPY17}.
 Most importantly, there is evidence showing behavioural changes in response to bandwidth expansion \cite{DBLP:conf/ndss/KangGS16}, suggesting similar artefacts might arise after throttling, packet drop, or other interference.
@@
 Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
-We assume that attacks will be bandwidth-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a \emph{transit-link attack}).
+We assume that attacks will be bandwidth-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
 The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.
@@
 %?? State the problem and what we want to achieve.
-To best discover how the use of RL techniques in the field of network intrusion prevention may yet be improved, it's prudent that we start by looking for weaknesses in work which exists today.
-The closest approach which is available within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant: their work helped to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
+To best discover how the use of RL techniques in the field of network intrusion prevention may be improved, it's prudent that we start by looking for weaknesses in work which exists today.
+The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant: their work helps to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
 We begin with a reimplementation and more principled re-examination of their work to further develop the state-of-the-art.
@@
 %?? So, why does this proliferate, exist? They effectively disavowed it in their own paper. Reading comprehension is hard, I guess.
-Remarkably, it seems to have propagated in a way which was not originally intended.
+%Remarkably, it seems to have propagated in a way which was not originally intended.
 It is clear that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network, but it is hard to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
@@
 Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
-While unrealistic with respect to packet content, it was assumed that this would provide representative load and packet inter-arrival characteristics, since the agents as initially defined are blind to such information.
+While unrealistic with respect to packet content, it was assumed that this would provide representative load and packet inter-arrival characteristics, since the agents as defined rely only upon load measurements.
 
@@
 The lack of a truly online reward function is sorely felt, however.
-This causes the loss of some of RL's main benefits, such as online learning with regards to an evolving problem context.
+This causes the loss of some of RL's main benefits, such as online learning with regard to an evolving problem context.
 
@@
 %\Cref{fig:mplex-tcp} shows that this trend is present.
-Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
+Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is based upon does not capture the key interactions between TCP flows and packet drop as a control action.
+The TCP traffic as it appears here is generated according to \Cref{sec:a-new-normal}.
 From the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}, through which the performance characteristics of TCP under packet drop are well-known and well-understood, their results should be considerably worse in practice:
@@
 for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions, and the probability $p \in (0, 1]$ that a packet is lost.
-Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
-Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
-\Cref{fig:nginx-coffin-nail} shows that this trend is present.
+Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant; good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
+\Cref{fig:nginx-coffin-nail} shows that this model performs poorly when attempting to safeguard TCP traffic, though remains better than leaving the network unprotected.
 %?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
-The TCP traffic as it appears here is generated according to \Cref{sec:a-new-normal}.
 UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.
 
-These results are of particular importance due to TCP's prevalence within the Internet: it carries \SIrange{79}{94}{\percent} of packets, corresponding to \SIrange{89}{98}{\percent} of data transferred by volume \cite{DBLP:conf/saint/ZhangDJC09}.
-As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP.
+These results are of particular importance due to TCP's prevalence within the Internet.
+The protocol is known to carry \SIrange{79}{94}{\percent} of packets, corresponding to \SIrange{89}{98}{\percent} of data transferred by volume \cite{DBLP:conf/saint/ZhangDJC09}.
+As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development.
 
@@
 	\item \emph{request-chain} testing has each host request several random files known to exist within the website, followed by any dependent resources for each (stylesheets, images, etc.).
-On completion, a host changes its IP to minimise downtime of inbound data while separating statistics per-flow.
+On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
 This presents a more balanced distribution of flow duration and size.
@@
 Finally, we update $\wvec{}$ according to \cref{eqn:sg-sarsa} with each available trace and the sole reward measurement an agent has at that time.
-As exploration still occurs for each action, this approach reduces $\epsilon$ multiple steps for each $t$.
-In turn, we increase the annealing window for $\epsilon$ to \num{8000} action choices to preserve exploration over time, while accounting for the greater volume of decisions being made.
+As exploration still occurs for each action, this approach reduces $\epsilon$ multiple notches every timestep.
+In turn, we increase the annealing window for $\epsilon$ to \num{8000} action choices so as to preserve exploration over time, by accounting for the greater volume of decisions being made.
 
@@
 Taking actions in this manner means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
-This introduces numerous potential issues: inability to response to unexpected changes in flow state, delayed service of new flows and risks that flow states become outdated; at their worst, these present additional attack surface to an adversary.
+This introduces numerous potential issues: inability to response to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated; at their worst, these present additional attack surface to an adversary.
 To adapt to these problems, we make use of \emph{timed random sequential} updates.
@@
 We present a second model, relying on the measurements of \textcite{DBLP:conf/ndss/KangGS16} which suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
-While it is not stated whether this behaviour applies to amplification attacks, we expect this to remain the case; the upload rate of any amplifier will be fixed at its upper bound given that upstream links of bots are saturated.
+While it is not stated whether this behaviour applies to amplification attacks, we expect this to remain the case; the upload rate of any amplifier will be fixed at its upper bound given that the upstream links of any connected bots are saturated.
 Moreover, bots receive no feedback on the effective throughput of their attack flows.
@@
 How might we apply their observations within the RL paradigm, while tailoring them towards non-ISP deployment?
-Our approach is to model how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, reducing the action set to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ in single steps.
+Our approach is to constrain how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, reducing the action set to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ in single steps.
 To enable temporary bandwidth expansion in resource constrained deployments, every flow is initially placed under light packet drop ($p=0.05$, or state 1), corresponding to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
@@
 
-This model is significantly different from the perspective of RL parameter tuning: each agent now requires the capability to plan ahead, since they may no longer choose $p$ directly from any state.
+This model is significantly different when considering how to tune the RL algorithm's parameters: each agent now requires the capability to plan ahead, since they may no longer choose $p$ directly from any state.
 This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.

### 2018-11-06 paper/paper.tex
Section V drafted

@@
 
-\subsection{Motivation}
+\subsection{Motivation}\label{sec:motivation}
 %?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
@@
 \cond{f} = [\load{s} > U_s],\\
-\cond{c}{(v)} = [\load{l_v} > U_s/k],\\
+\cond{c}{(v)} = \cond{f} \cdot{} \, [\load{l_v} > U_s/k],\\
 R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\label{eqn:reward-rt}\\
-R_{t}^{\mathit{CTL}}(v) = (1 - \cond{f}\cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{f} \cond{c}{(v)},\label{eqn:reward-rctl}
+R_{t}^{\mathit{CTL}}(v) = (1 - \cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{c}{(v)},\label{eqn:reward-rctl}
 \end{gather}
@@
 
-\subsection{TCP Multiplexing and Fallback Behaviour}\label{sec:tcp-multiplexing-and-fallback-behaviour}
+\subsection{TCP Multiplexing and Falloff Behaviour}\label{sec:tcp-multiplexing-and-fallback-behaviour}
 
@@
 The protocol is known to carry \SIrange{79}{94}{\percent} of packets, corresponding to \SIrange{89}{98}{\percent} of data transferred by volume \cite{DBLP:conf/saint/ZhangDJC09}.
-As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development.
+As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development to properly handle traffic with such characteristics.
 
@@
 \cond{f} = [\max(\uload{s}, \dload{s}) > U_s],\\
-\cond{c}{(v)} = [\max(\uload{l_v}, \dload{l_v}) > U_s/k],
+\cond{c}{(v)} = \cond{f}{} \cdot{} \, [\max(\uload{l_v}, \dload{l_v}) > U_s/k],
 \end{gather}
+\label{eqn:reward-but-better}
 \end{subequations}
@@
 
+%Explicitly mention \emph{why} we don't make use of neural networks, and instead favour more classical function approximation.
+Latency concerns, energy usage, graphics/TPU/whatever hardware availability or lack thereof across switches/routers/etc.
+To make decisions cheaply and at low latency, we maintain the usage of semi-gradient Sarsa with tile coding, rather than making use of neural networks or more complicated function approximators.
+Although this likely brings lower theoretical performance, there are more reasons than latency alone to favour classical methods; these include lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
+
 \subsection{SPF: SPIFFY Lite}
@@
 This often comes as a trade-off: expressive power and unfettered control, weighed against the smaller action spaces afforded by existing insight or structure.
-We present a second model, relying on the measurements of \textcite{DBLP:conf/ndss/KangGS16} which suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
+We propose a second model, relying on the measurements of \textcite{DBLP:conf/ndss/KangGS16} which suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
 While it is not stated whether this behaviour applies to amplification attacks, we expect this to remain the case; the upload rate of any amplifier will be fixed at its upper bound given that the upstream links of any connected bots are saturated.
@@
 We combine this with the enhancements described in \cref{sec:marl-plus}.
-%??
-Talk about SPF -- Influenced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
-Why?
-To demonstrate the design of a model which includes some aspects of real-world knowledge, and to see whether making an agent less ``free-form'' might allow us to trade power for learning speed (smaller action space etc.).
-Also, because we might be able to use its insights in an environ where we don't have spare bandwidth?
-Such a model means we need an agents need to ``plan ahead'', i.e.
-have $\gamma \ne 0$.
-We found $\gamma = 0.8$ to be the most effective choice.
-We also use most of the techniques from above.
 
@@
 This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
-We found $\gamma = 0.8$ to be the most effective choice during exploratory testing.
+We found $\gamma = 0.8$ to be the most effective choice during exploratory testing with UDP traffic.
 
@@
 	\caption{
-		Features: how good are they really? These are for UDP. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
+%		Features: how good are they really? These are for UDP. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
+		Learned performance of Marl++ agents when benign traffic is UDP-like, using only a single feature as a basis for decisions.
+		Mean IAT, inbound packet sizes, and global state offer the best predictive performance, while most features offer marginal advantage over the unprotected baseline.
+		Here, the link's bandwidth limit is removed, and so agents automatically learn to keep bandwidth below $U_s$ by negative reinforcement---this offers a slight per-feature performance increase compared to when this explicit cap is present (not pictured).
 		\label{fig:udp-feature-plots}
@@
 	\caption{
-		Features: how good are they really (part 2)? These are for TCP. These are less conclusive...
+%		Features: how good are they really (part 2)? These are for TCP. These are less conclusive...
+		Learned performance of Marl++ agents when benign traffic is TCP-like, using only a single feature as a basis for decisions (uncapped as in \cref{fig:udp-feature-plots}).
+		Global state and Mean IAT still offer the greatest improvement above baseline, but it is observed that different optimal features are expressed between different traffic models.
+		Packet-level statistics are considerably less effective for this class of traffic, with $\Delta$ Out Rate and Correspondence Ratio being proportionally better than they are alongside UDP.
 		\label{fig:tcp-feature-plots}
@@
 	\caption{
-		Features: how good are they really (part 3)? These are for TCP, but capped. These are less conclusive...
+%		Features: how good are they really (part 3)? These are for TCP, but capped. These are less conclusive...
+		Learned performance of Marl++ agents when benign traffic is TCP-like, using only a single feature as a basis for decisions (with bandwidth capped physically).
+		Although negative rewards (and thus negative reinforcement) cannot occur, all features achieve considerably better performance, a reversal of our observation for UDP traffic.
+%		Global state and Mean IAT still offer the greatest improvement above baseline, but packet-level statistics are considerably less effective for this class of traffic.
 		\label{fig:tcp-cap-feature-plots}
@@
 
-??
-Mention Colin/Qianru's concerns about Quic, IPv6.
-At this point, src/dst IP are the only ``traditional'' characteristics that can be relied upon.
-Even then, IPv6 can have a different IP for each application ON A HOST.
-Port number meaningless...
-
-??
-NEED TO SAY SOMEWHERE: Baselines use TC for establishing interleaving/buffers filling etc., when MARL+ or SPF are in play then disabling TC (thus allowing illegal states, with \num{-1} reward marking them as such---``negative reinforcement'') may accelerate learning.
-We have a choice---but the results show that for TCP we WANT TC enabled...
-For UDP, we actually get a slight degradation in performance.
-This suggests a kind of ``banding'', perhaps?
-
-?? What do we want?
-
-?? note---probably put in the experiments that Simon suggested at this point as justification for why we chose what we did...
-
-\textbf{Global state}---what came before (load statistics from various network locations)
-
-\textbf{Local state}---per-source data. Specifically, right now I'm using:
-
-?? \emph{IP}
-
-??
-\emph{Last action taken}: encodes the notion of current belief of flow maliciousness (default at 0).
-Potentially allows forgiveness: a mistakenly malicious flow's falloff after punishment could be used for such purpose (if we're lucky...)
-
-?? \emph{Flow duration, size}
-
-??
-\emph{Correspondence ratio}: ratio (in bytes) between upstream/downstream traffic for a src/dst pair.
-(Cite the work that suggested this) $C_X = \min(\mathit{in}, \mathit{out})/\max(\mathit{in}, \mathit{out})$
-
-?? \emph{Change in send/recv rates} since last action/check.
-
-?? \emph{Mean inter-arrival time}.
-
-?? \Cref{fig:udp-feature-plots} shows the effectiveness of each for UDP (resp.\ \cref{fig:udp-feature-plots} for TCP).
+The main element required to move to a per-source model is a feature set which offers high predictive power, such that behavioural differences are readily apparent to an agent.
+Elaborating further on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and offer an investigation into their use alongside different traffic types:
+%?? We use these features, and why...
+
+\fakepara{Global state ($\mathbf{4\cdot{}}$Load)}
+these are the existing set of load measurements introduced in \cref{sec:environment-and-rl-algorithm}.
+These indicate the overall health of the network, and crucially are all measurements which an agent directly controls.
+
+\fakepara{Source IP address}
+while ordinarily trivial to spoof, in real-world scenarios the addresses of a set of reflector nodes might exhibit similarity, most likely through geography.
+
+\fakepara{Last action taken}
+this encodes an agent's current belief in the maliciousness of a flow.
+This feature also potentially allows forgiveness, serving as a reference point for determining whether a source mistakenly marked as malicious exhibits different falloff behaviour after punishment.
+
+\fakepara{Flow duration and size}
+features which describe the length of time a connection has been active, and the amount of data transferred within that time.
+An extraordinarily long flow, having sent a lot of data, could be more likely to be an amplifier.
+
+\fakepara{Correspondence ratio}
+the ratio between upstream and downstream traffic for a source IP.
+We define this to be $C_X = \min(\mathit{up}, \mathit{down})/\max(\mathit{up}, \mathit{down})$, where a value close to 0 indicates strong asymmetry.
+
+\fakepara{$\mathbf{\Delta}$ Send/receive rate}
+the change in traffic rates caused by the last action.
+Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible in these fields.
+
+\fakepara{Mean inter-arrival time (IAT)}
+a measure of how often packets arrive at the agent's parent switch; low IATs indicate a high number of packets per second, and can be a possible marker of malicious behaviour.
+We only make use of the mean IAT of \emph{inbound} traffic.
+
+\fakepara{(Per-window) packet count}
+the amount of packets sent to/from a source over a flow's lifetime (or the current window of measurement), similar in use to flow size and mean IAT.
+
+\fakepara{Mean packet size per-window}
+legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
+Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.
+
+The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
+If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have little to no correlation with the class of traffic a flow might contain.
+Our aim was to design around this constraint as a form of future-proofing.
 
@@
 	\centering
-	\caption{Tile coding windows for each feature.}
+	\caption{Tile coding windows for each feature.\label{tab:codings}}
 
@@
 
-\subsection{What's done with these?}
-
-?? Tile coded together, action taken per-flow every timestep.
-
-?? TODO: everything
-
-??
-What concessions will we have to make in order to make per-flow processing more viable?
-Intelligent sampling/reanalysis of flows when needed (i.e.
-an external heuristic guiding method)?
-In SPIFFY's \cite{DBLP:conf/ndss/KangGS16} evaluation, we see clearly that it can take around \SI{2}{\second} for a flow to react fully to a rate increase---I think for the TCP step it may be wise to factor this in, too!
-
-??
-Explicitly mention \emph{why} we don't make use of neural networks, and instead favour more classical function approximation.
-Latency concerns, energy usage, graphics/TPU/whatever hardware availability or lack thereof across switches/routers/etc.
-
-??
+All of the above features, save for global state, are 1-dimensional.
+Each is tile-coded with 4 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
+We chose to take a separate tile-coding over each feature, ignoring combinatorial effects at present to minimise the amount of extra work an agent must perform.
+Indeed, combining too many features into one tile-coding can lead to poor performance due to an inability to generalise action values across similar states.
+On the other hand, treating each individually allows us to achieve a quick understanding of the importance of each while seeing adequate performance.
+
+\Cref{fig:udp-feature-plots} shows the effectiveness of each for UDP (resp.\ \cref{fig:udp-feature-plots} for TCP), using a topology with $n=2$ hosts per egress point.
+%?? Core findings---different protocols need different features, so everything we proposed above has a use!
+The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
+In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
+%?? How do they do when combined after individual training? Pretty well, especially for TCP.
+Additional testing shows that the learned policies may be effectively combined (by summing action values), and are particularly effective for TCP; these results are omitted to preserve space.
+
+%??
+NEED TO SAY SOMEWHERE: Baselines use TC for establishing interleaving/buffers filling etc., when MARL+ or SPF are in play then disabling TC (thus allowing illegal states, with \num{-1} reward marking them as such---``negative reinforcement'') may accelerate learning.
+We have a choice---but the results show that for TCP we WANT TC enabled...
+For UDP, we actually get a slight degradation in performance.
+This suggests a kind of ``banding'', perhaps?
+Important to mention that it easily learns how to maintain a logical bandwidth limit.
+During training of the above feature models, we elected to remove the physical bandwidth limit to allow agents to visit illegal states and receive \emph{negative reinforcement} for doing so.
+It is important to note that this did not affect the agents' ability to keep the total load below $U_s$, as this behaviour can be easily learned in all cases.
+%?? talk about cap difference: \cref{fig:tcp-cap-feature-plots}---we want cap enabled for TCP.
+We found this led to a marginal improvement in efficacy when UDP was the primary source of benign traffic, but observed that this hindered learning when TCP was the primary protocol under protection (\cref{fig:tcp-cap-feature-plots})---likely due to the loss of key system dynamics.
+It is anticipated that this difference might prove problematic for heterogeneous traffic mixtures one way of solving this may be to investigate the effect of replacing $U_s$ with e.g., $0.99 U_s$ in \cref{eqn:reward-but-better} to introduce a small band of negative reinforcement without actually removing the load restriction in the environment.
+Our intent is to potentially accelerate learning while preserving the effects of system dynamics---we term these \emph{banded} rewards.
+
+%\subsection{What's done with these?}
+
+%?? Tile coded together, action taken per-flow every timestep.
+
+%??
 NOTE: think about how I might need to change/alter the tile-coding when we add extra discriminative features...
@@
 
-??
+%??
 What other consequences are there?
 We're training each agent on all the flows it's acting on which means slight theoretical changes (we have multiple experience traces, and each uses its last state + new state + new target value (reward) as an update to the \emph{current model}).
+
+\section{The Results of Doing So}\label{sec:the-results-of-doing-so}
+
+?? They'd probably look real nice...
+
+?? How to organise? by traffic type, or by model?
+
+?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
+
+\section{Related Work}
+
+%?? Try and compare my work here when possible?
+
+\fakepara{DDoS Prevention}
+\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
+Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
+
+\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
+\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
+%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
+Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
+
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
+Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
+
+\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
+This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
+The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
+Because of this, traditional flooding-based attacks aren't fully alleviated.
+
+%?? Abuses of RL 
+\fakepara{RL in Networks}
+Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
+Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
+Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
+
+For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
+A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
+We feel there is a shortage of work in this category at present, at least in the field of networks; whether this is due to disinterest or difficulty is as-yet unknown.
+To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.
+
+An \emph{indirect-control} RL problem, where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
+A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
+The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
+Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
+
+\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
+Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
+To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
+The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
+The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.
+
+%These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
+
+\section{Discussion}
+
+Talk about flaws here, what could go wrong...
+
+?? What if there are too many flows and a decision set is never completed? Uh... trap door and skip to next set after 50 flow decisions or something?
 
@@
 
+?? Is the state space interpretable? Yes!
+
 ??
-Have some more things worth testing now.
-FSA with transitions based on the theory underlying SPIFFY.
-5 states (start in low pdrop) from \SIrange{0}{100}{\percent} drop, actions move up, down or hold on the level of drop.
-Why like this?
-Chosen deployment environments might not be able to reroute to get the requisite bandwidth epansion factor, so punish everyone to begin with.
-
-\section{The Results of Doing So}\label{sec:the-results-of-doing-so}
-
-?? They'd probably look real nice...
-
-??
-"We chose to take a separate tile-coding over each feature, ignoring combinatorial effects for now.
-Combining them all as one tile-coding led to poor performance due to an inability to generalise value across similar states."
-
-\section{Related Work}
-
-%?? Try and compare my work here when possible?
-
-\fakepara{DDoS Prevention}
-%?? Other recent work in the prevention of DDoS? (i.e., non-RL).
-\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
-Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
-%?? What does it do? detects state, not which flows are contributing...
-
-\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
-\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
-%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
-Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
-
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
-Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
-
-\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
-This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
-The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
-Because of this, traditional flooding-based attacks aren't fully alleviated.
-
-%?? Abuses of RL 
-\fakepara{RL in Networks}
-Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
-Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
-
-For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
-A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
-We feel there is a shortage of work in this category at present, at least in the field of networks; whether this is due to disinterest or difficulty is as-yet unknown.
-To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.
-
-An \emph{indirect-control} RL problem, where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
-%Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works. \cite{DBLP:conf/hotnets/ValadarskySST17, DBLP:conf/hotnets/MaoAMK16}.
-A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
-The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
-
-%?? Completely reword, yikes.
-\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
-Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
-To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
-The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
-The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.
-
-%These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
-
-\section{Discussion}
-
-Talk about flaws here, what could go wrong...
-
-?? What if there are too many flows and a decision set is never completed? Uh... trap door and skip to next set after 50 flow decisions or something?
+What concessions will we have to make in order to make per-flow processing more viable?
+Intelligent sampling/reanalysis of flows when needed (i.e.
+an external heuristic guiding method)?
+In SPIFFY's \cite{DBLP:conf/ndss/KangGS16} evaluation, we see clearly that it can take around \SI{2}{\second} for a flow to react fully to a rate increase---I think for the TCP step it may be wise to factor this in, too!
 

### 2018-11-06 tables/big-avg-reward.py
Tables of average values...

@@ -1,0 +1,102 @@
+import csv
+import numpy as np
+from os import path
+import sys
+
+traffics = ["udp", "tcp"]
+ns = [2, 4, 8, 16]
+models = ["m", "spf"]
+variants = [
+	("natural", "Capped"),
+	("uncap", "Uncapped"),
+	("banded", "Banded"),
+	("single", "Single"),
+]
+
+def get_average_reward(filename):
+	with open(filename, "rb") as csvfile:
+		reader = csv.reader(csvfile)
+		_ = reader.next()
+
+		dats = []
+		for row in reader:
+			dats.append(float(row[3]))
+
+		return np.mean(dats)
+
+def write_table(out_file, data):
+	lm = len(models)
+	lv = len(variants)
+
+	out_file.write(
+		"\\begin{{tabular}}{{@{{}}lc{}@{{}}}}\n".format("S"*((lm*lv)+1))
+	)
+	out_file.write(
+		"\\toprule\\multicolumn{1}{c}{Traffic} & " + \
+		"\\multicolumn{1}{c}{$n$} & " + \
+		"\\multicolumn{1}{c}{Marl} & " + \
+		"\\multicolumn{{{}}}{{c}}{{Marl++}} & ".format(lv) + \
+		"\\multicolumn{{{}}}{{c}}{{SPF}} ".format(lv) + \
+		"\\cmidrule(lr){{4-{}}}\\cmidrule(lr){{{}-{}}} \\\\\n".format(4 + lv-1, 4 + lv, 3 + lv + lv) + \
+		" & & "
+	)
+
+	for _ in models:
+		for (_fn, pres) in variants:
+			out_file.write("& \\multicolumn{{1}}{{c}}{{{}}} ".format(pres))
+
+	out_file.write("\\\\ \\midrule\n")
+
+	# Write file body.
+	for traffic in traffics:
+		for i, n in enumerate(ns):
+			key = (traffic,n)
+			out_file.write("{} & {} ".format("" if i>0 else traffic.upper(), n))
+
+			row = data[key]
+			to_bold = np.argmax(row)
+			for j, val in enumerate(row):
+				base = "& {} " if j != to_bold else "& \\textbf{{{}}} "
+				out_file.write(base.format(val))
+
+			out_file.write("\\\\ \n")
+
+	out_file.write("\\bottomrule\n\\end{{tabular}}")
+
+def get_data():
+	data = {}
+	for t in traffics:
+		for n in ns:
[…]

### 2018-11-06 paper/paper.tex
Improvements to table script, paper.

@@
 
+\usepackage{etoolbox}
 \usepackage[binary-units, per-mode=symbol]{siunitx}
-\sisetup{detect-all, range-phrase=--, range-units=single}
+\robustify\bfseries
+\sisetup{detect-all, range-phrase=--, range-units=single, detect-weight=true,detect-inline-weight=math}
 
@@
 \newcommand{\fakepara}[1]{\noindent\textbf{#1:}}
+
+\makeatletter\let\expandableinput\@@input\makeatother
 
@@
 
-{
-\newcommand{\load}[2]{\operatorname{load}^{#2}_{t}(#1)}
-\newcommand{\uload}[1]{\load{#1}{\uparrow}}
-\newcommand{\dload}[1]{\load{#1}{\downarrow}}
-\newcommand{\bload}[1]{\load{#1}{\updownarrow}}
+\newcommand{\arrload}[2]{\operatorname{load}^{#2}_{t}(#1)}
+\newcommand{\uload}[1]{\arrload{#1}{\uparrow}}
+\newcommand{\dload}[1]{\arrload{#1}{\downarrow}}
+\newcommand{\bload}[1]{\arrload{#1}{\updownarrow}}
 \newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
@@
 \end{subequations}
-replacing $\load{t}{}$ in \crefrange{eqn:reward-rt}{eqn:reward-rctl} with whichever directional load is prioritised.
+replacing $\arrload{t}{}$ in \crefrange{eqn:reward-rt}{eqn:reward-rctl} with whichever directional load is prioritised.
 We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, though we expect that $\bload{\cdot}$ would be the most suitable for general deployment or heterogeneous traffic patterns.
-}
 
@@
 the ratio between upstream and downstream traffic for a source IP.
-We define this to be $C_X = \min(\mathit{up}, \mathit{down})/\max(\mathit{up}, \mathit{down})$, where a value close to 0 indicates strong asymmetry.
+We define this to be $C_X = \min(\uload{\cdot}, \dload{\cdot})/\max(\uload{\cdot}, \dload{\cdot})$, where a value close to 0 indicates strong asymmetry.
 
@@
 Indeed, combining too many features into one tile-coding can lead to poor performance due to an inability to generalise action values across similar states.
-On the other hand, treating each individually allows us to achieve a quick understanding of the importance of each while seeing adequate performance.
-
-\Cref{fig:udp-feature-plots} shows the effectiveness of each for UDP (resp.\ \cref{fig:udp-feature-plots} for TCP), using a topology with $n=2$ hosts per egress point.
+On the other hand, treating features individually allows us to reach a quick understanding of the importance of each while seeing adequate performance.
+
+\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:udp-feature-plots} for TCP), using a topology with $n=2$ hosts per egress point.
 %?? Core findings---different protocols need different features, so everything we proposed above has a use!
@@
 %?? How do they do when combined after individual training? Pretty well, especially for TCP.
-Additional testing shows that the learned policies may be effectively combined (by summing action values), and are particularly effective for TCP; these results are omitted to preserve space.
+Additional testing shows that the learned per-feature policies may be easily combined (by summing action values), and that this technique is particularly effective for TCP; these results are omitted to preserve space.
 
@@
 %?? talk about cap difference: \cref{fig:tcp-cap-feature-plots}---we want cap enabled for TCP.
-We found this led to a marginal improvement in efficacy when UDP was the primary source of benign traffic, but observed that this hindered learning when TCP was the primary protocol under protection (\cref{fig:tcp-cap-feature-plots})---likely due to the loss of key system dynamics.
-It is anticipated that this difference might prove problematic for heterogeneous traffic mixtures one way of solving this may be to investigate the effect of replacing $U_s$ with e.g., $0.99 U_s$ in \cref{eqn:reward-but-better} to introduce a small band of negative reinforcement without actually removing the load restriction in the environment.
+We found this led to a marginal improvement in efficacy when UDP was the primary source of benign traffic, but observed that this hindered learning when TCP was the primary protocol under protection (\cref{fig:tcp-cap-feature-plots}), likely due to the loss of key system dynamics.
+It is anticipated that this difference might prove problematic for heterogeneous traffic mixtures.
+One way of solving this may be to investigate the effect of replacing $U_s$ with e.g., $0.99 U_s$ in \cref{eqn:reward-but-better}, to introduce a small band of negative reinforcement without actually removing the load restriction in the environment.
 Our intent is to potentially accelerate learning while preserving the effects of system dynamics---we term these \emph{banded} rewards.
@@
 
-?? They'd probably look real nice...
-
-?? How to organise? by traffic type, or by model?
-
-?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
-
-\section{Related Work}
-
-%?? Try and compare my work here when possible?
-
-\fakepara{DDoS Prevention}
-\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
-Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
-
-\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
-\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
-%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
-Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
-
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
-Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
-
-\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
-This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
-The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
-Because of this, traditional flooding-based attacks aren't fully alleviated.
-
-%?? Abuses of RL 
-\fakepara{RL in Networks}
-Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
-Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
-
-For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
-A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
-We feel there is a shortage of work in this category at present, at least in the field of networks; whether this is due to disinterest or difficulty is as-yet unknown.
-To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.
-
-An \emph{indirect-control} RL problem, where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
-A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
-The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
-
-\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
-Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
-To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
-The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
-The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.
-
-%These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
+We now examine the performance of our two new models as compared against existing RL work under different traffic behaviour, for varying $n$.
+%?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
+Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
+We present the average rewards for all combinations of these factors in \cref{tab:av-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold.
+
+Unless stated otherwise, these experiments use all \num{18} features introduced in \cref{sec:rethinking-the-state-space} (tile-coded individually), and make use of the physical bandwidth cap rather than pursuing negative reinforcement.
+Test system specifications match those given in \cref{sec:performance-in-an-emulated-environment}.
+
+\begin{table*}
+	\centering
+	\caption{Average reward for combinations of model, host density and traffic class.\label{tab:av-vals}}
+
+	\expandableinput ../tables/big-avg-reward.tex
+\end{table*}
+
+\subsection{Congestion-unaware traffic}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{../plots/udp-2}
+
+	\caption{
+		Online performance for $n=2$ hosts per egress point.
+		\label{fig:udp-2}
+	}
+\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{../plots/udp-16}
+
+	\caption{
+		It's a thing.
+		\label{fig:udp-16}
+	}
+\end{figure}
+
+Yeah, Marl ++ is good. So what. \Crefrange{fig:udp-2}{fig:udp-16}
+
+\subsection{Congestion-aware traffic}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{../plots/tcp-2}
+
+	\caption{
+		It's a thing.
+		\label{fig:tcp-2}
+	}
+\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{../plots/tcp-16}
+
+	\caption{
+		It's a thing.
+		\label{fig:tcp-16}
+	}
+\end{figure}
+
+Yeah, Marl ++ is still good. So what. \Crefrange{fig:tcp-2}{fig:tcp-16}
+
+\subsection{Banded rewards and negative reinforcement}
+
+?? Generally? Not good.
+?? While uncapped led to a per-feature improvement for UDP, Uncapped worse in all cases when all features included.
+
+\subsection{Single-agent performance}
+
+?? Curiously, MARL suffers but SPF actually become usable.
+
+\subsection{Per-feature training}
+
+?? Not yet performed.
 
@@
 Talk about flaws here, what could go wrong...
+
+?? Most importantly this still trails behind existing (exact) DDoS flow detection mechanisms
 
@@
 
-\subsection{Future Work}
+\section{Related Work}
+
+%?? Try and compare my work here when possible?
+
+\fakepara{DDoS Prevention}
+\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
+Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
+
+\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
+\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
+%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
+Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
+
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
+Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
+
+\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
+This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
+The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
+Because of this, traditional flooding-based attacks aren't fully alleviated.
+
+%?? Abuses of RL 
+\fakepara{RL in Networks}
+Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
+Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
+Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
+
+For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
+A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
+We feel there is a shortage of work in this category at present, at least in the field of networks; whether this is due to disinterest or difficulty is as-yet unknown.
+To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.
+
+An \emph{indirect-control} RL problem is one where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
+A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
+The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
+Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
+
+\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
+Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
+To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
+The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
+The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.
+
+%These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
+
+\section{Future Work}
 
 Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.
-
-\section{Conclusion}
-
-It all ends here...
-
-?? Please write me once everything else is in place.
 
@@
 THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
-
-??
-Overarching goal: works which \emph{respect the complexity of the network environment} (rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution.
-\emph{Call-to-action}?
 
@@
 
-?? Can probably fold a lot of this out into discussion or something.
+\section{Conclusion}
+
+It all ends here...
+
+?? Please write me once everything else is in place.
+
+??
+Overarching goal: works which \emph{respect the complexity of the network environment} (rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution.
+\emph{Call-to-action}?
 

### 2018-11-06 tables/big-avg-reward.py

@@ -14,12 +14,15 @@ variants = [
 ]
 
 def get_average_reward(filename):
-	with open(filename, "rb") as csvfile:
+	with open(filename, "r") as csvfile:
 		reader = csv.reader(csvfile)
-		_ = reader.next()
+		# _ = nereader.next()
 
 		dats = []
-		for row in reader:
+		for i, row in enumerate(reader):
+			if i == 0:
+				continue
+
 			dats.append(float(row[3]))
 
 		return np.mean(dats)
@@ -29,7 +32,7 @@ def write_table(out_file, data):
 	lv = len(variants)
 
 	out_file.write(
-		"\\begin{{tabular}}{{@{{}}lc{}@{{}}}}\n".format("S"*((lm*lv)+1))
+		"\\begin{{tabular}}{{@{{}}cc{}@{{}}}}\n".format("S"*((lm*lv)+1))
 	)
 	out_file.write(
 		"\\toprule\\multicolumn{1}{c}{Traffic} & " + \
@@ -37,7 +40,8 @@ def write_table(out_file, data):
 		"\\multicolumn{1}{c}{Marl} & " + \
 		"\\multicolumn{{{}}}{{c}}{{Marl++}} & ".format(lv) + \
 		"\\multicolumn{{{}}}{{c}}{{SPF}} ".format(lv) + \
-		"\\cmidrule(lr){{4-{}}}\\cmidrule(lr){{{}-{}}} \\\\\n".format(4 + lv-1, 4 + lv, 3 + lv + lv) + \
+		"\\\\\n" + \
+		"\\cmidrule(lr){{4-{}}}\\cmidrule(lr){{{}-{}}} ".format(4 + lv-1, 4 + lv, 3 + lv + lv) + \
 		" & & "
 	)
 
@@ -56,12 +60,12 @@ def write_table(out_file, data):
 			row = data[key]
 			to_bold = np.argmax(row)
 			for j, val in enumerate(row):
-				base = "& {} " if j != to_bold else "& \\textbf{{{}}} "
+				base = "& {:.3f} " if j != to_bold else "& \\bfseries {:.3f} "
 				out_file.write(base.format(val))
 
 			out_file.write("\\\\ \n")
 
-	out_file.write("\\bottomrule\n\\end{{tabular}}")
+	out_file.write("\\bottomrule\n\\end{tabular}")
 
 def get_data():
 	data = {}
@@ -93,7 +97,6 @@ def main():
[…]

### 2018-11-07 paper/paper.tex
Discussion, Results.

@@
 \documentclass[conference, letterpaper, 10pt, times]{IEEEtran}
+\usepackage{nopageno}
 
@@
 
+%\newcommand{\mytitle}{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention}
 \newcommand{\mytitle}{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention}
@@
 \usepackage{titling}
-\settowidth{\thanksmarkwidth}{*}
-\setlength{\thanksmargin}{-\thanksmarkwidth}
-
-%% Enable /thanks
-\IEEEoverridecommandlockouts
-\makeatletter
-\def\footnoterule{\relax%
-	\kern-5pt
-	\hbox to \columnwidth{\hfill\vrule width 0.5\columnwidth height 0.4pt\hfill}
-	\kern4.6pt}
-\makeatother
+%\settowidth{\thanksmarkwidth}{*}
+%\setlength{\thanksmargin}{-\thanksmarkwidth}
+%
+%%% Enable /thanks
+%\IEEEoverridecommandlockouts
+%\makeatletter
+%\def\footnoterule{\relax%
+%	\kern-5pt
+%	\hbox to \columnwidth{\hfill\vrule width 0.5\columnwidth height 0.4pt\hfill}
+%	\kern4.6pt}
+%\makeatother
 
@@
 %	\email{k.simpson.1@research.gla.ac.uk}
-Anonymous Giraffe\thanks{This work was supported by a nameless funder [grant number 7].},
+Anonymous Giraffe,
 Anonymous Badger\\
@@
 \emph{Reinforcement learning} (RL) may overcome the detection problem for certain classes of anomaly by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available).
+
+?? We explore the design space, and improve upon existing approaches.
 
@@
 \fakepara{Global state ($\mathbf{4\cdot{}}$Load)}
-these are the existing set of load measurements introduced in \cref{sec:environment-and-rl-algorithm}.
+These are the existing set of load measurements introduced in \cref{sec:environment-and-rl-algorithm}.
 These indicate the overall health of the network, and crucially are all measurements which an agent directly controls.
@@
 \fakepara{Source IP address}
-while ordinarily trivial to spoof, in real-world scenarios the addresses of a set of reflector nodes might exhibit similarity, most likely through geography.
+While ordinarily trivial to spoof, in real-world scenarios the addresses of a set of reflector nodes might exhibit similarity, most likely through geography.
 
 \fakepara{Last action taken}
-this encodes an agent's current belief in the maliciousness of a flow.
+This encodes an agent's current belief in the maliciousness of a flow.
 This feature also potentially allows forgiveness, serving as a reference point for determining whether a source mistakenly marked as malicious exhibits different falloff behaviour after punishment.
@@
 \fakepara{Flow duration and size}
-features which describe the length of time a connection has been active, and the amount of data transferred within that time.
+Features which describe the length of time a connection has been active, and the amount of data transferred within that time.
 An extraordinarily long flow, having sent a lot of data, could be more likely to be an amplifier.
@@
 \fakepara{Correspondence ratio}
-the ratio between upstream and downstream traffic for a source IP.
+The ratio between upstream and downstream traffic for a source IP.
 We define this to be $C_X = \min(\uload{\cdot}, \dload{\cdot})/\max(\uload{\cdot}, \dload{\cdot})$, where a value close to 0 indicates strong asymmetry.
@@
 \fakepara{$\mathbf{\Delta}$ Send/receive rate}
-the change in traffic rates caused by the last action.
+The change in traffic rates caused by the last action.
 Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible in these fields.
@@
 \fakepara{Mean inter-arrival time (IAT)}
-a measure of how often packets arrive at the agent's parent switch; low IATs indicate a high number of packets per second, and can be a possible marker of malicious behaviour.
+A measure of how often packets arrive at the agent's parent switch; low IATs indicate a high number of packets per second, and can be a possible marker of malicious behaviour.
 We only make use of the mean IAT of \emph{inbound} traffic.
@@
 \fakepara{(Per-window) packet count}
-the amount of packets sent to/from a source over a flow's lifetime (or the current window of measurement), similar in use to flow size and mean IAT.
+The amount of packets sent to/from a source over a flow's lifetime (or the current window of measurement), similar in use to flow size and mean IAT.
 
 \fakepara{Mean packet size per-window}
-legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
+Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
 Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.
@@
 All of the above features, save for global state, are 1-dimensional.
-Each is tile-coded with 4 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
+Each is tile-coded with 8 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
 We chose to take a separate tile-coding over each feature, ignoring combinatorial effects at present to minimise the amount of extra work an agent must perform.
@@
 
-\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:udp-feature-plots} for TCP), using a topology with $n=2$ hosts per egress point.
+\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-feature-plots} for TCP), using a topology with $n=2$ hosts per egress point averaged over 10 episodes.
 %?? Core findings---different protocols need different features, so everything we proposed above has a use!
@@
 
-\section{The Results of Doing So}\label{sec:the-results-of-doing-so}
-
-We now examine the performance of our two new models as compared against existing RL work under different traffic behaviour, for varying $n$.
+\section{The Results of Doing So}
+\label{sec:the-results-of-doing-so}
+
+We now examine the performance of our two new models (\emph{Marl++}, \emph{SPF}) as compared against existing RL work (\emph{Marl}) under different traffic behaviour and varying $n$ as before.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
 Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
+
 We present the average rewards for all combinations of these factors in \cref{tab:av-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold.
-
-Unless stated otherwise, these experiments use all \num{18} features introduced in \cref{sec:rethinking-the-state-space} (tile-coded individually), and make use of the physical bandwidth cap rather than pursuing negative reinforcement.
+Average rewards take into account any portions of time that an agent allows illegal system states.
+Several plots augment this, illustrating peak performance or the amount of time which an agent requires to learn.
+
+Unless stated otherwise, these experiments measure average performance over 10 episodes, use all \num{18} features introduced in \cref{sec:rethinking-the-state-space} (tile-coded individually), and make use of the physical bandwidth cap rather than pursuing negative reinforcement.
+In all cases, agents learn control of the network from scratch.
 Test system specifications match those given in \cref{sec:performance-in-an-emulated-environment}.
@@
 	\caption{
-		Online performance for $n=2$ hosts per egress point.
+		Online performance for $n=2$ hosts per egress point when benign traffic is UDP-like.
+		Although Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded), SPF significantly underperforms for this relatively simple topology.
+		Non-SPF agents start off reasonably well, slowly learning better policies.
 		\label{fig:udp-2}
@@
 	\caption{
-		It's a thing.
+		Online performance for $n=16$ hosts per egress point when benign traffic is UDP-like.
+		Marl++ remains marginally ahead of its predecessor, though both have undergone a significant drop in effectiveness.
+		SPF, remarkably, displays performance on par with Marl++ for this more difficult topology.
+		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
 		\label{fig:udp-16}
@@
 
-Yeah, Marl ++ is good. So what. \Crefrange{fig:udp-2}{fig:udp-16}
+Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
+\Cref{fig:udp-2,fig:udp-16} show the behaviour for $n \in \{2, 16\}$ in detail---the remaining choices of $n$ interpolate between these extremes, with the performance difference between Marl and Marl++ shrinking as $n$ increases.
 
@@
 	\caption{
-		It's a thing.
+		Online performance for $n=2$ hosts per egress point when benign traffic is TCP-like.
+		Marl++ and Marl achieve very similar performance, starting off similarly well without notable improvement over an episode.
+		SPF's performance is disappointingly close to baseline, indicating that it is as useful as having no defence system.
 		\label{fig:tcp-2}
@@
 	\caption{
-		It's a thing.
+		Online performance for $n=16$ hosts per egress point when benign traffic is UDP-like.
+		Although Marl++ takes longer to reach its most effective policy, we achieve visibly better peak performance.
+		SPF does manage to learn an improved policy compared to its starting state, but never manages to perform well.
 		\label{fig:tcp-16}
@@
 
-Yeah, Marl ++ is still good. So what. \Crefrange{fig:tcp-2}{fig:tcp-16}
+Marl++ consistently allows more legitimate traffic to reach its destination than Marl, with both significantly above the baseline (\cref{tab:av-vals}, \emph{Capped}, TCP).
+SPF performs consistently poorly across all choices of $n$.
+\Cref{fig:tcp-2}{fig:tcp-16} show two endpoints of a trend in our results: for larger $n$, Marl++ requires around twice as long as Marl to learn its best-achieving policy, underperforming for this period, but achieves notably better peak performance.
 
 \subsection{Banded rewards and negative reinforcement}
-
-?? Generally? Not good.
-?? While uncapped led to a per-feature improvement for UDP, Uncapped worse in all cases when all features included.
+In general, we find that explicitly training a policy to keep the system in a valid state via negative reinforcement (i.e., an \emph{uncapped} network) is most effective for UDP-like traffic with Marl++.
+For SPF, however, negative reinforcement becomes a worse choice for larger $n$.
+Where TCP-like traffic is concerned, we see that this strategy almost always leads to worse policies.
+
+The usefulness of banded rewards with regard to learning an effective policy appears to be unpredictable.
+For many of the configurations detailed in \cref{tab:av-vals}, for $n>2$ the average reward with banding is likely to lie between that of the capped and uncapped system models, but oscillates heavily.
+In a handful of UDP instances, banding can lead to \emph{better} average performance.
+More study is likely required into this; we choose to replace $U_s$ with $0.99U_s$ to minimise the impact on legitimate states, but this choice remains somewhat arbitrary.
 
 \subsection{Single-agent performance}
-
-?? Curiously, MARL suffers but SPF actually become usable.
+Making all decisions with a single agent is roughly equivalent to having a zero-cost communication channel between each pair of agents, theoretically allowing faster training by giving each agent more experience.
+Curiously, we observe that this often leads to drastically worse policies when used as part of Marl++ for small $n$, but makes SPF a considerably more competitive model---coming ahead of Marl for UDP traffic.
+We discuss our conjectures for why this reversal occurs in \cref{sec:discussion}.
 
@@
 
-?? Not yet performed.
-
-\section{Discussion}
-
-Talk about flaws here, what could go wrong...
-
-?? Most importantly this still trails behind existing (exact) DDoS flow detection mechanisms
-
-?? What if there are too many flows and a decision set is never completed? Uh... trap door and skip to next set after 50 flow decisions or something?
+?? Still being performed.
+
+\section{Discussion}\label{sec:discussion}
+
+%Talk about flaws here, what could go wrong...
+
+%?? Why does SPF only do well sometimes? Model is actually more difficult to learn, so it seems to do best when it has a larger set of decisions to learn from. But, it does worse for TCP?
+\fakepara{Model performance}
+Of the results presented, SPF's meandering performance is interesting given its far smaller action space.
+It's natural to expect that this would the model easier to learn, but the additional state required appears to make the model \emph{harder}, beyond even the gains we might expect from choosing a non-zero discount factor.
+Accordingly, we see that SPF performs best when agents learn from as much knowledge as possible: high $n$ and single-agent training.
+Paradoxically, Marl++ generally achieves the best performance yet actively suffers when treated as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.
+
+%?? May be hard to learn multiple features at once while controlling multiple flows while contending with many more agents, with harder dynamics like TCP. Does this hinder learning in the long run?
+Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect, or hope for.
+We suspect that some of these difficulties originate from the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to unpick the true value of each action), and controlling shared global state.
+It may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows, to explore this hypothesis.
+
+%It is likely that the design of Marl++
+
+%Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
+Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are an improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
+Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
+
+\fakepara{Security concerns and vulnerability}
+Can an agent be flooded with new flows to reduce their ability to make decisions?
+One of the risks introduced by our timed random sequential policy updates is that so much work can be queued up that an agent is never able to act on some attack flows.
+The natural solution is to impose an upper bound on the amount of action computations/policy updates that can be performed before a work list is discarded completely.
+This removes the guarantee that all flows will be visited fairly often, but if updates occur regularly then this random sampling my be sufficient to achieve good performance.
+
+%?? Is the state space interpretable? Yes!
+Machine learning algorithms have earned a reputation for eluding human interpretation, while being vulnerable to evasion and poisoning.
+Given the risks associated with introducing such techniques in the context of security, it is natural to be concerned with the interpretability of the models we have proposed.
+With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: $n_{\mathit{tilings}}n_{\mathit{tiles}}\dim{f}$ individual action value vectors per feature $f$, though considerably more combinations thereof.
+Furthermore, system state which is dependent on many signals drawn from across a wide network (such as our global state) is difficult to exert precise control over.
+These signals' topological separation, in concert with their burstiness and unpredictability, may have substantial effects on an attacker's capabilities.
+
+\section{Related Work}
+
+%?? Try and compare my work here when possible?
+
+\fakepara{DDoS Prevention}
+\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
+Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
+
+\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
+\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
+%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
+Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
+
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
+Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
+
+\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
+This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
+The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
+Because of this, traditional flooding-based attacks aren't fully alleviated.
+
+%?? Abuses of RL 
+\fakepara{RL in Networks}
+Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
+Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
+Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
+
+For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
+A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
+We feel there is a shortage of work in this category at present, at least in the field of networks; whether this is due to disinterest or difficulty is as-yet unknown.
+To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.
+
+An \emph{indirect-control} RL problem is one where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
+A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
+The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
+Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
+
+\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
+Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
+To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
+The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
+The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.
+
+%These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
+
+\section{Future Work}
+
+Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.
+
+?? Future Work? I.e., \emph{everything}: no one else is really looking at/interested in this specific kind of application of RL yet. \emph{Yet}.
+
+?? IDEA: try out average reward, TD($\lambda$) methods as future work...
 
@@
 
-?? Is the state space interpretable? Yes!
+??
+IDEA: Apply these techniques to programmable data planes etc.
+While it's pretty neat that what we have works assuming that ach router is a software (x86) switch running OVS, what might we need to consider when applying this to `real' switches?
+``PDP can allow this to be added to real routers to make it efficient to keep \& process state in the manner we require, as well as enabling more adaptive deployment''.
+Cite P4, BPFabric, other work on PDPs?
+
+??
+Benefit of the more realistic emulation environment is that it is far closer in behaviour and architecture (i.e.
+viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance.
+It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques.
+THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
 
@@
 
-\section{Related Work}
-
-%?? Try and compare my work here when possible?
-
-\fakepara{DDoS Prevention}
-\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
-Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
-
-\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
-\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
-%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
-Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
-
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
-Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
-
-\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
-This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
-The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
-Because of this, traditional flooding-based attacks aren't fully alleviated.
-
-%?? Abuses of RL 
-\fakepara{RL in Networks}
-Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
-Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
-
-For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
-A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
-We feel there is a shortage of work in this category at present, at least in the field of networks; whether this is due to disinterest or difficulty is as-yet unknown.
-To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.
-
-An \emph{indirect-control} RL problem is one where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
-A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
-The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
-
-\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
-Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
-To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
-The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
-The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.
-
-%These works emphasise the necessity of ingenuity in effectively handling how states and actions are represented.
-
-\section{Future Work}
-
-Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.
-
-?? Future Work? I.e., \emph{everything}: no one else is really looking at/interested in this specific kind of application of RL yet. \emph{Yet}.
-
-?? IDEA: try out average reward, TD($\lambda$) methods as future work...
-
-??
-IDEA: Apply these techniques to programmable data planes etc.
-While it's pretty neat that what we have works assuming that ach router is a software (x86) switch running OVS, what might we need to consider when applying this to `real' switches?
-``PDP can allow this to be added to real routers to make it efficient to keep \& process state in the manner we require, as well as enabling more adaptive deployment''.
-Cite P4, BPFabric, other work on PDPs?
-
-??
-Benefit of the more realistic emulation environment is that it is far closer in behaviour and architecture (i.e.
-viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance.
-It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques.
-THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
-
 ?? Security? I suspect that the very qualities that make inference difficult in IDS/IPS also increase the level of challenge an advanced threat must overcome.
-System state which is dependent on many signals drawn from across a wide network, in concert with their burstiness and unpredictability, may have substantial effects on an attacker's capabilities.
 ??
@@
 
+We stress once again that this work still trails behind existing (exact) DDoS flow detection mechanisms, yet ...
+
 \section*{Acknowledgements}
@@
 %Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
+This work was supported by a nameless funder [grant number 0].
 

### 2018-11-07 paper/paper.tex
Draft 1...

@@
 \emph{Reinforcement learning} (RL) may overcome the detection problem for certain classes of anomaly by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available).
-
-?? We explore the design space, and improve upon existing approaches.
-
-We present...
-?? Contribs
-
-?? Taking up space to figure out how much room I have for an intro
-
-?? still taking up space...
-
-?? still going...
-
-?? done...
+We examine how the paradigm has been applied to the task of DDoS prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.
+
+These flaws, poor handling of TCP traffic and high host density at egress points, catalyse and direct an exploration of the design space of RL models.
+We improve the state-of-the-art through two agent classes designed to act on a per-flow basis, supported by an in-depth investigation of feature suitability and empirical evaluation.
+
 \end{abstract}
@@
 \begin{itemize}
-	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:rethinking-the-state-space}).
-	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws, and risk factors concerning many traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
+%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
+	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
 	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
-	\item A source-level granularity approach to RL-driven DDoS prevention, improving upon past aggregate-based approaches (\cref{sec:rethinking-the-state-space}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
+	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
+	\item Two source-level granularity approaches to RL-driven DDoS prevention, improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
 \end{itemize}
@@
 
-\section{Acting On Individual Sources}
+\section{Acting On Individual Sources}\label{sec:acting-on-individual-sources}
 Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per source.
@@
 Accordingly, we see that SPF performs best when agents learn from as much knowledge as possible: high $n$ and single-agent training.
-Paradoxically, Marl++ generally achieves the best performance yet actively suffers when treated as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.
+Paradoxically, Marl++ generally achieves the best performance yet actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.
 
@@
 \section{Future Work}
-
-Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.
-
-?? Future Work? I.e., \emph{everything}: no one else is really looking at/interested in this specific kind of application of RL yet. \emph{Yet}.
-
-?? IDEA: try out average reward, TD($\lambda$) methods as future work...
-
-??
+%Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.
+%?? Future Work? I.e., \emph{everything}: no one else is really looking at/interested in this specific kind of application of RL yet. \emph{Yet}.
+
+%?? IDEA: try out average reward, TD($\lambda$) methods as future work...
+A natural research direction to enhance this work would be the combination of the classic function approximation techniques we make use of alongside the improved algorithms that the RL community has introduced in the past few years.
+Actor-critic methods, or algorithms based on eligibility traces are good candidates for investigation.
+
+%??
 What might we do for a reward function in the absence of heuristic estimates and/or explicit a priori knowledge?
@@
 only one in-out link).
-
-??
+A problem we raised (without a clear solution) was the design of reward functions which do not rely upon heuristic estimates or a priori knowledge of benign traffic content.
+If true online learning is desired (i.e., coping with a non-stationary environment), then such reward functions are sorely needed.
+While $\bload{\cdot}$ is likely to be a good candidate for many deployments, we believe that internal ``canary flows'' (or other active measurement) will be effective.
+
+%??
+Benefit of the more realistic emulation environment is that it is far closer in behaviour and architecture (i.e.
+viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance.
+It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques.
+THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
+Given that one of the advantages that RL methods can offer is the ability to handle non-stationary problems, it is important to propose and test sensible simulations or captures of evolving networks against these methods.
+These scenarios present ideal circumstances to apply work on adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}, changepoint detection, or intelligent/heuristic sampling methods to judge which flows are most worthy of consideration per tick.
+
+%??
 IDEA: Apply these techniques to programmable data planes etc.
@@
 Cite P4, BPFabric, other work on PDPs?
-
-??
-Benefit of the more realistic emulation environment is that it is far closer in behaviour and architecture (i.e.
-viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance.
-It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques.
-THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
-
-??
+%??
 What concessions will we have to make in order to make per-flow processing more viable?
@@
 In SPIFFY's \cite{DBLP:conf/ndss/KangGS16} evaluation, we see clearly that it can take around \SI{2}{\second} for a flow to react fully to a rate increase---I think for the TCP step it may be wise to factor this in, too!
-
-?? Security? I suspect that the very qualities that make inference difficult in IDS/IPS also increase the level of challenge an advanced threat must overcome.
-??
+%?? More future work --- share knowledge between agents. ``Knowledge bases'' for this purpose? (see: Qianru).
+Effective real-world deployment of RL-based defences cannot assume that switches in a network will support a custom version of OVS or other arbitrary software, introducing the question of whether agent training, execution and distribution may when using \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
+We also expect it will be fruitful to look into \emph{how} agents may share knowledge with one another, such as via \emph{knowledge bases} or some other efficient communication mechanism.
+
+%?? Security? I suspect that the very qualities that make inference difficult in IDS/IPS also increase the level of challenge an advanced threat must overcome.
+%??
 Might want to mention it in related work above, but the recent attention on adversarial examples/tricking models needs to be looked into for RL.
@@
 There is now examination of evasion attacks wrt.\ RL \cite{DBLP:journals/corr/HuangPGDA17}!
-??
+%??
 evasion attacks by \textcite{DBLP:conf/sp/Carlini017}---all of these are computed by way of a general stochastic optimiser, such as \emph{Adam} \cite{DBLP:journals/corr/KingmaB14}.
@@
 since they're monotonically increasing for the most part).
-
-?? More future work --- share knowledge between agents. ``Knowledge bases'' for this purpose? (see: Qianru).
+Although we believe that the security landscape for classical RL models is not \emph{identical} to that of neural-network based approaches (particularly with such noisy, volatile, and hard-to-control data), there is still immense value in determining the exact capabilities of a sufficiently powerful adversary as the risk of external control still exists.
+In particular, we believe that poisoning attacks and evasion attacks merit close consideration.
 
@@
 
-It all ends here...
-
-?? Please write me once everything else is in place.
-
-??
-Overarching goal: works which \emph{respect the complexity of the network environment} (rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution.
-\emph{Call-to-action}?
-
-We stress once again that this work still trails behind existing (exact) DDoS flow detection mechanisms, yet ...
+Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
+Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.
+
+We have identified weaknesses in past approaches arising from simulation assumptions, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
+Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
+Our evaluation shows that our new agent designs advance the state-of-the-art in RL-based DDoS prevention.
+
+We stress once again that this work still trails behind the performance of exact DDoS flow detection mechanisms, yet we hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
+It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
+%(rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?
 

### 2018-11-07 paper/paper.tex
Fix math, temporarily stop mentioning new sections.

@@
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
-Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
+Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, and the case where one agent makes all decisions.%, and pre-training on individual features.
 
@@
 
-\subsection{Per-feature training}
-
-?? Still being performed.
+%\subsection{Per-feature training}
+%
+%?? Still being performed.
 
@@
 Given the risks associated with introducing such techniques in the context of security, it is natural to be concerned with the interpretability of the models we have proposed.
-With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: $n_{\mathit{tilings}}n_{\mathit{tiles}}\dim{f}$ individual action value vectors per feature $f$, though considerably more combinations thereof.
+With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
 Furthermore, system state which is dependent on many signals drawn from across a wide network (such as our global state) is difficult to exert precise control over.

### 2018-11-07 code/marl/tcp-model-combo-channel-end.py
Per-feature training, but seperated out for each n.

@@ -1,0 +1,98 @@
+import cPickle
+from marl import *
+import sys
+from writer import writeResults, makeResultsAverage
+
+ft = __import__("tcp-model-combo-channel-prep")
+
+def run(restrict, state, rewards, good_traffic_percents, total_loads, contributors=None, store_sarsas=[], ep_len=10000, h_p=2):
+	results = marlExperiment(
+		n_teams = 2,
+
+		n_inters = 2,
+		n_learners = 3,
+		host_range = [h_p, h_p],
+
+		explore_episodes = 0.8,
+		episodes = 1,
+		episode_length = ep_len,
+		separate_episodes = True,
+
+		alpha = 0.05,
+		epsilon = 0.2,
+		discount = 0,
+
+		dt = 0.01,
+
+		rf = "ctl",
+		use_controller = True,
+		actions_target_flows = restrict is not None,
+
+		restrict = restrict,
+
+		rand_state = state,
+		rewards = rewards,
+		good_traffic_percents = good_traffic_percents,
+		total_loads = total_loads,
+
+		contributors = contributors,
+		store_sarsas = store_sarsas,
+
+		evil_range = [4,7],
+		model = "nginx",
+		reward_direction = "out",
+		randomise = True,
+		randomise_count = 3,
+		randomise_new_ip = True,
+
+		estimate_const_limit = True,
+	)
+
+	return results
+
+# For each episode, we want to...
+
+combo_result_sets = ([], [], [])
+
+configs = [
+	(combo_result_sets, False, "combine", 10000),
+]
+
+if __name__ == "__main__":
+	# read from first arg
+	hosts_p = ft.host_ps[int(sys.argv[1])]
+	file_dir = ft.file_dir.format(hosts_p)
+
+	with open(file_dir, "rb") as of:
+		data = cPickle.load(of)
+
+	for (randstate, contributors) in data:
+		g_contrib = (contributors[0][0], [a for a in xrange(4)])
+
+		feature_contribs = contributors[1:]
+
+		for (target_results, start_fresh, _, ep_len) in configs:
+			(rewards, good_traffic_percents, total_loads) = target_results
+			(rs, gs, ls, store_sarsas, rng_state, _) = run(
+				g_contrib[1],
+				randstate[1],
+				rewards, good_traffic_percents, total_loads,
+				# contributors=[g_contrib],
+				contributors=feature_contribs,
[…]

### 2018-11-07 code/marl/tcp-model-combo-channel-prep.py

@@ -1,0 +1,129 @@
+import cPickle
+from marl import *
+import sys
+from writer import writeResults, makeResultsAverage
+
+def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
+	results = marlExperiment(
+		n_teams = 2,#5,
+
+		n_inters = 2,
+		n_learners = 3,
+		host_range = [hosts_p, hosts_p],
+
+
+		explore_episodes = 0.8,
+		episodes = 1,
+		episode_length = 10000,
+		separate_episodes = True,
+
+		alpha = 0.05,
+		epsilon = 0.2,
+		discount = 0,
+
+
+		dt = 0.01,
+
+	#	old_style=True,
+
+		rf = "ctl",
+		use_controller = True,
+		actions_target_flows = restrict is not None,
+
+		restrict = restrict,
+
+		rand_state = state,
+		rewards = rewards,
+		good_traffic_percents = good_traffic_percents,
+		total_loads = total_loads,
+
+		evil_range = [4,7],
+		model = "nginx",
+		reward_direction = "out",
+		randomise = True,
+		randomise_count = 3,
+		randomise_new_ip = True,
+
+		estimate_const_limit = True,
+	)
+
+	return results
+
+# Run one set to old mode, run 8 on restrict [4+i for i in xrange(8)]
+# Build up results each time, should have 9 sets. collect RNG states for each set
+
+host_ps = [2, 4, 8, 16]
+results_dir = "../../results/"
+file_dir = results_dir + "tcp-combo-channel-prep-{}.pickle"
+in_progress_file_dir = results_dir + "tcp-combo-channel-prg-{}.pickle"
+chosen_features = [
+	5, # Mean IAT
+	7, # Delta Out Rate
+	6, # Delta In Rate
+]
+n_episodes = 10
+n_episodes_per_step = 5
+restrict_sets = [None] + [[4 + i] for i in chosen_features]
+out_names = ["g"] + ["f{}".format(i) for i in chosen_features]
+
+result_sets = [([], [], []) for i in xrange(1 + len(chosen_features))]
+
+if __name__ == "__main__":
+	# read from first arg
+	hosts_p = host_ps[int(sys.argv[1])]
+	file_dir = file_dir.format(hosts_p)
+	in_progress_file_dir = in_progress_file_dir.format(hosts_p)
+
+	# if second arg exists, then we're "in-progress"
+	# pickle has last experiment state...
+	start_i = 0
[…]

### 2018-11-07 code/marl/channel-run.sh
Runner, fixes.

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+tests=(0 3)
+#tests=(1 2)
+
+for v in "${tests[@]}"
+do
+	./purger.sh
+	python ./tcp-model-combo-channel-prep.py $v
+	./purger.sh
+	python ./tcp-model-combo-channel-prep.py $v "prog"
+	./purger.sh
+	python ./tcp-model-combo-channel-end.py $v
+	./purger.sh
+done

### 2018-11-07 code/marl/tcp-model-combo-channel-prep.py

@@ -94,7 +94,7 @@ if __name__ == "__main__":
 			(rs, gs, ls, store_sarsas, rng_state, _) = run(
 				restriction,
 				state[1],
-				rewards, good_traffic_percents, total_loads
+				rewards, good_traffic_percents, total_loads,
 				hosts_p
 			)
 

### 2018-11-08 paper/paper.tex
Ability to selectively include results in table.

@@
 
-These flaws, poor handling of TCP traffic and high host density at egress points, catalyse and direct an exploration of the design space of RL models.
+These flaws---poor handling of TCP traffic and high host density at egress points---catalyse and direct an exploration of the design space of RL models.
 We improve the state-of-the-art through two agent classes designed to act on a per-flow basis, supported by an in-depth investigation of feature suitability and empirical evaluation.

### 2018-11-08 tables/big-avg-reward.py

@@ -7,11 +7,17 @@ traffics = ["udp", "tcp"]
 ns = [2, 4, 8, 16]
 models = ["m", "spf"]
 variants = [
-	("natural", "Capped"),
-	("uncap", "Uncapped"),
-	("banded", "Banded"),
-	("single", "Single"),
+	("natural", None, "Capped"),
+	("uncap", None, "Uncapped"),
+	("banded", None, "Banded"),
+	("single", None, "Single"),
+	("channel", "../results/tcp-combo-{2}-{3}.csv", "Pretrain"),
 ]
+
+illegal_combos = set()
+illegal_combos.add(("m", "udp", "channel"))
+illegal_combos.add(("spf", "udp", "channel"))
+illegal_combos.add(("spf", "tcp", "channel"))
 
 def get_average_reward(filename):
 	with open(filename, "r") as csvfile:
@@ -46,7 +52,7 @@ def write_table(out_file, data):
 	)
 
 	for _ in models:
-		for (_fn, pres) in variants:
+		for (_fn, _replace_str, pres) in variants:
 			out_file.write("& \\multicolumn{{1}}{{c}}{{{}}} ".format(pres))
 
 	out_file.write("\\\\ \\midrule\n")
@@ -58,9 +64,25 @@ def write_table(out_file, data):
 			out_file.write("{} & {} ".format("" if i>0 else traffic.upper(), n))
 
 			row = data[key]
+			for j, _ in enumerate(row):
+				if j == 0:
+					continue
+
+				j -= 1
+				# print(j, len(variants))
+				model = models[int(j/len(variants))]
+				variant = variants[j%len(variants)][0]
+				if (model, traffic, variant) in illegal_combos:
+					row[j+1] = -1000.0
+
 			to_bold = np.argmax(row)
+
 			for j, val in enumerate(row):
 				base = "& {:.3f} " if j != to_bold else "& \\bfseries {:.3f} "
+
+				if val < -999.0:
+					base = "& \\multicolumn{{1}}{{c}}{{---}} "					
+
 				out_file.write(base.format(val))
 
 			out_file.write("\\\\ \n")
@@ -84,8 +106,9 @@ def get_data():
 			# do it for all parts now.
 			for m in models:
 				for v in variants:
+					f_str = "../results/{}-{}-{}-{}.csv" if v[1] is None else v[1]
 					row.append(get_average_reward(
[…]

### 2018-11-12 paper/paper.tex
Applied Feedback

--- 2018-11-07 paper/paper.tex
@@
 \begin{abstract}
-Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-heralded, but face problems hampering effective deployment.
-The detection problem in this domain is fraught with difficulty; it is an evolving, non-stationary problem as usage patterns shift, new protocols and applications are introduced, and is compounded by burstiness and seasonal variation.
+Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.
+The detection problem in this domain presents unique difficulties; it is an evolving, non-stationary problem as usage patterns shift and new protocols and applications are introduced, further compounded by burstiness and seasonal variation.
 
@@
 We examine how the paradigm has been applied to the task of DDoS prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.
-
-These flaws, poor handling of TCP traffic and high host density at egress points, catalyse and direct an exploration of the design space of RL models.
-We improve the state-of-the-art through two agent classes designed to act on a per-flow basis, supported by an in-depth investigation of feature suitability and empirical evaluation.
+These flaws---poor handling of TCP traffic and high host density at egress points---inspire and direct an exploration of the design space of RL models.
+We advance the state-of-the-art of DDoS mitigation using RL by introducing two agent classes designed to act on a per-flow basis, supported by an in-depth investigation of feature suitability and empirical evaluation.
+
+Our results show the existence of flow features with high predictive power for environments with predominantly TCP or UDP traffic, when such features are used as a basis for feedback-loop-like control.
+We show that the new RL agent models can offer a significant increase in goodput of legitimate UDP traffic for many choices of host density, alongside a marked increase in protected TCP traffic at high host density.
 
@@
 
-Network anomaly detection, intrusion detection and intrusion prevention are continually evolving problems, compounded by the partial, non-IID view of data at each point in the network.
-Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, or availability without being detected.
+Network anomaly detection and intrusion detection/prevention are continually evolving problems, compounded by the partial, non-IID view of data at each point in the network.
+Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, and availability without being detected \cite{DBLP:journals/comsur/BhuyanBK14}.
 These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
-This issue is inherent to \emph{misuse-} or \emph{signature-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making effective use of statistical measures.
-
-While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \textcite{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems---to quite a large extent, this remains the case today.
+This issue is inherent to \emph{misuse-} or \emph{signature-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making effective use of statistical measures \cite{DBLP:journals/comsur/BhuyanBK14}.
+
+While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \citeauthor{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems---to quite a large extent, this remains the case today \cite{DBLP:conf/sp/SommerP10}.
 Their application is made difficult due to significant operational differences from standard ML tasks, alongside certain characteristics of network traffic.
@@
 
-For certain classes of problem like flooding-based DDoS attacks, \emph{reinforcement learning} (RL) offers another perspective.
-The role of an RL agent differs from that of a standard classifier, adaptively reacting to threats by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
+For certain classes of problem e.g., flooding-based DDoS attacks, \emph{reinforcement learning} (RL) offers another perspective.
+The role of a RL agent differs from that of a standard classifier, adaptively reacting to threats by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
 In a sense, this allows us to ``overcome'' the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time; by looking for the effect rather than the cause.
@@
 
-To date, there have been few applications of this class of algorithms towards intrusion prevention which make use of their full potential for online control, rather than using them as the basis of a classifier.
+To date, there have been few applications of this class of algorithms towards intrusion detection and prevention which make use of their full potential for online control, rather than using them as the basis for a classifier.
 We aim to take steps to redress this and establish their proper capabilities, beyond simple ``blind application''.
+
+?? Expand as required
+
 In service of this goal, this paper contributes:
@@
 %	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
-	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
+	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}.
+We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling for deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
 	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
@@
 %?? Introduce RL, related definitions etc.
-\subsection{Reinforcement Learning}
-\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
-We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
-This model of system interaction is best described as a \emph{Markov decision process} (MDP).
-RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
+\subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
+\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
+We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$, it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
+This model of system interaction is best described as a \emph{Markov Decision Process} (MDP).
+RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal if an agent needs to further explore some region of the state space.
 The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
-From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+2} + \ldots$, choosing the discount factor $\gamma \in [0,1)$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
+In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new, custom-built training corpus.
+
+From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+3} + \ldots$, choosing the discount factor $\gamma \in [0,1)$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
 Formally, an agent's goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.
@@
 To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+ \dim{\mathcal{A}})$-dimensional grids with an optional bias component.
-The number of tilings has a direct effect on the resolution of the discretisation.
+The numbers of tilings and of tiles per dimension have a direct effect on the resolution of the discretisation.
 Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state-action pair exists within the corresponding tile.
@@
 
+This combination of algorithm and coding strategy is well-optimised.
 If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
-Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices).
 Action values for any state are then obtained by summation of the weight vectors from all activated tiles.
-Moreover, we need not store action values for tiles which have not yet been visited, conserving memory.
+Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
+This means that we need only perform $n_{\mathit{tilings}}$ floating point additions per model update:
+\begin{equation}
+	\bm{w}_{t+1}[i][\operatorname{index}(A_t)] = \bm{w}_{t}[i][\operatorname{index}(A_t)] + \alpha \delta_t, \forall i \in s_{\mathit{list}}.
+	\label{eqn:sg-sarsa-opt}
+\end{equation}
+If desired we may define a state-space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
+Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.
 
@@
 The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
-These are APT-level capabilities, whose exploration presents a rich source of future work.
-
-\section{Analysing Existing Work}\label{sec:environment-and-rl-algorithm}
+These are APT-level capabilities, whose exploration presents a rich source for future work.
+
+\section{Foundational Analysis}\label{sec:environment-and-rl-algorithm}
 %?? State the problem and what we want to achieve.
@@
 The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant: their work helps to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
-We begin with a reimplementation and more principled re-examination of their work to further develop the state-of-the-art.
+We begin with a reimplementation and more principled re-examination of their work to further develop the state-of-the-art, while demonstrating the viability of deployment and integration of these techniques within a software-defined network.
 
@@
 
-The network itself is tree-structured, where one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $l$ intermediate switches, which in turn each connect to $m$ egress switches.
+The network itself is tree-structured, where one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $\ell$ intermediate switches, which in turn each connect to $m$ egress switches.
+We say, then, that we have $k$ teams and $\ell m$ agents in each team.
 Agents are co-located with these egress switches, who control the proportion of upstream packets from $n$ external hosts to discard according to load statistics observed along their path to the server.
+
+We assume that the deployment environment is a moderately complex software-defined network, because the paradigm offers features which can directly benefit RL agents acting within.
+The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
+Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allows relocation and easy installation of learners.
+
 Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
 Each link has a delay of \SI{10}{\milli\second}.
-All links have unbounded capacity, save for the server-switch connection which is capped at a fixed $U_s$ \si{\mega\bit\per\second} in both directions.
+All links have unbounded capacity, save for the server-switch connection which is capped at a fixed $U_s$ \si{\mega\bit\per\second} in both directions; in practice, volumetric DDoS attacks exhibit a \emph{saturation zone} \cite{DBLP:journals/tnsm/SimpsonSMJPH18} which we aim to model while minimising the performance costs of emulation.
 \Cref{fig:marl-topol} demonstrates this visually.
@@
 \node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
-\node [labeltexts, right = -0.1cm of ldots] {$l$};
+\node [labeltexts, right = -0.1cm of ldots] {$\ell$};
 \node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
@@
 \caption{
-	Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $l$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
+	Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $\ell$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
 %	Empty nodes are considered to be internal.
@@
 
-Why use this topology, a fan-out tree, in particular?
-While this is taken verbatim from the work we use as a basis, it is presented within the confines of a framework given by \textcite{DBLP:journals/ton/YauLLY05}.
-This framework is, in turn, an extension of the class of topologies examined in the original pushback paper \cite{DBLP:journals/ccr/MahajanBFIPS02a}; this topology is far closer to the latter.
-Crucially, it seems to not actually represent or correspond to any realistic deployment at all.
-In the words of \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}:
-\begin{displaycquote}{DBLP:journals/ccr/MahajanBFIPS02a}
-	These simulations do not pretend to use realistic topologies or traffic mixes... the simple simulations in this scenario are instead intended... as a first step towards more rigorous evaluation.
-\end{displaycquote}
+%Why use this topology, a fan-out tree, in particular?
+%While this is taken verbatim from the work we use as a basis, it is presented within the confines of a framework given by \textcite{DBLP:journals/ton/YauLLY05}.
+%This framework is, in turn, an extension of the class of topologies examined in the original pushback paper \cite{DBLP:journals/ccr/MahajanBFIPS02a}; this topology is far closer to the latter.
+%Crucially, it seems to not actually represent or correspond to any realistic deployment at all.
+Although \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}, the originators of this topology, make it clear that it exists as a fairly unrepresentative example \cite{DBLP:journals/ccr/MahajanBFIPS02a},
+%\begin{displaycquote}{DBLP:journals/ccr/MahajanBFIPS02a}
+%	These simulations do not pretend to use realistic topologies or traffic mixes... the simple simulations in this scenario are instead intended... as a first step towards more rigorous evaluation.
+%\end{displaycquote}
 %The topology chosen to evaluate our predecessor, indeed, matches theirs.
@@
 %Remarkably, it seems to have propagated in a way which was not originally intended.
-It is clear that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network, but it is hard to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
-We aim to address this, to some extent.
+it remains the case that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
+It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
+We aim to address this through \cref{sec:performance-in-an-emulated-environment}.
 
@@
 Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
-While unrealistic with respect to packet content, it was assumed that this would provide representative load and packet inter-arrival characteristics, since the agents as defined rely only upon load measurements.
+This provides representative load and packet inter-arrival characteristics, since the agents as defined rely only upon load measurements.
 
@@
 Specifically, the agent chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
-In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}.
+In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so exhibit a risk of introducing collateral damage into the network.
+These choices primarily ensure that this rate reduction imposed on a source IP may never be performance.
 It's worth noting that there are various ways that this could be done, and that the application of \emph{programmable data planes} to this end are suggested as future work.
@@
 These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
-At time $t$ an agent $v$ (with team leader $l_v$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
+At time $t$, an agent $v$ (with team leader $l_v$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
 \begin{subequations}
@@
 
-The tile-coded state and chosen reward function are then used as input for the \emph{semi-gradient Sarsa} algorithm described in \cref{eqn:sg-sarsa}, making use of $\epsilon$-greedy action selection.
+The tile-coded state and chosen reward function are then used as inputs for the \emph{semi-gradient Sarsa} algorithm described in \cref{eqn:sg-sarsa}, making use of $\epsilon$-greedy action selection.
 Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their experience or weight vector updates with one another.
@@
 \section{Performance in an Emulated Environment}\label{sec:performance-in-an-emulated-environment}
-The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and assumes a rigid and unrepresentative topology.
-The latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
+The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and high densities of hosts at egress points.
+%?? Why? ...
+Of these, the latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
+
+These are important issues, particularly when we consider real-world deployment.
+Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most-used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
+Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network.
 
@@
 In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
+What this means is that datacentres and small-scale service deployments are more likely to have a host-to-learner ratio scaled heavily in favour of hosts, i.e. a large observed $n$.
 We feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions (as defined above) become less granular.
 
-To test this, we configured the network topology using $k=2$ teams, $l=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
+To test this, we configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
 This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single-server with a fan-out network.
@@
 Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
-$U_s$ was fixed at $klmn+2$ \si{\mega\bit\per\second}.
+$U_s$ was fixed at $k \ell mn+2$ \si{\mega\bit\per\second}.
 The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
 Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
+These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
 
@@
 
-\Cref{fig:marl-granularity} shows that as more hosts are allocated to each learner, the fraction of traffic believed to be good (as observed at the server) decreases.
+\Cref{fig:marl-granularity} shows that, as more hosts are allocated to each learner, the fraction of traffic believed to be good (as observed at the server) decreases.
 This manifests in two ways: the best-achievable performance drops, and so too does the learning rate.
@@
 \begin{thm}
-	As $n$ increases, it is more likely that a throttling switch will exhibit imperfect service: $\forall n \in \mathbb{Z}^{+}, P_{M,n} \le P_{M,n+1}$.
+	As the host/learner ratio $n$ increases, it is more likely that a throttling switch will exhibit imperfect service: $\forall n \in \mathbb{Z}^{+}, P_{M,n} \le P_{M,n+1}$.
 \end{thm}
@@
 %\Cref{fig:mplex-tcp} shows that this trend is present.
-Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is based upon does not capture the key interactions between TCP flows and packet drop as a control action.
-The TCP traffic as it appears here is generated according to \Cref{sec:a-new-normal}.
+Given the UDP-like behaviour of Tcpreplay, replicating the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is based upon does not capture the key interactions between TCP flows and packet drop as a control action.
+To the best of our knowledge, existing RL applications for this task employ the same formulation with incremental improvements, and so such a deficiency would be likely to occur in \emph{all} these prior works.
+
 From the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}, through which the performance characteristics of TCP under packet drop are well-known and well-understood, their results should be considerably worse in practice:
@@
 Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant; good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
-\Cref{fig:nginx-coffin-nail} shows that this model performs poorly when attempting to safeguard TCP traffic, though remains better than leaving the network unprotected.
+
+TCP traffic as it appears here is generated according to \Cref{sec:a-new-normal}.
+\Cref{fig:nginx-coffin-nail} shows that this model performs poorly when attempting to safeguard TCP traffic, though it remains better than leaving the network unprotected.
 %?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
@@
 
-Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
+%Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
 Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
@@
 Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
+We contribute a network model built around live testing of TCP traffic in an SDN-enabled environment, adaptable to arbitrary topologies, with an explicit focus on preserving the real-time dynamics of TCP traffic.
 In particular, we are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
-We make no claims that it is perfect or representative for all traffic, only that it captures some of the behaviour which we expect will plague most legitimate TCP flows; if need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.
+Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
+If need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.
 
@@
 %?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
-To model legitimate traffic the server node runs an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
+To model legitimate traffic, the server node runs an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
 Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
@@
 \end{itemize}
+We offer these to model two extremes: long-life host-to-server transfers (a characteristic we might e	xpect from e.g., streaming video), and common HTTP traffic.
 
@@
 Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that simulation remains tractable.
-When using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows.
+When using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows---a \SI{30}{\percent} rate increase in expectation, uploading \SIrange{2.03}{2.18}{$\!
+\times U_s$}.
+The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$
 
@@
 Taking actions in this manner means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
-This introduces numerous potential issues: inability to response to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated; at their worst, these present additional attack surface to an adversary.
+This introduces numerous potential issues: inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated; at their worst, these present additional attack surfaces to an adversary.
 To adapt to these problems, we make use of \emph{timed random sequential} updates.
@@
 We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, though we expect that $\bload{\cdot}$ would be the most suitable for general deployment or heterogeneous traffic patterns.
+These choices reflect where the bulk of transmitted bytes in each model are observed (and the lack of this knowledge in the general case).
 
@@
 Although this likely brings lower theoretical performance, there are more reasons than latency alone to favour classical methods; these include lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
+This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without the introduction of dedicated tensor processing hardware to networks.
+Simpler decision boundaries reduce the risk of overfitting and of unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.
+
+%?? Expand all this
 
@@
 An important question to ask while exploring the design space of RL agents is whether (and how) domain knowledge can improve the performance or learning rate of a defence system.
-This often comes as a trade-off: expressive power and unfettered control, weighed against the smaller action spaces afforded by existing insight or structure.
+This often comes as a trade-off: expressive power and control, weighed against the smaller action spaces afforded by existing insight or structure.
 We propose a second model, relying on the measurements of \textcite{DBLP:conf/ndss/KangGS16} which suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
@@
 How might we apply their observations within the RL paradigm, while tailoring them towards non-ISP deployment?
-Our approach is to constrain how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, reducing the action set to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ in single steps.
-To enable temporary bandwidth expansion in resource constrained deployments, every flow is initially placed under light packet drop ($p=0.05$, or state 1), corresponding to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
+Our approach is to constrain how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, choosing to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
+The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
 This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and non-congestion-aware traffic as required.
+
+The action set is then reduced such that agents can \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ in single steps.
+To enable temporary bandwidth expansion in resource constrained deployments, every flow is initially placed under light packet drop ($p=0.05$, or state 1); this is chosen above the equivalent for UDP due to TCP's higher prevalence.
 We combine this with the enhancements described in \cref{sec:marl-plus}.
@@
 This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
-We found $\gamma = 0.8$ to be the most effective choice during exploratory testing with UDP traffic.
+Recounting \cref{sec:reinforcement-learning}, the choice of $\gamma$ affects which fraction of the next current state-action pair's value is added to $R_t$ when updating the last state-action pair's value.
+We found the setting  $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing with UDP traffic, by observing average performance over $\gamma$ increments of $0.1$.
 
@@
 
-\begin{figure*}
+\begin{figure}
 	\centering
-	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-good}
-	\end{subfigure}
-	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
-	\end{subfigure}
-	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
-	\end{subfigure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-good}
+%	\end{subfigure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
+%	\end{subfigure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
+%	\end{subfigure}
+	\includegraphics[width=\linewidth]{../plots/ftprep-box}
 	\caption{
@@
 	}
-\end{figure*}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.8\linewidth]{../plots/ftprep-good-2}
+	\caption{
+		Demonstration of learning rate for several features included as part of \cref{fig:udp-feature-plots}.
+		Performance starts low, before achieving convergence at some later time; this explains the high volume of outliers occurring below the main distribution of many features.
+		\label{fig:udp-fp-converge}
+	}
+\end{figure}
 
@@
 
-\begin{figure*}
-	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good}
-	\end{subfigure}
-	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good-2}
-	\end{subfigure}
-	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good-3}
-	\end{subfigure}
+\begin{figure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good}
+%	\end{subfigure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good-2}
+%	\end{subfigure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good-3}
+%	\end{subfigure}
+	\includegraphics[width=\linewidth]{../plots/ftprep-tcp-box}
 
@@
 	}
-\end{figure*}
-
-\begin{figure*}
-	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good}
-	\end{subfigure}
-	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-2}
-	\end{subfigure}
-	\begin{subfigure}{0.32\linewidth}
-		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-3}
-	\end{subfigure}
+\end{figure}
+
+\begin{figure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good}
+%	\end{subfigure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-2}
+%	\end{subfigure}
+%	\begin{subfigure}{0.32\linewidth}
+%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-3}
+%	\end{subfigure}
+	\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-box}
 
@@
 	}
-\end{figure*}
+\end{figure}
 
@@
 
-\fakepara{Global state ($\mathbf{4\cdot{}}$Load)}
+\fakepara{Global state}
 These are the existing set of load measurements introduced in \cref{sec:environment-and-rl-algorithm}.
@@
 \fakepara{Source IP address}
-While ordinarily trivial to spoof, in real-world scenarios the addresses of a set of reflector nodes might exhibit similarity, most likely through geography.
+While ordinarily trivial to spoof, in real-world scenarios the addresses of a set of reflector nodes might exhibit similarity, e.g., similarly unhardened services exposed by a single organisation.
 
@@
 Features which describe the length of time a connection has been active, and the amount of data transferred within that time.
-An extraordinarily long flow, having sent a lot of data, could be more likely to be an amplifier.
+An extraordinarily long flow, having sent a lot of data, could be more likely to be an amplifier: though most (\SI{62}{\percent}) waves of amplifier traffic last shorter than \SI{15}{\minute} \cite{DBLP:conf/raid/KramerKMNKYR15}, this is considerably longer than the typical length of an HTTP request/response.
 
@@
 Additional testing shows that the learned per-feature policies may be easily combined (by summing action values), and that this technique is particularly effective for TCP; these results are omitted to preserve space.
+In no cases, however, do we manage to completely block attack traffic---at convergence, we observe that system load remains consistently at $U_s$.
 
@@
 During training of the above feature models, we elected to remove the physical bandwidth limit to allow agents to visit illegal states and receive \emph{negative reinforcement} for doing so.
-It is important to note that this did not affect the agents' ability to keep the total load below $U_s$, as this behaviour can be easily learned in all cases.
+While this means that agents must learn how to keep traffic below a logical limit, they are helped by the fact that strong negative reward signals offer explicit information about which actions are poor choices: observe from \cref{eqn:reward,eqn:reward-but-better} that if a physical link has limit $U_s$ then we will never see a reward below 0.
+It is important to note that neither setting affects the agents' ability to keep the total load below $U_s$.
 %?? talk about cap difference: \cref{fig:tcp-cap-feature-plots}---we want cap enabled for TCP.
@@
 One way of solving this may be to investigate the effect of replacing $U_s$ with e.g., $0.99 U_s$ in \cref{eqn:reward-but-better}, to introduce a small band of negative reinforcement without actually removing the load restriction in the environment.
+It is crucial to make this region as small as possible, hence the somewhat arbitrary choice; we observe that agents are always able to learn how to keep traffic consistently below the logical bandwidth limit (as in the original work).
 Our intent is to potentially accelerate learning while preserving the effects of system dynamics---we term these \emph{banded} rewards.
@@
 
-\section{The Results of Doing So}
+\section{Results}
 \label{sec:the-results-of-doing-so}
@@
 
-Unless stated otherwise, these experiments measure average performance over 10 episodes, use all \num{18} features introduced in \cref{sec:rethinking-the-state-space} (tile-coded individually), and make use of the physical bandwidth cap rather than pursuing negative reinforcement.
+Unless stated otherwise, these experiments measure average performance over 10 episodes, use all \num{14} features introduced in \cref{sec:rethinking-the-state-space} (tile-coded individually) alongside the global state, and make use of the physical bandwidth cap rather than pursuing negative reinforcement.
 In all cases, agents learn control of the network from scratch.
@@
 \subsection{Congestion-unaware traffic}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.9\linewidth]{../plots/udp-2}
+%	
+%	\caption{
+%		Online performance for $n=2$ hosts per egress point when benign traffic is UDP-like.
+%		Although Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded), SPF significantly underperforms for this relatively simple topology.
+%		Non-SPF agents start off reasonably well, slowly learning better policies.
+%		\label{fig:udp-2}
+%	}
+%\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.9\linewidth]{../plots/udp-16}
+%	
+%	\caption{
+%		Online performance for $n=16$ hosts per egress point when benign traffic is UDP-like.
+%		Marl++ remains marginally ahead of its predecessor, though both have undergone a significant drop in effectiveness.
+%		SPF, remarkably, displays performance on par with Marl++ for this more difficult topology.
+%		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
+%		\label{fig:udp-16}
+%	}
+%\end{figure}
 \begin{figure}
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/udp-2}
+	\includegraphics[width=0.9\linewidth]{../plots/udp-box}
 
 	\caption{
-		Online performance for $n=2$ hosts per egress point when benign traffic is UDP-like.
-		Although Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded), SPF significantly underperforms for this relatively simple topology.
-		Non-SPF agents start off reasonably well, slowly learning better policies.
-		\label{fig:udp-2}
+		Online performance  when benign traffic is UDP-like.
+		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
+		SPF, remarkably, displays performance on par with Marl++ for this more difficult topology in spite of performing poorly when $n=2$.
+		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
+		The longer tails of outliers typically indicate the longer training time the new models require.
+		\label{fig:udp-16}
 	}
 \end{figure}
+
+Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
+\Cref{fig:udp-2,fig:udp-16} show the behaviour for $n \in \{2, 16\}$ in detail---the remaining choices of $n$ interpolate between these extremes, with the performance difference between Marl and Marl++ shrinking as $n$ increases.
+
+\subsection{Congestion-aware traffic}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.9\linewidth]{../plots/tcp-2}
+%	
+%	\caption{
+%		Online performance for $n=2$ hosts per egress point when benign traffic is TCP-like.
+%		Marl++ and Marl achieve very similar performance, starting off similarly well without notable improvement over an episode.
+%		SPF's performance is disappointingly close to baseline, indicating that it is as useful as having no defence system.
+%		\label{fig:tcp-2}
+%	}
+%\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.9\linewidth]{../plots/tcp-16}
+%	
+%	\caption{
+%		Online performance for $n=16$ hosts per egress point when benign traffic is UDP-like.
+%		Although Marl++ takes longer to reach its most effective policy, we achieve visibly better peak performance.
+%		SPF does manage to learn an improved policy compared to its starting state, but never manages to perform well.
+%		\label{fig:tcp-16}
+%	}
+%\end{figure}
 \begin{figure}
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/udp-16}
+	\includegraphics[width=0.9\linewidth]{../plots/tcp-box}
 
 	\caption{
-		Online performance for $n=16$ hosts per egress point when benign traffic is UDP-like.
-		Marl++ remains marginally ahead of its predecessor, though both have undergone a significant drop in effectiveness.
-		SPF, remarkably, displays performance on par with Marl++ for this more difficult topology.
-		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
-		\label{fig:udp-16}
-	}
-\end{figure}
-
-Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
-\Cref{fig:udp-2,fig:udp-16} show the behaviour for $n \in \{2, 16\}$ in detail---the remaining choices of $n$ interpolate between these extremes, with the performance difference between Marl and Marl++ shrinking as $n$ increases.
-
-\subsection{Congestion-aware traffic}
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/tcp-2}
-
-	\caption{
-		Online performance for $n=2$ hosts per egress point when benign traffic is TCP-like.
-		Marl++ and Marl achieve very similar performance, starting off similarly well without notable improvement over an episode.
-		SPF's performance is disappointingly close to baseline, indicating that it is as useful as having no defence system.
-		\label{fig:tcp-2}
-	}
-\end{figure}
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/tcp-16}
-
-	\caption{
-		Online performance for $n=16$ hosts per egress point when benign traffic is UDP-like.
+		Online performance for when benign traffic is TCP-like.
+		Marl and Marl++ begin with essentially identical efficacy.
 		Although Marl++ takes longer to reach its most effective policy, we achieve visibly better peak performance.
@@
 SPF performs consistently poorly across all choices of $n$.
-\Cref{fig:tcp-2}{fig:tcp-16} show two endpoints of a trend in our results: for larger $n$, Marl++ requires around twice as long as Marl to learn its best-achieving policy, underperforming for this period, but achieves notably better peak performance.
+\Cref{fig:tcp-2,fig:tcp-16} show two endpoints of a trend in our results: for larger $n$, Marl++ requires around twice as long as Marl to learn its best-achieving policy, underperforming for this period, but achieves notably better peak performance.
 
@@
 \fakepara{Model performance}
-Of the results presented, SPF's meandering performance is interesting given its far smaller action space.
-It's natural to expect that this would the model easier to learn, but the additional state required appears to make the model \emph{harder}, beyond even the gains we might expect from choosing a non-zero discount factor.
+Of the results presented, SPF's unpredictable (often worse) performance is unexpected, given its far smaller action space.
+It's natural to expect that this would make the model easier to learn, but the additional state required appears to make the task \emph{harder}, beyond even the value of choosing a non-zero discount factor (adding forward-planning to explicitly mitigate this effect).
 Accordingly, we see that SPF performs best when agents learn from as much knowledge as possible: high $n$ and single-agent training.
@@
 Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
+The benefits we have at present are, however, substantial.
+What we offer above the alternate approaches we discussed in \cref{sec:related-work} are potentially more flexible deployments and the ability to adapt by design to changing network traffic characteristics, without requiring the network resources or capabilities that the most effective techniques rely upon.
 
@@
 
-\section{Related Work}
+\fakepara{Real-world Deployment}
+Currently, we assume that switches support an extension to OpenFlow to enable remotely installable packet-drop rules, either by running a modified version of OVS on commodity hardware at these locations or through custom firmware for one of these particular egress switch.
+Similar functionality could be employed by making use of OpenFlow's meter rules.
+As far as footprints are concerned, although the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, in practice our sparse representation typically leads to far smaller policies---\SIrange{100}{120}{\kibi\byte} from our experiments.
+Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}, with the necessary statement.)
+%?? Incorrect, pickling...
+%?? deployment/system guidelines, architecture, and overhead/footprint of doing this
+%?? Mention: probably fewer state measurements than in our emulations, but longer measurements means less noisy (so probably more accurate)
+A real world deployment would be unable to gather and transmit load/flow statistics quite as often as we can in an emulated environment, without inadvertently affecting client traffic.
+However, the measurements acquired in such a scenario are far more likely to be less noisy (by drawing state over longer periods of time), and so could greatly aid effective training.
+We do not forsee this overly harming training.
+%?? Potential for dividing up pre-feature training across different agents, or train like this locally (sets of flows trained by 1 sub-model).
+
+%?? Different global state per-agent? Seems this must be locally trained, just make it (everything up the chain to key destinations).
+For non-tree-structured topologies, we each a
+
+\section{Related Work}\label{sec:related-work}
 
@@
 Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.
+Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works don't apply the rich literature on the subject to its fullest potential.
 
@@
 A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
-We feel there is a shortage of work in this category at present, at least in the field of networks; whether this is due to disinterest or difficulty is as-yet unknown.
+We feel there is a shortage of work in this category at present, at least in the field of networks.
 To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.
@@
 A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
-The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
+The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
 Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
@@
 Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
-To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
+To overcome the high latency of action computation via a neural network, two agents are trained, handling aspects of short and long flows respectively.
 The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
@@
 
-\section{Future Work}
+\section{Conclusions and Future Work}
+
+Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
+Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.
+
+We have identified weaknesses in past approaches arising from simulation assumptions, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
+Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
+Our evaluation shows that our new agent designs advance the state-of-the-art in RL-based DDoS prevention.
+
+While this work still trails behind the performance of exact DDoS flow detection mechanisms, we hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
+Yet even now, we offer advantages above these works, such as online adaptation and potentially more flexible deployment, without requiring many of the same network resources or capabilities that they rely upon.
+It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
+%(rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?
+
+%\section{Future Work}
 %Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.
@@
 If true online learning is desired (i.e., coping with a non-stationary environment), then such reward functions are sorely needed.
-While $\bload{\cdot}$ is likely to be a good candidate for many deployments, we believe that internal ``canary flows'' (or other active measurement) will be effective.
+While $\bload{\cdot}$ is likely to be a good candidate for many deployments, we believe that internal ``canary flows'' (or some other active measurement) will be effective sources of data to base a new reward function upon.
 
@@
 These scenarios present ideal circumstances to apply work on adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}, changepoint detection, or intelligent/heuristic sampling methods to judge which flows are most worthy of consideration per tick.
+For estimating \emph{when} to explore, we believe that the intersection of signal processing and RL is as-yet unexplored.
 
@@
 %?? More future work --- share knowledge between agents. ``Knowledge bases'' for this purpose? (see: Qianru).
-Effective real-world deployment of RL-based defences cannot assume that switches in a network will support a custom version of OVS or other arbitrary software, introducing the question of whether agent training, execution and distribution may when using \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
+Effective real-world deployment of RL-based defences cannot assume that switches in a network will support a custom version of OVS or other arbitrary software, introducing the question of whether agent training, execution and distribution may be possible when using \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
 We also expect it will be fruitful to look into \emph{how} agents may share knowledge with one another, such as via \emph{knowledge bases} or some other efficient communication mechanism.
@@
 
-\section{Conclusion}
-
-Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
-Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.
-
-We have identified weaknesses in past approaches arising from simulation assumptions, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
-Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
-Our evaluation shows that our new agent designs advance the state-of-the-art in RL-based DDoS prevention.
-
-We stress once again that this work still trails behind the performance of exact DDoS flow detection mechanisms, yet we hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
-It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
-%(rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?
-
 \section*{Acknowledgements}

### 2018-11-12 tables/big-avg-reward.py

@@ -11,11 +11,11 @@ variants = [
 	("uncap", None, "Uncapped"),
 	("banded", None, "Banded"),
 	("single", None, "Single"),
-	("channel", "../results/tcp-combo-{2}-{3}.csv", "Pretrain"),
+	("channel", "../results/{1}-combo-{2}-{3}.csv", "Pretrain"),
 ]
 
 illegal_combos = set()
-illegal_combos.add(("m", "udp", "channel"))
+# illegal_combos.add(("m", "udp", "channel"))
 illegal_combos.add(("spf", "udp", "channel"))
 illegal_combos.add(("spf", "tcp", "channel"))
 

### 2018-11-12 paper/paper.tex
Cleanup...

@@
 Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
-\Cref{fig:udp-2,fig:udp-16} show the behaviour for $n \in \{2, 16\}$ in detail---the remaining choices of $n$ interpolate between these extremes, with the performance difference between Marl and Marl++ shrinking as $n$ increases.
+\Cref{fig:udp-16} shows the behaviour for all choices of $n$ at a high level---specifically, we see that Marl++ typically remains ahead of Marl, though this performance difference shrinks as $n$ increases.
 
@@
 SPF performs consistently poorly across all choices of $n$.
-\Cref{fig:tcp-2,fig:tcp-16} show two endpoints of a trend in our results: for larger $n$, Marl++ requires around twice as long as Marl to learn its best-achieving policy, underperforming for this period, but achieves notably better peak performance.
+\Cref{fig:tcp-16} shows a trend in our results: for larger $n$, Marl++ requires around twice as long as Marl to learn its best-achieving policy, underperforming for this period, but achieves notably better peak and average performance.
 
@@
 \fakepara{Real-world Deployment}
-Currently, we assume that switches support an extension to OpenFlow to enable remotely installable packet-drop rules, either by running a modified version of OVS on commodity hardware at these locations or through custom firmware for one of these particular egress switch.
+Currently, we assume that switches support an extension to OpenFlow to enable remotely installable packet-drop rules, either by running a modified version of OVS on commodity hardware at these locations or through custom firmware for egress switches.
 Similar functionality could be employed by making use of OpenFlow's meter rules.
-As far as footprints are concerned, although the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, in practice our sparse representation typically leads to far smaller policies---\SIrange{100}{120}{\kibi\byte} from our experiments.
-Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}, with the necessary statement.)
+
+Where overheads are concerned, the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies---\SIrange{100}{120}{\kibi\byte} from our experiments.
+Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
+We believe that it may be possible to implement similar behaviour through application of programmable data planes.
+
 %?? Incorrect, pickling...
@@
 %?? Mention: probably fewer state measurements than in our emulations, but longer measurements means less noisy (so probably more accurate)
-A real world deployment would be unable to gather and transmit load/flow statistics quite as often as we can in an emulated environment, without inadvertently affecting client traffic.
-However, the measurements acquired in such a scenario are far more likely to be less noisy (by drawing state over longer periods of time), and so could greatly aid effective training.
-We do not forsee this overly harming training.
+Gathering and transmission of load/flow statistics would be difficult to perform quite as often as we can in an emulated environment, without inadvertently affecting host traffic.
+However, the measurements acquired in such a scenario are far more likely to be less noisy (by being drawn over longer periods of time), which could aid effective training.
+The main bottlenecks are most likely in forwarding the load measurements from various aggregation points (which can be made more efficient by way of multicast) and in running some estimator to condition the reward function upon.
 %?? Potential for dividing up pre-feature training across different agents, or train like this locally (sets of flows trained by 1 sub-model).
-
 %?? Different global state per-agent? Seems this must be locally trained, just make it (everything up the chain to key destinations).
-For non-tree-structured topologies, we each a
+We expect that agents will be able to share policies for all fields except global state, which may help to offset the reduced rate of incoming experience.
+Regardless, it will take longer to achieve enough experience to converge on a good policy.
 

### 2018-11-12 code/marl/channel-run.sh
Final pretrainer expts...

@@ -1,7 +1,6 @@
 #!/bin/bash
 
-tests=(0 3)
-#tests=(1 2)
+tests=(0 1 2 3)
 
 for v in "${tests[@]}"
 do
@@ -11,5 +10,11 @@ do
 	python ./tcp-model-combo-channel-prep.py $v "prog"
 	./purger.sh
 	python ./tcp-model-combo-channel-end.py $v
+	./purger.sh
+	python ./udp-model-combo-channel-prep.py $v
+	./purger.sh
+	python ./udp-model-combo-channel-prep.py $v "prog"
+	./purger.sh
+	python ./udp-model-combo-channel-end.py $v
 	./purger.sh
 done

### 2018-11-12 code/marl/spf-channel-run.sh

@@ -1,0 +1,20 @@
+#!/bin/bash
+
+tests=(0 1 2 3)
+
+for v in "${tests[@]}"
+do
+	./purger.sh
+	python ./tcp-spf-combo-channel-prep.py $v
+	./purger.sh
+	python ./tcp-spf-combo-channel-prep.py $v "prog"
+	./purger.sh
+	python ./tcp-spf-combo-channel-end.py $v
+	./purger.sh
+	python ./udp-spf-combo-channel-prep.py $v
+	./purger.sh
+	python ./udp-spf-combo-channel-prep.py $v "prog"
+	./purger.sh
+	python ./udp-spf-combo-channel-end.py $v
+	./purger.sh
+done

### 2018-11-12 code/marl/statify-pickles.py

@@ -1,0 +1,87 @@
+from collections import Counter
+import cPickle as pickle
+import numpy as np
+
+results_dir = "../../results/"
+pickle_chooser = "tcp-combo-channel-prep-{}.pickle"
+out_file = "udp-action-f{}-{}.csv"
+
+restricts = [5, 6, 7]
+txrestricts = [a + 4 for a in restricts]
+windows = [
+	(0, 10000),
+	(-50, 50),
+	(-50, 50),
+]
+n_steps = 48
+
+ns = [2, 4, 8, 16]
+
+factor = 1/10.0
+
+for n in ns:
+	len_holder = {}
+	for i in txrestricts:
+		len_holder[(i,)] = []
+	len_holder[(0,)] = []
+
+	choice_holders = [{} for i in windows]
+		
+	f_dir = results_dir + pickle_chooser.format(n)
+	with open(f_dir, "rb") as of:
+		q = pickle.load(of)
+
+	for (_rng, ep) in q:
+		for (a_tree, restrict) in ep:
+			for team in a_tree:
+				for agent in team:
+					#print "n:{} -- {}".format(n, len(agent.values))
+					place = (0,) if restrict is None else tuple(restrict)
+					len_holder[place].append(len(agent.values))
+					if restrict is not None:
+						truth = restrict[0] - 4
+						ind = restricts.index(truth)
+						(lower, upper) = windows[ind]
+						start = float(lower)
+						move = (float(upper) - start)/float(n_steps)
+						space = choice_holders[ind]
+
+						curr = start - move
+						end = upper + move
+
+						#print curr, end, move
+						while curr < end:
+							if curr not in space:
+								space[curr] = ([], [])
+							s = agent.tc(np.array([curr]))
+							vals = agent.select_action(s)
+							space[curr][0].append(vals[0])
+							space[curr][1].append(vals[2])
+							#print curr, vals[2]
+							curr += move
+						
+	for (k, v) in len_holder.iteritems():
+		print "For n={} and feature {}, avg occ={}".format(n, k, np.mean(v))
+
+	for i, space in enumerate(choice_holders):
+		x_vals = []
+		mosts = []
+		means = []
+		avs = []
+		for (p_set, (action_choices, a_vals)) in space.iteritems():
+			c = Counter(action_choices)
+			x_vals.append(p_set)
+			q = c.most_common(4)
+			print q
+			mosts.append(q[0][0])
[…]

### 2018-11-12 code/marl/tcp-model-combo-channel-end.py

@@ -27,6 +27,7 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, contributo
 		rf = "ctl",
 		use_controller = True,
 		actions_target_flows = restrict is not None,
+		trs_maxtime = 0.001,
 
 		restrict = restrict,
 

### 2018-11-12 code/marl/tcp-spf-combo-channel-end.py

@@ -3,7 +3,7 @@ from marl import *
 import sys
 from writer import writeResults, makeResultsAverage
 
-ft = __import__("udp-model-combo-channel-prep")
+ft = __import__("tcp-spf-combo-channel-prep")
 
 def run(restrict, state, rewards, good_traffic_percents, total_loads, contributors=None, store_sarsas=[], ep_len=10000, h_p=2):
 	results = marlExperiment(
@@ -20,7 +20,7 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, contributo
 
 		alpha = 0.05,
 		epsilon = 0.2,
-		discount = 0,
+		discount = 0.8,
 
 		dt = 0.01,
 
@@ -39,15 +39,16 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, contributo
 		contributors = contributors,
 		store_sarsas = store_sarsas,
 
-		#evil_range = [4,7],
+		evil_range = [4,7],
 		model = "nginx",
-		submodel = "udp-flood",
-		#reward_direction = "out",
-		#randomise = True,
-		#randomise_count = 3,
-		#randomise_new_ip = True,
+		#submodel = "udp-flood",
+		reward_direction = "out",
+		randomise = True,
+		randomise_count = 3,
+		randomise_new_ip = True,
 
 		estimate_const_limit = True,
+		spiffy_mode = True,
 	)
 
 	return results
@@ -91,8 +92,8 @@ if __name__ == "__main__":
 	# Now, write out the results!
 	for (target_results, start_fresh, config_name, ep_len) in configs:
 		(rs, gs, ls) = target_results
-		csv_dir = ft.results_dir + "udp-combo-channel-{}.csv".format(hosts_p)
-		avg_csv_dir = ft.results_dir + "udp-combo-channel-{}-avg.csv".format(hosts_p)
+		csv_dir = ft.results_dir + "tcp-spf-combo-channel-{}.csv".format(hosts_p)
+		avg_csv_dir = ft.results_dir + "tcp-spf-combo-channel-{}-avg.csv".format(hosts_p)
 
 		results = (rs, gs, ls, None, None, None)
 

### 2018-11-12 code/marl/tcp-spf-combo-channel-prep.py

@@ -19,7 +19,7 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
 
 		alpha = 0.05,
 		epsilon = 0.2,
-		discount = 0,
+		discount = 0.8,
 
 
 		dt = 0.01,
@@ -40,12 +40,15 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
 
 		evil_range = [4,7],
 		model = "nginx",
+		#submodel = "udp-flood",
 		reward_direction = "out",
 		randomise = True,
 		randomise_count = 3,
+t
 		randomise_new_ip = True,
 
 		estimate_const_limit = True,
+		spiffy_mode = True,
 	)
 
 	return results
@@ -55,8 +58,8 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
 
 host_ps = [2, 4, 8, 16]
 results_dir = "../../results/"
-file_dir = results_dir + "tcp-combo-channel-prep-{}.pickle"
-in_progress_file_dir = results_dir + "tcp-combo-channel-prg-{}.pickle"
+file_dir = results_dir + "tcp-spf-combo-channel-prep-{}.pickle"
+in_progress_file_dir = results_dir + "tcp-spf-combo-channel-prg-{}.pickle"
 chosen_features = [
 	5, # Mean IAT
 	7, # Delta Out Rate
@@ -121,8 +124,8 @@ if __name__ == "__main__":
 
 	# Now, write out the results!
 	for ((rs, gs, ls), out_name) in zip(result_sets, out_names):
-		csv_dir = results_dir + "tcp-combo-channel-prep-{}-{}.csv".format(out_name, hosts_p)
-		avg_csv_dir = results_dir + "tcp-combo-channel-prep-{}-{}-avg.csv".format(out_name, hosts_p)
+		csv_dir = results_dir + "tcp-spf-combo-channel-prep-{}-{}.csv".format(out_name, hosts_p)
+		avg_csv_dir = results_dir + "tcp-spf-combo-channel-prep-{}-{}-avg.csv".format(out_name, hosts_p)
 
 		results = (rs, gs, ls, None, None, None)
 

### 2018-11-12 code/marl/udp-model-combo-channel-prep.py

@@ -29,6 +29,7 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
 		rf = "ctl",
 		use_controller = True,
 		actions_target_flows = restrict is not None,
+		trs_maxtime = 0.001,
 
 		restrict = restrict,
 
@@ -79,12 +80,12 @@ if __name__ == "__main__":
 	# pickle has last experiment state...
 	start_i = 0
 	in_progress = len(sys.argv) > 2
+	things_to_pickle = []
 
 	if in_progress:
 		with open(in_progress_file_dir, "rb") as of:
 			(start_i, result_sets, things_to_pickle) = cPickle.load(of)
 
-	things_to_pickle = []
 
 	randstate = None
 	for i in xrange(start_i, start_i + n_episodes_per_step):

### 2018-11-12 code/marl/udp-spf-combo-channel-end.py

@@ -3,7 +3,7 @@ from marl import *
 import sys
 from writer import writeResults, makeResultsAverage
 
-ft = __import__("udp-model-combo-channel-prep")
+ft = __import__("udp-spf-combo-channel-prep")
 
 def run(restrict, state, rewards, good_traffic_percents, total_loads, contributors=None, store_sarsas=[], ep_len=10000, h_p=2):
 	results = marlExperiment(
@@ -20,7 +20,7 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, contributo
 
 		alpha = 0.05,
 		epsilon = 0.2,
-		discount = 0,
+		discount = 0.8,
 
 		dt = 0.01,
 
@@ -48,6 +48,7 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, contributo
 		#randomise_new_ip = True,
 
 		estimate_const_limit = True,
+		spiffy_mode = True,
 	)
 
 	return results
@@ -91,8 +92,8 @@ if __name__ == "__main__":
 	# Now, write out the results!
 	for (target_results, start_fresh, config_name, ep_len) in configs:
 		(rs, gs, ls) = target_results
-		csv_dir = ft.results_dir + "udp-combo-channel-{}.csv".format(hosts_p)
-		avg_csv_dir = ft.results_dir + "udp-combo-channel-{}-avg.csv".format(hosts_p)
+		csv_dir = ft.results_dir + "udp-spf-combo-channel-{}.csv".format(hosts_p)
+		avg_csv_dir = ft.results_dir + "udp-spf-combo-channel-{}-avg.csv".format(hosts_p)
 
 		results = (rs, gs, ls, None, None, None)
 

### 2018-11-12 code/marl/udp-spf-combo-channel-prep.py

@@ -19,7 +19,7 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
 
 		alpha = 0.05,
 		epsilon = 0.2,
-		discount = 0,
+		discount = 0.8,
 
 
 		dt = 0.01,
@@ -38,14 +38,17 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
 		good_traffic_percents = good_traffic_percents,
 		total_loads = total_loads,
 
-		evil_range = [4,7],
+		#evil_range = [4,7],
 		model = "nginx",
-		reward_direction = "out",
-		randomise = True,
-		randomise_count = 3,
-		randomise_new_ip = True,
+		submodel = "udp-flood",
+		#reward_direction = "out",
+		#randomise = True,
+		#randomise_count = 3,
+t
+		#randomise_new_ip = True,
 
 		estimate_const_limit = True,
+		spiffy_mode = True,
 	)
 
 	return results
@@ -55,8 +58,8 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
 
 host_ps = [2, 4, 8, 16]
 results_dir = "../../results/"
-file_dir = results_dir + "tcp-combo-channel-prep-{}.pickle"
-in_progress_file_dir = results_dir + "tcp-combo-channel-prg-{}.pickle"
+file_dir = results_dir + "udp-spf-combo-channel-prep-{}.pickle"
+in_progress_file_dir = results_dir + "udp-spf-combo-channel-prg-{}.pickle"
 chosen_features = [
 	5, # Mean IAT
 	7, # Delta Out Rate
@@ -121,8 +124,8 @@ if __name__ == "__main__":
 
 	# Now, write out the results!
 	for ((rs, gs, ls), out_name) in zip(result_sets, out_names):
-		csv_dir = results_dir + "tcp-combo-channel-prep-{}-{}.csv".format(out_name, hosts_p)
-		avg_csv_dir = results_dir + "tcp-combo-channel-prep-{}-{}-avg.csv".format(out_name, hosts_p)
+		csv_dir = results_dir + "udp-spf-combo-channel-prep-{}-{}.csv".format(out_name, hosts_p)
+		avg_csv_dir = results_dir + "udp-spf-combo-channel-prep-{}-{}-avg.csv".format(out_name, hosts_p)
 
 		results = (rs, gs, ls, None, None, None)
 

### 2018-11-12 code/marl/tcp-spf-combo-channel-prep.py
Fixes...

@@ -44,7 +44,7 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
 		reward_direction = "out",
 		randomise = True,
 		randomise_count = 3,
-t
+
 		randomise_new_ip = True,
 
 		estimate_const_limit = True,

### 2018-11-12 code/marl/udp-spf-combo-channel-prep.py

@@ -44,7 +44,7 @@ def run(restrict, state, rewards, good_traffic_percents, total_loads, h_p):
 		#reward_direction = "out",
 		#randomise = True,
 		#randomise_count = 3,
-t
+
 		#randomise_new_ip = True,
 
 		estimate_const_limit = True,

### 2018-11-12 code/marl/marl.py
Fixed SPF with global state on its own

@@ -747,7 +747,8 @@ def marlExperiment(
 				if len(state) == 0:
 					action = default_machine_state
 				else:
-					(_, action) = state
+					((_svec, action), machine) = state
+					action = machine.action()
 				a = action if override_action is None else override_action
 				tx_ac = sarsa.actions[a] if isinstance(a, (int, long)) else a
 				updateUpstreamRoute(node, ac_prob=tx_ac)
@@ -1708,7 +1709,10 @@ def marlExperiment(
 					else:
 						prev_state = learner_traces[l_index]
 						if prev_state == {}:
-							prev_state = sarsa.last_act
+							machine = AcTrans()
+							prev_state = (sarsa.last_act, machine)
+
+						(last_act, machine) = prev_state
 
 						# Start time of action computation
 						s_t = time.time()
@@ -1717,8 +1721,9 @@ def marlExperiment(
 						state = sarsa.to_state(np.array(state_vec))
 
 						# Learn!
-						target_sarsa.update(state, reward, prev_state)
-						learner_traces[l_index] = target_sarsa.last_act
+						target_sarsa.update(state, reward, last_act)
+						machine.move(sarsa.last_act[1])
+						learner_traces[l_index] = (sarsa.last_act, machine)
 
 						# End time.
 						e_t = time.time()

### 2018-11-12 paper/paper.tex
Some fixes according to Simon

@@
 	pdftitle={\mytitle{}},
-%	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros}
-	pdfauthor={Anonymous Giraffe and Badger}
+%	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros, Simon D. Rogers}
+	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
 }
@@
 %	Kyle A.\ Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
-%	Dimitrios P.\ Pezaros\\
+%	Kyle A.\ Simpson,
+%	Dimitrios P.\ Pezaros,
+%   Simon D.\ Rogers\\
 %	University of Glasgow, Glasgow, Scotland,\\
@@
 Anonymous Giraffe,
-Anonymous Badger\\
+Anonymous Badger,
+Anonymous Ibex\\
 Unnamed Department, Nowhere\\
@@
 
-While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \citeauthor{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems---to quite a large extent, this remains the case today \cite{DBLP:conf/sp/SommerP10}.
+While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \citeauthor{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems, and to quite a large extent this remains the case today \cite{DBLP:conf/sp/SommerP10}.
 Their application is made difficult due to significant operational differences from standard ML tasks, alongside certain characteristics of network traffic.
@@
 We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling for deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
-	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
+	\item A reactive simulation of web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
 	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
@@
 If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
-Action values for any state are then obtained by summation of the weight vectors from all activated tiles.
+Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}})$ floating point additions per decision.
 Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
-This means that we need only perform $n_{\mathit{tilings}}$ floating point additions per model update:
+This means that we need only perform $n_{\mathit{tilings}} + 2$ floating point additions and \num{2} multiplications per model update:
 \begin{equation}
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=0.8\linewidth]{../plots/ftprep-good-2}
@@
 The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
+All features converge to their highest-observed performance within around \num{4000} timesteps: \cref{fig:udp-fp-converge} shows an example of this behaviour for differently performing features.
 In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
@@
 Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect, or hope for.
-We suspect that some of these difficulties originate from the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to unpick the true value of each action), and controlling shared global state.
+%??
+I don’t know if you want to also mention the problem of things getting “stuck” in locally optimal (but globally sub-optimal) policies.
+I think it’s fairly uncontroversial that an increase in state space would make this more likely.
+Also, the lack of discounting could play a large role (and more so as the state space becomes more complex…
+The most likely explanation is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
+The increased state-space size makes this a more likely occurrence, as is the unclear affect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state-space.
+We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to unpick the true value of each action), and controlling shared global state.
 It may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows, to explore this hypothesis.
@@
 The benefits we have at present are, however, substantial.
-What we offer above the alternate approaches we discussed in \cref{sec:related-work} are potentially more flexible deployments and the ability to adapt by design to changing network traffic characteristics, without requiring the network resources or capabilities that the most effective techniques rely upon.
+What we offer above the alternate approaches we discussed in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
 
@@
 Where overheads are concerned, the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies---\SIrange{100}{120}{\kibi\byte} from our experiments.
+As we have described earlier, action updates require a constant number of floating point operations---\num{122} floating point additions and \num{2} multiplications per update of $\wvec{}$, above the \num{140} additions required to choose an action.
+The vast majority of these operations can be vectorised trivially, if such hardware is present.
+Action computation for SPF is cheaper still, requiring only \num{42} additions per action required.
 Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
@@
 
-While this work still trails behind the performance of exact DDoS flow detection mechanisms, we hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
-Yet even now, we offer advantages above these works, such as online adaptation and potentially more flexible deployment, without requiring many of the same network resources or capabilities that they rely upon.
-It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
 %(rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?
@@
 %?? IDEA: try out average reward, TD($\lambda$) methods as future work...
+The extant weaknesses invite many further improvements worth investigation.
 A natural research direction to enhance this work would be the combination of the classic function approximation techniques we make use of alongside the improved algorithms that the RL community has introduced in the past few years.
@@
 
+%While this work still trails behind the performance of exact DDoS flow detection mechanisms, w
+We hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
+We still offer advantages above existing works, such as potentially more flexible deployment and automatically learned low-overhead decision-making, without requiring many of the same network resources or capabilities that they rely upon.
+It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
+
 \section*{Acknowledgements}
-%Our thanks go to Simon Rogers, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Richard Cziva for their advice, comments and technical assistance.
-Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
+%Our thanks go to Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Richard Cziva for their advice, comments and technical assistance.
+Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
 %Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.

### 2018-11-13 code/marl/statify-pickles.py
Whoopsie

@@ -106,7 +106,7 @@ for n in ns:
 				writ = csv.writer(of)
 				rest_header = [str(i) for i in xrange(data.shape[1])] if is_matrix else ["action"]
 				writ.writerow(["#x"] + rest_header)
-				for row in data:
+				for j, row in enumerate(data):
 					if not is_matrix:
 						row = [row]
-					writ.writerow(row)
+					writ.writerow([nx[j]] + list(row))

### 2018-11-13 code/marl/big-final-results.py
I live for data.

@@ -12,7 +12,7 @@ models = [
 ]
 
 # host_p, dt
-host_ps = [(2, 0.05), (4, 0.05), (8, 0.05), (16, 0.01)]
+host_ps = [(2, 0.05), (4, 0.05), (8, 0.05), (16, 0.05)]
 
 # prefix, tcp?
 traffic_types = [

### 2018-11-13 paper/paper.tex
Some more writing, sick new results. Less apologetic.

@@
 \DeclareFieldFormat[article]{url}{}
+\DeclareFieldFormat[inproceedings]{doi}{}
+\DeclareFieldFormat[article]{doi}{}
+\DeclareFieldFormat[inproceedings]{editor}{}
+\DeclareFieldFormat[proceedings]{editor}{}
+\DeclareFieldFormat[article]{editor}{}
+
+% Thanks to http://codydunne.blogspot.com/2012/01/suppressing-bibtex-fields-for-specific.html
+\AtEveryBibitem{% Clean up the bibtex rather than editing it
+	\clearlist{address}
+%	\clearfield{date}
+%	\clearfield{eprint}
+	\clearfield{isbn}
+	\clearfield{issn}
+	\clearlist{location}
+%	\clearfield{month}
+	\clearfield{series}
+
+	\ifentrytype{book}{}{% Remove publisher and editor except for books
+		\clearlist{publisher}
+		\clearname{editor}
+	}
+}
 
@@
 	pdftitle={\mytitle{}},
-%	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros, Simon D. Rogers}
+%	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros, Simon Rogers}
 	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
@@
 %	Dimitrios P.\ Pezaros,
-%   Simon D.\ Rogers\\
+%   Simon Rogers\\
 %	University of Glasgow, Glasgow, Scotland,\\
@@
 These flaws---poor handling of TCP traffic and high host density at egress points---inspire and direct an exploration of the design space of RL models.
-We advance the state-of-the-art of DDoS mitigation using RL by introducing two agent classes designed to act on a per-flow basis, supported by an in-depth investigation of feature suitability and empirical evaluation.
+We advance the state-of-the-art of DDoS mitigation using RL by introducing two agent classes designed to act on a per-flow basis, in a protocol-agnostic manner.
+This is supported by an in-depth investigation of feature suitability and empirical evaluation.
 
@@
 ?? Expand as required
+
+?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
+?? THAT IS A MAJOR CONTRIB, MENTION IT EVERYWHERE YOU CAN
 
@@
 If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
-Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}})$ floating point additions per decision.
+Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}-1})$ floating point additions per decision.
 Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
-This means that we need only perform $n_{\mathit{tilings}} + 2$ floating point additions and \num{2} multiplications per model update:
+This means that we need only perform $n_{\mathit{tilings}} + 2$ additions and \num{2} multiplications per model update:
 \begin{equation}
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{../plots/ftprep-good-2}
+	\includegraphics[width=0.9\linewidth]{../plots/ftprep-good-2}
 	\caption{
@@
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
-Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, and the case where one agent makes all decisions.%, and pre-training on individual features.
+Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
 
@@
 	\caption{
-		Online performance  when benign traffic is UDP-like.
+		Online performance for UDP-like benign traffic.
 		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
-		SPF, remarkably, displays performance on par with Marl++ for this more difficult topology in spite of performing poorly when $n=2$.
+		SPF, remarkably, displays performance on par with Marl++ for this more difficult topology in spite of underperforming when $n=2$.
 		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
-		The longer tails of outliers typically indicate the longer training time the new models require.
-		\label{fig:udp-16}
+		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
+		\label{fig:udp-box}
 	}
@@
 Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
-\Cref{fig:udp-16} shows the behaviour for all choices of $n$ at a high level---specifically, we see that Marl++ typically remains ahead of Marl, though this performance difference shrinks as $n$ increases.
+\Cref{fig:udp-box} shows the behaviour for all choices of $n$---specifically, we see that Marl++ typically remains ahead of Marl, though this performance difference shrinks as $n$ increases, while SPF performs better at high host-density.
 
@@
 %\end{figure}
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.9\linewidth]{../plots/tcp-16}
-%	
-%	\caption{
-%		Online performance for $n=16$ hosts per egress point when benign traffic is UDP-like.
-%		Although Marl++ takes longer to reach its most effective policy, we achieve visibly better peak performance.
-%		SPF does manage to learn an improved policy compared to its starting state, but never manages to perform well.
-%		\label{fig:tcp-16}
-%	}
-%\end{figure}
 \begin{figure}
@@
 	\caption{
-		Online performance for when benign traffic is TCP-like.
-		Marl and Marl++ begin with essentially identical efficacy.
-		Although Marl++ takes longer to reach its most effective policy, we achieve visibly better peak performance.
-		SPF does manage to learn an improved policy compared to its starting state, but never manages to perform well.
+		Online performance for TCP-like benign traffic.
+		Marl and Marl++ begin with essentially identical efficacy, with Marl++ having a small advantage at larger $n$.
+		SPF is slightly worse than Marl and Marl++ at low $n$, but achieves considerably better median performance and lower variance than the other models for higher $n$.
+		\label{fig:tcp-box}
+	}
+\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{../plots/tcp-16-single}
+
+	\caption{
+		Online performance for $n=16$ hosts per egress point when benign traffic is TCP-like.
+		At this level of host density, SPF reaches its peak performance sooner and is considerably more consistent throughout.
+		Marl++ takes longer to reach its most effective policy, we achieve visibly better peak performance.
+		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
 		\label{fig:tcp-16}
@@
 
-Marl++ consistently allows more legitimate traffic to reach its destination than Marl, with both significantly above the baseline (\cref{tab:av-vals}, \emph{Capped}, TCP).
-SPF performs consistently poorly across all choices of $n$.
-\Cref{fig:tcp-16} shows a trend in our results: for larger $n$, Marl++ requires around twice as long as Marl to learn its best-achieving policy, underperforming for this period, but achieves notably better peak and average performance.
+From \cref{fig:tcp-box}, Marl++ consistently allows more legitimate traffic to reach its destination than Marl, with both significantly above the baseline (\cref{tab:av-vals}, \emph{Capped}, TCP).
+SPF performs slightly worse at low $n$, but outperforms both models as host density increases.
+\Cref{fig:tcp-16} shows a trend in our results: for larger $n$, Marl++ requires around twice as long as Marl to learn its best-achieving policy, under-performing for this period, but achieves notably better peak and average performance.
+SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
 
@@
 In general, we find that explicitly training a policy to keep the system in a valid state via negative reinforcement (i.e., an \emph{uncapped} network) is most effective for UDP-like traffic with Marl++.
-For SPF, however, negative reinforcement becomes a worse choice for larger $n$.
-Where TCP-like traffic is concerned, we see that this strategy almost always leads to worse policies.
+For SPF, however, negative reinforcement is typically a worse choice.
+Where TCP-like traffic is concerned in both models, we see that this strategy almost always leads to worse policies.
 
@@
 Making all decisions with a single agent is roughly equivalent to having a zero-cost communication channel between each pair of agents, theoretically allowing faster training by giving each agent more experience.
-Curiously, we observe that this often leads to drastically worse policies when used as part of Marl++ for small $n$, but makes SPF a considerably more competitive model---coming ahead of Marl for UDP traffic.
+Curiously, we observe that this often leads to drastically worse policies when used as part of Marl++ for small $n$, but makes SPF a considerably more competitive model---especially for TCP, and as $n$ grows larger (\cref{fig:tcp-16}).
+Single-agent SPF almost consistently outperforms Marl.
 We discuss our conjectures for why this reversal occurs in \cref{sec:discussion}.
 
-%\subsection{Per-feature training}
-%
-%?? Still being performed.
+\subsection{Per-feature pretraining}
+To investigate the effects of pre-training individual features, we chose our four best-performing features (global state, mean IAT, $\Delta$ in/out rates) and trained each over \num{10000} iterations, before evaluating the policy obtained by summing the action values suggested by each.
+This may be achieved in a real network by randomly assigning new hosts to be controlled exclusively by one feature's policy, allowing agents to more easily learn \emph{which} information led to the correct action being taken.
+We reduce the timestep length to \SI{10}{\milli\second} to train each agent in less wall-clock time, at the cost of noisier data (and reduced learning rate).
+The results show no real change for Marl++, but a considerable gain for SPF against UDP (and performance drop against TCP).
+It is likely that, as did the different protocols in \cref{fig:udp-feature-plots,fig:tcp-cap-feature-plots}, Marl++ and SPF have different optimal feature sets for different traffic behaviours---many of which have been excluded.
 
@@
 \fakepara{Model performance}
-Of the results presented, SPF's unpredictable (often worse) performance is unexpected, given its far smaller action space.
+Of the results presented, SPF's unpredictable (often worse) starting performance is unexpected, given its far smaller action space.
 It's natural to expect that this would make the model easier to learn, but the additional state required appears to make the task \emph{harder}, beyond even the value of choosing a non-zero discount factor (adding forward-planning to explicitly mitigate this effect).
 Accordingly, we see that SPF performs best when agents learn from as much knowledge as possible: high $n$ and single-agent training.
+Why does it exhibit considerably lower variance?
+Our belief is that SPF is, in fact, a considerably stronger model for these reasons, and its successes give us particular cause to consider the best schemes for efficient information sharing.
+To filter incoming traffic from a source, it must decide to degrade their inbound traffic multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
 Paradoxically, Marl++ generally achieves the best performance yet actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.
@@
 We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to unpick the true value of each action), and controlling shared global state.
-It may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows, to explore this hypothesis.
+Our pre-training results support the idea that it may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows.
 
@@
 %Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
-Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are an improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
+Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
 Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
@@
 What we offer above the alternate approaches we discussed in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
+Moreover, our decision making processes are entirely agnostic with regards to the protocol or content of inbound traffic, offering future-proofing against the introduction of new transports.
 
@@
 The natural solution is to impose an upper bound on the amount of action computations/policy updates that can be performed before a work list is discarded completely.
-This removes the guarantee that all flows will be visited fairly often, but if updates occur regularly then this random sampling my be sufficient to achieve good performance.
+This removes the guarantee that all flows will be visited fairly often, but if updates occur regularly then this random sampling may be sufficient to achieve good performance.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{../plots/policy-16-tcp-f5-mean-log}
+
+	\caption{
+		Mean policy for Mean IAT when $n=16$.
+		The column at any row is the action most likely to be chosen: this isn't actually a policy because it's mean'd.
+		Repeated values originate from the bias tile, which is always active.
+		?? Shows... what?
+		\label{fig:intern-16-tcp-iat}
+	}
+\end{figure}
 
@@
 With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
+We may see this in \cref{fig:intern-16-tcp-iat}... ?? ELABORATE
 Furthermore, system state which is dependent on many signals drawn from across a wide network (such as our global state) is difficult to exert precise control over.
@@
 Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
-We believe that it may be possible to implement similar behaviour through application of programmable data planes.
+We believe that it may be possible to implement similar behaviour on plain switches through application of \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
 
@@
 %?? Mention: probably fewer state measurements than in our emulations, but longer measurements means less noisy (so probably more accurate)
-Gathering and transmission of load/flow statistics would be difficult to perform quite as often as we can in an emulated environment, without inadvertently affecting host traffic.
-However, the measurements acquired in such a scenario are far more likely to be less noisy (by being drawn over longer periods of time), which could aid effective training.
+Gathering and transmission of load/flow statistics would be difficult to perform quite as often as an emulated environment allows, without inadvertently affecting host traffic.
+However, the measurements acquired in such a scenario are far more likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
 The main bottlenecks are most likely in forwarding the load measurements from various aggregation points (which can be made more efficient by way of multicast) and in running some estimator to condition the reward function upon.
@@
 We expect that agents will be able to share policies for all fields except global state, which may help to offset the reduced rate of incoming experience.
-Regardless, it will take longer to achieve enough experience to converge on a good policy.
+Regardless, it will take longer to achieve enough state-state transitions to converge on a good policy.
 
@@
 Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
-Our evaluation shows that our new agent designs advance the state-of-the-art in RL-based DDoS prevention.
+Our evaluation shows that our new agent designs advance the state-of-the-art in RL-based DDoS prevention, with SPF showing the most promise for future evaluation.
 
@@
 We hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
-We still offer advantages above existing works, such as potentially more flexible deployment and automatically learned low-overhead decision-making, without requiring many of the same network resources or capabilities that they rely upon.
+It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, potentially more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques rely upon.
 It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.

### 2018-11-14 paper/paper.tex

@@
 	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}.
-We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling for deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
+We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
 	\item A reactive simulation of web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
@@
 If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
-Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}-1})$ floating point additions per decision.
+Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}}-1)$ floating point additions per decision.
 Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
@@
 The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
-Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allows relocation and easy installation of learners.
+Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners.
 
@@
 
-At every time step, each agent installs an action via OpenFlow, instructing its host switch to drop each packet with probability $p$.
+At every timestep, each agent installs an action via OpenFlow, instructing its host switch to drop each packet with probability $p$.
 Specifically, the agent chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
-In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so exhibit a risk of introducing collateral damage into the network.
-These choices primarily ensure that this rate reduction imposed on a source IP may never be performance.
-It's worth noting that there are various ways that this could be done, and that the application of \emph{programmable data planes} to this end are suggested as future work.
+These choices primarily ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
+In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
+It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
 
@@
 These are important issues, particularly when we consider real-world deployment.
-Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most-used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
+Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
 Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network.
@@
 What this means is that datacentres and small-scale service deployments are more likely to have a host-to-learner ratio scaled heavily in favour of hosts, i.e. a large observed $n$.
-We feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions (as defined above) become less granular.
+We feel is most worth examining aspects of such a case, where actions (as defined above) become less granular.
 
 To test this, we configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
-This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single-server with a fan-out network.
+This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
 The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
@@
 Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
-We contribute a network model built around live testing of TCP traffic in an SDN-enabled environment, adaptable to arbitrary topologies, with an explicit focus on preserving the real-time dynamics of TCP traffic.
+We contribute a network model built around live testing of TCP traffic in an SDN-enabled environment, which is adaptable to arbitrary topologies, with an explicit focus on preserving the real-time dynamics of TCP traffic.
 In particular, we are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
@@
 \subsection{Network Design}
-We make use of a fully software-defined network, built using Openflow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
+We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
 All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
@@
 \end{itemize}
-We offer these to model two extremes: long-life host-to-server transfers (a characteristic we might e	xpect from e.g., streaming video), and common HTTP traffic.
+We offer these to model two extremes: long-life host-to-server transfers (a characteristic we might expect from e.g., streaming video), and common HTTP traffic.
 
@@
 Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
-Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that simulation remains tractable.
+Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that emulation remains tractable.
 When using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows---a \SI{30}{\percent} rate increase in expectation, uploading \SIrange{2.03}{2.18}{$\!
 \times U_s$}.
-The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$
+The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
 
@@
 For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
-Uncompleted work is passed on to the next timestep, until it is emptied, at which point is is repopulated using the set of available measurements.
+Uncompleted work is passed on to the next timestep, until the list is emptied, at which point is is repopulated using the set of available measurements.
 To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
 State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
-Following \textcite{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 8 flows in this time.
+Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 8 flows in this time.
 We expect this should be tuned based on the frequency at which statistics arrive.
@@
 Although this likely brings lower theoretical performance, there are more reasons than latency alone to favour classical methods; these include lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
-This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without the introduction of dedicated tensor processing hardware to networks.
+This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without necessarily introducing dedicated tensor processing hardware to networks.
 Simpler decision boundaries reduce the risk of overfitting and of unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.
@@
 This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
-Recounting \cref{sec:reinforcement-learning}, the choice of $\gamma$ affects which fraction of the next current state-action pair's value is added to $R_t$ when updating the last state-action pair's value.
+Recounting \cref{sec:reinforcement-learning}, the choice of $\gamma$ affects which fraction of the current state-action pair's value is added to $R_t$ when updating the value of the last state-action pair.
 We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing with UDP traffic, by observing average performance over $\gamma$ increments of $0.1$.
@@
 		Global state and Mean IAT still offer the greatest improvement above baseline, but it is observed that different optimal features are expressed between different traffic models.
-		Packet-level statistics are considerably less effective for this class of traffic, with $\Delta$ Out Rate and Correspondence Ratio being proportionally better than they are alongside UDP.
+		Packet-level statistics are considerably less effective for this class of traffic, with $\Delta$ Out Rate and Correspondence Ratio being proportionally better than they are for UDP.
 		\label{fig:tcp-feature-plots}
@@
 
-\fakepara{Mean packet size per-window}
+\fakepara{Mean packet size per window}
 Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
@@
 One way of solving this may be to investigate the effect of replacing $U_s$ with e.g., $0.99 U_s$ in \cref{eqn:reward-but-better}, to introduce a small band of negative reinforcement without actually removing the load restriction in the environment.
-It is crucial to make this region as small as possible, hence the somewhat arbitrary choice; we observe that agents are always able to learn how to keep traffic consistently below the logical bandwidth limit (as in the original work).
+It is crucial to make this region as small as possible, hence the somewhat arbitrary choice of factor; we observe that agents are always able to learn how to keep traffic consistently below the logical bandwidth limit (as in the original work).
 Our intent is to potentially accelerate learning while preserving the effects of system dynamics---we term these \emph{banded} rewards.
@@
 Unless stated otherwise, these experiments measure average performance over 10 episodes, use all \num{14} features introduced in \cref{sec:rethinking-the-state-space} (tile-coded individually) alongside the global state, and make use of the physical bandwidth cap rather than pursuing negative reinforcement.
-In all cases, agents learn control of the network from scratch.
+In all cases, agents learn control of the network from scratch ($\wvec{}=\bm{0}$).
 Test system specifications match those given in \cref{sec:performance-in-an-emulated-environment}.
@@
 		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
-		SPF, remarkably, displays performance on par with Marl++ for this more difficult topology in spite of underperforming when $n=2$.
+		SPF, remarkably, slightly outperforms Marl++ (with lower variance) for this more difficult topology despite being worse for smaller $n$.
 		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
@@
 		Marl and Marl++ begin with essentially identical efficacy, with Marl++ having a small advantage at larger $n$.
-		SPF is slightly worse than Marl and Marl++ at low $n$, but achieves considerably better median performance and lower variance than the other models for higher $n$.
+		SPF is slightly worse than Marl and Marl++ at low $n$, but achieves considerably better median performance and lower variance than the other models for large $n$.
 		\label{fig:tcp-box}
@@
 	\caption{
-		Online performance for $n=16$ hosts per egress point when benign traffic is TCP-like.
-		At this level of host density, SPF reaches its peak performance sooner and is considerably more consistent throughout.
-		Marl++ takes longer to reach its most effective policy, we achieve visibly better peak performance.
+		Online performance of standard and single-agent models with $n=16$ hosts per egress point and TCP-like benign traffic.
+		At this level of host density, SPF reaches its peak performance sooner and is considerably more consistent throughout the episode.
+		Marl++ takes longer to reach its most effective policy, and achieves visibly better peak performance than the other standard models.
 		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
@@
 For SPF, however, negative reinforcement is typically a worse choice.
-Where TCP-like traffic is concerned in both models, we see that this strategy almost always leads to worse policies.
+Where TCP-like traffic is concerned in both models, we see that this strategy almost always leads to worse policies and higher collateral damage.
 
@@
 In a handful of UDP instances, banding can lead to \emph{better} average performance.
-More study is likely required into this; we choose to replace $U_s$ with $0.99U_s$ to minimise the impact on legitimate states, but this choice remains somewhat arbitrary.
+More study is likely required into this; we replace $U_s$ with $0.99U_s$ to minimise the impact on legitimate states, but this choice remains somewhat arbitrary.
 
@@
 This may be achieved in a real network by randomly assigning new hosts to be controlled exclusively by one feature's policy, allowing agents to more easily learn \emph{which} information led to the correct action being taken.
-We reduce the timestep length to \SI{10}{\milli\second} to train each agent in less wall-clock time, at the cost of noisier data (and reduced learning rate).
+To reflect this (and train all parts of an agent in the same time frame), we reduce the timestep length to \SI{10}{\milli\second} to train each agent in less wall-clock time, at the cost of noisier data (and reduced learning rate).
 The results show no real change for Marl++, but a considerable gain for SPF against UDP (and performance drop against TCP).
-It is likely that, as did the different protocols in \cref{fig:udp-feature-plots,fig:tcp-cap-feature-plots}, Marl++ and SPF have different optimal feature sets for different traffic behaviours---many of which have been excluded.
+It's most likely that SPF is sensitive to noisy data when trying to learn how to act on TCP, and that as the different protocols did in \cref{fig:udp-feature-plots,fig:tcp-cap-feature-plots}, Marl++ and SPF have different optimal feature sets for different traffic behaviours---many of which may have been excluded.
 
@@
 The increased state-space size makes this a more likely occurrence, as is the unclear affect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state-space.
-We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to unpick the true value of each action), and controlling shared global state.
+We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to observe the true value of each action), and controlling shared global state.
 Our pre-training results support the idea that it may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows.
@@
 The benefits we have at present are, however, substantial.
-What we offer above the alternate approaches we discussed in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
-Moreover, our decision making processes are entirely agnostic with regards to the protocol or content of inbound traffic, offering future-proofing against the introduction of new transports.
+What we offer above the alternate approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
+Moreover, our decision making processes are entirely agnostic of the protocol or content of inbound traffic, offering future-proofing against the introduction of new transports.
 
@@
 	\caption{
-		Mean policy for Mean IAT when $n=16$.
-		The column at any row is the action most likely to be chosen: this isn't actually a policy because it's mean'd.
+		Mean Marl action values for Mean IAT when $n=16$.
+		The lightest cell in any row would ordinarily be the action most likely to be chosen: this isn't actually a policy because it's mean'd.
 		Repeated values originate from the bias tile, which is always active.
@@
 With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
-We may see this in \cref{fig:intern-16-tcp-iat}... ?? ELABORATE
+\Cref{fig:intern-16-tcp-iat} shows how we may visualise the portion of a policy for a single feature.
 Furthermore, system state which is dependent on many signals drawn from across a wide network (such as our global state) is difficult to exert precise control over.
@@
 
-Where overheads are concerned, the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies---\SIrange{100}{120}{\kibi\byte} from our experiments.
+Where overheads are concerned, the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies: \SIrange{16.5}{17.0}{\kibi\byte} from our experiments.
 As we have described earlier, action updates require a constant number of floating point operations---\num{122} floating point additions and \num{2} multiplications per update of $\wvec{}$, above the \num{140} additions required to choose an action.
@@
 Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
-We believe that it may be possible to implement similar behaviour on plain switches through application of \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
+We believe that it may be possible to implement similar behaviour on standard commodity switches through application of \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
 
@@
 However, the measurements acquired in such a scenario are far more likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
-The main bottlenecks are most likely in forwarding the load measurements from various aggregation points (which can be made more efficient by way of multicast) and in running some estimator to condition the reward function upon.
+The main bottlenecks are most likely in forwarding the load measurements from various aggregation points (which can be made more efficient through multicast) and in running some estimator to condition the reward function upon.
 %?? Potential for dividing up pre-feature training across different agents, or train like this locally (sets of flows trained by 1 sub-model).
 %?? Different global state per-agent? Seems this must be locally trained, just make it (everything up the chain to key destinations).
-We expect that agents will be able to share policies for all fields except global state, which may help to offset the reduced rate of incoming experience.
+We expect that agents will be able to share policies for all features, which may help to offset the reduced rate of incoming experience.
 Regardless, it will take longer to achieve enough state-state transitions to converge on a good policy.
@@
 Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
-Our evaluation shows that our new agent designs advance the state-of-the-art in RL-based DDoS prevention, with SPF showing the most promise for future evaluation.
+Our evaluation shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, with SPF showing the most promise for future evaluation.
 

### 2018-11-14 paper/paper.tex
Everything squeezed into its rightful place. All's good in the world.
2 results to go...

@@
 We aim to take steps to redress this and establish their proper capabilities, beyond simple ``blind application''.
-
-?? Expand as required
-
-?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
-?? THAT IS A MAJOR CONTRIB, MENTION IT EVERYWHERE YOU CAN
-
-In service of this goal, this paper contributes:
+%?? Expand as required
+What approaches do exist are aimed towards the task of adaptive online DDoS mitigation, and rely upon learning control of probabilistic packet drop.
+%?? THAT IS A MAJOR CONTRIB, MENTION IT EVERYWHERE YOU CAN
+
+%?? Discuss the most important conclusions before the outline.
+We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
+The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting SDN architecture.
+In tandem with the development and evaluation of an effective state-space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
+We find that this our per-source decisions improve substantially upon the state-of-the-art, and observe that our second model achieves excellent performance for large $n$.
+Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
+%?? Also algorithmic enhancements such as multiple actions per timestep, 
+%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
+Briefly, this paper contributes:
 \begin{itemize}
@@
 	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
-	\item Two source-level granularity approaches to RL-driven DDoS prevention, improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
+	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Marl++}, \emph{SPF}), improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}) and discussion of security concerns and real-world deployment (\cref{sec:discussion}).
 \end{itemize}
@@
 \centering
-\resizebox{\linewidth}{!}{
+\resizebox{0.9\linewidth}{!}{
 				\begin{tikzpicture}[
@@
 \begin{subequations}
+	\vspace{-0.5cm}
 \newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
@@
 \begin{figure}
-	\includegraphics[width=\linewidth]{../plots/online-varyN-binary}
+	\centering
+	\includegraphics[width=0.90\linewidth]{../plots/online-varyN-binary}
 	\caption{
@@
 	1 - (P_G^n + P_B^n) &\le 1 - (P_G^{n+1} + P_B^{n+1})\\
-	P_{M,n} &\le P_{M,n+1}
+	P_{M,n} &\le P_{M,n+1} \qedhere
 	\end{align*}
@@
 \begin{figure}
-	\includegraphics[width=\linewidth]{../plots/online-varyN-nginx}
+	\centering
+	\includegraphics[width=0.95\linewidth]{../plots/online-varyN-nginx}
 	\caption{
@@
 	\includegraphics[width=\linewidth]{../plots/ftprep-box}
+	\vspace{-1.2cm}
 	\caption{
@@
 \begin{figure}
+	\centering
 %	\begin{subfigure}{0.32\linewidth}
@@
 	\includegraphics[width=\linewidth]{../plots/ftprep-tcp-box}
-
+	\vspace{-1.2cm}
 	\caption{
@@
 	\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-box}
-
+	\vspace{-1.2cm}
 	\caption{
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/udp-box}
+	\includegraphics[width=0.95\linewidth]{../plots/udp-box}
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/tcp-box}
+	\includegraphics[width=0.95\linewidth]{../plots/tcp-box}
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/tcp-16-single}
+	\includegraphics[width=0.95\linewidth]{../plots/tcp-16-single}
 
@@
 To reflect this (and train all parts of an agent in the same time frame), we reduce the timestep length to \SI{10}{\milli\second} to train each agent in less wall-clock time, at the cost of noisier data (and reduced learning rate).
-The results show no real change for Marl++, but a considerable gain for SPF against UDP (and performance drop against TCP).
-It's most likely that SPF is sensitive to noisy data when trying to learn how to act on TCP, and that as the different protocols did in \cref{fig:udp-feature-plots,fig:tcp-cap-feature-plots}, Marl++ and SPF have different optimal feature sets for different traffic behaviours---many of which may have been excluded.
+The results show no real change for Marl++, but a considerable gain for SPF (though smaller than a single learner).
+%It's most likely that SPF is sensitive to noisy data when trying to learn how to act on TCP, and that as the different protocols did in \cref{fig:udp-feature-plots,fig:tcp-cap-feature-plots}, Marl++ and SPF have different optimal feature sets for different traffic behaviours---many of which may have been excluded.
 
@@
 	\caption{
-		Mean Marl action values for Mean IAT when $n=16$.
-		The lightest cell in any row would ordinarily be the action most likely to be chosen: this isn't actually a policy because it's mean'd.
-		Repeated values originate from the bias tile, which is always active.
-		?? Shows... what?
+		Mean Marl action values from pretraining on mean IATs ($n=16$).
+		As these are mean values, lighter cells in each row indicate actions \emph{likely} to be taken by an agent's policy.
+		Repeated values originate from the bias tile, which is always active, and indicate regions of the state-space which have yet to be visited---here, this is most of the state space.
+		The measured effectiveness of this feature then suggests that a low-resolution coding over the standard region may, in fact, be a better choice.
+		We see that agents prefer to choose $p=0.0$ in most states, but higher $p$ when IATs are particularly small.
 		\label{fig:intern-16-tcp-iat}
+	}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\linewidth]{../plots/policy-4-tcp-f7-mean}
+
+	\caption{
+		Mean Marl action values from pretraining on $\Delta$ out rates ($n=4$).
+		A sudden ramp-up of server-to-host traffic is used as a strong indicator of flow legitimacy, while more punishing actions have comparatively higher value as $n$ decreases.
+		Furthermore, we see again that a region of the state space has gone somewhat unexplored---we observed after plotting this visualisation that many decreases in this feature are too small to hit the adjacent tile, which implies a mixed-resolution coding may improve an agent's policy further.
+		\label{fig:intern-16-tcp-something}
 	}
@@
 With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
-\Cref{fig:intern-16-tcp-iat} shows how we may visualise the portion of a policy for a single feature.
+\Cref{fig:intern-16-tcp-iat,fig:intern-16-tcp-something} show how we may visualise the portion of a policy for each feature, and describe what information can be gained from doing so.
 Furthermore, system state which is dependent on many signals drawn from across a wide network (such as our global state) is difficult to exert precise control over.

### 2018-11-14 paper/paper.tex
Consistency!

@@
 In tandem with the development and evaluation of an effective state-space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
-We find that this our per-source decisions improve substantially upon the state-of-the-art, and observe that our second model achieves excellent performance for large $n$.
+Our introduction of per-source decisions improves substantially upon the state-of-the-art, and we show that our second model achieves excellent performance for large $n$.
 Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
@@
 
-\subsection{Per-feature pretraining}
+\subsection{Per-feature pre-training}
 To investigate the effects of pre-training individual features, we chose our four best-performing features (global state, mean IAT, $\Delta$ in/out rates) and trained each over \num{10000} iterations, before evaluating the policy obtained by summing the action values suggested by each.
@@
 	\caption{
-		Mean Marl action values from pretraining on mean IATs ($n=16$).
+		Mean Marl action values from pre-training on mean IATs ($n=16$).
 		As these are mean values, lighter cells in each row indicate actions \emph{likely} to be taken by an agent's policy.
@@
 	\caption{
-		Mean Marl action values from pretraining on $\Delta$ out rates ($n=4$).
+		Mean Marl action values from pre-training on $\Delta$ out rates ($n=4$).
 		A sudden ramp-up of server-to-host traffic is used as a strong indicator of flow legitimacy, while more punishing actions have comparatively higher value as $n$ decreases.

### 2018-11-14 paper/paper.tex
And, done.
Some notes...

@@
 To reflect this (and train all parts of an agent in the same time frame), we reduce the timestep length to \SI{10}{\milli\second} to train each agent in less wall-clock time, at the cost of noisier data (and reduced learning rate).
-The results show no real change for Marl++, but a considerable gain for SPF (though smaller than a single learner).
+The results show little change for Marl++, but generally indicate a considerable gain for SPF (though smaller than what we'd measured for a single agent system).
 %It's most likely that SPF is sensitive to noisy data when trying to learn how to act on TCP, and that as the different protocols did in \cref{fig:udp-feature-plots,fig:tcp-cap-feature-plots}, Marl++ and SPF have different optimal feature sets for different traffic behaviours---many of which may have been excluded.
@@
 		Repeated values originate from the bias tile, which is always active, and indicate regions of the state-space which have yet to be visited---here, this is most of the state space.
+		Agents visibly learn action values for tiles which override the default bias tile's preferences.
 		The measured effectiveness of this feature then suggests that a low-resolution coding over the standard region may, in fact, be a better choice.
@@
 		Mean Marl action values from pre-training on $\Delta$ out rates ($n=4$).
-		A sudden ramp-up of server-to-host traffic is used as a strong indicator of flow legitimacy, while more punishing actions have comparatively higher value as $n$ decreases.
+		A sudden ramp-up of server-to-host traffic is used as a strong indicator of flow legitimacy, while more punishing actions have comparatively higher value as $\Delta$ out rate decreases.
 		Furthermore, we see again that a region of the state space has gone somewhat unexplored---we observed after plotting this visualisation that many decreases in this feature are too small to hit the adjacent tile, which implies a mixed-resolution coding may improve an agent's policy further.

### 2019-01-11 paper/paper.tex
Appease the PCs several months in advance.
WIP on explanations/justifications

@@
 % !BIB program = biber
-\documentclass[conference, letterpaper, 10pt, times]{IEEEtran}
+\documentclass[conference, a4paper, 10pt, times]{IEEEtran}
 \usepackage{nopageno}
@@
 \usepackage{hyperref}
-\usepackage{cleveref}
+\usepackage[nameinlink]{cleveref}
 \newcommand{\crefrangeconjunction}{--}
@@
 	pdftitle={\mytitle{}},
-%	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros, Simon Rogers}
+%	pdfauthor={Kyle A. Simpson, Simon Rogers, Dimitrios P. Pezaros}
 	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
@@
 
+\definecolor{inferno0}{rgb}{0.001462 0.000466 0.013866}
+\definecolor{inferno64}{rgb}{0.341500 0.062325 0.429425}
+\definecolor{inferno128}{rgb}{0.735683 0.215906 0.330245}
+\definecolor{inferno192}{rgb}{0.978422 0.557937 0.034931}
+\definecolor{inferno255}{rgb}{0.988362 0.998364 0.644924}
+
+\colorlet{lowac}{inferno128}
+\colorlet{midac}{inferno192}
+\colorlet{highac}{inferno255}
+
 %-------------------------------------%
@@
 %	Kyle A.\ Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
-%	Kyle A.\ Simpson,
-%	Dimitrios P.\ Pezaros,
-%   Simon Rogers\\
-%	University of Glasgow, Glasgow, Scotland,\\
-%	\email{k.simpson.1@research.gla.ac.uk}
-Anonymous Giraffe,
-Anonymous Badger,
-Anonymous Ibex\\
-Unnamed Department, Nowhere\\
-\email{giraffe.a@unnamed.com}
+	Kyle A.\ Simpson,
+	Simon Rogers,
+	Dimitrios P.\ Pezaros,\\
+	University of Glasgow, Glasgow, Scotland,\\
+	\email{k.simpson.1@research.gla.ac.uk}
+%Anonymous Giraffe,
+%Anonymous Badger,
+%Anonymous Ibex\\
+%Unnamed Department, Nowhere\\
+%\email{giraffe.a@unnamed.com}
 }
@@
 \emph{Reinforcement learning} (RL) may overcome the detection problem for certain classes of anomaly by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available).
-We examine how the paradigm has been applied to the task of DDoS prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.
+We examine how the paradigm has been applied to the task of \emph{Distributed Denial of Service} (DDoS) attack prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.
 These flaws---poor handling of TCP traffic and high host density at egress points---inspire and direct an exploration of the design space of RL models.
@@
 
-While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \citeauthor{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems, and to quite a large extent this remains the case today \cite{DBLP:conf/sp/SommerP10}.
-Their application is made difficult due to significant operational differences from standard ML tasks, alongside certain characteristics of network traffic.
-Briefly, these are the diversity of network traffic across varying timescales \cite{DBLP:conf/sp/SommerP10} and significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
+While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \citeauthor{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems---a distinct lack of real-world system deployments \cite{DBLP:conf/sp/SommerP10}.
+To quite a large extent, this remains the case today.
+?? RESTRUCTURE
+These authors posit that their application is made difficult due to significant operational differences from standard ML tasks.
+??
+(quoted: high cost of errors, lack of training data, semantic gap between results and operational interpretation, variability in input data, fundamental difficulties for conducting sound evaluation)
+This is further aggravated by certain characteristics of network traffic; the diversity of network traffic across varying timescales \cite{DBLP:conf/sp/SommerP10} and significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
 Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
@@
 
+??
+Make it clearer here what problem I specifically want to solve: principally a particular class of DDoS attacks; volume-based DDoS attacks.
+Amplification attacks are just a specialisation, this can be made more obvious.
+I think I need to be clearer about the \emph{intended deployment environment} of service hosts (i.e., not ISPs).
+
 For certain classes of problem e.g., flooding-based DDoS attacks, \emph{reinforcement learning} (RL) offers another perspective.
+?? Unclear explanation of RL here?
 The role of a RL agent differs from that of a standard classifier, adaptively reacting to threats by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
@@
 In tandem with the development and evaluation of an effective state-space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
-Our introduction of per-source decisions improves substantially upon the state-of-the-art, and we show that our second model achieves excellent performance for large $n$.
+Our introduction of per-source decisions improves substantially upon the state-of-the-art, and we show that our second model achieves excellent performance for high host density.
 Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
@@
 %?? Introduce RL, related definitions etc.
+
+\subsection{Distributed Denial of Service}
+%??
+DDoS attack variants (leading into characteristics, supporting features).
+Amplification (UDP \cite{DBLP:conf/ndss/Rossow14}, TCP \cite{DBLP:conf/uss/KuhrerHRH14}), Transit-link (Crossfire \cite{DBLP:conf/sp/KangLG13}, Coremelt \cite{DBLP:conf/esorics/StuderP09}).
+Mirai botnet's involvement \cite{DBLP:conf/uss/AntonakakisABBB17}.
+
+%?? Explain amplification attack, maybe transit-link?
+
+\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
+Attackers achieve this either by exploiting peculiarities of operating system or application behaviour (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{flooding-based attacks}).
+Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
+
+Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
+\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most exploitable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
+Malicious hosts send many small requests, spoofed to appear as though they originated from the victim, causing many large replies to be sent to the intended target---significantly increasing a botnet's throughput while masking the identity of each participant.
+\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
+
+?? EXPAND
+
 \subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
@@
 There is immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
-In particular, we shall focus on methods which choose actions according to their value estimates from the current state.
-To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+ \dim{\mathcal{A}})$-dimensional grids with an optional bias component.
-The numbers of tilings and of tiles per dimension have a direct effect on the resolution of the discretisation.
-Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state-action pair exists within the corresponding tile.
-This transform, as an example of linear function approximation, then necessitates the use of \emph{1-step semi-gradient Sarsa} \cite[pp.\ \numrange{243}{244}]{RL2E}.
-Given a weight vector $\wvec{0}=\bm{0}$ and learning rate $\alpha$, we may approximate an action $a$'s value in any state $s$ taking $\operatorname{q}(s, a) \approx \acval{s}{a}{\wvec{}}$ by continually updating $\wvec{t}$ as follows:
+In particular, we shall focus on methods which choose actions according to their value estimates from the current state: let $\operatorname{q}(s, a) \in \mathbb{R}$ denote the estimate of action $a$'s value if it were to be taken in state $s$.
+%?? Revisit the maths here, d, dim A, s...
+Exact (tabular) representations require that we store a value estimate for each action in every state---if state is real-valued or high-dimensional, then computation and storage quickly become infeasible.
+To cope with a continuous state and/or action space, one valuable technique is to employ linear function approximation backed by \emph{tile coding} \cite[pp.\ \numrange{217}{221}]{RL2E}.
+
+\begin{figure}
+	\centering
+	\begin{subfigure}{0.4\linewidth}
+		\resizebox{\linewidth}{!}{
+		\begin{tikzpicture}
+		\node at (0,0.3){$s = \begin{pmatrix}
+			0.7 \\
+			0.3
+			\end{pmatrix}$};
+		\node at (0,-1) {$\bm{x}(s,\cdot) = \begin{Bmatrix}
+			T_{1,9}, \\
+			T_{2,5}, \\
+			T_{\mathit{bias}}
+			\end{Bmatrix}$};
+
+		\node at (2.5,-0.5) {
+			\begin{tikzpicture}
+			\draw[step=0.5cm,color=uofgcobalt,opacity=0.7,shift={(0,0)},label=above:{Tiling 0}] (-0.5,-0.5) grid (1,1);
+			\fill[uofgcobalt,opacity=0.5] (0.5,-0.5) rectangle (1,0);
+			\node[color=uofgcobalt] at (0,1.1) {\footnotesize Tiling 1};
+
+			\draw[step=0.5cm,color=uofgpumpkin,opacity=0.7,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
+			\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
+			\node[color=uofgpumpkin] at (0.25,-0.95) {\footnotesize Tiling 2};
+
+			\node[circle, black, draw,
+			fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
+%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];
+
+			\draw[->] (-0.25,-0.5)--(-0.25,0.85);
+			\draw[->] (-0.25,-0.5)--(1.1,-0.5);
+
+			\node at (1,-0.7) {\footnotesize 1};
+			\node at (-0.4,0.75) {\footnotesize 1};
+			\node at (-0.35,-0.6) {\footnotesize 0};
+			\end{tikzpicture}
+		};
+
+		\end{tikzpicture}
+		}
+		\caption{Tile Coding\label{fig:tilecode-code}}
+	\end{subfigure}%
+	\begin{subfigure}{0.4\linewidth}
+		\resizebox{\linewidth}{!}{
+		\begin{tikzpicture}
+		% Top half
+		\def\topacs{
+			{-0.3,-0.5,-0.1,0.8},
+			{0.1,0.1,-0.2,0.3},
+			{0.3,0.4,0.2,-0.4}%
+		}
+
+		\foreach \line [count=\y] in \topacs {
+			\foreach \pix [count=\x] in \line {
+				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
+					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
+					\draw[shift={(-0.7,0)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
+				}{
+					\pgfmathtruncatemacro\lambda{\pix*100}
+					\draw[shift={(-0.7,0)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
+				}
+				\node[shift={(-0.35,0.175)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
+			}
+		}
+
+		\draw[xstep=0.7cm,ystep=0.35cm,shift={(0,0)}] (0,-1.06) grid (2.8,0);
+		\node[label=left:{$T_{1,9}$},shift={(0,-0.125)}] at (0,0) {}; 
+		\node[label=left:{$T_{2,5}$},shift={(0,-0.125)}] at (0,-0.375) {}; 
+		\node[label=left:{$T_{\mathit{bias}}$},shift={(0,-0.125)}] at (0,-0.75) {};
+
+		% bottom half
+		\def\botacs{
+			{0.1,0.0,-0.1,0.7}%
+		}
+
+		\foreach \line [count=\y] in \botacs {
+			\foreach \pix [count=\x] in \line {
+				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
+					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
+					\draw[shift={(-0.7,-1.275)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
+				}{
+					\pgfmathtruncatemacro\lambda{\pix*100}
+					\draw[shift={(-0.7,-1.275)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
+				}
+				\node[shift={(-0.35,-1.1)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
+			}
+		}
+
+		\draw[xstep=0.7cm,ystep=0.35cm,shift={(0,-1.275)}] (0,-0.35) grid (2.8,0);
+		\node[label=left:{$\mathit{Total}$},shift={(0,-0.175)}] at (0,-1.275) {};
+
+		\draw [->] (2.45, -1.9) -- (2.45, -1.65);
+		\end{tikzpicture}
+		}
+		\caption{Action Selection\label{fig:tilecode-select}}
+	\end{subfigure}
+
+	\caption{
+		An example of tile coding for 2-dimensional state and 4 actions, using 2 tilings, 3 tiles per dimension, and a bias tile.
+		All components of $s$ are clamped to the range $[0,1]$.
+		For simplicity, we denote $\bm{x}(s,\cdot)$ as a list of indices and represent the values of all actions at each tile with a vector.
+		(a) The state $s$ activates the bias tile and exactly one tile in each tiling.
+		(b) The action values of active tiles are summed to produce the current value estimate for each action available in $s$---for this state, local knowledge ensures that action 4 is chosen by the greedy policy despite typically being a poor choice elsewhere.
+		\label{fig:tilecode}
+	}
+\end{figure}
+
+Tile coding is a form of feature representation which converts a state-action pair into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing a $d$-dimensional subset of the space into a number of overlapping grids with an optional bias component.
+Each tile corresponds to an entry of $\operatorname{\mathbf{x}}(s, a)$ which is set to 1 if the state-action pair lies within it.
+\Cref{fig:tilecode-code} demonstrates the process for a 2-dimensional state space, and how the numbers of tilings and tiles per dimension control feature resolution and generalisation.
+Moreover, to capture combinatorial effects or create multi-scale representation we may combine codings by concatenating individual feature vectors.
+We may then approximate an action's value with respect to a policy parameter vector $\wvec{}$, defining some $\acval{s}{a}{\wvec{}} \approx \operatorname{q}(s, a)$:
+\begin{equation}
+%	\begin{gather}
+\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a)
+%	\end{gather}
+\label{eqn:lin-approx}
+\end{equation}
+As each component of $\wvec{}$ is the value estimate of the corresponding tile, learning an effective policy is equivalent to learning $\wvec{}$.
+Given a learning rate $\alpha \in \mathbb{R}$ and initialising $\wvec{0}=\bm{0}$, we may continually update $\wvec{t}$ using the \emph{1-step semi-gradient Sarsa} algorithm \cite[pp.\ \numrange{243}{244}]{RL2E}:
 \begin{subequations}
 	\begin{gather}
-	\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a),\\
 	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
@@
 	\label{eqn:sg-sarsa}
-	taking $\nabla$ with respect to $\wvec{}$.
+	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
 \end{subequations}
@@
 Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might use some other mechanism.
+\Cref{fig:tilecode-select} extends our working example to show how the value of each action is computed, combining a global estimate ($T_{\mathit{bias}}$) with knowledge particular to each state.
+
+\begin{figure}
+	\centering
+	\resizebox{0.5\linewidth}{!}{
+		\begin{tikzpicture}		
+		% Top half
+		\def\startacs{
+			{0.8},
+			{0.3},
+			{-0.4}%
+		}
+
+		\foreach \line [count=\y] in \startacs {
+			\foreach \pix [count=\x] in \line {
+				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
+					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
+					\draw[shift={(-0.7,0)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
+				}{
+					\pgfmathtruncatemacro\lambda{\pix*100}
+					\draw[shift={(-0.7,0)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
+				}
+				\node[shift={(-0.35,0.175)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
+			}
+		}
+
+		\draw[xstep=0.7cm,ystep=0.35cm,shift={(0,0)}] (0,-1.06) grid (0.7,0);
+		\node[label=left:{$T_{1,9}$},shift={(0,-0.125)}] at (0,0) {}; 
+		\node[label=left:{$T_{2,5}$},shift={(0,-0.125)}] at (0,-0.375) {}; 
+		\node[label=left:{$T_{\mathit{bias}}$},shift={(0,-0.125)}] at (0,-0.75) {};
+		\node[label=below:{\footnotesize Action 4},shift={(0.35,-0.125)}] at (0,-0.75) {};
+		\node[label=above:{\footnotesize $\wvec{t}$},shift={(0.35,-0.125)}] at (0,0) {};
+
+		\def\finalacs{
+			{0.7},
+			{0.2},
+			{-0.5}%
+		}
+
+		\foreach \line [count=\y] in \finalacs {
+			\foreach \pix [count=\x] in \line {
+				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
+					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
+					\draw[shift={(2-0.7,0)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
+				}{
+					\pgfmathtruncatemacro\lambda{\pix*100}
+					\draw[shift={(2-0.7,0)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
+				}
+				\node[shift={(2-0.35,0.175)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
+			}
+		}
+
+		\draw[xstep=0.7cm,ystep=0.35cm,shift={(2,0)}] (0,-1.06) grid +(0.7,1.06);
+		\node[label=below:{\footnotesize Action 4},shift={(2.35,-0.125)}] at (0,-0.75) {};
+		\node[label=above:{\footnotesize $\wvec{t+1}$},shift={(0.35,-0.125)}] at (2,0) {};
+
+		\draw[->] (0.9,-0.5) -- node[above] {$+ \alpha \delta_t$} (1.8,-0.5);
+		\end{tikzpicture}
+	}
+
+\caption{
+	The update step for \cref{fig:tilecode}, given an observed TD error $\delta_t=-0.2$ (indicating a lower observed reward than the expected long-term value of \num{0.7}) and $\alpha=0.5$.
+	Action 4's value is thus reduced in the tiles associated with state $s$, but remains the most likely choice; the negative $delta_t$ may have arisen from noise in the reward signal.
+	For illustrative purposes, we choose an unrealistically high $\alpha$ (typically, $\alpha=0.05$ or lower is a more reasonable choice).
+	\label{fig:tilecode-update}
+}
+\end{figure}
 
@@
 \end{equation}
+\Cref{fig:tilecode-update} shows how this applies to our prior example.
 If desired we may define a state-space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
@@
 %?? Discuss mininet? Networking terms? SDN stuff?
-
-\subsection{Distributed Denial of Service (DDoS)}
-%??
-DDoS attack variants (leading into characteristics, supporting features).
-Amplification (UDP \cite{DBLP:conf/ndss/Rossow14}, TCP \cite{DBLP:conf/uss/KuhrerHRH14}), Transit-link (Crossfire \cite{DBLP:conf/sp/KangLG13}, Coremelt \cite{DBLP:conf/esorics/StuderP09}).
-Mirai botnet's involvement \cite{DBLP:conf/uss/AntonakakisABBB17}.
-
-%?? Explain amplification attack, maybe transit-link?
-
-DDoS attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
-Attackers achieve this either by exploiting peculiarities of operating system or application behaviour (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{flooding-based attacks}).
-Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
-
-Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
-\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most exploitable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
-Malicious hosts send many small requests, spoofed to appear as though they originated from the victim, causing many large replies to be sent to the intended target---significantly increasing a botnet's throughput while masking the identity of each participant.
-\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
 
@@
 We assume that attacks will be bandwidth-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
+?? VERIFY AND RESEARCH THAT: REVIEWER 1 BELIEVES OTHERWISE.
 The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.
@@
 \section{Foundational Analysis}\label{sec:environment-and-rl-algorithm}
+?? REWORK AND GUT THIS SECTION + NEXT
+
 %?? State the problem and what we want to achieve.
@@
 \end{gather}
-where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.\label{eqn:reward}
+where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.
+$\cond{f}$ and $\cond{c}{(v)}$ represent the ``overloaded'' condition for these reward functions.\label{eqn:reward}
 \end{subequations}
@@
 Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
+?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
 Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network.
@@
 for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions, and the probability $p \in (0, 1]$ that a packet is lost.
+?? Does Mathis still hold? We observed similar trends empirically, but perhaps not the \emph{exact} curve or falloff behaviour...
 Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant; good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
@@
 The protocol is known to carry \SIrange{79}{94}{\percent} of packets, corresponding to \SIrange{89}{98}{\percent} of data transferred by volume \cite{DBLP:conf/saint/ZhangDJC09}.
+?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE TRAFFIC IN MODERN DAY WITH CAIDA.
 As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development to properly handle traffic with such characteristics.
@@
 \fakepara{Source IP address}
-While ordinarily trivial to spoof, in real-world scenarios the addresses of a set of reflector nodes might exhibit similarity, e.g., similarly unhardened services exposed by a single organisation.
+While ordinarily trivial to spoof (and thus of limited use for many classes of attack), reflectors are themselves legitimate services being abused by spoofing attackers.
+As a result, they communicate with attack victims using their own IP address.
+In real-world scenarios the addresses of a set of reflector nodes might exhibit similarity, e.g., similarly unhardened services exposed by a single organisation.
+?? Can we support this claim?
 
@@
 In no cases, however, do we manage to completely block attack traffic---at convergence, we observe that system load remains consistently at $U_s$.
+
+?? POTENTIALLY CUT THE ``SYSTEM DYNAMICS'' AND/OR BANDED REWARDS, THEY DON'T ADD MUCH ANYWAY.
 
@@
 \subsection{Banded rewards and negative reinforcement}
+?? PROBABLY CUT THIS
+
 In general, we find that explicitly training a policy to keep the system in a valid state via negative reinforcement (i.e., an \emph{uncapped} network) is most effective for UDP-like traffic with Marl++.
@@
 
+?? Actually talk about Marl (???) to appease reviewer \#1.
+
 \emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
@@
 %?? IDEA: try out average reward, TD($\lambda$) methods as future work...
-The extant weaknesses invite many further improvements worth investigation.
+The remaining weaknesses invite many further improvements worth investigation.
 A natural research direction to enhance this work would be the combination of the classic function approximation techniques we make use of alongside the improved algorithms that the RL community has introduced in the past few years.
@@
 This work was supported by a nameless funder [grant number 0].
+%This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].
 

### 2019-01-14 paper/paper.tex
Intro + RL Background fixes.

@@
 
-%\newcommand{\mytitle}{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention}
 \newcommand{\mytitle}{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention}
+%\newcommand{\mytitle}{Per-Host DDoS Mitigation by Direct-Control Reinforcement Learning}
 
@@
 %	Kyle A.\ Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
-	Kyle A.\ Simpson,
-	Simon Rogers,
-	Dimitrios P.\ Pezaros,\\
-	University of Glasgow, Glasgow, Scotland,\\
-	\email{k.simpson.1@research.gla.ac.uk}
-%Anonymous Giraffe,
-%Anonymous Badger,
-%Anonymous Ibex\\
-%Unnamed Department, Nowhere\\
-%\email{giraffe.a@unnamed.com}
+%	Kyle A.\ Simpson,
+%	Simon Rogers,
+%	Dimitrios P.\ Pezaros,\\
+%	University of Glasgow, Glasgow, Scotland,\\
+%	\email{k.simpson.1@research.gla.ac.uk}
+Anonymous Giraffe,
+Anonymous Badger,
+Anonymous Ibex\\
+Unnamed Department, Nowhere\\
+\email{giraffe.a@unnamed.com}
 }
@@
 
-Network anomaly detection and intrusion detection/prevention are continually evolving problems, compounded by the partial, non-IID view of data at each point in the network.
+Network anomaly detection and intrusion detection/prevention are continually evolving problems, compounded by the partial, non-\emph{independent and identically distributed} (IID) view of data at each point in the network.
 Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, and availability without being detected \cite{DBLP:journals/comsur/BhuyanBK14}.
@@
 To quite a large extent, this remains the case today.
-?? RESTRUCTURE
-These authors posit that their application is made difficult due to significant operational differences from standard ML tasks.
-??
-(quoted: high cost of errors, lack of training data, semantic gap between results and operational interpretation, variability in input data, fundamental difficulties for conducting sound evaluation)
-This is further aggravated by certain characteristics of network traffic; the diversity of network traffic across varying timescales \cite{DBLP:conf/sp/SommerP10} and significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
+These authors posit that their application is made difficult due to significant operational differences from standard ML tasks, including: the high cost of errors and extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}; a general lack of recent, openly available (and high-quality) training data; and diversity of network traffic across varying timescales combined with significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
 Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
-Further issues arise from the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges encountered when learning from unlabelled (often partial) data.
+Learning is made harder still by the challenges encountered when learning from unlabelled (often partial) data.
 All of these factors greatly inflate the difficulty of the detection problem.
 
-??
+%??
 Make it clearer here what problem I specifically want to solve: principally a particular class of DDoS attacks; volume-based DDoS attacks.
@@
 
-For certain classes of problem e.g., flooding-based DDoS attacks, \emph{reinforcement learning} (RL) offers another perspective.
-?? Unclear explanation of RL here?
-The role of a RL agent differs from that of a standard classifier, adaptively reacting to threats by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
-In a sense, this allows us to ``overcome'' the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time; by looking for the effect rather than the cause.
-The intent is to augment what existing misuse-based solutions can provide, by automatically alerting, recording and controlling what are believed to be illegal system states.
+For certain classes of problem e.g., volume-based \emph{distributed denial of service} (DDoS) attacks, \emph{reinforcement learning} (RL) offers another perspective.
+%?? Unclear explanation of RL here?
+RL agents operate by following a \emph{policy} to interact with or control a system, while at the same time using observed performance metrics and deliberate exploration to dynamically improve this policy.
+In this way the role of a RL agent differs from that of a standard classifier, adaptively reacting to threats by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
+In a sense, this allows us to ``overcome'' some of the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time; by looking for (and controlling) the effect rather than the cause.
+Long-term, we expect that the value of RL-based defence systems will be to augment what existing misuse-based solutions can provide, by automatically alerting, recording and controlling what are believed to be illegal system states.
+The goal of this work is much less general; we aim to prevent volume-based DDoS attacks with the aid of RL-based techniques (an important goal in its own right), while bringing to light the flexibility and applicability of these techniques in the security domain.
 %Whether it takes direct control of the network, or is used indirectly to optimise a key part of another system, more powerful `deep' RL techniques (and well-founded action spaces) aren't yet well explored for network IDS/IPS.
@@
 We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
-The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting SDN architecture.
+The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
 In tandem with the development and evaluation of an effective state-space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
@@
 		}
-		\caption{Action Selection\label{fig:tilecode-select}}
+		\caption{\centering Value Estimation and Action Selection\label{fig:tilecode-select}}
 	\end{subfigure}
@@
 Each tile corresponds to an entry of $\operatorname{\mathbf{x}}(s, a)$ which is set to 1 if the state-action pair lies within it.
-\Cref{fig:tilecode-code} demonstrates the process for a 2-dimensional state space, and how the numbers of tilings and tiles per dimension control feature resolution and generalisation.
+\Cref{fig:tilecode-code} demonstrates the process for a 2-dimensional state space, and that the numbers of tilings and tiles per dimension control feature resolution and generalisation.
 Moreover, to capture combinatorial effects or create multi-scale representation we may combine codings by concatenating individual feature vectors.
@@
 	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
-	\end{gather}
-	\label{eqn:sg-sarsa}
+	\end{gather}%
+	\label{eqn:sg-sarsa}%
 	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
@@
 Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might use some other mechanism.
-\Cref{fig:tilecode-select} extends our working example to show how the value of each action is computed, combining a global estimate ($T_{\mathit{bias}}$) with knowledge particular to each state.
+\Cref{fig:tilecode-select} extends the prior working example to show how the value of each action is computed (and which action would be chosen by a greedy policy), combining a global estimate ($T_{\mathit{bias}}$) with knowledge particular to each state.
 
@@
 	The update step for \cref{fig:tilecode}, given an observed TD error $\delta_t=-0.2$ (indicating a lower observed reward than the expected long-term value of \num{0.7}) and $\alpha=0.5$.
-	Action 4's value is thus reduced in the tiles associated with state $s$, but remains the most likely choice; the negative $delta_t$ may have arisen from noise in the reward signal.
-	For illustrative purposes, we choose an unrealistically high $\alpha$ (typically, $\alpha=0.05$ or lower is a more reasonable choice).
+	Action 4's value is thus reduced in the tiles associated with state $s$, but remains the most likely choice; the negative $\delta_t$ may have arisen from noise in the reward signal.
+	For illustrative purposes, we choose an unrealistically high $\alpha$ (typically, $\alpha\le0.05$ is a more reasonable choice).
 	\label{fig:tilecode-update}
@@
 
-This combination of algorithm and coding strategy is well-optimised.
-If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
+This combination of algorithm and coding strategy is well-optimised, if actions are discrete; this allows a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each tile.
 Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}}-1)$ floating point additions per decision.

### 2019-02-05 glass-talk/glass-talk.tex
New talk...

% !TeX program = lualatex

\documentclass[aspectratio=169,xcolor={dvipsnames}
,hide notes
%,show only notes
%,show notes on second screen=right
]{beamer}
\usetheme[background=light, numbering=fraction]{metropolis}
\usepackage{appendixnumberbeamer}

%\usepackage[T1]{fontenc}

\usepackage{bm}

\usepackage[labelfont=bf,textfont={it}]{caption}
\usepackage{subcaption}
\captionsetup[figure]{justification=centering}
\captionsetup[subfigure]{justification=centering}

\usepackage{tikz}
\usetikzlibrary{arrows.meta, calc, fit, positioning}

%\usepackage{fontspec}
%\setsansfont{Fira Sans Mono}

\usepackage{etoolbox}
\usepackage[binary-units]{siunitx}
\robustify\bfseries
\sisetup{detect-all, range-phrase=--, range-units=single}

\usepackage[UKenglish]{babel}
\usepackage{csquotes}

\usepackage{amssymb}

\usepackage{lipsum}
\usepackage[basic]{complexity}
\usepackage[super,negative]{nth}

\usepackage{booktabs}

%bib
\usepackage[maxnames=3,maxbibnames=99,mincrossrefs=5,sortcites
,backend=bibtex
,style=authortitle
]{biblatex}
\addbibresource{../paper/papers-off.bib}
\addbibresource{../paper/confs-off.bib}
\addbibresource{../paper/books-off.bib}
\addbibresource{../paper/rfc.bib}
\addbibresource{../paper/misc.bib}

\newcommand{\acval}[3]{\ensuremath{\operatorname{\hat{q}}(#1, #2, #3)}}
\newcommand{\wvec}[1]{\ensuremath{\bm{w}_{#1}}}

\makeatletter
\DeclareRobustCommand{\rvdots}{%
	\vbox{
		\baselineskip4\p@\lineskiplimit\z@
		\kern-\p@
		\hbox{.}\hbox{.}\hbox{.}
}}
\makeatother

% official colours
\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}

\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.317647, 0.2}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}

\definecolor{inferno0}{rgb}{0.001462 0.000466 0.013866}
\definecolor{inferno64}{rgb}{0.341500 0.062325 0.429425}
\definecolor{inferno128}{rgb}{0.735683 0.215906 0.330245}
\definecolor{inferno192}{rgb}{0.978422 0.557937 0.034931}
\definecolor{inferno255}{rgb}{0.988362 0.998364 0.644924}

\colorlet{lowac}{inferno128}
\colorlet{midac}{inferno192}
\colorlet{highac}{inferno255}

%picky abt et al.
\usepackage{xpatch}

\makeatletter\let\expandableinput\@@input\makeatother

\xpatchbibmacro{name:andothers}{%
	\bibstring{andothers}%
}{%
	\bibstring[\emph]{andothers}%
}{}{}

%opening!

\usepackage{cleveref}
\newcommand{\crefrangeconjunction}{--}

\usepackage{fontawesome}

\addtobeamertemplate{footnote}{\vspace{-6pt}\advance\hsize-0.5cm}{\vspace{6pt}}
\makeatletter
% Alternative A: footnote rule
\renewcommand*{\footnoterule}{\kern -3pt \hrule \@width 2in \kern 8.6pt}
% Alternative B: no footnote rule
% \renewcommand*{\footnoterule}{\kern 6pt}
\makeatother

%-------------------------------------%
%-------------------------------------%

\title{Improved Direct-Control Reinforcement Learning for DDoS Prevention}
\author{Kyle A. Simpson\\
	\small{\faGithub{} \href{https://github.com/felixmcfelix}{FelixMcFelix} \hspace{0.5em} \faGlobe{} \url{https://mcfelix.me}}}
\institute{University of Glasgow}
\date{\nth{5} February, 2019}

\begin{document}

\maketitle

\begin{frame}{Introduction (I)}
	\begin{itemize}
		\item Network IDS/IPS backed by machine learning haven't taken off as hoped---particularly anomaly-based work.
		\item Detection problem tricky in this domain:
		\begin{itemize}
			\item Evolving: usage shifts, new protocols, new applications.
			\item Burstiness, seasonal variation.
			\item Need for correctness, almost no false-positive tolerance.
			\item Labelling issues.
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Introduction (II)}
\begin{itemize}
	\item Classes of problem like flooding-based DDoS attacks manifest as a service degradation.
	\begin{itemize}
		\item Can these be controlled via feedback loop?
		\item \alert{``Overcome'' the difficulties of the detection problem} by monitoring and adapting to \emph{performance characteristics and consequences} in real-time.
	\end{itemize}
	\item Goal: prevent volume-based DDoS attacks.
	\item Long-term Goal: augment signature-based approaches to provide a last line of defence.
\end{itemize}
\end{frame}

\section{Background}

\begin{frame}{DDoS Attacks}
\begin{itemize}
	\item \textbf{Distributed denial of service}: many hosts act in such a way that a service/server/resource becomes unavailable.
	\item \alert{Volumetric} attacks---Amplification, Link Flooding.
	\item \alert{Protocol} attacks---SYN flood, handshake violations...
	\item \alert{Application} attacks---e.g., slowloris.
	\item \emph{Botnets} are the main driver, most need a \alert{constant stream of traffic} to remain in effect.
	\begin{itemize}
		\item This constant traffic is what makes RL suitable! Could generalise to \emph{any such attack}, but focus on amplification attacks.
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Reinforcement Learning: The Main Idea}
	\begin{itemize}
		\item Underlying theory: systems as (discrete-time) \alert{Markov Decision Processes}---states, actions, rewards and transition probabilities.
		\begin{itemize}
			\item I.e., choosing action $a_t$ from a policy in state $s_t$, $a_t \sim \pi(s_t)$, induces the next state $s_{t+1}$ and an associated reward $r_{t+1}$.
			\item Generalises to \alert{value} $q(s,a)$ or $V(s)$---how much reward can we \emph{eventually} expect from choosing an action in the current state?
		\end{itemize}

	\item Goal: train an agent to make optimal decisions based on observed state.
	\begin{itemize}
		\item Formally, learn a \alert{policy} to maximise the \alert{expected discounted reward}\footcite{RL2E}.
	\end{itemize}

%		\item The world/environment is believed to be modelled by stochastic state-transition probabilities, reward function distributions...
%		\begin{itemize}
%			\item But we don't need to model that!
%		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Reinforcement Learning: Benefits}
\begin{itemize}
	\item We can learn the optimal policy \alert{without modelling the world ourselves}.
	\item Formulation allows learning adaptively and online, so long as a reward signal is available.
	\item Many choices of algorithms, update mechanisms, function approximations, dependence on value functions, schemes for action selection/exploration...
	\begin{itemize}
		\item Orthogonal concerns, allowing tunable algorithm design.
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Reinforcement Learning: How it Works (I)}
\begin{itemize}
	\item How do we handle value estimation for real-valued, multi-dimensional state?
	\item Define (wrt. an arbitrary weight vector $\wvec{}$) some approximation function $\acval{s}{a}{\wvec{}} \approx \operatorname{q}(s, a)$.
	\begin{itemize}
		\item May use any function differentiable wrt $\wvec{}$ e.g., neural networks.
	\end{itemize}
\item For various reasons, we look at more classical methods: \alert{tile coding with linear function approximation}.
\begin{itemize}
	\item Convert state information to some sparse boolean vector $\operatorname{\mathbf{x}}(s, a)$.
	\item Subdivide space into tiles---each tile is an entry of $\bm{x}$, set to 1 if it contains $s$ and 0 otherwise.
	\item Then define $\acval{s}{a}{\wvec{}}=\wvec{}^{\top} \operatorname{\mathbf{x}}(s, a)$.
	\item Differentiable: $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Reinforcement Learning: How it Works (II)}
% make math easy
\begin{subequations}
	We can keep updating the policy by updating $\wvec{}$ as we receive state ($S_t, S_{t+1}$), action ($A_t, A_{t+1}$), reward information ($R_{t+1}$).
	The \alert{semi-gradient Sarsa algorithm} works here, starting from $\wvec{0}=\bm{0}$:
	\begin{gather}
	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla_{\wvec{}}{\acval{S_t}{A_t}{\wvec{t}}}
	\end{gather}
	Our \alert{policy} is then to choose whichever action has the highest estimated value in the current state, or choose at random with probability $\epsilon$ ($\epsilon$-greedy policy).
\end{subequations}
\end{frame}

\begin{frame}{Reinforcement Learning: An Example (I)}
\begin{figure}
	\centering
	\begin{subfigure}{0.5\linewidth}
		\resizebox{\linewidth}{!}{
			\begin{tikzpicture}
			\node at (0.2,0.4){$s = \begin{pmatrix}
				0.7 \\
				0.3
				\end{pmatrix}$};
			\node at (0,-1.1) {$\bm{x}(s,\cdot) = \begin{Bmatrix}
				T_{1,9}, \\
				T_{2,5}, \\
				T_{\mathit{bias}}
				\end{Bmatrix}$};

			\node at (2.5,-0.5) {
				\begin{tikzpicture}
				\draw[step=0.5cm,color=uofgcobalt,opacity=0.7,shift={(0,0)},label=above:{Tiling 0}] (-0.5,-0.5) grid (1,1);
				\fill[uofgcobalt,opacity=0.5] (0.5,-0.5) rectangle (1,0);
				\node[color=uofgcobalt] at (0,1.15) {\footnotesize Tiling 1};

				\draw[step=0.5cm,color=uofgpumpkin,opacity=0.7,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
				\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
				\node[color=uofgpumpkin] at (0.25,-0.95) {\footnotesize Tiling 2};

				\node[circle, black, draw,
				fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
				%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];

				\draw[->] (-0.25,-0.5)--(-0.25,0.85);
				\draw[->] (-0.25,-0.5)--(1.1,-0.5);

				\node at (1,-0.7) {\footnotesize 1};
				\node at (-0.4,0.75) {\footnotesize 1};
				\node at (-0.35,-0.6) {\footnotesize 0};
				\end{tikzpicture}
			};

		\end{tikzpicture}
	}
	\caption{Tile Coding\label{fig:tilecode-code}}
\end{subfigure}%
\begin{subfigure}{0.5\linewidth}
	\resizebox{\linewidth}{!}{
		\begin{tikzpicture}
		% Top half
		\def\topacs{
			{-0.3,-0.5,-0.1,0.8},
			{0.1,0.1,-0.2,0.3},
			{0.3,0.4,0.2,-0.4}%
		}

		\foreach \line [count=\y] in \topacs {
			\foreach \pix [count=\x] in \line {
				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
					\draw[shift={(-0.7,0)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
				}{
					\pgfmathtruncatemacro\lambda{\pix*100}
					\draw[shift={(-0.7,0)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
				}
				\node[shift={(-0.35,0.175)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
			}
		}

		\draw[xstep=0.7cm,ystep=0.35cm,shift={(0,0)}] (0,-1.06) grid (2.8,0);
		\node[label=left:{$T_{1,9}$},shift={(0,-0.125)}] at (0,0) {}; 
		\node[label=left:{$T_{2,5}$},shift={(0,-0.125)}] at (0,-0.375) {}; 
		\node[label=left:{$T_{\mathit{bias}}$},shift={(0,-0.125)}] at (0,-0.75) {};

		% bottom half
		\def\botacs{
			{0.1,0.0,-0.1,0.7}%
		}

		\foreach \line [count=\y] in \botacs {
			\foreach \pix [count=\x] in \line {
				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
					\draw[shift={(-0.7,-1.275)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
				}{
					\pgfmathtruncatemacro\lambda{\pix*100}
					\draw[shift={(-0.7,-1.275)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
				}
				\node[shift={(-0.35,-1.1)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
			}
		}

		\draw[xstep=0.7cm,ystep=0.35cm,shift={(0,-1.275)}] (0,-0.35) grid (2.8,0);
		\node[label=left:{$\mathit{Total}$},shift={(0,-0.175)}] at (0,-1.275) {};

		\draw [->] (2.45, -1.9) -- (2.45, -1.65);
		\end{tikzpicture}
	}
	\caption{\centering Value Estimation and Action Selection\label{fig:tilecode-select}}
\end{subfigure}

\caption{
	State-action value approximation, when tile coding used.
	\label{fig:tilecode}
}
\end{figure}
\end{frame}

\begin{frame}{Reinforcement Learning: An Example (II)}
\begin{figure}
	\centering
	\resizebox{0.5\linewidth}{!}{
		\begin{tikzpicture}		
		% Top half
		\def\startacs{
			{0.8},
			{0.3},
			{-0.4}%
		}

		\foreach \line [count=\y] in \startacs {
			\foreach \pix [count=\x] in \line {
				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
					\draw[shift={(-0.7,0)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
				}{
					\pgfmathtruncatemacro\lambda{\pix*100}
					\draw[shift={(-0.7,0)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
				}
				\node[shift={(-0.35,0.175)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
			}
		}

		\draw[xstep=0.7cm,ystep=0.35cm,shift={(0,0)}] (0,-1.06) grid (0.7,0);
		\node[label=left:{$T_{1,9}$},shift={(0,-0.125)}] at (0,0) {}; 
		\node[label=left:{$T_{2,5}$},shift={(0,-0.125)}] at (0,-0.375) {}; 
		\node[label=left:{$T_{\mathit{bias}}$},shift={(0,-0.125)}] at (0,-0.75) {};
		\node[label=below:{\footnotesize Action 4},shift={(0.35,-0.125)}] at (0,-0.75) {};
		\node[label=above:{\footnotesize $\wvec{t}$},shift={(0.35,-0.125)}] at (0,0) {};

		\def\finalacs{
			{0.7},
			{0.2},
			{-0.5}%
		}

		\foreach \line [count=\y] in \finalacs {
			\foreach \pix [count=\x] in \line {
				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
					\draw[shift={(2-0.7,0)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
				}{
					\pgfmathtruncatemacro\lambda{\pix*100}
					\draw[shift={(2-0.7,0)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
				}
				\node[shift={(2-0.35,0.175)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
			}
		}

		\draw[xstep=0.7cm,ystep=0.35cm,shift={(2,0)}] (0,-1.06) grid +(0.7,1.06);
		\node[label=below:{\footnotesize Action 4},shift={(2.35,-0.125)}] at (0,-0.75) {};
		\node[label=above:{\footnotesize $\wvec{t+1}$},shift={(0.35,-0.125)}] at (2,0) {};

		\draw[->] (0.9,-0.5) -- node[above] {$+ \alpha \delta_t$} (1.8,-0.5);
		\end{tikzpicture}
	}

	\caption{
		Update step, for observed TD error $\delta_t=-0.2$ and $\alpha=0.5$.
		\label{fig:tilecode-update}
	}
\end{figure}
\end{frame}

\begin{frame}{Reinforcement Learning: An Example (III)}
\begin{figure}
	\centering
		\resizebox{0.9\linewidth}{!}{
			\begin{tikzpicture}
			\node at (0.0,-1.2){$s = \begin{pmatrix}
				0.7 \\
				0.3
				\end{pmatrix}$};
			\node at (0,-2.7) {$\bm{x}(s,\cdot) = \bm{x}'(s,\cdot)+\!\!\!\!+ \bm{x}''(s,\cdot)$};
%			\node at (0,-1.1) {$\bm{x}(s,\cdot) = \begin{Bmatrix}
%				T_{1,9}, \\
%				T_{2,5}, \\
%				T_{\mathit{bias}}
%				\end{Bmatrix}$};

			\node at (5.5,-0.5) {
				\begin{tikzpicture}
				\draw[xstep=1.5cm,ystep=0.25cm,color=uofgcobalt,opacity=0.7,shift={(-0.25,-0.5)},label=above:{Tiling 0}] (0,0) grid (1.5,1.5);
				\fill[uofgcobalt,opacity=0.5,] (-0.25,-0.25) rectangle (1.25,0);
				\node[color=uofgcobalt] at (0,1.15) {\footnotesize Tiling 1};

%				\draw[step=0.5cm,color=uofgpumpkin,opacity=0.7,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
				\draw[xstep=1.5cm,ystep=0.25cm,color=uofgpumpkin,opacity=0.7,shift={(-0.25,-0.83)},label=above:{Tiling 0}] (0,0) grid (1.5,1.5);
%				\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
				\fill[uofgpumpkin,opacity=0.5,] (-0.25,-0.33) rectangle (1.25,-0.08);
				\node[color=uofgpumpkin] at (0.25,-1.05) {\footnotesize Tiling 2};

				\node[circle, black, draw,
				fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
				%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];

				\draw[->] (-0.25,-0.5)--(-0.25,0.85);
				\draw[->] (-0.25,-0.5)--(1.1,-0.5);

				\node at (1,-0.7) {\footnotesize 1};
				\node at (-0.4,0.75) {\footnotesize 1};
				\node at (-0.35,-0.6) {\footnotesize 0};

				\node at (3,0) {$\bm{x}'(s,\cdot) = \begin{Bmatrix}
					T_{1,5}, \\
					T_{2,4}
					\end{Bmatrix}$};
				\end{tikzpicture}
			};

		\node at (5.5,-3) {
			\begin{tikzpicture}
			\draw[xstep=0.75cm,ystep=1.5cm,color=uofglavendar,opacity=0.7,shift={(-0.5,-0.5)},label=above:{Tiling 0}] (0,0) grid (1.5,1.5);
			\fill[uofglavendar,opacity=0.5,] (0.25,-0.5) rectangle (1,1);
			\node[color=uofglavendar] at (0,1.15) {\footnotesize Tiling 3};

			\draw[xstep=0.75cm,ystep=1.5cm,color=uofglawn,opacity=0.7,shift={(-0.1,-0.5)},label=above:{Tiling 0}] (0,0) grid (1.5,1.5);
			\fill[uofglawn,opacity=0.5,] (-0.10,-0.5) rectangle (0.65,1);
			\node[color=uofglawn] at (1,-1) {\footnotesize Tiling 4};

			\node[circle, black, draw,
			fill, radius=0.5pt, inner sep=0pt,minimum size=1.5pt, label=above:{$s$}] at (0.625,-0.125) {};
			%			\filldraw (0.625,-0.125) circle[radius=1.5pt,label=above:{$s$}];

			\draw[->] (-0.25,-0.5)--(-0.25,0.85);
			\draw[->] (-0.25,-0.5)--(1.1,-0.5);

			\node at (1,-0.7) {\footnotesize 1};
			\node at (-0.4,0.75) {\footnotesize 1};
			\node at (-0.35,-0.6) {\footnotesize 0};
			\node at (3,0) {$\bm{x}''(s,\cdot) = \begin{Bmatrix}
				T_{3,2}, \\
				T_{4,1}
				\end{Bmatrix}$};
			\end{tikzpicture}
		};

		\end{tikzpicture}
	}

\caption{
	Other codings.
	\label{fig:tilecode-varied}
}
\end{figure}
\end{frame}

\begin{frame}{Where has RL succeeded in networks?}
	%Link to recent work in data-driven networking/optimisation?
	%
	%Link to that one GMM paper w/ RL communication.

	\begin{itemize}
		\item \alert{Data-driven networking.} Effectively applied to intra-domain routing \footcite{DBLP:conf/hotnets/ValadarskySST17}, task allocation \footcite{DBLP:conf/hotnets/MaoAMK16}, traffic optimisation \footcite{DBLP:conf/sigcomm/ChenL0L18}, video adaptive bitrate selection \footcite{DBLP:conf/sigcomm/MaoNA17} and more, each with general and domain-specific insights.

		\item \alert{In anomaly detection?} Optimising information sharing in distributed statistical model training \footcite{DBLP:conf/paisi/XuSH07}.
	\end{itemize}
\end{frame}

%\begin{frame}{What might we gain from RL in DDoS prevention?}
%content...
%\end{frame}

%\section{Existing work}

\section{The limits of existing work}

\begin{frame}{Multiagent RL for DDoS prevention}
%	Network Model \footcite{DBLP:journals/eaai/MalialisK15}
%	
%	Pushback \footcite{DBLP:journals/ccr/MahajanBFIPS02a}
%	
%	Group reward functions etc.
%
%	Its weaknesses? Strong assumptions about what knowledge the learners really have...

	\begin{itemize}
		\item Yet existing RL-based DDoS defences like MARL\footcite{DBLP:journals/eaai/MalialisK15} haven't been successful---why?
		% How? Violation of ISP-like criterion,

		\item Network model
		\begin{itemize}
			\item Hosts have a fixed probability of being benign/malicious.
			\item Tree overlay topology (split into teams) protecting one server.
			\item Per-team rewards: \alert{coordinated team learning}.
			\item Action: (per-timestep) choose $p\in\{0,10,...,90\}$, s.t. each learner (egress/ingress switch) drops $p\%$ of inbound traffic.
		\end{itemize}

		\item \alert{Actions not granular}---many hosts punished accidentally.
		\item \alert{Action affects congestion-aware traffic disproportionately}---TCP, QUIC \emph{actively slow down} if packets lost.

%		\item Implemented in mininet with Ryu controller, traffic generated by replaying traces.
%		\begin{itemize}
%			\item Packet content unimportant---only need accurate load stats/queuing behaviour.
%			\item Alternate model featuring live HTTP traffic.
%		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Multiagent RL (MARL) for DDoS prevention}
\begin{columns}
	\begin{column}{0.45\linewidth}
		\begin{itemize}
			\item Algorithm: Semi-gradient Sarsa, linear fn approx, $\epsilon$-greedy selection.
			\item Actions: Drop [0, 10, ... 90]\si{\percent} upstream traffic.
			\item State: load vectors of agent and parents ($\mathbb{R}^4$) $\rightarrow$ tile-coded (fixed-weight binary vector).
			\item Rewards: \num{-1} if $\mathit{load} > U_s$, else fraction of surviving legit traffic.
		\end{itemize}
	\end{column}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\centering
			\resizebox{\linewidth}{!}{
				\begin{tikzpicture}[
				texts/.style = {text=black},
				labeltexts/.style = {text=uofgsandstone},
				treeline/.style = {draw=uofgburgundy},
				treenode/.style = {texts, circle, centered, fill=white, treeline},
				load/.style = {fill=uofgcobalt},
				external/.style = {fill=uofgrust},
				hideline/.style = {draw=uofgsandstone!40!white},
				hidenode/.style = {treenode, hideline},
				grow'=right
				]
				\node[treenode, label={[texts]above:Server}] (root) {}
				child [treeline] { node [treenode, load, label={[texts]above:Core}] (sswitch) {}
					child [treeline] { node [treenode, load, label={[texts]above:Leader}] (teaml) {} 
						child [treeline] { node [treenode, load, label={[texts]above:Intermediate}] (inter) {}
							child [treeline] { node [treenode, load, label={[texts]above:Agent/Egress}] (agent) {}
								child [treeline] { node [treenode, external] (extern) {}
									child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
									child [hideline] { node [hidenode] (endhost) {} }
								}
							}
							child [hideline] { node [hidenode] (endagent) {} }
						}
						child [hideline] { node [hidenode] (endinter) {} }
					}
					child [hideline] { node [hidenode] (endteaml) {} }
					edge from parent
					node[below, labeltexts] {$U_s$}
				};

				%\draw[-] (teaml) -- (endteaml);
				\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
				\node [labeltexts, right = -0.1cm of kdots] {$k$};
				\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
				\node [labeltexts, right = -0.1cm of ldots] {$l$};
				\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
				\node [labeltexts, right = -0.1cm of mdots] {$m$};
				\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
				\node [labeltexts, right = -0.1cm of ndots] {$n$};
				\end{tikzpicture}
			}
			\caption{
				Network topology diagram.
				Red nodes are external, blue nodes feature in the state vector.
				Any packet drop occurs when forwarding packets from an egress switch to its parent (intermediate) switch.
			}
		\end{figure}
	\end{column}
\end{columns}
\end{frame}

%\begin{frame}{The case for finer granularity}
%	\begin{columns}
%	\begin{column}{0.45\linewidth}
%		\begin{itemize}
%			\item Learner/host ratio (action/host ratio) affects host QoS.
%			
%			\item Reduced service guarantees by nature of \emph{pushback} model.
%			\begin{itemize}
%				\item \alert{Worse with good-faith TCP congestion avoidance}.
%			\end{itemize}
%			%This is exacerbated by TCP congestion avoidance---legit hosts will be punished far more severely, bad actors don't care!
%			
%%			\item More granular $\implies$ should focus on flow stats, \alert{not aggregates} like now! %Alongside this higher-granularity view, reformulation to include flow characteristics in the decision-making process.
%		\end{itemize}
%	\end{column}
%	\begin{column}{0.5\linewidth}
%		\begin{figure}
%			\includegraphics[width=\linewidth]{../plots/online-varyN-binary-pres.pdf}
%			\caption{Service quality decreases as actions become less granular (affecting $n$ hosts at once).}
%		\end{figure}
%	\end{column}
%	\end{columns}
%\end{frame}
%
%\begin{frame}{The case for finer granularity}
%\centering
%\includegraphics[width=0.8\linewidth]{../plots/online-varyN-binary-pres.pdf}
%\end{frame}

\begin{frame}{Why care about congestion-aware traffic?}
\begin{columns}
	\begin{column}{0.5\linewidth}
		\begin{figure}
			\includegraphics[width=\linewidth]{../code/caida-stats/plots/ca-monthly-pres-bytes.pdf}
			\caption{Proportion of traffic volume which is congestion aware (TCP or QUIC) (CAIDA 2018 passive trace dataset).}
		\end{figure}
	\end{column}
	\begin{column}{0.45\linewidth}
		\begin{itemize}
			\item 9 years ago, TCP carried \SIrange{89}{98}{\percent} of data by volume\footnotemark.
			\item Nowadays, CAIDA traces suggest \SIrange{75}{84}{\percent} over both directions of a backbone link. If we include QUIC, \SIrange{76}{85}{\percent}.
			\item QUIC is a non-negligible fraction (peak \SI{4.5}{\percent} Dir A in Jul/Sep), and we can't examine anything below headers.
		\end{itemize}
	\end{column}
\end{columns}
\footcitetext{DBLP:conf/saint/ZhangDJC09}
\end{frame}

\begin{frame}{Why care about congestion-aware traffic?}
\centering
\includegraphics[width=0.8\linewidth]{../code/caida-stats/plots/ca-monthly-pres-bytes.pdf}
\end{frame}

%\begin{frame}{And the replication reveals...}	
%\begin{itemize}
%	\item Network topology has no basis in reality---admitted by its \emph{own} source work \footcite{DBLP:journals/ccr/MahajanBFIPS02a}.
%	
%	\item Action granularity causes more collateral damage than we'd like...
%	
%	\item ...and the picture is worse still for legitimate TCP flows.
%	
%	\item Reward function needs a priori knowledge/reliable estimates to learn online.
%	
%	\item \alert{But on the plus side, action computation is fast: \SIrange{80}{100}{\micro\second}.}
%\end{itemize}
%\end{frame}
%
%\begin{frame}{How can we use these observations? (The Immediate Future)}
%\begin{itemize}
%	\item \textbf{\alert{Why not take actions on a per-flow basis?}}
%	\begin{itemize}
%		\item Solves the granularity issues by construction.
%		\item Allows different treatment by flow features (i.e., protocol).
%	\end{itemize}
%	\item Need to rethink state space: more costly computation, but we have room to work in.
%	\begin{itemize}
%		\item We need any additions to be justified beyond just ``more data'', since \alert{changes affect training time and execution time}.
%	\end{itemize}
%	\item How do we select flows to act upon?
%\end{itemize}
%\end{frame}

\section{How do we adapt MARL to match the realities of network traffic?}

\begin{frame}{Adapting MARL}
\begin{itemize}
	\item Prior work with these techniques has two key flaws:
	\begin{itemize}
		\item Each action can (and will!) map to many flows (particularly in real networks), increasing the \alert{likelihood} of collateral damage.
		\item The majority of internet traffic is congestion-aware, increasing the \alert{cost} of collateral damage.
	\end{itemize}
	\item Naturally, we aim then to take decisions \emph{per-flow}.
\end{itemize}
\end{frame}

\begin{frame}{Network Model}
\begin{itemize}
	\item \textbf{Topology}: Identical to MARL's ``online'' experiments (single server, $k=2, \ell=3, m=2, \operatorname{P}_{\mathit{malicious}}=0.4$), but varying $n \in \{2, 4, 8, 16\}$.
	\begin{itemize}
		\item \numrange{24}{192} hosts, causing expected bandwidth load \SIrange{2.03}{2.18}{$\! \times U_s$}.
	\end{itemize}
	\item \textbf{TCP Traffic}: HTTP traffic (\emph{nginx} and \emph{libcurl}), random requests (long and short flows), TCP Cubic.
	\item \textbf{UDP Traffic (Malicious/Benign)}: \emph{hping3}, MTU-size packets (mapping to an amplification attack).
\end{itemize}
\end{frame}

\begin{frame}{An effective state space}
\begin{itemize}
	\item \textbf{Global state} The existing state space (load measurements).
	\item \textbf{Constants} Source IP.
	\item \textbf{Ongoing observations} Flow duration, flow size, packets in/out, correspondence ratio, mean inter-arrival time.
	\item \textbf{Per-timestep (window) observations} Last action taken, $\mathbf{\Delta}$ send/receive rate, mean packet size, packet in/out count.
\end{itemize}

	Why exclude port numbers, payload features? \alert{QUIC} and similar protocols \emph{will become ubiquitous}. \alert{This constraint is a form of protocol-agnostic future-proofing}.

%	Tile code each feature individually.
\end{frame}

\begin{frame}{Feature Performance---UDP}
\begin{figure}
	\centering
	%	\begin{subfigure}{0.32\linewidth}
	%		\includegraphics[width=\linewidth]{../plots/ftprep-good}
	%	\end{subfigure}
	%	\begin{subfigure}{0.32\linewidth}
	%		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
	%	\end{subfigure}
	%	\begin{subfigure}{0.32\linewidth}
	%		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
	%	\end{subfigure}
	\includegraphics[width=0.8\linewidth]{../plots/ftprep-box}
%	\vspace{-1.2cm}
	\caption{
		Learned performance of Marl++ agents when benign traffic is UDP-like, using only a single feature as a basis for decisions. Uncap.
		\label{fig:udp-feature-plots}
	}
\end{figure}
\end{frame}

\begin{frame}{Feature Performance---TCP}
\begin{figure}
	\centering
	%	\begin{subfigure}{0.32\linewidth}
	%		\includegraphics[width=\linewidth]{../plots/ftprep-good}
	%	\end{subfigure}
	%	\begin{subfigure}{0.32\linewidth}
	%		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
	%	\end{subfigure}
	%	\begin{subfigure}{0.32\linewidth}
	%		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
	%	\end{subfigure}
	\includegraphics[width=0.8\linewidth]{../plots/ftprep-tcp-cap-box}
%	\vspace{-1.2cm}
	\caption{
		Learned performance of Marl++ agents when benign traffic is TCP-like, using only a single feature as a basis for decisions.
		\label{fig:udp-feature-plots-cap}
	}
\end{figure}
\end{frame}

\begin{frame}{Implementation considerations}
	\begin{itemize}
		\item \alert{Timed random sequential decisions}
		\begin{itemize}
			\item Multiple actions per timestep---perform up to a deadline of \SI{1}{\milli\second} as suggested\footcite{DBLP:conf/sigcomm/ChenL0L18}.
			\item Carry over undone work to next timestep, fuse window measurements.
			\item Optionally drop after a few rounds to prevent starvation.
		\end{itemize}
		\item Reward functions modified for direction awareness.
		\item Optional model updates from other agents' traces/experience.
		\item Deliberately avoiding neural networks; training time, execution cost, energy usage, hardware...
		\item I'll refer to all these advancements with MARL-like actions as \alert{MARL++}.
	\end{itemize}
\end{frame}

\begin{frame}{Another look at actions---SPF}
\begin{itemize}
	\item Behavioural knowledge---SPIFFY\footcite{DBLP:conf/ndss/KangGS16} realised that \alert{volume-based attack flows can't scale up if they have their maximum bandwidth increased}.
	\item Can we constrain our model to speed up training/reduce variance?
	\item Limit $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, can only increase/decrease one step at a time.
	\begin{itemize}
		\item All flows start at $p = 0.05$.
	\end{itemize}
	\item \alert{3 actions}---\emph{punish, reward, no-op}.
	\item Now need us to choose $\gamma$ to \emph{plan ahead}: 0.8 an effective choice.
\end{itemize}
\end{frame}

\section{Evaluation}
\begin{frame}{UDP traffic performance}
\begin{columns}
	\begin{column}{0.7\linewidth}
	\begin{figure}
		\centering
		\includegraphics[width=0.95\linewidth]{../plots/udp-box}

		\caption{
			Online performance for UDP-like benign traffic.
			\label{fig:udp-box}
		}
	\end{figure}
	\end{column}
	\begin{column}{0.3\linewidth}
		\begin{itemize}
			\item Marl++ outperforms Marl---a success!
			\item SPF has low variance as expected, but performs poorly until large $n$.
		\end{itemize}
	\end{column}
\end{columns}
\end{frame}

\begin{frame}{TCP traffic performance}
\begin{columns}
	\begin{column}{0.3\linewidth}
		\begin{itemize}
			\item Harder to see statistically significant advantage until large $n$.
			\item SPF has low variance as expected, but performs poorly until large $n$.
		\end{itemize}
	\end{column}
	\begin{column}{0.7\linewidth}
		\begin{figure}
			\centering
			\includegraphics[width=0.95\linewidth]{../plots/tcp-box}

			\caption{
				Online performance for UDP-like benign traffic.
				\label{fig:tcp-box}
			}
		\end{figure}
	\end{column}
\end{columns}
\end{frame}

\begin{frame}{TCP traffic (n=16)---training in detail}
	\begin{figure}
		\centering
		\includegraphics[width=0.8\linewidth]{../plots/tcp-16-single}

		\caption{
			Online performance of standard and single-agent models with $n=16$ hosts per egress point and TCP-like benign traffic.
			At this level of host density, SPF reaches its peak performance sooner and is considerably more consistent throughout the episode.
			Marl++ takes longer to reach its most effective policy, and achieves visibly better peak performance than the other standard models.
			With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
			\label{fig:tcp-16}
		}
	\end{figure}
\end{frame}

\begin{frame}{All experiments}
	\begin{table}
		\centering
		\resizebox{\linewidth}{!}{\expandableinput ../tables/big-avg-reward.tex}
		\caption{Average reward for combinations of model, host density and traffic class.\label{tab:av-vals}}
	\end{table}
\end{frame}

\begin{frame}{Main take-homes}
\begin{itemize}
	\item The flow-level focus \alert{improves over past RL work in both problem cases}.
	\begin{itemize}
		\item Still a long way to go before it's competitive with established solutions.
		\item Likely converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
	\end{itemize}
	\item SPF relies on higher volumes of data to perform well (high $n$, experience sharing)---has potential to be useful if this need can be met.
	\begin{itemize}
		\item Why?
	\end{itemize}
\end{itemize}
\end{frame}

\section{The work yet undone}

\begin{frame}{Other environments}
\begin{itemize}
	\item Expansion to other topologies.
	\begin{itemize}
		\item Likely not viable or sensible to map this overlay topology to most networks: many sources/sinks, multiple paths make global state useless.
		\item All we need to change is global state.
	\end{itemize}
	\item So flip the problem on its head...
	\item This will enable direct comparison against other approaches.
\end{itemize}
\end{frame}

\begin{frame}{Other attacks}
\begin{itemize}
	\item More ``old-school'' attacks like TCP SYN/ACK floods, application-level attacks against emulated software...
	\item In AS-like networks/topologies, we can examine \alert{Link Flooding Attacks}.
	\item Crucially, need several of these to evaluate against attackers with \alert{evolving strategies} (and show generalisation beyond amplification attacks).
\end{itemize}
\end{frame}

\begin{frame}{The Near Future}
\begin{itemize}
	\item Reward functions without dependence on ahead-of-time knowledge.
	\begin{itemize}
		\item I.e., for certain distributions of communication we might want to maximise link utilisation in both directions.
		\item Active measurement---canary flows etc.
	\end{itemize}

\item Deriving normal model behaviour from traces.
\begin{itemize}
	\item We only need to simulate specific behaviour to test these enhancements, but that can become more representative.
	\item But we do need more diversity of services in the normal model.
\end{itemize}

\item Investigate other RL algorithms.
\begin{itemize}
	\item ``Deep learning'' probably not feasible.
	\item \alert{$\operatorname{TD}(\lambda)$, actor-critic methods}.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{The Far Future}
\begin{itemize}
	\item Other problems.
	\begin{itemize}
		\item New action spaces, careful consideration.
	\end{itemize}
\item Adversarial capabilities---evasion and poisoning attacks.
\item Knowledge-sharing between agents: cost-modelling and optimisation.
\item Test deployments in real networks.
\end{itemize}
\end{frame}

\begin{frame}[standout]{Conclusion}
	We've looked at...
	\begin{itemize}
		\item A quick introduction to RL, and its \alert{importance to future networks}.
		\item A recent `direct control' approach to intrusion prevention, and \alert{its significant weaknesses}.
		\item A \alert{new interaction model and state-space}, which can remedy these weaknesses.
		\item The \alert{next steps} towards proper generalisation.
	\end{itemize}

	\alert{Questions?}
\end{frame}

\appendix

\begin{frame}[allowframebreaks]{References}
\printbibliography[heading=none]
\end{frame}

\end{document}

### 2019-02-05 paper/paper.tex
New talk...

@@
 ?? VERIFY AND RESEARCH THAT: REVIEWER 1 BELIEVES OTHERWISE.
-The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.
+The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots such attacks are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
 The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
@@
 for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions, and the probability $p \in (0, 1]$ that a packet is lost.
+?? TCP Cubic WAS used on the lab machines. We do still see Mathis-like behaviour with TCP Cubic, which is a significant observation.
 ?? Does Mathis still hold? We observed similar trends empirically, but perhaps not the \emph{exact} curve or falloff behaviour...
@@
 Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
+??
+Flaws?
+what can't it do that we can do better?
+Long times to compute TBE routes on real networks, low expansion factor in real network can require more ``rounds'' of filtering.
+(``takes only'' \SI{14}{\second} for Cogent network?
+gross.)
 

### 2019-02-06 glass-talk/glass-talk.tex
Finished presentation

@@
 	\end{gather}
+	given $\alpha, \gamma \in [0,1]$.
+
 	Our \alert{policy} is then to choose whichever action has the highest estimated value in the current state, or choose at random with probability $\epsilon$ ($\epsilon$-greedy policy).
@@
 
-\begin{frame}{Multiagent RL for DDoS prevention}
+\begin{frame}{Multiagent RL (MARL) for DDoS prevention}
 %	Network Model \footcite{DBLP:journals/eaai/MalialisK15}
@@
 	\begin{itemize}
-		\item Yet existing RL-based DDoS defences like MARL\footcite{DBLP:journals/eaai/MalialisK15} haven't been successful---why?
+		\item There are existing RL-based DDoS defences like MARL\footcite{DBLP:journals/eaai/MalialisK15}.
 		% How? Violation of ISP-like criterion,
@@
 		\end{itemize}
-
-		\item \alert{Actions not granular}---many hosts punished accidentally.
-		\item \alert{Action affects congestion-aware traffic disproportionately}---TCP, QUIC \emph{actively slow down} if packets lost.
 
@@
 			\item Actions: Drop [0, 10, ... 90]\si{\percent} upstream traffic.
-			\item State: load vectors of agent and parents ($\mathbb{R}^4$) $\rightarrow$ tile-coded (fixed-weight binary vector).
+			\item State: load vectors of agent and parents ($\mathbb{R}^4$) $\rightarrow$ tile-coded.
 			\item Rewards: \num{-1} if $\mathit{load} > U_s$, else fraction of surviving legit traffic.
@@
 \end{columns}
+\end{frame}
+
+\begin{frame}{Multiagent RL (MARL) for DDoS prevention}
+%	Network Model \footcite{DBLP:journals/eaai/MalialisK15}
+%	
+%	Pushback \footcite{DBLP:journals/ccr/MahajanBFIPS02a}
+%	
+%	Group reward functions etc.
+%
+%	Its weaknesses? Strong assumptions about what knowledge the learners really have...
+
+\begin{itemize}
+	\item Yet existing RL-based DDoS defences like MARL haven't been successful---why?
+
+	\item \alert{Actions not granular}---many hosts punished accidentally.
+	\item \alert{Action affects congestion-aware traffic disproportionately}---TCP, QUIC \emph{actively slow down} if packets lost.
+\end{itemize}
 \end{frame}
@@
 	\end{itemize}
-	\item Naturally, we aim then to take decisions \emph{per-flow}.
+	\item Naturally, we aim then to take decisions \emph{per-flow} (or per-source, to be precise).
 \end{itemize}
@@
 	\centering
-	%	\begin{subfigure}{0.32\linewidth}
-	%		\includegraphics[width=\linewidth]{../plots/ftprep-good}
-	%	\end{subfigure}
-	%	\begin{subfigure}{0.32\linewidth}
-	%		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
-	%	\end{subfigure}
-	%	\begin{subfigure}{0.32\linewidth}
-	%		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
-	%	\end{subfigure}
-	\includegraphics[width=0.8\linewidth]{../plots/ftprep-box}
+	\includegraphics[width=0.8\linewidth]{../plots/ftprep-box-pres}
 %	\vspace{-1.2cm}
@@
 	\centering
-	%	\begin{subfigure}{0.32\linewidth}
-	%		\includegraphics[width=\linewidth]{../plots/ftprep-good}
-	%	\end{subfigure}
-	%	\begin{subfigure}{0.32\linewidth}
-	%		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
-	%	\end{subfigure}
-	%	\begin{subfigure}{0.32\linewidth}
-	%		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
-	%	\end{subfigure}
-	\includegraphics[width=0.8\linewidth]{../plots/ftprep-tcp-cap-box}
+	\includegraphics[width=0.8\linewidth]{../plots/ftprep-tcp-cap-box-pres}
 %	\vspace{-1.2cm}
@@
 		\centering
-		\includegraphics[width=0.95\linewidth]{../plots/udp-box}
+		\includegraphics[width=0.95\linewidth]{../plots/udp-box-pres}
 
@@
 			\item Harder to see statistically significant advantage until large $n$.
-			\item SPF has low variance as expected, but performs poorly until large $n$.
+			\item All have very high variance (due to poor actions on congestion-aware protocols), but SPF narrows it down.
 		\end{itemize}
@@
 			\centering
-			\includegraphics[width=0.95\linewidth]{../plots/tcp-box}
+			\includegraphics[width=0.95\linewidth]{../plots/tcp-box-pres}
 
@@
 		\centering
-		\includegraphics[width=0.8\linewidth]{../plots/tcp-16-single}
+		\includegraphics[width=0.8\linewidth]{../plots/tcp-16-single-pres}
 
@@
 		\item Likely not viable or sensible to map this overlay topology to most networks: many sources/sinks, multiple paths make global state useless.
-		\item All we need to change is global state.
+		\item \alert{Then what we need to change is global state!}
 	\end{itemize}
-	\item So flip the problem on its head...
+	\item So flip the problem on its head: even in multipath networks \emph{we know what path a flow will follow} (i.e., via ECMP).
+	\begin{itemize}
+		\item So, draw global state from the \emph{path}: endpoints + tertiles (still in $\mathbb{R}^{4}$).
+		\item Cost: possibly more network measurements, concept of teams no longer valid (so lose info from reward functions).
+	\end{itemize}
 	\item This will enable direct comparison against other approaches.
@@
 	\item Crucially, need several of these to evaluate against attackers with \alert{evolving strategies} (and show generalisation beyond amplification attacks).
+	\begin{itemize}
+		\item Evolving attackers are \emph{the norm}\footcite{DBLP:conf/spw/KangGS16}, though few works detail \emph{how} they evolve.
+	\end{itemize}
 \end{itemize}

### 2019-03-22 code/opus-voip-traffic/src/lib.rs
RTP-like (Opus) traffic generator v1

@@ -1,0 +1,167 @@
+#[macro_use]
+extern crate log;
+
+mod config;
+mod trace;
+
+pub use config::Config;
+
+use byteorder::{
+	NetworkEndian,
+	ReadBytesExt,
+	WriteBytesExt,
+};
+use net2::{
+	unix::UnixUdpBuilderExt,
+	UdpBuilder,
+};
+use rand::{
+	distributions::{
+		Uniform,
+	},
+	prelude::*,
+};
+use std::{
+	collections::HashMap,
+	io::{
+		Read,
+		Result as IoResult,
+	},
+	net::{
+		UdpSocket,
+		SocketAddr,
+	},
+	thread,
+	time::{
+		Duration,
+		Instant,
+	},
+};
+use trace::*;
+
+fn make_udp_socket(port: u16, non_block: bool) -> IoResult<UdpSocket> {
+	let out = UdpBuilder::new_v4()?;
+	
+	out.reuse_address(true)?;
+
+	if !cfg!(windows) {
+		out.reuse_port(true)?;
+	}
+
+	let out = out.bind(("127.0.0.1", port))?;
+	out.set_nonblocking(non_block)?;
+
+	Ok(out)
+}
+
+pub fn client(config: Config) {
+	let ts = trace::read_traces();
+
+	crossbeam::scope(|s| {
+		for i in 0..config.thread_count {
+			s.spawn(|_| {
+				inner_client(&config, &ts);
+			});
+		}
+	}).unwrap();
+}
+
+const CMAC_BYTES: usize = 16;
+const RTP_BYTES: usize = 12;
+
+fn inner_client(config: &Config, ts: &Vec<Trace>) {
+	let mut rng = thread_rng();
+
+	let draw = Uniform::new(0, ts.len());
+
+	let port_distrib = Uniform::new(30000, 54000);
+	let port = port_distrib.sample(&mut rng);
+
+	let end = if config.randomise_duration {
+		let time_distrib = Uniform::new(
+			config.duration_lb,
+			config.duration_ub
+				.expect("No upper bound set: cannot randomise call time."),
+		);
+		let port = time_distrib.sample(&mut rng);
+	} else {
+		config.duration_ub;
+	};
+
+	let el = &ts[draw.sample(&mut rng)];
+	
+	let socket = make_udp_socket(port, true).unwrap();
+
+	println!("Listening on port {:?}", port);
+
[…]

### 2019-03-22 code/opus-voip-traffic/src/trace/mod.rs

@@ -1,0 +1,43 @@
+mod packet_format;
+
+use rayon::prelude::*;
+use std::{
+	fmt::Debug,
+	fs::{
+		self,
+		File,
+	},
+	path::PathBuf,
+};
+
+pub(crate) use packet_format::*;
+
+const TRACE_DIR: &str = "traces/";
+
+pub type Trace = Vec<PacketChainLink>;
+
+pub fn read_traces() -> Vec<Trace> {
+	let file_entries: Vec<PathBuf> = fs::read_dir(TRACE_DIR)
+		.expect("Couldn't read files in trace directory...")
+		.filter_map(|x| if let Ok(x) = x {
+			Some(x.path())
+		} else {
+			warn!("Couldn't read entry: {:?}", x);
+			None
+		}).collect();
+
+	file_entries.par_iter()
+		.map(File::open)
+		.filter_map(warn_and_unpack)
+		.map(bincode::deserialize_from)
+		.filter_map(warn_and_unpack)
+		.filter(|x: &Trace| !x.is_empty())
+		.collect()
+}
+
+fn warn_and_unpack<T, E: Debug>(input: Result<T, E>) -> Option<T> {
+	input.map_err(|why| {
+		warn!("Couldn't parse entry: {:?}", why);
+		why
+	}).ok()
+}

### 2019-03-22 code/traffic-host/src/config.rs

@@ -1,32 +1,32 @@
 use std::{
-    borrow::Cow,
-    time::Duration,
+	borrow::Cow,
+	time::Duration,
 };
 
 #[derive(Debug)]
 pub struct Config<'a> {
 	/// Path to the dependency list structure.
-    pub dep_list_dir: Cow<'a, str>,
+	pub dep_list_dir: Cow<'a, str>,
 
-    /// Directory to use as basis for requests.
-    pub http_dir: Cow<'a, str>,
+	/// Directory to use as basis for requests.
+	pub http_dir: Cow<'a, str>,
 
-    /// Maximum download rate, in bytes/s.
-    pub max_down: u64,
+	/// Maximum download rate, in bytes/s.
+	pub max_down: u64,
 
-    /// Maximum upload rate, in bytes/s.
-    pub max_up: u64,
+	/// Maximum upload rate, in bytes/s.
+	pub max_up: u64,
 
-    /// Random request mode enabled.
-    pub randomise: bool,
+	/// Random request mode enabled.
+	pub randomise: bool,
 
-    /// Base url for the domain to make requests from.
-    pub url: Cow<'a, str>,
+	/// Base url for the domain to make requests from.
+	pub url: Cow<'a, str>,
 
-    /// Amount of requests to make (from a root object) before termination.
-    pub requests: Option<u64>,
+	/// Amount of requests to make (from a root object) before termination.
+	pub requests: Option<u64>,
 
-    /// Time to wait between requests (in ms)
-    pub wait_ms: Duration,
+	/// Time to wait between requests (in ms)
+	pub wait_ms: Duration,
 }
 

### 2019-04-28 code/marl/beast-validation.py
Reasonable (untested) first attempt at randomised single flow-feature isolation
* With a probability (modifier of current value of epsilon at the agent in question), RL agents can/will now make their decisions based upon a subset of the statespace for the given flow. This SHOULD work to reduce noise substantially, idea courtesy of Cristian Urlea back in February.

* Note: This is very likely to be BROKEN when combned with multiple subactors. I need to re-examine that particular interaction, and store/compute narrowings for each one. Either that, or limit computation to a single subactor when t

@@ -78,7 +78,7 @@ if __name__ == "__main__":
 		"alpha": 0.05,
 		"epsilon": 0.2,
 		"discount": 0.0,
-                "dt": 0.05,
+		"dt": 0.05,
 
 		"explore_episodes": 0.8,
 		"episodes": 10,
@@ -93,7 +93,8 @@ if __name__ == "__main__":
 		"trs_maxtime": 0.001,
 
 		"split_codings": True,
-		"feature_max": 18,
+		# "feature_max": 18,
+		"feature_max": 20,
 
 	}
 

### 2019-04-28 code/marl/marl.py

@@ -68,10 +68,11 @@ def marlExperiment(
 		algo = "sarsa",
 		trace_decay = 0.0,
 		trace_threshold = 0.0001,
+		use_path_measurements = True,
 
 		model = "tcpreplay",
 		submodel = None,
-                rescale_opus = False,
+		rescale_opus = False,
 
 		use_controller = False,
 		moralise_ips = True,
@@ -111,7 +112,7 @@ def marlExperiment(
 		bw_mon_socketed = False,
 		unix_sock = True,
 		print_times = False,
-                # Can set to "no", "udp", "always"
+		# Can set to "no", "udp", "always"
 		prevent_smart_switch_recording = "udp",
 		# FIXME: these two flags ae incompatible
 		record_times = False,
@@ -132,6 +133,10 @@ def marlExperiment(
 		split_codings = False,
 		extra_codings = [],
 		feature_max = 12,
+		combine_with_last_action = [12, 13, 14, 15, 16, 17, 18, 19],
+		strip_last_action = True,
+		explore_feature_isolation_modifier = 1.0,
+		explore_feature_isloation_duration = 5,
 
 		trs_maxtime = None,
 		reward_band = 1.0,
@@ -258,13 +263,13 @@ def marlExperiment(
 		# The stats observed show that flows occupy (in expectation)
 		# ~52kbps (median 49.906) with the below settings. If flows were constantly submitting,
 		# we'd see ~85kbps (median 97) accounting for a target 96kbps (w/ some per-server deviation)
-                divisor = 1.0
-                if rescale_opus:
-                    ub = host_range[1]
-                    # lerp between some observations at differing n
-                    pt = float(max(0, ub - 2)) / 14.0
-                    divisor = 0.6 + pt * (0.45 - 0.6)
-                    print "rescaled to", divisor
+		divisor = 1.0
+		if rescale_opus:
+			ub = host_range[1]
+			# lerp between some observations at differing n
+			pt = float(max(0, ub - 2)) / 14.0
+			divisor = 0.6 + pt * (0.45 - 0.6)
+			print "rescaled to", divisor
 		flow_bw = 52.39456 / (divisor * 1024.0) # to mbps.
 		subclient_count = int(math.ceil(max(1.0, bw / flow_bw)))
 		print subclient_count
@@ -278,8 +283,8 @@ def marlExperiment(
 			"-b", "../opus-voip-traffic",
 			"--ip-strategy", "even",
[…]

### 2019-04-28 code/marl/sarsa.py

@@ -21,6 +21,7 @@ class SarsaLearner:
 				trace_decay=0.0, trace_threshold=0.0001,
 				broken_math=False,
 				rescale_alpha=1.0,
+				increase_fixed_math_alpha_if_narrowed=True,
 				AcTrans=MarlMachine):
 		state_range = [
 			[0 for i in xrange(vec_size)] + extended_mins,
@@ -30,6 +31,8 @@ class SarsaLearner:
 		tc_indices = tc_indices if tc_indices is not None else [np.arange(vec_size)]
 		ntiles = [tile_c for _ in tc_indices]
 		ntilings = [tilings_c for _ in tc_indices]
+		self.ntilings = ntilings
+		self.tiling_set_count = len(tc_indices)
 
 		#print "tc has been configured with indices", tc_indices, "tiles", ntiles, "tilings", ntilings
 
@@ -42,6 +45,8 @@ class SarsaLearner:
 			state_range = state_range,
 			rnd_stream = np.random.RandomState(),
 		)
+		# We need access to this again if we perform feature fixation.
+		self.single_learn_rate = learn_rate
 
 		if rescale_alpha is not None and not broken_math:
 			learn_rate = (learn_rate * rescale_alpha) / float(sum(ntilings))
@@ -60,6 +65,7 @@ class SarsaLearner:
 
 		self._argmax_in_dt = False
 		self._wipe_trace_if_not_argmax = False
+		self.alpha_mod_fixed_math = increase_fixed_math_alpha_if_narrowed
 
 		self.trace_decay = trace_decay
 		self.trace_threshold = trace_threshold
@@ -78,19 +84,48 @@ class SarsaLearner:
 		self._ensure_state_vals_exist(state)
 		return [self.values[tile] for tile in state]
 
-	def _update_state_values(self, state, action, values):
+	def _update_state_values(self, state, action, values, narrowing):
 		self._ensure_state_vals_exist(state)
-		for tile, value in zip(state, values):
+
+		if narrowing is None:
+			narrowing = xrange(len(state))
+
+		# Why does this work?
+		# Action computed using only these values, we then compute the delta
+		# (using ALL values) and only propagate forward changes on the relevant tiles.
+		for i in narrowing:
+			tile = state[i]
+			value = value[i]
 			self.values[tile][action] = value
 
-	def select_action(self, state):
+	def _compute_relevant_narrowed_tilings(self, narrowing):
[…]

### 2019-04-28 code/marl/marl.py
Make narrowings apply per-subactor

@@ -1836,20 +1836,14 @@ def marlExperiment(
 									dm = [i] if record_deltas_in_times else None
 										
 									dat = flow_traces[ip]
-									(st, z) = dat[0][s_ac_num]
+									(st, z, narrowing_in_use) = dat[0][s_ac_num]
 									machine = dat[2]
-
-									# narrowing is on flow_trace[4]
-									# Explore (by fixating on idividual features) as a function of
-									# the agent's current epsilon...
-									narrowing_in_use = dat[4]
 
 									# handful of steps:
 									#  only draw new if none
 									#  decrement uses
 									#  don't use in update step if just generated
 									#  don't use in action step if about to expire
-									new_narrowing = False
 									allow_update_narrowing = False
 									allow_action_narrowing = False
 
@@ -1863,7 +1857,6 @@ def marlExperiment(
 												explore_feature_isolation_duration,
 												np.random.choice(s.tiling_set_count),
 											]
-											new_narrowing = True
 											allow_action_narrowing = True
 									else:
 										narrowing_in_use[0] -= 1
@@ -1889,13 +1882,13 @@ def marlExperiment(
 									narrowing_in_use = None
 
 								ac_vals += new_vals
-								substates.append((state, z_vec))
+								substates.append((state, z_vec, narrowing_in_use))
 
 								last_sarsa = s
 
 							l_action = last_sarsa.select_action_from_vals(ac_vals)
 							machine.move(l_action)
-							flow_traces[ip] = (substates, l_action, machine, z_vec, narrowing_in_use)
+							flow_traces[ip] = (substates, l_action, machine, z_vec)
 
 							if need_decay:
 								for (s, _) in subactors:

### 2019-05-03 paper/paper.tex
Reasonable (untested) first attempt at randomised single flow-feature isolation
* With a probability (modifier of current value of epsilon at the agent in question), RL agents can/will now make their decisions based upon a subset of the statespace for the given flow. This SHOULD work to reduce noise substantially, idea courtesy of Cristian Urlea back in February.

* Note: This is very likely to be BROKEN when combned with multiple subactors. I need to re-examine that particular interaction, and store/compute narrowings for each one. Either that, or limit computation to a single subactor when t

@@
 
+TCP-Cubic still has a dependence on the $1/p^{0.75}$ -- slightly slower falloff, but still significant \cite{rfc8312}.
+
 TCP traffic as it appears here is generated according to \Cref{sec:a-new-normal}.
@@
 Simpler decision boundaries reduce the risk of overfitting and of unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.
+
+??
+What other choices did we make?
+Taking actions with (and updating) fewer tiles according to probability $\epsilon$.
+Updating all tiles' action value estimates separately to produce different update targets.
 

### 2019-05-03 code/marl/marl.py

@@ -796,7 +796,11 @@ def marlExperiment(
 			cmd_list = []
 			# remove if groups fixed
 			os = p_drop_num # = None
-			msg = internal_choose_group(int(ac_prob * 10), target_ip=target_ip, force_old=os)
+			if target_ip is not None:
+				(src, dst) = target_ip
+				msg = internal_choose_group(int(ac_prob * 10), ip=dst, target_ip=src, force_old=os, subnet="255.255.255.255")
+			else:
+				msg = internal_choose_group(int(ac_prob * 10), target_ip=target_ip, force_old=os)
 
 		if alive:
 			updateOneRoute(switch, cmd_list, msg)
@@ -831,14 +835,20 @@ def marlExperiment(
 		if actions_target_flows:
 			intime = time.time()
 			for i, (node, sarsa, maps) in enumerate(zip(learners, sarsas, flow_action_sets)):
-				for ip, state in maps.iteritems():
+				for ip_pair, state in maps.iteritems():
+					# FIXME: don't send if action already taken...
 					#print "{}: {}".format(i, ip)
-					(_, action, machine, _) = state
+					(_, action, machine, _, guard) = state
+
+					if not guard[0]:
+						continue
+
 					a = action if override_action is None else override_action
 					# tx_ac = sarsa.actions[a] if isinstance(a, (int, long)) else a
 					tx_ac = machine.action() if isinstance(a, (int, long)) else a
 					#tx_ac = 0.0
 					updateUpstreamRoute(node, ac_prob=tx_ac, target_ip=ip)
+					guard[0] = False
 			outtime = time.time()
 			if print_times:
 				print "do_acs:", outtime-intime
@@ -1760,7 +1770,7 @@ def marlExperiment(
 
 								if ip_pair not in flow_space:
 									flow_space[ip_pair] = {
-										"ip": ip,
+										"ip": src_ip,
 										"last_act": 0.0 if not spiffy_mode else 0.05,
 										"last_rate_in": -1.0,
 										"last_rate_out": -1.0,
@@ -1829,18 +1839,18 @@ def marlExperiment(
 						flows_procd = 0
 						# ACT ON A SUBSET OF CURRENT
 						while can_act() and local_pos < len(local_work):
-							ip = local_work[local_pos]
-							local_visited_set.add(ip)
+							ip_pair = local_work[local_pos]
+							local_visited_set.add(ip_pair)
 
 							s_t = time.time()
-							l = flow_space[ip]
[…]

### 2019-05-07 code/marl/tnsm-results.py
First wave of results for mk.II paper...

@@ -38,7 +38,7 @@ host_ps = [
 # prefix, tcp?
 traffic_types = [
 	"tcp",
-#	"opus",
+	"opus",
 #	"mix",
 ]
 

### 2019-05-07 tables/tnsm-avg-reward.py
New table

@@ -1,0 +1,127 @@
+import csv
+import numpy as np
+from os import path
+import sys
+
+traffics = ["opus", "tcp"]
+ns = [2, 4, 8, 16]
+models = ["mpp", "spf"]
+variants = [
+	("separate", None, "Separate", None),
+	("single", None, "Single", None),
+]
+
+illegal_combos = set()
+# illegal_combos.add(("m", "udp", "channel"))
+# illegal_combos.add(("spf", "udp", "channel"))
+# illegal_combos.add(("spf", "tcp", "channel"))
+
+def get_average_reward(filename):
+	with open(filename, "r") as csvfile:
+		reader = csv.reader(csvfile)
+		# _ = nereader.next()
+
+		dats = []
+		for i, row in enumerate(reader):
+			if i == 0:
+				continue
+
+			dats.append(float(row[3]))
+
+		return np.mean(dats)
+
+def write_table(out_file, data):
+	lm = len(models)
+	lv = len(variants)
+
+	out_file.write(
+		"\\begin{{tabular}}{{@{{}}cc{}@{{}}}}\n".format("S"*((lm*lv)+1))
+	)
+	out_file.write(
+		"\\toprule\\multicolumn{1}{c}{Traffic} & " + \
+		"\\multicolumn{1}{c}{$n$} & " + \
+		"\\multicolumn{1}{c}{Marl} & " + \
+		"\\multicolumn{{{}}}{{c}}{{Marl++}} & ".format(lv) + \
+		"\\multicolumn{{{}}}{{c}}{{SPF}} ".format(lv) + \
+		"\\\\\n" + \
+		"\\cmidrule(lr){{4-{}}}\\cmidrule(lr){{{}-{}}} ".format(4 + lv-1, 4 + lv, 3 + lv + lv) + \
+		" & & "
+	)
+
+	for _ in models:
+		for (_fn, _replace_str, pres, _replace_m) in variants:
+			out_file.write("& \\multicolumn{{1}}{{c}}{{{}}} ".format(pres))
+
+	out_file.write("\\\\ \\midrule\n")
+
+	# Write file body.
+	for traffic in traffics:
+		for i, n in enumerate(ns):
+			key = (traffic,n)
+			out_file.write("{} & {} ".format("" if i>0 else traffic.upper(), n))
+
+			row = data[key]
+			for j, _ in enumerate(row):
+				if j == 0:
+					continue
+
+				j -= 1
+				# print(j, len(variants))
+				model = models[int(j/len(variants))]
[…]

### 2019-05-07 code/marl/marl.py
Disabled opus raw iface writing.
Generating vast numbers of clients (w/ differing IPs) has a very
definite cost as far as actions/updates are concerned...

@@ -288,7 +288,7 @@ def marlExperiment(
 			"-c", str(subclient_count),
 			"-b", "../opus-voip-traffic",
 			"--ip-strategy", "even",
-			"--iface", "{}-eth0".format(host.name),
+			#"--iface", "{}-eth0".format(host.name),
 			"--constant",
 			"--refresh",
 		]

### 2019-05-08 code/marl/marl.py
Fixes to the old way of doing things (and its intersection with the new way...)

@@ -257,6 +257,7 @@ def marlExperiment(
 	def th_cmd(dests, bw, target_ip=None):
 		if target_ip is None:
 			target_ip = random_target(dests)
+		#print "aiming for http://{}/".format(target_ip)
 
 		return [
 			"../traffic-host/target/release/traffic-host",
@@ -850,7 +851,6 @@ def marlExperiment(
 			for i, ((node_label, sarsa, _leader), maps) in enumerate(zip(actors, flow_action_sets)):
 				node = vertex_map[node_label]
 				for ip_pair, state in maps.iteritems():
-					#print "{}: {}".format(i, ip)
 					(_, action, machine, _, guard) = state
 
 					if not guard[0]:
@@ -860,14 +860,14 @@ def marlExperiment(
 					# tx_ac = sarsa.actions[a] if isinstance(a, (int, long)) else a
 					tx_ac = machine.action() if isinstance(a, (int, long)) else a
 					#tx_ac = 0.0
-					#print "placing {} in {}".format(tx_ac, ip_pair)
+					#print "agent {}: placing {} in {}".format(i, tx_ac, ip_pair)
 					updateUpstreamRoute(node, ac_prob=tx_ac, target_ip=ip_pair)
 					guard[0] = False
 			outtime = time.time()
 			if print_times:
 				print "do_acs:", outtime-intime
 		else:
-			for ((node_label, sarsa, _leader), maps) in zip(actors, flow_action_sets):
+			for i, ((node_label, sarsa, _leader), state) in enumerate(zip(actors, flow_action_sets)):
 				node = vertex_map[node_label]
 				if len(state) == 0:
 					action = default_machine_state
@@ -878,6 +878,7 @@ def marlExperiment(
 				tx_ac = sarsa.actions[a] if isinstance(a, (int, long)) else a
 				#print "chose {}".format(tx_ac)
 				#tx_ac = 0.9
+				#print "agent {}: placing {}".format(i, tx_ac)
 				updateUpstreamRoute(node, ac_prob=tx_ac)
 
 	def moralise(value, good, max_val=255, no_goods=[0, 255]):
@@ -1951,7 +1952,7 @@ def marlExperiment(
 
 					(local_work, local_work_set, local_visited_set) = queue_holder["curr"]
 					local_pos = queue_holder["pos"]
-					print "true", state_vec
+					#print "true", state_vec
 
 					# TODO: strip old entries?
 					# PROCESS ALL NEW DATA.
[…]

### 2019-05-12 paper/paper.tex
WIP massive paper overhaul.

--- 2019-01-14 paper/paper.tex
@@
 % !BIB program = biber
-\documentclass[conference, a4paper, 10pt, times]{IEEEtran}
+\documentclass[10pt, times, comsoc]{IEEEtran}
 \usepackage{nopageno}
@@
 %\usepackage[T1]{fontenc}
-\usepackage{amsfonts}
+%\usepackage{amsfonts}
 \usepackage{amsmath}
+\let\openbox\relax
 \usepackage{amsthm}
+\usepackage[cmintegrals]{newtxmath}
 \usepackage{bm}
@@
 
-\newcommand{\mytitle}{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention}
-%\newcommand{\mytitle}{Per-Host DDoS Mitigation by Direct-Control Reinforcement Learning}
+%\newcommand{\mytitle}{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention}
+\newcommand{\mytitle}{Per-Host DDoS Mitigation by Direct-Control Reinforcement Learning}
 
@@
 	pdftitle={\mytitle{}},
-%	pdfauthor={Kyle A. Simpson, Simon Rogers, Dimitrios P. Pezaros}
-	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
+	pdfauthor={Kyle A. Simpson, Simon Rogers, Dimitrios P. Pezaros}
+%	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
 }
@@
 \author{
-%	Kyle A.\ Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
-%	Kyle A.\ Simpson,
-%	Simon Rogers,
-%	Dimitrios P.\ Pezaros,\\
+%	Kyle~A.~Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
+	Kyle~A.~Simpson,
+	Simon~Rogers,
+	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
+	School of Computing Science, University of Glasgow, Glasgow, G12 8RZ, UK\\
 %	University of Glasgow, Glasgow, Scotland,\\
-%	\email{k.simpson.1@research.gla.ac.uk}
-Anonymous Giraffe,
-Anonymous Badger,
-Anonymous Ibex\\
-Unnamed Department, Nowhere\\
-\email{giraffe.a@unnamed.com}
+	\email{k.simpson.1@research.gla.ac.uk}
+%Anonymous Giraffe,
+%Anonymous Badger,
+%Anonymous Ibex\\
+%Unnamed Department, Nowhere\\
+%\email{giraffe.a@unnamed.com}
 }
@@
 \begin{abstract}
+?? Trim to 75--200 words, simplify.
+
 Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.
@@
 \end{abstract}
+
+\begin{IEEEkeywords}
+	Security services, distributed denial-of-service, software-defined networking, machine learning, reinforcement learning.
+\end{IEEEkeywords}
 
@@
 \emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
-Attackers achieve this either by exploiting peculiarities of operating system or application behaviour (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{flooding-based attacks}).
+Attackers achieve this either by exploiting peculiarities of operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
 Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
@@
 \emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
-
-?? EXPAND
 
@@
 Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.
+A caveat of tile coding remains, in that the value of $\alpha$ must be reduced according to the number of tilings to prevent divergence at the expense of slower learning ($\alpha \leftarrow \alpha / n_{\mathit{tilings}}$).
 
@@
 \subsection{Threat Model}
-
-%?? The fine folks at most security conferences will want to see one. I suppose it serves its purpose by delineating what an attacker can/cannot do.
-
-%?? Oultine attacker's goals, particularly wrt SPIFFY.
-%??
-They want to impact fair-share bandwidth allocation to flows (minimise), they have no worry about being unmasked (there are many available reflectors, which don't need to be bought etc...) [cite the rossow papers about amplification hell].
 An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
 Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
-We assume that attacks will be bandwidth-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
-?? VERIFY AND RESEARCH THAT: REVIEWER 1 BELIEVES OTHERWISE.
-The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.
+We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
+The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
 The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
@@
 
-\section{Foundational Analysis}\label{sec:environment-and-rl-algorithm}
-?? REWORK AND GUT THIS SECTION + NEXT
-
-%?? State the problem and what we want to achieve.
-To best discover how the use of RL techniques in the field of network intrusion prevention may be improved, it's prudent that we start by looking for weaknesses in work which exists today.
-The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant: their work helps to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
-We begin with a reimplementation and more principled re-examination of their work to further develop the state-of-the-art, while demonstrating the viability of deployment and integration of these techniques within a software-defined network.
-
-%?? Explicitly say the ``REIMPLEMENTING M\&K THING'' over here, to make later paras make a reasonable amount of sense.
-
-%??
-Why are we comparing against/building on/reimplementing M\&K?
-What was their main contribution?
-Answer: they helped to show the viability of live, adaptive control of the network to detect/prevent DDoS attacks...
-
-\subsection{Topology}\label{sec:topology}
-
-The network itself is tree-structured, where one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $\ell$ intermediate switches, which in turn each connect to $m$ egress switches.
-We say, then, that we have $k$ teams and $\ell m$ agents in each team.
-Agents are co-located with these egress switches, who control the proportion of upstream packets from $n$ external hosts to discard according to load statistics observed along their path to the server.
+\section{DDoS Mitigation with Per-flow Reinforcement Learning}
+Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
+However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
+We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
+
+\subsection{System Design and Assumptions}
+A deployment environment is defined as a network, with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
+These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
+Agents are co-located with each egress switch, and control the proportion of upstream packets from each external host to discard.
+Each destination node $s$ has a maximum capacity, $U_s$.
 
@@
 Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners.
-
 Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
-Each link has a delay of \SI{10}{\milli\second}.
-All links have unbounded capacity, save for the server-switch connection which is capped at a fixed $U_s$ \si{\mega\bit\per\second} in both directions; in practice, volumetric DDoS attacks exhibit a \emph{saturation zone} \cite{DBLP:journals/tnsm/SimpsonSMJPH18} which we aim to model while minimising the performance costs of emulation.
-\Cref{fig:marl-topol} demonstrates this visually.
+
+\subsection{Algorithm}
+To make decisions cheaply and at low latency, we use \emph{semi-gradient Sarsa with tile coding} as described in \cref{eqn:sg-sarsa} and \cref{sec:reinforcement-learning}, rather than making use of neural networks or more complicated function approximators.
+Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
+Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their weight vector updates with one another (but may share experience and traces with one another).
+
+Although the choice of a classical RL method likely brings lower theoretical performance, there are significant reasons to favour such methods; these include lower latency decision-making, lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
+This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
+Simpler decision boundaries reduce the risk of overfitting and unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.
+
+When choosing a learning algorithm, we compared against Q-learning as well as methods based on \emph{eligibility traces} such as Watkins's $\operatorname{Q}(\lambda)$ \cite[pp.
+312--314]{RL2E} and $\operatorname{Sarsa}(\lambda)$ \cite[pp.
+305]{RL2E}.
+Preliminary experiments found that Sarsa offered the best performance and behaviour.
+
+\subsubsection{Action rate}
+We chose to adapt the algorithm to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to facilitate effective learning.
+To this end, we allow agents to make many decisions per timestep.
+We maintain the last state-action pair associated with each source IP and destination, and calculate any actions for the flows which still exist.
+Finally, we update $\wvec{}$ using each available trace and the reward signal from the relevant destination.
+As exploration still occurs for each action, this approach reduces $\epsilon$ multiple notches every timestep.
+In turn, we increase the annealing window for $\epsilon$ by a factor of 8 so as to preserve exploration over time, by accounting for the greater volume of decisions being made.
+
+\subsubsection{Per-tile updates}
+While the standard formulation of \cref{eqn:sg-sarsa} updates the value of all tiles identically (by a scalar $\alpha \delta_t$), we found it more effective to compute a different temporal difference value \emph{for each tiling}.
+While we make use of the sum of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to set $\alpha$ to a higher value without divergence.
+A crucial observation is that value updates to each tile can move by different values in different directions, converging on effective estimates sooner.
+
+\subsubsection{Decision narrowings}
+When learning control on the basis of a high-dimensional, tile-coded state space, assignment of credit for each decision is difficult (because all tiles have identical gradient).
+To combat this, with probability $\epsilon$ an agent will mark a flow as being governed by a subset of the state space for the next 5 decisions.
+Each agent chooses actions on that source/destination pair using one element of local state, the global state, and the bias tile---we include the latter two to strike a balance between and accuracy and correct credit assignment.
+
+\subsection{Feature Space}\label{sec:feature-space}
 
@@
 	\centering
-\resizebox{0.9\linewidth}{!}{
+	\resizebox{0.8\linewidth}{!}{
 		\begin{tikzpicture}[
 		texts/.style = {text=black},
-labeltexts/.style = {text=uofgsandstone},
+		labeltexts/.style = {text=gray},
 		treeline/.style = {draw=uofgburgundy},
@@
 		load/.style = {fill=uofgcobalt},
+		loadline/.style = {draw=uofgcobalt, line width=0.75mm},
 		loadhide/.style = {fill=uofgcobalt!40!white},
@@
 		]
-\node[treenode, label={[texts]above:Server}] (root) {}
-child [treeline] { node [treenode, load, label={[texts]above:Core}] (sswitch) {}
-	child [treeline] { node [treenode, load, label={[texts]above:Leader}] (teaml) {} 
-		child [treeline] { node [treenode, load, label={[texts]above:Intermediate}] (inter) {}
-			child [treeline] { node [treenode, load, label={[texts]above:Agent/Egress}] (agent) {}
-				child [treeline] { node [treenode, external] (extern) {}
-					child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
-					child [hideline] { node [hidenode, externalhide] (endhost) {} }
-				}
-			}
-			child [hideline] { node [hidenode, loadhide] (endagent) {} }
-		}
-		child [hideline] { node [hidenode, loadhide] (endinter) {} }
-	}
-	child [hideline] { node [hidenode, loadhide] (endteaml) {} }
-	edge from parent
-	node[below, labeltexts] {$U_s$}
-};
-
-%\draw[-] (teaml) -- (endteaml);
-\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
-\node [labeltexts, right = -0.1cm of kdots] {$k$};
-\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
-\node [labeltexts, right = -0.1cm of ldots] {$\ell$};
-\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
-\node [labeltexts, right = -0.1cm of mdots] {$m$};
-\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
-\node [labeltexts, right = -0.1cm of ndots] {$n$};
+		\node [treenode, loadhide, label={[texts]above:Agent 1}] (mainagent) {};
+		\node [treenode, loadhide, right = of mainagent] (inner0) {};
+		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner0] (inner1) {};
+		\node [hidenode, below right = 0.5cm and 1cm of inner0] (inner2) {};
+		\node [treenode, loadhide, below right = 0.5cm and 1cm of inner1] (inner3) {};
+		\node [treenode, loadhide, right = of inner3] (inner4) {};
+		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
+		\node [hidenode, below right = 0.5cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};
+
+		\node [hidenode, above left = 0.5cm and 0.5cm of inner1, label={[labeltexts]above:Agent 2}] (agent2) {};
+		\node [hidenode, below left = 0.5cm and 0.5cm of inner2, label={[labeltexts]below:Agent 3}] (agent3) {};
+
+		\draw[hideline, -] ($ (mainagent) + (-1,0) $) -- (mainagent);
+		\draw[hideline, -] ($ (agent2) + (-1,0) $) -- (agent2);
+		\draw[hideline, -] ($ (agent3) + (-1,0) $) -- (agent3);
+		\draw[hideline, -] (inner1) -- (agent2);
+		\draw[hideline, -] (inner2) -- (agent3);
+
+		\draw[loadline, -] (mainagent) -- (inner0);
+		\draw[loadline, -] (inner0) -- (inner1);
+		\draw[hideline, -] (inner0) -- (inner2);
+		\draw[-] (inner1) -- (inner3);
+		\draw[hideline, -] (inner2) -- (inner3);
+		\draw[loadline, -] (inner3) -- (inner4);
+		\draw[loadline, -] (inner4) -- node [texts, above] {$U_{s_0}$} (s0) ;
+		\draw[hideline, -] (inner4) -- node [labeltexts, below] {$U_{s_1}$} (s1);
+
 		\end{tikzpicture}
@@
 	\caption{
-	Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $\ell$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
-%	Empty nodes are considered to be internal.
-	Red nodes are considered to be external, and the load of each blue node features in the state vector.
-	\label{fig:marl-topol}
+		Global state selection for a flow between an external host and server $s_0$ which passes over Agent 1.
+		All nodes in the path taken through the defended network are filled in blue, and all link load measurements which are chosen for action computation are indicated with a thick blue line.
+		\label{fig:global-state-path}
 	}
@@
 
-%Why use this topology, a fan-out tree, in particular?
-%While this is taken verbatim from the work we use as a basis, it is presented within the confines of a framework given by \textcite{DBLP:journals/ton/YauLLY05}.
-%This framework is, in turn, an extension of the class of topologies examined in the original pushback paper \cite{DBLP:journals/ccr/MahajanBFIPS02a}; this topology is far closer to the latter.
-%Crucially, it seems to not actually represent or correspond to any realistic deployment at all.
-Although \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}, the originators of this topology, make it clear that it exists as a fairly unrepresentative example \cite{DBLP:journals/ccr/MahajanBFIPS02a},
-%\begin{displaycquote}{DBLP:journals/ccr/MahajanBFIPS02a}
-%	These simulations do not pretend to use realistic topologies or traffic mixes... the simple simulations in this scenario are instead intended... as a first step towards more rigorous evaluation.
-%\end{displaycquote}
-%The topology chosen to evaluate our predecessor, indeed, matches theirs.
-%?? So, why does this proliferate, exist? They effectively disavowed it in their own paper. Reading comprehension is hard, I guess.
-%Remarkably, it seems to have propagated in a way which was not originally intended.
-it remains the case that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
-It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
-We aim to address this through \cref{sec:performance-in-an-emulated-environment}.
-
-Reimplementation of this network model was performed using mininet \cite{mininet}, a Python-based network emulator.
-Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
-This provides representative load and packet inter-arrival characteristics, since the agents as defined rely only upon load measurements.
-
-\subsection{Agent Definitions}
-
-Agents receive state input from monitors in the environment, a vector of the four load values (in \si{\mega\bit\per\second}) observed along their path to $s$: this is tile-coded with identical parameter choices to those of \textcite{DBLP:phd/ethos/Malialis14}.
-
-At every timestep, each agent installs an action via OpenFlow, instructing its host switch to drop each packet with probability $p$.
-Specifically, the agent chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
-These choices primarily ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
-In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
-It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
-
-We employ the reward functions introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of their global reward as a performance metric and their \emph{coordinated team learning} (CTL) reward for training.
-These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
-At time $t$, an agent $v$ (with team leader $l_v$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
+Our state space combines elements of global state (network link load observations) with per-flow measurements.
+Each is tile-coded with 8 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
+
+%?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
+Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}) depending upon the bandwidth measurements regularly received from monitors in the environment.
+For any flow, an agent then computes the path it would take through the network.
+The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
+In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
+\Cref{fig:global-state-path} illustrates the process.
+
+We build global state in this way to offer compatibility with multipath, multi-destination networks, offering support for diverse deployment environments from endpoint servers to transit ASes.
+Computing the path from agent to destination is not computationally expensive.
+Multipath routing is often fast since typical \emph{equal-cost multipath} (ECMP) routing algorithms simply hash a packet's flow key, and are deterministic to provide consistent quality-of-service to hosts.
+
+%?? Path computation fast and efficient due to deterministic routing based on hashes (ECMP)
+%?? Works for any arbitrary topology, even
+
+We describe and analyse each of the per-flow features included in the state vector throughout \cref{sec:rethinking-the-state-space}.
+Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
+
+\subsection{Reward Function}
+
+%?? Say ``we take $R_G$ from M and K''.
+%
+%?? Reward at each destination.
+
+\newcommand{\arrload}[2]{\operatorname{load}^{#2}_{t}(#1)}
+\newcommand{\uload}[1]{\arrload{#1}{\uparrow}}
+\newcommand{\dload}[1]{\arrload{#1}{\downarrow}}
+\newcommand{\bload}[1]{\arrload{#1}{\updownarrow}}
+\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
+%\fakepara{Reward function directionality}
+%The reward functions, as defined, do not take traffic direction into account.
+%We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
+Each destination node $s$ generates a reward signal, $R_{s,t}$, at every timestep $t$.
+Assume, for now, that each destination has access to some classification function $g(\cdot)$ which estimates the volume of legitimate traffic received, and expects to receive $\mathit{traffic}_s$.
+Denoting the upstream, downstream and combined loads $\uload{s}, \dload{s}, \bload{s}$ at this node:
+
 \begin{subequations}
 	\vspace{-0.5cm}
-\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
 	\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
 	\begin{gather}
-\cond{f} = [\load{s} > U_s],\\
-\cond{c}{(v)} = \cond{f} \cdot{} \, [\load{l_v} > U_s/k],\\
-R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\label{eqn:reward-rt}\\
-R_{t}^{\mathit{CTL}}(v) = (1 - \cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{c}{(v)},\label{eqn:reward-rctl}
+	\cond{s} = [\max(\uload{s}, \dload{s}) > U_s],\\
+	R_{s,t} = (1 - \cond{s}) \frac{g(\arrload{s}{-})}{\mathit{traffic}_s} - \cond{s},\label{eqn:reward-rtx}
 	\end{gather}
-where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.
-$\cond{f}$ and $\cond{c}{(v)}$ represent the ``overloaded'' condition for these reward functions.\label{eqn:reward}
+	replacing $\arrload{s}{-}$ in \cref{eqn:reward} with whichever directional load is prioritised according to the traffic characteristics of the deployment environment, where $\cond{s}$ represents the ``overloaded'' condition at destination $s$.\label{eqn:reward}
 \end{subequations}
-
-The tile-coded state and chosen reward function are then used as inputs for the \emph{semi-gradient Sarsa} algorithm described in \cref{eqn:sg-sarsa}, making use of $\epsilon$-greedy action selection.
-Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their experience or weight vector updates with one another.
-
-%?? M\&K Seem not to have a grasp on the importance of up/down traffic splits. Hmm......... (i.e., both could be relevant?) Man, who knows.
-
-%?? wrt reward function relying on heuristic elements or perfect knowledge:
-While it can be observed that reliance upon heuristic elements or perfect knowledge as in the above reward functions hampers their real-world applicability, a sufficiently well-trained agent needs only to follow the policy it has learned in a greedy manner, allowing pre-training by a simulated environment to transfer to reality.
-%?? BUT, if we don't have reward functions which are available at runtime, then we've lost out on what is arguably one of the main benefits that RL can give us!
-%?? Need to tie in the concept of ``meaningful work''?
-The lack of a truly online reward function is sorely felt, however.
-This causes the loss of some of RL's main benefits, such as online learning with regard to an evolving problem context.
-
-\section{Performance in an Emulated Environment}\label{sec:performance-in-an-emulated-environment}
-The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and high densities of hosts at egress points.
-%?? Why? ...
-Of these, the latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
-
-These are important issues, particularly when we consider real-world deployment.
-Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
-?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
-Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network.
-
-All experiments described were executed on Ubuntu 16.04.5 LTS (GNU/Linux 4.4.3-040403-generic x86\_64), using a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
-All code underpinning these findings is available on a public repository\footnote{Link removed for double-blind reviewing.}.
-%All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
-
-%?? TODO: Make repo public...
-
-\subsection{Service Guarantees and Host Density}
-The default topology chooses $n=2$ hosts per egress point, which isn't a parameter setting that can be described as realistic for many web service deployments.
-In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
-What this means is that datacentres and small-scale service deployments are more likely to have a host-to-learner ratio scaled heavily in favour of hosts, i.e. a large observed $n$.
-We feel is most worth examining aspects of such a case, where actions (as defined above) become less granular.
-
-To test this, we configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
-This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
-The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
-Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
-$U_s$ was fixed at $k \ell mn+2$ \si{\mega\bit\per\second}.
-The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
-Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
-These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.90\linewidth]{../plots/online-varyN-binary}
-	\caption{
-		MARL pushback control system performance plotted over time for various settings of $n$ hosts per agent.
-		This plot shows that, from the perspective of benign hosts, service guarantees degrade as inference and actions become less granular.
-		This generalises for all behavioural discriminators: even a perfect agent \emph{must} punish benign flows if they are grouped with malicious actors.
-		%		?? Effect on max perf AND learning rate??
-		Crucially, both the maximum achievable performance and learning rate are negatively impacted when granularity is coarse.
-		\label{fig:marl-granularity}
-	}
-\end{figure}
-
-\Cref{fig:marl-granularity} shows that, as more hosts are allocated to each learner, the fraction of traffic believed to be good (as observed at the server) decreases.
-This manifests in two ways: the best-achievable performance drops, and so too does the learning rate.
-The first consequence arises analytically.
+We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, though we expect that $\bload{\cdot}$ would be the most suitable for general deployment or heterogeneous traffic patterns.
+These choices reflect where the bulk of transmitted bytes in each traffic model are observed (and the lack of this knowledge in the general case).
+
+While our use and definition of $g(\cdot)$ appears nebulous, there are many ways to infer this quantity in practice.
+End-host servers may make use of canary flows or other active measurement schemes, or employ existing quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
+ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
+Even if such heuristics or perfect knowledge aren't available in deployment, a sufficiently well-trained agent needs only to follow the policy it has learned from training in a greedy manner, allowing pre-training by a simulated environment (with perfect knowledge) to transfer to reality.
+
+If a network is believed to be vulnerable to indirect attacks, such as link-flooding attacks, we may use the following reward:
+\begin{equation}
+	R_{s,t}^{\mathit{Cross}}(\beta) = \beta R_{s,t} + (1 - \beta) \min{\{R_{s',t} | s' \ne s\}} \label{eqn:lfa-reward}
+\end{equation}
+where the collaboration parameter $\beta \in [0,1]$ models the expected degree of interference between flows, and $s, s'$ are protected destination nodes in the network.
+The key insight underpinning LFAs is that flows can affect a target \emph{without communicating with that target}.
+$\beta$ then acts as a tunable parameter which can incentivise agents to remove flows which harm overall system health, by including the performance of the worst-performing destination.
+However, such attacks (and the effectiveness of $R_{s,t}^{\mathit{Cross}}$) are not examined by our work.
+
+\subsection{Action Space}
+When monitoring a source-destination pair, an agent uses its state vector to decide which proportion of that flow's \emph{inbound} traffic should be dropped.
+%?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
+This is implemented by installing an action via OpenFlow, instructing its host switch to drop each relevant packet with probability $p$.
+We choose to drop packets rather than impose traffic limits as it affords us a discrete action space without prior knowledge of traffic characteristics or measurement.
+Furthermore, we need not consider burstiness, fairness or implementation costs (such as per-flow bucket sizes) which could limit scalability.
+We offer two models on how to choose $p$:
+
+\subsubsection{Instant control}
+Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
+These choices primarily ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
+Since this model requires no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely opportunistic/myopic).
+
+\subsubsection{Guarded control}
+The measurements of \textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
+We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, choosing to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
+The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
+This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and non-congestion-aware traffic as required.
+
+The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
+To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$); this is chosen above the equivalent for UDP due to TCP's higher prevalence.
+Most importantly, an agent must choose to punish a flow multiple times in succession to cause rapid degradation, reducing variance while allowing an agent to see how a host reacts to structured changes in the environment.
+
+As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
+We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing with UDP traffic.
+
+\subsubsection{Risks}
+Our mode of action means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
+This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p$ is nonzero in any real network) while constant bitrate UDP traffic is proportional to $1 - p$.
+%It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
+This weakness still presents itself in modern flavours of TCP, such as TCP-Cubic which depends on $1/p^{0.75}$ \cite{rfc8312}.
+This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
+Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data transferred by volume \footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
+QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
+%As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development to properly handle traffic with such characteristics.
+
+%?? Make more sane. This is about ``Why go per-flow?''
+This further justifies our focus on per-flow decisions---real-world deployments see many flows pass over any egress point, making global actions (such as those chosen by \textcite{DBLP:journals/eaai/MalialisK15}) more likely to inflict collateral damage.
+%This manifests in two ways: the best-achievable performance drops, and so too does the learning rate.
 Given the probability that a host is legitimate, $P_G \in [0,1]$, it follows that a host will be malicious with probability $P_B = 1 - P_G$.
@@
 \end{corr}
-
-%?? Expand according to the summary you posted in the notes document.
-%That the use of UDP-like traffic replicates their experiments, compared against how real tcp behaves in mininet and according to the Mathis equation, shows that they really did not consider the behaviour of benign TCP in this environ (i.e., their results should be considerably worse in practice).
-
-\subsection{TCP Multiplexing and Falloff Behaviour}\label{sec:tcp-multiplexing-and-fallback-behaviour}
-
-%While Tcpreplay was useful in recreating the original work and examining the above condition, flow behaviour under packet loss is closer to that of UDP when using it to replay traffic.
-%To investigate the effects of collateral damage upon benign TCP flows, an iperf3 test was conducted using mininet.
-%Several hosts communicated concurrently to a single server over one switch, with maximum send rates in $\{1,...,5\}$ \si{\mega\bit\per\second} and a constant \SI{40}{\milli\second} latency.
-%The rate of packet drop was increased linearly from \SIrange{0}{40}{\percent} and the observed throughput for each host was measured, to show a trend and determine for the mininet environment whether throughput would decrease as expected, at a super-linear rate.
-
-%\begin{figure}
-%	\includegraphics[width=\linewidth]{../plots/mplex}
-%	\caption{
-%		Average TCP upload rate of hosts targeting different bandwidths in \emph{mininet}, computed over 10 runs, for choices of packet drop rate up until throughput is reduced to almost \SI{0}{\mega\bit\per\second}.
-%		As the rate of packet drop increases from \SIrange{0}{40}{\percent}, all flows converge on an extremely low throughput, regardless of the target bandwidth---a super-linear decrease for legitimate TCP traffic.
-%		Prior treatments have ignored such phenomena, which have strong implications on agent and network design.
-%		Malicious flows are not expected to obey the Mathis equation in this way, worsening collateral damage experienced by benign TCP flows.
-%		\label{fig:mplex-tcp}
-%	}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/online-varyN-nginx}
-	\caption{
-		Comparison of preserved traffic across different traffic models, choosing $n=2$ hosts per egress node.
-		No actions are taken in the ``baseline'' entries.
-		The agents examined here perform above the baseline in all cases, but perform significantly worse for live TCP traffic (nginx).
-		\label{fig:nginx-coffin-nail}
-	}
-\end{figure}
-
-%\Cref{fig:mplex-tcp} shows that this trend is present.
-Given the UDP-like behaviour of Tcpreplay, replicating the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is based upon does not capture the key interactions between TCP flows and packet drop as a control action.
-To the best of our knowledge, existing RL applications for this task employ the same formulation with incremental improvements, and so such a deficiency would be likely to occur in \emph{all} these prior works.
-
-From the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}, through which the performance characteristics of TCP under packet drop are well-known and well-understood, their results should be considerably worse in practice:
-\begin{equation}
-\operatorname{BW} = \frac{\operatorname{MSS}}{\operatorname{RTT}} \frac{C}{\sqrt{p}},
-\end{equation}
-for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions, and the probability $p \in (0, 1]$ that a packet is lost.
-?? Does Mathis still hold? We observed similar trends empirically, but perhaps not the \emph{exact} curve or falloff behaviour...
-Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant; good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
-
-TCP traffic as it appears here is generated according to \Cref{sec:a-new-normal}.
-\Cref{fig:nginx-coffin-nail} shows that this model performs poorly when attempting to safeguard TCP traffic, though it remains better than leaving the network unprotected.
-%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
-UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.
-
-These results are of particular importance due to TCP's prevalence within the Internet.
-The protocol is known to carry \SIrange{79}{94}{\percent} of packets, corresponding to \SIrange{89}{98}{\percent} of data transferred by volume \cite{DBLP:conf/saint/ZhangDJC09}.
-?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE TRAFFIC IN MODERN DAY WITH CAIDA.
-As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development to properly handle traffic with such characteristics.
-
-\subsection{Computational Cost}
-%??
-Consider talking about the execution times of the old MARL approach here?
-They're real nice (as expected), so we have lots of room to play around with while (hopefully) remaining under the 1ms target time given by \textcite{DBLP:conf/sigcomm/ChenL0L18}.
-
-%Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
-Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
-This is reassuring when measured alongside the insights from other work.
-\Textcite{DBLP:conf/sigcomm/ChenL0L18} observe that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
-%Most flows are short, and flow-size follows a heavy-tailed distribution.
-That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
-
-%\subsection{Synthesis}
-%%It'd be really good if I could figure out a rough proof for the service property...
-%%Okay... now say what this means for future viability! What changes can I make? Flow-specific metrics, of course! Increase granularity as much as I can!
-%Both of these experiments suggest that the most viable way to enhance and deploy this MARL approach is to consider flows individually to maximise performance for legitimate hosts and to detect malicious hosts in a more principled manner.
-%While this will add significant computational cost, this may be worked around with intelligent sampling and monitoring while taking multiple actions per-timestep \cite{DBLP:conf/hotnets/MaoAMK16}.
-%
-%?? Observations: training time lengthier by nature of emulated environment.
-%?? Risk of going for a simulation (i.e. numeric)? Always going to be interesting behaviour that is missed out on (in theory), at the cost of training time/limits of simulation speed
-
-\section{A New Normal}\label{sec:a-new-normal}
-
-%In establishing...
-%
-%?? How will I structure this?
-%?? Motivation -> Model -> Results?
-%?? OR Use the results of the last section to springboard into here?
-
-From what we have seen, it is difficult (or impossible) for trace-based or numerical simulations to correctly capture certain dynamics without an extraordinary amount of care or consideration.
-%As it turns out, 
-Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
-We contribute a network model built around live testing of TCP traffic in an SDN-enabled environment, which is adaptable to arbitrary topologies, with an explicit focus on preserving the real-time dynamics of TCP traffic.
-In particular, we are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
-Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
-If need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.
-
-%??
-ANGLE: set up an environment to test \emph{specific} behaviours to examine \emph{specific} problems in past work.
-I make no claims that it is perfect or representative for all traffic, just for this (likely common) behaviour which I expect to plague almost all legit TCP flows.
-
-%??
-Existing sims used for testing such applications reliant on traces, or not sophisticated enough to capture interactive, back-and-forth (correlated) behaviours---possibly discarded as second-hand effects by past work when these are so crucial given user traffic patterns (and the nature of the control signal we choose to enact).
-
-%??
-Remember, the motivation is clear.
-We don't care so much that it is "representative" wrt a specific deployment location or network type.
-The whole purpose of this is that we aim to test specific behaviour which traces cannot replicate (i.e., correlated back-and-forth, dynamics introduced to congestion-aware protocols, ...)
-%?? If we need to, we can condition the distribution of requests according to statistics mined from an existing trace if reviewer number 2 needs that extra push to be convinced.
-
-\subsection{Network Design}
-We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
-All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
-If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
-The controller is also responsible for computing how switches respond to ARP requests: the necessity of this arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
-We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.
-
-\subsection{Traffic Model}
-%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
-To model legitimate traffic, the server node runs an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
-Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
-Each host's download rate is limited to match whatever maximum bandwidth was assigned to it.
-This provides two modes of operation:
-\begin{itemize}
-	\item \emph{long flow} testing has each host constantly request a large binary file, such as a tarball of GCC8;
-	\item \emph{request-chain} testing has each host request several random files known to exist within the website, followed by any dependent resources for each (stylesheets, images, etc.).
-On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
-This presents a more balanced distribution of flow duration and size.
-\end{itemize}
-We offer these to model two extremes: long-life host-to-server transfers (a characteristic we might expect from e.g., streaming video), and common HTTP traffic.
-
-%?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
-Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
-Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that emulation remains tractable.
-When using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows---a \SI{30}{\percent} rate increase in expectation, uploading \SIrange{2.03}{2.18}{$\!
-\times U_s$}.
-The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
-
-\section{Acting On Individual Sources}\label{sec:acting-on-individual-sources}
-Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per source.
-However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
-We describe and justify our enhancements to the RL model of \textcite{DBLP:journals/eaai/MalialisK15}, present a new model which draws on domain knowledge leveraged by SPIFFY \cite{DBLP:conf/ndss/KangGS16}, and offer an empirical investigation into the effectiveness of flow-level statistics on live decision-making.
-
-\subsection{Marl++}\label{sec:marl-plus}
-%\fakepara{Multiple actions}
-%?? Talk about the modifications to standard: multiple actions per-timestep, updating many traces from one reward...
-Modifying the frequency with which actions and value functions are updated has important consequences for not only overall system performance, but also in \emph{how} an agent learns.
-We choose to adapt these algorithms to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to facilitate effective learning.
-To this end, we allow agents to make many decisions per timestep.
-We maintain the last state-action pair associated with each source IP, and calculate any actions for the flows which still exist.
-Finally, we update $\wvec{}$ according to \cref{eqn:sg-sarsa} with each available trace and the sole reward measurement an agent has at that time.
-As exploration still occurs for each action, this approach reduces $\epsilon$ multiple notches every timestep.
-In turn, we increase the annealing window for $\epsilon$ to \num{8000} action choices so as to preserve exploration over time, by accounting for the greater volume of decisions being made.
-
-%\fakepara{Timed Random Sequential actions}
-Taking actions in this manner means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
-This introduces numerous potential issues: inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated; at their worst, these present additional attack surfaces to an adversary.
+When considering that many hosts have an especially adverse reaction to our main means of control, flow-level granularity becomes an obvious choice.
+
+\subsection{Systems Considerations}
+Taking many actions per-timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
+This introduces some potential issues: inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated; at their worst, these present additional attack surfaces to an adversary.
 To adapt to these problems, we make use of \emph{timed random sequential} updates.
+
 Each agent begins with an empty work list.
@@
 State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
-Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 8 flows in this time.
+Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
 We expect this should be tuned based on the frequency at which statistics arrive.
-%?? Prune old? We can get through 8 per iteration, might use this as a basis for knowing when to say ``enough is enough''.
-
-\newcommand{\arrload}[2]{\operatorname{load}^{#2}_{t}(#1)}
-\newcommand{\uload}[1]{\arrload{#1}{\uparrow}}
-\newcommand{\dload}[1]{\arrload{#1}{\downarrow}}
-\newcommand{\bload}[1]{\arrload{#1}{\updownarrow}}
-\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
-%\fakepara{Reward function directionality}
-The reward functions, as defined, do not take traffic direction into account.
-We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
-Denoting the upstream, downstream and combined loads $\uload{s}, \dload{s}, \bload{s}$, we modify \cref{eqn:reward}:
-\begin{subequations}
-\begin{gather}
-\cond{f} = [\max(\uload{s}, \dload{s}) > U_s],\\
-\cond{c}{(v)} = \cond{f}{} \cdot{} \, [\max(\uload{l_v}, \dload{l_v}) > U_s/k],
-\end{gather}
-\label{eqn:reward-but-better}
-\end{subequations}
-replacing $\arrload{t}{}$ in \crefrange{eqn:reward-rt}{eqn:reward-rctl} with whichever directional load is prioritised.
-We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, though we expect that $\bload{\cdot}$ would be the most suitable for general deployment or heterogeneous traffic patterns.
-These choices reflect where the bulk of transmitted bytes in each model are observed (and the lack of this knowledge in the general case).
-
-%Explicitly mention \emph{why} we don't make use of neural networks, and instead favour more classical function approximation.
-Latency concerns, energy usage, graphics/TPU/whatever hardware availability or lack thereof across switches/routers/etc.
-To make decisions cheaply and at low latency, we maintain the usage of semi-gradient Sarsa with tile coding, rather than making use of neural networks or more complicated function approximators.
-Although this likely brings lower theoretical performance, there are more reasons than latency alone to favour classical methods; these include lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
-This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without necessarily introducing dedicated tensor processing hardware to networks.
-Simpler decision boundaries reduce the risk of overfitting and of unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.
-
-%?? Expand all this
-
-\subsection{SPF: SPIFFY Lite}
-
-An important question to ask while exploring the design space of RL agents is whether (and how) domain knowledge can improve the performance or learning rate of a defence system.
-This often comes as a trade-off: expressive power and control, weighed against the smaller action spaces afforded by existing insight or structure.
-We propose a second model, relying on the measurements of \textcite{DBLP:conf/ndss/KangGS16} which suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
-While it is not stated whether this behaviour applies to amplification attacks, we expect this to remain the case; the upload rate of any amplifier will be fixed at its upper bound given that the upstream links of any connected bots are saturated.
-Moreover, bots receive no feedback on the effective throughput of their attack flows.
-
-How might we apply their observations within the RL paradigm, while tailoring them towards non-ISP deployment?
-Our approach is to constrain how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, choosing to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
-The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
-This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and non-congestion-aware traffic as required.
-
-The action set is then reduced such that agents can \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ in single steps.
-To enable temporary bandwidth expansion in resource constrained deployments, every flow is initially placed under light packet drop ($p=0.05$, or state 1); this is chosen above the equivalent for UDP due to TCP's higher prevalence.
-We combine this with the enhancements described in \cref{sec:marl-plus}.
-
-This model is significantly different when considering how to tune the RL algorithm's parameters: each agent now requires the capability to plan ahead, since they may no longer choose $p$ directly from any state.
-This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
-Recounting \cref{sec:reinforcement-learning}, the choice of $\gamma$ affects which fraction of the current state-action pair's value is added to $R_t$ when updating the value of the last state-action pair.
-We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing with UDP traffic, by observing average performance over $\gamma$ increments of $0.1$.
-
-\subsection{Rethinking the State Space}\label{sec:rethinking-the-state-space}
+
+\section{Rethinking the State Space}\label{sec:rethinking-the-state-space}
 
@@
 	\centering
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-good}
-%	\end{subfigure}
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
-%	\end{subfigure}
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
-%	\end{subfigure}
-	\includegraphics[width=\linewidth]{../plots/ftprep-box}
+	\includegraphics[width=\linewidth]{../plots/ftprep-cap-box}
 	\vspace{-1.2cm}
 	\caption{
-%		Features: how good are they really? These are for UDP. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
-		Learned performance of Marl++ agents when benign traffic is UDP-like, using only a single feature as a basis for decisions.
+		Learned performance of Instant Control agents when benign traffic is UDP-like, using only a single feature as a basis for decisions.
 		Mean IAT, inbound packet sizes, and global state offer the best predictive performance, while most features offer marginal advantage over the unprotected baseline.
-		Here, the link's bandwidth limit is removed, and so agents automatically learn to keep bandwidth below $U_s$ by negative reinforcement---this offers a slight per-feature performance increase compared to when this explicit cap is present (not pictured).
 		\label{fig:udp-feature-plots}
@@
 \end{figure}
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\linewidth]{../plots/ftprep-laf-cap-box}
+%	\vspace{-1.2cm}
+%	\caption{
+%		?? UDP, combined with last action.
+%		\label{fig:udp-laf-feature-plots}
+%	}
+%\end{figure}
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/ftprep-good-2}
-	\caption{
-		Demonstration of learning rate for several features included as part of \cref{fig:udp-feature-plots}.
-		Performance starts low, before achieving convergence at some later time; this explains the high volume of outliers occurring below the main distribution of many features.
-		\label{fig:udp-fp-converge}
-	}
-\end{figure}
-
-%\begin{figure*}
-%	\centering
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-cap-good}
-%	\end{subfigure}
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-cap-good-2}
-%	\end{subfigure}
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-cap-good-3}
-%	\end{subfigure}
-%	\caption{
-%		Features: how good are they really? These are for UDP, but capped. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
-%		\label{fig:udp-cap-feature-plots}
-%	}
-%\end{figure*}
-
-\begin{figure}
-	\centering
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good}
-%	\end{subfigure}
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good-2}
-%	\end{subfigure}
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-good-3}
-%	\end{subfigure}
-	\includegraphics[width=\linewidth]{../plots/ftprep-tcp-box}
-	\vspace{-1.2cm}
-	\caption{
-%		Features: how good are they really (part 2)? These are for TCP. These are less conclusive...
-		Learned performance of Marl++ agents when benign traffic is TCP-like, using only a single feature as a basis for decisions (uncapped as in \cref{fig:udp-feature-plots}).
-		Global state and Mean IAT still offer the greatest improvement above baseline, but it is observed that different optimal features are expressed between different traffic models.
-		Packet-level statistics are considerably less effective for this class of traffic, with $\Delta$ Out Rate and Correspondence Ratio being proportionally better than they are for UDP.
-		\label{fig:tcp-feature-plots}
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good}
-%	\end{subfigure}
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-2}
-%	\end{subfigure}
-%	\begin{subfigure}{0.32\linewidth}
-%		\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-good-3}
-%	\end{subfigure}
 	\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-box}
@@
 	\caption{
-%		Features: how good are they really (part 3)? These are for TCP, but capped. These are less conclusive...
-		Learned performance of Marl++ agents when benign traffic is TCP-like, using only a single feature as a basis for decisions (with bandwidth capped physically).
-		Although negative rewards (and thus negative reinforcement) cannot occur, all features achieve considerably better performance, a reversal of our observation for UDP traffic.
-%		Global state and Mean IAT still offer the greatest improvement above baseline, but packet-level statistics are considerably less effective for this class of traffic.
+		Learned performance of Instant Control agents when benign traffic is TCP-like, using only a single feature as a basis for decisions.
+		All of the chosen features can offer a marked improvement over no protection at all.
+		Global state and Mean IAT still offer the greatest improvement above baseline, but packet-level statistics are considerably less effective for this class of traffic.
 		\label{fig:tcp-cap-feature-plots}
+	}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{../plots/ftprep-tcp-laf-cap-box}
+	\vspace{-1.2cm}
+	\caption{
+		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken as a basis for decisions.
+		This combination causes a significant improvement in the effectiveness of packet-level and per-window statistics.
+		\label{fig:tcp-laf-feature-plots}
 	}
@@
 \fakepara{Global state}
-These are the existing set of load measurements introduced in \cref{sec:environment-and-rl-algorithm}.
+These are the vector of load measurements along a flow's path introduced in \cref{sec:feature-space}.
 These indicate the overall health of the network, and crucially are all measurements which an agent directly controls.
@@
 As a result, they communicate with attack victims using their own IP address.
-In real-world scenarios the addresses of a set of reflector nodes might exhibit similarity, e.g., similarly unhardened services exposed by a single organisation.
-?? Can we support this claim?
+In real-world scenarios the addresses of a set of reflector nodes might exhibit similarity due to network uncleanliness \cite{DBLP:conf/imc/CollinsSFJWSK07}, e.g., similarly unhardened services exposed by a single organisation.
 
@@
 This feature also potentially allows forgiveness, serving as a reference point for determining whether a source mistakenly marked as malicious exhibits different falloff behaviour after punishment.
+It's important to note that this feature only makes sense once combined with another flow feature.
 
@@
 All of the above features, save for global state, are 1-dimensional.
-Each is tile-coded with 8 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
-We chose to take a separate tile-coding over each feature, ignoring combinatorial effects at present to minimise the amount of extra work an agent must perform.
-Indeed, combining too many features into one tile-coding can lead to poor performance due to an inability to generalise action values across similar states.
-On the other hand, treating features individually allows us to reach a quick understanding of the importance of each while seeing adequate performance.
-
-\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-feature-plots} for TCP), using a topology with $n=2$ hosts per egress point averaged over 10 episodes.
+\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), using a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
+\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends being observed when the same techniques is applied with UDP traffic (omitted).
 %?? Core findings---different protocols need different features, so everything we proposed above has a use!
 The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
-All features converge to their highest-observed performance within around \num{4000} timesteps: \cref{fig:udp-fp-converge} shows an example of this behaviour for differently performing features.
+All features converge to their highest-observed performance within around \num{4000} timesteps.
 In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
 %?? How do they do when combined after individual training? Pretty well, especially for TCP.
-Additional testing shows that the learned per-feature policies may be easily combined (by summing action values), and that this technique is particularly effective for TCP; these results are omitted to preserve space.
-In no cases, however, do we manage to completely block attack traffic---at convergence, we observe that system load remains consistently at $U_s$.
-
-?? POTENTIALLY CUT THE ``SYSTEM DYNAMICS'' AND/OR BANDED REWARDS, THEY DON'T ADD MUCH ANYWAY.
+%Additional testing shows that the learned per-feature policies may be easily combined (by summing action values), and that this technique is particularly effective for TCP; these results are omitted to preserve space.
+%In no cases, however, do we manage to completely block attack traffic---at convergence, we observe that system load remains consistently at $U_s$.
+
+\section{A New Normal}\label{sec:a-new-normal}
+
+%In establishing...
+%
+%?? How will I structure this?
+%?? Motivation -> Model -> Results?
+%?? OR Use the results of the last section to springboard into here?
+
+%From what we have seen, it is difficult (or impossible) for trace-based or numerical simulations to correctly capture certain dynamics without an extraordinary amount of care or consideration.
+%As it turns out, 
+%Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
+We contribute network models built around live testing of reactive TCP and UDP traffic in an SDN-enabled environment, which is adaptable to arbitrary topologies, with an explicit focus on preserving their real-time dynamics in a way that trace-based evaluation cannot.
+First and foremost, we are interested in representative load and packet inter-arrival characteristics and in how these characteristics evolve in response to actions.
+We introduce these models because are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
+%Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
+%If need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.
 
 %??
-NEED TO SAY SOMEWHERE: Baselines use TC for establishing interleaving/buffers filling etc., when MARL+ or SPF are in play then disabling TC (thus allowing illegal states, with \num{-1} reward marking them as such---``negative reinforcement'') may accelerate learning.
-We have a choice---but the results show that for TCP we WANT TC enabled...
-For UDP, we actually get a slight degradation in performance.
-This suggests a kind of ``banding'', perhaps?
-Important to mention that it easily learns how to maintain a logical bandwidth limit.
-During training of the above feature models, we elected to remove the physical bandwidth limit to allow agents to visit illegal states and receive \emph{negative reinforcement} for doing so.
-While this means that agents must learn how to keep traffic below a logical limit, they are helped by the fact that strong negative reward signals offer explicit information about which actions are poor choices: observe from \cref{eqn:reward,eqn:reward-but-better} that if a physical link has limit $U_s$ then we will never see a reward below 0.
-It is important to note that neither setting affects the agents' ability to keep the total load below $U_s$.
-%?? talk about cap difference: \cref{fig:tcp-cap-feature-plots}---we want cap enabled for TCP.
-We found this led to a marginal improvement in efficacy when UDP was the primary source of benign traffic, but observed that this hindered learning when TCP was the primary protocol under protection (\cref{fig:tcp-cap-feature-plots}), likely due to the loss of key system dynamics.
-It is anticipated that this difference might prove problematic for heterogeneous traffic mixtures.
-One way of solving this may be to investigate the effect of replacing $U_s$ with e.g., $0.99 U_s$ in \cref{eqn:reward-but-better}, to introduce a small band of negative reinforcement without actually removing the load restriction in the environment.
-It is crucial to make this region as small as possible, hence the somewhat arbitrary choice of factor; we observe that agents are always able to learn how to keep traffic consistently below the logical bandwidth limit (as in the original work).
-Our intent is to potentially accelerate learning while preserving the effects of system dynamics---we term these \emph{banded} rewards.
-
-%\subsection{What's done with these?}
-
-%?? Tile coded together, action taken per-flow every timestep.
+ANGLE: set up an environment to test \emph{specific} behaviours to examine \emph{specific} problems in past work.
+I make no claims that it is perfect or representative for all traffic, just for this (likely common) behaviour which I expect to plague almost all legit TCP flows.
 
 %??
-NOTE: think about how I might need to change/alter the tile-coding when we add extra discriminative features...
-(i.e., are there better ways to handle covariance between properties of system/flow state, properly capture generalisation?)
+Existing sims used for testing such applications reliant on traces, or not sophisticated enough to capture interactive, back-and-forth (correlated) behaviours---possibly discarded as second-hand effects by past work when these are so crucial given user traffic patterns (and the nature of the control signal we choose to enact).
 
 %??
-What other consequences are there?
-We're training each agent on all the flows it's acting on which means slight theoretical changes (we have multiple experience traces, and each uses its last state + new state + new target value (reward) as an update to the \emph{current model}).
+Remember, the motivation is clear.
+We don't care so much that it is "representative" wrt a specific deployment location or network type.
+The whole purpose of this is that we aim to test specific behaviour which traces cannot replicate (i.e., correlated back-and-forth, dynamics introduced to congestion-aware protocols, ...)
+%?? If we need to, we can condition the distribution of requests according to statistics mined from an existing trace if reviewer number 2 needs that extra push to be convinced.
+
+\subsection{Network Design}
+We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
+All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
+If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
+The controller is also responsible for computing how switches respond to ARP requests: the necessity of this arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
+%We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.
+
+\subsection{TCP (HTTP) Traffic Model}
+%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
+To model legitimate TCP traffic, the server nodes run an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
+Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
+Hosts and clients use the TCP-Cubic \cite{rfc8312} flavour of TCP.
+Each host's download rate is limited to match whatever maximum bandwidth was assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.).
+On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
+This presents a more balanced distribution of flow duration and size, with large files included to model elephant flows.
+
+\subsection{UDP (VoIP) Traffic Model}
+We model 
+?? Describe opus traffic model...\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}}.
+?? Based on Discord\footnote{\url{https://discord.gg}}
+?? Highly periodic, CBR
+?? Reflective factor among \emph{authenticated hosts}.
+
+%?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
+\subsection{Attack Traffic Model}
+Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
+Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that emulation remains tractable.
+
+\section{Evaluation}
+
+Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
+
+?? We use the traffic models introduced in \cref{sec:a-new-normal}.
+
+?? We evaluate in multi-agent mode, and in single-agent mode (0-cost perfect information-sharing).
+
+?? Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network.
+
+We compare our work most naturally against MARL \cite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
+
+The algorithm parameters were set at $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$ in the case of the \emph{Instant} model.
+
+?? Calculated expected malicious, keep with its other half.
+Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
+The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
+Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
+These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
+
+%The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and high densities of hosts at egress points.
+%?? Why? ...
+%Of these, the latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
+
+%These are important issues, particularly when we consider real-world deployment.
+%Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
+%?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
+
+??When UDP, \SIrange{2.5}{6}{\mega\bit\per\second}.
+When using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows---a \SI{30}{\percent} rate increase in expectation, uploading \SIrange{2.03}{2.18}{$\!
+\times U_s$}.
+The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
+
+?? $U_s$ was fixed at $N_{\mathit{hosts}}+2$ \si{\mega\bit\per\second}
+Each link has a delay of \SI{10}{\milli\second}.
+All links have unbounded capacity, save for the server-switch connection which is capped at a fixed $U_s$ \si{\mega\bit\per\second} in both directions.
+
+All experiments were executed on Ubuntu 18.04.2 LTS (GNU/Linux 4.4.3-040403-generic x86\_64), using a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
+%All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
+All code underpinning these findings is available on a public repository\footnote{Private until publication.}.
+
+\subsection{Single Destination}\label{sec:single-dest}
+%?? Move description of tree topol to here.
+The network is tree-structured, where one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $\ell$ intermediate switches, which in turn each connect to $m$ egress switches.
+We then have $N_{\mathit{hosts}} = k \ell m n$.
+\Cref{fig:marl-topol} demonstrates this visually.
+
+%Although \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}, the originators of this topology, make it clear that it exists as a fairly unrepresentative example \cite{DBLP:journals/ccr/MahajanBFIPS02a}, it remains the case that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
+%It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
+%We aim to address this through \cref{sec:performance-in-an-emulated-environment}.
+
+To test this, we configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
+This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
+%The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
+%Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
+%$U_s$ was fixed at $k \ell mn+2$ \si{\mega\bit\per\second}.
+%The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
+%Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
+These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
+
+\begin{figure}
+	\centering
+	\resizebox{0.9\linewidth}{!}{
+		\begin{tikzpicture}[
+		texts/.style = {text=black},
+		labeltexts/.style = {text=uofgsandstone},
+		treeline/.style = {draw=uofgburgundy},
+		treenode/.style = {texts, circle, centered, fill=white, treeline},
+		load/.style = {fill=uofgcobalt},
+		loadhide/.style = {fill=uofgcobalt!40!white},
+		external/.style = {fill=uofgrust},
+		externalhide/.style = {fill=uofgrust!40!white},
+		hideline/.style = {draw=uofgsandstone!40!white},
+		hidenode/.style = {treenode, hideline},
+		grow'=right
+		]
+		\node[treenode, label={[texts]above:Server}] (root) {}
+		child [treeline] { node [treenode, label={[texts]above:Core}] (sswitch) {}
+			child [treeline] { node [treenode, label={[texts]above:Leader}] (teaml) {} 
+				child [treeline] { node [treenode, label={[texts]above:Intermediate}] (inter) {}
+					child [treeline] { node [treenode, load, label={[texts]above:Agent/Egress}] (agent) {}
+						child [treeline] { node [treenode, external] (extern) {}
+							child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
+							child [hideline] { node [hidenode, externalhide] (endhost) {} }
+						}
+					}
+					child [hideline] { node [hidenode, loadhide] (endagent) {} }
+				}
+				child [hideline] { node [hidenode] (endinter) {} }
+			}
+			child [hideline] { node [hidenode] (endteaml) {} }
+			edge from parent
+			node[below, labeltexts] {$U_s$}
+		};
+
+		%\draw[-] (teaml) -- (endteaml);
+		\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
+		\node [labeltexts, right = -0.1cm of kdots] {$k$};
+		\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
+		\node [labeltexts, right = -0.1cm of ldots] {$\ell$};
+		\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
+		\node [labeltexts, right = -0.1cm of mdots] {$m$};
+		\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
+		\node [labeltexts, right = -0.1cm of ndots] {$n$};
+		\end{tikzpicture}
+	}
+	\caption{
+		Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $\ell$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
+		%	Empty nodes are considered to be internal.
+		Red nodes are considered to be external, and each blue node hosts an agent.
+		\label{fig:marl-topol}
+	}
+\end{figure}
+
+\subsection{Multiple Destinations}
+The previous topology allows for direct comparison against the state-of-the-art, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
+It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
+
+?? Come up with description of fat-tree (multi-dest) topol.
+?? Why fat tree? regularly appears in modern datacentres.
+
+?? $n \in \{6, 12, 24, 48\}$ hosts per learner (keeping $N_{\mathit{hosts}}$ identical to each tier of the single-host topology)
+?? Set $U_{s_0} = U_{s_1} = U_s / 2$
 
@@
 
-We now examine the performance of our two new models (\emph{Marl++}, \emph{SPF}) as compared against existing RL work (\emph{Marl}) under different traffic behaviour and varying $n$ as before.
+We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) under different traffic behaviour and topologies, varying the host-to-learner ratio $n$.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
-Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
-
+%Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
 We present the average rewards for all combinations of these factors in \cref{tab:av-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold.
@@
 
-Unless stated otherwise, these experiments measure average performance over 10 episodes, use all \num{14} features introduced in \cref{sec:rethinking-the-state-space} (tile-coded individually) alongside the global state, and make use of the physical bandwidth cap rather than pursuing negative reinforcement.
-In all cases, agents learn control of the network from scratch ($\wvec{}=\bm{0}$).
-Test system specifications match those given in \cref{sec:performance-in-an-emulated-environment}.
-
-\begin{table*}
+\begin{table}
 	\centering
-	\caption{Average reward for combinations of model, host density and traffic class.\label{tab:av-vals}}
-
-	\expandableinput ../tables/big-avg-reward.tex
-\end{table*}
+	\caption{Average reward for combinations of model, host density and traffic class with a single destination.\label{tab:av-vals}}
+
+%	\expandableinput ../tables/big-avg-reward.tex
+	\expandableinput ../tables/tnsm-avg-reward.tex
+\end{table}
 
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/udp-box}
+	\includegraphics[width=0.95\linewidth]{../plots/tnsm-udp-box-separate}
 
@@
 		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
+		?? REWORK/MAKE ACCURATE
 		\label{fig:udp-box}
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tcp-box}
+	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-separate}
 
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tcp-16-single}
+	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-16-single}
 
@@
 
-\subsection{Banded rewards and negative reinforcement}
-?? PROBABLY CUT THIS
-
-In general, we find that explicitly training a policy to keep the system in a valid state via negative reinforcement (i.e., an \emph{uncapped} network) is most effective for UDP-like traffic with Marl++.
-For SPF, however, negative reinforcement is typically a worse choice.
-Where TCP-like traffic is concerned in both models, we see that this strategy almost always leads to worse policies and higher collateral damage.
-
-The usefulness of banded rewards with regard to learning an effective policy appears to be unpredictable.
-For many of the configurations detailed in \cref{tab:av-vals}, for $n>2$ the average reward with banding is likely to lie between that of the capped and uncapped system models, but oscillates heavily.
-In a handful of UDP instances, banding can lead to \emph{better} average performance.
-More study is likely required into this; we replace $U_s$ with $0.99U_s$ to minimise the impact on legitimate states, but this choice remains somewhat arbitrary.
+%\subsection{Banded rewards and negative reinforcement}
+%?? PROBABLY CUT THIS
+%
+%In general, we find that explicitly training a policy to keep the system in a valid state via negative reinforcement (i.e., an \emph{uncapped} network) is most effective for UDP-like traffic with Marl++.
+%For SPF, however, negative reinforcement is typically a worse choice.
+%Where TCP-like traffic is concerned in both models, we see that this strategy almost always leads to worse policies and higher collateral damage.
+%
+%The usefulness of banded rewards with regard to learning an effective policy appears to be unpredictable.
+%For many of the configurations detailed in \cref{tab:av-vals}, for $n>2$ the average reward with banding is likely to lie between that of the capped and uncapped system models, but oscillates heavily.
+%In a handful of UDP instances, banding can lead to \emph{better} average performance.
+%More study is likely required into this; we replace $U_s$ with $0.99U_s$ to minimise the impact on legitimate states, but this choice remains somewhat arbitrary.
 
@@
 
-\subsection{Per-feature pre-training}
-To investigate the effects of pre-training individual features, we chose our four best-performing features (global state, mean IAT, $\Delta$ in/out rates) and trained each over \num{10000} iterations, before evaluating the policy obtained by summing the action values suggested by each.
-This may be achieved in a real network by randomly assigning new hosts to be controlled exclusively by one feature's policy, allowing agents to more easily learn \emph{which} information led to the correct action being taken.
-To reflect this (and train all parts of an agent in the same time frame), we reduce the timestep length to \SI{10}{\milli\second} to train each agent in less wall-clock time, at the cost of noisier data (and reduced learning rate).
-The results show little change for Marl++, but generally indicate a considerable gain for SPF (though smaller than what we'd measured for a single agent system).
+%\subsection{Per-feature pre-training}
+%To investigate the effects of pre-training individual features, we chose our four best-performing features (global state, mean IAT, $\Delta$ in/out rates) and trained each over \num{10000} iterations, before evaluating the policy obtained by summing the action values suggested by each.
+%This may be achieved in a real network by randomly assigning new hosts to be controlled exclusively by one feature's policy, allowing agents to more easily learn \emph{which} information led to the correct action being taken.
+%To reflect this (and train all parts of an agent in the same time frame), we reduce the timestep length to \SI{10}{\milli\second} to train each agent in less wall-clock time, at the cost of noisier data (and reduced learning rate).
+%The results show little change for Marl++, but generally indicate a considerable gain for SPF (though smaller than what we'd measured for a single agent system).
 %It's most likely that SPF is sensitive to noisy data when trying to learn how to act on TCP, and that as the different protocols did in \cref{fig:udp-feature-plots,fig:tcp-cap-feature-plots}, Marl++ and SPF have different optimal feature sets for different traffic behaviours---many of which may have been excluded.
+
+\subsection{Computational Cost}
+%??
+Consider talking about the execution times of the old MARL approach here?
+They're real nice (as expected), so we have lots of room to play around with while (hopefully) remaining under the 1ms target time given by \textcite{DBLP:conf/sigcomm/ChenL0L18}.
+
+%Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
+Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
+This is reassuring when measured alongside the insights from other work.
+\Textcite{DBLP:conf/sigcomm/ChenL0L18} observe that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
+%Most flows are short, and flow-size follows a heavy-tailed distribution.
+That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
+?? TODO: update with modern numbers...
 
@@
 ?? Actually talk about Marl (???) to appease reviewer \#1.
+The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15} (whom we have positioned our work against), and their contribution in applying RL to the task of intrusion prevention is significant: their work helps to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
+?? Recap their flaws, since they've been cut form every other aspect.
 
@@
 \section*{Acknowledgements}
-%Our thanks go to Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Richard Cziva for their advice, comments and technical assistance.
-Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
+Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice, comments and technical assistance.
+%Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
 %Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
-This work was supported by a nameless funder [grant number 0].
-%This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].
+%This work was supported by a nameless funder [grant number 0].
+This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].
 
@@
 
+\begin{IEEEbiography}
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/kyle}}]{Kyle A. Simpson}
+	received the MSci degree in computing science from the University of Glasgow in 2017.
+	He is currently a PhD student within the Networked Systems Research Laboratory at the School of Computing Science, University of Glasgow.
+
+	Mr.\ Kyle Simpson's research focusses on the use of machine learning and reinforcement learning techniques in cybersecurity and network management, with a core interest in evolving problems and defences.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/sdr}}]{Simon Rogers}
+	received the PhD from University of Bristol in 2006.
+Now he is a Senior Lecturer in the School of Computing Science at the University of Glasgow.
+Before starting a permanent  post in Glasgow he did a couple of Post Doctoral positions (under Mark Girolami).
+
+	Dr.\ Simon Rogers’s research involves the development of Machine Learning and Statistical techniques to help with the analysis of complex datasets, particularly within the field of Metabolomics but also other fields, Human-Computer Interaction and Information Retreival.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/dimitris}}]{Dimitrios P.\ Pezaros}
+	(S’01–M’04–SM’14) received the B.Sc.
+and Ph.D.
+degrees in computer science from Lancaster University.
+He is currently a Senior Lecturer (Associate Professor) and the Founding Director of the Networked Systems Research Laboratory at the School of Computing Science, University of Glasgow.
+He is also a visiting Professor at the University of Athens, Department of Informatics and Telecommunications.
+
+	Dr.\ Pezaros has published widely (over 80 peer-reviewed contributions to top-tier international conferences and journals) in the areas of computer communications, network and service management, and resilience of future networked infrastructures, and has received significant funding for his research in the above areas from public funding agencies and the industry.
+He is a Chartered Engineer, and a senior member of the IEEE and the ACM.
+
+\end{IEEEbiography}
+
 \end{document}

### 2019-05-12 paper/paper.tex
New model explanations etc.

@@
 \titlespacing*{\paragraph}{0pt}{1.5ex}{0.7ex}
+\titlespacing*{\subsubsection}{0pt}{1.5ex}{0.7ex}
 %\titleformat*{\filcenter\scshape}
@@
 	}
+
+	\ifentrytype{mics}{}{% Remove publisher and editor except for books
+		\clearfield{url}
+	}
 }
@@
 \newcommand{\crefrangeconjunction}{--}
+\crefname{table}{table}{tables}
 
@@
 	\centering
-	\resizebox{0.8\linewidth}{!}{
+	\resizebox{0.7\linewidth}{!}{
 		\begin{tikzpicture}[
@@
 %It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
-This weakness still presents itself in modern flavours of TCP, such as TCP-Cubic which depends on $1/p^{0.75}$ \cite{rfc8312}.
+This weakness still presents itself in modern flavours of TCP, such as TCP Cubic which depends on $1/p^{0.75}$ \cite{rfc8312}.
 This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
-Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data transferred by volume \footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
+Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
 QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
@@
 	\centering
-	\includegraphics[width=\linewidth]{../plots/ftprep-cap-box}
-	\vspace{-1.2cm}
+	\includegraphics[width=0.9\linewidth]{../plots/ftprep-cap-box}
+	\vspace{-1cm}
 	\caption{
@@
 	\centering
-	\includegraphics[width=\linewidth]{../plots/ftprep-tcp-cap-box}
-	\vspace{-1.2cm}
+	\includegraphics[width=0.9\linewidth]{../plots/ftprep-tcp-cap-box}
+	\vspace{-1cm}
 	\caption{
@@
 	\centering
-	\includegraphics[width=\linewidth]{../plots/ftprep-tcp-laf-cap-box}
-	\vspace{-1.2cm}
+	\includegraphics[width=0.9\linewidth]{../plots/ftprep-tcp-laf-cap-box}
+	\vspace{-1cm}
 	\caption{
@@
 Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
-Hosts and clients use the TCP-Cubic \cite{rfc8312} flavour of TCP.
+Hosts and clients use the TCP Cubic \cite{rfc8312} flavour of TCP.
 Each host's download rate is limited to match whatever maximum bandwidth was assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.).
@@
 
-\subsection{UDP (VoIP) Traffic Model}
-We model 
-?? Describe opus traffic model...\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}}.
-?? Based on Discord\footnote{\url{https://discord.gg}}
-?? Highly periodic, CBR
-?? Reflective factor among \emph{authenticated hosts}.
+\subsection{UDP (Opus) Traffic Model}
+VoIP traffic exhibits very different characteristics to the above model; packet arrivals are highly periodic due to real-time application requirements, flows have a constant bitrate, and flow behaviour which does not react substantially to lost packets.
+Interestingly, DDoS attack traffic is known to share many of these characteristics, offering an interesting detection problem.
+
+We present a VoIP traffic model\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward video game communities.
+We chose Discord as our prototype due to its publicly documented API, many open source bot frameworks for experimentation, large user base, and due to the lack of models for Opus-encoded traffic.
+%?? Highly periodic, CBR
+
+Hosts send RTP traffic with Salsa20 encrypted payloads---audio frames of length \SI{20}{\milli\second} at a target \SI{96}{\kilo\bit\per\second}.
+We generate similar traffic at hosts by replaying anonymised traces gathered in general and tabletop RPG servers; each trace contains only the size of each audio payload, entries denoting missed packets, and the duration of silent periods.
+We trim these silent periods to a maximum \SI{5}{\second} due to the lengthy talk/silence bursts introduced by users in RPG servers.
+Hosts punctuate audio frames with a 4-byte keepalive every \SI{5}{\second}.
+All traffic passes over a central server which groups hosts into rooms, and is forwarded to other participants.
+There is no peer-to-peer traffic---the server acts as a TURN relay \cite{rfc5766} for all hosts.
+%?? Reflective factor among \emph{authenticated hosts}.
+
+We find that each flow occupies an expected \SI{52.4}{\kilo\bit\per\second} upstream bandwidth.
+To match the target upload rate required at a host, we run enough sessions to meet the target data rate.
 
@@
 
-Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
-
-?? We use the traffic models introduced in \cref{sec:a-new-normal}.
-
-?? We evaluate in multi-agent mode, and in single-agent mode (0-cost perfect information-sharing).
-
-?? Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network.
-
-We compare our work most naturally against MARL \cite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
-
-The algorithm parameters were set at $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$ in the case of the \emph{Instant} model.
-
-?? Calculated expected malicious, keep with its other half.
-Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
-The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
-Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
-These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
-
-%The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and high densities of hosts at egress points.
-%?? Why? ...
-%Of these, the latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
-
-%These are important issues, particularly when we consider real-world deployment.
-%Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
-%?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
-
-??When UDP, \SIrange{2.5}{6}{\mega\bit\per\second}.
-When using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows---a \SI{30}{\percent} rate increase in expectation, uploading \SIrange{2.03}{2.18}{$\!
-\times U_s$}.
-The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
-
-?? $U_s$ was fixed at $N_{\mathit{hosts}}+2$ \si{\mega\bit\per\second}
-Each link has a delay of \SI{10}{\milli\second}.
-All links have unbounded capacity, save for the server-switch connection which is capped at a fixed $U_s$ \si{\mega\bit\per\second} in both directions.
+%Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
+
+We compare our work most naturally against the MARL approach introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
+We are most interested in seeing how their approach contrasts with ours across different topologies and workloads.
+Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network---we aim to see how these characteristics affect performance and learning rate.
+
+To test this, we made use of both traffic models introduced in \cref{sec:a-new-normal} (OPUS and TCP), both topologies discussed below (1-dest vs Fat-Tree), and vary the amount of hosts typically communicating over each agent's ingress/egress node.
+Additionally, we evaluated our new models in multi-agent mode (\emph{separate}, no model sharing), and in single-agent mode (\emph{single}, 0-cost perfect information sharing).
+In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
+Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes.
 
@@
 %All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
-All code underpinning these findings is available on a public repository\footnote{Private until publication.}.
+All code underpinning these findings is available on a public repository.\footnote{Private until publication.}
 
@@
 	\centering
-	\resizebox{0.9\linewidth}{!}{
+	\resizebox{0.6\linewidth}{!}{
 		\begin{tikzpicture}[
@@
 		%	Empty nodes are considered to be internal.
-		Red nodes are considered to be external, and each blue node hosts an agent.
+		Red nodes are external, and each blue node hosts an agent.
 		\label{fig:marl-topol}
@@
 It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
-
-?? Come up with description of fat-tree (multi-dest) topol.
-?? Why fat tree? regularly appears in modern datacentres.
-
-?? $n \in \{6, 12, 24, 48\}$ hosts per learner (keeping $N_{\mathit{hosts}}$ identical to each tier of the single-host topology)
-?? Set $U_{s_0} = U_{s_1} = U_s / 2$
+The fat-tree topology \cite{DBLP:conf/sigcomm/Al-FaresLV08} sees regular use in real-world datacentres and scales well horizontally.
+
+%?? Come up with description of fat-tree (multi-dest) topol.
+%?? Why fat tree? regularly appears in modern datacentres.
+%?? $k=4$ fat-tree , with one pod hosting two servers $s_0,s_1$.
+We use a $k=4$ fat-tree, with one pod hosting two servers $s_0$ and $s_1$.
+$n$ external hosts connect through each core switch (where agents are hosted), and communicate with $s_0, s_1$ uniformly randomly.
+Both servers host identical services.
+We set $n \in \{6, 12, 24, 48\}$ hosts per learner (keeping $N_{\mathit{hosts}}$ identical to each tier of the single-host topology), and restrict $U_{s_0} = U_{s_1} = U_s / 2$.
+
+\subsection{Parameters}
+The algorithm parameters were set at $\alpha=0.05$, linearly annealing $\epsilon=0.2 \rightarrow 0$ by $t=3000$ in the case of the Instant model (\num{12000} actions in the Guarded model).
+
+Benign hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second}, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
+%The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and high densities of hosts at egress points.
+%?? Why? ...
+%Of these, the latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
+%These are important issues, particularly when we consider real-world deployment.
+%Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
+%?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
+Malicious hosts sent \SIrange{2.5}{6}{\mega\bit\per\second} when testing UDP traffic, though when using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows.
+Given $n$ and $\operatorname{P}(\mathit{malicious})$, we see an expected malicious bandwidth \numrange{1.27}{1.87} and \SIrange{2.03}{2.18}{$\! \times U_s$} respectively.
+%The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
+
+$U_s$ was fixed at $N_{\mathit{hosts}}+2$ \si{\mega\bit\per\second} (to account for burstiness), and each link had a delay of \SI{10}{\milli\second}.
+All links had unbounded capacity, save for server-switch connections.
+These parameter choices match those of the original study to enable direct comparison, and many are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
 
@@
 
-We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) under different traffic behaviour and topologies, varying the host-to-learner ratio $n$.
+We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
 %Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
-We present the average rewards for all combinations of these factors in \cref{tab:av-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold.
+We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold.
 Average rewards take into account any portions of time that an agent allows illegal system states.
@@
 
-%	\expandableinput ../tables/big-avg-reward.tex
 	\expandableinput ../tables/tnsm-avg-reward.tex
+\end{table}
+\begin{table}
+	\centering
+	\caption{Average reward for combinations of model, host density and traffic class with multiple destinations.\label{tab:av-ecmp-vals}}
+
+	\expandableinput ../tables/tnsm-ecmp-avg-reward.tex
 \end{table}
@@
 	\caption{
-		Online performance for UDP-like benign traffic.
-		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
-		SPF, remarkably, slightly outperforms Marl++ (with lower variance) for this more difficult topology despite being worse for smaller $n$.
-		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
+		Online performance for Opus benign traffic in a single-destination network, multi-agent mode.
+%		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
+		\emph{Instant} outperforms Marl for $n \in \{4, 8\}$ (with higher variance), but performs similarly to Marl at $n\in \{2, 16\}$.
+		\emph{Guarded} always underperforms compared to the other agent designs in this problem variant.
+%		SPF, remarkably, slightly outperforms Marl++ (with lower variance) for this more difficult topology despite being worse for smaller $n$.
+%		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
 		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
-		?? REWORK/MAKE ACCURATE
-		\label{fig:udp-box}
+%		?? REWORK/MAKE ACCURATE
+		\label{fig:udp-tree-box}
+	}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.95\linewidth]{../plots/tnsm-udp-box-single}
+
+	\caption{
+		Online performance for Opus benign traffic in a single-destination network, single-agent mode.
+		?? DO I need this?
+		\label{fig:udp-tree-box-single}
 	}
@@
 \Cref{fig:udp-box} shows the behaviour for all choices of $n$---specifically, we see that Marl++ typically remains ahead of Marl, though this performance difference shrinks as $n$ increases, while SPF performs better at high host-density.
+
+?? Discuss multi-dest topol once results available.
 
@@
 	\caption{
-		Online performance for TCP-like benign traffic.
+		Online performance for HTTP benign traffic in a single-destination network, multi-agent mode.
 		Marl and Marl++ begin with essentially identical efficacy, with Marl++ having a small advantage at larger $n$.
 		SPF is slightly worse than Marl and Marl++ at low $n$, but achieves considerably better median performance and lower variance than the other models for large $n$.
-		\label{fig:tcp-box}
+		\label{fig:tcp-tree-box}
+	}
+\end{figure}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-single}
+
+	\caption{
+		Online performance for HTTP benign traffic in a single-destination network, single-agent mode.
+		?? DO I need this?
+		\label{fig-tcp-tree-box-single}
 	}
@@
 	\caption{
-		Online performance of standard and single-agent models with $n=16$ hosts per egress point and TCP-like benign traffic.
-		At this level of host density, SPF reaches its peak performance sooner and is considerably more consistent throughout the episode.
-		Marl++ takes longer to reach its most effective policy, and achieves visibly better peak performance than the other standard models.
-		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
-		\label{fig:tcp-16}
+		Online performance of standard and single-agent models in a single-destination network with $n=16$ hosts per egress point and HTTP benign traffic.
+		At this level of host density, Guarded reaches higher peak performance sooner and is considerably more consistent throughout the episode.
+		Guarded benefits greatly from information sharing, converging to protect around \SI{75}{\percent} of TCP traffic within \SI{100}{\second}.
+		The Instant model converges to perform similarly to Marl.
+%		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
+		\label{fig:tcp-tree-16}
 	}
@@
 
-%\subsection{Banded rewards and negative reinforcement}
-%?? PROBABLY CUT THIS
-%
-%In general, we find that explicitly training a policy to keep the system in a valid state via negative reinforcement (i.e., an \emph{uncapped} network) is most effective for UDP-like traffic with Marl++.
-%For SPF, however, negative reinforcement is typically a worse choice.
-%Where TCP-like traffic is concerned in both models, we see that this strategy almost always leads to worse policies and higher collateral damage.
-%
-%The usefulness of banded rewards with regard to learning an effective policy appears to be unpredictable.
-%For many of the configurations detailed in \cref{tab:av-vals}, for $n>2$ the average reward with banding is likely to lie between that of the capped and uncapped system models, but oscillates heavily.
-%In a handful of UDP instances, banding can lead to \emph{better} average performance.
-%More study is likely required into this; we replace $U_s$ with $0.99U_s$ to minimise the impact on legitimate states, but this choice remains somewhat arbitrary.
+?? Discuss multi-dest topol once results available.
 
@@
 We discuss our conjectures for why this reversal occurs in \cref{sec:discussion}.
-
-%\subsection{Per-feature pre-training}
-%To investigate the effects of pre-training individual features, we chose our four best-performing features (global state, mean IAT, $\Delta$ in/out rates) and trained each over \num{10000} iterations, before evaluating the policy obtained by summing the action values suggested by each.
-%This may be achieved in a real network by randomly assigning new hosts to be controlled exclusively by one feature's policy, allowing agents to more easily learn \emph{which} information led to the correct action being taken.
-%To reflect this (and train all parts of an agent in the same time frame), we reduce the timestep length to \SI{10}{\milli\second} to train each agent in less wall-clock time, at the cost of noisier data (and reduced learning rate).
-%The results show little change for Marl++, but generally indicate a considerable gain for SPF (though smaller than what we'd measured for a single agent system).
-%It's most likely that SPF is sensitive to noisy data when trying to learn how to act on TCP, and that as the different protocols did in \cref{fig:udp-feature-plots,fig:tcp-cap-feature-plots}, Marl++ and SPF have different optimal feature sets for different traffic behaviours---many of which may have been excluded.
 
@@
 That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
-?? TODO: update with modern numbers...
+%?? TODO: update with modern numbers...
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/policy-16-tcp-f5-mean-log}
-
-	\caption{
-		Mean Marl action values from pre-training on mean IATs ($n=16$).
-		As these are mean values, lighter cells in each row indicate actions \emph{likely} to be taken by an agent's policy.
-		Repeated values originate from the bias tile, which is always active, and indicate regions of the state-space which have yet to be visited---here, this is most of the state space.
-		Agents visibly learn action values for tiles which override the default bias tile's preferences.
-		The measured effectiveness of this feature then suggests that a low-resolution coding over the standard region may, in fact, be a better choice.
-		We see that agents prefer to choose $p=0.0$ in most states, but higher $p$ when IATs are particularly small.
-		\label{fig:intern-16-tcp-iat}
-	}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/policy-4-tcp-f7-mean}
-
-	\caption{
-		Mean Marl action values from pre-training on $\Delta$ out rates ($n=4$).
-		A sudden ramp-up of server-to-host traffic is used as a strong indicator of flow legitimacy, while more punishing actions have comparatively higher value as $\Delta$ out rate decreases.
-		Furthermore, we see again that a region of the state space has gone somewhat unexplored---we observed after plotting this visualisation that many decreases in this feature are too small to hit the adjacent tile, which implies a mixed-resolution coding may improve an agent's policy further.
-		\label{fig:intern-16-tcp-something}
-	}
-\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.9\linewidth]{../plots/policy-16-tcp-f5-mean-log}
+%	
+%	\caption{
+%		Mean Marl action values from pre-training on mean IATs ($n=16$).
+%		As these are mean values, lighter cells in each row indicate actions \emph{likely} to be taken by an agent's policy.
+%		Repeated values originate from the bias tile, which is always active, and indicate regions of the state-space which have yet to be visited---here, this is most of the state space.
+%		Agents visibly learn action values for tiles which override the default bias tile's preferences.
+%		The measured effectiveness of this feature then suggests that a low-resolution coding over the standard region may, in fact, be a better choice.
+%		We see that agents prefer to choose $p=0.0$ in most states, but higher $p$ when IATs are particularly small.
+%		\label{fig:intern-16-tcp-iat}
+%	}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.9\linewidth]{../plots/policy-4-tcp-f7-mean}
+%	
+%	\caption{
+%		Mean Marl action values from pre-training on $\Delta$ out rates ($n=4$).
+%		A sudden ramp-up of server-to-host traffic is used as a strong indicator of flow legitimacy, while more punishing actions have comparatively higher value as $\Delta$ out rate decreases.
+%		Furthermore, we see again that a region of the state space has gone somewhat unexplored---we observed after plotting this visualisation that many decreases in this feature are too small to hit the adjacent tile, which implies a mixed-resolution coding may improve an agent's policy further.
+%		\label{fig:intern-16-tcp-something}
+%	}
+%\end{figure}
 
@@
 gross.)
+?? I believe their assumptions may not extend to CBR traffic (e.g., UDP VoIP traffic), which we know makes up a sizeable proportion of network traffic. Also DASH traffic.
 
@@
 A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
-The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
+%The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
+The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
 Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
@@
 
-\renewcommand*{\bibfont}{\small}
+\renewcommand*{\bibfont}{\footnotesize}
 \printbibliography
@@
 	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/sdr}}]{Simon Rogers}
-	received the PhD from University of Bristol in 2006.
+	received the PhD degree from University of Bristol in 2006.
 Now he is a Senior Lecturer in the School of Computing Science at the University of Glasgow.
@@
 
-	Dr.\ Simon Rogers’s research involves the development of Machine Learning and Statistical techniques to help with the analysis of complex datasets, particularly within the field of Metabolomics but also other fields, Human-Computer Interaction and Information Retreival.
+	Dr.\ Simon Rogers’s research involves the development of machine learning and statistical techniques to help with the analysis of complex datasets, particularly within the field of metabolomics but also other fields, human-computer interaction and information retrieval.
 \end{IEEEbiography}

### 2019-05-12 tables/tnsm-ecmp-avg-reward.py

@@ -12,23 +12,28 @@ variants = [
 ]
 
 illegal_combos = set()
-# illegal_combos.add(("m", "udp", "channel"))
+illegal_combos.add(("spf", "opus", "single"))
 # illegal_combos.add(("spf", "udp", "channel"))
 # illegal_combos.add(("spf", "tcp", "channel"))
 
 def get_average_reward(filename):
-	with open(filename, "r") as csvfile:
-		reader = csv.reader(csvfile)
-		# _ = nereader.next()
+	output = 0.0
+	try:
+		with open(filename, "r") as csvfile:
+			reader = csv.reader(csvfile)
+			# _ = nereader.next()
 
-		dats = []
-		for i, row in enumerate(reader):
-			if i == 0:
-				continue
+			dats = []
+			for i, row in enumerate(reader):
+				if i == 0:
+					continue
 
-			dats.append(float(row[3]))
+				dats.append(float(row[3]))
 
-		return np.mean(dats)
+			output = np.mean(dats)
+	except:
+		print("file {} not found".format(filename))
+	return output
 
 def write_table(out_file, data):
 	lm = len(models)
@@ -94,7 +99,7 @@ def get_data():
 			row = []
 
 			row.append(get_average_reward(
-				"../results/tnsm-tree-{}-{}-m-separate.csv".format(
+				"../results/tnsm-ecmp-{}-{}-m-separate.csv".format(
 					n,
 					t,
 					)
@@ -103,7 +108,7 @@ def get_data():
 			# do it for all parts now.
 			for m_i, m in enumerate(models):
 				for v in variants:
-					f_str = "../results/tnsm-tree-{}-{}-{}-{}.csv" if v[1] is None else v[1]
+					f_str = "../results/tnsm-ecmp-{}-{}-{}-{}.csv" if v[1] is None else v[1]
 					m_str = m if v[3] is None else v[3][m_i]
 					row.append(get_average_reward(
 						# f_str.format(m_str, t, v[0], n)

### 2019-05-13 paper/paper.tex
Draft 1 for TNSM

@@
 \begin{abstract}
-?? Trim to 75--200 words, simplify.
-
-Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.
-The detection problem in this domain presents unique difficulties; it is an evolving, non-stationary problem as usage patterns shift and new protocols and applications are introduced, further compounded by burstiness and seasonal variation.
-
-\emph{Reinforcement learning} (RL) may overcome the detection problem for certain classes of anomaly by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available).
-We examine how the paradigm has been applied to the task of \emph{Distributed Denial of Service} (DDoS) attack prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.
-These flaws---poor handling of TCP traffic and high host density at egress points---inspire and direct an exploration of the design space of RL models.
-We advance the state-of-the-art of DDoS mitigation using RL by introducing two agent classes designed to act on a per-flow basis, in a protocol-agnostic manner.
+%?? Trim to 75--200 words, simplify.
+DDoS attacks plague the availability of online services today, yet like many cybersecurity problems are evolving and non-stationary.
+Normal and attack patterns shift as new protocols and applications are introduced, further compounded by burstiness and seasonal variation.
+Accordingly, application of machine learning-based techniques is difficult.
+%Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.
+
+\emph{Reinforcement learning} (RL) may overcome this detection problem for DDoS attacks by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available) in an online manner.
+%We examine how the paradigm has been applied to the task of \emph{Distributed Denial of Service} (DDoS) attack prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.
+We advance the state-of-the-art in RL-based DDoS mitigation by introducing two agent classes designed to act on a per-flow basis, in a protocol-agnostic manner for any network topology.
+%These flaws---poor handling of TCP traffic and high host density at egress points---inspire and direct an exploration of the design space of RL models.
 This is supported by an in-depth investigation of feature suitability and empirical evaluation.
 
-Our results show the existence of flow features with high predictive power for environments with predominantly TCP or UDP traffic, when such features are used as a basis for feedback-loop-like control.
-We show that the new RL agent models can offer a significant increase in goodput of legitimate UDP traffic for many choices of host density, alongside a marked increase in protected TCP traffic at high host density.
+Our results show the existence of flow features with high predictive power for different traffic classes, when a basis for feedback-loop-like control.
+We show that the new RL agent models can offer a significant increase in goodput of legitimate TCP traffic for many choices of host density.
 
@@
 In tandem with the development and evaluation of an effective state-space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
-Our introduction of per-source decisions improves substantially upon the state-of-the-art, and we show that our second model achieves excellent performance for high host density.
+Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
 Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
@@
 %	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
-	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}.
+%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}.
 We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
-	\item A reactive simulation of web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
+	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
 	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
-	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Marl++}, \emph{SPF}), improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}) and discussion of security concerns and real-world deployment (\cref{sec:discussion}).
+	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
+	\item An empirical evaluation of our new models alongside the state-of-the-art in RL-based DDoS mitigation with these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
 \end{itemize}
@@
 
-\section{DDoS Mitigation with Per-flow Reinforcement Learning}
+\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
 Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
@@
 %		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
-		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
 %		?? REWORK/MAKE ACCURATE
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-udp-box-single}
-
-	\caption{
-		Online performance for Opus benign traffic in a single-destination network, single-agent mode.
-		?? DO I need this?
-		\label{fig:udp-tree-box-single}
-	}
-\end{figure}
-
-Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
-\Cref{fig:udp-box} shows the behaviour for all choices of $n$---specifically, we see that Marl++ typically remains ahead of Marl, though this performance difference shrinks as $n$ increases, while SPF performs better at high host-density.
-
-?? Discuss multi-dest topol once results available.
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-udp-box-single}
+%	
+%	\caption{
+%		Online performance for Opus benign traffic in a single-destination network, single-agent mode.
+%		?? DO I need this?
+%		\label{fig:udp-tree-box-single}
+%	}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-udp-box-separate}
+%	
+%	\caption{
+%		Online performance for Opus benign traffic in a multi-destination network, multi-agent mode.
+%		?? DO I need this?
+%		\label{fig:udp-ecmp-box}
+%	}
+%\end{figure}
+%
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-udp-box-single}
+%	
+%	\caption{
+%		Online performance for Opus benign traffic in a multi-destination network, single-agent mode.
+%		?? DO I need this?
+%		\label{fig:udp-ecmp-box-single}
+%	}
+%\end{figure}
+
+In a single-destination network, we observe that Marl's performance degrades as $n$ increases.
+Typically, our \emph{Instant} agent design achieves the best performance in multi-agent mode, reducing collateral damage above the current state-of-the-art, but sharply degrades at low $n$ when agents share experience.
+We see a reversal of this trend for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
+%Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
+\Cref{fig:udp-tree-box} shows the preserved traffic in multi-agent mode.
+
+?? Discuss multi-dest topol once all results available.
 
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-separate}
-
-	\caption{
-		Online performance for HTTP benign traffic in a single-destination network, multi-agent mode.
-		Marl and Marl++ begin with essentially identical efficacy, with Marl++ having a small advantage at larger $n$.
-		SPF is slightly worse than Marl and Marl++ at low $n$, but achieves considerably better median performance and lower variance than the other models for large $n$.
-		\label{fig:tcp-tree-box}
-	}
-\end{figure}
-\begin{figure}
-	\centering
 	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-single}
@@
 		Online performance for HTTP benign traffic in a single-destination network, single-agent mode.
-		?? DO I need this?
-		\label{fig-tcp-tree-box-single}
+		\emph{Instant} and \emph{Guarded} exhibit essentially identical efficacy at $n~=~2$, protecting less traffic than Marl.
+		Only \emph{Guarded}'s performance rapidly increases with $n$, achieving considerably better median performance and lower variance than the other models.
+		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
+		\label{fig:tcp-tree-box}
 	}
 \end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-single}
+%	
+%	\caption{
+%		Online performance for HTTP benign traffic in a single-destination network, single-agent mode.
+%		?? DO I need this?
+%		\label{fig-tcp-tree-box-single}
+%	}
+%\end{figure}
 \begin{figure}
@@
 		Online performance of standard and single-agent models in a single-destination network with $n=16$ hosts per egress point and HTTP benign traffic.
-		At this level of host density, Guarded reaches higher peak performance sooner and is considerably more consistent throughout the episode.
-		Guarded benefits greatly from information sharing, converging to protect around \SI{75}{\percent} of TCP traffic within \SI{100}{\second}.
-		The Instant model converges to perform similarly to Marl.
+		At this level of host density, \emph{Guarded} reaches higher peak performance sooner and is considerably more consistent throughout the episode.
+		\emph{Guarded} benefits greatly from information sharing, converging to protect around \SI{75}{\percent} of TCP traffic within \SI{100}{\second}.
+		The \emph{Instant} model converges Marl's level of performance.
 %		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
@@
 
-From \cref{fig:tcp-box}, Marl++ consistently allows more legitimate traffic to reach its destination than Marl, with both significantly above the baseline (\cref{tab:av-vals}, \emph{Capped}, TCP).
-SPF performs slightly worse at low $n$, but outperforms both models as host density increases.
-\Cref{fig:tcp-16} shows a trend in our results: for larger $n$, Marl++ requires around twice as long as Marl to learn its best-achieving policy, under-performing for this period, but achieves notably better peak and average performance.
-SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
-
-?? Discuss multi-dest topol once results available.
-
-\subsection{Single-agent performance}
-Making all decisions with a single agent is roughly equivalent to having a zero-cost communication channel between each pair of agents, theoretically allowing faster training by giving each agent more experience.
-Curiously, we observe that this often leads to drastically worse policies when used as part of Marl++ for small $n$, but makes SPF a considerably more competitive model---especially for TCP, and as $n$ grows larger (\cref{fig:tcp-16}).
-Single-agent SPF almost consistently outperforms Marl.
-We discuss our conjectures for why this reversal occurs in \cref{sec:discussion}.
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-tcp-box-separate}
+%	
+%	\caption{
+%		Online performance for HTTP benign traffic in a multi-destination network, multi-agent mode.
+%		?? DO I need this?
+%		\label{fig:tcp-ecmp-box}
+%	}
+%\end{figure}
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-tcp-box-single}
+%	
+%	\caption{
+%		Online performance for HTTP benign traffic in a multi-destination network, single-agent mode.
+%		?? DO I need this?
+%		\label{fig:tcp-ecmp-box-single}
+%	}
+%\end{figure}
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-tcp-16-single}
+%	
+%	\caption{
+%		?? Eh
+%		\label{fig:tcp-ecmp-16}
+%	}
+%\end{figure}
+
+\Cref{tab:av-vals} shows that Marl offers a low (though fairly consistent) level of protection for TCP traffic, which the \emph{Instant} agent offers no substantial improvement over.
+However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
+\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
+%SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
+
+%?? Discuss multi-dest topol once results available.
+We observe that the same trends present themselves in the multi-destination topology: \emph{Guarded} remains the best fit for TCP, in both training modes.
+Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
+
+%\subsection{Single-agent performance}
+%Making all decisions with a single agent is roughly equivalent to having a zero-cost communication channel between each pair of agents, theoretically allowing faster training by giving each agent more experience.
+%Curiously, we observe that this often leads to drastically worse policies when used as part of Marl++ for small $n$, but makes SPF a considerably more competitive model---especially for TCP, and as $n$ grows larger (\cref{fig:tcp-16}).
+%Single-agent SPF almost consistently outperforms Marl.
+%We discuss our conjectures for why this reversal occurs in \cref{sec:discussion}.
 
@@
 
+%?? New angle -- \emph{Guarded} offers substantal improvements for \emph{most} internet traffic according to the stats we know. Is accurate protection of beningn UDP traffic actually more difficult?
+
 %Talk about flaws here, what could go wrong...
@@
 \fakepara{Model performance}
-Of the results presented, SPF's unpredictable (often worse) starting performance is unexpected, given its far smaller action space.
+Of the results presented, \emph{Guarded}'s unpredictable (often worse) starting performance is unexpected, given its far smaller action space.
 It's natural to expect that this would make the model easier to learn, but the additional state required appears to make the task \emph{harder}, beyond even the value of choosing a non-zero discount factor (adding forward-planning to explicitly mitigate this effect).
-Accordingly, we see that SPF performs best when agents learn from as much knowledge as possible: high $n$ and single-agent training.
-Why does it exhibit considerably lower variance?
-Our belief is that SPF is, in fact, a considerably stronger model for these reasons, and its successes give us particular cause to consider the best schemes for efficient information sharing.
+Accordingly, we see that this design performs best (and exhibits considerably lower variance) when agents learn from as much knowledge as possible: high $n$ and single-agent training.
 To filter incoming traffic from a source, it must decide to degrade their inbound traffic multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
-Paradoxically, Marl++ generally achieves the best performance yet actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.
+Our belief is that \emph{Guarded} is a considerably stronger model for these reasons, and its successes offer strong rationale to consider the best schemes for efficient information sharing.
+Paradoxically, \emph{Instant} generally achieves the best performance for UDP traffic yet actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.
 
 %?? May be hard to learn multiple features at once while controlling multiple flows while contending with many more agents, with harder dynamics like TCP. Does this hinder learning in the long run?
-Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect, or hope for.
+Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect, or hope for in the case of UDP traffic.
 %??
@@
 We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to observe the true value of each action), and controlling shared global state.
-Our pre-training results support the idea that it may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows.
+Moreover, benign UDP traffic shares many characteristics with attack traffic, suggesting that more training samples or some unknown feature might aid control, or that it may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows.
 
 %It is likely that the design of Marl++ 
+%?? Need to mention that Marl performs lower than their paper's numbers for TCP...
 
@@
 What we offer above the alternate approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
+The techniques we propose offer substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today.
 Moreover, our decision making processes are entirely agnostic of the protocol or content of inbound traffic, offering future-proofing against the introduction of new transports.
@@
 With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
-\Cref{fig:intern-16-tcp-iat,fig:intern-16-tcp-something} show how we may visualise the portion of a policy for each feature, and describe what information can be gained from doing so.
+%\Cref{fig:intern-16-tcp-iat,fig:intern-16-tcp-something} show how we may visualise the portion of a policy for each feature, and describe what information can be gained from doing so.
 Furthermore, system state which is dependent on many signals drawn from across a wide network (such as our global state) is difficult to exert precise control over.
@@
 
-Where overheads are concerned, the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies: \SIrange{16.5}{17.0}{\kibi\byte} from our experiments.
-As we have described earlier, action updates require a constant number of floating point operations---\num{122} floating point additions and \num{2} multiplications per update of $\wvec{}$, above the \num{140} additions required to choose an action.
+Where overheads are concerned, the state space sizes guarantee that an \emph{Instant} agent's policy remains under \SI{520}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies: $\sim$\SI{17.8}{\kibi\byte} from our experiments.
+\emph{Guarded} policies are $3/10$ of the size.
+As we have described earlier, action updates require a constant number of floating point operations---\num{160} floating point additions and \num{32} multiplications per update of $\wvec{}$ with per-tile updates, above the \num{160} additions required to choose an action.
 The vast majority of these operations can be vectorised trivially, if such hardware is present.
-Action computation for SPF is cheaper still, requiring only \num{42} additions per action required.
+Action computation for \emph{Guarded} agents is cheaper still, requiring only \num{48} additions per action required.
 Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
@@
 
-?? Actually talk about Marl (???) to appease reviewer \#1.
+%?? Actually talk about Marl (???) to appease reviewer \#1.
 The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15} (whom we have positioned our work against), and their contribution in applying RL to the task of intrusion prevention is significant: their work helps to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
-?? Recap their flaws, since they've been cut form every other aspect.
+They create a tree overlay topology (subdivided into teams), where each agent applies packet drop to \emph{all} flows inbound to a protected server.
+%?? Recap their flaws, since they've been cut form every other aspect.
+Our results show that their technique underperforms at high host density and when congestion-aware traffic dominates---that their results do not demonstrate this suggests an evaluation driven purely by traces (rather than live application dynamics).
 
@@
 %Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
-Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
-??
-Flaws?
-what can't it do that we can do better?
-Long times to compute TBE routes on real networks, low expansion factor in real network can require more ``rounds'' of filtering.
-(``takes only'' \SI{14}{\second} for Cogent network?
-gross.)
-?? I believe their assumptions may not extend to CBR traffic (e.g., UDP VoIP traffic), which we know makes up a sizeable proportion of network traffic. Also DASH traffic.
+Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ASes, although some of our feature choices are backed by similar observations.
+A weakness of their approach is that computing a route to measure bandwidth expansion on real networks can be costly (up to \SI{14}{\second} for the Cogent topology), and that the low expansion factors in real network can require more ``rounds'' of filtering.
+Their assumptions about traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), which we know makes up a sizeable proportion of network traffic.
 
@@
 A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
-We feel there is a shortage of work in this category at present, at least in the field of networks.
-To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.
+%We feel there is a shortage of work in this category at present, at least in the field of networks.
+To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}.
 
@@
 
-\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
+\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation.
 Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
@@
 \section{Conclusions and Future Work}
-
 Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
@@
 Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
-Our evaluation shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, with SPF showing the most promise for future evaluation.
+Our evaluation shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, with \emph{Guarded} agents showing the most promise for future evaluation.
 
@@
 %?? IDEA: try out average reward, TD($\lambda$) methods as future work...
+%A natural research direction to enhance this work would be the combination of the classic function approximation techniques we make use of alongside the improved algorithms that the RL community has introduced in the past few years.
+%Actor-critic methods, or algorithms based on eligibility traces are good candidates for investigation. 
 The remaining weaknesses invite many further improvements worth investigation.
-A natural research direction to enhance this work would be the combination of the classic function approximation techniques we make use of alongside the improved algorithms that the RL community has introduced in the past few years.
-Actor-critic methods, or algorithms based on eligibility traces are good candidates for investigation.
-
 %??
@@
 If true online learning is desired (i.e., coping with a non-stationary environment), then such reward functions are sorely needed.
-While $\bload{\cdot}$ is likely to be a good candidate for many deployments, we believe that internal ``canary flows'' (or some other active measurement) will be effective sources of data to base a new reward function upon.
+While $\bload{\cdot}$ is likely to be a good candidate for many deployments, we believe that finding an effective metric derived from the individual statistics we suggested serves as an interesting research problem.
 
@@
 Given that one of the advantages that RL methods can offer is the ability to handle non-stationary problems, it is important to propose and test sensible simulations or captures of evolving networks against these methods.
-These scenarios present ideal circumstances to apply work on adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}, changepoint detection, or intelligent/heuristic sampling methods to judge which flows are most worthy of consideration per tick.
+It is known, for instance, that DDoS attack strategies evolve in real time \cite{DBLP:conf/spw/KangGS16}; evaluation is difficult at present, since no works detail what patterns such evolution might take.
+Regardless, these scenarios present ideal circumstances to apply work on adaptive exploration \cite{DBLP:conf/annpr/TokicP12}, changepoint detection, or intelligent/heuristic sampling methods to judge which flows are most worthy of consideration per tick.
 For estimating \emph{when} to explore, we believe that the intersection of signal processing and RL is as-yet unexplored.
@@
 Effective real-world deployment of RL-based defences cannot assume that switches in a network will support a custom version of OVS or other arbitrary software, introducing the question of whether agent training, execution and distribution may be possible when using \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
-We also expect it will be fruitful to look into \emph{how} agents may share knowledge with one another, such as via \emph{knowledge bases} or some other efficient communication mechanism.
+We also expect it will be fruitful to look into \emph{how} agents may share knowledge with one another.
 
@@
 \printbibliography
-
+\vspace{-0.5cm}
+\vskip 0pt plus -1fil
 \begin{IEEEbiography}
@@
 \end{IEEEbiography}
-
+\vspace{-0.5cm}
+\vskip 0pt plus -1fil
 \begin{IEEEbiography}

### 2019-05-13 tables/tnsm-ecmp-avg-reward.py

@@ -63,7 +63,7 @@ def write_table(out_file, data):
 	for traffic in traffics:
 		for i, n in enumerate(ns):
 			key = (traffic,n)
-			out_file.write("{} & {} ".format("" if i>0 else traffic.upper(), n))
+			out_file.write("{} & {} ".format("" if i>0 else traffic.upper(), n * 3))
 
 			row = data[key]
 			for j, _ in enumerate(row):

### 2019-05-13 tables/tnsm-ecmp-avg-reward.py
Last multi-dest results

@@ -12,7 +12,7 @@ variants = [
 ]
 
 illegal_combos = set()
-illegal_combos.add(("spf", "opus", "single"))
+#illegal_combos.add(("spf", "opus", "single"))
 # illegal_combos.add(("spf", "udp", "channel"))
 # illegal_combos.add(("spf", "tcp", "channel"))
 

### 2019-05-14 paper/paper.tex
Post proof-reading

@@
 Normal and attack patterns shift as new protocols and applications are introduced, further compounded by burstiness and seasonal variation.
-Accordingly, application of machine learning-based techniques is difficult.
+Accordingly, it is difficult to apply machine learning-based techniques and defences in practice.
 %Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.
@@
 
-Our results show the existence of flow features with high predictive power for different traffic classes, when a basis for feedback-loop-like control.
+Our results show the existence of flow features with high predictive power for different traffic classes, when used as a basis for feedback-loop-like control.
 We show that the new RL agent models can offer a significant increase in goodput of legitimate TCP traffic for many choices of host density.
@@
 %?? Expand as required
-What approaches do exist are aimed towards the task of adaptive online DDoS mitigation, and rely upon learning control of probabilistic packet drop.
+What approaches do exist are aimed towards the task of adaptive online DDoS mitigation, and rely upon learning to control probabilistic packet drop.
 %?? THAT IS A MAJOR CONTRIB, MENTION IT EVERYWHERE YOU CAN
@@
 The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
-In tandem with the development and evaluation of an effective state-space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
+In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
 Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
@@
 	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
-	\item An empirical evaluation of our new models alongside the state-of-the-art in RL-based DDoS mitigation with these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
+	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
 \end{itemize}
@@
 \emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
-Attackers achieve this either by exploiting peculiarities of operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
+Attackers achieve this by either exploiting peculiarities of operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
 Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
@@
 	\centering
-	\begin{subfigure}{0.4\linewidth}
+	\begin{subfigure}{0.45\linewidth}
 		\resizebox{\linewidth}{!}{
@@
 
-			\draw[step=0.5cm,color=uofgpumpkin,opacity=0.7,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
+			\draw[step=0.5cm,color=uofgpumpkin,opacity=0.9,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
 			\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
-			\node[color=uofgpumpkin] at (0.25,-0.95) {\footnotesize Tiling 2};
+			\node[color=uofgpumpkin!50!uofgrust] at (0.25,-0.95) {\footnotesize Tiling 2};
 
@@
 	\end{subfigure}%
-	\begin{subfigure}{0.4\linewidth}
+	\begin{subfigure}{0.45\linewidth}
 		\resizebox{\linewidth}{!}{
@@
 This combination of algorithm and coding strategy is well-optimised, if actions are discrete; this allows a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each tile.
-Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}}-1)$ floating point additions per decision.
+Action values for any state are then obtained by summing the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}}-1)$ floating point additions per decision.
 Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
@@
 \Cref{fig:tilecode-update} shows how this applies to our prior example.
-If desired we may define a state-space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
+If desired we may define a state space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
 Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.
@@
 As exploration still occurs for each action, this approach reduces $\epsilon$ multiple notches every timestep.
-In turn, we increase the annealing window for $\epsilon$ by a factor of 8 so as to preserve exploration over time, by accounting for the greater volume of decisions being made.
+In turn, we increase the annealing window for $\epsilon$ by a factor of \num{2.67} so as to preserve exploration over time, by accounting for the greater volume of decisions being made.
 
@@
 	\centering
-	\resizebox{0.7\linewidth}{!}{
+	\resizebox{0.8\linewidth}{!}{
 		\begin{tikzpicture}[
@@
 This is implemented by installing an action via OpenFlow, instructing its host switch to drop each relevant packet with probability $p$.
-We choose to drop packets rather than impose traffic limits as it affords us a discrete action space without prior knowledge of traffic characteristics or measurement.
+We choose to drop packets rather than impose traffic limits as it offers us a discrete action space without prior knowledge of traffic characteristics or measurement.
 Furthermore, we need not consider burstiness, fairness or implementation costs (such as per-flow bucket sizes) which could limit scalability.
@@
 The measurements of \textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
-We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, choosing to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
+We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$,.
+The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
+
+We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
 The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
@@
 
-The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
 To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$); this is chosen above the equivalent for UDP due to TCP's higher prevalence.
-Most importantly, an agent must choose to punish a flow multiple times in succession to cause rapid degradation, reducing variance while allowing an agent to see how a host reacts to structured changes in the environment.
+Most importantly, an agent must now choose to punish a flow multiple times in succession to cause rapid degradation, reducing variance while allowing an agent to see how a host reacts to structured changes in the environment.
 
 As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
-We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing with UDP traffic.
+We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing.
 
@@
 Our mode of action means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
-This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p$ is nonzero in any real network) while constant bitrate UDP traffic is proportional to $1 - p$.
+This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p$ is nonzero in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
 %It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
-This weakness still presents itself in modern flavours of TCP, such as TCP Cubic which depends on $1/p^{0.75}$ \cite{rfc8312}.
+This weakness still presents itself in modern flavours of TCP, such as TCP Cubic which in turn has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
 This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
@@
 Taking many actions per-timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
-This introduces some potential issues: inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated; at their worst, these present additional attack surfaces to an adversary.
+This introduces some potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated.
+At their worst, these risks present additional attack surface to an adversary.
 To adapt to these problems, we make use of \emph{timed random sequential} updates.
@@
 \fakepara{Global state}
-These are the vector of load measurements along a flow's path introduced in \cref{sec:feature-space}.
-These indicate the overall health of the network, and crucially are all measurements which an agent directly controls.
+This is the vector of load measurements along a flow's path introduced in \cref{sec:feature-space}.
+These values indicate the overall health of the network, and crucially are all measurements which an agent directly controls.
 
@@
 This feature also potentially allows forgiveness, serving as a reference point for determining whether a source mistakenly marked as malicious exhibits different falloff behaviour after punishment.
-It's important to note that this feature only makes sense once combined with another flow feature.
+It's important to note that this feature only makes sense once combined with another flow feature, and never appears individually tile-coded.
 
@@
 \Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), using a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
-\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends being observed when the same techniques is applied with UDP traffic (omitted).
+\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends being observed when the same technique is applied with UDP traffic (omitted).
 %?? Core findings---different protocols need different features, so everything we proposed above has a use!
@@
 First and foremost, we are interested in representative load and packet inter-arrival characteristics and in how these characteristics evolve in response to actions.
-We introduce these models because are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
+We introduce these models because we are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
 %Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
@@
 All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
+If several ports offer different (equal-length) paths to a destination, a consistent random port is chosen from the flow-hash by an OpenFlow \emph{Group action} (in \emph{select} mode).
 If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
-The controller is also responsible for computing how switches respond to ARP requests: the necessity of this arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
+The controller is also responsible for computing how switches respond to ARP requests: this need arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
 %We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.
@@
 %?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
-To model legitimate TCP traffic, the server nodes run an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
+To model legitimate TCP traffic, server nodes run an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
 Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
@@
 \subsection{UDP (Opus) Traffic Model}
-VoIP traffic exhibits very different characteristics to the above model; packet arrivals are highly periodic due to real-time application requirements, flows have a constant bitrate, and flow behaviour which does not react substantially to lost packets.
+VoIP traffic exhibits very different characteristics to the above model; packet arrivals are highly periodic due to real-time application requirements, flows have a constant bitrate, and do not react substantially to lost packets.
 Interestingly, DDoS attack traffic is known to share many of these characteristics, offering an interesting detection problem.
@@
 Hosts send RTP traffic with Salsa20 encrypted payloads---audio frames of length \SI{20}{\milli\second} at a target \SI{96}{\kilo\bit\per\second}.
-We generate similar traffic at hosts by replaying anonymised traces gathered in general and tabletop RPG servers; each trace contains only the size of each audio payload, entries denoting missed packets, and the duration of silent periods.
-We trim these silent periods to a maximum \SI{5}{\second} due to the lengthy talk/silence bursts introduced by users in RPG servers.
+We generate similar traffic at hosts by replaying anonymised traces gathered in general use and tabletop RPG servers; each trace contains only the size of each audio payload, entries denoting missed packets, and the duration of silent periods.
+We trim these silent periods to a maximum \SI{5}{\second} due to the lengthy talk/silence bursts introduced by users in RPG servers, and estimate the size of missed packets by taking an exponentially-weighted moving average over known sizes.
 Hosts punctuate audio frames with a 4-byte keepalive every \SI{5}{\second}.
-All traffic passes over a central server which groups hosts into rooms, and is forwarded to other participants.
+All traffic passes over a central server which groups hosts into rooms, and is forwarded to other participants; we do not replicate pre-call Websocket traffic which would be used for authentication.
 There is no peer-to-peer traffic---the server acts as a TURN relay \cite{rfc5766} for all hosts.
@@
 We find that each flow occupies an expected \SI{52.4}{\kilo\bit\per\second} upstream bandwidth.
-To match the target upload rate required at a host, we run enough sessions to meet the target data rate.
+To match the target upload rate assigned to each host, it runs enough individual sessions to meet the target data rate.
 
@@
 %Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
-These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
+%These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
 
@@
 	\centering
-	\resizebox{0.6\linewidth}{!}{
+	\resizebox{0.8\linewidth}{!}{
 		\begin{tikzpicture}[
@@
 It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
-The fat-tree topology \cite{DBLP:conf/sigcomm/Al-FaresLV08} sees regular use in real-world datacentres and scales well horizontally.
+In contrast, the fat-tree topology \cite{DBLP:conf/sigcomm/Al-FaresLV08} sees regular use in real-world datacentres and scales well horizontally.
 
@@
 \subsection{Parameters}
-The algorithm parameters were set at $\alpha=0.05$, linearly annealing $\epsilon=0.2 \rightarrow 0$ by $t=3000$ in the case of the Instant model (\num{12000} actions in the Guarded model).
+The algorithm parameters were set at $\alpha=0.05$, linearly annealing $\epsilon=0.2 \rightarrow 0$ by $t=3000$ in the case of Marl (\num{8000} actions per agent in the \emph{Instant/Guarded} models).
 
@@
 %?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
-Malicious hosts sent \SIrange{2.5}{6}{\mega\bit\per\second} when testing UDP traffic, though when using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows.
+Malicious hosts sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, though when using TCP-like traffic, each attacker's upload range had to be increased to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows.
 Given $n$ and $\operatorname{P}(\mathit{malicious})$, we see an expected malicious bandwidth \numrange{1.27}{1.87} and \SIrange{2.03}{2.18}{$\! \times U_s$} respectively.
@@
 
-?? Discuss multi-dest topol once all results available.
+%?? Discuss multi-dest topol once all results available.
+When defending multiple destinations, we see a sharp decrease in the effectiveness of all agent designs.
+Our \emph{Instant} and \emph{Guarded} designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
 
@@
 		Only \emph{Guarded}'s performance rapidly increases with $n$, achieving considerably better median performance and lower variance than the other models.
-		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
+		The longer tails of outliers typically indicate the longer training time the new models require---we observe that \emph{Guarded} typically has considerably lower variance once it has converged on a stable policy.
 		\label{fig:tcp-tree-box}
@@
 		\emph{Guarded} benefits greatly from information sharing, converging to protect around \SI{75}{\percent} of TCP traffic within \SI{100}{\second}.
-		The \emph{Instant} model converges Marl's level of performance.
+		The \emph{Instant} model converges to Marl's level of performance.
 %		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
@@
 \Cref{tab:av-vals} shows that Marl offers a low (though fairly consistent) level of protection for TCP traffic, which the \emph{Instant} agent offers no substantial improvement over.
-However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
-\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
+However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$\!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
+\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
 %SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
@@
 %?? May be hard to learn multiple features at once while controlling multiple flows while contending with many more agents, with harder dynamics like TCP. Does this hinder learning in the long run?
-Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect, or hope for in the case of UDP traffic.
+Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect in the case of UDP traffic.
 %??
@@
 The most likely explanation is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
-The increased state-space size makes this a more likely occurrence, as is the unclear affect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state-space.
+The increased state space size makes this a more likely occurrence, as is the unclear affect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state space.
 We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to observe the true value of each action), and controlling shared global state.
+Although our design does take steps to counteract such effects, these mitigations may not be enough.
 Moreover, benign UDP traffic shares many characteristics with attack traffic, suggesting that more training samples or some unknown feature might aid control, or that it may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows.
@@
 Where overheads are concerned, the state space sizes guarantee that an \emph{Instant} agent's policy remains under \SI{520}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies: $\sim$\SI{17.8}{\kibi\byte} from our experiments.
-\emph{Guarded} policies are $3/10$ of the size.
+\emph{Guarded} policies are \SI{30}{\percent} of this size.
 As we have described earlier, action updates require a constant number of floating point operations---\num{160} floating point additions and \num{32} multiplications per update of $\wvec{}$ with per-tile updates, above the \num{160} additions required to choose an action.
 The vast majority of these operations can be vectorised trivially, if such hardware is present.
-Action computation for \emph{Guarded} agents is cheaper still, requiring only \num{48} additions per action required.
+Action computation for \emph{Guarded} agents is cheaper still, requiring only \num{48} additions per action.
 Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
@@
 However, the measurements acquired in such a scenario are far more likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
-The main bottlenecks are most likely in forwarding the load measurements from various aggregation points (which can be made more efficient through multicast) and in running some estimator to condition the reward function upon.
+The main bottlenecks are most likely in forwarding the load measurements from various aggregation points (which can be made more efficient through multicast) and in running some estimator $\operatorname{g}(\cdot)$ to condition the reward function.
 %?? Potential for dividing up pre-feature training across different agents, or train like this locally (sets of flows trained by 1 sub-model).
@@
 A weakness of their approach is that computing a route to measure bandwidth expansion on real networks can be costly (up to \SI{14}{\second} for the Cogent topology), and that the low expansion factors in real network can require more ``rounds'' of filtering.
-Their assumptions about traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), which we know makes up a sizeable proportion of network traffic.
+Their assumptions about traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), both of which make up a sizeable proportion of network traffic.
 
@@
 Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works don't apply the rich literature on the subject to its fullest potential.
+Given that one of the main strengths of RL techniques is the ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works don't apply the rich literature on the subject to its fullest potential.
 
@@
 Our evaluation shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, with \emph{Guarded} agents showing the most promise for future evaluation.
+
+The most direct improvements to be made lie in the correct protection of legitimate UDP traffic, which our agent designs have difficulty safeguarding.
+Outside of this, there is scope to test these new techniques against link-flooding attacks in large-scale topologies using reward functions such as \cref{eqn:lfa-reward}.
+Simulation is the most likely avenue for such evaluation.
 
@@
 Given that one of the advantages that RL methods can offer is the ability to handle non-stationary problems, it is important to propose and test sensible simulations or captures of evolving networks against these methods.
-It is known, for instance, that DDoS attack strategies evolve in real time \cite{DBLP:conf/spw/KangGS16}; evaluation is difficult at present, since no works detail what patterns such evolution might take.
+While it is known that DDoS attack strategies evolve in real time \cite{DBLP:conf/spw/KangGS16}, evaluation is difficult at present since no works detail what patterns such evolution might take.
 Regardless, these scenarios present ideal circumstances to apply work on adaptive exploration \cite{DBLP:conf/annpr/TokicP12}, changepoint detection, or intelligent/heuristic sampling methods to judge which flows are most worthy of consideration per tick.
@@
 We hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
-It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, potentially more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques rely upon.
+It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques rely upon.
 It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.

### 2019-05-14 paper/paper.tex
Final typo

@@
 The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
-This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and non-congestion-aware traffic as required.
+This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.
 

### 2019-07-23 code/marl/tnsm-baseline-runner.sh
Baseline results for topology/traffic combos...

@@ -1,0 +1,23 @@
+#!/bin/bash
+SCRIPT="./tnsm-baseline-results.py"
+
+getCount() {
+	python $SCRIPT $1
+}
+runExpt() {
+	python $SCRIPT $1 $2
+}
+
+./purger.sh
+
+for model in `seq 0 2`;
+do
+	for i in $(seq 0 $(expr $(getCount $model) - 1));
+	do
+		./purger.sh
+		runExpt $model $i
+	done
+done
+
+./purger.sh
+

### 2019-07-23 paper/infocom.tex
Begin Infocom paper
Some fixes applied, new section laid out. Then it will be culling
time...

--- 2019-05-14 paper/paper.tex
@@
 % !BIB program = biber
-\documentclass[10pt, times, comsoc]{IEEEtran}
+\documentclass[10pt, times, conference, letterpaper]{IEEEtran}
 \usepackage{nopageno}
@@
 tracking=true,
-%kerning=true,spacing=true
+kerning=true,spacing=true
 ]{microtype}
@@
 %\newcommand{\mytitle}{Improving Direct-Control Reinforcement Learning for Network Intrusion Prevention}
-\newcommand{\mytitle}{Per-Host DDoS Mitigation by Direct-Control Reinforcement Learning}
+\newcommand{\mytitle}{Protocol-Agnostic DDoS Mitigation through Reinforcement Learning}
 
@@
 	pdftitle={\mytitle{}},
-	pdfauthor={Kyle A. Simpson, Simon Rogers, Dimitrios P. Pezaros}
-%	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
+%	pdfauthor={Kyle A. Simpson, Simon Rogers, Dimitrios P. Pezaros}
+	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
 }
@@
 %	Kyle~A.~Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
-	Kyle~A.~Simpson,
-	Simon~Rogers,
-	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
-	School of Computing Science, University of Glasgow, Glasgow, G12 8RZ, UK\\
+%	Kyle~A.~Simpson,
+%	Simon~Rogers,
+%	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
+%	School of Computing Science, University of Glasgow, Glasgow, G12 8RZ, UK\\
 %	University of Glasgow, Glasgow, Scotland,\\
-	\email{k.simpson.1@research.gla.ac.uk}
-%Anonymous Giraffe,
-%Anonymous Badger,
-%Anonymous Ibex\\
-%Unnamed Department, Nowhere\\
-%\email{giraffe.a@unnamed.com}
+%	\email{k.simpson.1@research.gla.ac.uk}
+Anonymous Giraffe,
+Anonymous Badger,
+Anonymous Ibex\\
+Unnamed Department, Nowhere\\
+\email{giraffe.a@unnamed.com}
 }
@@
 \begin{abstract}
-%?? Trim to 75--200 words, simplify.
-DDoS attacks plague the availability of online services today, yet like many cybersecurity problems are evolving and non-stationary.
-Normal and attack patterns shift as new protocols and applications are introduced, further compounded by burstiness and seasonal variation.
-Accordingly, it is difficult to apply machine learning-based techniques and defences in practice.
-%Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.
-
-\emph{Reinforcement learning} (RL) may overcome this detection problem for DDoS attacks by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available) in an online manner.
-%We examine how the paradigm has been applied to the task of \emph{Distributed Denial of Service} (DDoS) attack prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.
-We advance the state-of-the-art in RL-based DDoS mitigation by introducing two agent classes designed to act on a per-flow basis, in a protocol-agnostic manner for any network topology.
-%These flaws---poor handling of TCP traffic and high host density at egress points---inspire and direct an exploration of the design space of RL models.
+DDoS attacks still cause significant harm to online services to this day.
+Their detection is difficult, because like many cybersecurity problems they \emph{evolve}: normal and attack patterns shift due to new protocols and applications.
+Such evolution makes it difficult to apply machine learning techniques and defences.
+
+Reinforcement learning (RL) agents manage and monitor system health in an online manner.
+Unsupervised, online learning is a key part of overcoming this detection problem.
+We present the architecture of a system designed to host such RL agents in a typical network.
+Furthermore, we introduce two agents which advance the state-of-the-art in using RL for DDoS mitigation.
+Both of these act per flow, in a protocol-agnostic manner for any network topology.
 This is supported by an in-depth investigation of feature suitability and empirical evaluation.
 
-Our results show the existence of flow features with high predictive power for different traffic classes, when used as a basis for feedback-loop-like control.
-We show that the new RL agent models can offer a significant increase in goodput of legitimate TCP traffic for many choices of host density.
-
+Our results show the existence of flow features with high predictive power for bulk TCP and VoIP traffic.
+We show that the new agents offer a significant increase in throughput of legitimate TCP traffic for many choices of host density.
 \end{abstract}
 
-\begin{IEEEkeywords}
-	Security services, distributed denial-of-service, software-defined networking, machine learning, reinforcement learning.
-\end{IEEEkeywords}
+%\begin{IEEEkeywords}
+%	Security services, distributed denial-of-service, software-defined networking, machine learning, reinforcement learning.
+%\end{IEEEkeywords}
 
 \section{Introduction}
+?? Redo intro to be more system focussed? Needs to be suitably different.
 
@@
 Preliminary experiments found that Sarsa offered the best performance and behaviour.
+?? Link plot or table w Sarsa vs. Q results.
 
@@
 While the standard formulation of \cref{eqn:sg-sarsa} updates the value of all tiles identically (by a scalar $\alpha \delta_t$), we found it more effective to compute a different temporal difference value \emph{for each tiling}.
+This adjusts the value of each activated tile with a different magnitude and direction, correcting tilings which contribute too much or too little value without disturbing estimates which are already close to correct.
 While we make use of the sum of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to set $\alpha$ to a higher value without divergence.
@@
 When learning control on the basis of a high-dimensional, tile-coded state space, assignment of credit for each decision is difficult (because all tiles have identical gradient).
-To combat this, with probability $\epsilon$ an agent will mark a flow as being governed by a subset of the state space for the next 5 decisions.
+Rephrasing this, assigning credit to the \emph{features} which drove each decision is difficult.
+To combat this, with probability $\epsilon$ an agent will mark a flow as being governed by a subset of the state space for the next five decisions.
 Each agent chooses actions on that source/destination pair using one element of local state, the global state, and the bias tile---we include the latter two to strike a balance between and accuracy and correct credit assignment.
@@
 Furthermore, we need not consider burstiness, fairness or implementation costs (such as per-flow bucket sizes) which could limit scalability.
-We offer two models on how to choose $p$:
+We offer two models, which differ only on how to choose $p$.
+The intuition behind these designs is that an agent may be able to directly choose the level of packet drop at the expense of a larger action space (\emph{Instant control}), or that we may be able to reduce the action set (and learn a policy within fewer actions) by making use of domain knowledge (\emph{Guarded control}).
 
@@
 The measurements of \textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
-We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$,.
+We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
 The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
@@
 We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing.
+While beneficial for legitimate flows who behave similarly to attack traffic, this introduces a lag of an extra two actions in blocking illegitimate flows.
 
@@
 This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
-Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
+Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{Repository link removed for double-blind reviewing.}.
+%Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
 QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
@@
 %This manifests in two ways: the best-achievable performance drops, and so too does the learning rate.
-Given the probability that a host is legitimate, $P_G \in [0,1]$, it follows that a host will be malicious with probability $P_B = 1 - P_G$.
-Defining \emph{imperfect service} to mean any case where all $n$ hosts connecting over a switch do not share the same classification (i.e., a mixture), then the probability that a switch is delivering imperfect service is $P_{M,n} = 1 - (P_G^n + P_B^n)$.
+Given the probability that a host is legitimate, $P_G \in [0,1]$,  if it is assumed that a host may be only legitimate or malicious then it follows that it will be malicious with probability $P_B = 1 - P_G$.
+Supposing we have a set of $n$ hosts, the probability that all hosts will be legitimate is $P_G^n$ (resp. $P_B$ for malicious).
+Defining \emph{imperfect service} $P_{M,n}$ to mean any case where all $n$ hosts connecting over a switch do not share the same classification (i.e., a mixture), then the probability that a switch is delivering imperfect service is $P_{M,n} = 1 - (P_G^n + P_B^n)$.
 \begin{thm}
@@
 \begin{proof}
-	\emph{Base case:} $P_{M,1}=0, P_{M,2} = 1 - P_G^2 - P_B^2 > 0$.
+	\emph{Base case:} $P_{M,1}=0, P_{M,2} = 1 - P_G^2 - P_B^2 \ge 0$.
+
 	\emph{Inductive step:} Assume that the theorem holds for $n$. Observe that $P_G^n \ge P_G^{n+1}$ (resp.\ $P_B$). It then follows that:
@@
 
-\subsection{Systems Considerations}
+\subsection{Systems Considerations}\label{sec:systems-considerations}
 Taking many actions per-timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
@@
 We expect this should be tuned based on the frequency at which statistics arrive.
+The amount of processed flows per deadline depends on the agent design (FLOP count according to policy size, as in \cref{sec:discussion}), but also on the amount of flow telemetry data needing processed---our current implementation is written in python, restricting most of this handling to a single thread.
+An implementation in a systems language such as Rust or C++ would allow faster processing.
+
+\section{System Architecture}\label{sec:system-architecture}
+
+?? Here is where I would put my ramblings about system architecture...
 
@@
 \fakepara{Global state}
-This is the vector of load measurements along a flow's path introduced in \cref{sec:feature-space}.
+This is the vector of current load measurements along a flow's path introduced in \cref{sec:feature-space}.
 These values indicate the overall health of the network, and crucially are all measurements which an agent directly controls.
@@
 As a result, they communicate with attack victims using their own IP address.
-In real-world scenarios the addresses of a set of reflector nodes might exhibit similarity due to network uncleanliness \cite{DBLP:conf/imc/CollinsSFJWSK07}, e.g., similarly unhardened services exposed by a single organisation.
+In real-world scenarios the addresses of a set of reflector nodes might exhibit similarity (such as the same CIDR block) due to network uncleanliness \cite{DBLP:conf/imc/CollinsSFJWSK07}, e.g., similarly unhardened services exposed by a single organisation.
 
@@
 
-We present a VoIP traffic model\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward video game communities.
+We present a VoIP traffic model\footnote{Repository link removed for double-blind reviewing.} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward video game communities.
+%We present a VoIP traffic model\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward video game communities.
 We chose Discord as our prototype due to its publicly documented API, many open source bot frameworks for experimentation, large user base, and due to the lack of models for Opus-encoded traffic.
+All parameters (save for shortening of silent periods) are derived from passive measurement, API documentation, and the bot community's observations.
 %?? Highly periodic, CBR
@@
 
-	\expandableinput ../tables/tnsm-avg-reward.tex
+	\resizebox{\linewidth}{!}{
+		\expandableinput ../tables/infocom-tree-avg-reward.tex
+	}
 \end{table}
@@
 
-	\expandableinput ../tables/tnsm-ecmp-avg-reward.tex
+	\resizebox{\linewidth}{!}{
+		\expandableinput ../tables/infocom-ecmp-avg-reward.tex
+	}
 \end{table}
@@
 However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$\!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
-\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
+\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement over Marl, \SI{8}{$\!\times$} more traffic than no defence system) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
 %SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
@@
 Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
-This is reassuring when measured alongside the insights from other work.
+This is reassuring when measured alongside the insights from other work, since as we discussed in \cref{sec:systems-considerations} the python language bottlenecks performance.
 \Textcite{DBLP:conf/sigcomm/ChenL0L18} observe that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
@@
 \section*{Acknowledgements}
-Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice, comments and technical assistance.
-%Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
+%Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice, comments and technical assistance.
+Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
 %Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
-%This work was supported by a nameless funder [grant number 0].
-This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].
+This work was supported by a nameless funder with a name approximately as long as this line [grant number 0].
+%This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].
 
@@
 \printbibliography
-\vspace{-0.5cm}
-\vskip 0pt plus -1fil
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/kyle}}]{Kyle A. Simpson}
-	received the MSci degree in computing science from the University of Glasgow in 2017.
-	He is currently a PhD student within the Networked Systems Research Laboratory at the School of Computing Science, University of Glasgow.
-
-	Mr.\ Kyle Simpson's research focusses on the use of machine learning and reinforcement learning techniques in cybersecurity and network management, with a core interest in evolving problems and defences.
-\end{IEEEbiography}
-\vspace{-0.5cm}
-\vskip 0pt plus -1fil
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/sdr}}]{Simon Rogers}
-	received the PhD degree from University of Bristol in 2006.
-Now he is a Senior Lecturer in the School of Computing Science at the University of Glasgow.
-Before starting a permanent  post in Glasgow he did a couple of Post Doctoral positions (under Mark Girolami).
-
-	Dr.\ Simon Rogers’s research involves the development of machine learning and statistical techniques to help with the analysis of complex datasets, particularly within the field of metabolomics but also other fields, human-computer interaction and information retrieval.
-\end{IEEEbiography}
-\vspace{-0.5cm}
-\vskip 0pt plus -1fil
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/dimitris}}]{Dimitrios P.\ Pezaros}
-	(S’01–M’04–SM’14) received the B.Sc.
-and Ph.D.
-degrees in computer science from Lancaster University.
-He is currently a Senior Lecturer (Associate Professor) and the Founding Director of the Networked Systems Research Laboratory at the School of Computing Science, University of Glasgow.
-He is also a visiting Professor at the University of Athens, Department of Informatics and Telecommunications.
-
-	Dr.\ Pezaros has published widely (over 80 peer-reviewed contributions to top-tier international conferences and journals) in the areas of computer communications, network and service management, and resilience of future networked infrastructures, and has received significant funding for his research in the above areas from public funding agencies and the industry.
-He is a Chartered Engineer, and a senior member of the IEEE and the ACM.
-
-\end{IEEEbiography}
 

### 2019-07-23 tables/infocom-ecmp-avg-reward.py

@@ -1,0 +1,141 @@
+import csv
+import numpy as np
+from os import path
+import sys
+
+traffics = ["opus", "tcp"]
+ns = [2, 4, 8, 16]
+models = ["mpp", "spf"]
+variants = [
+	("separate", None, "Separate", None),
+	("single", None, "Single", None),
+]
+
+illegal_combos = set()
+#illegal_combos.add(("spf", "opus", "single"))
+# illegal_combos.add(("spf", "udp", "channel"))
+# illegal_combos.add(("spf", "tcp", "channel"))
+
+def get_average_reward(filename):
+	output = 0.0
+	try:
+		with open(filename, "r") as csvfile:
+			reader = csv.reader(csvfile)
+			# _ = nereader.next()
+
+			dats = []
+			for i, row in enumerate(reader):
+				if i == 0:
+					continue
+
+				dats.append(float(row[3]))
+
+			output = np.mean(dats)
+	except:
+		print("file {} not found".format(filename))
+	return output
+
+def write_table(out_file, data):
+	lm = len(models)
+	lv = len(variants)
+
+	out_file.write(
+		"\\begin{{tabular}}{{@{{}}cc{}@{{}}}}\n".format("S"*((lm*lv)+2))
+	)
+	out_file.write(
+		"\\toprule\\multicolumn{1}{c}{Traffic} & " + \
+		"\\multicolumn{1}{c}{$n$} & " + \
+		"\\multicolumn{1}{c}{Unprotected} & " + \
+		"\\multicolumn{1}{c}{Marl} & " + \
+		"\\multicolumn{{{}}}{{c}}{{Instant}} & ".format(lv) + \
+		"\\multicolumn{{{}}}{{c}}{{Guarded}} ".format(lv) + \
+		"\\\\\n" + \
+		"\\cmidrule(lr){{5-{}}}\\cmidrule(lr){{{}-{}}} ".format(5 + lv-1, 5 + lv, 4 + lv + lv) + \
+		" & & & "
+	)
+
+	for _ in models:
+		for (_fn, _replace_str, pres, _replace_m) in variants:
+			out_file.write("& \\multicolumn{{1}}{{c}}{{{}}} ".format(pres))
+
+	out_file.write("\\\\ \\midrule\n")
+
+	# Write file body.
+	for traffic in traffics:
+		for i, n in enumerate(ns):
+			key = (traffic,n)
[…]

### 2019-07-23 tables/infocom-tree-avg-reward.py

@@ -1,0 +1,141 @@
+import csv
+import numpy as np
+from os import path
+import sys
+
+traffics = ["opus", "tcp"]
+ns = [2, 4, 8, 16]
+models = ["mpp", "spf"]
+variants = [
+	("separate", None, "Separate", None),
+	("single", None, "Single", None),
+]
+
+illegal_combos = set()
+#illegal_combos.add(("spf", "opus", "single"))
+# illegal_combos.add(("spf", "udp", "channel"))
+# illegal_combos.add(("spf", "tcp", "channel"))
+
+def get_average_reward(filename):
+	output = 0.0
+	try:
+		with open(filename, "r") as csvfile:
+			reader = csv.reader(csvfile)
+			# _ = nereader.next()
+
+			dats = []
+			for i, row in enumerate(reader):
+				if i == 0:
+					continue
+
+				dats.append(float(row[3]))
+
+			output = np.mean(dats)
+	except:
+		print("file {} not found".format(filename))
+	return output
+
+def write_table(out_file, data):
+	lm = len(models)
+	lv = len(variants)
+
+	out_file.write(
+		"\\begin{{tabular}}{{@{{}}cc{}@{{}}}}\n".format("S"*((lm*lv)+2))
+	)
+	out_file.write(
+		"\\toprule\\multicolumn{1}{c}{Traffic} & " + \
+		"\\multicolumn{1}{c}{$n$} & " + \
+		"\\multicolumn{1}{c}{Unprotected} & " + \
+		"\\multicolumn{1}{c}{Marl} & " + \
+		"\\multicolumn{{{}}}{{c}}{{Instant}} & ".format(lv) + \
+		"\\multicolumn{{{}}}{{c}}{{Guarded}} ".format(lv) + \
+		"\\\\\n" + \
+		"\\cmidrule(lr){{5-{}}}\\cmidrule(lr){{{}-{}}} ".format(5 + lv-1, 5 + lv, 4 + lv + lv) + \
+		" & & & "
+	)
+
+	for _ in models:
+		for (_fn, _replace_str, pres, _replace_m) in variants:
+			out_file.write("& \\multicolumn{{1}}{{c}}{{{}}} ".format(pres))
+
+	out_file.write("\\\\ \\midrule\n")
+
+	# Write file body.
+	for traffic in traffics:
+		for i, n in enumerate(ns):
+			key = (traffic,n)
[…]

### 2019-07-24 paper/infocom.tex
Starting arch plot...
Diagram done.

@@
 \usepackage{tikz}
-\usetikzlibrary{arrows.meta, calc, fit, positioning}
+\usepackage{varwidth}
+\usetikzlibrary{arrows.meta, calc, fit, positioning, shapes.misc}
 
@@
 
-?? Here is where I would put my ramblings about system architecture...
+%?? Here is where I would put my ramblings about system architecture...
+We present a description...
+
+\subsection{Stats API}
+?? API between agent and switches/other agents? Push/Pull API on stats? How do they share experience? Could broadcasts aid both?
+
+?? How do stats get from switcehs to the agent?
+
+?? What does experience transfer look like? Why?
+
+\begin{figure}
+	\centering
+	\resizebox{\linewidth}{!}{\begin{tikzpicture}
+		\node(remote){Remote};
+
+		%%%
+
+		\node[below=0.2 of remote](swpos){};
+		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
+		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
+		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
+		\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
+		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
+
+		%
+
+		\node[right=2 of swpos](epos){};
+		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
+		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
+		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
+		\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.85,0)$) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\small \centering Estimator\\$g(\cdot)$\end{varwidth}};
+
+		%
+
+		\node[right=3.5 of epos](apos){};
+		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
+		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
+		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
+		\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent VF};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
+
+		%
+
+		\node[below=2.5 of remote.west](linestart){};
+		\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
+		\draw [dashed] (linestart) -- (lineend);
+
+		%%%
+
+		\node[below=2.6 of remote](local){Local};
+
+		%%%
+
+		\node[below=0.2 of local](aswpos){};
+		\node[fill=white!80!uofgthistle, draw=black, minimum height=3.5cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
+		\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\small \centering OpenFlow\\Tables\end{varwidth}};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\small \centering Stats\\Collector\end{varwidth}};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
+
+		%
+
+		\node[right=3 of aswpos](avfpos){};
+		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
+		\node[below right, inner sep=2pt] at (avf.north west) {\small Agent VF};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\small \centering Flowstate\\Database\end{varwidth}};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\small \centering TRSE\\Scheduler\end{varwidth}};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};
+
+		%%%
+
+		\tikzset{>=stealth}
+
+		\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
+		\draw[thick, ->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};
+
+		\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
+
+		\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
+		\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};
+
+		\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};
+
+		\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
+		\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};
+
+		\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
+		\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
+		\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
+		\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
+		\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};
+
+		\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
+		\node[right=0.3] at (avfsa.east) {\tiny Experience};
+	\end{tikzpicture}}
+	\caption{System architecture  for our RL-driven DDoS defence system.}
+\end{figure}
 

### 2019-07-26 paper/infocom.tex
The butchering begins...

@@
 \subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
-\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
-We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$, it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
+\emph{Reinforcement learning} (RL) is a variant of machine learning concerned with training an agent to choose an optimal sequence of actions for a given task \cite{RL2E}.
+We assume at any point in time $t$, an agent knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
 This model of system interaction is best described as a \emph{Markov Decision Process} (MDP).
-RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal if an agent needs to further explore some region of the state space.
-The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
-In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new, custom-built training corpus.
-
-From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+3} + \ldots$, choosing the discount factor $\gamma \in [0,1)$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
-Formally, an agent's goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.
+RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken.
+The policy is refined as more decisions and rewards become available, learning adaptively and online.
+In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new training corpus.
 
 %?? Include some details of function approximation in the formalisation? I.e. tile coding, stability and convergence guarantees...
-
-There is immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
-In particular, we shall focus on methods which choose actions according to their value estimates from the current state: let $\operatorname{q}(s, a) \in \mathbb{R}$ denote the estimate of action $a$'s value if it were to be taken in state $s$.
+We focus on methods based on action value estimates from the current state: let $\operatorname{q}(s, a) \in \mathbb{R}$ denote the estimate of action $a$'s value if it were to be taken in state $s$.
 %?? Revisit the maths here, d, dim A, s...
-Exact (tabular) representations require that we store a value estimate for each action in every state---if state is real-valued or high-dimensional, then computation and storage quickly become infeasible.
-To cope with a continuous state and/or action space, one valuable technique is to employ linear function approximation backed by \emph{tile coding} \cite[pp.\ \numrange{217}{221}]{RL2E}.
+To cope with a continuous state and/or action space, one technique to obtain this value is to employ linear function approximation backed by \emph{tile coding} \cite[pp.\ \numrange{217}{221}]{RL2E}.
 
@@
 		An example of tile coding for 2-dimensional state and 4 actions, using 2 tilings, 3 tiles per dimension, and a bias tile.
-		All components of $s$ are clamped to the range $[0,1]$.
-		For simplicity, we denote $\bm{x}(s,\cdot)$ as a list of indices and represent the values of all actions at each tile with a vector.
-		(a) The state $s$ activates the bias tile and exactly one tile in each tiling.
-		(b) The action values of active tiles are summed to produce the current value estimate for each action available in $s$---for this state, local knowledge ensures that action 4 is chosen by the greedy policy despite typically being a poor choice elsewhere.
 		\label{fig:tilecode}
@@
 
-Tile coding is a form of feature representation which converts a state-action pair into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing a $d$-dimensional subset of the space into a number of overlapping grids with an optional bias component.
+Tile coding is a form of feature representation which converts a state-action pair into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing a $d$-dimensional subset of the space into a number of overlapping grids.
 Each tile corresponds to an entry of $\operatorname{\mathbf{x}}(s, a)$ which is set to 1 if the state-action pair lies within it.
 \Cref{fig:tilecode-code} demonstrates the process for a 2-dimensional state space, and that the numbers of tilings and tiles per dimension control feature resolution and generalisation.
-Moreover, to capture combinatorial effects or create multi-scale representation we may combine codings by concatenating individual feature vectors.
 We may then approximate an action's value with respect to a policy parameter vector $\wvec{}$, defining some $\acval{s}{a}{\wvec{}} \approx \operatorname{q}(s, a)$:
@@
 	\label{eqn:sg-sarsa}%
-	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
+	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, the discount factor $\gamma \in [0,1)$ determines how crucial future rewards are, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
 \end{subequations}
@@
 Computing the approximate value of every action available in the current state forms the basis of a policy.
-Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might use some other mechanism.
-\Cref{fig:tilecode-select} extends the prior working example to show how the value of each action is computed (and which action would be chosen by a greedy policy), combining a global estimate ($T_{\mathit{bias}}$) with knowledge particular to each state.
-
-\begin{figure}
-	\centering
-	\resizebox{0.5\linewidth}{!}{
-		\begin{tikzpicture}		
-		% Top half
-		\def\startacs{
-			{0.8},
-			{0.3},
-			{-0.4}%
-		}
-
-		\foreach \line [count=\y] in \startacs {
-			\foreach \pix [count=\x] in \line {
-				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
-					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
-					\draw[shift={(-0.7,0)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
-				}{
-					\pgfmathtruncatemacro\lambda{\pix*100}
-					\draw[shift={(-0.7,0)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
-				}
-				\node[shift={(-0.35,0.175)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
-			}
-		}
-
-		\draw[xstep=0.7cm,ystep=0.35cm,shift={(0,0)}] (0,-1.06) grid (0.7,0);
-		\node[label=left:{$T_{1,9}$},shift={(0,-0.125)}] at (0,0) {}; 
-		\node[label=left:{$T_{2,5}$},shift={(0,-0.125)}] at (0,-0.375) {}; 
-		\node[label=left:{$T_{\mathit{bias}}$},shift={(0,-0.125)}] at (0,-0.75) {};
-		\node[label=below:{\footnotesize Action 4},shift={(0.35,-0.125)}] at (0,-0.75) {};
-		\node[label=above:{\footnotesize $\wvec{t}$},shift={(0.35,-0.125)}] at (0,0) {};
-
-		\def\finalacs{
-			{0.7},
-			{0.2},
-			{-0.5}%
-		}
-
-		\foreach \line [count=\y] in \finalacs {
-			\foreach \pix [count=\x] in \line {
-				\ifthenelse{\lengthtest{\pix pt < 0 pt}}{
-					\pgfmathtruncatemacro\lambda{(\pix+1)*100}
-					\draw[shift={(2-0.7,0)},fill=midac!\lambda!lowac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
-				}{
-					\pgfmathtruncatemacro\lambda{\pix*100}
-					\draw[shift={(2-0.7,0)},fill=highac!\lambda!midac] (0.7*\x,-0.35*\y) rectangle +(0.7,0.35);
-				}
-				\node[shift={(2-0.35,0.175)}] at (0.7*\x,-0.35*\y) {\footnotesize \pix};
-			}
-		}
-
-		\draw[xstep=0.7cm,ystep=0.35cm,shift={(2,0)}] (0,-1.06) grid +(0.7,1.06);
-		\node[label=below:{\footnotesize Action 4},shift={(2.35,-0.125)}] at (0,-0.75) {};
-		\node[label=above:{\footnotesize $\wvec{t+1}$},shift={(0.35,-0.125)}] at (2,0) {};
-
-		\draw[->] (0.9,-0.5) -- node[above] {$+ \alpha \delta_t$} (1.8,-0.5);
-		\end{tikzpicture}
-	}
-
-\caption{
-	The update step for \cref{fig:tilecode}, given an observed TD error $\delta_t=-0.2$ (indicating a lower observed reward than the expected long-term value of \num{0.7}) and $\alpha=0.5$.
-	Action 4's value is thus reduced in the tiles associated with state $s$, but remains the most likely choice; the negative $\delta_t$ may have arisen from noise in the reward signal.
-	For illustrative purposes, we choose an unrealistically high $\alpha$ (typically, $\alpha\le0.05$ is a more reasonable choice).
-	\label{fig:tilecode-update}
-}
-\end{figure}
-
-This combination of algorithm and coding strategy is well-optimised, if actions are discrete; this allows a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each tile.
-Action values for any state are then obtained by summing the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}}-1)$ floating point additions per decision.
-Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
-This means that we need only perform $n_{\mathit{tilings}} + 2$ additions and \num{2} multiplications per model update:
-\begin{equation}
-\bm{w}_{t+1}[i][\operatorname{index}(A_t)] = \bm{w}_{t}[i][\operatorname{index}(A_t)] + \alpha \delta_t, \forall i \in s_{\mathit{list}}.
-\label{eqn:sg-sarsa-opt}
-\end{equation}
-\Cref{fig:tilecode-update} shows how this applies to our prior example.
-If desired we may define a state space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
-Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.
-A caveat of tile coding remains, in that the value of $\alpha$ must be reduced according to the number of tilings to prevent divergence at the expense of slower learning ($\alpha \leftarrow \alpha / n_{\mathit{tilings}}$).
-
-%??
-Is this \emph{actually} just sarsa?
-We're using fn approx (of course), but this is fraught with its own difficulties.
-Is it strictly speaking correct to describe it as Sarsa at this point?
-It's, at the very least, 1-step semi-gradient Sarsa given that it is clearly on-policy w/ fn approx...
-
-%\subsection{Intrusion Detection}
-%Probably want to talk about NIDS/IPS,
-%?? Discuss mininet? Networking terms? SDN stuff?
+To encourage early exploration, actions with maximal value can be chosen each time except when taking random actions with probability $\epsilon$---an \emph{$\epsilon$-greedy policy}).
+\Cref{fig:tilecode-select} extends the prior working example to show how the value of each action is computed (and which action would be chosen by a greedy policy).
+
+This combination of algorithm and coding strategy is well-optimised, if actions are discrete.
+A particularly efficient (vectorised) implementation is described by \textcite{RL2E}---briefly, we need only perform $n_{\mathit{tilings}} + 2$ additions and \num{2} multiplications per model update and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
 
@@
 
-\subsection{Systems Considerations}\label{sec:systems-considerations}
-Taking many actions per-timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
-This introduces some potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated.
+\section{System Architecture}\label{sec:system-architecture}
+
+%?? Here is where I would put my ramblings about system architecture...
+We present our design of a system which operates and support the effective real-world deployment of RL-based agents, acting to mitigate DDoS attacks.
+\Cref{fig:sys-arch} displays this visually, separating system elements which are local to each agent from those which reside elsewhere in the network.
+We describe the main purpose and operation of each module within an agent's \emph{virtual function} (VF), and discuss techniques to make deployment more efficient using existing technologies.
+
+\subsection{Core and RL Executor}
+The core module is the main loop for decision-making in an agent's architecture.
+At each timestep, the core receives information about which flows have arrived and should be acted upon from the \emph{TRS Scheduler}.
+The core then retrieves the current and previous state vector associated with each flow from the \emph{Flowstate Database}, passing them into the RL algorithm alongside the last action chosen for that flow (if available).
+
+The RL algorithm then returns an action.
+Each action is passed to the database, which computes and returns the desired packet drop rate according to the agent model (instant/guarded) while updating flow state.
+This is then converted into an OpenFlow message carrying the packet drop rules; these are sent in a single message to the agent's switch using the same groupings produced by the scheduler.
+Finally, timing information is passed back into the scheduler to refine its estimates about how much work should be scheduled in the next timestep.
+
+\subsection{Stats API and Collectors}
+To act effectively, the agents we have designed require a rich assortment of information from the network and other agents.
+Each agent then exposes an API through which statistics and policy information may arrive.
+%?? API between agent and switches/other agents? Push/Pull API on stats? How do they share experience? Could broadcasts aid both?
+
+%?? What does experience transfer look like? Why? Explain that every agent has a different policy, and why this is beneficial
+Our agents can act either independently, having no agent-to-agent communication, or cooperatively.
+In the latter case agents transfer, when possible, \emph{experience} to one another---lists of state-action-state transitions with associated rewards.
+It's noted that a transition may be high-value or surprising to one agent, while well-known to another, causing each to produce different policy updates from the same unit of experience.
+For this reason we choose not to transfer policy deltas between agents, causing each to learn its own policy.
+We leave the study of which scheme achieves better performance to future work.
+
+%?? How do stats get from switcehs to the agent?
+Load collectors and estimators periodically push their observations to each active agent VF.
+In our current implementation, load statistics are gathered via VFs active at each network switch, though we expect that OpenFlow stats requests, NetFlow or SNMP data may be used to derive load measures.
+Transferring state to agents and experience sharing can both be made more efficient through effective use of broadcast addressing in a target network.
+Depending on the capabilities of switches in the network, the estimator can either send actual benign traffic estimates or parameters for use in a reward function.
+
+\subsection{Flowstate Database}
+For each flow 5-tuple, we hold two state vectors containing the features described in \cref{sec:rethinking-the-state-space}---the current state, and the state which induced the last action.
+To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set.
+State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
+For flows outside of the current work list, we simply replace the stored vector.
+
+\subsection{TRS Scheduler}
+Acting on a possibly unbounded set of flows in each timestep introduces some potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and the risk that flow states become outdated.
 At their worst, these risks present additional attack surface to an adversary.
-To adapt to these problems, we make use of \emph{timed random sequential} updates.
-
-Each agent begins with an empty work list.
-For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
-Uncompleted work is passed on to the next timestep, until the list is emptied, at which point is is repopulated using the set of available measurements.
-To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
-State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
+To tackle these problems, we make use of \emph{Timed Random Sequential} updates.
+
+The scheduler begins with a shuffled work list of active flows.
+When requested, the scheduler estimates the cost of an action computation using the timing information received from the core, proceeding down the list to send a set of 5-tuples to the core which can be handled in a set time limit.
+The scheduler continues until the list is empty, at which point it is repopulated and reshuffled with active flows.
+
 Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
@@
 
-\section{System Architecture}\label{sec:system-architecture}
-
-%?? Here is where I would put my ramblings about system architecture...
-We present a description...
-
-\subsection{Stats API}
-?? API between agent and switches/other agents? Push/Pull API on stats? How do they share experience? Could broadcasts aid both?
-
-?? How do stats get from switcehs to the agent?
-
-?? What does experience transfer look like? Why?
+\subsection{Agent Switches}
+%?? How do we do stat collection?
+Our agent switches operate a modified version of Open vSwitch, implementing an action which requests that each matched packet be dropped with a certain probability.
+To get around the lack of floating-point support in many environments, we represent this probability using a 32-bit unsigned integer (scaling \num{1.0} to $2^{32}-1$).
+On commodity hardware, we believe that a similar effect can be achieved using OpenFlow meters (at the expense of being stateful measures).
+
+We use OpenFlow groups to simplify control messages: premade tables with permitted levels of packetdrop.
+This saves some overhead compared to using experimenter/extension headers.
+Flows are automatically given a group with the default level of packet drop (according to the chosen agent design), meaning that switches don't need to refer to a controller or the agent VF.
+
+%?? What do our modifications to openflow look like? How do we get around the lack of floating point math?
 
@@
 		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\small \centering Flowstate\\Database\end{varwidth}};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\small \centering TRSE\\Scheduler\end{varwidth}};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\small \centering TRS\\Scheduler\end{varwidth}};
 		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
@@
 	\end{tikzpicture}}
-	\caption{System architecture  for our RL-driven DDoS defence system.}
+	\caption{
+		System architecture  for our RL-driven DDoS defence system.
+		\label{fig:sys-arch}
+	}
 \end{figure}
@@
 %All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
-All code underpinning these findings is available on a public repository.\footnote{Private until publication.}
+%All code underpinning these findings is available on a public repository.\footnote{Private until publication.}
 

### 2019-07-26 code/marl/marl.py
Attempt at mixing models...

@@ -77,6 +77,7 @@ def marlExperiment(
 		model = "tcpreplay",
 		submodel = None,
 		rescale_opus = False,
+		mix_model = None,
 
 		use_controller = False,
 		moralise_ips = True,
@@ -937,8 +938,21 @@ def marlExperiment(
 
 			new_host.setIP(ip, 24)
 
+			if mix_model is None:
+				sm = submodel
+			else:
+				draw = random.random()
+				total = 0.0
+				i = 0
+				found_sm = False
+				while (not found_sm) && i < len(mix_model):
+					total += p
+					if draw < total:
+						sm = m["submodel"]
+						found_sm = True
+
 			hosts.append(
-				(new_host, good, bw, link, ip, extern_no)
+				(new_host, good, bw, link, ip, extern_no, sm)
 			)
 		return hosts
 
@@ -1432,7 +1446,7 @@ def marlExperiment(
 
 		host_ip_mac_map = {}
 
-		for (host, good, bw, _, _, extern_no) in all_hosts:
+		for (host, good, bw, _, _, extern_no, sm) in all_hosts:
 			(lhs,) = struct.unpack("I", socket.inet_aton(host.IP()))
 			host_ip_mac_map[lhs] = host.MAC()
 			(_a_node, _a_sarsa, a_leader) = actors[extern_no]
@@ -1485,7 +1499,7 @@ def marlExperiment(
 
 		# What do the hosts believe their IP to be?
 		# i.e. this won't set correctly until the net has started.
-		for (h, i) in [(host, ip) for (host, _, _, _, ip, _) in all_hosts] + [(vertex_map[d], d[0][0]) for d in dests]:
+		for (h, i) in [(host, ip) for (host, _, _, _, ip, _, _) in all_hosts] + [(vertex_map[d], d[0][0]) for d in dests]:
 			h.setIP(i, 24)
 			h.setDefaultRoute(h.intf())
 
@@ -1722,38 +1736,47 @@ def marlExperiment(
 		for i, dest_node in enumerate(dests):
 			dest = vertex_map[dest_node]
 			if model == "nginx":
-				if submodel is None:
-					# How to have a sane config for each?
-					# Config needs to include the IP so that listening works as intended...
-					fname = "temp{}.conf".format(i)
+				cmds = []
+				
+				if mix_model is None:
+					sms = [submodel]
+				else:
+					sms = [m["submodel"] for (p, m) in mix_model]
+				for sm in sms:
+					if sm is None:
[…]

### 2019-07-26 code/marl/tnsm-results.py

@@ -147,6 +147,23 @@ if __name__ == "__main__":
 		params["randomise"] = True
 		params["randomise_new_ip"] = True
 		params["reward_direction"] = "in"
+	elif traffic_prefix == "mix":
+		params["randomise"] = True
+		params["randomise_count"] = 1
+		params["randomise_new_ip"] = True
+		params["reward_direction"] = "inout"
+		params["mix_model"] = [
+			# TCP
+			(0.8, {
+				"model": "nginx",
+				"submodel": None
+			}),
+			# Opus
+			(0.2, {
+				"model": "nginx",
+				"submodel": "opus-voip",
+			}),
+		]
 	deps.append(traffic_types)
 
 	broken_math = expt_part(maths, deps)

### 2019-07-27 paper/infocom.tex
Work-in-progress trimming/rewording/new stuff...

@@
 \DeclareFieldFormat[article]{editor}{}
+
+\usepackage{etoolbox}
+\expandafter\patchcmd\csname citeauthor \endcsname
+{\begingroup}{\begingroup\aftergroup\@}{}{}
 
@@
 \section{Introduction}
-?? Redo intro to be more system focussed? Needs to be suitably different.
+?? Redo intro to be more system focussed? Needs to be suitably different. Perhaps reduce focus on evolving, make it way more about "protocol-agnostic"...
 
@@
 	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
-	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
-	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
-	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
+	\item A description of the system architecture which supports the real-world deployment of such agents (\cref{sec:system-architecture}).
+	\item An investigation into suitable features for an automatic DDoS mitigation system (\cref{sec:rethinking-the-state-space}).
+%	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
+	\item A comparison of the effectiveness of different RL algorithms in these agents (\cref{sec:the-results-of-doing-so}).
+	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
 \end{itemize}
 
-\section{Background and Threat Model}
+\section{Background}
 %?? Introduce RL, related definitions etc.
@@
 \emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
-Attackers achieve this by either exploiting peculiarities of operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
+Attackers achieve this by exploiting operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through volume of requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
 Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
 
-Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
-\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most exploitable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
-Malicious hosts send many small requests, spoofed to appear as though they originated from the victim, causing many large replies to be sent to the intended target---significantly increasing a botnet's throughput while masking the identity of each participant.
+Volume-based attacks are the most relevant to our work, with two varieties being especially prevalent today.
+\emph{Amplification attacks} exploit services who send large replies in response to small requests, with UDP-based services like DNS and NTP seeing most abuse \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
+Malicious hosts send many small requests, spoofed using the victim's address, causing many large replies to be sent to the intended target---increasing a botnet's throughput while leaving each participant anonymous.
 \emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
@@
 We assume at any point in time $t$, an agent knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
-This model of system interaction is best described as a \emph{Markov Decision Process} (MDP).
+This model of system interaction is best described as a \emph{Markov Decision Process} (MDP), where agents aim to maximise the \emph{expected reward} received.
 RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken.
-The policy is refined as more decisions and rewards become available, learning adaptively and online.
+The policy is refined as more decisions and rewards become available, learning online.
 In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new training corpus.
@@
 	\label{eqn:sg-sarsa}%
-	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, the discount factor $\gamma \in [0,1)$ determines how crucial future rewards are, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
+	where $\delta_t$ is known as the \emph{temporal-difference} (TD) error, the \emph{discount factor} $\gamma \in [0,1)$ determines how crucial future rewards are, and the vector gradient $\nabla$ is taken with respect to $\wvec{}$.
 \end{subequations}
@@
 an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
-Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete reasons for their use here.
-We have seen that for other domains in particular, misclassification is a serious problem, which can introduce \emph{collateral damage} in the context of DDoS prevention.
-In theory, the feedback-loop-like model allows us to monitor flows \emph{after} an action is taken to allow forgiveness of mistakenly punished flows.
-This does rely upon the ability to take a flow-by-flow view of the state space, but if we can combine knowledge of current state with the last applied action, then perhaps a flow which falls off identically to a legitimate flow can be rescued.
+There are concrete reasons to use RL methods in this domain, beyond handling non-stationarity.
+Misclassification can be a serious problem, introducing \emph{collateral damage} in a DDoS prevention system.
+RL's feedback-loop model lets us monitor flows \emph{after} an action is taken to perhap allow forgiveness of mistakenly punished flows.
+This relies upon a flow-by-flow view of the state space, but by combining knowledge of current state with the last applied action, a punished flow which falls off identically to a legitimate flow can be rescued.
 
@@
 Other studies suggest that there are particularly useful features which make the task of online DDoS flow identification feasible.
-Aggregate network load observed at various locations suggests the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest asymmetry and in many cases illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
-Generic volume-based statistics (counts, counts per duration, average packet sizes) have seen effectiveness in such as $k$-nearest neighbours classifiers trained to detect DDoS attacks in progress \cite{DBLP:conf/dsn/LeeKSPY17}.
-Most importantly, there is evidence showing behavioural changes in response to bandwidth expansion \cite{DBLP:conf/ndss/KangGS16}, suggesting similar artefacts might arise after throttling, packet drop, or other interference.
+Aggregate network load observed at various locations suggests the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
+Volume-based statistics (counts, counts per duration, average packet sizes) have seen use in $k$-nearest neighbours classifiers trained to detect DDoS attacks in progress \cite{DBLP:conf/dsn/LeeKSPY17}.
+Most importantly, there is evidence showing that bandwidth expansion may cause behavioural changes  which differ between benign and attack traffic \cite{DBLP:conf/ndss/KangGS16}.
 %?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
@@
 
-\subsection{Threat Model}
-An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
-Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
-We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
-The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
-The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
-
-We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
-While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
-The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
-These are APT-level capabilities, whose exploration presents a rich source for future work.
+%\subsection{Threat Model}
+%An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
+%Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
+%We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
+%The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
+%The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
+%
+%We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
+%While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
+%The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
+%These are APT-level capabilities, whose exploration presents a rich source for future work.
 
 \section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
-Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
-However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
+%Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
+We hypothesise that per-flow agent designs will offer better accuracy than current RL-based agents for DDoS mitigation.
+%However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
 We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
@@
 \subsection{System Design and Assumptions}
-A deployment environment is defined as a network, with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
+A deployment environment is a network, with a set of \emph{ingress/egress points} from its domain of control through which traffic can flow, and a set of protected \emph{destination nodes}.
 These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
@@
 
-We assume that the deployment environment is a moderately complex software-defined network, because the paradigm offers features which can directly benefit RL agents acting within.
-The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
-Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners.
-Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
-
 \subsection{Algorithm}
-To make decisions cheaply and at low latency, we use \emph{semi-gradient Sarsa with tile coding} as described in \cref{eqn:sg-sarsa} and \cref{sec:reinforcement-learning}, rather than making use of neural networks or more complicated function approximators.
+To make decisions and learn cheaply with low latency, we use \emph{semi-gradient Sarsa with tile coding} \ref{eqn:sg-sarsa} (\cref{sec:reinforcement-learning}), rather than make use of neural networks.
 Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
-Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their weight vector updates with one another (but may share experience and traces with one another).
-
-Although the choice of a classical RL method likely brings lower theoretical performance, there are significant reasons to favour such methods; these include lower latency decision-making, lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
+Each agent has its own internal parameter vector $\wvec{}$.
+
+While classical RL methods likely have lower accuracy, there are reasons to favour them; faster decisions, lower energy usage, reduced model complexity (and training time), and the availability of compatible hardware.
 This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
-Simpler decision boundaries reduce the risk of overfitting and unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.
-
-When choosing a learning algorithm, we compared against Q-learning as well as methods based on \emph{eligibility traces} such as Watkins's $\operatorname{Q}(\lambda)$ \cite[pp.
-312--314]{RL2E} and $\operatorname{Sarsa}(\lambda)$ \cite[pp.
-305]{RL2E}.
-Preliminary experiments found that Sarsa offered the best performance and behaviour.
-?? Link plot or table w Sarsa vs. Q results.
+Approaches like this have simpler decision boundaries, reducing the risk of overfitting and unexpected behaviour. 
+Simpler methods of computing policies may also aid in interpreting anomalous action choices.
 
@@
 		Global state selection for a flow between an external host and server $s_0$ which passes over Agent 1.
-		All nodes in the path taken through the defended network are filled in blue, and all link load measurements which are chosen for action computation are indicated with a thick blue line.
+		Nodes in the path taken are filled in blue, and link load measurements which are used for decisions are indicated with a thick blue line.
 		\label{fig:global-state-path}
@@
 QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
-%As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development to properly handle traffic with such characteristics.
-
-%?? Make more sane. This is about ``Why go per-flow?''
-This further justifies our focus on per-flow decisions---real-world deployments see many flows pass over any egress point, making global actions (such as those chosen by \textcite{DBLP:journals/eaai/MalialisK15}) more likely to inflict collateral damage.
-%This manifests in two ways: the best-achievable performance drops, and so too does the learning rate.
-Given the probability that a host is legitimate, $P_G \in [0,1]$,  if it is assumed that a host may be only legitimate or malicious then it follows that it will be malicious with probability $P_B = 1 - P_G$.
-Supposing we have a set of $n$ hosts, the probability that all hosts will be legitimate is $P_G^n$ (resp. $P_B$ for malicious).
-Defining \emph{imperfect service} $P_{M,n}$ to mean any case where all $n$ hosts connecting over a switch do not share the same classification (i.e., a mixture), then the probability that a switch is delivering imperfect service is $P_{M,n} = 1 - (P_G^n + P_B^n)$.
-\begin{thm}
-	As the host/learner ratio $n$ increases, it is more likely that a throttling switch will exhibit imperfect service: $\forall n \in \mathbb{Z}^{+}, P_{M,n} \le P_{M,n+1}$.
-\end{thm}
-\begin{proof}
-	\emph{Base case:} $P_{M,1}=0, P_{M,2} = 1 - P_G^2 - P_B^2 \ge 0$.
-
-	\emph{Inductive step:} Assume that the theorem holds for $n$. Observe that $P_G^n \ge P_G^{n+1}$ (resp.\ $P_B$). It then follows that:
-	\begin{align*}
-	P_G^n + P_B^n &\ge P_G^{n+1} + P_B^{n+1}\\
-	1 - (P_G^n + P_B^n) &\le 1 - (P_G^{n+1} + P_B^{n+1})\\
-	P_{M,n} &\le P_{M,n+1} \qedhere
-	\end{align*}
-\end{proof}
-\begin{corr}
-	Restricting $P_G \in (0,1)$ so that both $P_G$ and $P_B$ are non-zero ensures strict inequality: $P_{M,n} < P_{M,n+1}$.
-\end{corr}
+Furthermore, applying a single rule to many flows (as in \textcite{DBLP:journals/eaai/MalialisK15}) increases the likelihood of mistaken collateral damage.
 When considering that many hosts have an especially adverse reaction to our main means of control, flow-level granularity becomes an obvious choice.
@@
 \Cref{fig:sys-arch} displays this visually, separating system elements which are local to each agent from those which reside elsewhere in the network.
-We describe the main purpose and operation of each module within an agent's \emph{virtual function} (VF), and discuss techniques to make deployment more efficient using existing technologies.
-
-\subsection{Core and RL Executor}
+In particular, we assume that each agent operates as a \emph{virtual function} (VF) adjacent to a software-defined switch.
+Software-defined networks allow a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
+Furthermore, networks of this kind more naturally enable the future relocation and easy installation of learners.
+Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
+We describe the main purpose and operation of each module within an agent's VF, and discuss techniques to make deployment more efficient using existing technologies.
+
+\subsection{Core and RL Executor}\label{sec:core-and-rl-executor}
 The core module is the main loop for decision-making in an agent's architecture.
@@
 Finally, timing information is passed back into the scheduler to refine its estimates about how much work should be scheduled in the next timestep.
+
+State space sizes guarantee that an \emph{Instant} agent's policy remains under \SI{520}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies: $\sim$\SI{17.8}{\kibi\byte} from our experiments.
+\emph{Guarded} policies are \SI{30}{\percent} of this size.
+As we have described earlier, action updates require a constant number of floating point operations---\num{160} floating point additions and \num{32} multiplications per update of $\wvec{}$ with per-tile updates, above the \num{160} additions required to choose an action.
+The vast majority of these operations can be vectorised.
+Action computation for \emph{Guarded} agents is cheaper still, requiring only \num{48} additions per action.
 
@@
 
+Gathering and transmission of load/flow statistics would be difficult to perform quite as often as an emulated environment allows.
+The measurements acquired in such a scenario are far more likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
+
 \subsection{Flowstate Database}
@@
 We expect this should be tuned based on the frequency at which statistics arrive.
-The amount of processed flows per deadline depends on the agent design (FLOP count according to policy size, as in \cref{sec:discussion}), but also on the amount of flow telemetry data needing processed---our current implementation is written in python, restricting most of this handling to a single thread.
-An implementation in a systems language such as Rust or C++ would allow faster processing.
+The amount of processed flows per deadline depends on the agent design (FLOP count according to policy size), but also on the amount of flow telemetry data needing processed---our current implementation is written in python, restricting most of this handling to a single thread.
+An implementation in a systems language such as Rust or C++ would allow faster concurrent processing.
+
+There is a risk that so much work can be queued up that an agent is never able to act on some attack flows.
+A solution is to impose an upper bound on the amount of action computations/policy updates that can be performed before a work list is regenerated.
+This removes the guarantee that all flows will be visited fairly often, but if updates occur regularly then this random sampling may be sufficient to achieve good performance.
 
@@
 
-\section{Results}
+\section{Results and Discussion}
 \label{sec:the-results-of-doing-so}
@@
 Our \emph{Instant} and \emph{Guarded} designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
+Paradoxically, \emph{Instant} actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between widely different actions, due to shared characteristics between legitimate and malicious flows.
+
+Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect in the case of UDP traffic.
+The most likely explanation for low UDP protection is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
+The increased state space size and similar characteristics between benign and attack traffic make this more likely.
+These difficulties may be further worsened by the competitive nature of multi-agent learning.
 
@@
 
+\emph{Guarded}'s unpredictable (often worse) starting performance is unexpected, given its far smaller action space.
+We expected that this would make the model easier to learn, but the additional state required appears to make the task \emph{harder} initially.
+This design performs best (and exhibits considerably lower variance) when agents learn from as much knowledge as possible: high $n$ and single-agent training.
+To filter incoming traffic, it must degrade a flow multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
+We claim \emph{Guarded} is a considerably stronger model for these reasons, and its successes offer strong rationale to consider the best schemes for efficient information sharing.
+
 %\subsection{Single-agent performance}
@@
 
-\subsection{Computational Cost}
-%??
+%\subsection{Computational Cost}
+%%??
 Consider talking about the execution times of the old MARL approach here?
 They're real nice (as expected), so we have lots of room to play around with while (hopefully) remaining under the 1ms target time given by \textcite{DBLP:conf/sigcomm/ChenL0L18}.
-
-%Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
-Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
-This is reassuring when measured alongside the insights from other work, since as we discussed in \cref{sec:systems-considerations} the python language bottlenecks performance.
-\Textcite{DBLP:conf/sigcomm/ChenL0L18} observe that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
-%Most flows are short, and flow-size follows a heavy-tailed distribution.
-That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
+%
+%%Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
+%Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
+%This is reassuring when measured alongside the insights from other work, since as we discussed in \cref{sec:systems-considerations} the python language bottlenecks performance.
+%\Textcite{DBLP:conf/sigcomm/ChenL0L18} observe that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
+%%Most flows are short, and flow-size follows a heavy-tailed distribution.
+%That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
 %?? TODO: update with modern numbers...
 
-\section{Discussion}\label{sec:discussion}
-
-%?? New angle -- \emph{Guarded} offers substantal improvements for \emph{most} internet traffic according to the stats we know. Is accurate protection of beningn UDP traffic actually more difficult?
-
-%Talk about flaws here, what could go wrong...
-
-%?? Why does SPF only do well sometimes? Model is actually more difficult to learn, so it seems to do best when it has a larger set of decisions to learn from. But, it does worse for TCP?
-\fakepara{Model performance}
-Of the results presented, \emph{Guarded}'s unpredictable (often worse) starting performance is unexpected, given its far smaller action space.
-It's natural to expect that this would make the model easier to learn, but the additional state required appears to make the task \emph{harder}, beyond even the value of choosing a non-zero discount factor (adding forward-planning to explicitly mitigate this effect).
-Accordingly, we see that this design performs best (and exhibits considerably lower variance) when agents learn from as much knowledge as possible: high $n$ and single-agent training.
-To filter incoming traffic from a source, it must decide to degrade their inbound traffic multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
-Our belief is that \emph{Guarded} is a considerably stronger model for these reasons, and its successes offer strong rationale to consider the best schemes for efficient information sharing.
-Paradoxically, \emph{Instant} generally achieves the best performance for UDP traffic yet actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.
-
-%?? May be hard to learn multiple features at once while controlling multiple flows while contending with many more agents, with harder dynamics like TCP. Does this hinder learning in the long run?
-Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect in the case of UDP traffic.
-%??
-I don’t know if you want to also mention the problem of things getting “stuck” in locally optimal (but globally sub-optimal) policies.
-I think it’s fairly uncontroversial that an increase in state space would make this more likely.
-Also, the lack of discounting could play a large role (and more so as the state space becomes more complex…
-The most likely explanation is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
-The increased state space size makes this a more likely occurrence, as is the unclear affect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state space.
-We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to observe the true value of each action), and controlling shared global state.
-Although our design does take steps to counteract such effects, these mitigations may not be enough.
-Moreover, benign UDP traffic shares many characteristics with attack traffic, suggesting that more training samples or some unknown feature might aid control, or that it may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows.
-
-%It is likely that the design of Marl++ 
-%?? Need to mention that Marl performs lower than their paper's numbers for TCP...
-
-%Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
-Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
-Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
-The benefits we have at present are, however, substantial.
-What we offer above the alternate approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
-The techniques we propose offer substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today.
-Moreover, our decision making processes are entirely agnostic of the protocol or content of inbound traffic, offering future-proofing against the introduction of new transports.
-
-\fakepara{Security concerns and vulnerability}
-Can an agent be flooded with new flows to reduce their ability to make decisions?
-One of the risks introduced by our timed random sequential policy updates is that so much work can be queued up that an agent is never able to act on some attack flows.
-The natural solution is to impose an upper bound on the amount of action computations/policy updates that can be performed before a work list is discarded completely.
-This removes the guarantee that all flows will be visited fairly often, but if updates occur regularly then this random sampling may be sufficient to achieve good performance.
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.9\linewidth]{../plots/policy-16-tcp-f5-mean-log}
-%	
-%	\caption{
-%		Mean Marl action values from pre-training on mean IATs ($n=16$).
-%		As these are mean values, lighter cells in each row indicate actions \emph{likely} to be taken by an agent's policy.
-%		Repeated values originate from the bias tile, which is always active, and indicate regions of the state-space which have yet to be visited---here, this is most of the state space.
-%		Agents visibly learn action values for tiles which override the default bias tile's preferences.
-%		The measured effectiveness of this feature then suggests that a low-resolution coding over the standard region may, in fact, be a better choice.
-%		We see that agents prefer to choose $p=0.0$ in most states, but higher $p$ when IATs are particularly small.
-%		\label{fig:intern-16-tcp-iat}
-%	}
-%\end{figure}
-%
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.9\linewidth]{../plots/policy-4-tcp-f7-mean}
-%	
-%	\caption{
-%		Mean Marl action values from pre-training on $\Delta$ out rates ($n=4$).
-%		A sudden ramp-up of server-to-host traffic is used as a strong indicator of flow legitimacy, while more punishing actions have comparatively higher value as $\Delta$ out rate decreases.
-%		Furthermore, we see again that a region of the state space has gone somewhat unexplored---we observed after plotting this visualisation that many decreases in this feature are too small to hit the adjacent tile, which implies a mixed-resolution coding may improve an agent's policy further.
-%		\label{fig:intern-16-tcp-something}
-%	}
-%\end{figure}
-
-%?? Is the state space interpretable? Yes!
-Machine learning algorithms have earned a reputation for eluding human interpretation, while being vulnerable to evasion and poisoning.
-Given the risks associated with introducing such techniques in the context of security, it is natural to be concerned with the interpretability of the models we have proposed.
-With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
-%\Cref{fig:intern-16-tcp-iat,fig:intern-16-tcp-something} show how we may visualise the portion of a policy for each feature, and describe what information can be gained from doing so.
-Furthermore, system state which is dependent on many signals drawn from across a wide network (such as our global state) is difficult to exert precise control over.
-These signals' topological separation, in concert with their burstiness and unpredictability, may have substantial effects on an attacker's capabilities.
-
-\fakepara{Real-world Deployment}
-Currently, we assume that switches support an extension to OpenFlow to enable remotely installable packet-drop rules, either by running a modified version of OVS on commodity hardware at these locations or through custom firmware for egress switches.
-Similar functionality could be employed by making use of OpenFlow's meter rules.
-
-Where overheads are concerned, the state space sizes guarantee that an \emph{Instant} agent's policy remains under \SI{520}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies: $\sim$\SI{17.8}{\kibi\byte} from our experiments.
-\emph{Guarded} policies are \SI{30}{\percent} of this size.
-As we have described earlier, action updates require a constant number of floating point operations---\num{160} floating point additions and \num{32} multiplications per update of $\wvec{}$ with per-tile updates, above the \num{160} additions required to choose an action.
-The vast majority of these operations can be vectorised trivially, if such hardware is present.
-Action computation for \emph{Guarded} agents is cheaper still, requiring only \num{48} additions per action.
-Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
-We believe that it may be possible to implement similar behaviour on standard commodity switches through application of \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
-
-%?? Incorrect, pickling...
-%?? deployment/system guidelines, architecture, and overhead/footprint of doing this
-%?? Mention: probably fewer state measurements than in our emulations, but longer measurements means less noisy (so probably more accurate)
-Gathering and transmission of load/flow statistics would be difficult to perform quite as often as an emulated environment allows, without inadvertently affecting host traffic.
-However, the measurements acquired in such a scenario are far more likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
-The main bottlenecks are most likely in forwarding the load measurements from various aggregation points (which can be made more efficient through multicast) and in running some estimator $\operatorname{g}(\cdot)$ to condition the reward function.
-%?? Potential for dividing up pre-feature training across different agents, or train like this locally (sets of flows trained by 1 sub-model).
-%?? Different global state per-agent? Seems this must be locally trained, just make it (everything up the chain to key destinations).
-We expect that agents will be able to share policies for all features, which may help to offset the reduced rate of incoming experience.
-Regardless, it will take longer to achieve enough state-state transitions to converge on a good policy.
+\subsection{RL algorithms}
+When choosing a learning algorithm, we compared semi-gradient Sarsa against Q-learning.
+Preliminary experiments found that Sarsa offered the best performance and behaviour.
+?? Link plot or table w Sarsa vs. Q results.
 
@@
 \fakepara{DDoS Prevention}
-\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
+\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of volume-based DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
 Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
@@
 %?? Actually talk about Marl (???) to appease reviewer \#1.
-The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15} (whom we have positioned our work against), and their contribution in applying RL to the task of intrusion prevention is significant: their work helps to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
-They create a tree overlay topology (subdivided into teams), where each agent applies packet drop to \emph{all} flows inbound to a protected server.
+The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, whom we have positioned our work against.
+They create a tree overlay topology subdivided into teams, where each agent applies packet drop to \emph{all} inbound traffic.
 %?? Recap their flaws, since they've been cut form every other aspect.
-Our results show that their technique underperforms at high host density and when congestion-aware traffic dominates---that their results do not demonstrate this suggests an evaluation driven purely by traces (rather than live application dynamics).
+Their technique suffers at high host density and when congestion-aware traffic dominates.
+These effects are not present in their results, suggesting an evaluation driven purely by traces.
 
 \emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
-\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
+Bots participating in an attack are often unable to match any bandwidth expansion in an AS due to having already saturated the capacity of their local outbound links.
 %Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
-Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ASes, although some of our feature choices are backed by similar observations.
-A weakness of their approach is that computing a route to measure bandwidth expansion on real networks can be costly (up to \SI{14}{\second} for the Cogent topology), and that the low expansion factors in real network can require more ``rounds'' of filtering.
-Their assumptions about traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), both of which make up a sizeable proportion of network traffic.
-
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
-Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
+Unlike our approach, SPIFFY is intended to be deployed within ASes.
+A weakness of their approach is that computing a route with enough spare bandwidth can be costly, and that low expansion factors in real network can require more ``rounds'' of testing. 
+Traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), which make up a sizeable portion of network traffic.
+
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a generalised SDN framework for intrusion detection, but shows the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
+Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the some evidence needed to understand how the system compares.
 
@@
 This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
-The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
+The approach uses upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
 Because of this, traditional flooding-based attacks aren't fully alleviated.
@@
 \fakepara{RL in Networks}
-Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
-Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-Given that one of the main strengths of RL techniques is the ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works don't apply the rich literature on the subject to its fullest potential.
-
-For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
-A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
-%We feel there is a shortage of work in this category at present, at least in the field of networks.
-To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}.
-
-An \emph{indirect-control} RL problem is one where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
-A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
+Application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
+Past work uses the approach as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
+Given that one of the main strengths of RL techniques is the ability to control (and adapt through) ongoing interaction, these works don't apply the paradigm at its fullest potential.
+
+A past example which embodies a different use learns when it is best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
 %The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+%The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
+Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
 Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
@@
 
-We have identified weaknesses in past approaches arising from simulation assumptions, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
-Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
+We have put forth an RL agent design which acts per-flow, describing the algorithmic modifications and engineering choices needed to make its deployment feasible.
+Supporting this, we've demonstrated the value of the features these agents use as evidence.
 Our evaluation shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, with \emph{Guarded} agents showing the most promise for future evaluation.
@@
 Simulation is the most likely avenue for such evaluation.
-
-%(rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?
-
-%\section{Future Work}
-%Airlift half of the ``conclusion'' and paste it in here, so that it can be a lot neater.
-%?? Future Work? I.e., \emph{everything}: no one else is really looking at/interested in this specific kind of application of RL yet. \emph{Yet}.
-
-%?? IDEA: try out average reward, TD($\lambda$) methods as future work...
-%A natural research direction to enhance this work would be the combination of the classic function approximation techniques we make use of alongside the improved algorithms that the RL community has introduced in the past few years.
-%Actor-critic methods, or algorithms based on eligibility traces are good candidates for investigation. 
-The remaining weaknesses invite many further improvements worth investigation.
-%??
-What might we do for a reward function in the absence of heuristic estimates and/or explicit a priori knowledge?
-I think a good candidate is the sum of up and down throughputs (normalised by capacity sum), so long as \emph{neither exceeds the link capacity}.
-We can extend the team-based formulations similarly.
-This, in theory, promotes traffic diversity since it's not like flooding-based DDoS attacks are going to submit meaningful work to a server.
-The intuition, I suppose, is that certain classes of flow will have a small footprint in one direction which causes a sizeable increase in the other!
-Alternatively, monitor the health of canary flows which cross the team boundary (i.e.
-only one in-out link).
-A problem we raised (without a clear solution) was the design of reward functions which do not rely upon heuristic estimates or a priori knowledge of benign traffic content.
-If true online learning is desired (i.e., coping with a non-stationary environment), then such reward functions are sorely needed.
-While $\bload{\cdot}$ is likely to be a good candidate for many deployments, we believe that finding an effective metric derived from the individual statistics we suggested serves as an interesting research problem.
-
-%??
-Benefit of the more realistic emulation environment is that it is far closer in behaviour and architecture (i.e.
-viable) to a real SDN-enabled deployment, captures some dynamics which were otherwise hidden/lost by human ignorance.
-It also allows me to develop the system towards evolving traffic models where it is expected that RL should shine over and above standard ML techniques.
-THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
-Given that one of the advantages that RL methods can offer is the ability to handle non-stationary problems, it is important to propose and test sensible simulations or captures of evolving networks against these methods.
-While it is known that DDoS attack strategies evolve in real time \cite{DBLP:conf/spw/KangGS16}, evaluation is difficult at present since no works detail what patterns such evolution might take.
-Regardless, these scenarios present ideal circumstances to apply work on adaptive exploration \cite{DBLP:conf/annpr/TokicP12}, changepoint detection, or intelligent/heuristic sampling methods to judge which flows are most worthy of consideration per tick.
-For estimating \emph{when} to explore, we believe that the intersection of signal processing and RL is as-yet unexplored.
-
-%??
-IDEA: Apply these techniques to programmable data planes etc.
-While it's pretty neat that what we have works assuming that ach router is a software (x86) switch running OVS, what might we need to consider when applying this to `real' switches?
-``PDP can allow this to be added to real routers to make it efficient to keep \& process state in the manner we require, as well as enabling more adaptive deployment''.
-Cite P4, BPFabric, other work on PDPs?
-%??
-What concessions will we have to make in order to make per-flow processing more viable?
-Intelligent sampling/reanalysis of flows when needed (i.e.
-an external heuristic guiding method)?
-In SPIFFY's \cite{DBLP:conf/ndss/KangGS16} evaluation, we see clearly that it can take around \SI{2}{\second} for a flow to react fully to a rate increase---I think for the TCP step it may be wise to factor this in, too!
-%?? More future work --- share knowledge between agents. ``Knowledge bases'' for this purpose? (see: Qianru).
-Effective real-world deployment of RL-based defences cannot assume that switches in a network will support a custom version of OVS or other arbitrary software, introducing the question of whether agent training, execution and distribution may be possible when using \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
-We also expect it will be fruitful to look into \emph{how} agents may share knowledge with one another.
-
-%?? Security? I suspect that the very qualities that make inference difficult in IDS/IPS also increase the level of challenge an advanced threat must overcome.
-%??
-Might want to mention it in related work above, but the recent attention on adversarial examples/tricking models needs to be looked into for RL.
-Poisoning attacks relevant for online techniques: old bounds exist \textcite{DBLP:journals/jmlr/KloftL10}, new stuff concerns collaborative learners \cite{DBLP:conf/acsac/ShenTS16}, nothing for rl.
-Hot topic in deep networks \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18}, but naturally still relevant with even linear approximations or exact tabular case due to limits of the PAC assumption.
-There is now examination of evasion attacks wrt.\ RL \cite{DBLP:journals/corr/HuangPGDA17}!
-%??
-evasion attacks by \textcite{DBLP:conf/sp/Carlini017}---all of these are computed by way of a general stochastic optimiser, such as \emph{Adam} \cite{DBLP:journals/corr/KingmaB14}.
-possible to apply something similar to our learned model to assess its security?
-would the suggested states even be valid?
-(i.e.
-since they're monotonically increasing for the most part).
-Although we believe that the security landscape for classical RL models is not \emph{identical} to that of neural-network based approaches (particularly with such noisy, volatile, and hard-to-control data), there is still immense value in determining the exact capabilities of a sufficiently powerful adversary as the risk of external control still exists.
-In particular, we believe that poisoning attacks and evasion attacks merit close consideration.
-
-%While this work still trails behind the performance of exact DDoS flow detection mechanisms, w
-We hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
-It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques rely upon.
-It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
+The remaining weaknesses invite many further improvements worth investigation: the design of reward functions derived from QoE metrics, agent operation via \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}, and evaluation against evolving DDoS attacks \cite{DBLP:conf/spw/KangGS16}.
+
+We hope it is clear that reinforcement learning holds promise and can inspire further innovation.
+It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques discussed in \cref{sec:related-work} rely upon.
+Our proposed approach offers substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today, while offering future-proofing against the introduction of new transports.
+It's hoped that more research in this direction will open the door to works which capture the complex and fluid nature of modern networks; evolving topologies, traffic and protocol distributions, and attacks.
 
 \section*{Acknowledgements}
-%Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice, comments and technical assistance.
-Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
+%Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou and Cristian Urlea for their advice, comments and technical assistance.
+Our thanks go to Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
 %Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.

### 2019-07-27 paper/infocom.tex
Intro still needs to be gutted...

@@
 		An example of tile coding for 2-dimensional state and 4 actions, using 2 tilings, 3 tiles per dimension, and a bias tile.
+		\vspace{-1em}
 		\label{fig:tilecode}
@@
 \subsection{System Design and Assumptions}
-A deployment environment is a network, with a set of \emph{ingress/egress points} from its domain of control through which traffic can flow, and a set of protected \emph{destination nodes}.
+Agents are deployed in a network, with a set of \emph{ingress/egress points} from its domain of control through which traffic can flow, and a set of protected \emph{destination nodes}.
 These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
@@
 \subsection{Algorithm}
-To make decisions and learn cheaply with low latency, we use \emph{semi-gradient Sarsa with tile coding} \ref{eqn:sg-sarsa} (\cref{sec:reinforcement-learning}), rather than make use of neural networks.
+To make decisions and learn cheaply with low latency, we use \emph{semi-gradient Sarsa with tile coding} \ref{eqn:sg-sarsa} (\cref{sec:reinforcement-learning}).
 Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
@@
 
-While classical RL methods likely have lower accuracy, there are reasons to favour them; faster decisions, lower energy usage, reduced model complexity (and training time), and the availability of compatible hardware.
-This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
-Approaches like this have simpler decision boundaries, reducing the risk of overfitting and unexpected behaviour. 
-Simpler methods of computing policies may also aid in interpreting anomalous action choices.
+While classical RL methods likely have lower accuracy than more complex approximators like neural networks, there are reasons to favour them; faster decisions, lower energy usage, reduced model complexity (and training time), and the availability of compatible hardware.
+This aligns with the goal of quick online learning and fast adaptation to traffic changes, without introducing dedicated tensor processing hardware.
+Approaches like this also have simpler decision boundaries, reducing the risks of overfitting and unexpected behaviour, while possibly aiding in interpreting anomalous action choices.
 
 \subsubsection{Action rate}
-We chose to adapt the algorithm to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to facilitate effective learning.
-To this end, we allow agents to make many decisions per timestep.
-We maintain the last state-action pair associated with each source IP and destination, and calculate any actions for the flows which still exist.
-Finally, we update $\wvec{}$ using each available trace and the reward signal from the relevant destination.
-As exploration still occurs for each action, this approach reduces $\epsilon$ multiple notches every timestep.
+The algorithm is adapted to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to learn quickly.
+Agents thus make many decisions per timestep.
+We update $\wvec{}$ using each decision made and the reward signal from the relevant destination.
+As exploration still occurs for each action, $\epsilon$ must be reduced multiple notches every timestep.
 In turn, we increase the annealing window for $\epsilon$ by a factor of \num{2.67} so as to preserve exploration over time, by accounting for the greater volume of decisions being made.
@@
 \subsubsection{Per-tile updates}
-While the standard formulation of \cref{eqn:sg-sarsa} updates the value of all tiles identically (by a scalar $\alpha \delta_t$), we found it more effective to compute a different temporal difference value \emph{for each tiling}.
-This adjusts the value of each activated tile with a different magnitude and direction, correcting tilings which contribute too much or too little value without disturbing estimates which are already close to correct.
+To adjust the value of each activated tile with a different magnitude and direction, we found it effective to compute a different temporal difference value \emph{for each tiling}.
+This corrects tilings which contribute too much or too little value without disturbing estimates which are already close to correct.
 While we make use of the sum of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to set $\alpha$ to a higher value without divergence.
-A crucial observation is that value updates to each tile can move by different values in different directions, converging on effective estimates sooner.
+Compared to the standard algorithm, convergence on effective estimates occurs sooner.
 
 \subsubsection{Decision narrowings}
-When learning control on the basis of a high-dimensional, tile-coded state space, assignment of credit for each decision is difficult (because all tiles have identical gradient).
-Rephrasing this, assigning credit to the \emph{features} which drove each decision is difficult.
+When learning a high-dimensional tile-coding, assignment of credit for each decision is difficult because all tiles have identical gradient--i.e., assigning credit to the \emph{features} which drove each decision is difficult.
 To combat this, with probability $\epsilon$ an agent will mark a flow as being governed by a subset of the state space for the next five decisions.
-Each agent chooses actions on that source/destination pair using one element of local state, the global state, and the bias tile---we include the latter two to strike a balance between and accuracy and correct credit assignment.
+Each agent chooses actions for that flow using \emph{one element of local state}, the global state, and the bias tile---we include the latter two to strike a balance between and accuracy and correct credit assignment.
 
@@
 	\centering
-	\resizebox{0.7\linewidth}{!}{
+	\resizebox{0.65\linewidth}{!}{
 		\begin{tikzpicture}[
@@
 		Nodes in the path taken are filled in blue, and link load measurements which are used for decisions are indicated with a thick blue line.
+		\vspace{-1em}
 		\label{fig:global-state-path}
@@
 
-Our state space combines elements of global state (network link load observations) with per-flow measurements.
+Our state space combines global state (network link loads) with per-flow measurements.
 Each is tile-coded with 8 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
@@
 %?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
-Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}) depending upon the bandwidth measurements regularly received from monitors in the environment.
+Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}): the bandwidth measurements regularly received from monitors in the environment.
 For any flow, an agent then computes the path it would take through the network.
-The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
+The incoming load along the first hop, last hop, and tertiles of the path may then be tile-coded together.
 In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
@@
 
-We build global state in this way to offer compatibility with multipath, multi-destination networks, offering support for diverse deployment environments from endpoint servers to transit ASes.
+Global state built in this way is compatible with multipath, multi-destination networks, offering support for diverse deployment environments from endpoint servers up to transit ASes.
 Computing the path from agent to destination is not computationally expensive.
-Multipath routing is often fast since typical \emph{equal-cost multipath} (ECMP) routing algorithms simply hash a packet's flow key, and are deterministic to provide consistent quality-of-service to hosts.
+Multipath routing is fast since typical ECMP routing algorithms simply hash a packet's flow tuple and are deterministic to provide consistent quality-of-service.
 
@@
 
-We describe and analyse each of the per-flow features included in the state vector throughout \cref{sec:rethinking-the-state-space}.
+We describe and analyse the per-flow features in the state vector throughout \cref{sec:rethinking-the-state-space}.
 Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
@@
 %We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
+We denote the upstream, downstream and combined loads at a node $s$ as $\uload{s}, \dload{s}, \bload{s}$ respectively.
 Each destination node $s$ generates a reward signal, $R_{s,t}$, at every timestep $t$.
-Assume, for now, that each destination has access to some classification function $g(\cdot)$ which estimates the volume of legitimate traffic received, and expects to receive $\mathit{traffic}_s$.
-Denoting the upstream, downstream and combined loads $\uload{s}, \dload{s}, \bload{s}$ at this node:
-
+Assuming each destination has access to some classification function $g(\cdot)$ which estimates the volume of legitimate traffic received, and expects to receive $\mathit{traffic}_s$, we may compute the reward:
 \begin{subequations}
-	\vspace{-0.5cm}
+%	\vspace{-0.5cm}
 	\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
@@
 \end{subequations}
-We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, though we expect that $\bload{\cdot}$ would be the most suitable for general deployment or heterogeneous traffic patterns.
-These choices reflect where the bulk of transmitted bytes in each traffic model are observed (and the lack of this knowledge in the general case).
+We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, and suggest $\bload{\cdot}$ for general deployment or heterogeneous traffic.
+These choices reflect where the bulk of transmitted bytes in each traffic model are observed (or the lack of this knowledge).
 
@@
 ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
-Even if such heuristics or perfect knowledge aren't available in deployment, a sufficiently well-trained agent needs only to follow the policy it has learned from training in a greedy manner, allowing pre-training by a simulated environment (with perfect knowledge) to transfer to reality.
+We leave the use of such QoE metrics as future work.
+Even without online rewards, a well-trained agent needs only to follow the greedy policy it has learned from offline (simulated) training to transfer to reality.
 
@@
 $\beta$ then acts as a tunable parameter which can incentivise agents to remove flows which harm overall system health, by including the performance of the worst-performing destination.
-However, such attacks (and the effectiveness of $R_{s,t}^{\mathit{Cross}}$) are not examined by our work.
+However, we do not examine such attacks (and the effectiveness of $R_{s,t}^{\mathit{Cross}}$) at present.
 
 \subsection{Action Space}
-When monitoring a source-destination pair, an agent uses its state vector to decide which proportion of that flow's \emph{inbound} traffic should be dropped.
+When monitoring a flow, an agent uses its state vector to decide which portion of its \emph{inbound} traffic should be dropped.
 %?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
-This is implemented by installing an action via OpenFlow, instructing its host switch to drop each relevant packet with probability $p$.
 We choose to drop packets rather than impose traffic limits as it offers us a discrete action space without prior knowledge of traffic characteristics or measurement.
-Furthermore, we need not consider burstiness, fairness or implementation costs (such as per-flow bucket sizes) which could limit scalability.
+Furthermore, we need not consider burstiness, fairness or implementation costs (such as bucket sizes and tuning) which could limit scalability.
 We offer two models, which differ only on how to choose $p$.
@@
 \subsubsection{Instant control}
-Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
-These choices primarily ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
+Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set.
+These choices ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
 Since this model requires no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely opportunistic/myopic).
@@
 \subsubsection{Guarded control}
-The measurements of \textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
-We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
+\Textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
+We apply their observations by constraining how an agent treats each flow via simple finite state automata: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
 The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
@@
 This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.
-
-To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$); this is chosen above the equivalent for UDP due to TCP's higher prevalence.
-Most importantly, an agent must now choose to punish a flow multiple times in succession to cause rapid degradation, reducing variance while allowing an agent to see how a host reacts to structured changes in the environment.
-
 As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
 We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing.
+
+To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$), set this low due to TCP's higher prevalence.
+An agent must now choose to punish a flow multiple times in succession to remove its effects, reducing variance while allowing an agent to see how a host reacts to controlled conditions.
 While beneficial for legitimate flows who behave similarly to attack traffic, this introduces a lag of an extra two actions in blocking illegitimate flows.
@@
 \subsubsection{Risks}
-Our mode of action means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
+Our mode of action means that each agent carries a heavy risk of introducing collateral damage into the network.
 This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p$ is nonzero in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
 %It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
-This weakness still presents itself in modern flavours of TCP, such as TCP Cubic which in turn has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
-This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
+This weakness remains in modern flavours of TCP---TCP Cubic has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
+The prevalence of congestion-aware protocols within the Internet makes carefulness critical.
 Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{Repository link removed for double-blind reviewing.}.
@@
 %?? Here is where I would put my ramblings about system architecture...
-We present our design of a system which operates and support the effective real-world deployment of RL-based agents, acting to mitigate DDoS attacks.
+We present our design of a system which supports the effective real-world deployment of RL-based DDoS mitigation.
 \Cref{fig:sys-arch} displays this visually, separating system elements which are local to each agent from those which reside elsewhere in the network.
-In particular, we assume that each agent operates as a \emph{virtual function} (VF) adjacent to a software-defined switch.
+We assume that each agent operates as a \emph{virtual function} (VF) adjacent to a software-defined switch.
 Software-defined networks allow a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
 Furthermore, networks of this kind more naturally enable the future relocation and easy installation of learners.
-Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
+Agents communicate with their co-hosted OpenFlow-enabled switches to install probabilistic packet-drop rules.
 We describe the main purpose and operation of each module within an agent's VF, and discuss techniques to make deployment more efficient using existing technologies.
@@
 	\centering
-	\resizebox{\linewidth}{!}{\begin{tikzpicture}
+	\resizebox{0.9\linewidth}{!}{\begin{tikzpicture}
 		\node(remote){Remote};
@@
 	\caption{
-		Learned performance of Instant Control agents when benign traffic is UDP-like, using only a single feature as a basis for decisions.
-		Mean IAT, inbound packet sizes, and global state offer the best predictive performance, while most features offer marginal advantage over the unprotected baseline.
+		Learned performance of Instant Control agents when benign traffic is UDP-like, using only a single feature.
+		Mean IAT, inbound packet sizes, and global state offer the best predictive performance.
+		\vspace{-1em}
 		\label{fig:udp-feature-plots}
@@
 	\caption{
-		Learned performance of Instant Control agents when benign traffic is TCP-like, using only a single feature as a basis for decisions.
+		Learned performance of Instant Control agents when benign traffic is TCP-like, using only a single feature.
 		All of the chosen features can offer a marked improvement over no protection at all.
-		Global state and Mean IAT still offer the greatest improvement above baseline, but packet-level statistics are considerably less effective for this class of traffic.
+		Global state and Mean IAT still offer the greatest improvement above baseline.
+		\vspace{-1em}
 		\label{fig:tcp-cap-feature-plots}
@@
 	\caption{
-		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken as a basis for decisions.
+		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken.
 		This combination causes a significant improvement in the effectiveness of packet-level and per-window statistics.
@@
 
+	\resizebox{0.5\linewidth}{!}{
 	\begin{tabular}{@{}ll@{}}
@@
 	\end{tabular}
+	}
+	\vspace{-1em}
 \end{table}
@@
 		Red nodes are external, and each blue node hosts an agent.
+		\vspace{-1em}
 		\label{fig:marl-topol}
@@
 
-We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
+We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) compared against existing RL work (\emph{Marl}) with different traffic and topologies, varying the host-to-learner ratio $n$.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
 %Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
-We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold.
-Average rewards take into account any portions of time that an agent allows illegal system states.
-Several plots augment this, illustrating peak performance or the amount of time which an agent requires to learn.
+We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---these provide a rough idea of expected performance, with the highest-performing model in bold.
+We include plots which paint a different picture, showing peak performance and the amount of time an agent needs to learn.
 
@@
 
-	\resizebox{\linewidth}{!}{
+	\resizebox{0.9\linewidth}{!}{
 		\expandableinput ../tables/infocom-tree-avg-reward.tex
 	}
+\vspace{-1em}
 \end{table}
@@
 
-	\resizebox{\linewidth}{!}{
+	\resizebox{0.9\linewidth}{!}{
 		\expandableinput ../tables/infocom-ecmp-avg-reward.tex
 	}
+\vspace{-1em}
 \end{table}
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-udp-box-separate}
+	\includegraphics[width=0.9\linewidth]{../plots/tnsm-udp-box-separate}
 
@@
 		\emph{Instant} outperforms Marl for $n \in \{4, 8\}$ (with higher variance), but performs similarly to Marl at $n\in \{2, 16\}$.
-		\emph{Guarded} always underperforms compared to the other agent designs in this problem variant.
+		Here, \emph{Guarded} always underperforms compared to the other agent designs.
 %		SPF, remarkably, slightly outperforms Marl++ (with lower variance) for this more difficult topology despite being worse for smaller $n$.
@@
 %		?? REWORK/MAKE ACCURATE
+		\vspace{-1em}
 		\label{fig:udp-tree-box}
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-single}
+	\includegraphics[width=0.9\linewidth]{../plots/tnsm-tcp-box-single}
 
@@
 		Online performance for HTTP benign traffic in a single-destination network, single-agent mode.
-		\emph{Instant} and \emph{Guarded} exhibit essentially identical efficacy at $n~=~2$, protecting less traffic than Marl.
+%		\emph{Instant} and \emph{Guarded} exhibit essentially identical efficacy at $n~=~2$, protecting less traffic than Marl.
 		Only \emph{Guarded}'s performance rapidly increases with $n$, achieving considerably better median performance and lower variance than the other models.
-		The longer tails of outliers typically indicate the longer training time the new models require---we observe that \emph{Guarded} typically has considerably lower variance once it has converged on a stable policy.
+		Longer tails of outliers typically indicate longer training times---we observe that \emph{Guarded} has considerably lower variance once it has converged on a stable policy.
+		\vspace{-1em}
 		\label{fig:tcp-tree-box}
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-16-single}
+	\includegraphics[width=0.9\linewidth]{../plots/tnsm-tcp-16-single}
 
 	\caption{
-		Online performance of standard and single-agent models in a single-destination network with $n=16$ hosts per egress point and HTTP benign traffic.
-		At this level of host density, \emph{Guarded} reaches higher peak performance sooner and is considerably more consistent throughout the episode.
-		\emph{Guarded} benefits greatly from information sharing, converging to protect around \SI{75}{\percent} of TCP traffic within \SI{100}{\second}.
-		The \emph{Instant} model converges to Marl's level of performance.
+		Online performance of standard and single-agent models over time in a single-destination network with $n=16$ hosts per egress point and HTTP benign traffic.
+%		At this level of host density, \emph{Guarded} reaches higher peak performance sooner and is considerably more consistent throughout the episode.
+%		\emph{Guarded} benefits greatly from information sharing, converging to protect around \SI{75}{\percent} of TCP traffic within \SI{100}{\second}.
+%		The \emph{Instant} model converges to Marl's level of performance.
 %		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
+		\vspace{-1em}
 		\label{fig:tcp-tree-16}
@@
 
-\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
+\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} remedies transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
 Bots participating in an attack are often unable to match any bandwidth expansion in an AS due to having already saturated the capacity of their local outbound links.
@@
 Unlike our approach, SPIFFY is intended to be deployed within ASes.
-A weakness of their approach is that computing a route with enough spare bandwidth can be costly, and that low expansion factors in real network can require more ``rounds'' of testing. 
+Its weakness is that computing a route with enough spare bandwidth can be costly, and that low expansion factors in real network can require more ``rounds'' of testing. 
 Traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), which make up a sizeable portion of network traffic.
 
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a generalised SDN framework for intrusion detection, but shows the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
-Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the some evidence needed to understand how the system compares.
-
-\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a generalised SDN framework for intrusion detection, using a \emph{k-nearest neighbours} classifier to detect individual attack flows.
+Although heavyweight (and effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the evidence needed to understand how the systems compare.
+
+\Textcite{DBLP:conf/sp/SmithS18} use AS-level routing to tackle both transit-link and flooding-based attacks.
 This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
-The approach uses upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
-Because of this, traditional flooding-based attacks aren't fully alleviated.
+The approach uses BGP advertisements to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
+Because of this, volume-based attacks aren't fully alleviated.
 
@@
 \fakepara{RL in Networks}
-Application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
-Past work uses the approach as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
+RL has seen comparatively little examination in the challenge of intrusion detection/prevention.
+Past works use the approach as a traditional classifier \cite{shamshirband2014anomaly,DBLP:conf/mates/ServinK08}.
 Given that one of the main strengths of RL techniques is the ability to control (and adapt through) ongoing interaction, these works don't apply the paradigm at its fullest potential.
 
-A past example which embodies a different use learns when it is best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
+An effective use has been to learn when it is best to \emph{communicate} and share knowledge between anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
 %The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
 %The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
+Outside of this, there has been growing interest in the use of RL in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming.
 
@@
 Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
-To overcome the high latency of action computation via a neural network, two agents are trained, handling aspects of short and long flows respectively.
-The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
-The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.
+To overcome the high latency of action computation via a neural network, two agents are trained, handling aspects of short (priority thresholds) and long flows (bespoke allocations) respectively.
 
@@
 \section{Conclusions and Future Work}
-Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
-Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.
+Through this paper, we have introduced reinforcement learning and its relevance to network intrusion prevention.
+We believe that the ability to learn network control as an online feedback loop makes it particularly effective in modern networks.
 

### 2019-07-29 paper/infocom.tex
New intro for Infocom...

@@
 \section{Introduction}
-?? Redo intro to be more system focussed? Needs to be suitably different. Perhaps reduce focus on evolving, make it way more about "protocol-agnostic"...
-
-Network anomaly detection and intrusion detection/prevention are continually evolving problems, compounded by the partial, non-\emph{independent and identically distributed} (IID) view of data at each point in the network.
-Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, and availability without being detected \cite{DBLP:journals/comsur/BhuyanBK14}.
-These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
-This issue is inherent to \emph{misuse-} or \emph{signature-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making effective use of statistical measures \cite{DBLP:journals/comsur/BhuyanBK14}.
-
-While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \citeauthor{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems---a distinct lack of real-world system deployments \cite{DBLP:conf/sp/SommerP10}.
-To quite a large extent, this remains the case today.
-These authors posit that their application is made difficult due to significant operational differences from standard ML tasks, including: the high cost of errors and extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}; a general lack of recent, openly available (and high-quality) training data; and diversity of network traffic across varying timescales combined with significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
-Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
-Learning is made harder still by the challenges encountered when learning from unlabelled (often partial) data.
-All of these factors greatly inflate the difficulty of the detection problem.
-
-%??
-Make it clearer here what problem I specifically want to solve: principally a particular class of DDoS attacks; volume-based DDoS attacks.
-Amplification attacks are just a specialisation, this can be made more obvious.
-I think I need to be clearer about the \emph{intended deployment environment} of service hosts (i.e., not ISPs).
-
-For certain classes of problem e.g., volume-based \emph{distributed denial of service} (DDoS) attacks, \emph{reinforcement learning} (RL) offers another perspective.
+%?? Redo intro to be more system focussed? Needs to be suitably different. Perhaps reduce focus on evolving, make it way more about "protocol-agnostic"...
+%?? Just focus on DDoS, not anomaly detection writ-large
+
+%?? DDoS attacks are bad...
+Distributed Denial-of-Service (DDoS) attacks are a constant threat to online services.
+Botnets like Mirai \cite{DBLP:conf/uss/AntonakakisABBB17} have been the primary driver of attacks against businesses like GitHub \cite{github-ddos-2018} (\SI{1.2}{\tera\bit\per\second}) and Dyn \cite{dyn-ddos-2016} (\SI{1.35}{\tera\bit\per\second}) in the past few years, bringing them offline by completely swamping the capacity of their servers and transit links.
+%Memcrashed UDP amplification, 51000 factor for GitHub.
+Existing DDoS defences typically rely on protocol behaviours and signatures, or upon proprietary middleboxes and traffic scrubbing services.
+%?? What happens when we roll out new protocols? Particularly those which focus on obfuscation? Behaviours become completely opaque, and ossification works against us.
+These defence strategies are effective around their time of invention, but rely upon network ossification to remain relevant.
+
+However, the internet is always in flux.
+The constant deployment of new services and protocols means that traffic is \emph{non-stationary} and evolving.
+This ranges from subtle variation, such as new TCP congestion control algorithms \cite{rfc8312}, to the rapid rollout \cite{DBLP:conf/pam/RuthPDH18} of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
+%?? What happens when we roll out new protocols? Particularly those which focus on obfuscation? Behaviours become completely opaque, and ossification works against us.
+Reliance on a fixed network is at odds with reality, when QUIC causes a large proportion of (traditionally) constant bitrate UDP traffic to become congestion-aware.
+Behaviours and signatures become harder still to extract if the main purpose of a protocol is obfuscation (again, as in QUIC).
+%?? Also attacks evolve.
+DDoS attacks themselves also evolve in both short and long timescales (i.e., a sequence of attack strategies vs.\@ unseen attacks) \cite{DBLP:conf/spw/KangGS16}.
+Thus, we must look at defence models which can identify characteristics and behaviours of DDoS traffic \emph{automatically}.
+
+%Network anomaly detection and intrusion detection/prevention are continually evolving problems, compounded by the partial, non-\emph{independent and identically distributed} (IID) view of data at each point in the network.
+%Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, and availability without being detected \cite{DBLP:journals/comsur/BhuyanBK14}.
+%These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
+%This issue is inherent to \emph{misuse-} or \emph{signature-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making effective use of statistical measures \cite{DBLP:journals/comsur/BhuyanBK14}.
+
+%?? Then talk about ML. And its costs, say that its ADS/IDS weaknesses extend to DDoS prevention.
+While \emph{machine learning} (ML) approaches seem suitable for this problem, there is a distinct lack of real-world deployments in the broader field of intrusion detection/prevention.
+\Textcite{DBLP:conf/sp/SommerP10} posit that this stems from greater risks than other ML tasks: the high cost of errors, low tolerance for false positives; a lack of recent, high-quality training data; and traffic diversity and evolution.
+It is immediately clear that these challenges also apply to DDoS prevention.
+Learning is made harder still by the challenges encountered when learning from unlabelled/partial data.
+%DDoS prevention exhibits many of these issues.
+%?? Then Bam! Introduce RL.
+
+\emph{Reinforcement learning} (RL) offers another perspective.
 %?? Unclear explanation of RL here?
-RL agents operate by following a \emph{policy} to interact with or control a system, while at the same time using observed performance metrics and deliberate exploration to dynamically improve this policy.
-In this way the role of a RL agent differs from that of a standard classifier, adaptively reacting to threats by assuming the role of a feedback loop for network optimisation, typically to safeguard service guarantees.
-In a sense, this allows us to ``overcome'' some of the difficulties of the detection problem by monitoring \emph{performance characteristics and consequences} in real-time; by looking for (and controlling) the effect rather than the cause.
-Long-term, we expect that the value of RL-based defence systems will be to augment what existing misuse-based solutions can provide, by automatically alerting, recording and controlling what are believed to be illegal system states.
-The goal of this work is much less general; we aim to prevent volume-based DDoS attacks with the aid of RL-based techniques (an important goal in its own right), while bringing to light the flexibility and applicability of these techniques in the security domain.
+RL agents follow a \emph{policy} to interact with or control a system, while at the same time using observed performance to improve this policy online.
+An agent then adaptively reacts to threats by acting as a feedback loop for network optimisation---this allows us to overcome some of these difficulties by monitoring the consequences of each action in real-time.
+We aim to prevent volume-based DDoS attacks with the aid of RL-based techniques, while bringing to light the flexibility of these techniques in the security domain.
 %Whether it takes direct control of the network, or is used indirectly to optimise a key part of another system, more powerful `deep' RL techniques (and well-founded action spaces) aren't yet well explored for network IDS/IPS.
 %These range from more modern training algorithms \cite{DBLP:journals/corr/SchulmanWDRK17, DBLP:conf/icml/SchulmanLAJM15}, to evolutionary strategies \cite{DBLP:journals/corr/SalimansHCS17, DBLP:journals/corr/abs-1802-08842}, hierarchical action composition \cite{DBLP:journals/corr/abs-1710-09767}, and competitive multi-agent learning \cite{DBLP:journals/corr/abs-1710-03748}.
-
-To date, there have been few applications of this class of algorithms towards intrusion detection and prevention which make use of their full potential for online control, rather than using them as the basis for a classifier.
-We aim to take steps to redress this and establish their proper capabilities, beyond simple ``blind application''.
-%?? Expand as required
-What approaches do exist are aimed towards the task of adaptive online DDoS mitigation, and rely upon learning to control probabilistic packet drop.
-%?? THAT IS A MAJOR CONTRIB, MENTION IT EVERYWHERE YOU CAN
+%?? This raises the question: how do we do this effectively? Agents need state and reward info, which can be a lot of data transfer. THis gets worse (hint: we have the answer...)
+This introduces a new set of problems: what effective state/action spaces should look like, how state and reward information should be efficiently propagated, and how this can be achieved on commodity hardware.
 
 %?? Discuss the most important conclusions before the outline.
-We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
-The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
-In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
-Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
-Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
+We find that the existing RL work for this task fails to account for congestion-aware traffic (i.e., TCP, QUIC) and environments with high host density at ingress points.
+To remedy this, our agents make throttling decisions per-source, and show how an action model built on domain-specific knowledge can increase performance.
+This requires additional engineering: updating RL agents from multiple traces per timestep, smart work scheduling, and an architecture featuring (but minimising) the use of software-defined networks (SDN).
+%In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
+This improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that the use of domain knowledge achieves excellent performance at high host density.
+Crucially, the models are protocol- and content-agnostic to offer future-proofing against the rollout of future protocols.
 %?? Also algorithmic enhancements such as multiple actions per timestep, 
@@
 We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
-	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
+	\item Two source-level granularity models for RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models) (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
 	\item A description of the system architecture which supports the real-world deployment of such agents (\cref{sec:system-architecture}).
-	\item An investigation into suitable features for an automatic DDoS mitigation system (\cref{sec:rethinking-the-state-space}).
+	\item An investigation of suitable features for this task (\cref{sec:rethinking-the-state-space}).
 %	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
-	\item A comparison of the effectiveness of different RL algorithms in these agents (\cref{sec:the-results-of-doing-so}).
-	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
+	\item A comparison of the effectiveness of different RL algorithms in these agents, and against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
+%	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
 \end{itemize}

### 2019-07-29 paper/infocom.tex
Somehow reduced to the Infocom page limit...

@@
 We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
-	\item Two source-level granularity models for RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models) (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
+	\item Two source-level granularity models for RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models) and suitable features (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
 	\item A description of the system architecture which supports the real-world deployment of such agents (\cref{sec:system-architecture}).
-	\item An investigation of suitable features for this task (\cref{sec:rethinking-the-state-space}).
+%	\item An investigation of suitable features for this task (\cref{sec:rethinking-the-state-space}).
 %	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
-	\item A comparison of the effectiveness of different RL algorithms in these agents, and against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
+	\item A performance comparison of different RL algorithms within these agents, and against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
 %	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
@@
 
-\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
-Attackers achieve this by exploiting operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through volume of requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
+\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, to inflict financial harm or as an act of vandalism.
+Attackers achieve this by exploiting system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
 Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
 
-Volume-based attacks are the most relevant to our work, with two varieties being especially prevalent today.
+Volume-based attacks are the most relevant to our work, with two varieties being especially prevalent.
 \emph{Amplification attacks} exploit services who send large replies in response to small requests, with UDP-based services like DNS and NTP seeing most abuse \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
-Malicious hosts send many small requests, spoofed using the victim's address, causing many large replies to be sent to the intended target---increasing a botnet's throughput while leaving each participant anonymous.
+Malicious hosts send many small requests, spoofed using the victim's address, causing many large replies to be sent to the intended target---increasing a botnet's throughput while keeping participants anonymous.
 \emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
@@
 We assume at any point in time $t$, an agent knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
-This model of system interaction is best described as a \emph{Markov Decision Process} (MDP), where agents aim to maximise the \emph{expected reward} received.
+This model of system interaction is a \emph{Markov Decision Process} (MDP), where agents aim to maximise the \emph{expected reward} received.
 RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken.
@@
 To encourage early exploration, actions with maximal value can be chosen each time except when taking random actions with probability $\epsilon$---an \emph{$\epsilon$-greedy policy}).
-\Cref{fig:tilecode-select} extends the prior working example to show how the value of each action is computed (and which action would be chosen by a greedy policy).
+\Cref{fig:tilecode-select} shows how the value of each action is computed (and which action would be chosen by a greedy policy).
 
@@
 There are concrete reasons to use RL methods in this domain, beyond handling non-stationarity.
-Misclassification can be a serious problem, introducing \emph{collateral damage} in a DDoS prevention system.
-RL's feedback-loop model lets us monitor flows \emph{after} an action is taken to perhap allow forgiveness of mistakenly punished flows.
-This relies upon a flow-by-flow view of the state space, but by combining knowledge of current state with the last applied action, a punished flow which falls off identically to a legitimate flow can be rescued.
+Misclassification is a serious problem, introducing \emph{collateral damage} in a DDoS prevention system.
+RL's feedback-loop model lets us monitor flows \emph{after} an action is taken to perhaps allow forgiveness of mistakenly punished flows.
+This relies upon a flow-by-flow view of the state space, but knowledge of the current state and last applied action detect whether punished flows fall off identically to legitimate flows.
 
@@
 Other studies suggest that there are particularly useful features which make the task of online DDoS flow identification feasible.
-Aggregate network load observed at various locations suggests the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
-Volume-based statistics (counts, counts per duration, average packet sizes) have seen use in $k$-nearest neighbours classifiers trained to detect DDoS attacks in progress \cite{DBLP:conf/dsn/LeeKSPY17}.
+Link loads observed at various locations suggest the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
+Volume-based statistics (counts, counts per duration, average packet sizes) have seen use in $k$-nearest neighbours classifiers trained to detect DDoS attacks \cite{DBLP:conf/dsn/LeeKSPY17}.
 Most importantly, there is evidence showing that bandwidth expansion may cause behavioural changes which differ between benign and attack traffic \cite{DBLP:conf/ndss/KangGS16}.
-%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
-%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
-
-%\section{A Plan, of Sorts}
-%
-%\begin{enumerate}
-%	\item The main case for contribution in what I have so far:
-%	\begin{itemize}
-%		\item Past work reliant on unrealistic network models: tcp-like behaviour (and its effects on collateral damage) not captured, disjoint ranges of traffic distribution (no benign heavy-hitters), ISP-like topology.
-%		\item I offer more realistic network emulation environment, better treatment of protocol/traffic characteristics.
-%	\end{itemize}
-%	\item Forthcoming: rethinking state/action spaces to operate at a finer level of granularity.
-New network model (live tcp back-and-forth), allows us to test collateral damage assumptions in a more realistic manner (and show clear case for moving beyond work of malialis and kudenko)
-%\end{enumerate}
-
-%\subsection{Threat Model}
-%An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
-%Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
-%We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
-%The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
-%The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
-%
-%We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
-%While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
-%The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
-%These are APT-level capabilities, whose exploration presents a rich source for future work.
 
 \section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
-%Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
 We hypothesise that per-flow agent designs will offer better accuracy than current RL-based agents for DDoS mitigation.
-%However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
-We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
+We describe and justify our approach and algorithmic improvements, presenting two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
 
 \subsection{System Design and Assumptions}
-Agents are deployed in a network, with a set of \emph{ingress/egress points} from its domain of control through which traffic can flow, and a set of protected \emph{destination nodes}.
+Agents are deployed in a network with a set of \emph{ingress/egress points} from its domain of control through which traffic can flow, and a set of protected \emph{destination nodes}.
 These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
@@
 \subsubsection{Per-tile updates}
-To adjust the value of each activated tile with a different magnitude and direction, we found it effective to compute a different temporal difference value \emph{for each tiling}.
-This corrects tilings which contribute too much or too little value without disturbing estimates which are already close to correct.
-While we make use of the sum of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to set $\alpha$ to a higher value without divergence.
-Compared to the standard algorithm, convergence on effective estimates occurs sooner.
+To adjust the value of each activated tile with a different magnitude and direction, we compute a different temporal difference value \emph{for each tiling}.
+This corrects tiles which contribute too much or too little value without disturbing estimates which are almost correct.
+While we make use of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to increase $\alpha$ without divergence.
+Compared to the standard algorithm, effective policies are reached sooner.
 
@@
 		\node [treenode, loadhide, right = of mainagent] (inner0) {};
-		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner0] (inner1) {};
-		\node [hidenode, below right = 0.5cm and 1cm of inner0] (inner2) {};
-		\node [treenode, loadhide, below right = 0.5cm and 1cm of inner1] (inner3) {};
+		\node [treenode, loadhide, above right = 0.2cm and 1cm of inner0] (inner1) {};
+		\node [hidenode, below right = 0.2cm and 1cm of inner0] (inner2) {};
+		\node [treenode, loadhide, below right = 0.2cm and 1cm of inner1] (inner3) {};
 		\node [treenode, loadhide, right = of inner3] (inner4) {};
-		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
-		\node [hidenode, below right = 0.5cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};
-
-		\node [hidenode, above left = 0.5cm and 0.5cm of inner1, label={[labeltexts]above:Agent 2}] (agent2) {};
-		\node [hidenode, below left = 0.5cm and 0.5cm of inner2, label={[labeltexts]below:Agent 3}] (agent3) {};
+		\node [treenode, loadhide, above right = 0.2cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
+		\node [hidenode, below right = 0.2cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};
+
+		\node [hidenode, above left = 0.2cm and 0.5cm of inner1, label={[labeltexts]above:Agent 2}] (agent2) {};
+		\node [hidenode, below left = 0.2cm and 0.5cm of inner2, label={[labeltexts]below:Agent 3}] (agent3) {};
 
@@
 
-Our state space combines global state (network link loads) with per-flow measurements.
+Our state space combines global state (network load) with per-flow measurements.
 Each is tile-coded with 8 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
+Features are tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
 
@@
 Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}): the bandwidth measurements regularly received from monitors in the environment.
-For any flow, an agent then computes the path it would take through the network.
-The incoming load along the first hop, last hop, and tertiles of the path may then be tile-coded together.
-In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
+For any flow, an agent computes the path it would take through the network.
+The incoming load along the first hop, last hop, and tertiles of the path are then tile-coded together.
+In the event that this path is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
 \Cref{fig:global-state-path} illustrates the process.
-
 Global state built in this way is compatible with multipath, multi-destination networks, offering support for diverse deployment environments from endpoint servers up to transit ASes.
-Computing the path from agent to destination is not computationally expensive.
-Multipath routing is fast since typical ECMP routing algorithms simply hash a packet's flow tuple and are deterministic to provide consistent quality-of-service.
+Computing a flow's path is not computationally expensive: multipath routing is fast since typical ECMP routing algorithms simply hash a packet's 5-tuple and are deterministic to provide consistent quality-of-service.
 
@@
 
-We describe and analyse the per-flow features in the state vector throughout \cref{sec:rethinking-the-state-space}.
-Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
+\Cref{fig:udp-feature-plots} shows the effectiveness of per-flow features for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), on a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
+The plots demonstrate that different protocols and traffic classes are best defended by different features---every feature presented has value in a complete model.
+Packet-level and per-window statistics are at their based when combined with the last action for both UDP-like and TCP-like traffic (omitted).
+All features converge to their highest-observed performance within around \num{4000} timesteps.
+In general, some of the most effective features are global state, mean IAT, mean inbound packet size and $\Delta$ rates.
+The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
+If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have no correlation with the class of traffic a flow contains.
 
@@
 While our use and definition of $g(\cdot)$ appears nebulous, there are many ways to infer this quantity in practice.
-End-host servers may make use of canary flows or other active measurement schemes, or employ existing quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
-ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
-We leave the use of such QoE metrics as future work.
-Even without online rewards, a well-trained agent needs only to follow the greedy policy it has learned from offline (simulated) training to transfer to reality.
+End-host servers may make use of canary flows or active measurement schemes, and can employ quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
+ASes and transit networks may make use of reports received from downstream, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
+We leave the use of such data as future work.
+Even without online rewards, a well-trained agent needs only to follow the greedy policy it has learned from offline (simulated) training.
 
@@
 %?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
-We choose to drop packets rather than impose traffic limits as it offers us a discrete action space without prior knowledge of traffic characteristics or measurement.
-Furthermore, we need not consider burstiness, fairness or implementation costs (such as bucket sizes and tuning) which could limit scalability.
+We choose to drop packets ()rather than impose traffic limits) to have a discrete action space without prior knowledge of traffic characteristics.
+Furthermore, we need not consider burstiness, fairness and parameters such as bucket sizes which could require tuning.
 We offer two models, which differ only on how to choose $p$.
-The intuition behind these designs is that an agent may be able to directly choose the level of packet drop at the expense of a larger action space (\emph{Instant control}), or that we may be able to reduce the action set (and learn a policy within fewer actions) by making use of domain knowledge (\emph{Guarded control}).
+The intuition behind these designs is that an agent may directly choose the level of packet drop at the expense of a larger action space (\emph{Instant control}), or that we may reduce the action set (and learn a policy within fewer actions) by making use of domain knowledge (\emph{Guarded control}).
 
 \subsubsection{Instant control}
-Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set.
+Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$.
 These choices ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
@@
 \Textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
-We apply their observations by constraining how an agent treats each flow via simple finite state automata: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
+We apply their observations, constraining how an agent treats each flow via simple finite state automata: restricting $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
 The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
 
-We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
+We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limits (\SI{25}{\percent} increments for UDP traffic).
 The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
@@
 As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
-We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing.
-
-To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$), set this low due to TCP's higher prevalence.
-An agent must now choose to punish a flow multiple times in succession to remove its effects, reducing variance while allowing an agent to see how a host reacts to controlled conditions.
+We found the setting $\gamma = 0.8$ to be the most effective choice during exploratory testing.
+
+To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$), chosen due to TCP's higher prevalence.
+An agent must now choose to punish a flow multiple times in succession to remove it, reducing variance while allowing an agent to see how a host reacts to controlled conditions.
 While beneficial for legitimate flows who behave similarly to attack traffic, this introduces a lag of an extra two actions in blocking illegitimate flows.
@@
 \subsubsection{Risks}
-Our mode of action means that each agent carries a heavy risk of introducing collateral damage into the network.
-This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p$ is nonzero in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
-%It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
+Our mode of action means that each agent carries a risk of introducing collateral damage into the network.
+This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p\ne0$ in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
 This weakness remains in modern flavours of TCP---TCP Cubic has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
-The prevalence of congestion-aware protocols within the Internet makes carefulness critical.
-Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{Repository link removed for double-blind reviewing.}.
+Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets and \SIrange{77}{84}{\percent} of data volume\footnote{Repository link removed for double-blind reviewing.}.
 %Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
 QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
-Furthermore, applying a single rule to many flows (as in \textcite{DBLP:journals/eaai/MalialisK15}) increases the likelihood of mistaken collateral damage.
+Furthermore, applying a single rule to many flows (as in \textcite{DBLP:journals/eaai/MalialisK15}) increases the likelihood of collateral damage.
 When considering that many hosts have an especially adverse reaction to our main means of control, flow-level granularity becomes an obvious choice.
@@
 We present our design of a system which supports the effective real-world deployment of RL-based DDoS mitigation.
-\Cref{fig:sys-arch} displays this visually, separating system elements which are local to each agent from those which reside elsewhere in the network.
-We assume that each agent operates as a \emph{virtual function} (VF) adjacent to a software-defined switch.
-Software-defined networks allow a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
-Furthermore, networks of this kind more naturally enable the future relocation and easy installation of learners.
-Agents communicate with their co-hosted OpenFlow-enabled switches to install probabilistic packet-drop rules.
+\Cref{fig:sys-arch} displays this, separating system elements which are local to each agent from those which reside elsewhere in the network.
+We operate each agent as a \emph{virtual function} (VF) adjacent to a software-defined switch.
+Software-defined networks allow a controller (or authorised hosts) to install actions, forwarding rules and logic into a switch at runtime.
+Moreover, networks of this kind more naturally enable the future relocation and easy installation of learners.
+Agents communicate with these co-hosted switches to install probabilistic packet-drop rules.
 We describe the main purpose and operation of each module within an agent's VF, and discuss techniques to make deployment more efficient using existing technologies.
@@
 \subsection{Core and RL Executor}\label{sec:core-and-rl-executor}
-The core module is the main loop for decision-making in an agent's architecture.
+The core module is the main loop in an agent's architecture.
 At each timestep, the core receives information about which flows have arrived and should be acted upon from the \emph{TRS Scheduler}.
@@
 The RL algorithm then returns an action.
-Each action is passed to the database, which computes and returns the desired packet drop rate according to the agent model (instant/guarded) while updating flow state.
-This is then converted into an OpenFlow message carrying the packet drop rules; these are sent in a single message to the agent's switch using the same groupings produced by the scheduler.
+Each action is passed to the database, which computes and returns a packet drop rate according to the agent model (instant/guarded) while updating flow state.
+This is then converted into an OpenFlow message carrying packet drop rules; these are batched to the agent's switch using the same groupings produced by the scheduler.
 Finally, timing information is passed back into the scheduler to refine its estimates about how much work should be scheduled in the next timestep.
 
-State space sizes guarantee that an \emph{Instant} agent's policy remains under \SI{520}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies: $\sim$\SI{17.8}{\kibi\byte} from our experiments.
+State space sizes guarantee that an \emph{Instant} policy remains under \SI{520}{\kibi\byte}, though our sparse representation typically leads to far smaller policies: $\sim$\SI{17.8}{\kibi\byte} from our experiments.
 \emph{Guarded} policies are \SI{30}{\percent} of this size.
-As we have described earlier, action updates require a constant number of floating point operations---\num{160} floating point additions and \num{32} multiplications per update of $\wvec{}$ with per-tile updates, above the \num{160} additions required to choose an action.
+As described earlier, action updates require a constant number of floating point operations---\num{160} floating point additions and \num{32} multiplications per update of $\wvec{}$ with per-tile updates, with \num{160} additions required to choose an action.
 The vast majority of these operations can be vectorised.
@@
 \subsection{Stats API and Collectors}
-To act effectively, the agents we have designed require a rich assortment of information from the network and other agents.
-Each agent then exposes an API through which statistics and policy information may arrive.
+Agents require information from the network and one another to be effective.
+%Each agent then exposes an API through which statistics and policy information may arrive.
 %?? API between agent and switches/other agents? Push/Pull API on stats? How do they share experience? Could broadcasts aid both?
-
 %?? What does experience transfer look like? Why? Explain that every agent has a different policy, and why this is beneficial
@@
 It's noted that a transition may be high-value or surprising to one agent, while well-known to another, causing each to produce different policy updates from the same unit of experience.
-For this reason we choose not to transfer policy deltas between agents, causing each to learn its own policy.
-We leave the study of which scheme achieves better performance to future work.
+For this reason we do not transfer policy deltas between agents, causing each to learn its own policy.
+Which scheme achieves better performance is left to future work.
 
 %?? How do stats get from switcehs to the agent?
-Load collectors and estimators periodically push their observations to each active agent VF.
-In our current implementation, load statistics are gathered via VFs active at each network switch, though we expect that OpenFlow stats requests, NetFlow or SNMP data may be used to derive load measures.
+Load collectors and estimators periodically push observations to each active agent VF.
+In our current implementation, load statistics are gathered via VFs active at each network switch, though we expect that OpenFlow stats requests, NetFlow or SNMP data may be used to derive these cheaply.
 Transferring state to agents and experience sharing can both be made more efficient through effective use of broadcast addressing in a target network.
-Depending on the capabilities of switches in the network, the estimator can either send actual benign traffic estimates or parameters for use in a reward function.
+Depending on the capabilities of switches in the network, the estimator can either send benign traffic estimates or parameters for use in a reward function.
 
 Gathering and transmission of load/flow statistics would be difficult to perform quite as often as an emulated environment allows.
-The measurements acquired in such a scenario are far more likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
+However, the measurements acquired in such a scenario are far more likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
 
 \subsection{Flowstate Database}
-For each flow 5-tuple, we hold two state vectors containing the features described in \cref{sec:rethinking-the-state-space}---the current state, and the state which induced the last action.
+For each flow 5-tuple, we hold two state vectors containing the features described in \cref{sec:feature-space}---the current state, and the state which induced the last action.
 To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set.
@@
 \subsection{TRS Scheduler}
-Acting on a possibly unbounded set of flows in each timestep introduces some potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and the risk that flow states become outdated.
+Acting on an unbounded set of flows in each timestep introduces potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and the risk that flow states become outdated.
 At their worst, these risks present additional attack surface to an adversary.
@@
 
-Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
-We expect this should be tuned based on the frequency at which statistics arrive.
-The amount of processed flows per deadline depends on the agent design (FLOP count according to policy size), but also on the amount of flow telemetry data needing processed---our current implementation is written in python, restricting most of this handling to a single thread.
+Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations of short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}--- an agent is typically able to process around 3 flows in this time.
+We expect deadlines should be tuned based on the frequency at which statistics arrive.
+The amount of processed flows per deadline depends on the agent design (FLOP count, policy size), but also on the amount of flow telemetry data needing processed---our current implementation is written in python, restricting this handling to a single thread.
 An implementation in a systems language such as Rust or C++ would allow faster concurrent processing.
@@
 There is a risk that so much work can be queued up that an agent is never able to act on some attack flows.
-A solution is to impose an upper bound on the amount of action computations/policy updates that can be performed before a work list is regenerated.
-This removes the guarantee that all flows will be visited fairly often, but if updates occur regularly then this random sampling may be sufficient to achieve good performance.
+A solution is to impose an upper bound on the amount of action computations/policy updates that can be performed before the work list is regenerated.
+This removes the guarantee that all flows will be visited often, but if updates occur regularly then this sampling may be sufficient to achieve good performance.
 
@@
 
-		\node[below=0.2 of remote](swpos){};
+		\node[below=0.05 of remote](swpos){};
 		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
@@
 
-		\node[below=2.5 of remote.west](linestart){};
+		\node[below=2.3 of remote.west](linestart){};
 		\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
@@
 
-		\node[below=2.6 of remote](local){Local};
+		\node[below=2.4 of remote](local){Local};
 
@@
 
-		\node[below=0.2 of local](aswpos){};
+		\node[below=0.05 of local](aswpos){};
 		\node[fill=white!80!uofgthistle, draw=black, minimum height=3.5cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
@@
 	}
+\vspace{-1em}
 \end{figure}
-
-\section{Rethinking the State Space}\label{sec:rethinking-the-state-space}
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/ftprep-cap-box}
+	\includegraphics[width=0.8\linewidth]{../plots/ftprep-cap-box}
 	\vspace{-1cm}
@@
 		Learned performance of Instant Control agents when benign traffic is UDP-like, using only a single feature.
-		Mean IAT, inbound packet sizes, and global state offer the best predictive performance.
+%		Mean IAT, inbound packet sizes, and global state offer the best predictive performance.
 		\vspace{-1em}
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/ftprep-tcp-cap-box}
+	\includegraphics[width=0.8\linewidth]{../plots/ftprep-tcp-cap-box}
 	\vspace{-1cm}
@@
 		Learned performance of Instant Control agents when benign traffic is TCP-like, using only a single feature.
-		All of the chosen features can offer a marked improvement over no protection at all.
-		Global state and Mean IAT still offer the greatest improvement above baseline.
+%		All of the chosen features can offer a marked improvement over no protection at all.
+%		Global state and Mean IAT still offer the greatest improvement above baseline.
 		\vspace{-1em}
@@
 \end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/ftprep-tcp-laf-cap-box}
-	\vspace{-1cm}
-	\caption{
-		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken.
-		This combination causes a significant improvement in the effectiveness of packet-level and per-window statistics.
-		\label{fig:tcp-laf-feature-plots}
-	}
-\end{figure}
-
-The main element required to move to a per-source model is a feature set which offers high predictive power, such that behavioural differences are readily apparent to an agent.
-Elaborating further on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and offer an investigation into their use alongside different traffic types:
-%?? We use these features, and why...
-
-\fakepara{Global state}
-This is the vector of current load measurements along a flow's path introduced in \cref{sec:feature-space}.
-These values indicate the overall health of the network, and crucially are all measurements which an agent directly controls.
-
-\fakepara{Source IP address}
-While ordinarily trivial to spoof (and thus of limited use for many classes of attack), reflectors are themselves legitimate services being abused by spoofing attackers.
-As a result, they communicate with attack victims using their own IP address.
-In real-world scenarios the addresses of a set of reflector nodes might exhibit similarity (such as the same CIDR block) due to network uncleanliness \cite{DBLP:conf/imc/CollinsSFJWSK07}, e.g., similarly unhardened services exposed by a single organisation.
-
-\fakepara{Last action taken}
-This encodes an agent's current belief in the maliciousness of a flow.
-This feature also potentially allows forgiveness, serving as a reference point for determining whether a source mistakenly marked as malicious exhibits different falloff behaviour after punishment.
-It's important to note that this feature only makes sense once combined with another flow feature, and never appears individually tile-coded.
-
-\fakepara{Flow duration and size}
-Features which describe the length of time a connection has been active, and the amount of data transferred within that time.
-An extraordinarily long flow, having sent a lot of data, could be more likely to be an amplifier: though most (\SI{62}{\percent}) waves of amplifier traffic last shorter than \SI{15}{\minute} \cite{DBLP:conf/raid/KramerKMNKYR15}, this is considerably longer than the typical length of an HTTP request/response.
-
-\fakepara{Correspondence ratio}
-The ratio between upstream and downstream traffic for a source IP.
-We define this to be $C_X = \min(\uload{\cdot}, \dload{\cdot})/\max(\uload{\cdot}, \dload{\cdot})$, where a value close to 0 indicates strong asymmetry.
-
-\fakepara{$\mathbf{\Delta}$ Send/receive rate}
-The change in traffic rates caused by the last action.
-Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible in these fields.
-
-\fakepara{Mean inter-arrival time (IAT)}
-A measure of how often packets arrive at the agent's parent switch; low IATs indicate a high number of packets per second, and can be a possible marker of malicious behaviour.
-We only make use of the mean IAT of \emph{inbound} traffic.
-
-\fakepara{(Per-window) packet count}
-The amount of packets sent to/from a source over a flow's lifetime (or the current window of measurement), similar in use to flow size and mean IAT.
-
-\fakepara{Mean packet size per window}
-Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
-Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.
-
-The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
-If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have little to no correlation with the class of traffic a flow might contain.
-Our aim was to design around this constraint as a form of future-proofing.
 
@@
 
-All of the above features, save for global state, are 1-dimensional.
-\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), using a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
-\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends being observed when the same technique is applied with UDP traffic (omitted).
-%?? Core findings---different protocols need different features, so everything we proposed above has a use!
-The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
-All features converge to their highest-observed performance within around \num{4000} timesteps.
-In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.8\linewidth]{../plots/ftprep-tcp-laf-cap-box}
+%	\vspace{-1cm}
+%	\caption{
+%		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken.
+%		\label{fig:tcp-laf-feature-plots}
+%	}
+%\end{figure}
+
+%\section{Rethinking the State Space}\label{sec:rethinking-the-state-space}
+%
+%Per-flow agents require features which offer high predictive power, making behavioural differences readily apparent.
+%Expanding on the measures discussed in \cref{sec:motivation}, we believe the following features to be useful (and humanly justifiable), and investigate their use on different traffic types:
+%%?? We use these features, and why...
+%
+%\fakepara{Global state}
+%This is the vector of current load measurements along a flow's path introduced in \cref{sec:feature-space}.
+%These values indicate the overall health of the network, and crucially are all measurements which an agent directly controls.
+%
+%\fakepara{Source IP address}
+%Ordinarily trivial to spoof, though reflectors are themselves legitimate services being abused by spoofing attackers.
+%As a result, they communicate with attack victims using their own IP address.
+%In real-world scenarios the addresses of a set of reflector nodes might exhibit CIDR block similarity \cite{DBLP:conf/imc/CollinsSFJWSK07}, e.g., similarly unhardened services exposed by a single organisation.
+%
+%\fakepara{Last action taken}
+%An agent's current belief in the maliciousness of a flow.
+%This feature also serves as a reference point for determining whether a flow is actually malicious based on its response.
+%This feature only makes sense when combined with another flow feature, and never appears individually tile-coded.
+%
+%\fakepara{Flow duration and size}
+%Features which describe the length of time a connection has been active, and the amount of data transferred within that time.
+%An extraordinarily long flow, having sent a lot of data, could be more likely to be an amplifier: though most (\SI{62}{\percent}) waves of amplifier traffic last shorter than \SI{15}{\minute} \cite{DBLP:conf/raid/KramerKMNKYR15}, this is considerably longer than the typical length of an HTTP request/response.
+%
+%\fakepara{Correspondence ratio}
+%The ratio between upstream and downstream traffic for a source IP.
+%We define this to be $C_X = \min(\uload{\cdot}, \dload{\cdot})/\max(\uload{\cdot}, \dload{\cdot})$, where a value close to 0 indicates strong asymmetry.
+%
+%\fakepara{$\mathbf{\Delta}$ Send/receive rate}
+%The change in traffic rates caused by the last action.
+%Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible in these fields.
+%
+%\fakepara{Mean inter-arrival time (IAT)}
+%A measure of how often packets arrive at the agent's parent switch; low IATs indicate a high number of packets per second, and can be a possible marker of malicious behaviour.
+%We only make use of the mean IAT of \emph{inbound} traffic.
+%
+%\fakepara{(Per-window) packet count}
+%The amount of packets sent to/from a source over a flow's lifetime (or the current window of measurement), similar in use to flow size and mean IAT.
+%
+%\fakepara{Mean packet size per window}
+%Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
+%Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.
+%
+%The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
+%If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have no correlation with the class of traffic a flow contains.
+%
+%All of the above features, save for global state, are 1-dimensional.
+%\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), using a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
+%%\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends being observed when the same technique is applied with UDP traffic (omitted).
+%%?? Core findings---different protocols need different features, so everything we proposed above has a use!
+%The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
+%Packet-level and per-window statistics are at their based when combined with the last action for both UDP-like and TCP-like traffic (omitted).
+%All features converge to their highest-observed performance within around \num{4000} timesteps.
+%In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
 %?? How do they do when combined after individual training? Pretty well, especially for TCP.
@@
 %Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
-We contribute network models built around live testing of reactive TCP and UDP traffic in an SDN-enabled environment, which is adaptable to arbitrary topologies, with an explicit focus on preserving their real-time dynamics in a way that trace-based evaluation cannot.
-First and foremost, we are interested in representative load and packet inter-arrival characteristics and in how these characteristics evolve in response to actions.
-We introduce these models because we are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
+Our network model allows live testing of reactive TCP and UDP traffic in an SDN environment, over arbitrary topologies, with an explicit focus on preserving real-time dynamics in a way that trace-based evaluation cannot.
+We are primarily interested in load and packet inter-arrival characteristics, and in how they evolve in response to agent actions.
+%We aim to capture interactive, correlated back-and-forth exchanges associated with live traffic.
 %Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
@@
 We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
-All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
-If several ports offer different (equal-length) paths to a destination, a consistent random port is chosen from the flow-hash by an OpenFlow \emph{Group action} (in \emph{select} mode).
-If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
-The controller is also responsible for computing how switches respond to ARP requests: this need arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
+Internal routers begin with knowledge of the shortest path to each protected host, while inbound flows register their return path at each hop, ensuring consistent traffic conditions for each flow.
+OpenFlow \emph{group actions} provide consistent multipath routing where needed. 
+%If several ports offer different (equal-length) paths to a destination, a consistent random port is chosen from the flow-hash by an OpenFlow \emph{Group action} (in \emph{select} mode).
+%If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
+The controller is also responsible for computing how switches respond to ARP requests: this need arises due to the reliance upon Linux's networking stack for live applications.
 %We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.
@@
 To model legitimate TCP traffic, server nodes run an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
-Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
-Hosts and clients use the TCP Cubic \cite{rfc8312} flavour of TCP.
-Each host's download rate is limited to match whatever maximum bandwidth was assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.).
-On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
-This presents a more balanced distribution of flow duration and size, with large files included to model elephant flows.
+Benign hosts repeatedly request resources from the server via libcurl.
+Both use the TCP Cubic \cite{rfc8312} flavour of TCP.
+A host's download rate is limited to match its assigned maximum bandwidth, and requests several files known to exist within a website, followed by dependencies for each (stylesheets, images, etc.).
+This presents a balanced distribution of flow duration and size, with large files included to model bulk transfer.
+On completion, a host changes its IP to generate separate statistics per-flow.
 
 \subsection{UDP (Opus) Traffic Model}
-VoIP traffic exhibits very different characteristics to the above model; packet arrivals are highly periodic due to real-time application requirements, flows have a constant bitrate, and do not react substantially to lost packets.
-Interestingly, DDoS attack traffic is known to share many of these characteristics, offering an interesting detection problem.
-
-We present a VoIP traffic model\footnote{Repository link removed for double-blind reviewing.} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward video game communities.
+VoIP traffic is heavily different to the above model; packet arrivals are highly periodic due to real-time requirements, flows are constant bitrate, and do not react substantially to lost packets.
+DDoS attack traffic is known to share many of these characteristics, offering an interesting detection problem.
+We present a traffic model\footnote{Repository link removed for double-blind reviewing.} based on Discord\footnote{\url{https://discord.gg}}, a messaging and VoIP platform.
 %We present a VoIP traffic model\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward video game communities.
-We chose Discord as our prototype due to its publicly documented API, many open source bot frameworks for experimentation, large user base, and due to the lack of models for Opus-encoded traffic.
-All parameters (save for shortening of silent periods) are derived from passive measurement, API documentation, and the bot community's observations.
+We use this as our prototype due to its public API, open source frameworks, and the lack of models for Opus-encoded traffic.
+All parameters (save for truncating silent periods) are derived from passive measurement, API documentation, and community observations.
 %?? Highly periodic, CBR
 
-Hosts send RTP traffic with Salsa20 encrypted payloads---audio frames of length \SI{20}{\milli\second} at a target \SI{96}{\kilo\bit\per\second}.
-We generate similar traffic at hosts by replaying anonymised traces gathered in general use and tabletop RPG servers; each trace contains only the size of each audio payload, entries denoting missed packets, and the duration of silent periods.
-We trim these silent periods to a maximum \SI{5}{\second} due to the lengthy talk/silence bursts introduced by users in RPG servers, and estimate the size of missed packets by taking an exponentially-weighted moving average over known sizes.
-Hosts punctuate audio frames with a 4-byte keepalive every \SI{5}{\second}.
-All traffic passes over a central server which groups hosts into rooms, and is forwarded to other participants; we do not replicate pre-call Websocket traffic which would be used for authentication.
-There is no peer-to-peer traffic---the server acts as a TURN relay \cite{rfc5766} for all hosts.
+Discord hosts send RTP traffic with Salsa20 encrypted payloads---audio frames of length \SI{20}{\milli\second} at a target \SI{96}{\kilo\bit\per\second}.
+Hosts generate similar traffic by replaying anonymised traces; each contains the size of an audio payload, missed packets, and silent durations.
+We trim silent periods to a maximum \SI{5}{\second} due to lengthy talk/silence bursts of users in RPG servers, and estimate the size of missed packets with an exponentially-weighted moving average.
+Hosts transmit a 4-byte keepalive every \SI{5}{\second}.
+A central server groups hosts into rooms, and forwards packets to other participants; we do not replicate pre-call Websocket traffic which would be used for authentication.
+There is no peer-to-peer traffic---the server acts as a TURN relay for all hosts.
 %?? Reflective factor among \emph{authenticated hosts}.
-
 We find that each flow occupies an expected \SI{52.4}{\kilo\bit\per\second} upstream bandwidth.
@@
 %?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
-\subsection{Attack Traffic Model}
-Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
-Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that emulation remains tractable.
 
@@
 
-We compare our work most naturally against the MARL approach introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
-We are most interested in seeing how their approach contrasts with ours across different topologies and workloads.
-Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network---we aim to see how these characteristics affect performance and learning rate.
-
-To test this, we made use of both traffic models introduced in \cref{sec:a-new-normal} (OPUS and TCP), both topologies discussed below (1-dest vs Fat-Tree), and vary the amount of hosts typically communicating over each agent's ingress/egress node.
-Additionally, we evaluated our new models in multi-agent mode (\emph{separate}, no model sharing), and in single-agent mode (\emph{single}, 0-cost perfect information sharing).
-In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
+We compare our work's performance and learning rate against Marl \cite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention, across different topologies and workloads.
+We made use of the OPUS and TCP traffic models introduced in \cref{sec:a-new-normal}, both topologies discussed below (1-dest vs Fat-Tree), and varied the amount of hosts communicating over each ingress/egress node---different networks will impose different levels of host density, e.g., popular web servers may have many more clients than egress points from their network.
+Additionally, we evaluated our new models in multi-agent (\emph{separate}, no model sharing) and i single-agent modes (\emph{single}, 0-cost perfect information sharing).
+In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (resetting policies between episodes).
 Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes.
+Malicious traffic is generated by \emph{hping3} program, generating UDP-flood traffic targeting random ports with MTU-sized packets, mapping to the large frame sizes expected in an amplification attack.
 
@@
 \subsection{Single Destination}\label{sec:single-dest}
-%?? Move description of tree topol to here.
-The network is tree-structured, where one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $\ell$ intermediate switches, which in turn each connect to $m$ egress switches.
-We then have $N_{\mathit{hosts}} = k \ell m n$.
-\Cref{fig:marl-topol} demonstrates this visually.
-
-%Although \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}, the originators of this topology, make it clear that it exists as a fairly unrepresentative example \cite{DBLP:journals/ccr/MahajanBFIPS02a}, it remains the case that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
-%It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
-%We aim to address this through \cref{sec:performance-in-an-emulated-environment}.
-
-To test this, we configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
-This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
-%The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
-%Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
-%$U_s$ was fixed at $k \ell mn+2$ \si{\mega\bit\per\second}.
-%The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
-%Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
-%These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
-
-\begin{figure}
-	\centering
-	\resizebox{0.8\linewidth}{!}{
-		\begin{tikzpicture}[
-		texts/.style = {text=black},
-		labeltexts/.style = {text=uofgsandstone},
-		treeline/.style = {draw=uofgburgundy},
-		treenode/.style = {texts, circle, centered, fill=white, treeline},
-		load/.style = {fill=uofgcobalt},
-		loadhide/.style = {fill=uofgcobalt!40!white},
-		external/.style = {fill=uofgrust},
-		externalhide/.style = {fill=uofgrust!40!white},
-		hideline/.style = {draw=uofgsandstone!40!white},
-		hidenode/.style = {treenode, hideline},
-		grow'=right
-		]
-		\node[treenode, label={[texts]above:Server}] (root) {}
-		child [treeline] { node [treenode, label={[texts]above:Core}] (sswitch) {}
-			child [treeline] { node [treenode, label={[texts]above:Leader}] (teaml) {} 
-				child [treeline] { node [treenode, label={[texts]above:Intermediate}] (inter) {}
-					child [treeline] { node [treenode, load, label={[texts]above:Agent/Egress}] (agent) {}
-						child [treeline] { node [treenode, external] (extern) {}
-							child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
-							child [hideline] { node [hidenode, externalhide] (endhost) {} }
-						}
-					}
-					child [hideline] { node [hidenode, loadhide] (endagent) {} }
-				}
-				child [hideline] { node [hidenode] (endinter) {} }
-			}
-			child [hideline] { node [hidenode] (endteaml) {} }
-			edge from parent
-			node[below, labeltexts] {$U_s$}
-		};
-
-		%\draw[-] (teaml) -- (endteaml);
-		\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
-		\node [labeltexts, right = -0.1cm of kdots] {$k$};
-		\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
-		\node [labeltexts, right = -0.1cm of ldots] {$\ell$};
-		\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
-		\node [labeltexts, right = -0.1cm of mdots] {$m$};
-		\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
-		\node [labeltexts, right = -0.1cm of ndots] {$n$};
-		\end{tikzpicture}
-	}
-	\caption{
-		Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $\ell$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
-		%	Empty nodes are considered to be internal.
-		Red nodes are external, and each blue node hosts an agent.
-		\vspace{-1em}
-		\label{fig:marl-topol}
-	}
-\end{figure}
+The network is tree-structured: one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $\ell$ intermediate switches, which in turn each connect to $m$ egress switches.
+$n$ hosts then connect through each egress switch.
+We thus have $N_{\mathit{hosts}} = k \ell m n$.
+We configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
+This is a slight simplification of the \textquote{online} experiment of \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but similar structure.
+
+%\begin{figure}
+%	\centering
+%	\resizebox{0.8\linewidth}{!}{
+%		\begin{tikzpicture}[
+%		texts/.style = {text=black},
+%		labeltexts/.style = {text=uofgsandstone},
+%		treeline/.style = {draw=uofgburgundy},
+%		treenode/.style = {texts, circle, centered, fill=white, treeline},
+%		load/.style = {fill=uofgcobalt},
+%		loadhide/.style = {fill=uofgcobalt!40!white},
+%		external/.style = {fill=uofgrust},
+%		externalhide/.style = {fill=uofgrust!40!white},
+%		hideline/.style = {draw=uofgsandstone!40!white},
+%		hidenode/.style = {treenode, hideline},
+%		grow'=right
+%		]
+%		\node[treenode, label={[texts]above:Server}] (root) {}
+%		child [treeline] { node [treenode, label={[texts]above:Core}] (sswitch) {}
+%			child [treeline] { node [treenode, label={[texts]above:Leader}] (teaml) {} 
+%				child [treeline] { node [treenode, label={[texts]above:Intermediate}] (inter) {}
+%					child [treeline] { node [treenode, load, label={[texts]above:Agent/Egress}] (agent) {}
+%						child [treeline] { node [treenode, external] (extern) {}
+%							child [treeline] { node [treenode, external, label={[texts]above:Host}] (host) {} }
+%							child [hideline] { node [hidenode, externalhide] (endhost) {} }
+%						}
+%					}
+%					child [hideline] { node [hidenode, loadhide] (endagent) {} }
+%				}
+%				child [hideline] { node [hidenode] (endinter) {} }
+%			}
+%			child [hideline] { node [hidenode] (endteaml) {} }
+%			edge from parent
+%			node[below, labeltexts] {$U_s$}
+%		};
+%		
+%		%\draw[-] (teaml) -- (endteaml);
+%		\node [labeltexts] (kdots) at ($(teaml)!0.5!(endteaml)$) {$\rvdots$};
+%		\node [labeltexts, right = -0.1cm of kdots] {$k$};
+%		\node [labeltexts] (ldots) at ($(inter)!0.5!(endinter)$) {$\rvdots$};
+%		\node [labeltexts, right = -0.1cm of ldots] {$\ell$};
+%		\node [labeltexts] (mdots) at ($(agent)!0.5!(endagent)$) {$\rvdots$};
+%		\node [labeltexts, right = -0.1cm of mdots] {$m$};
+%		\node [labeltexts] (ndots) at ($(host)!0.5!(endhost)$) {$\rvdots$};
+%		\node [labeltexts, right = -0.1cm of ndots] {$n$};
+%		\end{tikzpicture}
+%	}
+%	\caption{
+%		Network topology diagram, showing how the server and its core switch's $k$ teams are structured, with $\ell$ intermediate routers per team, connected to $m$ agents which each moderate $n$ hosts beyond a single external switch.
+%		%	Empty nodes are considered to be internal.
+%		Red nodes are external, and each blue node hosts an agent.
+%		\vspace{-1em}
+%		\label{fig:marl-topol}
+%	}
+%\end{figure}
 
 \subsection{Multiple Destinations}
-The previous topology allows for direct comparison against the state-of-the-art, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
-It is hard, however, to argue its relevance to specific classes of victim or to reason about the interactions it might have with dependent applications.
+The previous topology allows for direct comparison against the state-of-the-art, however it is hard to argue its relevance to specific classes of victim or how it might interact with applications.
 In contrast, the fat-tree topology \cite{DBLP:conf/sigcomm/Al-FaresLV08} sees regular use in real-world datacentres and scales well horizontally.
-
-%?? Come up with description of fat-tree (multi-dest) topol.
-%?? Why fat tree? regularly appears in modern datacentres.
-%?? $k=4$ fat-tree , with one pod hosting two servers $s_0,s_1$.
 We use a $k=4$ fat-tree, with one pod hosting two servers $s_0$ and $s_1$.
@@
 \subsection{Parameters}
-The algorithm parameters were set at $\alpha=0.05$, linearly annealing $\epsilon=0.2 \rightarrow 0$ by $t=3000$ in the case of Marl (\num{8000} actions per agent in the \emph{Instant/Guarded} models).
-
-Benign hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second}, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
-%The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and high densities of hosts at egress points.
-%?? Why? ...
-%Of these, the latter two are most deserving of a closer investigation, as they have stronger implications for wide-scale deployment.
-%These are important issues, particularly when we consider real-world deployment.
-%Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
-%?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
-Malicious hosts sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, though when using TCP-like traffic, each attacker's upload range had to be increased to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows.
+Algorithm parameters were set at $\alpha=0.05$, linearly annealing $\epsilon=0.2 \rightarrow 0$ by $t=3000$ in the case of Marl (\num{8000} actions per agent in the \emph{Instant/Guarded} models).
+Benign hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second}, and hosts were redrawn each episode with $\operatorname{P}(\mathit{malicious})=0.4$.
+Malicious hosts sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, increased to \SIrange{4}{7}{\mega\bit\per\second} when using TCP-like or mixed traffic to meaningfully impact benign flows.
 Given $n$ and $\operatorname{P}(\mathit{malicious})$, we see an expected malicious bandwidth \numrange{1.27}{1.87} and \SIrange{2.03}{2.18}{$\! \times U_s$} respectively.
-%The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
-
 $U_s$ was fixed at $N_{\mathit{hosts}}+2$ \si{\mega\bit\per\second} (to account for burstiness), and each link had a delay of \SI{10}{\milli\second}.
 All links had unbounded capacity, save for server-switch connections.
-These parameter choices match those of the original study to enable direct comparison, and many are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
+These parameter choices match those of the original study for direct comparison, but we justify our range of $n$ as capturing increasing scales of host activity.
 
@@
 
-We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) compared against existing RL work (\emph{Marl}) with different traffic and topologies, varying the host-to-learner ratio $n$.
+We now compare the performance of our two new models (\emph{Instant}, \emph{Guarded}) against existing RL work (\emph{Marl}) with different traffic and topologies, varying the host-to-learner ratio $n$.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
 %Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
-We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---these provide a rough idea of expected performance, with the highest-performing model in bold.
+We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold.
 We include plots which paint a different picture, showing peak performance and the amount of time an agent needs to learn.
@@
 
-	\resizebox{0.9\linewidth}{!}{
+	\resizebox{0.8\linewidth}{!}{
 		\expandableinput ../tables/infocom-tree-avg-reward.tex
@@
 
-	\resizebox{0.9\linewidth}{!}{
+	\resizebox{0.8\linewidth}{!}{
 		\expandableinput ../tables/infocom-ecmp-avg-reward.tex
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/tnsm-udp-box-separate}
-
+	\includegraphics[width=0.8\linewidth]{../plots/tnsm-udp-box-separate}
+	\vspace{-1em}
 	\caption{
@@
 In a single-destination network, we observe that Marl's performance degrades as $n$ increases.
-Typically, our \emph{Instant} agent design achieves the best performance in multi-agent mode, reducing collateral damage above the current state-of-the-art, but sharply degrades at low $n$ when agents share experience.
+Typically, \emph{Instant} agents achieves the best performance in multi-agent mode, reducing collateral damage above the current state-of-the-art, but sharply degrade at low $n$ when agents share experience.
 We see a reversal of this trend for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
@@
 \Cref{fig:udp-tree-box} shows the preserved traffic in multi-agent mode.
-
 %?? Discuss multi-dest topol once all results available.
 When defending multiple destinations, we see a sharp decrease in the effectiveness of all agent designs.
-Our \emph{Instant} and \emph{Guarded} designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
-Paradoxically, \emph{Instant} actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between widely different actions, due to shared characteristics between legitimate and malicious flows.
-
-Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect in the case of UDP traffic.
+\emph{Instant} and \emph{Guarded} agents become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
+
+%Although we improve upon Marl in both identified problem cases, the improvements are small for UDP traffic.
 The most likely explanation for low UDP protection is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
@@
 These difficulties may be further worsened by the competitive nature of multi-agent learning.
+Paradoxically, \emph{Instant} actively suffers when trained as a single learner---this may occur due to even spreads of values between widely different actions, due to shared characteristics between legitimate and malicious flows.
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/tnsm-tcp-box-single}
-
+	\includegraphics[width=0.8\linewidth]{../plots/tnsm-tcp-box-single}
+	\vspace{-1em}
 	\caption{
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/tnsm-tcp-16-single}
-
+	\includegraphics[width=0.8\linewidth]{../plots/tnsm-tcp-16-single}
+	\vspace{-1em}
 	\caption{
@@
 
-\Cref{tab:av-vals} shows that Marl offers a low (though fairly consistent) level of protection for TCP traffic, which the \emph{Instant} agent offers no substantial improvement over.
-However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$\!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
+\Cref{tab:av-vals} shows that Marl offers a low level of protection for TCP traffic, which the \emph{Instant} agent offers little improvement over.
+However, \emph{Guarded} agents fare substantially better, particularly when experience can be shared---offering a \SI{2.21}{$\!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
 \Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement over Marl, \SI{8}{$\!\times$} more traffic than no defence system) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
 %SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
-
 %?? Discuss multi-dest topol once results available.
-We observe that the same trends present themselves in the multi-destination topology: \emph{Guarded} remains the best fit for TCP, in both training modes.
+The same trends are present in the multi-destination topology: \emph{Guarded} remains the best fit for TCP, in both training modes.
 Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
 
-\emph{Guarded}'s unpredictable (often worse) starting performance is unexpected, given its far smaller action space.
-We expected that this would make the model easier to learn, but the additional state required appears to make the task \emph{harder} initially.
-This design performs best (and exhibits considerably lower variance) when agents learn from as much knowledge as possible: high $n$ and single-agent training.
-To filter incoming traffic, it must degrade a flow multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
-We claim \emph{Guarded} is a considerably stronger model for these reasons, and its successes offer strong rationale to consider the best schemes for efficient information sharing.
+\emph{Guarded}'s unpredictable and worse starting performance is unexpected, given its smaller action space.
+We expected that this would make the model easier to learn, but the extra state required appears to make the task initially \emph{harder}.
+This design performs best (with considerably lower variance) when agents gain as much experience as possible: high $n$ and single-agent training.
+To filter traffic, it must degrade a flow multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
+We claim \emph{Guarded} is a stronger model for these reasons, and its successes offer strong rationale to consider the best schemes for efficient information sharing.
 
@@
 \Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of volume-based DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
-Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
+Many of their features aren't relevant to our work, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
 
 %?? Actually talk about Marl (???) to appease reviewer \#1.
-The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, whom we have positioned our work against.
+The closest approach in this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, whom we have positioned our work against.
 They create a tree overlay topology subdivided into teams, where each agent applies packet drop to \emph{all} inbound traffic.
@@
 Their technique suffers at high host density and when congestion-aware traffic dominates.
-These effects are not present in their results, suggesting an evaluation driven purely by traces.
+%These effects are not present in their results, suggesting an evaluation driven purely by traces.
 
 \emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} remedies transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
-Bots participating in an attack are often unable to match any bandwidth expansion in an AS due to having already saturated the capacity of their local outbound links.
+Bots participating in an attack cannot often match any bandwidth expansion due to having already saturated the capacity of their local outbound links.
 %Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
 Unlike our approach, SPIFFY is intended to be deployed within ASes.
-Its weakness is that computing a route with enough spare bandwidth can be costly, and that low expansion factors in real network can require more ``rounds'' of testing. 
-Traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), which make up a sizeable portion of network traffic.
-
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a generalised SDN framework for intrusion detection, using a \emph{k-nearest neighbours} classifier to detect individual attack flows.
+Its weakness is that computing a route with enough spare bandwidth can be costly, and that low expansion factors in real network can require more rounds of testing. 
+Traffic response to bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), which make up a sizeable portion of network traffic.
+
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a general SDN framework for intrusion detection, using a \emph{k-nearest neighbours} classifier to detect individual attack flows.
 Although heavyweight (and effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the evidence needed to understand how the systems compare.
 
-\Textcite{DBLP:conf/sp/SmithS18} use AS-level routing to tackle both transit-link and flooding-based attacks.
+\Textcite{DBLP:conf/sp/SmithS18} use AS-level routing to tackle transit-link and direct attacks.
 This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
@@
 \fakepara{RL in Networks}
-RL has seen comparatively little examination in the challenge of intrusion detection/prevention.
+RL has seen little examination in the challenge of intrusion detection/prevention.
 Past works use the approach as a traditional classifier \cite{shamshirband2014anomaly,DBLP:conf/mates/ServinK08}.
@@
 
-An effective use has been to learn when it is best to \emph{communicate} and share knowledge between anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
-%The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-%The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-Outside of this, there has been growing interest in the use of RL in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming.
+Effective uses have lain in learning when it is best to \emph{communicate} and share knowledge between anomaly detectors \cite{DBLP:conf/paisi/XuSH07}, and in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+\Textcite{DBLP:conf/sigcomm/MaoNA17} employ client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming.
 
 \emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation.
-Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
-To overcome the high latency of action computation via a neural network, two agents are trained, handling aspects of short (priority thresholds) and long flows (bespoke allocations) respectively.
+They find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
+To overcome the high latency of neural network computations, two agents are trained, handling aspects of short (priority thresholds) and long flows (bespoke allocations) respectively.
 

### 2019-07-29 tables/infocom-tree-avg-reward.py

@@ -64,7 +64,7 @@ def write_table(out_file, data):
 	for traffic in traffics:
 		for i, n in enumerate(ns):
 			key = (traffic,n)
-			out_file.write("{} & {} ".format("" if i>0 else traffic.upper(), n * 3))
+			out_file.write("{} & {} ".format("" if i>0 else traffic.upper(), n))
 
 			row = data[key]
 			for j, _ in enumerate(row):

### 2019-07-29 paper/infocom.tex
Algo compare, even more trimming.

@@
 	\centering
-	\resizebox{0.9\linewidth}{!}{\begin{tikzpicture}
+	\resizebox{0.8\linewidth}{!}{\begin{tikzpicture}
 		\node(remote){Remote};
@@
 
-	\resizebox{0.5\linewidth}{!}{
+	\resizebox{0.45\linewidth}{!}{
 		\begin{tabular}{@{}ll@{}}
@@
 We compare our work's performance and learning rate against Marl \cite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention, across different topologies and workloads.
-We made use of the OPUS and TCP traffic models introduced in \cref{sec:a-new-normal}, both topologies discussed below (1-dest vs Fat-Tree), and varied the amount of hosts communicating over each ingress/egress node---different networks will impose different levels of host density, e.g., popular web servers may have many more clients than egress points from their network.
-Additionally, we evaluated our new models in multi-agent (\emph{separate}, no model sharing) and i single-agent modes (\emph{single}, 0-cost perfect information sharing).
+We made use of the OPUS and TCP traffic models introduced in \cref{sec:a-new-normal}, both topologies discussed below (1-dest vs Fat-Tree), and varied the amount of hosts communicating over each egress node.
+Additionally, we evaluated these models in multi-agent (\emph{separate}, no model sharing) and single-agent modes (\emph{single}, 0-cost perfect information sharing).
 In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (resetting policies between episodes).
 Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes.
-Malicious traffic is generated by \emph{hping3} program, generating UDP-flood traffic targeting random ports with MTU-sized packets, mapping to the large frame sizes expected in an amplification attack.
-
+Malicious traffic is provided by \emph{hping3}, generating UDP-flood traffic targeting random ports with MTU-sized packets.
 All experiments were executed on Ubuntu 18.04.2 LTS (GNU/Linux 4.4.3-040403-generic x86\_64), using a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
@@
 Benign hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second}, and hosts were redrawn each episode with $\operatorname{P}(\mathit{malicious})=0.4$.
-Malicious hosts sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, increased to \SIrange{4}{7}{\mega\bit\per\second} when using TCP-like or mixed traffic to meaningfully impact benign flows.
+Malicious hosts sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, increased to \SIrange{4}{7}{\mega\bit\per\second} for TCP-like or mixed traffic to meaningfully impact benign flows.
 Given $n$ and $\operatorname{P}(\mathit{malicious})$, we see an expected malicious bandwidth \numrange{1.27}{1.87} and \SIrange{2.03}{2.18}{$\! \times U_s$} respectively.
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{../plots/tnsm-udp-box-separate}
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-udp-box-separate}
 	\vspace{-1em}
@@
 %		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
-		\emph{Instant} outperforms Marl for $n \in \{4, 8\}$ (with higher variance), but performs similarly to Marl at $n\in \{2, 16\}$.
-		Here, \emph{Guarded} always underperforms compared to the other agent designs.
+%		\emph{Instant} outperforms Marl for $n \in \{4, 8\}$ (with higher variance), but performs similarly to Marl at $n\in \{2, 16\}$.
+		\emph{Guarded} always underperforms.
 %		SPF, remarkably, slightly outperforms Marl++ (with lower variance) for this more difficult topology despite being worse for smaller $n$.
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{../plots/tnsm-tcp-box-single}
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-box-single}
 	\vspace{-1em}
@@
 %		\emph{Instant} and \emph{Guarded} exhibit essentially identical efficacy at $n~=~2$, protecting less traffic than Marl.
-		Only \emph{Guarded}'s performance rapidly increases with $n$, achieving considerably better median performance and lower variance than the other models.
-		Longer tails of outliers typically indicate longer training times---we observe that \emph{Guarded} has considerably lower variance once it has converged on a stable policy.
+%		Longer tails of outliers typically indicate longer training.
+%		\emph{Guarded}'s performance rapidly increases with $n$, achieving considerably better median performance and lower variance than the other models.
+%		Longer tails of outliers typically indicate longer training times---we observe that \emph{Guarded} has considerably lower variance once it has converged on a stable policy.
+		\label{fig:tcp-tree-box}
+	}
 \vspace{-1em}
-		\label{fig:tcp-tree-box}
-	}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{../plots/tnsm-tcp-16-single}
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-16-single}
 	\vspace{-1em}
 	\caption{
-		Online performance of standard and single-agent models over time in a single-destination network with $n=16$ hosts per egress point and HTTP benign traffic.
+		Online performance of standard and single-agent models over time (TCP, $n=16$, tree).
 %		At this level of host density, \emph{Guarded} reaches higher peak performance sooner and is considerably more consistent throughout the episode.
@@
 %		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
+		\label{fig:tcp-tree-16}
+	}
 \vspace{-1em}
-		\label{fig:tcp-tree-16}
-	}
 \end{figure}
@@
 \subsection{RL algorithms}
-When choosing a learning algorithm, we compared semi-gradient Sarsa against Q-learning.
-Preliminary experiments found that Sarsa offered the best performance and behaviour.
-?? Link plot or table w Sarsa vs. Q results.
+When choosing a learning algorithm, we compared semi-gradient Sarsa against Q-learning, and variants of these algorithms based on \emph{eligibility traces} (Sarsa($\lambda$), Watkins's Q($\lambda$)).
+Our experiments found that Sarsa offered the best performance and behaviour; \cref{fig:algo-spf-tcp} shows a trend observed across model choices and communication except for multiagent \emph{Instant} protection of TCP where Q-learning succeeds.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-algo-spf-tcp}
+	\vspace{-1em}
+	\caption{
+		Comparison of RL algorithms for \emph{Guarded} (TCP, $n=16$, tree).
+		%		At this level of host density, \emph{Guarded} reaches higher peak performance sooner and is considerably more consistent throughout the episode.
+		%		\emph{Guarded} benefits greatly from information sharing, converging to protect around \SI{75}{\percent} of TCP traffic within \SI{100}{\second}.
+		%		The \emph{Instant} model converges to Marl's level of performance.
+		%		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
+		\label{fig:algo-spf-tcp}
+	}
+\vspace{-1em}
+\end{figure}
 
@@
 \emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a general SDN framework for intrusion detection, using a \emph{k-nearest neighbours} classifier to detect individual attack flows.
-Although heavyweight (and effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the evidence needed to understand how the systems compare.
+Although heavyweight (and effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the evidence needed to draw conclusions.
 

### 2019-07-29 paper/infocom.tex
Plot fun times...
Policy description, typos...

@@
 \newcommand{\wvec}[1]{\ensuremath{\bm{w}_{#1}}}
+\DeclareMathOperator*{\argmax}{argmax}
 
@@
 It is immediately clear that these challenges also apply to DDoS prevention.
-Learning is made harder still by the challenges encountered when learning from unlabelled/partial data.
+Learning is made harder still by the challenges encountered when learning from unlabelled or partial data.
 %DDoS prevention exhibits many of these issues.
@@
 %	\item An investigation of suitable features for this task (\cref{sec:rethinking-the-state-space}).
-%	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
+	\item Reactive simulations of HTTP and VoIP web-server traffic (\cref{sec:a-new-normal}).
 	\item A performance comparison of different RL algorithms within these agents, and against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
@@
 	\centering
-	\begin{subfigure}{0.45\linewidth}
+	\begin{subfigure}{0.4\linewidth}
 		\resizebox{\linewidth}{!}{
@@
 	\end{subfigure}%
-	\begin{subfigure}{0.45\linewidth}
+	\begin{subfigure}{0.4\linewidth}
 		\resizebox{\linewidth}{!}{
@@
 To encourage early exploration, actions with maximal value can be chosen each time except when taking random actions with probability $\epsilon$---an \emph{$\epsilon$-greedy policy}).
+Each action $a \in \operatorname{A}(s)$ has probability $\pi(a|s,\wvec{})$ of being chosen:
+\begin{equation}
+	\pi(a|s,\wvec{}) =
+	\begin{cases}
+	1 - \epsilon + \epsilon/{|}{\operatorname{A}(s)}{|} & \text{if $a = \operatorname*{argmax}\limits_{a'} \acval{s}{a'}{\wvec{}}$} \\
+	\epsilon/{|}{\operatorname{A}(s)}{|} & \text{otherwise}
+	\end{cases}
+\end{equation}
 \Cref{fig:tilecode-select} shows how the value of each action is computed (and which action would be chosen by a greedy policy).
@@
 \subsection{Algorithm}
-To make decisions and learn cheaply with low latency, we use \emph{semi-gradient Sarsa with tile coding} \ref{eqn:sg-sarsa} (\cref{sec:reinforcement-learning}).
+To make decisions and learn cheaply with low latency, we use \emph{semi-gradient Sarsa with tile coding} (\cref{eqn:sg-sarsa}).
 Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
@@
 \subsubsection{Decision narrowings}
-When learning a high-dimensional tile-coding, assignment of credit for each decision is difficult because all tiles have identical gradient--i.e., assigning credit to the \emph{features} which drove each decision is difficult.
+When learning a high-dimensional tile-coding, assignment of credit for each decision is difficult because all tiles have identical gradient---i.e., assigning credit to the \emph{features} which drove each decision is difficult.
 To combat this, with probability $\epsilon$ an agent will mark a flow as being governed by a subset of the state space for the next five decisions.
@@
 The plots demonstrate that different protocols and traffic classes are best defended by different features---every feature presented has value in a complete model.
-Packet-level and per-window statistics are at their based when combined with the last action for both UDP-like and TCP-like traffic (omitted).
+Packet-level and per-window statistics are at their best when combined with the last action for both UDP-like and TCP-like traffic (omitted).
 All features converge to their highest-observed performance within around \num{4000} timesteps.
@@
 End-host servers may make use of canary flows or active measurement schemes, and can employ quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
-ASes and transit networks may make use of reports received from downstream, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
+ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
 We leave the use of such data as future work.
@@
 %?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
-We choose to drop packets ()rather than impose traffic limits) to have a discrete action space without prior knowledge of traffic characteristics.
+We choose to drop packets (rather than impose traffic limits) to have a discrete action space without prior knowledge of traffic characteristics.
 Furthermore, we need not consider burstiness, fairness and parameters such as bucket sizes which could require tuning.
@@
 To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$), chosen due to TCP's higher prevalence.
-An agent must now choose to punish a flow multiple times in succession to remove it, reducing variance while allowing an agent to see how a host reacts to controlled conditions.
+An agent must now choose to punish a flow multiple times in succession to block it, reducing variance while allowing an agent to see how a host reacts to controlled conditions.
 While beneficial for legitimate flows who behave similarly to attack traffic, this introduces a lag of an extra two actions in blocking illegitimate flows.
@@
 %Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
-QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
+QUIC now comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
 Furthermore, applying a single rule to many flows (as in \textcite{DBLP:journals/eaai/MalialisK15}) increases the likelihood of collateral damage.
@@
 Moreover, networks of this kind more naturally enable the future relocation and easy installation of learners.
-Agents communicate with these co-hosted switches to install probabilistic packet-drop rules.
+Agent VFs communicate with these co-hosted switches to install probabilistic packet-drop rules.
 We describe the main purpose and operation of each module within an agent's VF, and discuss techniques to make deployment more efficient using existing technologies.
@@
 The RL algorithm then returns an action.
-Each action is passed to the database, which computes and returns a packet drop rate according to the agent model (instant/guarded) while updating flow state.
+Each action is passed to the database, which computes and returns a packet drop rate according to the agent model (\emph{Instant/Guarded}) while updating flow state.
 This is then converted into an OpenFlow message carrying packet drop rules; these are batched to the agent's switch using the same groupings produced by the scheduler.
@@
 Gathering and transmission of load/flow statistics would be difficult to perform quite as often as an emulated environment allows.
-However, the measurements acquired in such a scenario are far more likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
+However, the measurements acquired in such a scenario are likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
 
@@
 
-Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations of short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}--- an agent is typically able to process around 3 flows in this time.
+Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations on handling short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---an agent is typically able to process around 3 flows in this time.
 We expect deadlines should be tuned based on the frequency at which statistics arrive.
@@
 To get around the lack of floating-point support in many environments, we represent this probability using a 32-bit unsigned integer (scaling \num{1.0} to $2^{32}-1$).
-On commodity hardware, we believe that a similar effect can be achieved using OpenFlow meters (at the expense of being stateful measures).
-
-We use OpenFlow groups to simplify control messages: premade tables with permitted levels of packetdrop.
+On commodity hardware, we believe that a similar effect can be achieved using OpenFlow meters (at the expense of these being stateful measures).
+
+We use OpenFlow groups to simplify control messages: premade tables with permitted levels of packet drop.
 This saves some overhead compared to using experimenter/extension headers.
@@
 This presents a balanced distribution of flow duration and size, with large files included to model bulk transfer.
-On completion, a host changes its IP to generate separate statistics per-flow.
+On completion, a host changes its IP to generate separate statistics per flow.
 
@@
 In a single-destination network, we observe that Marl's performance degrades as $n$ increases.
-Typically, \emph{Instant} agents achieves the best performance in multi-agent mode, reducing collateral damage above the current state-of-the-art, but sharply degrade at low $n$ when agents share experience.
+Typically, \emph{Instant} agents achieve the best performance in multi-agent mode, reducing collateral damage above the current state-of-the-art, but sharply degrade at low $n$ when agents share experience.
 We see a reversal of this trend for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
@@
 	\centering
-	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-16-single}
+	\includegraphics[width=0.75\linewidth]{../plots/infocom-tcp-16-single}
 	\vspace{-1em}
@@
 Unlike our approach, SPIFFY is intended to be deployed within ASes.
-Its weakness is that computing a route with enough spare bandwidth can be costly, and that low expansion factors in real network can require more rounds of testing. 
+Its weakness is that computing a route with enough spare bandwidth can be costly, and that low expansion factors in real networks can require more rounds of testing. 
 Traffic response to bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), which make up a sizeable portion of network traffic.

### 2019-07-31 paper/infocom.tex
Just the conclusion + new results to be handled...

@@
 %	pdfauthor={Kyle A. Simpson, Simon Rogers, Dimitrios P. Pezaros}
-	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
+	pdfauthor={Anonymous Giraffe, Anonymous Ibex, Anonymous Badger}
 }
@@
 Anonymous Giraffe,
-Anonymous Badger,
-Anonymous Ibex\\
+Anonymous Ibex,
+Anonymous Badger\\
 Unnamed Department, Nowhere\\
@@
 Unsupervised, online learning is a key part of overcoming this detection problem.
-We present the architecture of a system designed to host such RL agents in a typical network.
-Furthermore, we introduce two agents which advance the state-of-the-art in using RL for DDoS mitigation.
+Transfer of training data and execution can be expensive, hampering deployment.
+In this paper, we show how to deploy such RL agents in a typical network at low cost through a novel architecture.
+Our two agent designs make better use of flow information to advance the state-of-the-art in using RL for DDoS mitigation.
 Both of these act per flow, in a protocol-agnostic manner for any network topology.
@@
 
-Our results show the existence of flow features with high predictive power for bulk TCP and VoIP traffic.
-We show that the new agents offer a significant increase in throughput of legitimate TCP traffic for many choices of host density.
+Our results show that flow features effective on bulk TCP and VoIP traffic, combined with our agent designs, protect \SI{8}{$\!\times$} more TCP traffic than no defence at high host density.
+This is a \SI{2.5}{$\!\times$} improvement over RL's state-of-the-art in DDoS defence.
+
+%(\SI{2.5}{$\!\times$} improvement over Marl, \SI{8}{$\!\times$} more traffic than no defence system)
 \end{abstract}
@@
 
-However, the internet is always in flux.
+However, the Internet is always in flux.
 The constant deployment of new services and protocols means that traffic is \emph{non-stationary} and evolving.
@@
 
-%Network anomaly detection and intrusion detection/prevention are continually evolving problems, compounded by the partial, non-\emph{independent and identically distributed} (IID) view of data at each point in the network.
-%Attacks and anomalous behaviours evolve, becoming more sophisticated or employing new vectors to harm a network or system's confidentiality, integrity, and availability without being detected \cite{DBLP:journals/comsur/BhuyanBK14}.
-%These attacks and anomalies have measurable consequences and symptoms which allow a skilled analyst to infer new signatures for detection by misuse-based classifiers, but unseen attacks may only be defended against after-the-fact.
-%This issue is inherent to \emph{misuse-} or \emph{signature-based} intrusion detectors, and it has been long-hoped that \emph{anomaly-based} detectors would surpass this by making effective use of statistical measures \cite{DBLP:journals/comsur/BhuyanBK14}.
-
-%?? Then talk about ML. And its costs, say that its ADS/IDS weaknesses extend to DDoS prevention.
-While \emph{machine learning} (ML) approaches seem suitable for this problem, there is a distinct lack of real-world deployments in the broader field of intrusion detection/prevention.
-\Textcite{DBLP:conf/sp/SommerP10} posit that this stems from greater risks than other ML tasks: the high cost of errors, low tolerance for false positives; a lack of recent, high-quality training data; and traffic diversity and evolution.
-It is immediately clear that these challenges also apply to DDoS prevention.
-Learning is made harder still by the challenges encountered when learning from unlabelled or partial data.
-%DDoS prevention exhibits many of these issues.
-%?? Then Bam! Introduce RL.
-
+While \emph{machine learning} (ML) approaches seem suitable for this problem, their deployment has been sparse in the broad field of intrusion prevention due to key risks: the high cost of errors; a low tolerance for false positives; and traffic diversity and evolution \cite{DBLP:conf/sp/SommerP10}.
+These challenges also apply to DDoS prevention, made harder when learning from unlabelled or partial data.
 \emph{Reinforcement learning} (RL) offers another perspective.
-%?? Unclear explanation of RL here?
-RL agents follow a \emph{policy} to interact with or control a system, while at the same time using observed performance to improve this policy online.
-An agent then adaptively reacts to threats by acting as a feedback loop for network optimisation---this allows us to overcome some of these difficulties by monitoring the consequences of each action in real-time.
-We aim to prevent volume-based DDoS attacks with the aid of RL-based techniques, while bringing to light the flexibility of these techniques in the security domain.
-%Whether it takes direct control of the network, or is used indirectly to optimise a key part of another system, more powerful `deep' RL techniques (and well-founded action spaces) aren't yet well explored for network IDS/IPS.
-%These range from more modern training algorithms \cite{DBLP:journals/corr/SchulmanWDRK17, DBLP:conf/icml/SchulmanLAJM15}, to evolutionary strategies \cite{DBLP:journals/corr/SalimansHCS17, DBLP:journals/corr/abs-1802-08842}, hierarchical action composition \cite{DBLP:journals/corr/abs-1710-09767}, and competitive multi-agent learning \cite{DBLP:journals/corr/abs-1710-03748}.
-%?? This raises the question: how do we do this effectively? Agents need state and reward info, which can be a lot of data transfer. THis gets worse (hint: we have the answer...)
+RL agents follow a \emph{policy} to control a system, using observed performance to improve this policy \emph{online}.
+An agent then adapts to threats by acting as a feedback loop for network optimisation---allowing us to overcome some of these difficulties by monitoring the consequences of each action in real-time.
+We aim to prevent volume-based DDoS attacks with the aid of RL techniques, while bringing to light their flexibility in the security domain.
 This introduces a new set of problems: what effective state/action spaces should look like, how state and reward information should be efficiently propagated, and how this can be achieved on commodity hardware.
@@
 %?? Discuss the most important conclusions before the outline.
-We find that the existing RL work for this task fails to account for congestion-aware traffic (i.e., TCP, QUIC) and environments with high host density at ingress points.
+We find that the existing RL work for this task \cite{DBLP:journals/eaai/MalialisK15} fails to account for congestion-aware traffic (i.e., TCP, QUIC) and environments with high host density at ingress points.
 To remedy this, our agents make throttling decisions per-source, and show how an action model built on domain-specific knowledge can increase performance.
-This requires additional engineering: updating RL agents from multiple traces per timestep, smart work scheduling, and an architecture featuring (but minimising) the use of software-defined networks (SDN).
+This requires additional engineering: updating RL agents from multiple traces per timestep, smart work scheduling, and an architecture featuring the use of Software-defined Networks (SDN).
+Our design minimises the use of (and thus attack surface introduced by) new technologies like SDNs, to increase compatibility with existing infrastructure and simplify operation.
+These allow us to demonstrate how much value each feature offers when detecting attack flows.
 %In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
@@
 
+\subsection{Motivation}\label{sec:motivation}
+%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
+%??
+Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows.
+This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e.
+an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
+There are concrete reasons to use RL methods in this domain, beyond handling non-stationarity.
+Misclassification is a serious problem, introducing \emph{collateral damage} in a DDoS prevention system.
+RL's feedback-loop model lets us monitor flows \emph{after} an action is taken to perhaps allow forgiveness of mistakenly punished flows.
+This relies upon a flow-by-flow view of the state space, but knowledge of the current state and last applied action detect whether punished flows fall off identically to legitimate flows.
+
+%Which features might be best suited to this problem?
+%??
+Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
+Other studies suggest that there are particularly useful features which make the task of online DDoS flow identification feasible.
+Link loads observed at various locations suggest the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
+Volume-based statistics (counts, counts per duration, average packet sizes) have seen use in $k$-nearest neighbours classifiers trained to detect DDoS attacks \cite{DBLP:conf/dsn/LeeKSPY17}.
+Most importantly, there is evidence showing that bandwidth expansion may cause behavioural changes which differ between benign and attack traffic \cite{DBLP:conf/ndss/KangGS16}.
+
 \subsection{Distributed Denial of Service}
@@
 
-\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, to inflict financial harm or as an act of vandalism.
-Attackers achieve this by exploiting system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
-Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
-
-Volume-based attacks are the most relevant to our work, with two varieties being especially prevalent.
-\emph{Amplification attacks} exploit services who send large replies in response to small requests, with UDP-based services like DNS and NTP seeing most abuse \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
-Malicious hosts send many small requests, spoofed using the victim's address, causing many large replies to be sent to the intended target---increasing a botnet's throughput while keeping participants anonymous.
-\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
+\emph{Distributed denial of service} (DDoS) attacks may be broken down further---exploiting system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
+Hosts often participate unwillingly, having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.
+Volume-based attacks are most relevant to our work, with two varieties being especially prevalent.
+\emph{Amplification attacks} exploit services who send large responses to small requests, with UDP-based services like DNS and NTP seeing most abuse \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
+Malicious hosts send many small requests, spoofed using the victim's address, causing the intended target to receive many large replies---increasing a botnet's throughput while keeping participants anonymous.
+\emph{Link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.
 
@@
 	\end{cases}
+	\label{eqn:egreedy-pi}
 \end{equation}
@@
 
-\subsection{Motivation}\label{sec:motivation}
-%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
-%??
-Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows.
-This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e.
-an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
-There are concrete reasons to use RL methods in this domain, beyond handling non-stationarity.
-Misclassification is a serious problem, introducing \emph{collateral damage} in a DDoS prevention system.
-RL's feedback-loop model lets us monitor flows \emph{after} an action is taken to perhaps allow forgiveness of mistakenly punished flows.
-This relies upon a flow-by-flow view of the state space, but knowledge of the current state and last applied action detect whether punished flows fall off identically to legitimate flows.
-
-%Which features might be best suited to this problem?
-%??
-Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
-Other studies suggest that there are particularly useful features which make the task of online DDoS flow identification feasible.
-Link loads observed at various locations suggest the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
-Volume-based statistics (counts, counts per duration, average packet sizes) have seen use in $k$-nearest neighbours classifiers trained to detect DDoS attacks \cite{DBLP:conf/dsn/LeeKSPY17}.
-Most importantly, there is evidence showing that bandwidth expansion may cause behavioural changes which differ between benign and attack traffic \cite{DBLP:conf/ndss/KangGS16}.
-
 \section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
-We hypothesise that per-flow agent designs will offer better accuracy than current RL-based agents for DDoS mitigation.
-We describe and justify our approach and algorithmic improvements, presenting two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
+We hypothesise that per-flow agent designs offer better accuracy than current RL-based agents for DDoS mitigation, because existing models apply one action over several flows.
+Congestion-aware traffic is both more common and more vulnerable to this.
+We describe and justify our approach and algorithmic improvements, presenting two different action models.
 
@@
 To make decisions and learn cheaply with low latency, we use \emph{semi-gradient Sarsa with tile coding} (\cref{eqn:sg-sarsa}).
-Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
-Each agent has its own internal parameter vector $\wvec{}$.
+Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time (e.g., \num{10000} timesteps).
+Each agent has its own internal parameter vector $\wvec{}$---giving its policy $\pi$.
 
@@
 We update $\wvec{}$ using each decision made and the reward signal from the relevant destination.
-As exploration still occurs for each action, $\epsilon$ must be reduced multiple notches every timestep.
-In turn, we increase the annealing window for $\epsilon$ by a factor of \num{2.67} so as to preserve exploration over time, by accounting for the greater volume of decisions being made.
+As exploration still occurs for each action according to \cref{eqn:egreedy-pi}, and we take \numrange{2}{3} actions per timestep, $\epsilon$ must be reduced multiple notches every timestep.
+We increase the annealing window for $\epsilon$ by a factor of \num{2.67} to preserve exploration over time, accounting for the greater volume of decisions being made.
 
 \subsubsection{Per-tile updates}
-To adjust the value of each activated tile with a different magnitude and direction, we compute a different temporal difference value \emph{for each tiling}.
-This corrects tiles which contribute too much or too little value without disturbing estimates which are almost correct.
+To adjust the value of activated tiles with different magnitudes and directions, we compute a different temporal difference value $\delta_t$ \emph{for each tiling}.
+Ordinarily, all tiles would change identically.
+This corrects tiles which contribute too much or little value without disturbing estimates which are almost correct.
 While we make use of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to increase $\alpha$ without divergence.
@@
 \end{figure}
-
-Our state space combines global state (network load) with per-flow measurements.
-Each is tile-coded with 8 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
-Features are tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
-
-%?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
-Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}): the bandwidth measurements regularly received from monitors in the environment.
-For any flow, an agent computes the path it would take through the network.
-The incoming load along the first hop, last hop, and tertiles of the path are then tile-coded together.
-In the event that this path is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
-\Cref{fig:global-state-path} illustrates the process.
-Global state built in this way is compatible with multipath, multi-destination networks, offering support for diverse deployment environments from endpoint servers up to transit ASes.
-Computing a flow's path is not computationally expensive: multipath routing is fast since typical ECMP routing algorithms simply hash a packet's 5-tuple and are deterministic to provide consistent quality-of-service.
-
-%?? Path computation fast and efficient due to deterministic routing based on hashes (ECMP)
-%?? Works for any arbitrary topology, even
-
-\Cref{fig:udp-feature-plots} shows the effectiveness of per-flow features for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), on a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
-The plots demonstrate that different protocols and traffic classes are best defended by different features---every feature presented has value in a complete model.
-Packet-level and per-window statistics are at their best when combined with the last action for both UDP-like and TCP-like traffic (omitted).
-All features converge to their highest-observed performance within around \num{4000} timesteps.
-In general, some of the most effective features are global state, mean IAT, mean inbound packet size and $\Delta$ rates.
-The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
-If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have no correlation with the class of traffic a flow contains.
-
-\subsection{Reward Function}
-
-%?? Say ``we take $R_G$ from M and K''.
-%
-%?? Reward at each destination.
-
-\newcommand{\arrload}[2]{\operatorname{load}^{#2}_{t}(#1)}
-\newcommand{\uload}[1]{\arrload{#1}{\uparrow}}
-\newcommand{\dload}[1]{\arrload{#1}{\downarrow}}
-\newcommand{\bload}[1]{\arrload{#1}{\updownarrow}}
-\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
-%\fakepara{Reward function directionality}
-%The reward functions, as defined, do not take traffic direction into account.
-%We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
-We denote the upstream, downstream and combined loads at a node $s$ as $\uload{s}, \dload{s}, \bload{s}$ respectively.
-Each destination node $s$ generates a reward signal, $R_{s,t}$, at every timestep $t$.
-Assuming each destination has access to some classification function $g(\cdot)$ which estimates the volume of legitimate traffic received, and expects to receive $\mathit{traffic}_s$, we may compute the reward:
-\begin{subequations}
-%	\vspace{-0.5cm}
-	\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
-	\begin{gather}
-	\cond{s} = [\max(\uload{s}, \dload{s}) > U_s],\\
-	R_{s,t} = (1 - \cond{s}) \frac{g(\arrload{s}{-})}{\mathit{traffic}_s} - \cond{s},\label{eqn:reward-rtx}
-	\end{gather}
-	replacing $\arrload{s}{-}$ in \cref{eqn:reward} with whichever directional load is prioritised according to the traffic characteristics of the deployment environment, where $\cond{s}$ represents the ``overloaded'' condition at destination $s$.\label{eqn:reward}
-\end{subequations}
-We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, and suggest $\bload{\cdot}$ for general deployment or heterogeneous traffic.
-These choices reflect where the bulk of transmitted bytes in each traffic model are observed (or the lack of this knowledge).
-
-While our use and definition of $g(\cdot)$ appears nebulous, there are many ways to infer this quantity in practice.
-End-host servers may make use of canary flows or active measurement schemes, and can employ quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
-ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
-We leave the use of such data as future work.
-Even without online rewards, a well-trained agent needs only to follow the greedy policy it has learned from offline (simulated) training.
-
-If a network is believed to be vulnerable to indirect attacks, such as link-flooding attacks, we may use the following reward:
-\begin{equation}
-	R_{s,t}^{\mathit{Cross}}(\beta) = \beta R_{s,t} + (1 - \beta) \min{\{R_{s',t} | s' \ne s\}} \label{eqn:lfa-reward}
-\end{equation}
-where the collaboration parameter $\beta \in [0,1]$ models the expected degree of interference between flows, and $s, s'$ are protected destination nodes in the network.
-The key insight underpinning LFAs is that flows can affect a target \emph{without communicating with that target}.
-$\beta$ then acts as a tunable parameter which can incentivise agents to remove flows which harm overall system health, by including the performance of the worst-performing destination.
-However, we do not examine such attacks (and the effectiveness of $R_{s,t}^{\mathit{Cross}}$) at present.
-
-\subsection{Action Space}
-When monitoring a flow, an agent uses its state vector to decide which portion of its \emph{inbound} traffic should be dropped.
-%?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
-We choose to drop packets (rather than impose traffic limits) to have a discrete action space without prior knowledge of traffic characteristics.
-Furthermore, we need not consider burstiness, fairness and parameters such as bucket sizes which could require tuning.
-We offer two models, which differ only on how to choose $p$.
-The intuition behind these designs is that an agent may directly choose the level of packet drop at the expense of a larger action space (\emph{Instant control}), or that we may reduce the action set (and learn a policy within fewer actions) by making use of domain knowledge (\emph{Guarded control}).
-
-\subsubsection{Instant control}
-Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$.
-These choices ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
-Since this model requires no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely opportunistic/myopic).
-
-\subsubsection{Guarded control}
-\Textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
-We apply their observations, constraining how an agent treats each flow via simple finite state automata: restricting $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
-The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
-
-We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limits (\SI{25}{\percent} increments for UDP traffic).
-The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
-This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.
-As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
-We found the setting $\gamma = 0.8$ to be the most effective choice during exploratory testing.
-
-To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$), chosen due to TCP's higher prevalence.
-An agent must now choose to punish a flow multiple times in succession to block it, reducing variance while allowing an agent to see how a host reacts to controlled conditions.
-While beneficial for legitimate flows who behave similarly to attack traffic, this introduces a lag of an extra two actions in blocking illegitimate flows.
-
-\subsubsection{Risks}
-Our mode of action means that each agent carries a risk of introducing collateral damage into the network.
-This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p\ne0$ in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
-This weakness remains in modern flavours of TCP---TCP Cubic has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
-Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets and \SIrange{77}{84}{\percent} of data volume\footnote{Repository link removed for double-blind reviewing.}.
-%Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
-QUIC now comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
-Furthermore, applying a single rule to many flows (as in \textcite{DBLP:journals/eaai/MalialisK15}) increases the likelihood of collateral damage.
-When considering that many hosts have an especially adverse reaction to our main means of control, flow-level granularity becomes an obvious choice.
-
-\section{System Architecture}\label{sec:system-architecture}
-
-%?? Here is where I would put my ramblings about system architecture...
-We present our design of a system which supports the effective real-world deployment of RL-based DDoS mitigation.
-\Cref{fig:sys-arch} displays this, separating system elements which are local to each agent from those which reside elsewhere in the network.
-We operate each agent as a \emph{virtual function} (VF) adjacent to a software-defined switch.
-Software-defined networks allow a controller (or authorised hosts) to install actions, forwarding rules and logic into a switch at runtime.
-Moreover, networks of this kind more naturally enable the future relocation and easy installation of learners.
-Agent VFs communicate with these co-hosted switches to install probabilistic packet-drop rules.
-We describe the main purpose and operation of each module within an agent's VF, and discuss techniques to make deployment more efficient using existing technologies.
-
-\subsection{Core and RL Executor}\label{sec:core-and-rl-executor}
-The core module is the main loop in an agent's architecture.
-At each timestep, the core receives information about which flows have arrived and should be acted upon from the \emph{TRS Scheduler}.
-The core then retrieves the current and previous state vector associated with each flow from the \emph{Flowstate Database}, passing them into the RL algorithm alongside the last action chosen for that flow (if available).
-
-The RL algorithm then returns an action.
-Each action is passed to the database, which computes and returns a packet drop rate according to the agent model (\emph{Instant/Guarded}) while updating flow state.
-This is then converted into an OpenFlow message carrying packet drop rules; these are batched to the agent's switch using the same groupings produced by the scheduler.
-Finally, timing information is passed back into the scheduler to refine its estimates about how much work should be scheduled in the next timestep.
-
-State space sizes guarantee that an \emph{Instant} policy remains under \SI{520}{\kibi\byte}, though our sparse representation typically leads to far smaller policies: $\sim$\SI{17.8}{\kibi\byte} from our experiments.
-\emph{Guarded} policies are \SI{30}{\percent} of this size.
-As described earlier, action updates require a constant number of floating point operations---\num{160} floating point additions and \num{32} multiplications per update of $\wvec{}$ with per-tile updates, with \num{160} additions required to choose an action.
-The vast majority of these operations can be vectorised.
-Action computation for \emph{Guarded} agents is cheaper still, requiring only \num{48} additions per action.
-
-\subsection{Stats API and Collectors}
-Agents require information from the network and one another to be effective.
-%Each agent then exposes an API through which statistics and policy information may arrive.
-%?? API between agent and switches/other agents? Push/Pull API on stats? How do they share experience? Could broadcasts aid both?
-%?? What does experience transfer look like? Why? Explain that every agent has a different policy, and why this is beneficial
-Our agents can act either independently, having no agent-to-agent communication, or cooperatively.
-In the latter case agents transfer, when possible, \emph{experience} to one another---lists of state-action-state transitions with associated rewards.
-It's noted that a transition may be high-value or surprising to one agent, while well-known to another, causing each to produce different policy updates from the same unit of experience.
-For this reason we do not transfer policy deltas between agents, causing each to learn its own policy.
-Which scheme achieves better performance is left to future work.
-
-%?? How do stats get from switcehs to the agent?
-Load collectors and estimators periodically push observations to each active agent VF.
-In our current implementation, load statistics are gathered via VFs active at each network switch, though we expect that OpenFlow stats requests, NetFlow or SNMP data may be used to derive these cheaply.
-Transferring state to agents and experience sharing can both be made more efficient through effective use of broadcast addressing in a target network.
-Depending on the capabilities of switches in the network, the estimator can either send benign traffic estimates or parameters for use in a reward function.
-
-Gathering and transmission of load/flow statistics would be difficult to perform quite as often as an emulated environment allows.
-However, the measurements acquired in such a scenario are likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
-
-\subsection{Flowstate Database}
-For each flow 5-tuple, we hold two state vectors containing the features described in \cref{sec:feature-space}---the current state, and the state which induced the last action.
-To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set.
-State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
-For flows outside of the current work list, we simply replace the stored vector.
-
-\subsection{TRS Scheduler}
-Acting on an unbounded set of flows in each timestep introduces potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and the risk that flow states become outdated.
-At their worst, these risks present additional attack surface to an adversary.
-To tackle these problems, we make use of \emph{Timed Random Sequential} updates.
-
-The scheduler begins with a shuffled work list of active flows.
-When requested, the scheduler estimates the cost of an action computation using the timing information received from the core, proceeding down the list to send a set of 5-tuples to the core which can be handled in a set time limit.
-The scheduler continues until the list is empty, at which point it is repopulated and reshuffled with active flows.
-
-Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations on handling short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---an agent is typically able to process around 3 flows in this time.
-We expect deadlines should be tuned based on the frequency at which statistics arrive.
-The amount of processed flows per deadline depends on the agent design (FLOP count, policy size), but also on the amount of flow telemetry data needing processed---our current implementation is written in python, restricting this handling to a single thread.
-An implementation in a systems language such as Rust or C++ would allow faster concurrent processing.
-
-There is a risk that so much work can be queued up that an agent is never able to act on some attack flows.
-A solution is to impose an upper bound on the amount of action computations/policy updates that can be performed before the work list is regenerated.
-This removes the guarantee that all flows will be visited often, but if updates occur regularly then this sampling may be sufficient to achieve good performance.
-
-\subsection{Agent Switches}
-%?? How do we do stat collection?
-Our agent switches operate a modified version of Open vSwitch, implementing an action which requests that each matched packet be dropped with a certain probability.
-To get around the lack of floating-point support in many environments, we represent this probability using a 32-bit unsigned integer (scaling \num{1.0} to $2^{32}-1$).
-On commodity hardware, we believe that a similar effect can be achieved using OpenFlow meters (at the expense of these being stateful measures).
-
-We use OpenFlow groups to simplify control messages: premade tables with permitted levels of packet drop.
-This saves some overhead compared to using experimenter/extension headers.
-Flows are automatically given a group with the default level of packet drop (according to the chosen agent design), meaning that switches don't need to refer to a controller or the agent VF.
-
-%?? What do our modifications to openflow look like? How do we get around the lack of floating point math?
-
-\begin{figure}
-	\centering
-	\resizebox{0.8\linewidth}{!}{\begin{tikzpicture}
-		\node(remote){Remote};
-
-		%%%
-
-		\node[below=0.05 of remote](swpos){};
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
-		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
-		\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
-		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
-
-		%
-
-		\node[right=2 of swpos](epos){};
-		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
-		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
-		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
-		\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.85,0)$) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\small \centering Estimator\\$g(\cdot)$\end{varwidth}};
-
-		%
-
-		\node[right=3.5 of epos](apos){};
-		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
-		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
-		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
-		\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent VF};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
-
-		%
-
-		\node[below=2.3 of remote.west](linestart){};
-		\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
-		\draw [dashed] (linestart) -- (lineend);
-
-		%%%
-
-		\node[below=2.4 of remote](local){Local};
-
-		%%%
-
-		\node[below=0.05 of local](aswpos){};
-		\node[fill=white!80!uofgthistle, draw=black, minimum height=3.5cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
-		\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\small \centering OpenFlow\\Tables\end{varwidth}};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\small \centering Stats\\Collector\end{varwidth}};
-		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
-
-		%
-
-		\node[right=3 of aswpos](avfpos){};
-		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
-		\node[below right, inner sep=2pt] at (avf.north west) {\small Agent VF};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\small \centering Flowstate\\Database\end{varwidth}};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\small \centering TRS\\Scheduler\end{varwidth}};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
-		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};
-
-		%%%
-
-		\tikzset{>=stealth}
-
-		\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
-		\draw[thick, ->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};
-
-		\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
-
-		\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
-		\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};
-
-		\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};
-
-		\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
-		\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};
-
-		\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
-		\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
-		\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
-		\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
-		\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};
-
-		\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
-		\node[right=0.3] at (avfsa.east) {\tiny Experience};
-	\end{tikzpicture}}
-	\caption{
-		System architecture  for our RL-driven DDoS defence system.
-		\label{fig:sys-arch}
-	}
-\vspace{-1em}
-\end{figure}
-
 \begin{figure}
@@
 		%		Mean IAT, inbound packet sizes, and global state offer the best predictive performance.
-		\vspace{-1em}
 		\label{fig:udp-feature-plots}
 	}
+\vspace{-1em}
 \end{figure}
@@
 		%		Global state and Mean IAT still offer the greatest improvement above baseline.
-		\vspace{-1em}
 		\label{fig:tcp-cap-feature-plots}
 	}
+	\vspace{-1em}
 \end{figure}
@@
 
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.8\linewidth]{../plots/ftprep-tcp-laf-cap-box}
-%	\vspace{-1cm}
-%	\caption{
-%		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken.
-%		\label{fig:tcp-laf-feature-plots}
-%	}
-%\end{figure}
-
-%\section{Rethinking the State Space}\label{sec:rethinking-the-state-space}
-%
-%Per-flow agents require features which offer high predictive power, making behavioural differences readily apparent.
-%Expanding on the measures discussed in \cref{sec:motivation}, we believe the following features to be useful (and humanly justifiable), and investigate their use on different traffic types:
-%%?? We use these features, and why...
-%
-%\fakepara{Global state}
-%This is the vector of current load measurements along a flow's path introduced in \cref{sec:feature-space}.
-%These values indicate the overall health of the network, and crucially are all measurements which an agent directly controls.
-%
-%\fakepara{Source IP address}
-%Ordinarily trivial to spoof, though reflectors are themselves legitimate services being abused by spoofing attackers.
-%As a result, they communicate with attack victims using their own IP address.
-%In real-world scenarios the addresses of a set of reflector nodes might exhibit CIDR block similarity \cite{DBLP:conf/imc/CollinsSFJWSK07}, e.g., similarly unhardened services exposed by a single organisation.
-%
-%\fakepara{Last action taken}
-%An agent's current belief in the maliciousness of a flow.
-%This feature also serves as a reference point for determining whether a flow is actually malicious based on its response.
-%This feature only makes sense when combined with another flow feature, and never appears individually tile-coded.
-%
-%\fakepara{Flow duration and size}
-%Features which describe the length of time a connection has been active, and the amount of data transferred within that time.
-%An extraordinarily long flow, having sent a lot of data, could be more likely to be an amplifier: though most (\SI{62}{\percent}) waves of amplifier traffic last shorter than \SI{15}{\minute} \cite{DBLP:conf/raid/KramerKMNKYR15}, this is considerably longer than the typical length of an HTTP request/response.
-%
-%\fakepara{Correspondence ratio}
-%The ratio between upstream and downstream traffic for a source IP.
-%We define this to be $C_X = \min(\uload{\cdot}, \dload{\cdot})/\max(\uload{\cdot}, \dload{\cdot})$, where a value close to 0 indicates strong asymmetry.
-%
-%\fakepara{$\mathbf{\Delta}$ Send/receive rate}
-%The change in traffic rates caused by the last action.
-%Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible in these fields.
-%
-%\fakepara{Mean inter-arrival time (IAT)}
-%A measure of how often packets arrive at the agent's parent switch; low IATs indicate a high number of packets per second, and can be a possible marker of malicious behaviour.
-%We only make use of the mean IAT of \emph{inbound} traffic.
-%
-%\fakepara{(Per-window) packet count}
-%The amount of packets sent to/from a source over a flow's lifetime (or the current window of measurement), similar in use to flow size and mean IAT.
-%
-%\fakepara{Mean packet size per window}
-%Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
-%Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.
-%
-%The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
-%If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have no correlation with the class of traffic a flow contains.
-%
-%All of the above features, save for global state, are 1-dimensional.
-%\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), using a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
-%%\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends being observed when the same technique is applied with UDP traffic (omitted).
-%%?? Core findings---different protocols need different features, so everything we proposed above has a use!
-%The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
-%Packet-level and per-window statistics are at their based when combined with the last action for both UDP-like and TCP-like traffic (omitted).
-%All features converge to their highest-observed performance within around \num{4000} timesteps.
-%In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
-%?? How do they do when combined after individual training? Pretty well, especially for TCP.
-%Additional testing shows that the learned per-feature policies may be easily combined (by summing action values), and that this technique is particularly effective for TCP; these results are omitted to preserve space.
-%In no cases, however, do we manage to completely block attack traffic---at convergence, we observe that system load remains consistently at $U_s$.
-
-\section{A New Normal}\label{sec:a-new-normal}
-
-%In establishing...
-%
-%?? How will I structure this?
-%?? Motivation -> Model -> Results?
-%?? OR Use the results of the last section to springboard into here?
-
-%From what we have seen, it is difficult (or impossible) for trace-based or numerical simulations to correctly capture certain dynamics without an extraordinary amount of care or consideration.
-%As it turns out, 
-%Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
+Our state space combines global state (network load) with per-flow measurements.
+Each is tile-coded with 8 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
+Features are tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
+
+%?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
+Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}): the bandwidth measurements regularly received from monitors in the environment.
+For any flow, an agent computes the path it would take through the network.
+The incoming load along the first hop, last hop, and tertiles of the path are then tile-coded together.
+In the event that this path is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
+\Cref{fig:global-state-path} illustrates the process.
+Global state built in this way is compatible with multipath, multi-destination networks, offering support for diverse deployment environments from endpoint servers up to transit ASes.
+Computing a flow's path is not computationally expensive: multipath routing is fast since typical ECMP routing algorithms simply hash a packet's 5-tuple and are deterministic to provide consistent quality-of-service.
+
+%?? Path computation fast and efficient due to deterministic routing based on hashes (ECMP)
+%?? Works for any arbitrary topology, even
+
+\Cref{fig:udp-feature-plots} shows the effectiveness of per-flow features for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), on a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
+``Baseline'' denotes an unprotected network.
+The plots demonstrate that different protocols and traffic classes are best defended by different features---every feature presented has value.
+Packet-level and per-window statistics are at their best when combined with the last action for both UDP-like and TCP-like traffic (omitted).
+All features converge to their highest-observed performance within around \num{4000} timesteps.
+In general, some of the most effective features are global state, mean IAT, mean inbound packet size and $\Delta$ rates.
+The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
+If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have no correlation with the class of traffic a flow contains.
+
+\subsection{Reward Function}
+
+%?? Say ``we take $R_G$ from M and K''.
+%
+%?? Reward at each destination.
+
+\newcommand{\arrload}[2]{\operatorname{load}^{#2}_{t}(#1)}
+\newcommand{\uload}[1]{\arrload{#1}{\uparrow}}
+\newcommand{\dload}[1]{\arrload{#1}{\downarrow}}
+\newcommand{\bload}[1]{\arrload{#1}{\updownarrow}}
+\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
+%\fakepara{Reward function directionality}
+%The reward functions, as defined, do not take traffic direction into account.
+%We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
+We denote the upstream, downstream and combined loads at a node $s$ as $\uload{s}, \dload{s}, \bload{s} \in \mathbb{R}$ respectively.
+Each destination node $s$ generates a reward signal, $R_{s,t}$, at every timestep $t$.
+Assuming each destination has access to some classification function $g(\cdot)$ which estimates the volume of legitimate traffic received, and expects to receive $\mathit{traffic}_s$, we may compute the reward:
+\begin{subequations}
+%	\vspace{-0.5cm}
+	\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
+	\begin{gather}
+	\cond{s} = [\max(\uload{s}, \dload{s}) > U_s],\\
+	R_{s,t} = (1 - \cond{s}) \frac{g(\arrload{s}{-})}{\mathit{traffic}_s} - \cond{s},\label{eqn:reward-rtx}
+	\end{gather}
+	replacing $\arrload{s}{-}$ in \cref{eqn:reward} with whichever directional load is prioritised according to the traffic characteristics of the deployment environment, where $\cond{s}$ represents the ``overloaded'' condition at destination $s$.\label{eqn:reward}
+\end{subequations}
+We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, and suggest $\bload{\cdot}$ for general deployment or heterogeneous traffic.
+These choices reflect where the bulk of transmitted bytes in each traffic model are observed (or the lack of this knowledge).
+
+While our use and definition of $g(\cdot)$ appears nebulous, there are many ways to infer this quantity in practice.
+End-host servers may make use of canary flows or active measurement schemes, and can employ quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
+ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
+For testing, we use \emph{a priori} knowledge of which flows are legitimate; we leave the use of noisier real-world data as future work.
+Even without online rewards, a well-trained agent needs only to follow the greedy policy it has learned from offline (simulated) training.
+
+If a network is believed to be vulnerable to indirect attacks, such as link-flooding attacks, we may use the following reward:
+\begin{equation}
+	R_{s,t}^{\mathit{Cross}}(\beta) = \beta R_{s,t} + (1 - \beta) \min{\{R_{s',t} | s' \ne s\}} \label{eqn:lfa-reward}
+\end{equation}
+where the collaboration parameter $\beta \in [0,1]$ models the expected degree of interference between flows, and $s, s'$ are protected destination nodes in the network.
+The key insight underpinning LFAs is that flows can affect a target \emph{without communicating with that target}.
+$\beta$ then acts as a tunable parameter which can incentivise agents to remove flows which harm overall system health, by including the performance of the worst-performing destination.
+However, we do not examine such attacks (and the effectiveness of $R_{s,t}^{\mathit{Cross}}$) at present.
+
+\subsection{Action Space}
+When monitoring a flow, an agent uses its state vector to decide which portion of its \emph{inbound} traffic should be dropped.
+%?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
+We choose to drop packets (rather than impose traffic limits) to have a discrete action space without prior knowledge of traffic characteristics.
+Furthermore, we need not consider burstiness, fairness and parameters such as bucket sizes which could require tuning.
+We offer two models, which differ only on how to choose $p$.
+The intuition behind these designs is that an agent may directly choose the level of packet drop at the expense of a larger action space (\emph{Instant control}), or that we may reduce the action set (and learn a policy within fewer actions) by making use of domain knowledge (\emph{Guarded control}).
+
+\subsubsection{Instant control}
+Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$.
+These choices ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
+Since this model requires no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely opportunistic/myopic).
+
+\subsubsection{Guarded control}
+\Textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
+We apply their observations, constraining how an agent treats each flow via simple finite state automata: restricting $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
+The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
+This stops agents from wildly varying $p$ beyond their ability to correlate a host's response to its matching action.
+
+We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limits (\SI{25}{\percent} increments for UDP traffic).
+The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
+This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.
+As each agent now needs the capability to plan ahead, we require a discount factor $\gamma \ne 0$, a parameter which allows future value to influence value updates (\cref{eqn:sg-sarsa}).
+We found $\gamma = 0.8$ to be the most effective choice during exploratory testing.
+
+To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$), chosen due to TCP's higher prevalence.
+An agent must now choose to punish a flow multiple times in succession to block it, reducing variance while allowing an agent to see how a host reacts to controlled conditions.
+While beneficial for legitimate flows who behave similarly to attack traffic, this introduces a lag of an extra two actions in blocking illegitimate flows.
+
+\subsubsection{Risks}
+Our mode of action means that each agent carries a risk of introducing collateral damage into the network.
+This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p\ne0$ in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
+This weakness remains in modern flavours of TCP---TCP Cubic has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
+Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets and \SIrange{77}{84}{\percent} of data volume\footnote{Repository link removed for double-blind reviewing.}.
+%Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
+QUIC now comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
+Furthermore, applying a single rule to many flows (as in \textcite{DBLP:journals/eaai/MalialisK15}) increases the likelihood of collateral damage.
+When considering that many hosts have an especially adverse reaction to our main means of control, flow-level granularity becomes an obvious choice.
+
+\section{System Architecture}\label{sec:system-architecture}
+
+%?? Here is where I would put my ramblings about system architecture...
+We present our design of a system which supports the effective real-world deployment of RL-based DDoS mitigation.
+\Cref{fig:sys-arch} displays this, separating system elements which are local to each agent from those which reside elsewhere in the network.
+We operate each agent as a \emph{virtualised Network Function} (vNF) adjacent to a software-defined switch.
+SDN allows a controller (or authorised hosts) to install actions, forwarding rules and logic into a switch at runtime.
+Moreover, networks of this kind more naturally enable the future relocation and easy installation of learners.
+Agent vNFs communicate with these co-hosted switches to install probabilistic packet-drop rules.
+We describe the main purpose and operation of each module within an agent's vNF, and discuss techniques to make deployment more efficient using existing technologies.
+
+\subsection{Core and RL Executor}\label{sec:core-and-rl-executor}
+The core module is the main loop in an agent's architecture.
+At each timestep, the core receives information about which flows have arrived and should be acted upon from the \emph{TRS Scheduler}.
+The core then retrieves the current and previous state vector associated with each flow from the \emph{Flowstate Database}, passing them into the RL algorithm alongside the last action chosen for that flow (if available).
+
+The RL algorithm then returns an action.
+Each action is passed to the database, which computes and returns a packet drop rate according to the agent model (\emph{Instant/Guarded}) while updating flow state.
+This is then converted into an OpenFlow message carrying packet drop rules; these are batched to the agent's switch using the same groupings produced by the scheduler.
+Finally, timing information is passed back into the scheduler to refine its estimates about how much work should be scheduled in the next timestep.
+
+State space sizes guarantee that an \emph{Instant} policy remains under \SI{520}{\kibi\byte}, though our sparse representation typically leads to far smaller policies: $\sim$\SI{17.8}{\kibi\byte} from our experiments.
+\emph{Guarded} policies are \SI{30}{\percent} of this size.
+As described earlier, action updates require a constant number of floating point operations---\num{160} floating point additions and \num{32} multiplications per update of $\wvec{}$ with per-tile updates, with \num{160} additions required to choose an action.
+The vast majority of these operations can be vectorised.
+Action computation for \emph{Guarded} agents is cheaper still, requiring only \num{48} additions per action.
+
+\subsection{Stats API and Collectors}
+Agents require information from the network and one another to be effective.
+Our agents can act either independently, having no agent-to-agent communication, or cooperatively.
+In the latter case agents transfer, when possible, \emph{experience} to one another---lists of state-action-state transitions with associated rewards.
+It's noted that a transition may be high-value or surprising to one agent, while well-known to another, causing each to produce different policy updates from the same unit of experience.
+For this reason we do not transfer policy deltas between agents, causing each to learn its own policy.
+Which scheme achieves better performance is left to future work.
+
+%?? How do stats get from switcehs to the agent?
+Load collectors and estimators periodically push observations to each active agent vNF.
+In our current implementation, load statistics are gathered via vNFs active at each network switch, though we expect that OpenFlow stats requests, NetFlow or SNMP data may be used to derive these cheaply.
+Transferring state to agents and experience sharing can both be made more efficient through effective use of broadcast addressing in a target network.
+Depending on the capabilities of switches in the network, the estimator can either send benign traffic estimates or parameters for use in a reward function.
+
+Gathering and transmission of load/flow statistics would be difficult to perform quite as often as an emulated environment allows.
+However, the measurements acquired in such a scenario are likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
+
+\subsection{Flowstate Database}
+For each flow 5-tuple, we hold two state vectors containing the features described in \cref{sec:feature-space}---the current state, and the state which induced the last action.
+To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set.
+State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
+For flows outside of the current work list, we simply replace the stored vector.
+
+\subsection{TRS Scheduler}
+Acting on an unbounded set of flows in each timestep introduces potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and the risk that flow states become outdated.
+At their worst, these risks present additional attack surface to an adversary.
+To tackle these problems, we make use of \emph{Timed Random Sequential} updates.
+
+The scheduler begins with a shuffled work list of active flows.
+When requested, the scheduler estimates the cost of an action computation using the timing information received from the core, proceeding down the list to send a set of 5-tuples to the core which can be handled in a set time limit.
+The scheduler continues until the list is empty, at which point it is repopulated and reshuffled with active flows.
+
+Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations on handling short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---an agent is typically able to process around 3 flows in this time.
+We expect deadlines should be tuned based on the frequency at which statistics arrive.
+The amount of processed flows per deadline depends on the agent design (FLOP count, policy size), but also on the amount of flow telemetry data needing processed---our current implementation is written in python, restricting this handling to a single thread.
+An implementation in a systems language such as Rust or C++ would allow faster concurrent processing.
+
+There is a risk that so much work can be queued up that an agent is never able to act on some attack flows.
+A solution is to impose an upper bound on the amount of action computations/policy updates that can be performed before the work list is regenerated.
+This removes the guarantee that all flows will be visited often, but if updates occur regularly then this sampling may be sufficient to achieve good performance.
+
+\subsection{Agent Switches}
+%?? How do we do stat collection?
+Our agent switches operate a modified version of Open vSwitch, implementing an action which requests that each matched packet be dropped with a certain probability.
+To get around the lack of floating-point support in many environments, we represent this probability using a 32-bit unsigned integer (scaling \num{1.0} to $2^{32}-1$).
+On commodity hardware, we believe that a similar effect can be achieved using OpenFlow meters (at the expense of these being stateful measures).
+
+We use OpenFlow groups to simplify control messages: premade tables with permitted levels of packet drop.
+This saves some overhead compared to using experimenter/extension headers.
+Flows are automatically given a group with the default level of packet drop (according to the chosen agent design), meaning that switches don't need to refer to a controller or the agent vNF.
+
+%?? What do our modifications to openflow look like? How do we get around the lack of floating point math?
+
+\begin{figure}
+	\centering
+	\resizebox{0.8\linewidth}{!}{\begin{tikzpicture}
+		\node(remote){Remote};
+
+		%%%
+
+		\node[below=0.05 of remote](swpos){};
+		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of swpos.north west](sw1){};
+		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw1.north west](sw2){};
+		\node[fill=white!80!uofgcobalt, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of sw2.north west](switch){};
+		\node[below right, inner sep=2pt] at (switch.north west) {\small Switch};
+		\node[fill=white!90!uofgcobalt, draw, rectangle, rounded corners=0.05cm, above=0.1] (oswlc) at (switch.south) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
+
+		%
+
+		\node[right=2 of swpos](epos){};
+		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of epos.north west](e1){};
+		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e1.north west](e2){};
+		\node[fill=white!80!uofgthistle, draw=black, minimum height=1.5cm, minimum width=3.5cm, below right= 0.1 of e2.north west](egress){};
+		\node[below right, inner sep=2pt] at (egress.north west) {\small Egress Switch};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (eglc) at ($(egress.south) + (-0.85,0)$) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, right=0.1] (egest) at (eglc.east) {\begin{varwidth}{1.5 cm}\small \centering Estimator\\$g(\cdot)$\end{varwidth}};
+
+		%
+
+		\node[right=3.5 of epos](apos){};
+		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of apos.north west](a1){};
+		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a1.north west](a2){};
+		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=1.5cm, minimum width=2cm, below right= 0.1 of a2.north west](otheragent){};
+		\node[below right, inner sep=2pt] at (otheragent.north west) {\small Agent vNF};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, above=0.3] (oasa) at (otheragent.south) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
+
+		%
+
+		\node[below=2.3 of remote.west](linestart){};
+		\path let \p1 = (linestart) in node (lineend) at (9,\y1){};
+		\draw [dashed] (linestart) -- (lineend);
+
+		%%%
+
+		\node[below=2.4 of remote](local){Local};
+
+		%%%
+
+		\node[below=0.05 of local](aswpos){};
+		\node[fill=white!80!uofgthistle, draw=black, minimum height=3.5cm, minimum width=2.2cm, below right= 0.1 of aswpos.north west](aswitch){};
+		\node[below right, inner sep=2pt] at (aswitch.north west) {\small Agent Switch};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswoft) at (aswitch.south) {\begin{varwidth}{1.5 cm}\small \centering OpenFlow\\Tables\end{varwidth}};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswsc) at (aswoft.north) {\begin{varwidth}{1.5 cm}\small \centering Stats\\Collector\end{varwidth}};
+		\node[fill=white!90!uofgthistle, draw, rectangle, rounded corners=0.05cm, above=0.1] (aswlc) at (aswsc.north) {\begin{varwidth}{1.5 cm}\small \centering Load\\Collector\end{varwidth}};
+
+		%
+
+		\node[right=3 of aswpos](avfpos){};
+		\node[fill=white!60!uofgpumpkin, draw=black, minimum height=3.5cm, minimum width=4.5cm, below right= 0.1 of avfpos.north west](avf){};
+		\node[below right, inner sep=2pt] at (avf.north west) {\small Agent vNF};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.15] (avfsa) at ($(avf.north) + (0.2,0)$) {\begin{varwidth}{1.5 cm}\small \centering Stats API\end{varwidth}};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=0.9] (avfdb) at (avfsa.south west) {\begin{varwidth}{1.5 cm}\small \centering Flowstate\\Database\end{varwidth}};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.1] (avfsched) at (avfdb.east) {\begin{varwidth}{1.5 cm}\small \centering TRS\\Scheduler\end{varwidth}};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, below=2.3] (avfcore) at (avfsa.south) {\begin{varwidth}{1.5 cm}\small \centering Core\end{varwidth}};
+		\node[fill=white!90!uofgpumpkin, draw, rectangle, rounded corners=0.05cm, right=0.8] (avfrl) at (avfcore.east) {\begin{varwidth}{1.5 cm}\small \centering RL\end{varwidth}};
+
+		%%%
+
+		\tikzset{>=stealth}
+
+		\draw[thick, ->] (aswlc) -- (avfsa.west) node[midway,above] {\tiny Current load};
+		\draw[thick, ->] (aswsc) -- (avfsa.west) node[midway,sloped, above] {\tiny Flow stats};
+
+		\draw[thick, ->] (oswlc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
+
+		\draw[thick, ->] (eglc) -- (avfsa) node[midway,above, sloped] {\tiny Current load};
+		\draw[thick, ->] (egest) -- (avfsa) node[midway,above, sloped] {\tiny Estimation data};
+
+		\draw[thick, <->] (oasa) -- (avfsa) node[midway,above, sloped] {\tiny Experience};
+
+		\draw[thick, ->] (avfsa) -- (avfdb) node[midway,above, sloped] {\tiny State};
+		\draw[thick, ->] (avfsa) -- (avfsched) node[midway,above, sloped] {\tiny Live flows};
+
+		\draw[thick, ->] (avfcore) -- (aswoft) node[midway,above, sloped] {\tiny Packet drop rules};
+		\draw[thick, <->] (avfcore) -- (avfdb) node[midway,below, sloped] {\tiny State};
+		\draw[thick, <->] (avfcore) -- (avfsched) node[midway,below, sloped] {\tiny Work};
+		\draw[thick, <-] ($(avfcore.east) + (0,0.1)$) -- ($(avfrl.west) + (0,0.1)$) node[midway,above, sloped] {\tiny Actions};
+		\draw[thick, ->] (avfcore) -- (avfrl) node[midway,below, sloped] {\tiny State};
+
+		\draw[thick, ->] (avfsa.east) to [out=0, in=45] (avfrl.north);
+		\node[right=0.3] at (avfsa.east) {\tiny Experience};
+	\end{tikzpicture}}
+	\caption{
+		System architecture  for our RL-driven DDoS defence system.
+		\label{fig:sys-arch}
+	}
+\vspace{-1em}
+\end{figure}
+
+\section{Evaluation}
+We compare our work's performance and learning rate against Marl \cite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention, across different topologies and workloads.
+Marl builds an overlay tree topology, choosing a packet drop rate $P$ to be applied to \emph{all} flows passing over an agent.
+We made use of the OPUS and TCP traffic models introduced in \cref{sec:a-new-normal}, both topologies discussed below (1-dest vs Fat-Tree), and varied the amount of hosts communicating over each egress node.
+Additionally, we evaluated these models in multi-agent (\emph{separate}, no model sharing) and single-agent modes (\emph{single}, 0-cost perfect information sharing).
+In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (resetting policies between episodes).
+Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes.
+Malicious traffic is provided by \emph{hping3}, generating UDP-flood traffic targeting random ports with MTU-sized packets.
+All experiments were executed on Ubuntu 18.04.2 LTS (GNU/Linux 4.4.3-040403-generic x86\_64), using a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
+
+\subsection{Modelling Normal Traffic}\label{sec:a-new-normal}
 Our network model allows live testing of reactive TCP and UDP traffic in an SDN environment, over arbitrary topologies, with an explicit focus on preserving real-time dynamics in a way that trace-based evaluation cannot.
 We are primarily interested in load and packet inter-arrival characteristics, and in how they evolve in response to agent actions.
-%We aim to capture interactive, correlated back-and-forth exchanges associated with live traffic.
-%Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
-%If need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.
-
-%??
-ANGLE: set up an environment to test \emph{specific} behaviours to examine \emph{specific} problems in past work.
-I make no claims that it is perfect or representative for all traffic, just for this (likely common) behaviour which I expect to plague almost all legit TCP flows.
-
-%??
-Existing sims used for testing such applications reliant on traces, or not sophisticated enough to capture interactive, back-and-forth (correlated) behaviours---possibly discarded as second-hand effects by past work when these are so crucial given user traffic patterns (and the nature of the control signal we choose to enact).
-
-%??
-Remember, the motivation is clear.
-We don't care so much that it is "representative" wrt a specific deployment location or network type.
-The whole purpose of this is that we aim to test specific behaviour which traces cannot replicate (i.e., correlated back-and-forth, dynamics introduced to congestion-aware protocols, ...)
-%?? If we need to, we can condition the distribution of requests according to statistics mined from an existing trace if reviewer number 2 needs that extra push to be convinced.
-
-\subsection{Network Design}
-We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
+
+\fakepara{Network Design}
+We make use of a full SDN, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
 Internal routers begin with knowledge of the shortest path to each protected host, while inbound flows register their return path at each hop, ensuring consistent traffic conditions for each flow.
 OpenFlow \emph{group actions} provide consistent multipath routing where needed. 
-%If several ports offer different (equal-length) paths to a destination, a consistent random port is chosen from the flow-hash by an OpenFlow \emph{Group action} (in \emph{select} mode).
-%If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
 The controller is also responsible for computing how switches respond to ARP requests: this need arises due to the reliance upon Linux's networking stack for live applications.
-%We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.
-
-\subsection{TCP (HTTP) Traffic Model}
+
+\fakepara{TCP (HTTP) Traffic Model}
 %?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
-To model legitimate TCP traffic, server nodes run an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
+To model legitimate TCP traffic, server nodes run an nginx v1.10.3 HTTP daemon, serving static web pages alongside various large files and binaries.
 Benign hosts repeatedly request resources from the server via libcurl.
 Both use the TCP Cubic \cite{rfc8312} flavour of TCP.
-A host's download rate is limited to match its assigned maximum bandwidth, and requests several files known to exist within a website, followed by dependencies for each (stylesheets, images, etc.).
+A host's download rate is limited to its assigned maximum bandwidth, and requests several files which exist within a website, followed by any dependencies (stylesheets, images, etc.).
 This presents a balanced distribution of flow duration and size, with large files included to model bulk transfer.
-On completion, a host changes its IP to generate separate statistics per flow.
-
-\subsection{UDP (Opus) Traffic Model}
-VoIP traffic is heavily different to the above model; packet arrivals are highly periodic due to real-time requirements, flows are constant bitrate, and do not react substantially to lost packets.
-DDoS attack traffic is known to share many of these characteristics, offering an interesting detection problem.
+On completion, a host changes its IP to generate separate flow statistics.
+
+\fakepara{UDP (Opus) Traffic Model}
+VoIP traffic is heavily different from HTTP; packet arrivals are highly periodic due to real-time requirements, flows are constant bitrate, and do not react significantly to lost packets.
+DDoS attack traffic is known to share many of these traits, posing an interesting detection problem.
 We present a traffic model\footnote{Repository link removed for double-blind reviewing.} based on Discord\footnote{\url{https://discord.gg}}, a messaging and VoIP platform.
@@
 
-\section{Evaluation}
-
-%Traffic is played back from hosts via Tcpreplay at a bandwidth assigned uniformly from a `good' or `bad' distribution, each using the same pcap file with source and destination IP addresses rewritten.
-
-We compare our work's performance and learning rate against Marl \cite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention, across different topologies and workloads.
-We made use of the OPUS and TCP traffic models introduced in \cref{sec:a-new-normal}, both topologies discussed below (1-dest vs Fat-Tree), and varied the amount of hosts communicating over each egress node.
-Additionally, we evaluated these models in multi-agent (\emph{separate}, no model sharing) and single-agent modes (\emph{single}, 0-cost perfect information sharing).
-In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (resetting policies between episodes).
-Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes.
-Malicious traffic is provided by \emph{hping3}, generating UDP-flood traffic targeting random ports with MTU-sized packets.
-All experiments were executed on Ubuntu 18.04.2 LTS (GNU/Linux 4.4.3-040403-generic x86\_64), using a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
-%All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
-%All code underpinning these findings is available on a public repository.\footnote{Private until publication.}
-
-\subsection{Single Destination}\label{sec:single-dest}
+\subsection{Single Destination Topology}\label{sec:single-dest}
 The network is tree-structured: one server $s$ connects through a dedicated switch to $k$ team leader switches, each connected to $\ell$ intermediate switches, which in turn each connect to $m$ egress switches.
@@
 
-\subsection{Multiple Destinations}
+\subsection{Multiple Destination Topology}
 The previous topology allows for direct comparison against the state-of-the-art, however it is hard to argue its relevance to specific classes of victim or how it might interact with applications.
@@
 We now compare the performance of our two new models (\emph{Instant}, \emph{Guarded}) against existing RL work (\emph{Marl}) with different traffic and topologies, varying the host-to-learner ratio $n$.
-%?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
-%Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
 We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold.
@@
 We see a reversal of this trend for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
-%Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
 \Cref{fig:udp-tree-box} shows the preserved traffic in multi-agent mode.
-%?? Discuss multi-dest topol once all results available.
 When defending multiple destinations, we see a sharp decrease in the effectiveness of all agent designs.
@@
 
-%Although we improve upon Marl in both identified problem cases, the improvements are small for UDP traffic.
-The most likely explanation for low UDP protection is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
+We observe that low UDP protection arises because agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
 The increased state space size and similar characteristics between benign and attack traffic make this more likely.
@@
 \subsection{RL algorithms}
-When choosing a learning algorithm, we compared semi-gradient Sarsa against Q-learning, and variants of these algorithms based on \emph{eligibility traces} (Sarsa($\lambda$), Watkins's Q($\lambda$)).
-Our experiments found that Sarsa offered the best performance and behaviour; \cref{fig:algo-spf-tcp} shows a trend observed across model choices and communication except for multiagent \emph{Instant} protection of TCP where Q-learning succeeds.
+When choosing a learning algorithm, we compared semi-gradient Sarsa against Q-learning, and variants of these algorithms based on \emph{eligibility traces} (ETs) (Sarsa($\lambda$), Watkins's Q($\lambda$)).
+Sarsa and Q-learning represent \emph{on-} and \emph{off-policy} algorithms respectively, a theoretical distinction affecting which state-action pair contributes the discounted value to $\delta_t$, where off-policy algorithms are associated with being able to better learn from distinct streams of experience.
+ET algorithms allow past gradient values to slowly decay, having a lasting effect on learning by offering another means for knowledge to propagate.
+Supposing $\lambda=0.5$, a tile which was active three timesteps ago would receive an update of $\lambda^3\delta_t=\frac{1}{8} \delta_t$.
+Our experiments found that Sarsa offered the best performance and behaviour; \cref{fig:algo-spf-tcp} shows a trend observed across model choices and communication strategies---that Sarsa leads by a few percentage points, and that ET methods diverge over time.
+Inter-related mechanisms for learning often introduce bias---we suspect that learning from many actions per timestep causes this when combined with ETs.
+An exception is in multiagent \emph{Instant} protection of TCP, where Q-learning succeeds (omitted).
 
@@
 \fakepara{RL in Networks}
-RL has seen little examination in the challenge of intrusion detection/prevention.
-Past works use the approach as a traditional classifier \cite{shamshirband2014anomaly,DBLP:conf/mates/ServinK08}.
-Given that one of the main strengths of RL techniques is the ability to control (and adapt through) ongoing interaction, these works don't apply the paradigm at its fullest potential.
-
-Effective uses have lain in learning when it is best to \emph{communicate} and share knowledge between anomaly detectors \cite{DBLP:conf/paisi/XuSH07}, and in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-\Textcite{DBLP:conf/sigcomm/MaoNA17} employ client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming.
-
-\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation.
-They find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
-To overcome the high latency of neural network computations, two agents are trained, handling aspects of short (priority thresholds) and long flows (bespoke allocations) respectively.
+RL has seen little examination in the challenge of intrusion detection/prevention---past works use the approach as a traditional classifier \cite{shamshirband2014anomaly,DBLP:conf/mates/ServinK08}.
+Effective uses have lain in learning when it is best to \emph{communicate} and share knowledge between anomaly detectors \cite{DBLP:conf/paisi/XuSH07}, and in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17}, for process allocation \cite{DBLP:conf/hotnets/MaoAMK16}, traffic optimisation \cite{DBLP:conf/sigcomm/ChenL0L18}, and in multimedia bitrate selection \cite{DBLP:conf/sigcomm/MaoNA17}.
+%\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation.
+%They find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
+%To overcome the high latency of neural network computations, two agents are trained, handling aspects of short (priority thresholds) and long flows (bespoke allocations) respectively.
 
@@
 
-\section{Conclusions and Future Work}
-Through this paper, we have introduced reinforcement learning and its relevance to network intrusion prevention.
-We believe that the ability to learn network control as an online feedback loop makes it particularly effective in modern networks.
-
-We have put forth an RL agent design which acts per-flow, describing the algorithmic modifications and engineering choices needed to make its deployment feasible.
+\section{Conclusions}
+In this paper we have put forth RL agent designs which act per-flow, describing the algorithmic modifications and engineering choices needed to make their deployment feasible.
 Supporting this, we've demonstrated the value of the features these agents use as evidence.
-Our evaluation shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, with \emph{Guarded} agents showing the most promise for future evaluation.
+Our evaluation against realistic normal traffic shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, protecting \SI{2.5}{$\!\times$} more TCP traffic than past RL work, with \emph{Guarded} agents showing the most promise for future evaluation.
 
@@
 
-We hope it is clear that reinforcement learning holds promise and can inspire further innovation.
+We hope it is clear that RL holds promise and can inspire further innovation.
 It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques discussed in \cref{sec:related-work} rely upon.
@@
 %Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
-This work was supported by a nameless funder with a name approximately as long as this line [grant number 0].
-%This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].
+This work was supported by a nameless funder with a name approximately as long as this line [grants 00000000, 00000000 and 00000000], and another nameless funder with a project name and code bringing us closer to the amount of space required here to prevent future unplanned changes.
+%This work was supported in part by the Engineering and Physical Sciences Research Council [grants EP/M508056/1, EP/N033957/1 and EP/P004024/1] and European Cooperation in Science and Technology (COST) Action CA15127: RECODIS---Resilient communication and services.
 

### 2019-07-31 paper/paper.tex
Just the conclusion + new results to be handled...

@@
 \section{Introduction}
-
 Network anomaly detection and intrusion detection/prevention are continually evolving problems, compounded by the partial, non-\emph{independent and identically distributed} (IID) view of data at each point in the network.
@@
 %?? Discuss the most important conclusions before the outline.
-We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
+We find that the existing work for this task \cite{DBLP:journals/eaai/MalialisK15} fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
 The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.

### 2019-07-31 paper/infocom.tex
Finalest changes.

@@
 \begin{itemize}
-%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
-%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}.
-We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
 	\item Two source-level granularity models for RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models) and suitable features (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
 	\item A description of the system architecture which supports the real-world deployment of such agents (\cref{sec:system-architecture}).
-%	\item An investigation of suitable features for this task (\cref{sec:rethinking-the-state-space}).
 	\item Reactive simulations of HTTP and VoIP web-server traffic (\cref{sec:a-new-normal}).
 	\item A performance comparison of different RL algorithms within these agents, and against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
-%	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
 \end{itemize}
@@
 Supporting this, we've demonstrated the value of the features these agents use as evidence.
+We've shown that RL algorithms can be modified to suit the learning environment, and how an architecture featuring the (minimal) use of SDNs can aid deployment in diverse networks.
 Our evaluation against realistic normal traffic shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, protecting \SI{2.5}{$\!\times$} more TCP traffic than past RL work, with \emph{Guarded} agents showing the most promise for future evaluation.
@@
 The remaining weaknesses invite many further improvements worth investigation: the design of reward functions derived from QoE metrics, agent operation via \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}, and evaluation against evolving DDoS attacks \cite{DBLP:conf/spw/KangGS16}.
-
-We hope it is clear that RL holds promise and can inspire further innovation.
-It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques discussed in \cref{sec:related-work} rely upon.
-Our proposed approach offers substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today, while offering future-proofing against the introduction of new transports.
-It's hoped that more research in this direction will open the door to works which capture the complex and fluid nature of modern networks; evolving topologies, traffic and protocol distributions, and attacks.
+Even now, our approach offers distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, more flexible deployment, and automatically learned low-overhead decision-making on the most prevalent class of traffic.
+
+%We hope it is clear that RL holds promise and can inspire further innovation.
+%It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques discussed in \cref{sec:related-work} rely upon.
+%Our proposed approach offers substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today, while offering future-proofing against the introduction of new transports.
+%It's hoped that more research in this direction will open the door to works which capture the complex and fluid nature of modern networks; evolving topologies, traffic and protocol distributions, and attacks.
 

### 2019-09-23 paper/paper-revised.tex
Progress on revised paper for TNSM
Now includes results on SPIFFY, with some light discussion as to their
relevance. WIP on the rest of the changes mandated...

--- 2019-07-31 paper/paper.tex
@@
-% !TeX document-id = {e7fef4b9-afbc-4e81-94aa-c4847183d047}
+% !TeX document-id = {b3181d9a-ffe0-4331-a65c-c7fb8f275531}
 % !BIB program = biber
@@
 
+\usepackage[normalem]{ulem}
+
 \usepackage[dvipsnames]{xcolor}
@@
 \robustify\bfseries
-\sisetup{detect-all, range-phrase=--, range-units=single, detect-weight=true,detect-inline-weight=math}
+\robustify\emph
+\robustify\uline
+\sisetup{detect-all, range-phrase=--, range-units=single, detect-weight=true,detect-inline-weight=math, table-format=1.3}
 
@@
 %bib
-\usepackage[style=ieee,maxnames=3,mincitenames=1,maxcitenames=2,maxbibnames=99,
+\usepackage[style=ieee,minnames=1,maxcitenames=2,maxbibnames=2,
 mincrossrefs=99,minxrefs=99,
@@
 %?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
-Briefly, this paper contributes:
-\begin{itemize}
-%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
-%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}.
-We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
-	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
-	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
-	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
-	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
-\end{itemize}
+
+\subsection{Contributions}
+This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
+These are designed to make effective decisions irrespective of protocol, and to act on individual flows at the edge of network topologies with arbitrary shape and size.
+
+We offer an in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
+These features have been suggested by past studies into DDoS prevention, and have been independently tested within their own contexts.
+Our study is the first attempt to quantify the individual efficacy of each in an RL-focussed setting.
+
+We implement reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
+To our knowledge, this is the first attempt to study or replicate Opus-based VoIP traffic, which has become commonplace since the codec's release in 2012.
+These new traffic models inform an empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
+Observing that past work has been compared only against AIMD \cite{DBLP:journals/ton/YauLLY05}, we additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS defences.
 
@@
 The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
-Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners.
+Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners (e.g., as examined by \textcite{TNSM-that-guys-paper}).
 Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
@@
 	\centering
-	\resizebox{0.8\linewidth}{!}{
+	\resizebox{0.65\linewidth}{!}{
 		\begin{tikzpicture}[
@@
 \subsection{Systems Considerations}
-Taking many actions per-timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
+Taking many actions per timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
 This introduces some potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated.
@@
 For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
-Uncompleted work is passed on to the next timestep, until the list is emptied, at which point is is repopulated using the set of available measurements.
+Uncompleted work is passed on to the next timestep, until the list is emptied, at which point it is repopulated using the set of available measurements.
 To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
@@
 
-\section{A New Normal}\label{sec:a-new-normal}
+\section{Modelling Normal Traffic}\label{sec:a-new-normal}
 
@@
 All traffic passes over a central server which groups hosts into rooms, and is forwarded to other participants; we do not replicate pre-call Websocket traffic which would be used for authentication.
-There is no peer-to-peer traffic---the server acts as a TURN relay \cite{rfc5766} for all hosts.
+There is no peer-to-peer traffic---the server acts as a TURN relay for all hosts.
 %?? Reflective factor among \emph{authenticated hosts}.
@@
 Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network---we aim to see how these characteristics affect performance and learning rate.
+Marl is already known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} approach to DDoS mitigation, although the state of the art has long since moved on.
+
+To paint a more current picture, we compare our work against an effective modern-day algorithmic approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
+SPIFFY tests a proportion of flows by rerouting them through an alternate path with higher bandwidth, and observing how their speed has changed some time later.
+We compare ourselves against this approach to position our new RL agent designs against the modern state of the art, observing that SPIFFY has a similar mode of interaction to RL-based systems (taking action, observing an effect, and acting once again) and does not rely on protocol characteristics or signatures.
+We make the simplifying assumption that a suitable unused path exists (with identical bandwidth to the server's link).
+We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion factor), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\times$} bandwidth expansion from legitimate flows, making a judgement after \SI{5}{\second}.
 
@@
 %Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
-We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold.
+We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold and underlining the best RL-based model.
 Average rewards take into account any portions of time that an agent allows illegal system states.
@@
 
-	\expandableinput ../tables/tnsm-avg-reward.tex
+	\resizebox{0.8\linewidth}{!}{
+		\expandableinput ../tables/tnsm-tree-avg-reward-spiffy.tex
+	}
 \end{table}
@@
 
-	\expandableinput ../tables/tnsm-ecmp-avg-reward.tex
+	\resizebox{0.8\linewidth}{!}{
+		\expandableinput ../tables/tnsm-ecmp-avg-reward-spiffy.tex
+	}
 \end{table}
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-udp-box-separate}
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-udp-box-separate}
 
@@
 \Cref{fig:udp-tree-box} shows the preserved traffic in multi-agent mode.
+Interestingly, SPIFFY is unable to effectively protect constant bitrate legitimate traffic.
 
@@
 Our \emph{Instant} and \emph{Guarded} designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
+SPIFFY remains ineffective at protecting this class of traffic.
 
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-single}
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-box-single}
 
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-16-single}
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-16-single}
 
@@
 \Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
+Here, SPIFFY greatly outperforms all of the RL-based methods.
 %SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
@@
 Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
+SPIFFY continues to outperform the RL-based methods.
 
@@
 %Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
-Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
+Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this strand still trails behind existing (exact) DDoS flow detection mechanisms where TCP traffic is concerned.
+The ability to better protect VoIP traffic when compared against one of these approaches is a curious observation, which suggests that other (exact) protocol-agnostic approaches may carry hidden assumptions and is a promising direction for future investigation.
 Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
 The benefits we have at present are, however, substantial.
-What we offer above the alternate approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
+What we offer above many of the alternate approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
 The techniques we propose offer substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today.

### 2019-09-23 tables/tnsm-ecmp-avg-reward-spiffy.py

@@ -1,0 +1,144 @@
+import csv
+import numpy as np
+from os import path
+import sys
+
+traffics = ["opus", "tcp"]
+ns = [2, 4, 8, 16]
+models = ["mpp", "spf"]
+variants = [
+	("separate", None, "Separate", None),
+	("single", None, "Single", None),
+]
+
+illegal_combos = set()
+#illegal_combos.add(("spf", "opus", "single"))
+# illegal_combos.add(("spf", "udp", "channel"))
+# illegal_combos.add(("spf", "tcp", "channel"))
+
+def get_average_reward(filename):
+	output = 0.0
+	try:
+		with open(filename, "r") as csvfile:
+			reader = csv.reader(csvfile)
+			# _ = nereader.next()
+
+			dats = []
+			for i, row in enumerate(reader):
+				if i == 0:
+					continue
+
+				dats.append(float(row[3]))
+
+			output = np.mean(dats)
+	except:
+		print("file {} not found".format(filename))
+	return output
+
+def write_table(out_file, data):
+	lm = len(models)
+	lv = len(variants)
+
+	out_file.write(
+		"\\begin{{tabular}}{{@{{}}cc{}@{{}}}}\n".format("S"*((lm*lv)+2))
+	)
+	out_file.write(
+		"\\toprule\\multicolumn{1}{c}{Traffic} & " + \
+		"\\multicolumn{1}{c}{$n$} & " + \
+		"\\multicolumn{1}{c}{SPIFFY} & " + \
+		"\\multicolumn{1}{c}{Marl} & " + \
+		"\\multicolumn{{{}}}{{c}}{{Instant}} & ".format(lv) + \
+		"\\multicolumn{{{}}}{{c}}{{Guarded}} ".format(lv) + \
+		"\\\\\n" + \
+		"\\cmidrule(lr){{5-{}}}\\cmidrule(lr){{{}-{}}} ".format(5 + lv-1, 5 + lv, 4 + lv + lv) + \
+		" & & & "
+	)
+
+	for _ in models:
+		for (_fn, _replace_str, pres, _replace_m) in variants:
+			out_file.write("& \\multicolumn{{1}}{{c}}{{{}}} ".format(pres))
+
+	out_file.write("\\\\ \\midrule\n")
+
+	# Write file body.
+	for traffic in traffics:
+		for i, n in enumerate(ns):
+			key = (traffic,n)
[…]

### 2019-09-23 tables/tnsm-tree-avg-reward-spiffy.py

@@ -1,0 +1,144 @@
+import csv
+import numpy as np
+from os import path
+import sys
+
+traffics = ["opus", "tcp"]
+ns = [2, 4, 8, 16]
+models = ["mpp", "spf"]
+variants = [
+	("separate", None, "Separate", None),
+	("single", None, "Single", None),
+]
+
+illegal_combos = set()
+#illegal_combos.add(("spf", "opus", "single"))
+# illegal_combos.add(("spf", "udp", "channel"))
+# illegal_combos.add(("spf", "tcp", "channel"))
+
+def get_average_reward(filename):
+	output = 0.0
+	try:
+		with open(filename, "r") as csvfile:
+			reader = csv.reader(csvfile)
+			# _ = nereader.next()
+
+			dats = []
+			for i, row in enumerate(reader):
+				if i == 0:
+					continue
+
+				dats.append(float(row[3]))
+
+			output = np.mean(dats)
+	except:
+		print("file {} not found".format(filename))
+	return output
+
+def write_table(out_file, data):
+	lm = len(models)
+	lv = len(variants)
+
+	out_file.write(
+		"\\begin{{tabular}}{{@{{}}cc{}@{{}}}}\n".format("S"*((lm*lv)+2))
+	)
+	out_file.write(
+		"\\toprule\\multicolumn{1}{c}{Traffic} & " + \
+		"\\multicolumn{1}{c}{$n$} & " + \
+		"\\multicolumn{1}{c}{SPIFFY} & " + \
+		"\\multicolumn{1}{c}{Marl} & " + \
+		"\\multicolumn{{{}}}{{c}}{{Instant}} & ".format(lv) + \
+		"\\multicolumn{{{}}}{{c}}{{Guarded}} ".format(lv) + \
+		"\\\\\n" + \
+		"\\cmidrule(lr){{5-{}}}\\cmidrule(lr){{{}-{}}} ".format(5 + lv-1, 5 + lv, 4 + lv + lv) + \
+		" & & & "
+	)
+
+	for _ in models:
+		for (_fn, _replace_str, pres, _replace_m) in variants:
+			out_file.write("& \\multicolumn{{1}}{{c}}{{{}}} ".format(pres))
+
+	out_file.write("\\\\ \\midrule\n")
+
+	# Write file body.
+	for traffic in traffics:
+		for i, n in enumerate(ns):
+			key = (traffic,n)
[…]

### 2019-09-25 code/marl/tnsm-extras-runner.sh
More auxilliary experiments...

@@ -1,0 +1,25 @@
+#!/bin/bash
+SCRIPT="./tnsm-results.py"
+Scripts=( "./tnsm-results-highrate.py" "./tnsm-results-highdensity.py" "./tnsm-results-highdensitylowwidth.py" )
+#Scripts=( "./tnsm-results-highdensity.py" "./tnsm-results-highdensitylowwidth.py" )
+
+getCount() {
+	python $2 $1
+}
+runExpt() {
+	python $3 $1 $2
+}
+
+./purger.sh
+
+for s in ${Scripts[*]};
+do
+	for i in $(seq 0 $(expr $(getCount 0 $s) - 1));
+	do
+		./purger.sh
+		runExpt 0 $i $s
+	done
+done
+
+./purger.sh
+

### 2019-09-25 code/marl/tnsm-results-highdensity.py

@@ -1,0 +1,230 @@
+import cPickle
+from marl import *
+import sys
+from writer import writeResults, makeResultsAverage, dumbWriter
+
+results_dir = "../../results/"
+
+# prefix, spiffy_mode, discount
+models = [
+#	"m", # MARL
+#	"spiffy", # Real SPIFFY
+#	"mpp", # MARL++
+	"spf", # SPF
+]
+
+# Use this to try out the effect of a sensible discount on Marl++...
+discounts = [
+	#0.0,
+	0.8,
+]
+maths = [
+	True,
+#	False,
+]
+
+algos = [
+	"sarsa",
+#	"q",
+]
+
+# host_p, dt
+host_ps = [
+#	(2, 0.05),
+#	(4, 0.05),
+#	(8, 0.05),
+	(32, 0.05),
+	(64, 0.05),
+	(96, 0.05),
+	(128, 0.05),
+]
+
+# prefix, tcp?
+traffic_types = [
+	"tcp",
+#	"opus",
+#	"mix",
+]
+
+# prefix, single_learner
+single_learners = [
+#	("separate", False),
+	("single", True),
+]
+
+topols = [
+	"tree",
+#	"ecmp",
+]
+
+def total_per_model(model):
+	return len(discounts) * len(algos) * len(host_ps) * len(traffic_types) * (1 if model < 2 else len(single_learners)) * len(maths) * len(topols)
+
+if __name__ == "__main__":
+	# If <2 args, return the upper bound of expts per model.
+	if len(sys.argv) < 3:
+		print total_per_model(int(sys.argv[1]))
+		sys.exit(0)
+
+	# if 2 args (both int), choose that model (0/1) and then that sub-experiment
+	model = int(sys.argv[1])
+	experiment = int(sys.argv[2])
+
+	def expt_part(source, dependents=[]):
+		fold_len = 1
+		for el in dependents:
+			fold_len *= len(el)
+		return source[(experiment/fold_len) % len(source)]
+
+	params = {
+		#"P_good": 1.0,
+		"n_teams": 2,
+		"n_inters": 2,
+		"n_learners": 3,
+
+		"alpha": 0.05,
+		"epsilon": 0.2,
+		"discount": 0.0,
+		"dt": 0.05,
+
+		"ecmp_servers": 2,
+
+		"explore_episodes": 0.8,
+		"episodes": 5,
+		"episode_length": 10000,
+		"separate_episodes": True,
+
+		"rf": "marl",
[…]

### 2019-09-25 code/marl/tnsm-results-highdensitylowwidth.py

@@ -1,0 +1,230 @@
+import cPickle
+from marl import *
+import sys
+from writer import writeResults, makeResultsAverage, dumbWriter
+
+results_dir = "../../results/"
+
+# prefix, spiffy_mode, discount
+models = [
+#	"m", # MARL
+#	"spiffy", # Real SPIFFY
+#	"mpp", # MARL++
+	"spf", # SPF
+]
+
+# Use this to try out the effect of a sensible discount on Marl++...
+discounts = [
+	#0.0,
+	0.8,
+]
+maths = [
+	True,
+#	False,
+]
+
+algos = [
+	"sarsa",
+#	"q",
+]
+
+# host_p, dt
+host_ps = [
+#	(2, 0.05),
+#	(4, 0.05),
+	(16, 0.05),
+	(32, 0.05),
+	(64, 0.05),
+	(128, 0.05),
+	(192, 0.05),
+]
+
+# prefix, tcp?
+traffic_types = [
+	"tcp",
+#	"opus",
+#	"mix",
+]
+
+# prefix, single_learner
+single_learners = [
+#	("separate", False),
+	("single", True),
+]
+
+topols = [
+	"tree",
+#	"ecmp",
+]
+
+def total_per_model(model):
+	return len(discounts) * len(algos) * len(host_ps) * len(traffic_types) * (1 if model < 2 else len(single_learners)) * len(maths) * len(topols)
+
+if __name__ == "__main__":
+	# If <2 args, return the upper bound of expts per model.
+	if len(sys.argv) < 3:
+		print total_per_model(int(sys.argv[1]))
+		sys.exit(0)
+
+	# if 2 args (both int), choose that model (0/1) and then that sub-experiment
+	model = int(sys.argv[1])
+	experiment = int(sys.argv[2])
+
+	def expt_part(source, dependents=[]):
+		fold_len = 1
+		for el in dependents:
+			fold_len *= len(el)
+		return source[(experiment/fold_len) % len(source)]
+
+	params = {
+		#"P_good": 1.0,
+		"n_teams": 1,
+		"n_inters": 1,
+		"n_learners": 1,
+
+		"alpha": 0.05,
+		"epsilon": 0.2,
+		"discount": 0.0,
+		"dt": 0.05,
+
+		"ecmp_servers": 2,
+
+		"explore_episodes": 0.8,
+		"episodes": 5,
+		"episode_length": 10000,
+		"separate_episodes": True,
+
+		"rf": "marl",
[…]

### 2019-09-25 code/marl/tnsm-results-highrate.py

@@ -1,0 +1,238 @@
+import cPickle
+from marl import *
+import sys
+from writer import writeResults, makeResultsAverage, dumbWriter
+
+results_dir = "../../results/"
+
+# prefix, spiffy_mode, discount
+models = [
+#	"m", # MARL
+#	"spiffy", # Real SPIFFY
+#	"mpp", # MARL++
+	"spf", # SPF
+]
+
+# Use this to try out the effect of a sensible discount on Marl++...
+discounts = [
+	#0.0,
+	0.8,
+]
+maths = [
+	True,
+#	False,
+]
+
+algos = [
+	"sarsa",
+#	"q",
+]
+
+# host_p, dt
+host_ps = [
+#	(2, 0.05),
+#	(4, 0.05),
+#	(8, 0.05),
+	(16, 0.05),
+]
+
+# prefix, tcp?
+traffic_types = [
+	"tcp",
+#	"opus",
+#	"mix",
+]
+
+# prefix, single_learner
+single_learners = [
+#	("separate", False),
+	("single", True),
+]
+
+topols = [
+	"tree",
+#	"ecmp",
+]
+
+evil_factors = [
+	2.0,
+	2.5,
+	3.0,
+	3.5,
+]
+
+def total_per_model(model):
+	return len(evil_factors) * len(discounts) * len(algos) * len(host_ps) * len(traffic_types) * (1 if model < 2 else len(single_learners)) * len(maths) * len(topols)
+
+if __name__ == "__main__":
+	# If <2 args, return the upper bound of expts per model.
+	if len(sys.argv) < 3:
+		print total_per_model(int(sys.argv[1]))
+		sys.exit(0)
+
+	# if 2 args (both int), choose that model (0/1) and then that sub-experiment
+	model = int(sys.argv[1])
+	experiment = int(sys.argv[2])
+
+	def expt_part(source, dependents=[]):
+		fold_len = 1
+		for el in dependents:
+			fold_len *= len(el)
+		return source[(experiment/fold_len) % len(source)]
+
+	params = {
+		#"P_good": 1.0,
+		"n_teams": 2,
+		"n_inters": 2,
+		"n_learners": 3,
+
+		"alpha": 0.05,
+		"epsilon": 0.2,
+		"discount": 0.0,
+		"dt": 0.05,
+
+		"ecmp_servers": 2,
+
+		"explore_episodes": 0.8,
+		"episodes": 10,
+		"episode_length": 10000,
[…]

### 2019-09-25 paper/paper-revised.tex

@@
 \colorlet{highac}{inferno255}
+
+%\usepackage{framed}
+%\usepackage{adjustbox}
+%\colorlet{shadecolor}{uofgcobalt!20}
+\usepackage[color]{changebar}
+\setlength\changebarsep{10pt}
+\setlength\changebarwidth{6pt}
+\cbcolor{uofgcobalt}
 
@@
 %?? Discuss the most important conclusions before the outline.
-We find that the existing work for this task \cite{DBLP:journals/eaai/MalialisK15} fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
-The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
+We find that the existing work for this task \cite{DBLP:journals/eaai/MalialisK15} fails to account for congestion-aware traffic (i.e., TCP) and environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
+To remedy this, we make throttling decisions on a per-source basis and present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
 In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
@@
 
+\cbstart
 \subsection{Contributions}
 This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
-These are designed to make effective decisions irrespective of protocol, and to act on individual flows at the edge of network topologies with arbitrary shape and size.
-
-We offer an in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
-These features have been suggested by past studies into DDoS prevention, and have been independently tested within their own contexts.
+These are designed to make effective decisions irrespective of protocol, and act on individual flows at the edge of any network topology.
+We offer an in-depth investigation into suitable features for automatic DDoS mitigation, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
+These features have been suggested by past studies into DDoS prevention, and have been independently tested in their own contexts.
 Our study is the first attempt to quantify the individual efficacy of each in an RL-focussed setting.
@@
 Observing that past work has been compared only against AIMD \cite{DBLP:journals/ton/YauLLY05}, we additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS defences.
+\cbend
 
@@
 Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
-\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most exploitable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
+\emph{Amplification attacks} exploit services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most exploitable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
 Malicious hosts send many small requests, spoofed to appear as though they originated from the victim, causing many large replies to be sent to the intended target---significantly increasing a botnet's throughput while masking the identity of each participant.
@@
 There is immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
-In particular, we shall focus on methods which choose actions according to their value estimates from the current state: let $\operatorname{q}(s, a) \in \mathbb{R}$ denote the estimate of action $a$'s value if it were to be taken in state $s$.
+In particular, we focus on methods which choose actions according to their value estimates from the current state: let $\operatorname{q}(s, a) \in \mathbb{R}$ be the estimate of action $a$'s value if it were to be taken in state $s$.
 %?? Revisit the maths here, d, dim A, s...
@@
 		An example of tile coding for 2-dimensional state and 4 actions, using 2 tilings, 3 tiles per dimension, and a bias tile.
-		All components of $s$ are clamped to the range $[0,1]$.
+		All components of $s$ are clamped to $[0,1]$.
 		For simplicity, we denote $\bm{x}(s,\cdot)$ as a list of indices and represent the values of all actions at each tile with a vector.
@@
 	}
+	\vspace{-0.6cm}
 \end{figure}
@@
 
-Computing the approximate value of every action available in the current state forms the basis of a policy.
+Computing the approximate value of every available action forms the basis of a policy.
 Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might use some other mechanism.
@@
 	}
-
+\vspace{-0.25cm}
 \caption{
@@
 }
+\vspace{-0.6cm}
 \end{figure}
@@
 \subsection{Threat Model}
-An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
-Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
+An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to a host which connects to it, and they are expected to act rationally in its pursuit.
+Threat actors are external and act intentionally, aren't expected to be 
+\emph{advanced persistent threats}, and likely range from hacktivists to moderately funded adversaries.
 We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
-The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
+The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication with \nth{3} party organisations such as upstream ISPs.
 The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
 
-We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
-While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
-The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
+We do not assume that an attacker has white-box access to an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
+While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in volatile network traffic statistics falls outside of the scope of this work.
+The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/corr/abs-1902-09062}.
 These are APT-level capabilities, whose exploration presents a rich source for future work.
@@
 \section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
-Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
-However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
+Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to design agents such that filtering decisions are computed per flow.
+However, these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
 We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
@@
 \subsection{System Design and Assumptions}
-A deployment environment is defined as a network, with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
-These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
-Agents are co-located with each egress switch, and control the proportion of upstream packets from each external host to discard.
+A deployment environment is a network with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
+These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
+\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model or agent design, and control the proportion of upstream packets from each external host to discard.
 Each destination node $s$ has a maximum capacity, $U_s$.
+\cbend
 
@@
 
+\cbstart
+Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
+Only the ingress/egress nodes from a network need to be OpenFlow-enabled, because it is advantageous to perform filtering as close to a flow's source as possible.
+In a traditional network, each agent has exclusive control over its switch's tables.
+In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
+The main difference is that a traditional network needs this additional hardware, and does not allow for an operator to dynamically determine where the "edge" of their network lies through vNF relocation.
+\cbend
+
 \subsection{Algorithm}
-To make decisions cheaply and at low latency, we use \emph{semi-gradient Sarsa with tile coding} as described in \cref{eqn:sg-sarsa} and \cref{sec:reinforcement-learning}, rather than making use of neural networks or more complicated function approximators.
-Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
+To make decisions cheaply and at low latency, we use \emph{semi-gradient Sarsa with tile coding} as described in \cref{eqn:sg-sarsa} and \cref{sec:reinforcement-learning}, rather than using neural networks or more complicated function approximators.
+Exploration is introduced via $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
 Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their weight vector updates with one another (but may share experience and traces with one another).
@@
 \subsubsection{Action rate}
-We chose to adapt the algorithm to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to facilitate effective learning.
+We adapt the algorithm to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible for effective learning.
 To this end, we allow agents to make many decisions per timestep.
@@
 	}
+\vspace{-0.25cm}
 	\caption{
@@
 	}
+\vspace{-0.5cm}
 \end{figure}
@@
 The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
-In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
+\cbstart In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) the load measurement of a middle hop or the last hop. \cbend
 \Cref{fig:global-state-path} illustrates the process.
@@
 Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
+\cbstart Rather than having the network push the data to an agent, the agent requests this information about active flows periodically to isolate it from non-control-plane traffic and to eliminate the risk of resource exhaustion by excessive requests.
+\cbend
 
@@
 While our use and definition of $g(\cdot)$ appears nebulous, there are many ways to infer this quantity in practice.
-End-host servers may make use of canary flows or other active measurement schemes, or employ existing quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
+End-host servers may use canary flows or other active measurements, or employ existing quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
 ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
-Even if such heuristics or perfect knowledge aren't available in deployment, a sufficiently well-trained agent needs only to follow the policy it has learned from training in a greedy manner, allowing pre-training by a simulated environment (with perfect knowledge) to transfer to reality.
+Even if such heuristics or perfect knowledge aren't available in deployment, a sufficiently well-trained agent needs only to greedily follow the policy it has learned from training, allowing pre-training by a simulated environment (with perfect knowledge) to transfer to reality.
 
@@
 We choose to drop packets rather than impose traffic limits as it offers us a discrete action space without prior knowledge of traffic characteristics or measurement.
-Furthermore, we need not consider burstiness, fairness or implementation costs (such as per-flow bucket sizes) which could limit scalability.
+Furthermore, we need not consider burstiness, fairness or tuning (such as per-flow bucket sizes) which could limit scalability.
 We offer two models on how to choose $p$:
@@
 Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
-These choices primarily ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
-Since this model requires no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely opportunistic/myopic).
+These choices ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
+Since this model needs no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely myopic).
 
@@
 The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
-
 We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
@@
 
-\subsection{Systems Considerations}
+\subsection{Systems Considerations}\label{sec:systems-considerations}
 Taking many actions per timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
@@
 We expect this should be tuned based on the frequency at which statistics arrive.
+Naturally, this implies that an agent must carry work forward (and coalesce state updates) when the \emph{host density} $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/ftprep-cap-box}
-	\vspace{-1cm}
+	\includegraphics[width=0.65\linewidth]{../plots/ftprep-cap-box}
+	\vspace{-0.75cm}
 	\caption{
@@
 	}
+\vspace{-0.4cm}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/ftprep-tcp-cap-box}
-	\vspace{-1cm}
+	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-cap-box}
+	\vspace{-0.75cm}
 	\caption{
@@
 	}
+\vspace{-0.7cm}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/ftprep-tcp-laf-cap-box}
-	\vspace{-1cm}
+	\vspace{-0.75cm}
+	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-laf-cap-box}
+	\vspace{-0.75cm}
 	\caption{
@@
 
-The main element required to move to a per-source model is a feature set which offers high predictive power, such that behavioural differences are readily apparent to an agent.
+The main element required to move to a per-source model is a feature set which offers high predictive power, so that behavioural differences are apparent to an agent.
 Elaborating further on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and offer an investigation into their use alongside different traffic types:
@@
 The change in traffic rates caused by the last action.
-Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible in these fields.
+Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible here.
 
@@
 Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
-Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.
+Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other flooding attacks.
 
@@
 
+	\resizebox{0.45\linewidth}{!}{
 	\begin{tabular}{@{}ll@{}}
@@
 	\end{tabular}
+	}
+\vspace{-0.5cm}
 \end{table}
@@
 
-\section{Modelling Normal Traffic}\label{sec:a-new-normal}
+\section{Traffic Modelling}\label{sec:a-new-normal}
 
@@
 \subsection{Network Design}
-We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
+We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller based on \emph{Ryu} \cite{ryu}.
 All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
 If several ports offer different (equal-length) paths to a destination, a consistent random port is chosen from the flow-hash by an OpenFlow \emph{Group action} (in \emph{select} mode).
-If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
-The controller is also responsible for computing how switches respond to ARP requests: this need arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
+If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random outbound port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub ASes).
+The controller is also responsible for computing how switches respond to ARP requests: this need arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for trace-based evaluation.
 %We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.
 
-\subsection{TCP (HTTP) Traffic Model}
+\subsection{TCP (HTTP) Traffic Model}\label{sec:tcp-http-traffic-model}
 %?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
@@
 Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
-Hosts and clients use the TCP Cubic \cite{rfc8312} flavour of TCP.
+Hosts and clients both use TCP Cubic \cite{rfc8312}.
 Each host's download rate is limited to match whatever maximum bandwidth was assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.).
@@
 
-\subsection{UDP (Opus) Traffic Model}
-VoIP traffic exhibits very different characteristics to the above model; packet arrivals are highly periodic due to real-time application requirements, flows have a constant bitrate, and do not react substantially to lost packets.
+\subsection{UDP (Opus/VoIP) Traffic Model}\label{sec:udp-opus-traffic-model}
+VoIP traffic exhibits very different characteristics to the above model; packet arrivals are highly periodic due to real-time requirements, flows have a constant bitrate, and do not react substantially to lost packets.
 Interestingly, DDoS attack traffic is known to share many of these characteristics, offering an interesting detection problem.
-
-We present a VoIP traffic model\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward video game communities.
-We chose Discord as our prototype due to its publicly documented API, many open source bot frameworks for experimentation, large user base, and due to the lack of models for Opus-encoded traffic.
+We present a VoIP traffic model\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward gaming communities.
+We chose Discord as our prototype due to its publicly documented API, many open source bot frameworks, large user base, and due to the lack of models for Opus-encoded traffic.
 %?? Highly periodic, CBR
 
-Hosts send RTP traffic with Salsa20 encrypted payloads---audio frames of length \SI{20}{\milli\second} at a target \SI{96}{\kilo\bit\per\second}.
+Hosts send RTP traffic with Salsa20 encrypted payloads---\SI{20}{\milli\second} audio frames at \SI{96}{\kilo\bit\per\second}.
 We generate similar traffic at hosts by replaying anonymised traces gathered in general use and tabletop RPG servers; each trace contains only the size of each audio payload, entries denoting missed packets, and the duration of silent periods.
@@
 %?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
-\subsection{Attack Traffic Model}
+\cbstart
+\subsection{Attack Traffic Model}\label{sec:attack-traffic-model}
 Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
-Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that emulation remains tractable.
-
-\section{Evaluation}
+We configure each instance of \emph{hping3} to generate ethernet MTU-sized packets (\SI{1500}{\byte}) with a random source and destination port towards a target server, and configure the output rate $r$ (in \si{\mega\bit\per\second}) by setting the inter-arrival time $t_{\mathit{attack}}=\frac{1500 \cdot 8}{r\cdot10^6}$.
+This fulfils certain characteristics of many types of amplification DDoS traffic: it is congestion-unaware \cite{DBLP:conf/ndss/Rossow14}, and packets are larger than the minimum frame size and identically-sized (e.g., NTP amplification traffic is fragmented at the application layer into \SI{482}{\byte} chunks \cite{cisco-ntp-amp}).
+We differ from NTP amplification in frame size so that inter-arrival times are larger, to keep emulation of the network feasible at high traffic rates.
+\cbend
+
+\section{Evaluation}\label{sec:evaluation}
 
@@
 
-We compare our work most naturally against the MARL approach introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
+We compare our work most naturally against MARL, introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
 We are most interested in seeing how their approach contrasts with ours across different topologies and workloads.
 Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network---we aim to see how these characteristics affect performance and learning rate.
+\cbstart
 Marl is already known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} approach to DDoS mitigation, although the state of the art has long since moved on.
-
-To paint a more current picture, we compare our work against an effective modern-day algorithmic approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
-SPIFFY tests a proportion of flows by rerouting them through an alternate path with higher bandwidth, and observing how their speed has changed some time later.
+To paint a more current picture, we compare our work against an effective modern-day approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
+SPIFFY tests a proportion of flows by routing them through an alternate path with higher bandwidth, and observing how their speed has changed some time later.
 We compare ourselves against this approach to position our new RL agent designs against the modern state of the art, observing that SPIFFY has a similar mode of interaction to RL-based systems (taking action, observing an effect, and acting once again) and does not rely on protocol characteristics or signatures.
@@
 We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion factor), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\times$} bandwidth expansion from legitimate flows, making a judgement after \SI{5}{\second}.
+\cbend
 
@@
 In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
-Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes.
+Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes---a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign otherwise.
+Benign hosts generate traffic from either \cref{sec:tcp-http-traffic-model,sec:udp-opus-traffic-model} depending on the experiment, while malicious hosts generate traffic according to \cref{sec:attack-traffic-model} (both at experiment-dependent rates).
 
@@
 We then have $N_{\mathit{hosts}} = k \ell m n$.
-\Cref{fig:marl-topol} demonstrates this visually.
-
+\Cref{fig:marl-topol} demonstrates this.
 %Although \citeauthor{DBLP:journals/ccr/MahajanBFIPS02a}, the originators of this topology, make it clear that it exists as a fairly unrepresentative example \cite{DBLP:journals/ccr/MahajanBFIPS02a}, it remains the case that such a network topology allows for functional testing, and indeed is illustrative of one way in which attack traffic might aggregate in the network.
@@
 %We aim to address this through \cref{sec:performance-in-an-emulated-environment}.
-
-To test this, we configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
-This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
+We configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
+This is a slight simplification of \citeauthor{DBLP:journals/eaai/MalialisK15}'s \textquote{online} experiment \cite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
 %The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
@@
 
-Benign hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second}, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
+Benign hosts each occupied \SIrange{0}{1}{\mega\bit\per\second}, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
 %The original introduction of this approach to direct-control reinforcement learning as introduced by \textcite{DBLP:journals/eaai/MalialisK15} fails to consider key cases: the absence of a suitable heuristic classifier $g(\cdot)$, disjoint ranges of traffic distribution (i.e., the presence of benign heavy-hitters), the accurate simulation of TCP-like behaviour (and its effects on collateral damage), and high densities of hosts at egress points.
@@
 %?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
-Malicious hosts sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, though when using TCP-like traffic, each attacker's upload range had to be increased to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows.
+Malicious hosts each sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, though this was increased to \SIrange{4}{7}{\mega\bit\per\second} when using TCP-like traffic (to meaningfully impact benign flows).
 Given $n$ and $\operatorname{P}(\mathit{malicious})$, we see an expected malicious bandwidth \numrange{1.27}{1.87} and \SIrange{2.03}{2.18}{$\! \times U_s$} respectively.
 %The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
+\cbstart For our choices of $n$ in both topologies, we observe $N_{\mathit{hosts}} \in \left\{24, 48, 96, 192\right\}$, and an expected number of malicious hosts $\mathbb{E}\left[N_{\mathit{attackers}}\right] \in \left\{9.6, 19.2, 38.4, 76.8\right\}$.
+For the largest choice of $n$, we see an expected total attack traffic $\mathbb{E}\left[V_{\mathit{attack}}\right] = 334.05$ and \SI{422.4}{\mega\bit\per\second} for Opus and HTTP traffic respectively.
+\cbend
 
@@
 
+	\cbstart
 	\resizebox{0.8\linewidth}{!}{
@@
 	}
+	\cbend
+\vspace{-0.25cm}
 \end{table}
@@
 	\includegraphics[width=0.75\linewidth]{../plots/tnsm-udp-box-separate}
-
+	\vspace{-0.3cm}
 	\caption{
@@
 		\emph{Instant} outperforms Marl for $n \in \{4, 8\}$ (with higher variance), but performs similarly to Marl at $n\in \{2, 16\}$.
-		\emph{Guarded} always underperforms compared to the other agent designs in this problem variant.
+		\emph{Guarded} underperforms compared to the other agent designs in this problem variant.
 %		SPF, remarkably, slightly outperforms Marl++ (with lower variance) for this more difficult topology despite being worse for smaller $n$.
@@
 In a single-destination network, we observe that Marl's performance degrades as $n$ increases.
-Typically, our \emph{Instant} agent design achieves the best performance in multi-agent mode, reducing collateral damage above the current state-of-the-art, but sharply degrades at low $n$ when agents share experience.
-We see a reversal of this trend for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
+Typically, our \emph{Instant} agent design achieves the best performance in multi-agent mode, having lower collateral damage than the current state-of-the-art, but sharply degrades at low $n$ when agents share experience.
+This trend reverses for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
 %Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
 \Cref{fig:udp-tree-box} shows the preserved traffic in multi-agent mode.
-Interestingly, SPIFFY is unable to effectively protect constant bitrate legitimate traffic.
-
 %?? Discuss multi-dest topol once all results available.
 When defending multiple destinations, we see a sharp decrease in the effectiveness of all agent designs.
-Our \emph{Instant} and \emph{Guarded} designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
-SPIFFY remains ineffective at protecting this class of traffic.
+Our new agent designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
+\cbstart Interestingly, SPIFFY is unable to effectively protect constant bitrate traffic. \cbend
 
@@
 	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-box-single}
-
+	\vspace{-0.3cm}
 	\caption{
@@
 	}
+\vspace{-0.45cm}
 \end{figure}
@@
 	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-16-single}
-
+	\vspace{-0.3cm}
 	\caption{
-		Online performance of standard and single-agent models in a single-destination network with $n=16$ hosts per egress point and HTTP benign traffic.
+		Online performance of standard and single-agent models in a single-destination network with $n=16$ hosts per egress point, HTTP traffic.
 		At this level of host density, \emph{Guarded} reaches higher peak performance sooner and is considerably more consistent throughout the episode.
@@
 \Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
-Here, SPIFFY greatly outperforms all of the RL-based methods.
+\cbstart Here, SPIFFY greatly outperforms all of the RL-based methods. \cbend
 %SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
-
 %?? Discuss multi-dest topol once results available.
@@
 Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
-SPIFFY continues to outperform the RL-based methods.
+\cbstart SPIFFY continues to outperform the RL-based methods. \cbend
 
@@
 
+\cbstart
+\subsection{Increased Attack Volume}\label{sec:results-attack-volume}
+To assess the effect of larger volumes of attack traffic, we increase an attacker's output by various factors, supposing $n=16$ with HTTP traffic (\emph{Guarded}, Single); \cref{tab:atk-vol} records the expected rate of attack and average performance.
+We see that the initial increase in traffic volume causes the steepest reduction in performance (due to the increased cost of incorrect action), though performance levels out as attack traffic increases.
+\cbend
+
+\begin{table}
+\cbstart
+	\centering
+%	\vspace{-0.5cm}
+	\caption{Average reward versus attack volume.\label{tab:atk-vol}}
+\resizebox{0.45\linewidth}{!}{
+\begin{tabular}{@{}SSS@{}}
+	\toprule\multicolumn{1}{c}{Factor} & \multicolumn{1}{c}{$\mathbb{E}\left[V_{\mathit{attack}}\right]$ (\si{\mega\bit\per\second})} & \multicolumn{1}{c}{Reward} \\ \midrule
+	2.0 & 844.8 & 0.625 \\
+	2.5 & 1056.0 & 0.620 \\
+	3.0 & 1267.2 & 0.619 \\
+	3.5 & 1478.4 & 0.600 \\
+	\bottomrule
+\end{tabular}
+}
+\cbend
+\end{table}
+
 \subsection{Computational Cost}
@@
 %Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
-Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
+Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second} per flow.
 This is reassuring when measured alongside the insights from other work.
@@
 That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
+\cbstart As discussed in \cref{sec:systems-considerations}, we are able to judge 3 flows before this deadline: this difference is primarily accounted for by serialisation/communication delays and single-threaded processing in the Python language.
+\cbend
 %?? TODO: update with modern numbers...
@@
 Accordingly, we see that this design performs best (and exhibits considerably lower variance) when agents learn from as much knowledge as possible: high $n$ and single-agent training.
-To filter incoming traffic from a source, it must decide to degrade their inbound traffic multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
+To filter incoming traffic from a source, it must decide to degrade inbound traffic multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
 Our belief is that \emph{Guarded} is a considerably stronger model for these reasons, and its successes offer strong rationale to consider the best schemes for efficient information sharing.
@@
 The benefits we have at present are, however, substantial.
-What we offer above many of the alternate approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
+What we offer above many of the approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
 The techniques we propose offer substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today.
-Moreover, our decision making processes are entirely agnostic of the protocol or content of inbound traffic, offering future-proofing against the introduction of new transports.
+Moreover, our decision making processes are entirely agnostic of the protocol or content of traffic, offering future-proofing against the introduction of new transports.
 
@@
 Can an agent be flooded with new flows to reduce their ability to make decisions?
-One of the risks introduced by our timed random sequential policy updates is that so much work can be queued up that an agent is never able to act on some attack flows.
+One of the risks introduced by our policy update strategy is that so much work can be queued up that an agent is never able to act on some attack flows.
 The natural solution is to impose an upper bound on the amount of action computations/policy updates that can be performed before a work list is discarded completely.
@@
 
+\cbstart
+Can an attack on the controller can impact the effectiveness of our approach?
+This question hinges upon whether the deployment environment is a traditional network or is fully SDN-enabled---each agent is, in a sense, \emph{a} controller alongside the network's controller.
+In a traditional network, only the agents act as controllers, but since agents periodically request per-flow data (rather than continuously receiving it) no amount of flows generates more requests or messages to the agent.
+More work is generated, but we discuss how to handle this safely above.
+Accordingly, the agent can never be stalled by request volume: the only communication with elsewhere in the network (load measurements) comes from trusted nodes, is highly periodic, and has constant size.
+The same logic holds for a fully software-defined network.
+Recalling that we do not employ the network's controller to install filtering rules on edge switches, an agent's ability to act is unimpeded.
+Thus, the controller is made no more vulnerable than in any other SDN.
+The only necessary change for such a scenario is that a load measurement which has not been updated (due to a timeout or missed deadline) should be set at $R_t=-1$.
+\cbend
 %\begin{figure}
@@
 Action computation for \emph{Guarded} agents is cheaper still, requiring only \num{48} additions per action.
-Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
+Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this.
 We believe that it may be possible to implement similar behaviour on standard commodity switches through application of \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.
@@
 %?? Mention: probably fewer state measurements than in our emulations, but longer measurements means less noisy (so probably more accurate)
-Gathering and transmission of load/flow statistics would be difficult to perform quite as often as an emulated environment allows, without inadvertently affecting host traffic.
-However, the measurements acquired in such a scenario are far more likely to be less noisy (by being collected over longer periods of time), which could aid effective training.
-The main bottlenecks are most likely in forwarding the load measurements from various aggregation points (which can be made more efficient through multicast) and in running some estimator $\operatorname{g}(\cdot)$ to condition the reward function.
+Gathering and transmission of load/flow statistics would be difficult to perform as often as an emulated environment allows, without inadvertently affecting host traffic.
+However, the measurements acquired in such a scenario are likely to be less noisy (by being collected over longer periods of time), which could aid training.
+The main bottlenecks are likely in forwarding the load measurements from various aggregation points (which can be made more efficient through multicast) and in running some estimator $\operatorname{g}(\cdot)$ to condition the reward function.
 %?? Potential for dividing up pre-feature training across different agents, or train like this locally (sets of flows trained by 1 sub-model).
@@
 
+\cbstart
+A limitation of SDN-capable hardware is that OpenFlow table entries occupy \SI{6}{$\!
+\times$} as much space as traditional routing rules---the cost and power demands of TCAM mean that commercial switches only have space for \SIrange{2}{20}{\kilo} rules \cite{DBLP:journals/comsur/NguyenSBT16}.
+Our approach consumes an amount of rules equal to the number of active flows (i.e., the host density), and by the end of an experiment a switch may have accrued around ?? rules.
+While we use a default action as a fallback to maintain connectivity, eviction of high-value decisions (those which filter the highest-bandwidth attackers) poses a significant risk.
+Given that most flows are small (with the majority of bytes coming from a few large ``heavy-hitters'') \cite{DBLP:journals/ccr/PanBPS03}, it may suffice to only apply RL-based analysis to larger flows.
+%?? What could we do in the meantime? I propose manually assigning elephant flows a high importance rule, while increasing the importance of "block" decisions even higher.
+OpenFlow rules have an \emph{importance}, controlling which rules may be evicted by a new entry (preventing entries from evicting those with higher importance).
+If an agent is to act on all flows, a solution is to assign an importance of 0 to mice flows, 1 to elephant flows, and 2 to total filtering (leaving agents to time out and remove elephant flow rules to prevent bloat).
+Given the high churn and prevalence of mice flows, eviction here is most likely to affect flows which no longer exist.
+In both cases, extra rules can be made available by upgrading rules which completely filter a flow into an upstream blackholing request (as used in collaborative DDoS approaches \cite{DBLP:conf/acsac/RamanathanMYZ18}), having the agent remove this rule once blackholing is active.
+\cbend
+
 \section{Related Work}\label{sec:related-work}
@@
 Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
-
 %?? Actually talk about Marl (???) to appease reviewer \#1.
@@
 A weakness of their approach is that computing a route to measure bandwidth expansion on real networks can be costly (up to \SI{14}{\second} for the Cogent topology), and that the low expansion factors in real network can require more ``rounds'' of filtering.
-Their assumptions about traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), both of which make up a sizeable proportion of network traffic.
-
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
+\cbstart Their assumptions about traffic response to such bandwidth expansion do not hold for constant bitrate flows (e.g., VoIP) and may not extend to HTTP DASH flows, both of which make up a sizeable proportion of network traffic.
+\cbend
+
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
 Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
@@
 For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
-A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
+A \emph{direct-control} RL problem is one where the RL agent(s) learn optimal control over a set of actions as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
 %We feel there is a shortage of work in this category at present, at least in the field of networks.
@@
 \section{Conclusions and Future Work}
-Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
+Through this paper, we have discussed reinforcement learning and its relevance to network intrusion prevention.
 Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.
-
 We have identified weaknesses in past approaches arising from simulation assumptions, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
@@
 \section*{Acknowledgements}
-Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice, comments and technical assistance.
+Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice and technical assistance, and to our anonymous reviewers.
 %Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
-%Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
 %This work was supported by a nameless funder [grant number 0].
-This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].
+This work was supported in part by the Engineering and Physical Sciences Research Council [grants EP/M508056/1, EP/N033957/1 and EP/P004024/1] and European Cooperation in Science and Technology (COST) Action CA15127: RECODIS---Resilient communication and services.
 
@@
 \printbibliography
-\vspace{-0.5cm}
+\vspace{-1cm}
 \vskip 0pt plus -1fil
@@
 \end{IEEEbiography}
-\vspace{-0.5cm}
+\vspace{-1cm}
 \vskip 0pt plus -1fil
@@
 
-	Dr.\ Simon Rogers’s research involves the development of machine learning and statistical techniques to help with the analysis of complex datasets, particularly within the field of metabolomics but also other fields, human-computer interaction and information retrieval.
+	Dr.\ Simon Rogers’s research involves the development of machine learning and statistical techniques to help with the analysis of complex datasets, particularly within the field of metabolomics, human-computer interaction and information retrieval.
 \end{IEEEbiography}
-\vspace{-0.5cm}
+\vspace{-1cm}
 \vskip 0pt plus -1fil
@@
 degrees in computer science from Lancaster University.
-He is currently a Senior Lecturer (Associate Professor) and the Founding Director of the Networked Systems Research Laboratory at the School of Computing Science, University of Glasgow.
+He is currently a Professor and the Founding Director of the Networked Systems Research Laboratory at the School of Computing Science, University of Glasgow.
 He is also a visiting Professor at the University of Athens, Department of Informatics and Telecommunications.
 
-	Dr.\ Pezaros has published widely (over 80 peer-reviewed contributions to top-tier international conferences and journals) in the areas of computer communications, network and service management, and resilience of future networked infrastructures, and has received significant funding for his research in the above areas from public funding agencies and the industry.
+	Professor Pezaros has published widely (over 80 peer-reviewed contributions to top-tier international conferences and journals) in the areas of computer communications, network and service management, and resilience of future networked infrastructures, and has received significant funding for his research in the above areas from public funding agencies and the industry.
 He is a Chartered Engineer, and a senior member of the IEEE and the ACM.

### 2019-09-25 paper/paper-revised.tex
Almost there...

@@
 
+\subsection{Contributions}
 \cbstart
-\subsection{Contributions}
 This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
@@
 We offer an in-depth investigation into suitable features for automatic DDoS mitigation, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
-These features have been suggested by past studies into DDoS prevention, and have been independently tested in their own contexts.
-Our study is the first attempt to quantify the individual efficacy of each in an RL-focussed setting.
+These features have been suggested by past studies, and independently tested in their own contexts.
+Our study is the first attempt to quantify the individual efficacy of each in an RL setting.
 
@@
 These new traffic models inform an empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
-Observing that past work has been compared only against AIMD \cite{DBLP:journals/ton/YauLLY05}, we additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS defences.
-\cbend
+We additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS \cbend defences.
 
@@
 	\centering
-	\resizebox{0.5\linewidth}{!}{
+	\resizebox{0.45\linewidth}{!}{
 		\begin{tikzpicture}		
@@
 \subsection{Threat Model}
-An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to a host which connects to it, and they are expected to act rationally in its pursuit.
+An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to hosts, and they are expected to act rationally in its pursuit.
 Threat actors are external and act intentionally, aren't expected to be 
 \emph{advanced persistent threats}, and likely range from hacktivists to moderately funded adversaries.
-We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
+We assume that attacks will be volumetric DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
 The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication with \nth{3} party organisations such as upstream ISPs.
@@
 These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
-\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model or agent design, and control the proportion of upstream packets from each external host to discard.
-Each destination node $s$ has a maximum capacity, $U_s$.
-\cbend
+\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design, and control the proportion of upstream packets from each external host to discard.
+Each destination node $s$ has a maximum capacity, \cbend $U_s$.
 
@@
 Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
-Only the ingress/egress nodes from a network need to be OpenFlow-enabled, because it is advantageous to perform filtering as close to a flow's source as possible.
+Only the ingress/egress nodes from a network need to be OpenFlow-enabled, as it is advantageous to perform filtering as close to a flow's source as possible.
 In a traditional network, each agent has exclusive control over its switch's tables.
 In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
-The main difference is that a traditional network needs this additional hardware, and does not allow for an operator to dynamically determine where the "edge" of their network lies through vNF relocation.
+The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the "edge" of their network lies through vNF relocation.
 \cbend
@@
 Although the choice of a classical RL method likely brings lower theoretical performance, there are significant reasons to favour such methods; these include lower latency decision-making, lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
-This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
-Simpler decision boundaries reduce the risk of overfitting and unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.
+This aligns with our goal of quick online learning, and faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
+Simpler decision boundaries reduce the risk of overfitting and unexpected behaviour, and we expect that the simplicity of tile-based policy computation will also greatly aid interpretability of anomalous action choices.
 
@@
 To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
-State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
+State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields.
 Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
 We expect this should be tuned based on the frequency at which statistics arrive.
+\cbstart
 Naturally, this implies that an agent must carry work forward (and coalesce state updates) when the \emph{host density} $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.
+\cbend
 
@@
 	\centering
-	\vspace{-0.75cm}
+%	\vspace{-0.25cm}
 	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-laf-cap-box}
@@
 	}
+\vspace{-0.6cm}
 \end{figure}
@@
 %?? Reflective factor among \emph{authenticated hosts}.
-
 We find that each flow occupies an expected \SI{52.4}{\kilo\bit\per\second} upstream bandwidth.
@@
 %?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
+\subsection{Attack Traffic Model}\label{sec:attack-traffic-model}
 \cbstart
-\subsection{Attack Traffic Model}\label{sec:attack-traffic-model}
 Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
@@
 We make the simplifying assumption that a suitable unused path exists (with identical bandwidth to the server's link).
-We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion factor), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\times$} bandwidth expansion from legitimate flows, making a judgement after \SI{5}{\second}.
-\cbend
+We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion factor), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\times$} bandwidth expansion from legitimate flows, making a judgement after \cbend \SI{5}{\second}.
 
@@
 %Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
-%$U_s$ was fixed at $k \ell mn+2$ \si{\mega\bit\per\second}.
+%$U_s$  $k \ell mn+2$ \si{\mega\bit\per\second}.
 %The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
@@
 	}
+\vspace{-0.75cm}
 \end{figure}
@@
 In contrast, the fat-tree topology \cite{DBLP:conf/sigcomm/Al-FaresLV08} sees regular use in real-world datacentres and scales well horizontally.
-
 %?? Come up with description of fat-tree (multi-dest) topol.
@@
 $U_s$ was fixed at $N_{\mathit{hosts}}+2$ \si{\mega\bit\per\second} (to account for burstiness), and each link had a delay of \SI{10}{\milli\second}.
-All links had unbounded capacity, save for server-switch connections.
-These parameter choices match those of the original study to enable direct comparison, and many are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
+All links had unbounded capacity, save for each server-switch.
+These parameters match those of the original study to enable direct comparison, and many are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
 
@@
 
-We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
+We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
 %Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
-We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold and underlining the best RL-based model.
+We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold and the best RL-based model underlined.
 Average rewards take into account any portions of time that an agent allows illegal system states.
@@
 	\centering
+	\cbstart
 	\includegraphics[width=0.75\linewidth]{../plots/tnsm-udp-box-separate}
+	\cbend
 	\vspace{-0.3cm}
@@
 \begin{figure}
+%	\vspace{-0.5cm}
 	\centering
+	\cbstart
 	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-box-single}
+	\cbend
 	\vspace{-0.3cm}
@@
 	}
-\vspace{-0.45cm}
+\vspace{-0.5cm}
 \end{figure}
@@
 	\centering
+	\vspace{-0.6cm}
 	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-16-single}
@@
 \Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$\!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
-\cbstart Here, SPIFFY greatly outperforms all of the RL-based methods. \cbend
 %SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
@@
 Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
-\cbstart SPIFFY continues to outperform the RL-based methods. \cbend
+In both cases, SPIFFY greatly outperforms the RL-based methods.
 
@@
 
+\subsection{Increased Attack Volume}\label{sec:results-attack-volume}
 \cbstart
-\subsection{Increased Attack Volume}\label{sec:results-attack-volume}
 To assess the effect of larger volumes of attack traffic, we increase an attacker's output by various factors, supposing $n=16$ with HTTP traffic (\emph{Guarded}, Single); \cref{tab:atk-vol} records the expected rate of attack and average performance.
-We see that the initial increase in traffic volume causes the steepest reduction in performance (due to the increased cost of incorrect action), though performance levels out as attack traffic increases.
-\cbend
+The initial increase in traffic volume causes the steepest reduction in performance (due to the increased cost of incorrect action), though performance levels out as attack traffic \cbend increases.
 
@@
 	\centering
-%	\vspace{-0.5cm}
+%	\vspace{-0.1cm}
 	\caption{Average reward versus attack volume.\label{tab:atk-vol}}
@@
 \cbend
+\vspace{-0.75cm}
 \end{table}
@@
 %Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
-Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second} per flow.
+Measurements from each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second} per flow.
 This is reassuring when measured alongside the insights from other work.
@@
 That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
-\cbstart As discussed in \cref{sec:systems-considerations}, we are able to judge 3 flows before this deadline: this difference is primarily accounted for by serialisation/communication delays and single-threaded processing in the Python language.
+\cbstart
+This cost is constant and independent of network size.
+As discussed in \cref{sec:systems-considerations}, we are able to judge 3 flows before this deadline: the difference is primarily accounted for by serialisation/communication delays and single-threaded processing in the Python language.
 \cbend
@@
 %?? May be hard to learn multiple features at once while controlling multiple flows while contending with many more agents, with harder dynamics like TCP. Does this hinder learning in the long run?
-Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect in the case of UDP traffic.
+Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect for UDP traffic.
 %??
@@
 The most likely explanation is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
-The increased state space size makes this a more likely occurrence, as is the unclear affect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state space.
+The increased state space size makes this a more likely occurrence, as does the unclear effect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state space.
 We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to observe the true value of each action), and controlling shared global state.
@@
 %Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
+\cbstart
 Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this strand still trails behind existing (exact) DDoS flow detection mechanisms where TCP traffic is concerned.
 The ability to better protect VoIP traffic when compared against one of these approaches is a curious observation, which suggests that other (exact) protocol-agnostic approaches may carry hidden assumptions and is a promising direction for future investigation.
-Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
+Similar traffic makes up a significant fraction of network load today (\SIrange{18}{27}{\percent}).
+Although we have conducted work to map the territory, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
 The benefits we have at present are, however, substantial.
-What we offer above many of the approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
-The techniques we propose offer substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today.
+What we offer above many of the approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments, low-overhead and fixed-cost decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
 Moreover, our decision making processes are entirely agnostic of the protocol or content of traffic, offering future-proofing against the introduction of new transports.
+\cbend
 
@@
 \cbstart
-Can an attack on the controller can impact the effectiveness of our approach?
+Can an attack on the controller can impact our approach?
 This question hinges upon whether the deployment environment is a traditional network or is fully SDN-enabled---each agent is, in a sense, \emph{a} controller alongside the network's controller.
-In a traditional network, only the agents act as controllers, but since agents periodically request per-flow data (rather than continuously receiving it) no amount of flows generates more requests or messages to the agent.
+In a traditional network, only the agents act as controllers, but since they periodically request per-flow data (rather than continuously receiving it) no amount of flows generates more requests or messages to the agent.
 More work is generated, but we discuss how to handle this safely above.
-Accordingly, the agent can never be stalled by request volume: the only communication with elsewhere in the network (load measurements) comes from trusted nodes, is highly periodic, and has constant size.
+Accordingly, agents can never be stalled by request volume: their only remote communication (load measurements) comes from trusted nodes, is highly periodic, and has constant size.
 The same logic holds for a fully software-defined network.
@@
 \cbstart
-A limitation of SDN-capable hardware is that OpenFlow table entries occupy \SI{6}{$\!
-\times$} as much space as traditional routing rules---the cost and power demands of TCAM mean that commercial switches only have space for \SIrange{2}{20}{\kilo} rules \cite{DBLP:journals/comsur/NguyenSBT16}.
-Our approach consumes an amount of rules equal to the number of active flows (i.e., the host density), and by the end of an experiment a switch may have accrued around ?? rules.
+One limit of SDN-capable hardware is that OpenFlow rules occupy \SI{6}{$\!
+\times$} the space of traditional rules---the cost of TCAM mean that commercial switches only have space for \SIrange{2}{20}{\kilo} rules \cite{DBLP:journals/comsur/NguyenSBT16}.
+Our approach consumes a rule for each active flow (i.e., the host density), and by the end of an experiment a switch may have accrued around \num{900} rules.
 While we use a default action as a fallback to maintain connectivity, eviction of high-value decisions (those which filter the highest-bandwidth attackers) poses a significant risk.
@@
 If an agent is to act on all flows, a solution is to assign an importance of 0 to mice flows, 1 to elephant flows, and 2 to total filtering (leaving agents to time out and remove elephant flow rules to prevent bloat).
-Given the high churn and prevalence of mice flows, eviction here is most likely to affect flows which no longer exist.
-In both cases, extra rules can be made available by upgrading rules which completely filter a flow into an upstream blackholing request (as used in collaborative DDoS approaches \cite{DBLP:conf/acsac/RamanathanMYZ18}), having the agent remove this rule once blackholing is active.
+Given the high churn and prevalence of mice flows, eviction here is most likely to affect flows which are complete.
+In both cases, extra rules can be made available by upgrading rules which completely filter a flow into an upstream blackholing request (as in collaborative approaches \cite{DBLP:conf/acsac/RamanathanMYZ18}), having the agent remove this rule once blackholing is active.
 \cbend
@@
 \fakepara{DDoS Prevention}
-\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
+\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of flooding DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
 Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
@@
 %Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
-Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ASes, although some of our feature choices are backed by similar observations.
+%Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ASes, although .
 A weakness of their approach is that computing a route to measure bandwidth expansion on real networks can be costly (up to \SI{14}{\second} for the Cogent topology), and that the low expansion factors in real network can require more ``rounds'' of filtering.
-\cbstart Their assumptions about traffic response to such bandwidth expansion do not hold for constant bitrate flows (e.g., VoIP) and may not extend to HTTP DASH flows, both of which make up a sizeable proportion of network traffic.
+\cbstart
+By contrast, our approach takes a constant time to compute an action for a flow regardless of topology size.
+Their assumptions about traffic response to such bandwidth expansion do not hold for constant bitrate flows (e.g., VoIP) and may not extend to HTTP DASH flows, both of which make up a sizeable proportion of network traffic.
 \cbend
@@
 
-\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
-This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
+\Textcite{DBLP:conf/sp/SmithS18} use AS-level routing to tackle both transit-link and flooding-based attacks.
+This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples.
 The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
-Because of this, traditional flooding-based attacks aren't fully alleviated.
+Because of this, flooding-based attacks aren't fully alleviated.
 
@@
 Through this paper, we have discussed reinforcement learning and its relevance to network intrusion prevention.
-Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.
-We have identified weaknesses in past approaches arising from simulation assumptions, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
+Our belief is that the potential to learn feedback-loop-like control online and against non-stationarity makes it particularly suited to the problems endemic to the field.
+We have identified weaknesses in past approaches, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
 Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
@@
 THEN: Room to introduce/roll-in dynamic changepoint detection or adaptive exploration \cite{DBLP:conf/ki/Tokic10, DBLP:conf/ki/TokicP11, DBLP:conf/annpr/TokicP12}?
-Given that one of the advantages that RL methods can offer is the ability to handle non-stationary problems, it is important to propose and test sensible simulations or captures of evolving networks against these methods.
+Given that one of the advantages of RL methods is the ability to handle non-stationary problems, it is important to propose and test sensible simulations or captures of evolving networks.
 While it is known that DDoS attack strategies evolve in real time \cite{DBLP:conf/spw/KangGS16}, evaluation is difficult at present since no works detail what patterns such evolution might take.
-Regardless, these scenarios present ideal circumstances to apply work on adaptive exploration \cite{DBLP:conf/annpr/TokicP12}, changepoint detection, or intelligent/heuristic sampling methods to judge which flows are most worthy of consideration per tick.
+Regardless, these scenarios present ideal circumstances to apply adaptive exploration \cite{DBLP:conf/annpr/TokicP12}, changepoint detection, or intelligent sampling methods to judge which flows are most worthy of consideration.
 For estimating \emph{when} to explore, we believe that the intersection of signal processing and RL is as-yet unexplored.
@@
 %While this work still trails behind the performance of exact DDoS flow detection mechanisms, w
-We hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
-It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques rely upon.
-It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
+We hope it is clear that reinforcement learning holds promise and can inspire further innovation.
+It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the
+network resources or capabilities that other techniques rely upon.
+It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
 
@@
 \printbibliography
-\vspace{-1cm}
+\vspace{-4.5cm}
 \vskip 0pt plus -1fil
@@
 \end{IEEEbiography}
-\vspace{-1cm}
+\vspace{-1.1cm}
 \vskip 0pt plus -1fil

### 2019-09-26 paper/paper-revised.tex
Barely making the page count.

@@
 To quite a large extent, this remains the case today.
-These authors posit that their application is made difficult due to significant operational differences from standard ML tasks, including: the high cost of errors and extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}; a general lack of recent, openly available (and high-quality) training data; and diversity of network traffic across varying timescales combined with significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
-Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
-Learning is made harder still by the challenges encountered when learning from unlabelled (often partial) data.
+They posit that their use is made difficult due to significant operational differences from standard ML tasks, including: the high cost of errors and extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}; a general lack of recent, openly available (and high-quality) training data; and diversity of network traffic across varying timescales combined with significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
+Above the aggregate level, the constant deployment of new services and protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
+Learning is made harder still by the challenges encountered with unlabelled (often partial) data.
 All of these factors greatly inflate the difficulty of the detection problem.
@@
 \subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
-\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
-We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$, it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
-This model of system interaction is best described as a \emph{Markov Decision Process} (MDP).
+\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with training an agent to choose an optimal sequence of actions in pursuit of a given task \cite{RL2E}.
+We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$, it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
+This model of system interaction is a \emph{Markov Decision Process} (MDP).
 RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal if an agent needs to further explore some region of the state space.
-The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
+The policy is refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
 In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new, custom-built training corpus.
@@
 
-The main element required to move to a per-source model is a feature set which offers high predictive power, so that behavioural differences are apparent to an agent.
-Elaborating further on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and offer an investigation into their use alongside different traffic types:
+The main element required by a per-source model is a feature set with high predictive power, so that behavioural differences are apparent to an agent.
+Elaborating on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and investigate their use alongside different traffic types:
 %?? We use these features, and why...
@@
 \fakepara{Source IP address}
-While ordinarily trivial to spoof (and thus of limited use for many classes of attack), reflectors are themselves legitimate services being abused by spoofing attackers.
+While trivial to spoof (and thus of limited use for many classes of attack), reflectors are themselves legitimate services being abused by spoofing attackers.
 As a result, they communicate with attack victims using their own IP address.
-In real-world scenarios the addresses of a set of reflector nodes might exhibit similarity due to network uncleanliness \cite{DBLP:conf/imc/CollinsSFJWSK07}, e.g., similarly unhardened services exposed by a single organisation.
+In real-world scenarios the addresses of reflector nodes might exhibit similarity due to network uncleanliness \cite{DBLP:conf/imc/CollinsSFJWSK07}, e.g., unhardened services exposed by a single organisation.
 
@@
 All of the above features, save for global state, are 1-dimensional.
-\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), using a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
-\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends being observed when the same technique is applied with UDP traffic (omitted).
+\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), on a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 runs.
+\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends observed when applied to UDP traffic (omitted).
 %?? Core findings---different protocols need different features, so everything we proposed above has a use!
-The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
+The plots show that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
 All features converge to their highest-observed performance within around \num{4000} timesteps.
@@
 \cbstart
-Marl is already known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} approach to DDoS mitigation, although the state of the art has long since moved on.
-To paint a more current picture, we compare our work against an effective modern-day approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
-SPIFFY tests a proportion of flows by routing them through an alternate path with higher bandwidth, and observing how their speed has changed some time later.
-We compare ourselves against this approach to position our new RL agent designs against the modern state of the art, observing that SPIFFY has a similar mode of interaction to RL-based systems (taking action, observing an effect, and acting once again) and does not rely on protocol characteristics or signatures.
+Marl is known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} strategy, yet the state of the art has long since moved on.
+To paint a more current picture, we compare our work against an effective modern approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
+SPIFFY tests a proportion of flows by routing them through an alternate path with higher bandwidth, observing how their speed changes some time later.
+This comparison lets us position our new agent designs against the state of the art, observing that SPIFFY has a similar mode of interaction to RL-based systems (taking action, observing an effect, and acting once again) and does not rely on protocol characteristics or signatures.
 We make the simplifying assumption that a suitable unused path exists (with identical bandwidth to the server's link).
-We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion factor), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\times$} bandwidth expansion from legitimate flows, making a judgement after \cbend \SI{5}{\second}.
+We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\!
+\times$} expansion from legitimate flows, making a judgement after \cbend \SI{5}{\second}.
 
@@
 	\cbend
-\vspace{-0.25cm}
+\vspace{-0.5cm}
 \end{table}
@@
 		Online performance for HTTP benign traffic in a single-destination network, single-agent mode.
-		\emph{Instant} and \emph{Guarded} exhibit essentially identical efficacy at $n~=~2$, protecting less traffic than Marl.
-		Only \emph{Guarded}'s performance rapidly increases with $n$, achieving considerably better median performance and lower variance than the other models.
+		\emph{Instant} and \emph{Guarded} exhibit similar efficacy at $n=2$, protecting less traffic than Marl.
+		Only \emph{Guarded}'s performance rapidly increases with $n$, achieving a considerably better median and lower variance than the other models.
 		The longer tails of outliers typically indicate the longer training time the new models require---we observe that \emph{Guarded} typically has considerably lower variance once it has converged on a stable policy.
@@
 One limit of SDN-capable hardware is that OpenFlow rules occupy \SI{6}{$\!
-\times$} the space of traditional rules---the cost of TCAM mean that commercial switches only have space for \SIrange{2}{20}{\kilo} rules \cite{DBLP:journals/comsur/NguyenSBT16}.
-Our approach consumes a rule for each active flow (i.e., the host density), and by the end of an experiment a switch may have accrued around \num{900} rules.
-While we use a default action as a fallback to maintain connectivity, eviction of high-value decisions (those which filter the highest-bandwidth attackers) poses a significant risk.
-Given that most flows are small (with the majority of bytes coming from a few large ``heavy-hitters'') \cite{DBLP:journals/ccr/PanBPS03}, it may suffice to only apply RL-based analysis to larger flows.
+\times$} the space of standard rules---commercial switches only have TCAM space for \SIrange{2}{20}{\kilo} rules \cite{DBLP:journals/comsur/NguyenSBT16}.
+Our approach consumes a rule for each active flow (the host density), and by the end of an experiment a switch can accrue around \num{900} rules.
+While we use a default fallback action to maintain connectivity, eviction of high-value decisions which filter high-bandwidth attackers poses a significant risk.
+Given that most flows are small (with the majority of bytes coming from a few ``heavy-hitters'') \cite{DBLP:journals/ccr/PanBPS03}, it may suffice to only apply RL-based analysis to larger flows.
 %?? What could we do in the meantime? I propose manually assigning elephant flows a high importance rule, while increasing the importance of "block" decisions even higher.
@@
 Given the high churn and prevalence of mice flows, eviction here is most likely to affect flows which are complete.
-In both cases, extra rules can be made available by upgrading rules which completely filter a flow into an upstream blackholing request (as in collaborative approaches \cite{DBLP:conf/acsac/RamanathanMYZ18}), having the agent remove this rule once blackholing is active.
+In both cases, extra rules can be made available by upgrading rules which completely filter a flow into upstream blackholing (as in collaborative approaches \cite{DBLP:conf/acsac/RamanathanMYZ18}), having the agent remove this rule once blackholing is active.
 \cbend
@@
 \fakepara{DDoS Prevention}
-\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of flooding DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
-Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
+\Textcite{DBLP:conf/lcn/BragaMP10} examine the detection of flooding DDoS attacks through \emph{self-organising maps}, using SDN to gather statistics effectively.
+Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts contribute to an attack.
 %?? Actually talk about Marl (???) to appease reviewer \#1.
@@
 \emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
-\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
+\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion (having already saturated the capacity of their outbound links), while legitimate flows typically speed up to match the new fair-share rate.
 %Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
@@
 \fakepara{RL in Networks}
-Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
-Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-Given that one of the main strengths of RL techniques is the ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works don't apply the rich literature on the subject to its fullest potential.
+Earnest, well-considered application of RL towards the challenge of intrusion prevention has seen comparatively little examination.
+Past work treats the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
+Given that the main strengths of RL techniques are the ability to control ongoing interaction and adapt by observing the concrete effects of actions, such works don't apply the rich literature on the subject to its fullest potential.
 
@@
 
-An \emph{indirect-control} RL problem is one where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
+An \emph{indirect-control} RL problem is one where agents act in service to \emph{another technique} responsible for decision-making, optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
 A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
 %The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
+This work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated.
+Outside of intrusion detection, there has been growing interest in the use of RL in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
 
@@
 Through this paper, we have discussed reinforcement learning and its relevance to network intrusion prevention.
-Our belief is that the potential to learn feedback-loop-like control online and against non-stationarity makes it particularly suited to the problems endemic to the field.
-We have identified weaknesses in past approaches, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
-Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
+We believe the potential to learn feedback loop-like control online and against non-stationarity makes it particularly suited to the problems endemic to the field.
+We identified weaknesses in past work, recommending an RL agent which acts per flow, and have outlined the algorithmic and engineering choices needed to make its deployment feasible.
+Supporting this, we've presented an in-depth examination of our feature space, offering quantitative and qualitative justification for our choices.
 Our evaluation shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, with \emph{Guarded} agents showing the most promise for future evaluation.
@@
 \printbibliography
-\vspace{-4.5cm}
-\vskip 0pt plus -1fil
+\enlargethispage{1cm}
+\vspace{-1.1cm}
+\vskip -6pt plus -1fil
 \begin{IEEEbiography}
@@
 \end{IEEEbiography}
+\vskip 0pt plus -1fil
 

### 2019-09-26 paper/paper-revised.tex
Actually make the page count.

@@
 
-For certain classes of problem e.g., volume-based \emph{distributed denial of service} (DDoS) attacks, \emph{reinforcement learning} (RL) offers another perspective.
+For certain classes of problem e.g., volumetric \emph{distributed denial of service} (DDoS) attacks, \emph{reinforcement learning} (RL) offers another perspective.
 %?? Unclear explanation of RL here?
@@
 %It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
-This weakness still presents itself in modern flavours of TCP, such as TCP Cubic which in turn has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
+This weakness is still present in modern TCP flavours, such as TCP Cubic which in turn has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
 This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
@@
 	}
-\vspace{-0.5cm}
+\vspace{-0.6cm}
 \end{figure}
@@
 	\centering
-	\vspace{-0.6cm}
+	\vspace{-0.8cm}
 	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-16-single}
-	\vspace{-0.3cm}
+	\vspace{-0.35cm}
 	\caption{
@@
 	}
-\vspace{-0.5cm}
+%\vspace{-0.5cm}
 \end{figure}
@@
 Machine learning algorithms have earned a reputation for eluding human interpretation, while being vulnerable to evasion and poisoning.
-Given the risks associated with introducing such techniques in the context of security, it is natural to be concerned with the interpretability of the models we have proposed.
-With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
+Given the security risks associated with introducing such techniques, it is natural to be concerned with the interpretability of the models we have proposed.
+With the exception of global state, the tile coding parameters we make use of ensure that the set of outputs for each feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
 %\Cref{fig:intern-16-tcp-iat,fig:intern-16-tcp-something} show how we may visualise the portion of a policy for each feature, and describe what information can be gained from doing so.
@@
 By contrast, our approach takes a constant time to compute an action for a flow regardless of topology size.
-Their assumptions about traffic response to such bandwidth expansion do not hold for constant bitrate flows (e.g., VoIP) and may not extend to HTTP DASH flows, both of which make up a sizeable proportion of network traffic.
-\cbend
+Their assumptions about traffic response to such bandwidth expansion do not hold for constant bitrate flows (e.g., VoIP) and may not extend to HTTP DASH flows, both of which make up a sizeable proportion of network \cbend traffic.
 
@@
 Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
-
 \Textcite{DBLP:conf/sp/SmithS18} use AS-level routing to tackle both transit-link and flooding-based attacks.
@@
 To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}.
-
 An \emph{indirect-control} RL problem is one where agents act in service to \emph{another technique} responsible for decision-making, optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
-A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
+A past example includes learning when best to share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
 %The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
 This work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated.
-Outside of intrusion detection, there has been growing interest in the use of RL in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
-
+Outside of intrusion detection, there has been growing interest in the use of RL in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+\textcite{DBLP:conf/sigcomm/MaoNA17} employ client-side observations of network state and video performance with RL to optimise bitrate selection for multimedia streaming.
 \emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation.
@@
 To overcome the high latency of action computation via a neural network, two agents are trained, handling aspects of short and long flows respectively.
-The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
-The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.
+The first learns to optimise the flow size thresholds to demarcate long and short flows; these short flows are routed by ECMP.
+The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows.
 
@@
 %Actor-critic methods, or algorithms based on eligibility traces are good candidates for investigation. 
-The remaining weaknesses invite many further improvements worth investigation.
+The remaining weaknesses invite many improvements worth investigation.
 %??
@@
 \printbibliography
-\enlargethispage{1cm}
+%\enlargethispage{-1cm}
 \vspace{-1.1cm}

### 2019-09-27 paper/paper-revised.tex
A few tweaks...

@@
 In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
-The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the "edge" of their network lies through vNF relocation.
+The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the ``edge'' of their network lies through vNF relocation.
 \cbend
@@
 \cbstart
-Naturally, this implies that an agent must carry work forward (and coalesce state updates) when the \emph{host density} $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.
+Naturally, this implies that an agent must carry work forward (and coalesce state updates) when \emph{host density} is $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.
 \cbend
@@
 In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
-Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes---a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign otherwise.
+Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between \cbstart episodes---a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign \cbend otherwise.
 Benign hosts generate traffic from either \cref{sec:tcp-http-traffic-model,sec:udp-opus-traffic-model} depending on the experiment, while malicious hosts generate traffic according to \cref{sec:attack-traffic-model} (both at experiment-dependent rates).
@@
 
-We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
+We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and \cbstart \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
 %Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
-We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold and the best RL-based model underlined.
+We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold and the best RL-based model \cbend underlined.
 Average rewards take into account any portions of time that an agent allows illegal system states.

### 2019-10-01 paper/paper-revised.tex
Final paper and in blue!

--- 2019-09-26 paper/paper-revised.tex
@@
 \cbstart
+{\color{uofgcobalt}
 This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
@@
 These new traffic models inform an empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
-We additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS \cbend defences.
+We additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS \cbend defences.}
 
@@
 These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
-\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design, and control the proportion of upstream packets from each external host to discard.
+{\color{uofgcobalt}\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design}, and control the proportion of upstream packets from each external host to discard.
 Each destination node $s$ has a maximum capacity, \cbend $U_s$.
@@
 \cbstart
+{\color{uofgcobalt}
 Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
@@
 In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
-The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the "edge" of their network lies through vNF relocation.
+The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the ``edge'' of their network lies through vNF relocation.}
 \cbend
@@
 The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
-\cbstart In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) the load measurement of a middle hop or the last hop. \cbend
+\cbstart {\color{uofgcobalt}In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) the load measurement of a middle hop or the last hop.} \cbend
 \Cref{fig:global-state-path} illustrates the process.
@@
 Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
-\cbstart Rather than having the network push the data to an agent, the agent requests this information about active flows periodically to isolate it from non-control-plane traffic and to eliminate the risk of resource exhaustion by excessive requests.
-\cbend
+\cbstart {\color{uofgcobalt}Rather than having the network push the data to an agent, the agent requests this information about active flows periodically to isolate it from non-control-plane traffic and to eliminate the risk of resource exhaustion by excessive requests.} \cbend
 
@@
 \cbstart
-Naturally, this implies that an agent must carry work forward (and coalesce state updates) when the \emph{host density} $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.
+{\color{uofgcobalt}Naturally, this implies that an agent must carry work forward (and coalesce state updates) when \emph{host density} is $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.}
 \cbend
@@
 Hosts and clients both use TCP Cubic \cite{rfc8312}.
-Each host's download rate is limited to match whatever maximum bandwidth was assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.).
+Each host's download rate is limited to match the maximum bandwidth assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.) \cbstart {\color{uofgcobalt}as a browser \cbend might.}
 On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
-This presents a more balanced distribution of flow duration and size, with large files included to model elephant flows.
+This presents a balanced distribution of flow duration and size, with large files included to model elephant flows.
 
@@
 \cbstart
-Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
+{\color{uofgcobalt}Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
 We configure each instance of \emph{hping3} to generate ethernet MTU-sized packets (\SI{1500}{\byte}) with a random source and destination port towards a target server, and configure the output rate $r$ (in \si{\mega\bit\per\second}) by setting the inter-arrival time $t_{\mathit{attack}}=\frac{1500 \cdot 8}{r\cdot10^6}$.
 This fulfils certain characteristics of many types of amplification DDoS traffic: it is congestion-unaware \cite{DBLP:conf/ndss/Rossow14}, and packets are larger than the minimum frame size and identically-sized (e.g., NTP amplification traffic is fragmented at the application layer into \SI{482}{\byte} chunks \cite{cisco-ntp-amp}).
-We differ from NTP amplification in frame size so that inter-arrival times are larger, to keep emulation of the network feasible at high traffic rates.
+We differ from NTP amplification in frame size so that inter-arrival times are larger, to keep emulation of the network feasible at high traffic rates.}
 \cbend
@@
 \cbstart
-Marl is known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} strategy, yet the state of the art has long since moved on.
+{\color{uofgcobalt}Marl is known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} strategy, yet the state of the art has long since moved on.
 To paint a more current picture, we compare our work against an effective modern approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
@@
 We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\!
-\times$} expansion from legitimate flows, making a judgement after \cbend \SI{5}{\second}.
+\times$} expansion from legitimate flows, making a judgement after \cbend \SI{5}{\second}.}
 
@@
 In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
-Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes---a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign otherwise.
+Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between \cbstart episodes---{\color{uofgcobalt}a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign \cbend otherwise.}
 Benign hosts generate traffic from either \cref{sec:tcp-http-traffic-model,sec:udp-opus-traffic-model} depending on the experiment, while malicious hosts generate traffic according to \cref{sec:attack-traffic-model} (both at experiment-dependent rates).
@@
 %The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
-\cbstart For our choices of $n$ in both topologies, we observe $N_{\mathit{hosts}} \in \left\{24, 48, 96, 192\right\}$, and an expected number of malicious hosts $\mathbb{E}\left[N_{\mathit{attackers}}\right] \in \left\{9.6, 19.2, 38.4, 76.8\right\}$.
+{\color{uofgcobalt}\cbstart For our choices of $n$ in both topologies, we observe $N_{\mathit{hosts}} \in \left\{24, 48, 96, 192\right\}$, and an expected number of malicious hosts $\mathbb{E}\left[N_{\mathit{attackers}}\right] \in \left\{9.6, 19.2, 38.4, 76.8\right\}$.
 For the largest choice of $n$, we see an expected total attack traffic $\mathbb{E}\left[V_{\mathit{attack}}\right] = 334.05$ and \SI{422.4}{\mega\bit\per\second} for Opus and HTTP traffic respectively.
-\cbend
+\cbend}
 
@@
 
-We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
+We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and {\color{uofgcobalt} \cbstart \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
 %Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
-We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold and the best RL-based model underlined.
+We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold and the best RL-based model \cbend underlined.}
 Average rewards take into account any portions of time that an agent allows illegal system states.
@@
 Our new agent designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
-\cbstart Interestingly, SPIFFY is unable to effectively protect constant bitrate traffic. \cbend
+\cbstart {\color{uofgcobalt}Interestingly, SPIFFY is unable to effectively protect constant bitrate traffic.} \cbend
 
@@
 Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
-In both cases, SPIFFY greatly outperforms the RL-based methods.
+{\color{uofgcobalt}In both cases, SPIFFY greatly outperforms the RL-based methods.}
 
@@
 \cbstart
-To assess the effect of larger volumes of attack traffic, we increase an attacker's output by various factors, supposing $n=16$ with HTTP traffic (\emph{Guarded}, Single); \cref{tab:atk-vol} records the expected rate of attack and average performance.
-The initial increase in traffic volume causes the steepest reduction in performance (due to the increased cost of incorrect action), though performance levels out as attack traffic \cbend increases.
+{\color{uofgcobalt}To assess the effect of larger volumes of attack traffic, we increase an attacker's output by various factors, supposing $n=16$ with HTTP traffic (\emph{Guarded}, Single); \cref{tab:atk-vol} records the expected rate of attack and average performance.
+The initial increase in traffic volume causes the steepest reduction in performance (due to the increased cost of incorrect action), though performance levels out as attack traffic \cbend increases.}
 
@@
 	\toprule\multicolumn{1}{c}{Factor} & \multicolumn{1}{c}{$\mathbb{E}\left[V_{\mathit{attack}}\right]$ (\si{\mega\bit\per\second})} & \multicolumn{1}{c}{Reward} \\ \midrule
+	1.5 & 633.6 & 0.671 \\
 	2.0 & 844.8 & 0.625 \\
@@
 \cbstart
-This cost is constant and independent of network size.
-As discussed in \cref{sec:systems-considerations}, we are able to judge 3 flows before this deadline: the difference is primarily accounted for by serialisation/communication delays and single-threaded processing in the Python language.
+{\color{uofgcobalt}This cost is constant and independent of network size.
+As discussed in \cref{sec:systems-considerations}, we are able to judge 3 flows before this deadline: the difference is primarily accounted for by serialisation/communication delays and single-threaded processing in the Python language.}
 \cbend
@@
 \cbstart
-Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this strand still trails behind existing (exact) DDoS flow detection mechanisms where TCP traffic is concerned.
+{\color{uofgcobalt}Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this strand still trails behind existing (exact) DDoS flow detection mechanisms where TCP traffic is concerned.
 The ability to better protect VoIP traffic when compared against one of these approaches is a curious observation, which suggests that other (exact) protocol-agnostic approaches may carry hidden assumptions and is a promising direction for future investigation.
@@
 What we offer above many of the approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments, low-overhead and fixed-cost decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
-Moreover, our decision making processes are entirely agnostic of the protocol or content of traffic, offering future-proofing against the introduction of new transports.
+Moreover, our decision making processes are entirely agnostic of the protocol or content of traffic, offering future-proofing against the introduction of new transports.}
 \cbend
@@
 \cbstart
-Can an attack on the controller can impact our approach?
+{\color{uofgcobalt}Can an attack on the controller can impact our approach?
 This question hinges upon whether the deployment environment is a traditional network or is fully SDN-enabled---each agent is, in a sense, \emph{a} controller alongside the network's controller.
@@
 Thus, the controller is made no more vulnerable than in any other SDN.
-The only necessary change for such a scenario is that a load measurement which has not been updated (due to a timeout or missed deadline) should be set at $R_t=-1$.
+The only necessary change for such a scenario is that a load measurement which has not been updated (due to a timeout or missed deadline) should be set at $R_t=-1$.}
 \cbend
@@
 \cbstart
-One limit of SDN-capable hardware is that OpenFlow rules occupy \SI{6}{$\!
+{\color{uofgcobalt}One limit of SDN-capable hardware is that OpenFlow rules occupy \SI{6}{$\!
 \times$} the space of standard rules---commercial switches only have TCAM space for \SIrange{2}{20}{\kilo} rules \cite{DBLP:journals/comsur/NguyenSBT16}.
@@
 Given the high churn and prevalence of mice flows, eviction here is most likely to affect flows which are complete.
-In both cases, extra rules can be made available by upgrading rules which completely filter a flow into upstream blackholing (as in collaborative approaches \cite{DBLP:conf/acsac/RamanathanMYZ18}), having the agent remove this rule once blackholing is active.
+In both cases, extra rules can be made available by upgrading rules which completely filter a flow into upstream blackholing (as in collaborative approaches \cite{DBLP:conf/acsac/RamanathanMYZ18}), having the agent remove this rule once blackholing is active.}
 \cbend
@@
 \cbstart
-By contrast, our approach takes a constant time to compute an action for a flow regardless of topology size.
-Their assumptions about traffic response to such bandwidth expansion do not hold for constant bitrate flows (e.g., VoIP) and may not extend to HTTP DASH flows, both of which make up a sizeable proportion of network \cbend traffic.
+{\color{uofgcobalt}By contrast, our approach takes a constant time to compute an action for a flow regardless of topology size.
+Their assumptions about traffic response to such bandwidth expansion do not hold for constant bitrate flows (e.g., VoIP) and may not extend to HTTP DASH flows, both of which make up a sizeable proportion of network \cbend traffic.}
 

### 2019-11-11 paper/paper-revised.tex
Hide revision indicators.

@@
 \setlength\changebarwidth{6pt}
-\cbcolor{uofgcobalt}
+%\cbcolor{uofgcobalt}
+\cbcolor{white}
+\colorlet{revisiontext}{black}
+%\colorlet{revisiontext}{uofgcobalt}
 
@@
 \cbstart
-{\color{uofgcobalt}
+{\color{revisiontext}
 This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
@@
 These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
-{\color{uofgcobalt}\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design}, and control the proportion of upstream packets from each external host to discard.
+{\color{revisiontext}\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design}, and control the proportion of upstream packets from each external host to discard.
 Each destination node $s$ has a maximum capacity, \cbend $U_s$.
@@
 \cbstart
-{\color{uofgcobalt}
+{\color{revisiontext}
 Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
@@
 The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
-\cbstart {\color{uofgcobalt}In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) the load measurement of a middle hop or the last hop.} \cbend
+\cbstart {\color{revisiontext}In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) the load measurement of a middle hop or the last hop.} \cbend
 \Cref{fig:global-state-path} illustrates the process.
@@
 Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
-\cbstart {\color{uofgcobalt}Rather than having the network push the data to an agent, the agent requests this information about active flows periodically to isolate it from non-control-plane traffic and to eliminate the risk of resource exhaustion by excessive requests.} \cbend
+\cbstart {\color{revisiontext}Rather than having the network push the data to an agent, the agent requests this information about active flows periodically to isolate it from non-control-plane traffic and to eliminate the risk of resource exhaustion by excessive requests.} \cbend
 
@@
 \cbstart
-{\color{uofgcobalt}Naturally, this implies that an agent must carry work forward (and coalesce state updates) when \emph{host density} is $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.}
+{\color{revisiontext}Naturally, this implies that an agent must carry work forward (and coalesce state updates) when \emph{host density} is $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.}
 \cbend
@@
 Hosts and clients both use TCP Cubic \cite{rfc8312}.
-Each host's download rate is limited to match the maximum bandwidth assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.) \cbstart {\color{uofgcobalt}as a browser \cbend might.}
+Each host's download rate is limited to match the maximum bandwidth assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.) \cbstart {\color{revisiontext}as a browser \cbend might.}
 On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
@@
 \cbstart
-{\color{uofgcobalt}Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
+{\color{revisiontext}Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
 We configure each instance of \emph{hping3} to generate ethernet MTU-sized packets (\SI{1500}{\byte}) with a random source and destination port towards a target server, and configure the output rate $r$ (in \si{\mega\bit\per\second}) by setting the inter-arrival time $t_{\mathit{attack}}=\frac{1500 \cdot 8}{r\cdot10^6}$.
@@
 \cbstart
-{\color{uofgcobalt}Marl is known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} strategy, yet the state of the art has long since moved on.
+{\color{revisiontext}Marl is known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} strategy, yet the state of the art has long since moved on.
 To paint a more current picture, we compare our work against an effective modern approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
@@
 In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
-Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between \cbstart episodes---{\color{uofgcobalt}a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign \cbend otherwise.}
+Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between \cbstart episodes---{\color{revisiontext}a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign \cbend otherwise.}
 Benign hosts generate traffic from either \cref{sec:tcp-http-traffic-model,sec:udp-opus-traffic-model} depending on the experiment, while malicious hosts generate traffic according to \cref{sec:attack-traffic-model} (both at experiment-dependent rates).
@@
 %The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
-{\color{uofgcobalt}\cbstart For our choices of $n$ in both topologies, we observe $N_{\mathit{hosts}} \in \left\{24, 48, 96, 192\right\}$, and an expected number of malicious hosts $\mathbb{E}\left[N_{\mathit{attackers}}\right] \in \left\{9.6, 19.2, 38.4, 76.8\right\}$.
+{\color{revisiontext}\cbstart For our choices of $n$ in both topologies, we observe $N_{\mathit{hosts}} \in \left\{24, 48, 96, 192\right\}$, and an expected number of malicious hosts $\mathbb{E}\left[N_{\mathit{attackers}}\right] \in \left\{9.6, 19.2, 38.4, 76.8\right\}$.
 For the largest choice of $n$, we see an expected total attack traffic $\mathbb{E}\left[V_{\mathit{attack}}\right] = 334.05$ and \SI{422.4}{\mega\bit\per\second} for Opus and HTTP traffic respectively.
@@
 
-We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and {\color{uofgcobalt} \cbstart \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
+We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and {\color{revisiontext} \cbstart \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
@@
 Our new agent designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
-\cbstart {\color{uofgcobalt}Interestingly, SPIFFY is unable to effectively protect constant bitrate traffic.} \cbend
+\cbstart {\color{revisiontext}Interestingly, SPIFFY is unable to effectively protect constant bitrate traffic.} \cbend
 
@@
 Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
-{\color{uofgcobalt}In both cases, SPIFFY greatly outperforms the RL-based methods.}
+{\color{revisiontext}In both cases, SPIFFY greatly outperforms the RL-based methods.}
 
@@
 \cbstart
-{\color{uofgcobalt}To assess the effect of larger volumes of attack traffic, we increase an attacker's output by various factors, supposing $n=16$ with HTTP traffic (\emph{Guarded}, Single); \cref{tab:atk-vol} records the expected rate of attack and average performance.
+{\color{revisiontext}To assess the effect of larger volumes of attack traffic, we increase an attacker's output by various factors, supposing $n=16$ with HTTP traffic (\emph{Guarded}, Single); \cref{tab:atk-vol} records the expected rate of attack and average performance.
 The initial increase in traffic volume causes the steepest reduction in performance (due to the increased cost of incorrect action), though performance levels out as attack traffic \cbend increases.}
@@
 \cbstart
-{\color{uofgcobalt}This cost is constant and independent of network size.
+{\color{revisiontext}This cost is constant and independent of network size.
 As discussed in \cref{sec:systems-considerations}, we are able to judge 3 flows before this deadline: the difference is primarily accounted for by serialisation/communication delays and single-threaded processing in the Python language.}
@@
 \cbstart
-{\color{uofgcobalt}Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this strand still trails behind existing (exact) DDoS flow detection mechanisms where TCP traffic is concerned.
+{\color{revisiontext}Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this strand still trails behind existing (exact) DDoS flow detection mechanisms where TCP traffic is concerned.
 The ability to better protect VoIP traffic when compared against one of these approaches is a curious observation, which suggests that other (exact) protocol-agnostic approaches may carry hidden assumptions and is a promising direction for future investigation.
@@
 \cbstart
-{\color{uofgcobalt}Can an attack on the controller can impact our approach?
+{\color{revisiontext}Can an attack on the controller can impact our approach?
 This question hinges upon whether the deployment environment is a traditional network or is fully SDN-enabled---each agent is, in a sense, \emph{a} controller alongside the network's controller.
@@
 \cbstart
-{\color{uofgcobalt}One limit of SDN-capable hardware is that OpenFlow rules occupy \SI{6}{$\!
+{\color{revisiontext}One limit of SDN-capable hardware is that OpenFlow rules occupy \SI{6}{$\!
 \times$} the space of standard rules---commercial switches only have TCAM space for \SIrange{2}{20}{\kilo} rules \cite{DBLP:journals/comsur/NguyenSBT16}.
@@
 \cbstart
-{\color{uofgcobalt}By contrast, our approach takes a constant time to compute an action for a flow regardless of topology size.
+{\color{revisiontext}By contrast, our approach takes a constant time to compute an action for a flow regardless of topology size.
 Their assumptions about traffic response to such bandwidth expansion do not hold for constant bitrate flows (e.g., VoIP) and may not extend to HTTP DASH flows, both of which make up a sizeable proportion of network \cbend traffic.}

### 2019-12-12 paper/paper.tex
Fix URLs in online cites.

@@
 
-	\ifentrytype{mics}{}{% Remove publisher and editor except for books
+	\ifentrytype{misc}{}{% Remove url except for misc and online.
+		\ifentrytype{online}{}{
 			\clearfield{url}
@@
 	}
+}
 
 %opening!
+
+\usepackage{scalerel}
+\usetikzlibrary{svg.path}
+
+\definecolor{orcidlogocol}{HTML}{A6CE39}
+\tikzset{
+	orcidlogo/.pic={
+		\fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
+		\fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
+		svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
+		svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
+	}
+}
+
+\newcommand\orcidicon[1]{
+	\href{https://orcid.org/#1}{\mbox{\scalerel*{
+		\begin{tikzpicture}[yscale=-1,transform shape]
+		\pic{orcidlogo};
+		\end{tikzpicture}
+	}{|}}}}
 
@@
 
+%\usepackage{framed}
+%\usepackage{adjustbox}
+%\colorlet{shadecolor}{uofgcobalt!20}
+\usepackage[color]{changebar}
+\setlength\changebarsep{10pt}
+\setlength\changebarwidth{6pt}
+%\cbcolor{uofgcobalt}
+\cbcolor{white}
+\colorlet{revisiontext}{black}
+%\colorlet{revisiontext}{uofgcobalt}
+
 %-------------------------------------%
@@
 %	Kyle~A.~Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
-	Kyle~A.~Simpson,
-	Simon~Rogers,
-	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
+	\protect Kyle~A.~Simpson\protect\orcidicon{0000-0001-8068-9909},
+%	Kyle~A.~Simpson,
+	Simon~Rogers\protect\orcidicon{0000-0003-3578-4477},
+%	Simon~Rogers,
+	Dimitrios~P.~Pezaros\protect\orcidicon{0000-0003-0939-378X},~\IEEEmembership{Senior~Member,~IEEE,},\\
+%	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
 	School of Computing Science, University of Glasgow, Glasgow, G12 8RZ, UK\\
@@
 %?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
-Briefly, this paper contributes:
-\begin{itemize}
-%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
-%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}.
-We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
-	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
-	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
-	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
-	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
-\end{itemize}
+
+\subsection{Contributions}
+\cbstart
+{\color{revisiontext}
+This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
+These are designed to make effective decisions irrespective of protocol, and act on individual flows at the edge of any network topology.
+We offer an in-depth investigation into suitable features for automatic DDoS mitigation, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
+These features have been suggested by past studies, and independently tested in their own contexts.
+Our study is the first attempt to quantify the individual efficacy of each in an RL setting.
+
+We implement reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
+To our knowledge, this is the first attempt to study or replicate Opus-based VoIP traffic, which has become commonplace since the codec's release in 2012.
+These new traffic models inform an empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
+We additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS \cbend defences.}
 
@@
 \subsection{Threat Model}
-An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
-Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
-We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
-The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
+An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to hosts, and they are expected to act rationally in its pursuit.
+Threat actors are external and act intentionally, aren't expected to be 
+\emph{advanced persistent threats}, and likely range from hacktivists to moderately funded adversaries.
+We assume that attacks will be volumetric DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
+The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication with \nth{3} party organisations such as upstream ISPs.
 The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
 
-We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
-While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
-The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
+We do not assume that an attacker has white-box access to an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
+While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in volatile network traffic statistics falls outside of the scope of this work.
+The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/corr/abs-1902-09062}.
 These are APT-level capabilities, whose exploration presents a rich source for future work.
@@
 \subsection{System Design and Assumptions}
-A deployment environment is defined as a network, with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
-These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
-Agents are co-located with each egress switch, and control the proportion of upstream packets from each external host to discard.
-Each destination node $s$ has a maximum capacity, $U_s$.
+A deployment environment is a network with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
+These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
+{\color{revisiontext}\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design}, and control the proportion of upstream packets from each external host to discard.
+Each destination node $s$ has a maximum capacity, \cbend $U_s$.
 
@@
 The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
-Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners.
+Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners (e.g., as examined by \textcite{TNSM-that-guys-paper}).
 Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.
 
+\cbstart
+{\color{revisiontext}
+Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
+Only the ingress/egress nodes from a network need to be OpenFlow-enabled, as it is advantageous to perform filtering as close to a flow's source as possible.
+In a traditional network, each agent has exclusive control over its switch's tables.
+In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
+The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the ``edge'' of their network lies through vNF relocation.}
+\cbend
+
 \subsection{Algorithm}
-To make decisions cheaply and at low latency, we use \emph{semi-gradient Sarsa with tile coding} as described in \cref{eqn:sg-sarsa} and \cref{sec:reinforcement-learning}, rather than making use of neural networks or more complicated function approximators.
-Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
+To make decisions cheaply and at low latency, we use \emph{semi-gradient Sarsa with tile coding} as described in \cref{eqn:sg-sarsa} and \cref{sec:reinforcement-learning}, rather than using neural networks or more complicated function approximators.
+Exploration is introduced via $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
 Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their weight vector updates with one another (but may share experience and traces with one another).
@@
 The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
-In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
+\cbstart {\color{revisiontext}In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) the load measurement of a middle hop or the last hop.} \cbend
 \Cref{fig:global-state-path} illustrates the process.
@@
 Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
+\cbstart {\color{revisiontext}Rather than having the network push the data to an agent, the agent requests this information about active flows periodically to isolate it from non-control-plane traffic and to eliminate the risk of resource exhaustion by excessive requests.} \cbend
 
@@
 The measurements of \textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
-We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$,.
+We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
 The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
-
 We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
@@
 
-\subsection{Systems Considerations}
-Taking many actions per-timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
+\subsection{Systems Considerations}\label{sec:systems-considerations}
+Taking many actions per timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
 This introduces some potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated.
@@
 For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
-Uncompleted work is passed on to the next timestep, until the list is emptied, at which point is is repopulated using the set of available measurements.
+Uncompleted work is passed on to the next timestep, until the list is emptied, at which point it is repopulated using the set of available measurements.
 To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
-State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
+State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields.
 Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
 We expect this should be tuned based on the frequency at which statistics arrive.
+\cbstart
+{\color{revisiontext}Naturally, this implies that an agent must carry work forward (and coalesce state updates) when \emph{host density} is $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.}
+\cbend
 
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{../plots/ftprep-tcp-laf-cap-box}
-	\vspace{-1cm}
+%	\vspace{-0.25cm}
+	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-laf-cap-box}
+	\vspace{-0.75cm}
 	\caption{
@@
 	}
+\vspace{-0.6cm}
 \end{figure}
 
-The main element required to move to a per-source model is a feature set which offers high predictive power, such that behavioural differences are readily apparent to an agent.
-Elaborating further on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and offer an investigation into their use alongside different traffic types:
+The main element required by a per-source model is a feature set with high predictive power, so that behavioural differences are apparent to an agent.
+Elaborating on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and investigate their use alongside different traffic types:
 %?? We use these features, and why...
@@
 
-\section{A New Normal}\label{sec:a-new-normal}
+\section{Traffic Modelling}\label{sec:a-new-normal}
 
@@
 Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
-Hosts and clients use the TCP Cubic \cite{rfc8312} flavour of TCP.
-Each host's download rate is limited to match whatever maximum bandwidth was assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.).
+Hosts and clients both use TCP Cubic \cite{rfc8312}.
+Each host's download rate is limited to match the maximum bandwidth assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.) \cbstart {\color{revisiontext}as a browser \cbend might.}
 On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
-This presents a more balanced distribution of flow duration and size, with large files included to model elephant flows.
-
-\subsection{UDP (Opus) Traffic Model}
-VoIP traffic exhibits very different characteristics to the above model; packet arrivals are highly periodic due to real-time application requirements, flows have a constant bitrate, and do not react substantially to lost packets.
+This presents a balanced distribution of flow duration and size, with large files included to model elephant flows.
+
+\subsection{UDP (Opus/VoIP) Traffic Model}\label{sec:udp-opus-traffic-model}
+VoIP traffic exhibits very different characteristics to the above model; packet arrivals are highly periodic due to real-time requirements, flows have a constant bitrate, and do not react substantially to lost packets.
 Interestingly, DDoS attack traffic is known to share many of these characteristics, offering an interesting detection problem.
-
-We present a VoIP traffic model\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward video game communities.
-We chose Discord as our prototype due to its publicly documented API, many open source bot frameworks for experimentation, large user base, and due to the lack of models for Opus-encoded traffic.
+We present a VoIP traffic model\footnote{\url{https://github.com/FelixMcFelix/opus-voip-traffic}} based on Discord\footnote{\url{https://discord.gg}}, a freely-available messaging and VoIP platform geared toward gaming communities.
+We chose Discord as our prototype due to its publicly documented API, many open source bot frameworks, large user base, and due to the lack of models for Opus-encoded traffic.
 %?? Highly periodic, CBR
 
-Hosts send RTP traffic with Salsa20 encrypted payloads---audio frames of length \SI{20}{\milli\second} at a target \SI{96}{\kilo\bit\per\second}.
+Hosts send RTP traffic with Salsa20 encrypted payloads---\SI{20}{\milli\second} audio frames at \SI{96}{\kilo\bit\per\second}.
 We generate similar traffic at hosts by replaying anonymised traces gathered in general use and tabletop RPG servers; each trace contains only the size of each audio payload, entries denoting missed packets, and the duration of silent periods.
@@
 All traffic passes over a central server which groups hosts into rooms, and is forwarded to other participants; we do not replicate pre-call Websocket traffic which would be used for authentication.
-There is no peer-to-peer traffic---the server acts as a TURN relay \cite{rfc5766} for all hosts.
+There is no peer-to-peer traffic---the server acts as a TURN relay for all hosts.
 %?? Reflective factor among \emph{authenticated hosts}.
-
 We find that each flow occupies an expected \SI{52.4}{\kilo\bit\per\second} upstream bandwidth.
@@
 %?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
-\subsection{Attack Traffic Model}
-Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
-Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that emulation remains tractable.
-
-\section{Evaluation}
+\subsection{Attack Traffic Model}\label{sec:attack-traffic-model}
+\cbstart
+{\color{revisiontext}Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
+We configure each instance of \emph{hping3} to generate ethernet MTU-sized packets (\SI{1500}{\byte}) with a random source and destination port towards a target server, and configure the output rate $r$ (in \si{\mega\bit\per\second}) by setting the inter-arrival time $t_{\mathit{attack}}=\frac{1500 \cdot 8}{r\cdot10^6}$.
+This fulfils certain characteristics of many types of amplification DDoS traffic: it is congestion-unaware \cite{DBLP:conf/ndss/Rossow14}, and packets are larger than the minimum frame size and identically-sized (e.g., NTP amplification traffic is fragmented at the application layer into \SI{482}{\byte} chunks \cite{cisco-ntp-amp}).
+We differ from NTP amplification in frame size so that inter-arrival times are larger, to keep emulation of the network feasible at high traffic rates.}
+\cbend
+
+\section{Evaluation}\label{sec:evaluation}
 
@@
 
-We compare our work most naturally against the MARL approach introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
+We compare our work most naturally against MARL, introduced by \textcite{DBLP:journals/eaai/MalialisK15}, the state-of-the-art in RL-based DDoS prevention.
 We are most interested in seeing how their approach contrasts with ours across different topologies and workloads.
 Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network---we aim to see how these characteristics affect performance and learning rate.
+\cbstart
+{\color{revisiontext}Marl is known to outperform the AIMD \cite{DBLP:journals/ton/YauLLY05} strategy, yet the state of the art has long since moved on.
+To paint a more current picture, we compare our work against an effective modern approach, SPIFFY \cite{DBLP:conf/ndss/KangGS16}.
+SPIFFY tests a proportion of flows by routing them through an alternate path with higher bandwidth, observing how their speed changes some time later.
+This comparison lets us position our new agent designs against the state of the art, observing that SPIFFY has a similar mode of interaction to RL-based systems (taking action, observing an effect, and acting once again) and does not rely on protocol characteristics or signatures.
+We make the simplifying assumption that a suitable unused path exists (with identical bandwidth to the server's link).
+We test \SI{10}{\percent} of active flows at a time (according to the authors' observation that there is a factor of \num{10} difference between the ideal and achieved bandwidth expansion), excluding flows below \SI{50}{\kilo\bit\per\second} and requiring a \SI{3}{$\!
+\times$} expansion from legitimate flows, making a judgement after \cbend \SI{5}{\second}.}
 
@@
 In each case, the algorithm's performance was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
-Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between episodes.
+Host allocations at the beginning of each episode were generated pseudorandomly to ensure fairness between \cbstart episodes---{\color{revisiontext}a host is malicious with probability $\operatorname{P}(\mathit{malicious})$, and is benign \cbend otherwise.}
+Benign hosts generate traffic from either \cref{sec:tcp-http-traffic-model,sec:udp-opus-traffic-model} depending on the experiment, while malicious hosts generate traffic according to \cref{sec:attack-traffic-model} (both at experiment-dependent rates).
 
 All experiments were executed on Ubuntu 18.04.2 LTS (GNU/Linux 4.4.3-040403-generic x86\_64), using a 4-core Intel Core i7-6700K (clocked at \SI{4.2}{\giga\hertz}) and \SI{32}{\gibi\byte} of RAM.
-%All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
-All code underpinning these findings is available on a public repository.\footnote{Private until publication.}
+All code underpinning these findings is available on a public repository\footnote{\url{https://github.com/FelixMcFelix/rln-dc-ddos-paper}}.
+%All code underpinning these findings is available on a public repository.\footnote{Private until publication.}
 
@@
 %We aim to address this through \cref{sec:performance-in-an-emulated-environment}.
-
-To test this, we configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
-This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
+We configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
+This is a slight simplification of \citeauthor{DBLP:journals/eaai/MalialisK15}'s \textquote{online} experiment \cite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
 %The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
 %Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
-%$U_s$ was fixed at $k \ell mn+2$ \si{\mega\bit\per\second}.
+%$U_s$  $k \ell mn+2$ \si{\mega\bit\per\second}.
 %The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
@@
 %?? NEED TO VERIFY VOLUME OF CONGESTION-AWARE PROTOCOLS
-Malicious hosts sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, though when using TCP-like traffic, each attacker's upload range had to be increased to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows.
+Malicious hosts each sent \SIrange{2.5}{6}{\mega\bit\per\second} when attacking UDP traffic, though this was increased to \SIrange{4}{7}{\mega\bit\per\second} when using TCP-like traffic (to meaningfully impact benign flows).
 Given $n$ and $\operatorname{P}(\mathit{malicious})$, we see an expected malicious bandwidth \numrange{1.27}{1.87} and \SIrange{2.03}{2.18}{$\! \times U_s$} respectively.
 %The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.
+{\color{revisiontext}\cbstart For our choices of $n$ in both topologies, we observe $N_{\mathit{hosts}} \in \left\{24, 48, 96, 192\right\}$, and an expected number of malicious hosts $\mathbb{E}\left[N_{\mathit{attackers}}\right] \in \left\{9.6, 19.2, 38.4, 76.8\right\}$.
+For the largest choice of $n$, we see an expected total attack traffic $\mathbb{E}\left[V_{\mathit{attack}}\right] = 334.05$ and \SI{422.4}{\mega\bit\per\second} for Opus and HTTP traffic respectively.
+\cbend}
 
 $U_s$ was fixed at $N_{\mathit{hosts}}+2$ \si{\mega\bit\per\second} (to account for burstiness), and each link had a delay of \SI{10}{\milli\second}.
-All links had unbounded capacity, save for server-switch connections.
-These parameter choices match those of the original study to enable direct comparison, and many are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
+All links had unbounded capacity, save for each server-switch.
+These parameters match those of the original study to enable direct comparison, and many are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.
 
@@
 
-We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
+We now examine the performance of our two new models (\emph{Instant}, \emph{Guarded}) as compared against existing RL work (\emph{Marl}) and {\color{revisiontext} \cbstart \emph{SPIFFY} under different traffic behaviour and topologies, varying the host-to-learner ratio $n$ and environment.
 %?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).
 %Additionally, we examine the performance effects of environmental characteristics and potential improvements: negative reinforcement, the case where one agent makes all decisions, and pre-training on individual features.
-We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---these provide a rough idea of expected performance, giving the highest-performing model in bold.
+We present the average rewards for all combinations of these factors in \crefrange{tab:av-vals}{tab:av-ecmp-vals}---providing a rough idea of expected performance, with the highest-performing model in bold and the best RL-based model \cbend underlined.}
 Average rewards take into account any portions of time that an agent allows illegal system states.
@@
 
-	\expandableinput ../tables/tnsm-avg-reward.tex
+	\cbstart
+	\resizebox{0.8\linewidth}{!}{
+		\expandableinput ../tables/tnsm-tree-avg-reward-spiffy.tex
+	}
+	\cbend
+\vspace{-0.25cm}
 \end{table}
@@
 
-	\expandableinput ../tables/tnsm-ecmp-avg-reward.tex
+	\cbstart
+	\resizebox{0.8\linewidth}{!}{
+		\expandableinput ../tables/tnsm-ecmp-avg-reward-spiffy.tex
+	}
+	\cbend
+\vspace{-0.5cm}
 \end{table}
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-udp-box-separate}
-
+	\cbstart
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-udp-box-separate}
+	\cbend
+	\vspace{-0.3cm}
 	\caption{
@@
 In a single-destination network, we observe that Marl's performance degrades as $n$ increases.
-Typically, our \emph{Instant} agent design achieves the best performance in multi-agent mode, reducing collateral damage above the current state-of-the-art, but sharply degrades at low $n$ when agents share experience.
-We see a reversal of this trend for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
+Typically, our \emph{Instant} agent design achieves the best performance in multi-agent mode, having lower collateral damage than the current state-of-the-art, but sharply degrades at low $n$ when agents share experience.
+This trend reverses for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
 %Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
 \Cref{fig:udp-tree-box} shows the preserved traffic in multi-agent mode.
-
 %?? Discuss multi-dest topol once all results available.
 When defending multiple destinations, we see a sharp decrease in the effectiveness of all agent designs.
-Our \emph{Instant} and \emph{Guarded} designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
+Our new agent designs become more effective as $n$ increases, while Marl's effectiveness is roughly constant (aside from the outlier at $n=12$).
+\cbstart {\color{revisiontext}Interestingly, SPIFFY is unable to effectively protect constant bitrate traffic.} \cbend
 
@@
 \begin{figure}
+%	\vspace{-0.5cm}
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-single}
-
+	\cbstart
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-box-single}
+	\cbend
+	\vspace{-0.3cm}
 	\caption{
 		Online performance for HTTP benign traffic in a single-destination network, single-agent mode.
-		\emph{Instant} and \emph{Guarded} exhibit essentially identical efficacy at $n~=~2$, protecting less traffic than Marl.
-		Only \emph{Guarded}'s performance rapidly increases with $n$, achieving considerably better median performance and lower variance than the other models.
+		\emph{Instant} and \emph{Guarded} exhibit similar efficacy at $n=2$, protecting less traffic than Marl.
+		Only \emph{Guarded}'s performance rapidly increases with $n$, achieving a considerably better median and lower variance than the other models.
 		The longer tails of outliers typically indicate the longer training time the new models require---we observe that \emph{Guarded} typically has considerably lower variance once it has converged on a stable policy.
@@
 	\centering
-	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-16-single}
-
+	\vspace{-0.8cm}
+	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-16-single}
+	\vspace{-0.35cm}
 	\caption{
-		Online performance of standard and single-agent models in a single-destination network with $n=16$ hosts per egress point and HTTP benign traffic.
+		Online performance of standard and single-agent models in a single-destination network with $n=16$ hosts per egress point, HTTP traffic.
 		At this level of host density, \emph{Guarded} reaches higher peak performance sooner and is considerably more consistent throughout the episode.
@@
 	}
+%\vspace{-0.5cm}
 \end{figure}
@@
 %SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.
-
 %?? Discuss multi-dest topol once results available.
@@
 Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.
+{\color{revisiontext}In both cases, SPIFFY greatly outperforms the RL-based methods.}
 
@@
 
+\subsection{Increased Attack Volume}\label{sec:results-attack-volume}
+\cbstart
+{\color{revisiontext}To assess the effect of larger volumes of attack traffic, we increase an attacker's output by various factors, supposing $n=16$ with HTTP traffic (\emph{Guarded}, Single); \cref{tab:atk-vol} records the expected rate of attack and average performance.
+The initial increase in traffic volume causes the steepest reduction in performance (due to the increased cost of incorrect action), though performance levels out as attack traffic \cbend increases.}
+
+\begin{table}
+\cbstart
+	\centering
+%	\vspace{-0.1cm}
+	\caption{Average reward versus attack volume.\label{tab:atk-vol}}
+\resizebox{0.45\linewidth}{!}{
+\begin{tabular}{@{}SSS@{}}
+	\toprule\multicolumn{1}{c}{Factor} & \multicolumn{1}{c}{$\mathbb{E}\left[V_{\mathit{attack}}\right]$ (\si{\mega\bit\per\second})} & \multicolumn{1}{c}{Reward} \\ \midrule
+	1.5 & 633.6 & 0.671 \\
+	2.0 & 844.8 & 0.625 \\
+	2.5 & 1056.0 & 0.620 \\
+	3.0 & 1267.2 & 0.619 \\
+	3.5 & 1478.4 & 0.600 \\
+	\bottomrule
+\end{tabular}
+}
+\cbend
+\vspace{-0.75cm}
+\end{table}
+
 \subsection{Computational Cost}
@@
 %Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
-Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
+Measurements from each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second} per flow.
 This is reassuring when measured alongside the insights from other work.
@@
 That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.
+\cbstart
+{\color{revisiontext}This cost is constant and independent of network size.
+As discussed in \cref{sec:systems-considerations}, we are able to judge 3 flows before this deadline: the difference is primarily accounted for by serialisation/communication delays and single-threaded processing in the Python language.}
+\cbend
 %?? TODO: update with modern numbers...
@@
 %Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
-Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
-Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
+\cbstart
+{\color{revisiontext}Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this strand still trails behind existing (exact) DDoS flow detection mechanisms where TCP traffic is concerned.
+The ability to better protect VoIP traffic when compared against one of these approaches is a curious observation, which suggests that other (exact) protocol-agnostic approaches may carry hidden assumptions and is a promising direction for future investigation.
+Similar traffic makes up a significant fraction of network load today (\SIrange{18}{27}{\percent}).
+Although we have conducted work to map the territory, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
 The benefits we have at present are, however, substantial.
-What we offer above the alternate approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
-The techniques we propose offer substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today.
-Moreover, our decision making processes are entirely agnostic of the protocol or content of inbound traffic, offering future-proofing against the introduction of new transports.
+What we offer above many of the approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments, low-overhead and fixed-cost decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
+Moreover, our decision making processes are entirely agnostic of the protocol or content of traffic, offering future-proofing against the introduction of new transports.}
+\cbend
 
@@
 
+\cbstart
+{\color{revisiontext}Can an attack on the controller can impact our approach?
+This question hinges upon whether the deployment environment is a traditional network or is fully SDN-enabled---each agent is, in a sense, \emph{a} controller alongside the network's controller.
+In a traditional network, only the agents act as controllers, but since they periodically request per-flow data (rather than continuously receiving it) no amount of flows generates more requests or messages to the agent.
+More work is generated, but we discuss how to handle this safely above.
+Accordingly, agents can never be stalled by request volume: their only remote communication (load measurements) comes from trusted nodes, is highly periodic, and has constant size.
+The same logic holds for a fully software-defined network.
+Recalling that we do not employ the network's controller to install filtering rules on edge switches, an agent's ability to act is unimpeded.
+Thus, the controller is made no more vulnerable than in any other SDN.
+The only necessary change for such a scenario is that a load measurement which has not been updated (due to a timeout or missed deadline) should be set at $R_t=-1$.}
+\cbend
 %\begin{figure}
@@
 
+\cbstart
+{\color{revisiontext}One limit of SDN-capable hardware is that OpenFlow rules occupy \SI{6}{$\!
+\times$} the space of standard rules---commercial switches only have TCAM space for \SIrange{2}{20}{\kilo} rules \cite{DBLP:journals/comsur/NguyenSBT16}.
+Our approach consumes a rule for each active flow (the host density), and by the end of an experiment a switch can accrue around \num{900} rules.
+While we use a default fallback action to maintain connectivity, eviction of high-value decisions which filter high-bandwidth attackers poses a significant risk.
+Given that most flows are small (with the majority of bytes coming from a few ``heavy-hitters'') \cite{DBLP:journals/ccr/PanBPS03}, it may suffice to only apply RL-based analysis to larger flows.
+%?? What could we do in the meantime? I propose manually assigning elephant flows a high importance rule, while increasing the importance of "block" decisions even higher.
+OpenFlow rules have an \emph{importance}, controlling which rules may be evicted by a new entry (preventing entries from evicting those with higher importance).
+If an agent is to act on all flows, a solution is to assign an importance of 0 to mice flows, 1 to elephant flows, and 2 to total filtering (leaving agents to time out and remove elephant flow rules to prevent bloat).
+Given the high churn and prevalence of mice flows, eviction here is most likely to affect flows which are complete.
+In both cases, extra rules can be made available by upgrading rules which completely filter a flow into upstream blackholing (as in collaborative approaches \cite{DBLP:conf/acsac/RamanathanMYZ18}), having the agent remove this rule once blackholing is active.}
+\cbend
+
 \section{Related Work}\label{sec:related-work}
@@
 \fakepara{DDoS Prevention}
-\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
-Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
-
+\Textcite{DBLP:conf/lcn/BragaMP10} examine the detection of flooding DDoS attacks through \emph{self-organising maps}, using SDN to gather statistics effectively.
+Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts contribute to an attack.
 %?? Actually talk about Marl (???) to appease reviewer \#1.
@@
 \emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
-\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
+\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion (having already saturated the capacity of their outbound links), while legitimate flows typically speed up to match the new fair-share rate.
 %Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
-Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ASes, although some of our feature choices are backed by similar observations.
+%Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ASes, although .
 A weakness of their approach is that computing a route to measure bandwidth expansion on real networks can be costly (up to \SI{14}{\second} for the Cogent topology), and that the low expansion factors in real network can require more ``rounds'' of filtering.
-Their assumptions about traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), both of which make up a sizeable proportion of network traffic.
-
-\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
+\cbstart
+{\color{revisiontext}By contrast, our approach takes a constant time to compute an action for a flow regardless of topology size.
+Their assumptions about traffic response to such bandwidth expansion do not hold for constant bitrate flows (e.g., VoIP) and may not extend to HTTP DASH flows, both of which make up a sizeable proportion of network \cbend traffic.}
+
+\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a generalised SDN framework for intrusion detection, but has shown the use of a \emph{k-nearest neighbours} classifier to detect individual attack flows.
 Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.
-
-\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
-This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
+\Textcite{DBLP:conf/sp/SmithS18} use AS-level routing to tackle both transit-link and flooding-based attacks.
+This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples.
 The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
-Because of this, traditional flooding-based attacks aren't fully alleviated.
+Because of this, flooding-based attacks aren't fully alleviated.
 
@@
 \fakepara{RL in Networks}
-Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
-Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
-Given that one of the main strengths of RL techniques is the ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works don't apply the rich literature on the subject to its fullest potential.
+Earnest, well-considered application of RL towards the challenge of intrusion prevention has seen comparatively little examination.
+Past work treats the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
+Given that the main strengths of RL techniques are the ability to control ongoing interaction and adapt by observing the concrete effects of actions, such works don't apply the rich literature on the subject to its fullest potential.
 
 For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
-A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
+A \emph{direct-control} RL problem is one where the RL agent(s) learn optimal control over a set of actions as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
 %We feel there is a shortage of work in this category at present, at least in the field of networks.
 To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}.
-
-An \emph{indirect-control} RL problem is one where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
-A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
+An \emph{indirect-control} RL problem is one where agents act in service to \emph{another technique} responsible for decision-making, optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
+A past example includes learning when best to share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
 %The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-The position of this work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
-Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
-Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.
-
+This work is weakened by its reliance on the problematic `DARPA99' dataset \cite{DBLP:conf/sp/SommerP10}, but the idea itself is well-treated.
+Outside of intrusion detection, there has been growing interest in the use of RL in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
+\textcite{DBLP:conf/sigcomm/MaoNA17} employ client-side observations of network state and video performance with RL to optimise bitrate selection for multimedia streaming.
 \emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation.
@@
 \section{Conclusions and Future Work}
-Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
-Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.
-
-We have identified weaknesses in past approaches arising from simulation assumptions, recommending an RL agent design which acts per-flow, and have outlined the algorithmic modifications and engineering choices needed to make its deployment feasible.
-Supporting this, we've presented an in-depth examination of the design of our feature space, offering quantitative and qualitative justification for our choices.
+Through this paper, we have discussed reinforcement learning and its relevance to network intrusion prevention.
+We believe the potential to learn feedback loop-like control online and against non-stationarity makes it particularly suited to the problems endemic to the field.
+We identified weaknesses in past work, recommending an RL agent which acts per flow, and have outlined the algorithmic and engineering choices needed to make its deployment feasible.
+Supporting this, we've presented an in-depth examination of our feature space, offering quantitative and qualitative justification for our choices.
 Our evaluation shows that our new agent designs considerably advance the state-of-the-art in RL-based DDoS prevention, with \emph{Guarded} agents showing the most promise for future evaluation.
@@
 %While this work still trails behind the performance of exact DDoS flow detection mechanisms, w
-We hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
-It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, more flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the same network resources or capabilities that other techniques rely upon.
-It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
-
+We hope it is clear that reinforcement learning holds promise and can inspire further innovation.
+It allows us to offer distinct advantages above existing works, such as protocol-agnostic DDoS flow detection, flexible deployment, and automatically learned low-overhead decision-making---without requiring many of the
+network resources or capabilities that other techniques rely upon.
+It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
+
+\vspace{-0.2cm}
 \section*{Acknowledgements}
-Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice, comments and technical assistance.
+\vspace{-0.1cm}
+Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice and technical assistance, and to our anonymous reviewers.
 %Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
-%Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
 %This work was supported by a nameless funder [grant number 0].
-This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].
+This work was supported in part by the Engineering and Physical Sciences Research Council [grants EP/M508056/1, EP/N033957/1 and EP/P004024/1] and European Cooperation in Science and Technology (COST) Action CA15127: RECODIS---Resilient communication and services.
+\vspace{-0.2cm}
 
@@
 \printbibliography
-\vspace{-0.5cm}
-\vskip 0pt plus -1fil
+%\enlargethispage{-1cm}
+\vspace{-1.1cm}
+\vskip -6pt plus -1fil
 \begin{IEEEbiography}
@@
 \end{IEEEbiography}
-\vspace{-0.5cm}
+\vspace{-1.3cm}
 \vskip 0pt plus -1fil
@@
 
-	Dr.\ Simon Rogers’s research involves the development of machine learning and statistical techniques to help with the analysis of complex datasets, particularly within the field of metabolomics but also other fields, human-computer interaction and information retrieval.
+	Dr.\ Simon Rogers’s research involves the development of machine learning and statistical techniques to help with the analysis of complex datasets, particularly within the field of metabolomics, human-computer interaction and information retrieval.
 \end{IEEEbiography}
-\vspace{-0.5cm}
+\vspace{-1.25cm}
 \vskip 0pt plus -1fil

### 2019-12-13 paper/paper.tex
Move key source to paper folder---author final.

@@
 % !TeX document-id = {b3181d9a-ffe0-4331-a65c-c7fb8f275531}
+% !TeX program = pdflatex
 % !BIB program = biber
@@
 }
-\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} }
+\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}}}
 
@@
 %	Simon~Rogers,
-	Dimitrios~P.~Pezaros\protect\orcidicon{0000-0003-0939-378X},~\IEEEmembership{Senior~Member,~IEEE,},\\
+	Dimitrios~P.~Pezaros\protect\orcidicon{0000-0003-0939-378X},~\IEEEmembership{Senior~Member,~IEEE},\\
 %	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
@@
 %	University of Glasgow, Glasgow, Scotland,\\
-	\email{k.simpson.1@research.gla.ac.uk}
+	\email{k.simpson.1@research.gla.ac.uk}, \email{Simon.Rogers@glasgow.ac.uk}, \email{Dimitrios.Pezaros@glasgow.ac.uk}
 %Anonymous Giraffe,
@@
 	\centering
-	\includegraphics[width=0.65\linewidth]{../plots/ftprep-cap-box}
+	\includegraphics[width=0.65\linewidth]{ftprep-cap-box}
 	\vspace{-0.75cm}
@@
 	\centering
-	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-cap-box}
+	\includegraphics[width=0.65\linewidth]{ftprep-tcp-cap-box}
 	\vspace{-0.75cm}
@@
 %	\vspace{-0.25cm}
-	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-laf-cap-box}
+	\includegraphics[width=0.65\linewidth]{ftprep-tcp-laf-cap-box}
 	\vspace{-0.75cm}
@@
 	\resizebox{0.8\linewidth}{!}{
-		\expandableinput ../tables/tnsm-tree-avg-reward-spiffy.tex
+		\expandableinput tnsm-tree-avg-reward-spiffy.tex
 	}
@@
 	\resizebox{0.8\linewidth}{!}{
-		\expandableinput ../tables/tnsm-ecmp-avg-reward-spiffy.tex
+		\expandableinput tnsm-ecmp-avg-reward-spiffy.tex
 	}
@@
 	\cbstart
-	\includegraphics[width=0.75\linewidth]{../plots/tnsm-udp-box-separate}
+	\includegraphics[width=0.75\linewidth]{tnsm-udp-box-separate}
 	\cbend
@@
 	\cbstart
-	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-box-single}
+	\includegraphics[width=0.75\linewidth]{tnsm-tcp-box-single}
 	\cbend
@@
 	\vspace{-0.8cm}
-	\includegraphics[width=0.75\linewidth]{../plots/tnsm-tcp-16-single}
+	\includegraphics[width=0.75\linewidth]{tnsm-tcp-16-single}
 	\vspace{-0.35cm}
@@
 \vspace{-0.1cm}
-Our thanks go to Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice and technical assistance, and to our anonymous reviewers.
+The authors would like to thank Colin Perkins, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Cristian Urlea for their advice and technical assistance, and their anonymous reviewers.
 %Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
 %This work was supported by a nameless funder [grant number 0].
-This work was supported in part by the Engineering and Physical Sciences Research Council [grants EP/M508056/1, EP/N033957/1 and EP/P004024/1] and European Cooperation in Science and Technology (COST) Action CA15127: RECODIS---Resilient communication and services.
-\vspace{-0.2cm}
+This work has been supported in part by the UK Engineering and Physical Sciences Research Council [grants EP/N509668/1, EP/N033957/1 and EP/P004024/1] and European Cooperation in Science and Technology (COST) Action CA15127: RECODIS---Resilient communication and services.
+\vspace{-0.55cm}
 
@@
 \begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/kyle}}]{Kyle A. Simpson}
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/kyle}}]{Kyle~A.~Simpson}
 	received the MSci degree in computing science from the University of Glasgow in 2017.
 	He is currently a PhD student within the Networked Systems Research Laboratory at the School of Computing Science, University of Glasgow.
-
-	Mr.\ Kyle Simpson's research focusses on the use of machine learning and reinforcement learning techniques in cybersecurity and network management, with a core interest in evolving problems and defences.
+	His research focusses on the use of machine learning and reinforcement learning techniques in cybersecurity and network management, with a core interest in evolving problems and defences.
 \end{IEEEbiography}
-\vspace{-1.3cm}
+\vspace{-1.1cm}
 \vskip 0pt plus -1fil
 \begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/sdr}}]{Simon Rogers}
-	received the PhD degree from University of Bristol in 2006.
-Now he is a Senior Lecturer in the School of Computing Science at the University of Glasgow.
-Before starting a permanent post in Glasgow he did a couple of Post Doctoral positions (under Mark Girolami).
-
-	Dr.\ Simon Rogers’s research involves the development of machine learning and statistical techniques to help with the analysis of complex datasets, particularly within the field of metabolomics, human-computer interaction and information retrieval.
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/sdr}}]{Simon~Rogers}
+	is a senior lecturer in the School of Computing Science, University of Glasgow.
+He received his PhD from the department of Engineering mathematics at the University of Bristol in 2005 and has been a permanent member of academic staff at the University of Glasgow since 2009.
+His work focuses on the development of machine learning and statistical methods for the analysis of complex data, particularly within the field of computational biology.
 \end{IEEEbiography}
-\vspace{-1.25cm}
+\vspace{-1.1cm}
 \vskip 0pt plus -1fil
 \begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{authorpics/dimitris}}]{Dimitrios P.\ Pezaros}
+	[{\includegraphics[width=1in,height=1.25in,clip]{authorpics/dimitris}}]{Dimitrios~P.~Pezaros}
 	(S’01–M’04–SM’14) received the B.Sc.
 and Ph.D.
-degrees in computer science from Lancaster University.
-He is currently a Professor and the Founding Director of the Networked Systems Research Laboratory at the School of Computing Science, University of Glasgow.
+degrees in Computer Science from Lancaster University.
+He is currently (full) Professor and the founding director of the Networked Systems Research Laboratory at the School of Computing Science, University of Glasgow.
 He is also a visiting Professor at the University of Athens, Department of Informatics and Telecommunications.
-
-	Professor Pezaros has published widely (over 80 peer-reviewed contributions to top-tier international conferences and journals) in the areas of computer communications, network and service management, and resilience of future networked infrastructures, and has received significant funding for his research in the above areas from public funding agencies and the industry.
+Professor Pezaros has published widely in the areas of computer communications, network and service management, and resilience of future networked infrastructures, and has received significant funding for his research from public funding agencies and the industry.
 He is a Chartered Engineer, and a senior member of the IEEE and the ACM.
